### 2022-05-25 SIMPaper/SIMPaper.Rmd
cleaned, licensed version of cbnn

---
title: "Case-Base Neural Networks: survival analysis with time-varying, higher-order interactions"
author:
- name: Jesse Islam
  num: a
- name: Maxime Turgeon
  num: b
- name: Robert Sladek
  num: c
- name: Sahir Bhatnagar
  num: d
address:
- num: a
  org: Department of Quantitative life sciences, McGill University 
- num: b
  org: Department of Statistics, University of Manitoba
- num: c
  org: Department of Human Genetics, McGill University
- num: d
  org: Department of Biostatistics, McGill University
corres: "\\email{jesse.islam@mail.mcgill.ca}"
authormark: Islam \emph{et al}.
articletype: Research article
received: 2022-04-29
revised: 2022-06-01
accepted: 2022-07-01
abstract: "In survival analysis, covariate interactions are difficult to model using regression-based methods without prior knowledge. Neural network-based survival models have been developed for data-driven interaction modeling, however, these approaches lack implementations for time-varying interactions or a flexible baseline hazard. To address this, we extend the case-base sampling framework and propose Case-Base Neural Networks (CBNN), which introduce time directly into the model, using commonly available neural network components for a simple implementation. First, we case-base sample the survival data, converting the survival time of each participant into discrete moments in time. This allows us to estimate the probability an event occurs at a given moment using a neural network to get the hazard. We compare CBNN to survival methods based on regression and neural networks in two simulations and two real data applications with a fixed neural network architecture. We report metrics for each model over follow-up time using both a proper (Index of Prediction Accuracy) and non-proper (inverse probability censoring weights-adjusted concordance index) scoring rule. A simple simulation shows CBNN outperforming the other neural network methods. The complex simulation highlights the ability of CBNN to model both a complex baseline hazard and time-varying interactions, outperforming all competitors. The first real data application shows CBNN outperforming all neural network competitors, while a second real data application shows competitive performance. CBNN has shown that it is more consistent in prediction performance across time in the simulations and real data applications compared to the competing neural network approaches."

keywords: survival analysis, machine learning, case-base, neural network
bibliography: bibfile.bib
output: 
  rticles::sim_article:
    includes:
      in_header: preamble.tex
    keep_tex: true
header-includes: 
  - \usepackage{placeins}
  - \newcommand{\sigmoid}{\mathrm{sigmoid}}
  - \renewcommand{\thefootnote}{\alph{footnote}}
---

In survival analysis, covariate interactions are difficult to model using regression-based methods without prior knowledge. Neural network-based survival models have been developed for data-driven interaction modeling, however, these approaches can not model both time-varying interactions and complex baseline hazards. To address this, we propose Case-Base Neural Networks (CBNN) which extends the case-base sampling framework which introduce time directly into the model. We case-base sample the survival data, converting the survival time of each participant into discrete moments in time. This allows us to use a feed-forward neural network that estimate the probability of an event occurring at a given moment, which can be used to estimate the hazard function. We then compare CBNN to survival methods based on regression and neural networks in two simulations and two real data applications with a fixed neural network architecture. We report metrics for each model over follow-up time using both proper (Index of Prediction Accuracy) and non-proper (inverse probability censoring weights-adjusted concordance index) scoring rules. A simple simulation with an exponential hazard model shows CBNN consistently outperforming the other neural network methods. The complex simulation highlights the ability of CBNN to model both a complex baseline hazard and time-varying interactions, outperforming all competitors. The first real data application shows CBNN outperforming all neural network competitors, while a second real data application shows competitive performance. In the simulations and real data applications, CBNN provides a more consistent predictive performance across time than the competing neural network approaches.

# Introduction

what is aft, dont like assuming distribution, smooth. Less used than cox. Proportional hazards assumptions hard to maintain in long study, all regression require prior knowledge.

The widespread use of the Cox proportional hazards model has led to a focus on survival models based on hazard ratios and relative risks rather than on survival curves and absolute risks [@hanley2009].
In clinical settings, Cox proportional hazards models are frequently prioritized over smooth-in-time accelerated failure time (AFT) models [@hanley2009], which can estimate absolute risks by modeling the hazard directly through user specification of a baseline hazard distribution. Using splines as the baseline hazard is an alternative to assuming a specific distribution [@royston2002flexible]. The risk associated with each covariate may not be constant over time, making it difficult to maintain the proportional hazards assumption. Previous studies related to breast cancer have discovered time-varying interactions with covariates of interest, like tumor size [@coradini2000time]. A limitation of regression-based models is the requirement of prior knowledge; the analyst must know potential time-varying interactions and explicitly evaluate them.

Smooth-in-time accelerated failure time (AFT) models can estimate absolute risks by modeling the hazard directly through user specification of a baseline hazard distribution [@kleinbaum2012survival]. The assumption of a specific distribution tends to be a point of contention. A method has been developped using splines, providing flexibility to the baseline hazard  [@royston2002flexible]. AFT models are underutilized compared to Cox proportional hazards, resulting in a focus on survival models based on hazard ratios and relative risks rather than on survival curves and absolute risks [@hanley2009]. However, The risk associated with each covariate may not be constant over time, making it difficult to maintain the proportional hazards assumption. Previous studies related to breast cancer have discovered time-varying interactions with covariates of interest, like tumor size [@coradini2000time]. A limitation of all regression-based models is the requirement of prior knowledge; the analyst must know potential time-varying interactions and explicitly evaluate them.

Neural networks provide a data-driven approach to approximating interaction terms. DeepSurv is a neural network-based proportional hazards model that implements the Cox partial log-likelihood as a custom loss function [@katzman2018DeepSurv]. This results in a stepwise absolute risk curve that cannot accommodate time-varying interactions. Compared to Cox regression, DeepSurv has shown an improvement in performance on the Study to Understand Prognoses Preferences and Risk Treatments (SUPPORT) dataset [@knaus1995SUPPORT]. Farraggi et al. suggested a modification to the loss function required for a Cox neural network that can handle non-proportional hazards [@faraggi1995neural]. Wang et al. took the concept of extreme learning machines and applied them in a survival context, removing the need for tuning the number of layers and nodes [@wang2019extreme].

DeepHit requires that each survival time of interest be specified in the model and directly estimates survival curves, rather than deriving a hazard function [@lee2018DeepHit]. It assumes an inverse Gaussian distribution as the baseline hazard and its concordance index (c-index) score outperforms DeepSurv [@katzman2018DeepSurv] on the Molecular Taxonomy of Breast Cancer International Consortium (METABRIC) dataset [@curtis2012genomic].

Deep Survival Machines (DSM) provides a flexible, parametric survival model using neural networks with a mixture of distributions as the baseline hazard [@dsmPaper]. On both the SUPPORT and METABRIC datasets, DSM outperforms DeepSurv and DeepHit [@dsmPaper]. However, like DeepHit and DeepSurv, DSM cannot model time-varying interactions.

These methods need custom loss and activation functions that are difficult to implement in practice. We note the complexity of these alternative neural network approaches; DeepSurv, DeepHit, and DSM all require custom loss functions [@katzman2018DeepSurv] [@lee2018DeepHit] [@dsmPaper], while DSM requires a two-phase learning process and user implementations for distributions beyond Log-Normal or Weibull [@dsmPaper]. Regression-based approaches require prior specification of all interaction terms, which makes it challenging to model covariate effects that change over time. In this article, we propose Case-Base Neural Networks (CBNN) as a method that models time-varying interactions and a flexible baseline hazard using commonly available neural network components. Our approach to modeling the full hazard uses case-base sampling [@hanley2009]. This sampling technique allows probabilistic models to predict survival outcomes. Following the case-base framework, we can use transformations of time as a feature (covariate) to specify different baseline hazards. For example, by including splines of time as covariates, we can approximate the Royston-Parmar flexible baseline hazard model [@royston2002flexible], however, this still requires user specification of time-varying interactions. CBNN can model both without user specification; only a sufficiently deep network is required.

In Section \ref{methods}, we describe how case-base sampling and neural networks are combined both conceptually and algebraically, along with our hyperparameter choices and software implementation. In Section \ref{sims}, we describe our metrics and compare the performance of CBNN, DeepSurv, DeepHit, DSM, Cox regression, and case-base using logistic regression on simulated data. Section \ref{casestudies} describes the real-data analysis, while in Section \ref{discussion}, we explore the implications of our results and contextualize them within neural network survival analysis in a single event setting.

# Case-base neural network methodology, metrics, and software {#methods}

In this section, we define case-base sampling, which converts the total survival time into discrete person-specific moments (person-moments). Then, we detail how neural networks can be used within this framework, explicitly incorporating time as a feature while adjusting for the sampling bias. Finally, we report on the software versions used. A package is available for use at [GITHUBLINK]. The entire code base to generate the analyses in this paper is available at [GITHUB LINK].

## Case-base sampling

Case-base sampling is an alternative framework for survival analysis [@hanley2009]. In case-base sampling, we sample from the continuous survival time of each person in our dataset to create a *base series* of *person-moments*. This *base series* complements the *case series*, which contains all person-moments at which the event of interest occurs.

For each person-moment sampled, let $X_i$ be the corresponding covariate profile $\left(x_{i1},x_{i2},...,x_{ip} \right)$, $T_i$ be the time of the person-moment, and $Y_i$ be the indicator variable for whether the event of interest occurred at time $T_i$. We estimate the hazard function $h(t \mid X_i)$ using the sampled person-moments. Recall that $h(t \mid X_i)$ is the instantaneous potential of experiencing the event at time $t$ for a given set of covariates $X_i$, assuming $T_i \geq t$.

Now, let $b$ be the (user-defined) size of the *base series*, and let $B$ be the sum of all follow-up times for the individuals in the study. If we sample the *base series* uniformly, then the hazard function of the sampling process is equal to $b/B$. Therefore, we have the following equality \footnote{We are abusing notation here, conflating hazards with probabilities. For a rigorous treatment, see Saarela \& Hanley (2015) section 3 \cite{saarela2015} .}: \begin{align}\label{eqn:main}
\frac{P\left(Y_i=1 \mid X_i, T_i\right)}{P\left(Y_i = 0 \mid X_i, T_i\right)} = \frac{h\left(T_i \mid X_i\right)}{b/B}.
\end{align} 
The odds of a person-moment being a part of the *case series* is the ratio of the hazard $h(T_i \mid X_i)$ and the uniform rate $b/B$. Using \eqref{eqn:main}, we can see how the log-hazard function can be estimated from the log-odds arising from case-base sampling: \begin{align}\label{eqn:offset}
\log \left( h\left(t \mid X_i\right)\right) = \log \left(\frac{P\left(Y_i = 1 \mid X_i, t\right)}{P\left(Y_i = 0 \mid X_i, t\right)}\right) + \log\left(\frac{b}{B}\right).
\end{align}

We may estimate the hazard function if we adjust for the bias introduced when sampling a fraction of the study base $B$ by adding the offset term $\log\left(\frac{b}{B} \right)$ (\eqref{eqn:offset}). Next, we propose using neural networks to model the odds.

## Neural networks to model the hazard function

After case-base sampling, we pass all features, including time, into any user-defined feed-forward component, to which an offset term (to adjust for bias from case-base sampling) is added, then passed through a sigmoid activation function (Figure \ref{fig:NNarch}). We use the sigmoid function as its inverse is the odds, which we can use to calculate the hazard. The general form for the neural network using CBNN is:

\begin{align}\label{eqn:nnProb}
P\left(Y=1|X,T\right)&=\sigmoid\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right) \right),
\end{align}

where $T$ is a random variable representing the event time, $X$ is the random variable for a covariate profile, $f_{\theta}(X, T)$ represents any feed-forward neural network architecture, $\log\left(\frac{B}{b}\right)$ is bias term set by case-base sampling, $\theta$ is the set of parameters learned by the neural network and $\sigmoid(x)=\frac{1}{1+e^{-x}}$. By allowing a neural network to approximate a higher-order polynomial of time, the baseline hazard specification is now data-driven, where user-defined hyperparameters such as regularization, number of nodes, and layers control the flexibility of the model. We provide a detailed description of the choices we made in the next sub-section.

```{r NNarch,echo = FALSE, fig.align = 'center', out.width = "100%", fig.cap = "Steps involved in CBNN from case-base sampling to the model framework we use for training. The first step is case-base sampling, completed before training begins. Next, we pass this sampled data through a feed-forward neural network. We add the offset and pass that through a sigmoid activation function, whose output is a probability. Once the neural network model completes its training, we can convert the probability output to a hazard, using it for our survival outcomes of interest."}
library(knitr)
knitr::include_graphics(here::here("figures", "nnarch2.PNG"))
```

The following derivation shows how our probability estimate is converted to odds:
\begin{align*}
 \log\left( h(t \mid X_i) \right) &= \log\left(\frac{\sigmoid \left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}{1-\sigmoid \left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}\right) + \log\left(\frac{b}{B}\right) \\
 &= \log\left( \frac{\frac{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)+1}}{1-\frac{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)+1}}\right) + \log\left(\frac{b}{B}\right) \\
 &= \log\left(\exp\left( f_{\theta}(X, T) + \log\left(\frac{B}{b}\right) \right) \right) + \log\left(\frac{b}{B}\right) \\
 &= f_{\theta}(X, T) + \log\left(\frac{B}{b}\right) + \log\left(\frac{b}{B}\right) \\
&= f_{\theta}(X, T). 
\end{align*}

We use binary cross-entropy as our loss function [@gulli2017]:
\begin{align*}
L(\theta)=-\frac{1}{N} \sum^{N}_{i=1} y_{i} \cdot \log(\hat{f}_{\theta}(x_{i}, t_{i}) ) + (1-y_{i} )\cdot \log(1-\hat{f}_{\theta}(x_{i}, t_{i}) ),
\end{align*}
where $\hat{f}_{\theta}(x_{i}, t_{i})$ is our estimate for a given covariate profile and time, $y_{i}$ is our target value specifying whether an event occurred, and $N$ represents the number of individuals in our training set.

Backpropagation is used to optimize the parameters in the model with an appropriate minimization algorithm (e.g. Adam, RMSPropagation, stochastic gradient descent)[@gulli2017]. For our analysis, we use Adam. Note that the size of the *case series* is fixed as the number of events, but we can make the *base series* as large as we want. A ratio of 100:1 *base series* to *case series* is sufficient [@hanley2009]. We pass our feed-forward neural network through a sigmoid activation function (Figure \ref{fig:NNarch}). Finally, we can convert this model output to a hazard. When using our model for predictions, we manually set the offset term to 0 in the new data, as we account for the bias during the fitting process.

Since we are directly modeling the hazard, we can readily estimate the risk function ($F$) at time $t$ for a covariate profile $X$, viz.\begin{align}\label{eqn:ci2}
F\left(t|X\right)& = 1 - \exp\left(-\int_{0}^{t}h(u|X) \,\textrm du\right).
\end{align} 
We use Riemann's Sum [@hughes2020calculus] to approximate the integral in \eqref{eqn:ci2}.

## Hyperparameter selection

Neural networks provide flexibility when defining the architecture and optimization parameters of our models. These hyperparameter choices can affect the estimated parameters and were chosen during a set of initial simulations to determine if CBNN can learn complex interactions in practice. We set a max number of epochs to be $2000$, batch size as $512$, learning rate as $10^{-3}$, decay as $10^{-7}$, patience as $10$ epochs, $\{50,50,25,25\}$ nodes in each hidden layer with $50\%$ dropout at each layer, a validation split of $20\%$, a testing split of $10%$, a minimum delta loss on the validation set of $10^{-7}$ over 10 epochs, and Adam [@gulli2017] as our optimizer. These choices may permit the model to approximate higher-order interactions while preventing over-fitting [@srivastava2014dropout]. We fix a train-validation-test split that allows us to update the weights with a subset of the data (training), assess performance at each epoch (validation) and gauge performance of the final model (test). We select the best weights after training based on validation loss [@gulli2017].

## Software implementation

R [@Rsoft] and python [@py] are used to evaluate methods from both languages. We fit the Cox model using the **survival** package [@survpkg], the CBLR model using the **casebase** package [@cbpkg], the DeepSurv model using the **survivalmodels** package [@survmods], the DSM model using **DeepSurvivalMachines** [@dsmPaper] and the DeepHit model using **pycox** [@lee2018DeepHit]. We made the components of CBNN using the **keras** [@keras] package and the **casebase** package for the sampling step. The **simsurv** package [@simsurv] is used to simulate our simulation studies, while **flexsurv** [@flexsurv] is used to fit a flexible baseline hazard using splines for our complex simulation. We use the implementation of c-ipcw from the python package **sksurv** [@sksurv]. The **riskRegression** package [@riskRegression] is used to get the Brier score and IPA metrics. We changed the **riskRegression** package to be used with any user supplied risk function $F$. To ensure that both R and Python-based models are running in unison on the same data through our simulations and bootstrap, we use the **reticulate** package [@reticulate].

# Simulation studies {#sims}

In this section, we use simulated data to evaluate the performance of CBNN and compare our approach with existing methods based on regression (Cox, CBLR) and neural network (DeepHit, DeepSurv, DSM) methods. We specify a linear combination of each covariate as the linear predictor in regression-based approaches (Cox, CBLR), which contrasts with neural network approaches that allow for non-linear interactions. We simulate data under a simple exponential model, and a complex baseline hazard with time-varying interactions, each with 10% random censoring. For both settings, we simulate three covariates:

$$
z_{1} \sim \textrm{Bernoulli}(0.5) \qquad \qquad
z_{2} \sim \begin{cases}
 N(0,0.5) & \textrm{if } z_{1}=0\\ 
 N(10,0.5) & \textrm{if } z_{1}=1
\end{cases} \qquad \qquad
z_{3} \sim \begin{cases}
 N(8,0.5) & \textrm{if } z_{1}=0\\ 
 N(-3,0.5) & \textrm{if } z_{1}=1.
\end{cases}
$$

The DeepHit-specific hyperparameter alpha is set to $0.5$ (equal weight between its negative log-likelihood and ranking losses [@lee2018DeepHit]). We change the **DeepSurvivalMachines** [@dsmPaper] package to include dropout and a minimum delta loss during the fitting process. For DSM, we define a mixture model of six Weibull distributions for the baseline hazard. All other hyperparameters are held constant across all neural network methods in both the simulation studies and real data applications.

Besides the methods mentioned above, we include the Optimal model in our comparisons using CBLR. That is, we include the exact functional form of the covariates in a CBLR model (referred to as Optimal for simplicity). We calculate $t$-based 95% confidence intervals using 100 replications of the simulated data For all analyses, we use 80% for training and 20% for the test set. 20% of the training set is kept for validation at each epoch. We predict risk functions $F$ using \eqref{eqn:ci2} for individuals in the test set, which are used to calculate our c-ipcw and IPA scores.

## Performance metrics

We use two metrics to assess the performance of the different methods of interest: 1) index of prediction accuracy (IPA) [@kattan2018index] and 2) inverse probability censoring weights-adjusted concordance index (c-ipcw) [@uno2011], which we define below. Both these time-dependent metrics provide transparency as to when in follow-up time each model may perform better than the others.

### Index of prediction accuracy (IPA)

The IPA is a function of the Brier score ($BS(t)$) [@graf1999], which is defined as
\begin{align}
BS(t)=\frac{1}{N}\sum^{N}_{i=1}\left(\frac{\left(\widehat{F}(X_{i},t)-1\right)^{2}\cdot I_{T_{i}\leq t,\delta_{i}=1}}{\widehat{G}(T_{i})} + \frac{\left(\widehat{F}(X_{i},t)\right)^{2}\cdot I_{T_{i}>t}}{\widehat{G}(t)}\right),
\end{align}
where $\delta_{i}=1$ shows individuals who have experienced the event, $N$ represents the number of samples in our dataset over which we calculate $BS(t)$, $\widehat{G}(t)=P[C>t]$ is a non-parametric estimate of the censoring distribution, $C$ is censoring time and $T_{i}$ is an individual’s survival or censoring time. The Brier score provides a score that accounts for the information loss because of censoring. There are three categories of individuals that may appear within the dataset once we fix our $t$ of interest. Individuals who experienced the event before $t$ are present in the first term of the equation. The second term of the equation includes individuals who experience the event or are censored after $t$. Those censored before $t$ are the third category of people. The  inverse probability censoring weights (IPCW) adjustment ($G(\cdot)$) is to account for these category three individuals whose information is missing. The IPA score as a function of time is given by
\begin{align}
\textrm{IPA}(t) &= 1-\frac{BS_{model}(t)}{BS_{null}(t)}, \nonumber
\end{align} where $BS_{model}(t)$ represents the Brier score over time $t$ for the model of interest, and $BS_{null}(t)$ represents the Brier score if we use an unadjusted Kaplan-Meier (KM) curve as the prediction for all observations [@kattan2018index]. Note that IPA has an upper bound of one, where positive values show an increase in performance over the null model and negative values show that the null model performs better. These scores demonstrate how performance changes over follow-up time.

A potential artifact of IPA is that the score is unstable at earlier and later survival times. This is because of a near equivalent Brier score between each model and the null model. At small values, a difference of $0.1$ creates a more significant fold change than at larger values. As the Brier score is potentially small at the start and at the end of their respective curves, we may see instability of the IPA score at those ends.

### Inverse probability censoring weights-adjusted concordance index

The c-ipcw is a non-proper, rank-based metric that does not depend on the censoring times in the test data [@uno2011]. The c-ipcw is given by

\begin{align} \label{eq:cidx}
c-ipcw(t) &= \frac{\sum^{N}_{i=1}\sum^{N}_{j=1}\delta_{i}\left\{\widehat{G}(T_{i})\right\}^{-2} I(T_{i}<T_{j},T_{i}<t) I\left(\widehat{F}(t|X_{i})>\widehat{F}(t|X_{j})\right)}{\sum^{N}_{i=1}\sum^{N}_{j=1}\delta_{i}\left\{\widehat{G}(T_{i})\right\}^{-2} I(T_{i}<T_{j},T_{i}<t)},
\end{align} where the follow-up period of interest is (0,$t$), $I(\cdot)$ is an indicator function, $\widehat{F}(X_{i},t)$ is the risk function estimated for everyone in the study at time $t$ and c-ipcw can compare the performance of different models, where a higher score is better. Note that the c-ipcw may produce misleading performance, as it ranks based on survival times, not event status [@cindexfails2019]. This metric is considered an unbiased population concordance measure because of the IPCW adjustment [@uno2011].

\srb{in figure 2 i changed $C_{IPCW}$ to c-ipcw, as capital C is used for censoring. in the plots, would it make sense to leave it as a small letter?}

```{r megaPlot, echo = FALSE , fig.align = 'center', out.width = "100%", fig.cap = "Summarizes the simple simulation (A, B), complex simulation (C, D), SUPPORT case study (E, F) and METABRIC case study (G, H) results. The first row shows the IPA score for each model in each study over follow-up time. Negative values mean our model performs worse than the null model and positive values mean the model performs better. The second row shows the c-ipcw for each model in each study over follow-up time. A score of 1 is the maximum performance for either metric. Each model-specific metric in each study shows a 95-percent confidence interval over 100 iterations. The models of interest are case-base with logistic regression (CBLR), Case-Base Neural Networks (CBNN), Cox, DeepHit, DeepSurv, Deep Survival Machines (DSM), Optimal (a CBLR model with the exact interaction terms and baseline hazard specified) and Kaplan-Meier (to serve as a baseline, predicting the average for all individuals)."}
#readRDS(here::here("figures", "megaPlot.rds"))
knitr::include_graphics(here::here("../pmnnPlayground/reviewedAnalysisPMNN/figures", "megaPlot.pdf"))
```

## Simple simulation: constant baseline hazard

We simulate data from a simple model that depends on a constant baseline hazard:
$$h(t \mid X_i) = \lambda\cdot e^{\beta_{{1}}z_{1}+\beta_{{2}}z_{2}+\beta_{{3}}z_{3}},$$
with minimal covariate effects given by $\beta_{{1}}=0.1, \beta_{{2}}=0.1, \beta_{{3}}=0.1$ and $\lambda=1.0$. Once we simulate survival times, we introduce 10% random censoring.

### Performance comparison in simple simulation

Figure \ref{fig:megaPlot} A, B and Table \ref{tab:megaTable} A show the results for the simple simulation. The regression-based methods (CBLR, Cox, Optimal) outperform the neural network ones in the simple simulation setting. Among the neural network approaches, CBNN outperforms all other methods in terms of both IPA and c-ipcw (Figure \ref{fig:megaPlot} A, B). Specifically, we see CBNN is consistent across time with smaller confidence intervals compared to DeepHit, DeepSurv and DSM.

In this simple setting, the regression models are much closer to the Optimal model, while the neural network models perform worse than the null model. The wide confidence bands in c-ipcw suggest the neural network models may be over-parameterized.

## Complex simulation: flexible baseline hazard, time-varying interactions

This simulation demonstrates performance with the presence of a complex baseline hazard and a time-varying interaction. Originally used to demonstrate the spline-based hazard model proposed by Royston and Parmar [@royston2002flexible], the breast cancer dataset provides a complex hazard from which we simulate, available in the **flexsurv** R package [@flexsurv]. To increase the complexity of our data-generating mechanism for this simulation, we design the model as follows:
\begin{align}
h(t \mid X_i) =\sum_{i=1}^{5} (\gamma_{i} \cdot \psi_{i}) + \beta_{{1}} (z_{1}) + \beta_{{2}} (z_{2})+ \beta_{{3}} (z_{3})+ \tau_{1} ( z_{1} \cdot z_{2} \cdot t)+ \tau_{2} ( z_{1} \cdot z_{3})+ \tau_{3} (z_{2} \cdot z_{3}), \nonumber
\end{align}

```{=latex}
\FloatBarrier
\begin{table}
\caption{Four tables representing performance at certain follow-up times for the simple simulation, complex simulation, SUPPORT and METABRIC. Each table shows performance for each method in each study at $25\%$, $50\%$. $75\%$ and $100\%$ of follow-up time. The bold elements show the best model for each study, at each follow-up time of interest. These tables are included to provide exact measures at certain intervals. The models of interest are: Cox, case-base with logistic regression (CBLR), DeepSurv, DeepHit, Case-Base Neural Networks (CBNN), Optimal (a CBLR model with the exact interaction terms and baseline hazard specified) and Deep Survival Machines (DSM).}
\label{tab:megaTable}
```

```{r megaTable, echo = FALSE , fig.align = 'center', out.width = "100%"}
#readRDS(here::here("figures", "megaPlot.rds"))
knitr::include_graphics(here::here("../pmnnPlayground/reviewedAnalysisPMNN/figures", "megaTable.pdf"))
```

```{=latex}
\end{table}
\FloatBarrier
```

where $\gamma_{1}=3.9, \gamma_{2}=3, \gamma_{3}=-0.43, \gamma_{4}=1.33,\gamma_{5}=-0.86, \beta_{{1}}=1, \beta_{{2}}=1, \beta_{{3}}=1, \tau_{1}=10, \tau_{2}=2, \tau_{3}=2$ and $\psi$ are basis splines. To generate the $gamma$ coefficients, let $g$: $\mathbb{R} \rightarrow \mathbb{R}$ be a smoothing method for variable $t$ by a projection on to a set of basis functions on splines: 
\srb{i used gt as the function for splines. Specifically, natural cubic spline function of log time https://www.rdocumentation.org/packages/flexsurv/versions/2.1/topics/flexsurvspline . therefore, should i use g(log(t)) and and replace "smoothing method" with natural cubic spline? SAHIR: will fix any issues in this section in the final version.}
\begin{align}
g(t)=\sum^{m}_{l=1}\psi_{l} (t)\gamma_{l}.
\end{align}

An initial model with three knots ($m=5$) is fit using the *flexsurvspline* function with an intercept-only model on the data, which returns the coefficients ($\gamma$) and basis of time ($\psi$) we use in our model. Note that we fix these values for the analysis. The $\beta$ coefficients represent direct effects, $\tau_{2}$ and ${\tau_3}$ represent interactions and $\tau_{1}$ is a time-varying interaction.

### Performance comparison in complex simulation

Figure \ref{fig:megaPlot} C, D and Table \ref{tab:megaTable} B show the performance over time on a test set in the complex simulation. Apart from the Optimal regression model, CBNN outperforms the competing models when examining IPA, and up to the 75-percentile of follow-up time for c-ipcw. We expected the Optimal model to perform best in both metrics. However, this was not the case for c-ipcw and may be due to an artifact of concordance-based metrics, where a misspecified model may perform better than a correctly specified one [@cindexfails2019]. We attribute this difference in performance to the flexibility of CBNN in terms of time-varying interactions and baseline hazard, flexibility the other neural network models do not have, explaining their drop in performance.

# Application to SUPPORT and METABRIC data {#casestudies}

Our complex simulation demonstrates the superior performance of CBNN in ideal conditions with clean data. To obtain a more realistic performance assessment, we compared models using two real datasets with a time-to-event outcome. The first case study examines the SUPPORT dataset [@knaus1995SUPPORT]. The second case study examines the METABRIC dataset [@curtis2012genomic]. We use the same hyperparameters as in the simulation studies. As we do not know the true model for the real data, we exclude the Optimal model. We split the datasets keeping 20% of the observations as a test set. 20% of the training set is kept aside for validation at each epoch. We predict risk functions for everyone in the test set, which are used to calculate our metrics. We conduct 100 bootstrap re-samples for the real data applications to obtain confidence intervals.

## Performance evaluation using the SUPPORT dataset

The SUPPORT dataset tracks the time until death for seriously ill patients at five American hospitals [@knaus1995SUPPORT]. We use a pre-processed version of the dataset made available in the DeepSurv package [@katzman2018DeepSurv]. This dataset contains 9104 samples and 14 covariates (age, sex, race, number of comorbidities, presence of diabetes, presence of dementia, presence of cancer, mean arterial blood pressure, heart rate, respiration rate, temperature, white blood cell count, serum’s sodium, and serum’s creatinine) [@katzman2018DeepSurv]. Patients with missing features were excluded and 68.10% of the patients died during the 5.56-year study period [@katzman2018DeepSurv].

Figure \ref{fig:megaPlot} E, F and Table \ref{tab:megaTable} C demonstrates the performance over time on a test set. The regression models (CBLR, Cox) perform best considering IPA, followed by CBNN from the $25^{th}$ to $100^{th}$ percentile of follow-up time. We note a drop in performance for CBNN before the $25^{th}$ percentile of follow-up. For c-ipcw, CBNN outperforms the competing models consistently over follow-up time. Note that performance is similar for all models, aside from DeepSurv whose c-ipcw is lower than the rest (Figure \ref{fig:megaPlot} E, F and Table \ref{tab:megaTable} C).

## Performance evaluation using the METABRIC dataset

METABRIC is a 30-yearlong study aiming to discover the molecular drivers of breast tumors, following 2000 individuals with breast cancer until death [@curtis2012genomic]. They described these growths as primary invasive breast carcinomas, with the goal of discovering both genetic and clinical risk factors for breast cancer survival [@curtis2012genomic]. We used the processed dataset made available through DeepSurv [@katzman2018DeepSurv], which includes 1980 patients of which 57.72% die due to breast cancer within a median 10 years of follow-up [@katzman2018DeepSurv]. There are 11 covariates in total: 4 RNA-Seq gene expressions (MKI67, EGFR, PGR, and ERBB2) and 5 clinical features (hormone treatment indicator, radiotherapy indicator, chemotherapy indicator, ER-positive indicator, and age at diagnosis) [@katzman2018DeepSurv].

Figure \ref{fig:megaPlot} G, H and Table \ref{tab:megaTable} D shows the performance on a test set over time. The IPA scores suggest that regression models outperform competing models on this dataset, as all the neural network models are equal to or perform worse than KM over follow-up time. Our CBNN model is comparable to KM until around the 50-percentile of follow-up time, after which CBNN, DeepSurv and DSM drop in performance. The c-ipcw produces a different ranking. Our CBNN model outperforms the other models up to the $25^{th}$ percentile of follow-up time, whereas DeepHit performs best from the $50^{th}$ to $75^{th}$ percentile of follow-up. With this metric, CBNN and DeepHit outperform the regression models. This disagreement between IPA and c-ipcw between may be due to the misspecification issue of concordance-based metrics [@cindexfails2019]. The neural network models may be over-parameterized, as shown by the wide confidence bands in c-ipcw.

# Discussion {#discussion}

CBNN models survival outcomes by using neural networks on case-base sampled data. While estimating the hazard function, we incorporate follow-up-time as a feature, providing a data-driven estimation of a flexible baseline hazard and time-varying interactions. Based on our complex simulation results (Figure \ref{fig:megaPlot} C, D and Table \ref{tab:megaTable} B), CBNN outperforms the competitors when time-varying interactions and a complex baseline hazard are present.

The competing neural network model we evaluated can not model time varying interactions. DeepSurv is a proportional hazards model and does not estimate the baseline hazard [@katzman2018DeepSurv]. DeepHit requires an alpha hyperparameter, is restricted to a single distribution for the baseline hazard, and models the survival function directly [@lee2018DeepHit]. DSM requires the user to specify the component distributions in the mixture for a flexible baseline hazard to be fit [@dsmPaper]. These alternative neural network approaches match on time, while CBNN models time directly.

In the simple simulation, all neural network-based approaches performed worse than KM (IPA scores), while the regression-based approaches did not. We attribute this to potential over-parameterization in the neural network models. The wide confidence intervals suggest over-fitting may be the reason for the loss in performance, even with dropout. Both DeepSurv and DSM are affected, while DeepHit and CBNN are less so.

In the complex simulation, CBNN has a distinct advantage over the other methods while examining the IPA score. The regression models do show improvement over the null model, while the competing neural network approaches do not. The time-varying interactions and complex baseline hazard may have caused a drop in performance for the competing neural network methods, compared to the regression models without interactions.

In the SUPPORT application, the overall performance considering IPA is near identical. The regression models outperform CBNN, which performs better than the competing neural network models. As the confidence intervals are tight, it is unlikely that the loss in performance is because of over-fitting. Flexibility in both interaction modeling and baseline hazard is not particularly helpful with the SUPPORT dataset.

In the METABRIC application IPA results, DSM and DeepSurv had a steep drop in performance. As DeepSurv does not depend on the baseline hazard during the fitting process and DSM is the most flexible competing model in our comparison, the baseline hazard is unlikely to cause this drop. With tight confidence intervals, over-fitting is unlikely as well. Further research would be required to determine the cause of this drop in performance. We do not see the same drop in CBNN demonstrating that, in the METABRIC application, CBNN outperforms DeepSurv and DSM. Considering IPA score, regression models consistently outperform the neural network models. From a biological perspective, the 30-year follow-up time in the METABRIC study may contain competing causes of death. The signal from the start of the study may not match the signal towards the end, potentially explaining the drop in performance as we reach later survival times with fewer individuals.

Concordance-based measures are commonly used in survival analysis to compare models and we opt to keep them in our analyses. However, c-ipcw is a non-proper metric and may cause misspecified models to appear better than they should [@cindexfails2019]. The model rankings between c-ipcw and IPA differed for both the complex simulation and METABRIC application. In the complex simulation, the c-ipcw of a correctly specified model (Optimal) underperforms CBNN and CBLR; models without the exact interaction terms specified. We base our METABRIC assessment of this disagreement between metrics on the complex simulation. The IPA score ranks regression models first, followed by DeepHit and CBNN, but the c-ipcw score ranks regression models after DeepHit and CBNN. Wide confidence intervals for neural network models in c-ipcw show potential misspecification, suggesting that regression models perform best in the METABRIC application.

During our testing for the complex simulation and SUPPORT application, DSM did not converge with our choice of hyper-parameters and model design. As the goal is to fix the design and compare performance, we did not change any shared hyperparameters. For DSM, a different number and combination of component distributions did not help the model converge, though we did not conduct an exhaustive test. This may be a limitation to their method or early-stage implementation issues. We include this method where it did function as it shares a similar goal of a flexible baseline hazard.

If the user suspects that there may be unknown time-varying interactions and a complex baseline hazard, CBNN should strongly be considered. Once we perform case-base sampling and adjust for the sampling bias, we can use a sigmoid activation function to predict our hazard function. The user can approach their problem as they would for probabilistic modeling while interpreting their prediction in a survival context. In terms of performance, we saw that CBNN consistently performs better than the other neural network-based approaches over follow-up time. This may be because of the hyperparameters chosen, as they differ from each of the respective methods papers. CBNN only requires the use of standard components in machine learning libraries (add layer and sigmoid activation function). Due to the simplicity in its implementation and by extension user experience, we expect CBNN to be both a user-friendly approach to data-driven survival analysis and easily extendable to any feed-forward neural network frameworks.

Though CBNN can model time-varying interactions and a flexible baseline hazard, all models require substantial amounts of data to learn real, complex relationships effectively. Larger datasets will appear with time, but the cost of generating longitudinal datasets remains. A viable alternative strategy is the use of transfer learning, where independent and identically distributed data can function as a preliminary model to learn related features in a parallel dataset, initializing weights for our survival task [@bozinovski2020reminder].

This paper describes CBNN in the single event setting. A next step would be to extend this approach to competing risks. A promising use case would be high-dimensional data with complex correlation structures, such as imaging data.

### 2022-05-25 SIMPaper/SIMPaper.tex
cleaned, licensed version of cbnn

--- 2022-05-25 SIMPaper/SIMPaper.Rmd
@@
----
-title: "Case-Base Neural Networks: survival analysis with time-varying, higher-order interactions"
-author:
-- name: Jesse Islam
-  num: a
-- name: Maxime Turgeon
-  num: b
-- name: Robert Sladek
-  num: c
-- name: Sahir Bhatnagar
-  num: d
-address:
-- num: a
-  org: Department of Quantitative life sciences, McGill University 
-- num: b
-  org: Department of Statistics, University of Manitoba
-- num: c
-  org: Department of Human Genetics, McGill University
-- num: d
-  org: Department of Biostatistics, McGill University
-corres: "\\email{jesse.islam@mail.mcgill.ca}"
-authormark: Islam \emph{et al}.
-articletype: Research article
-received: 2022-04-29
-revised: 2022-06-01
-accepted: 2022-07-01
-abstract: "In survival analysis, covariate interactions are difficult to model using regression-based methods without prior knowledge.
-Neural network-based survival models have been developed for data-driven interaction modeling, however, these approaches lack implementations for time-varying interactions or a flexible baseline hazard.
-To address this, we extend the case-base sampling framework and propose Case-Base Neural Networks (CBNN), which introduce time directly into the model, using commonly available neural network components for a simple implementation.
-First, we case-base sample the survival data, converting the survival time of each participant into discrete moments in time.
-This allows us to estimate the probability an event occurs at a given moment using a neural network to get the hazard.
-We compare CBNN to survival methods based on regression and neural networks in two simulations and two real data applications with a fixed neural network architecture.
-We report metrics for each model over follow-up time using both a proper (Index of Prediction Accuracy) and non-proper (inverse probability censoring weights-adjusted concordance index) scoring rule.
-A simple simulation shows CBNN outperforming the other neural network methods.
-The complex simulation highlights the ability of CBNN to model both a complex baseline hazard and time-varying interactions, outperforming all competitors.
-The first real data application shows CBNN outperforming all neural network competitors, while a second real data application shows competitive performance.
-CBNN has shown that it is more consistent in prediction performance across time in the simulations and real data applications compared to the competing neural network approaches."
+\documentclass[AMA,STIX1COL,]{WileyNJD-v2}
 
-keywords: survival analysis, machine learning, case-base, neural network
-bibliography: bibfile.bib
-output: 
-  rticles::sim_article:
-    includes:
-      in_header: preamble.tex
-    keep_tex: true
-header-includes: 
-  - \usepackage{placeins}
-  - \newcommand{\sigmoid}{\mathrm{sigmoid}}
-  - \renewcommand{\thefootnote}{\alph{footnote}}
----
+% tightlist command for lists without linebreak
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
-In survival analysis, covariate interactions are difficult to model using regression-based methods without prior knowledge.
-Neural network-based survival models have been developed for data-driven interaction modeling, however, these approaches can not model both time-varying interactions and complex baseline hazards.
-To address this, we propose Case-Base Neural Networks (CBNN) which extends the case-base sampling framework which introduce time directly into the model.
-We case-base sample the survival data, converting the survival time of each participant into discrete moments in time.
-This allows us to use a feed-forward neural network that estimate the probability of an event occurring at a given moment, which can be used to estimate the hazard function.
-We then compare CBNN to survival methods based on regression and neural networks in two simulations and two real data applications with a fixed neural network architecture.
-We report metrics for each model over follow-up time using both proper (Index of Prediction Accuracy) and non-proper (inverse probability censoring weights-adjusted concordance index) scoring rules.
-A simple simulation with an exponential hazard model shows CBNN consistently outperforming the other neural network methods.
-The complex simulation highlights the ability of CBNN to model both a complex baseline hazard and time-varying interactions, outperforming all competitors.
-The first real data application shows CBNN outperforming all neural network competitors, while a second real data application shows competitive performance.
-In the simulations and real data applications, CBNN provides a more consistent predictive performance across time than the competing neural network approaches.
+\usepackage{longtable}
+\newcommand{\srb}[1]{\textsf{\textcolor{red}{#1}}}
+\usepackage{ulem}
+\usepackage{placeins}
+\newcommand{\sigmoid}{\mathrm{sigmoid}}
+\renewcommand{\thefootnote}{\alph{footnote}}
 
-# Introduction
+\articletype{Research article}
 
-what is aft, dont like assuming distribution, smooth. Less used than cox. Proportional hazards assumptions hard to maintain in long study, all regression require prior knowledge.
+\received{2022-04-29}
 
-The widespread use of the Cox proportional hazards model has led to a focus on survival models based on hazard ratios and relative risks rather than on survival curves and absolute risks [@hanley2009].
-In clinical settings, Cox proportional hazards models are frequently prioritized over smooth-in-time accelerated failure time (AFT) models [@hanley2009], which can estimate absolute risks by modeling the hazard directly through user specification of a baseline hazard distribution.
-Using splines as the baseline hazard is an alternative to assuming a specific distribution [@royston2002flexible].
-The risk associated with each covariate may not be constant over time, making it difficult to maintain the proportional hazards assumption.
-Previous studies related to breast cancer have discovered time-varying interactions with covariates of interest, like tumor size [@coradini2000time].
-A limitation of regression-based models is the requirement of prior knowledge; the analyst must know potential time-varying interactions and explicitly evaluate them.
+\revised{2022-06-01}
 
-Smooth-in-time accelerated failure time (AFT) models can estimate absolute risks by modeling the hazard directly through user specification of a baseline hazard distribution [@kleinbaum2012survival].
-The assumption of a specific distribution tends to be a point of contention.
-A method has been developped using splines, providing flexibility to the baseline hazard  [@royston2002flexible].
-AFT models are underutilized compared to Cox proportional hazards, resulting in a focus on survival models based on hazard ratios and relative risks rather than on survival curves and absolute risks [@hanley2009].
-However, The risk associated with each covariate may not be constant over time, making it difficult to maintain the proportional hazards assumption.
-Previous studies related to breast cancer have discovered time-varying interactions with covariates of interest, like tumor size [@coradini2000time].
-A limitation of all regression-based models is the requirement of prior knowledge; the analyst must know potential time-varying interactions and explicitly evaluate them.
+\accepted{2022-07-01}
 
-Neural networks provide a data-driven approach to approximating interaction terms.
-DeepSurv is a neural network-based proportional hazards model that implements the Cox partial log-likelihood as a custom loss function [@katzman2018DeepSurv].
-This results in a stepwise absolute risk curve that cannot accommodate time-varying interactions.
-Compared to Cox regression, DeepSurv has shown an improvement in performance on the Study to Understand Prognoses Preferences and Risk Treatments (SUPPORT) dataset [@knaus1995SUPPORT].
-Farraggi et al.
-suggested a modification to the loss function required for a Cox neural network that can handle non-proportional hazards [@faraggi1995neural].
-Wang et al.
-took the concept of extreme learning machines and applied them in a survival context, removing the need for tuning the number of layers and nodes [@wang2019extreme].
+\raggedbottom
 
-DeepHit requires that each survival time of interest be specified in the model and directly estimates survival curves, rather than deriving a hazard function [@lee2018DeepHit].
-It assumes an inverse Gaussian distribution as the baseline hazard and its concordance index (c-index) score outperforms DeepSurv [@katzman2018DeepSurv] on the Molecular Taxonomy of Breast Cancer International Consortium (METABRIC) dataset [@curtis2012genomic].
+\begin{document}
 
-Deep Survival Machines (DSM) provides a flexible, parametric survival model using neural networks with a mixture of distributions as the baseline hazard [@dsmPaper].
-On both the SUPPORT and METABRIC datasets, DSM outperforms DeepSurv and DeepHit [@dsmPaper].
-However, like DeepHit and DeepSurv, DSM cannot model time-varying interactions.
+\title{Case-Base Neural Networks: survival analysis with time-varying,
+higher-order interactions}
 
-These methods need custom loss and activation functions that are difficult to implement in practice.
-We note the complexity of these alternative neural network approaches; DeepSurv, DeepHit, and DSM all require custom loss functions [@katzman2018DeepSurv] [@lee2018DeepHit] [@dsmPaper], while DSM requires a two-phase learning process and user implementations for distributions beyond Log-Normal or Weibull [@dsmPaper].
-Regression-based approaches require prior specification of all interaction terms, which makes it challenging to model covariate effects that change over time.
-In this article, we propose Case-Base Neural Networks (CBNN) as a method that models time-varying interactions and a flexible baseline hazard using commonly available neural network components.
-Our approach to modeling the full hazard uses case-base sampling [@hanley2009].
-This sampling technique allows probabilistic models to predict survival outcomes.
-Following the case-base framework, we can use transformations of time as a feature (covariate) to specify different baseline hazards.
-For example, by including splines of time as covariates, we can approximate the Royston-Parmar flexible baseline hazard model [@royston2002flexible], however, this still requires user specification of time-varying interactions.
-CBNN can model both without user specification; only a sufficiently deep network is required.
+\author[a]{Jesse Islam}
+\author[b]{Maxime Turgeon}
+\author[c]{Robert Sladek}
+\author[d]{Sahir Bhatnagar}
 
-In Section \ref{methods}, we describe how case-base sampling and neural networks are combined both conceptually and algebraically, along with our hyperparameter choices and software implementation.
-In Section \ref{sims}, we describe our metrics and compare the performance of CBNN, DeepSurv, DeepHit, DSM, Cox regression, and case-base using logistic regression on simulated data.
-Section \ref{casestudies} describes the real-data analysis, while in Section \ref{discussion}, we explore the implications of our results and contextualize them within neural network survival analysis in a single event setting.
+\authormark{Islam \emph{et al}.}
 
-# Case-base neural network methodology, metrics, and software {#methods}
+\address[a]{Department of Quantitative life sciences, McGill University}
+\address[b]{Department of Statistics, University of Manitoba}
+\address[c]{Department of Human Genetics, McGill University}
+\address[d]{Department of Biostatistics, McGill University}
 
-In this section, we define case-base sampling, which converts the total survival time into discrete person-specific moments (person-moments).
-Then, we detail how neural networks can be used within this framework, explicitly incorporating time as a feature while adjusting for the sampling bias.
-Finally, we report on the software versions used.
-A package is available for use at [GITHUBLINK].
-The entire code base to generate the analyses in this paper is available at [GITHUB LINK].
+\corres{\email{jesse.islam@mail.mcgill.ca}}
 
-## Case-base sampling
+\abstract{In survival analysis, covariate interactions are difficult to
+model without prior knowledge using regression-based methods. Neural
+network-based survival models have been developed for data-driven
+interaction modeling, however, these approaches lack implementations for
+time-varying interactions or a flexible baseline hazard. To address
+this, we extend the case-base sampling framework and propose Case-Base
+Neural Networks (CBNN), which introduce time directly into the model,
+using commonly available neural network components for a simple
+implementation. First, we case-base sample the survival data, converting
+the survival time of each participant into discrete moments in time.
+This allows us to estimate the probability an event occurs at a given
+moment using a neural network to get the hazard. We compare CBNN to
+survival methods based on regression and neural networks in two
+simulations and two real data applications with a fixed neural network
+architecture. We report metrics for each model over follow-up time using
+both a proper (Index of Prediction Accuracy) and non-proper (inverse
+probability censoring weights-adjusted concordance index) scoring rule.
+A simple simulation shows CBNN outperforming the other neural network
+methods. The complex simulation highlights the ability of CBNN to model
+both a complex baseline hazard and time-varying interactions,
+outperforming all competitors. The first real data application shows
+CBNN outperforming all neural network competitors, while a second real
+data application shows competitive performance. CBNN has shown that it
+is more consistent in prediction performance across time in the
+simulations and real data applications compared to the competing neural
+network approaches.}
 
-Case-base sampling is an alternative framework for survival analysis [@hanley2009].
-In case-base sampling, we sample from the continuous survival time of each person in our dataset to create a *base series* of *person-moments*.
-This *base series* complements the *case series*, which contains all person-moments at which the event of interest occurs.
+\keywords{survival analysis, machine learning, case-base, neural
+network}
 
-For each person-moment sampled, let $X_i$ be the corresponding covariate profile $\left(x_{i1},x_{i2},...,x_{ip} \right)$, $T_i$ be the time of the person-moment, and $Y_i$ be the indicator variable for whether the event of interest occurred at time $T_i$.
-We estimate the hazard function $h(t \mid X_i)$ using the sampled person-moments.
-Recall that $h(t \mid X_i)$ is the instantaneous potential of experiencing the event at time $t$ for a given set of covariates $X_i$, assuming $T_i \geq t$.
+\maketitle
 
-Now, let $b$ be the (user-defined) size of the *base series*, and let $B$ be the sum of all follow-up times for the individuals in the study.
-If we sample the *base series* uniformly, then the hazard function of the sampling process is equal to $b/B$.
-Therefore, we have the following equality \footnote{We are abusing notation here, conflating hazards with probabilities.
-For a rigorous treatment, see Saarela \& Hanley (2015) section 3 \cite{saarela2015} .}: \begin{align}\label{eqn:main}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+The widespread use of the Cox proportional hazards model has led to a
+focus on survival models based on hazard ratios and relative risks
+rather than on survival curves and absolute risks \citep{hanley2009}. In
+clinical settings, Cox models are frequently prioritized over
+smooth-in-time accelerated failure time (AFT) models \citep{hanley2009},
+which can estimate absolute risks by modeling the hazard directly
+through user specification of a baseline hazard distribution. The user
+may forgo this choice by implementing a flexible set of parameters using
+splines, as proposed by Royston and Palmar \citep{royston2002flexible}.
+These regression models also require user-specified interactions, which
+may limit risk prediction accuracy.
+
+Over longer follow-up, it is challenging to maintain the proportional
+hazards assumption as the risk associated with each covariate may not be
+constant with time. Previous studies related to breast cancer have
+discovered time-varying interactions with covariates of interest, like
+tumor size \citep{coradini2000time} \citep{hilsenbeck1998time}. Another
+limitation of regression-based models is the requirement of prior
+knowledge; the analyst must know potential time-varying interactions and
+explicitly evaluate them.
+
+Data-driven neural network approaches bypass the need for prior
+knowledge of interaction terms. DeepSurv is a semi-parametric
+proportional hazards neural network model that implements the Cox
+partial log-likelihood as a custom loss function
+\citep{katzman2018DeepSurv}. This results in a stepwise absolute risk
+curve that cannot accommodate time-varying interactions. Compared to Cox
+regression, DeepSurv has shown a small improvement in performance using
+a single layer neural network on the Study to Understand Prognoses
+Preferences and Risk Treatments (SUPPORT) dataset
+\citep{knaus1995SUPPORT}. Farraggi et al.~suggested a modification to
+the loss function required for a Cox neural network that can model
+non-proportional hazards \citep{faraggi1995neural}. Wang et al.~took the
+concept of extreme learning machines and applied them in a survival
+context, removing the need for tuning the number of layers and nodes
+\citep{wang2019extreme}. A Cox-based neural network combined with Local
+Interpretable Model-Agnostic Explanations (LIME)
+\citep{ribeiro2016model} interprets the association of each Single
+Nucleotide Polymorphism (SNP) \citep{sun2020genome} to a phenotype.
+
+DeepHit requires each survival time of interest to be specified in the
+model and directly estimates survival curves, rather than deriving a
+hazard function \citep{lee2018DeepHit}. It assumes an inverse Gaussian
+distribution as the baseline hazard and its concordance index (c-index)
+score outperforms DeepSurv \citep{katzman2018DeepSurv} on the Molecular
+Taxonomy of Breast Cancer International Consortium (METABRIC) dataset
+\citep{curtis2012genomic}.
+
+Ripley et al.~presented seven different distributions and modeling
+paradigms based on neural networks \citep{ripley2004non}. They provided
+the methodology behind parametric log-logistic and log-normal neural
+network models. Like DeepHit, these models are restricted to one of the
+seven baseline hazard distributions, reducing the flexibility of the
+average survival prediction.
+
+Deep Survival Machines (DSM) provides a flexible, parametric survival
+model using neural networks with a mixture of distributions as the
+baseline hazard \citep{dsmPaper}. DSM allows the user to specify a
+mixture of distributions for a flexible baseline hazard. On both the
+SUPPORT and METABRIC datasets, DSM outperforms DeepSurv and DeepHit
+\citep{dsmPaper}. However, like DeepHit and DeepSurv, DSM cannot model
+time-varying interactions.
+
+These methods need custom loss and activation functions that are
+difficult to implement in practice. We note the complexity of these
+alternative neural network approaches; DeepSurv, DeepHit, and DSM all
+require custom loss functions \citep{katzman2018DeepSurv}
+\citep{lee2018DeepHit} \citep{dsmPaper}, while DSM requires user
+implementations for distributions beyond Log-Normal or Weibull and a
+two-phase learning process \citep{dsmPaper}. Regression-based approaches
+require prior specification of all interaction terms, which makes it
+challenging to model covariate effects that change over time. In this
+article, we propose Case-Base Neural Networks (CBNN) as a method that
+models time-varying interactions and a flexible baseline hazard using
+commonly available neural network components. Our approach to modeling
+the full hazard uses case-base sampling \citep{hanley2009}, which allows
+probabilistic models to predict survival outcomes. Following the
+case-base framework, we can use transformations of time as a feature
+(covariate) to specify different baseline hazards. For example, by
+including splines of time as covariates, we can approximate the
+Royston-Parmar flexible baseline hazard model
+\citep{royston2002flexible}. An alternative approach to Royston-Parmar
+is case-base with logistic regression (CBLR), which provides some
+flexibility in modeling, but still requires user specification of
+time-varying interactions and baseline hazard, while CBNN can model both
+without user specification; only a sufficiently deep network is
+required.
+
+In Section \ref{methods}, we describe how case-base sampling and neural
+networks are combined both conceptually and algebraically, along with
+our hyperparameter choice and software implementation. Then, in Section
+\ref{sims}, we describe our metrics and compare the performance of CBNN,
+DeepSurv, DeepHit, DSM, Cox regression, and case-base using logistic
+regression on simulated data. Next, Section \ref{casestudies} describes
+the real-data analysis, while in Section \ref{discussion}, we explore
+the implications of our results and contextualize them within neural
+network survival analysis in a single event setting.
+
+\hypertarget{methods}{%
+\section{Case-base neural network methodology, metrics, and
+software}\label{methods}}
+
+In this section, we define case-base sampling, which converts the total
+survival time into discrete person-specific moments (person-moments).
+Then, we detail how neural networks can be used within this framework,
+explicitly incorporating time as a feature while adjusting for the
+sampling bias. Finally, we report on the software versions used. A
+package is available for use at {[}GITHUBLINK{]}. The entire code base
+to generate the analyses in this paper is available at {[}GITHUB
+LINK{]}.
+
+\hypertarget{case-base-sampling}{%
+\subsection{Case-base sampling}\label{case-base-sampling}}
+
+Case-base sampling is an alternative framework for survival analysis
+\citep{hanley2009}. In case-base sampling, we sample from the continuous
+survival time of each person in our dataset to create a \emph{base
+series} of \emph{person-moments}. This \emph{base series} complements
+the \emph{case series}, which contains all person-moments at which the
+event of interest occurs.
+
+For each person-moment sampled, let \(X_i\) be the corresponding
+covariate profile \(\left(x_{i1},x_{i2},...,x_{ip} \right)\), \(T_i\) be
+the time of the person-moment, and \(Y_i\) be the indicator variable for
+whether the event of interest occurred at time \(T_i\). We estimate the
+hazard function \(h(t \mid X_i)\) using the sampled person-moments.
+Recall that \(h(t \mid X_i)\) is the instantaneous potential of
+experiencing the event at time \(t\) for a given set of covariates
+\(X_i\), assuming \(T_i \geq t\).
+
+Now, let \(b\) be the (user-defined) size of the \emph{base series}, and
+let \(B\) be the sum of all follow-up times for the individuals in the
+study. If we sample the \emph{base series} uniformly, then the hazard
+function of the sampling process is equal to \(b/B\). Therefore, we have
+the following equality
+\footnote{We are abusing notation here, conflating hazards with probabilities. For a rigorous treatment, see Saarela \& Hanley (2015) section 3 \cite{saarela2015} .}:
+\begin{align}\label{eqn:main}
 \frac{P\left(Y_i=1 \mid X_i, T_i\right)}{P\left(Y_i = 0 \mid X_i, T_i\right)} = \frac{h\left(T_i \mid X_i\right)}{b/B}.
-\end{align} 
-The odds of a person-moment being a part of the *case series* is the ratio of the hazard $h(T_i \mid X_i)$ and the uniform rate $b/B$.
-Using \eqref{eqn:main}, we can see how the log-hazard function can be estimated from the log-odds arising from case-base sampling: \begin{align}\label{eqn:offset}
+\end{align} The odds of a person-moment being a part of the \emph{case
+series} is the ratio of the hazard \(h(T_i \mid X_i)\) and the uniform
+rate \(b/B\). Using \eqref{eqn:main}, we can see how the log-hazard
+function can be estimated from the log-odds arising from case-base
+sampling: \begin{align}\label{eqn:offset}
 \log \left( h\left(t \mid X_i\right)\right) = \log \left(\frac{P\left(Y_i = 1 \mid X_i, t\right)}{P\left(Y_i = 0 \mid X_i, t\right)}\right) + \log\left(\frac{b}{B}\right).
@@
 
-We may estimate the hazard function if we adjust for the bias introduced when sampling a fraction of the study base $B$ by adding the offset term $\log\left(\frac{b}{B} \right)$ (\eqref{eqn:offset}).
-Next, we propose using neural networks to model the odds.
+We may estimate the hazard function if we adjust for the bias introduced
+when sampling a fraction of the study base \(B\) by adding the offset
+term \(\log\left(\frac{b}{B} \right)\) (\eqref{eqn:offset}). Next, we
+propose using neural networks to model the odds.
 
-## Neural networks to model the hazard function
+\hypertarget{neural-networks-to-model-the-hazard-function}{%
+\subsection{Neural networks to model the hazard
+function}\label{neural-networks-to-model-the-hazard-function}}
 
-After case-base sampling, we pass all features, including time, into any user-defined feed-forward component, to which an offset term (to adjust for bias from case-base sampling) is added, then passed through a sigmoid activation function (Figure \ref{fig:NNarch}).
-We use the sigmoid function as its inverse is the odds, which we can use to calculate the hazard.
-The general form for the neural network using CBNN is:
+After case-base sampling, we pass all features, including time, into any
+user-defined feed-forward component, to which an offset term (to adjust
+for bias from case-base sampling) is added, then passed through a
+sigmoid activation function (Figure \ref{fig:NNarch}). We use the
+sigmoid function as its inverse is the odds, which we can use to
+calculate the hazard. The general form for the neural network using CBNN
+is:
 
 \begin{align}\label{eqn:nnProb}
-P\left(Y=1|X,T\right)&=\sigmoid\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right) \right),
+P\left(Y=1|X,T\right)&=\mathrm{sigmoid}\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right) \right),
 \end{align}
 
-where $T$ is a random variable representing the event time, $X$ is the random variable for a covariate profile, $f_{\theta}(X, T)$ represents any feed-forward neural network architecture, $\log\left(\frac{B}{b}\right)$ is bias term set by case-base sampling, $\theta$ is the set of parameters learned by the neural network and $\sigmoid(x)=\frac{1}{1+e^{-x}}$.
-By allowing a neural network to approximate a higher-order polynomial of time, the baseline hazard specification is now data-driven, where user-defined hyperparameters such as regularization, number of nodes, and layers control the flexibility of the model.
-We provide a detailed description of the choices we made in the next sub-section.
+where \(T\) is a random variable representing the event time, \(X\) is
+the random variable for a covariate profile, \(f_{\theta}(X, T)\)
+represents any feed-forward neural network architecture,
+\(\log\left(\frac{B}{b}\right)\) is bias term set by case-base sampling,
+\(\theta\) is the set of parameters learned by the neural network and
+\(\mathrm{sigmoid}(x)=\frac{1}{1+e^{-x}}\). By allowing a neural network
+to approximate a higher-order polynomial of time, the baseline hazard
+specification is now data-driven, where user-defined hyperparameters
+such as regularization, number of nodes, and layers control the
+flexibility of the model. We provide a detailed description of the
+choices we made in the next sub-section.
 
-```{r NNarch,echo = FALSE, fig.align = 'center', out.width = "100%", fig.cap = "Steps involved in CBNN from case-base sampling to the model framework we use for training.
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{/home/bhatnagar-lab/jislam/pmnnProjects/paper/figures/nnarch2}
+
+}
+
+\caption{Steps involved in CBNN from case-base sampling to the model framework we use for training.
 The first step is case-base sampling, completed before training begins.
@@
 We add the offset and pass that through a sigmoid activation function, whose output is a probability.
-Once the neural network model completes its training, we can convert the probability output to a hazard, using it for our survival outcomes of interest."}
-library(knitr)
-knitr::include_graphics(here::here("figures", "nnarch2.PNG"))
-```
+Once the neural network model completes its training, we can convert the probability output to a hazard, using it for our survival outcomes of interest.}\label{fig:NNarch}
+\end{figure}
 
-The following derivation shows how our probability estimate is converted to odds:
-\begin{align*}
- \log\left( h(t \mid X_i) \right) &= \log\left(\frac{\sigmoid \left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}{1-\sigmoid \left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}\right) + \log\left(\frac{b}{B}\right) \\
+The following derivation shows how our probability estimate is converted
+to odds: \begin{align*}
+ \log\left( h(t \mid X_i) \right) &= \log\left(\frac{\mathrm{sigmoid}\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}{1-\mathrm{sigmoid}\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}\right) + \log\left(\frac{b}{B}\right) \\
  &= \log\left( \frac{\frac{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)+1}}{1-\frac{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)+1}}\right) + \log\left(\frac{b}{B}\right) \\
@@
 
-We use binary cross-entropy as our loss function [@gulli2017]:
+We use binary cross-entropy as our loss function \citep{gulli2017}:
 \begin{align*}
 L(\theta)=-\frac{1}{N} \sum^{N}_{i=1} y_{i} \cdot \log(\hat{f}_{\theta}(x_{i}, t_{i}) ) + (1-y_{i} )\cdot \log(1-\hat{f}_{\theta}(x_{i}, t_{i}) ),
-\end{align*}
-where $\hat{f}_{\theta}(x_{i}, t_{i})$ is our estimate for a given covariate profile and time, $y_{i}$ is our target value specifying whether an event occurred, and $N$ represents the number of individuals in our training set.
+\end{align*} where \(\hat{f}_{\theta}(x_{i}, t_{i})\) is our estimate
+for a given covariate profile and time, \(y_{i}\) is our target value
+specifying whether an event occurred, and \(N\) represents the number of
+individuals in our training set.
 
-Backpropagation is used to optimize the parameters in the model with an appropriate minimization algorithm (e.g.
-Adam, RMSPropagation, stochastic gradient descent)[@gulli2017].
-For our analysis, we use Adam.
-Note that the size of the *case series* is fixed as the number of events, but we can make the *base series* as large as we want.
-A ratio of 100:1 *base series* to *case series* is sufficient [@hanley2009].
-We pass our feed-forward neural network through a sigmoid activation function (Figure \ref{fig:NNarch}).
-Finally, we can convert this model output to a hazard.
-When using our model for predictions, we manually set the offset term to 0 in the new data, as we account for the bias during the fitting process.
+Backpropagation is used to optimize the parameters in the model with an
+appropriate minimization algorithm (e.g.~Adam, RMSPropagation,
+stochastic gradient descent)\citep{gulli2017}. For our analysis, we use
+Adam. Note that the size of the \emph{case series} is fixed as the
+number of events, but we can make the \emph{base series} as large as we
+want. A ratio of 100:1 \emph{base series} to \emph{case series} is
+sufficient \citep{hanley2009}. We pass our feed-forward neural network
+through a sigmoid activation function (Figure \ref{fig:NNarch}).
+Finally, we can convert this model output to a hazard. When using our
+model for predictions, we manually set the offset term to 0 in the new
+data, as we account for the bias during the fitting process.
 
-Since we are directly modeling the hazard, we can readily estimate the risk function ($F$) at time $t$ for a covariate profile $X$, viz.\begin{align}\label{eqn:ci2}
+Since we are directly modeling the hazard, we can readily estimate the
+risk function (\(F\)) at time \(t\) for a covariate profile \(X\),
+viz.\begin{align}\label{eqn:ci2}
 F\left(t|X\right)& = 1 - \exp\left(-\int_{0}^{t}h(u|X) \,\textrm du\right).
-\end{align} 
-We use Riemann's Sum [@hughes2020calculus] to approximate the integral in \eqref{eqn:ci2}.
+\end{align} We use Riemann's Sum \citep{hughes2020calculus} to
+approximate the integral in \eqref{eqn:ci2}.
 
-## Hyperparameter selection
+\hypertarget{hyperparameter-selection}{%
+\subsection{Hyperparameter selection}\label{hyperparameter-selection}}
 
-Neural networks provide flexibility when defining the architecture and optimization parameters of our models.
-These hyperparameter choices can affect the estimated parameters and were chosen during a set of initial simulations to determine if CBNN can learn complex interactions in practice.
-We set a max number of epochs to be $2000$, batch size as $512$, learning rate as $10^{-3}$, decay as $10^{-7}$, patience as $10$ epochs, $\{50,50,25,25\}$ nodes in each hidden layer with $50\%$ dropout at each layer, a validation split of $20\%$, a testing split of $10%$, a minimum delta loss on the validation set of $10^{-7}$ over 10 epochs, and Adam [@gulli2017] as our optimizer.
-These choices may permit the model to approximate higher-order interactions while preventing over-fitting [@srivastava2014dropout].
-We fix a train-validation-test split that allows us to update the weights with a subset of the data (training), assess performance at each epoch (validation) and gauge performance of the final model (test).
-We select the best weights after training based on validation loss [@gulli2017].
+Neural networks provide flexibility when defining the architecture and
+optimization parameters of our models. These hyperparameter choices can
+affect the estimated parameters and were chosen during a set of initial
+simulations to determine if CBNN can learn complex interactions in
+practice. We set a max number of epochs to be \(2000\), batch size as
+\(512\), learning rate as \(10^{-3}\), decay as \(10^{-7}\), patience as
+\(10\) epochs, \(\{50,50,25,25\}\) nodes in each hidden layer with
+\(50\%\) dropout at each layer, a validation split of \(20\%\), a
+testing split of \(10%
+\), a minimum delta loss on the validation set of \(10^{-7}\) over 10
+epochs, and Adam \citep{gulli2017} as our optimizer. These choices may
+permit the model to approximate higher-order interactions while
+preventing over-fitting \citep{srivastava2014dropout}. We fix a
+train-validation-test split that allows us to update the weights with a
+subset of the data (training), assess performance at each epoch
+(validation) and gauge performance of the final model (test). We select
+the best weights after training based on validation loss
+\citep{gulli2017}.
 
-## Software implementation
+\hypertarget{software-implementation}{%
+\subsection{Software implementation}\label{software-implementation}}
 
-R [@Rsoft] and python [@py] are used to evaluate methods from both languages.
-We fit the Cox model using the **survival** package [@survpkg], the CBLR model using the **casebase** package [@cbpkg], the DeepSurv model using the **survivalmodels** package [@survmods], the DSM model using **DeepSurvivalMachines** [@dsmPaper] and the DeepHit model using **pycox** [@lee2018DeepHit].
-We made the components of CBNN using the **keras** [@keras] package and the **casebase** package for the sampling step.
-The **simsurv** package [@simsurv] is used to simulate our simulation studies, while **flexsurv** [@flexsurv] is used to fit a flexible baseline hazard using splines for our complex simulation.
-We use the implementation of c-ipcw from the python package **sksurv** [@sksurv].
-The **riskRegression** package [@riskRegression] is used to get the Brier score and IPA metrics.
-We changed the **riskRegression** package to be used with any user supplied risk function $F$.
-To ensure that both R and Python-based models are running in unison on the same data through our simulations and bootstrap, we use the **reticulate** package [@reticulate].
+R \citep{Rsoft} and python \citep{py} are used to evaluate methods from
+both languages. We fit the Cox model using the \textbf{survival} package
+\citep{survpkg}, the CBLR model using the \textbf{casebase} package
+\citep{cbpkg}, the DeepSurv model using the \textbf{survivalmodels}
+package \citep{survmods}, the DSM model using
+\textbf{DeepSurvivalMachines} \citep{dsmPaper} and the DeepHit model
+using \textbf{pycox} \citep{lee2018DeepHit}. We made the components of
+CBNN using the \textbf{keras} \citep{keras} package and the
+\textbf{casebase} package for the sampling step. The \textbf{simsurv}
+package \citep{simsurv} is used to simulate our simulation studies,
+while \textbf{flexsurv} \citep{flexsurv} is used to fit a flexible
+baseline hazard using splines for our complex simulation. We use the
+implementation of c-ipcw from the python package \textbf{sksurv}
+\citep{sksurv}. The \textbf{riskRegression} package
+\citep{riskRegression} is used to get the Brier score and IPA metrics.
+We changed the \textbf{riskRegression} package to be used with any user
+supplied risk function \(F\). To ensure that both R and Python-based
+models are running in unison on the same data through our simulations
+and bootstrap, we use the \textbf{reticulate} package
+\citep{reticulate}.
 
-# Simulation studies {#sims}
+\hypertarget{sims}{%
+\section{Simulation studies}\label{sims}}
 
-In this section, we use simulated data to evaluate the performance of CBNN and compare our approach with existing methods based on regression (Cox, CBLR) and neural network (DeepHit, DeepSurv, DSM) methods.
-We specify a linear combination of each covariate as the linear predictor in regression-based approaches (Cox, CBLR), which contrasts with neural network approaches that allow for non-linear interactions.
-We simulate data under a simple exponential model, and a complex baseline hazard with time-varying interactions, each with 10% random censoring.
-For both settings, we simulate three covariates:
+In this section, we use simulated data to evaluate the performance of
+CBNN and compare our approach with existing methods based on regression
+(Cox, CBLR) and neural network (DeepHit, DeepSurv, DSM) methods. We
+specify a linear combination of each covariate as the linear predictor
+in regression-based approaches (Cox, CBLR), which contrasts with neural
+network approaches that allow for non-linear interactions. We simulate
+data under a simple exponential model, and a complex baseline hazard
+with time-varying interactions, each with 10\% random censoring. For
+both settings, we simulate three covariates:
 
-$$
+\[
 z_{1} \sim \textrm{Bernoulli}(0.5) \qquad \qquad
@@
 \end{cases}
-$$
+\]
 
-The DeepHit-specific hyperparameter alpha is set to $0.5$ (equal weight between its negative log-likelihood and ranking losses [@lee2018DeepHit]).
-We change the **DeepSurvivalMachines** [@dsmPaper] package to include dropout and a minimum delta loss during the fitting process.
-For DSM, we define a mixture model of six Weibull distributions for the baseline hazard.
-All other hyperparameters are held constant across all neural network methods in both the simulation studies and real data applications.
+The DeepHit-specific hyperparameter alpha is set to \(0.5\) (equal
+weight between its negative log-likelihood and ranking losses
+\citep{lee2018DeepHit}). We change the \textbf{DeepSurvivalMachines}
+\citep{dsmPaper} package to include dropout and a minimum delta loss
+during the fitting process. For DSM, we define a mixture model of six
+Weibull distributions for the baseline hazard. All other hyperparameters
+are held constant across all neural network methods in both the
+simulation studies and real data applications.
 
-Besides the methods mentioned above, we include the Optimal model in our comparisons using CBLR.
-That is, we include the exact functional form of the covariates in a CBLR model (referred to as Optimal for simplicity).
-We calculate $t$-based 95% confidence intervals using 100 replications of the simulated data For all analyses, we use 80% for training and 20% for the test set.
-20% of the training set is kept for validation at each epoch.
-We predict risk functions $F$ using \eqref{eqn:ci2} for individuals in the test set, which are used to calculate our c-ipcw and IPA scores.
+Besides the methods mentioned above, we include the Optimal model in our
+comparisons using CBLR. That is, we include the exact functional form of
+the covariates in a CBLR model (referred to as Optimal for simplicity).
+We calculate \(t\)-based 95\% confidence intervals using 100
+replications of the simulated data For all analyses, we use 80\% for
+training and 20\% for the test set. 20\% of the training set is kept for
+validation at each epoch. We predict risk functions \(F\) using
+\eqref{eqn:ci2} for individuals in the test set, which are used to
+calculate our c-ipcw and IPA scores.
 
-## Performance metrics
+\hypertarget{performance-metrics}{%
+\subsection{Performance metrics}\label{performance-metrics}}
 
-We use two metrics to assess the performance of the different methods of interest: 1) index of prediction accuracy (IPA) [@kattan2018index] and 2) inverse probability censoring weights-adjusted concordance index (c-ipcw) [@uno2011], which we define below.
-Both these time-dependent metrics provide transparency as to when in follow-up time each model may perform better than the others.
+We use two metrics to assess the performance of the different methods of
+interest: 1) index of prediction accuracy (IPA) \citep{kattan2018index}
+and 2) inverse probability censoring weights-adjusted concordance index
+(c-ipcw) \citep{uno2011}, which we define below. Both these
+time-dependent metrics provide transparency as to when in follow-up time
+each model may perform better than the others.
 
-### Index of prediction accuracy (IPA)
+\hypertarget{index-of-prediction-accuracy-ipa}{%
+\subsubsection{Index of prediction accuracy
+(IPA)}\label{index-of-prediction-accuracy-ipa}}
 
-The IPA is a function of the Brier score ($BS(t)$) [@graf1999], which is defined as
-\begin{align}
+The IPA is a function of the Brier score (\(BS(t)\)) \citep{graf1999},
+which is defined as \begin{align}
 BS(t)=\frac{1}{N}\sum^{N}_{i=1}\left(\frac{\left(\widehat{F}(X_{i},t)-1\right)^{2}\cdot I_{T_{i}\leq t,\delta_{i}=1}}{\widehat{G}(T_{i})} + \frac{\left(\widehat{F}(X_{i},t)\right)^{2}\cdot I_{T_{i}>t}}{\widehat{G}(t)}\right),
-\end{align}
-where $\delta_{i}=1$ shows individuals who have experienced the event, $N$ represents the number of samples in our dataset over which we calculate $BS(t)$, $\widehat{G}(t)=P[C>t]$ is a non-parametric estimate of the censoring distribution, $C$ is censoring time and $T_{i}$ is an individual’s survival or censoring time.
-The Brier score provides a score that accounts for the information loss because of censoring.
-There are three categories of individuals that may appear within the dataset once we fix our $t$ of interest.
-Individuals who experienced the event before $t$ are present in the first term of the equation.
-The second term of the equation includes individuals who experience the event or are censored after $t$.
-Those censored before $t$ are the third category of people.
-The  inverse probability censoring weights (IPCW) adjustment ($G(\cdot)$) is to account for these category three individuals whose information is missing.
-The IPA score as a function of time is given by
-\begin{align}
+\end{align} where \(\delta_{i}=1\) shows individuals who have
+experienced the event, \(N\) represents the number of samples in our
+dataset over which we calculate \(BS(t)\), \(\widehat{G}(t)=P[C>t]\) is
+a non-parametric estimate of the censoring distribution, \(C\) is
+censoring time and \(T_{i}\) is an individual's survival or censoring
+time. The Brier score provides a score that accounts for the information
+loss because of censoring. There are three categories of individuals
+that may appear within the dataset once we fix our \(t\) of interest.
+Individuals who experienced the event before \(t\) are present in the
+first term of the equation. The second term of the equation includes
+individuals who experience the event or are censored after \(t\). Those
+censored before \(t\) are the third category of people. The inverse
+probability censoring weights (IPCW) adjustment (\(G(\cdot)\)) is to
+account for these category three individuals whose information is
+missing. The IPA score as a function of time is given by \begin{align}
 \textrm{IPA}(t) &= 1-\frac{BS_{model}(t)}{BS_{null}(t)}, \nonumber
-\end{align} where $BS_{model}(t)$ represents the Brier score over time $t$ for the model of interest, and $BS_{null}(t)$ represents the Brier score if we use an unadjusted Kaplan-Meier (KM) curve as the prediction for all observations [@kattan2018index].
-Note that IPA has an upper bound of one, where positive values show an increase in performance over the null model and negative values show that the null model performs better.
-These scores demonstrate how performance changes over follow-up time.
+\end{align} where \(BS_{model}(t)\) represents the Brier score over time
+\(t\) for the model of interest, and \(BS_{null}(t)\) represents the
+Brier score if we use an unadjusted Kaplan-Meier curve as the prediction
+for all observations \citep{kattan2018index}. Note that IPA has an upper
+bound of one, where positive values show an increase in performance over
+the null model and negative values show that the null model performs
+better. These scores demonstrate how performance changes over follow-up
+time.
 
-A potential artifact of IPA is that the score is unstable at earlier and later survival times.
-This is because of a near equivalent Brier score between each model and the null model.
-At small values, a difference of $0.1$ creates a more significant fold change than at larger values.
-As the Brier score is potentially small at the start and at the end of their respective curves, we may see instability of the IPA score at those ends.
+A potential artifact of IPA is that the score is unstable at earlier and
+later survival times. This is because of a near equivalent Brier score
+between each model and the null model. At small values, a difference of
+\(0.1\) creates a more significant fold change than at larger values. As
+the Brier score is potentially small at the start and at the end of
+their respective curves, we may see instability of the IPA score at
+those ends.
 
-### Inverse probability censoring weights-adjusted concordance index
+\hypertarget{inverse-probability-censoring-weights-adjusted-concordance-index}{%
+\subsubsection{Inverse probability censoring weights-adjusted
+concordance
+index}\label{inverse-probability-censoring-weights-adjusted-concordance-index}}
 
-The c-ipcw is a non-proper, rank-based metric that does not depend on the censoring times in the test data [@uno2011]. The c-ipcw is given by
+The c-ipcw is a non-proper, rank-based metric that does not depend on
+the censoring times in the test data \citep{uno2011}. The c-ipcw is
+given by
 
@@
 c-ipcw(t) &= \frac{\sum^{N}_{i=1}\sum^{N}_{j=1}\delta_{i}\left\{\widehat{G}(T_{i})\right\}^{-2} I(T_{i}<T_{j},T_{i}<t) I\left(\widehat{F}(t|X_{i})>\widehat{F}(t|X_{j})\right)}{\sum^{N}_{i=1}\sum^{N}_{j=1}\delta_{i}\left\{\widehat{G}(T_{i})\right\}^{-2} I(T_{i}<T_{j},T_{i}<t)},
-\end{align} where the follow-up period of interest is (0,$t$), $I(\cdot)$ is an indicator function, $\widehat{F}(X_{i},t)$ is the risk function estimated for everyone in the study at time $t$ and c-ipcw can compare the performance of different models, where a higher score is better.
-Note that the c-ipcw may produce misleading performance, as it ranks based on survival times, not event status [@cindexfails2019].
-This metric is considered an unbiased population concordance measure because of the IPCW adjustment [@uno2011].
+\end{align} where the follow-up period of interest is (0,\(t\)),
+\(I(\cdot)\) is an indicator function, \(\widehat{F}(X_{i},t)\) is the
+risk function estimated for everyone in the study at time \(t\) and
+c-ipcw can compare the performance of different models, where a higher
+score is better. Note that the c-ipcw may produce misleading
+performance, as it ranks based on survival times, not event status
+\citep{cindexfails2019}. This metric is considered an unbiased
+population concordance measure because of the IPCW adjustment
+\citep{uno2011}.
 
@@
 
-```{r megaPlot, echo = FALSE , fig.align = 'center', out.width = "100%", fig.cap = "Summarizes the simple simulation (A, B), complex simulation (C, D), SUPPORT case study (E, F) and METABRIC case study (G, H) results.
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{/home/bhatnagar-lab/jislam/pmnnProjects/paper/../pmnnPlayground/reviewedAnalysisPMNN/figures/megaPlot}
+
+}
+
+\caption{Summarizes the simple simulation (A, B), complex simulation (C, D), SUPPORT case study (E, F) and METABRIC case study (G, H) results.
 The first row shows the IPA score for each model in each study over follow-up time.
@@
 Each model-specific metric in each study shows a 95-percent confidence interval over 100 iterations.
-The models of interest are case-base with logistic regression (CBLR), Case-Base Neural Networks (CBNN), Cox, DeepHit, DeepSurv, Deep Survival Machines (DSM), Optimal (a CBLR model with the exact interaction terms and baseline hazard specified) and Kaplan-Meier (to serve as a baseline, predicting the average for all individuals)."}
-#readRDS(here::here("figures", "megaPlot.rds"))
-knitr::include_graphics(here::here("../pmnnPlayground/reviewedAnalysisPMNN/figures", "megaPlot.pdf"))
-```
+The models of interest are case-base with logistic regression (CBLR), Case-Base Neural Networks (CBNN), Cox, DeepHit, DeepSurv, Deep Survival Machines (DSM), Optimal (a CBLR model with the exact interaction terms and baseline hazard specified) and Kaplan-Meier (to serve as a baseline, predicting the average for all individuals).}\label{fig:megaPlot}
+\end{figure}
 
-## Simple simulation: constant baseline hazard
+\hypertarget{simple-simulation-constant-baseline-hazard}{%
+\subsection{Simple simulation: constant baseline
+hazard}\label{simple-simulation-constant-baseline-hazard}}
 
-We simulate data from a simple model that depends on a constant baseline hazard:
-$$h(t \mid X_i) = \lambda\cdot e^{\beta_{{1}}z_{1}+\beta_{{2}}z_{2}+\beta_{{3}}z_{3}},$$
-with minimal covariate effects given by $\beta_{{1}}=0.1, \beta_{{2}}=0.1, \beta_{{3}}=0.1$ and $\lambda=1.0$. Once we simulate survival times, we introduce 10% random censoring.
+We simulate data from a simple model that primarily depends on a
+constant baseline hazard:
+\[h(t \mid X_i) = \lambda\cdot e^{\beta_{{1}}z_{1}+\beta_{{2}}z_{2}+\beta_{{3}}z_{3}},\]
+with minimal covariate effects given by
+\(\beta_{{1}}=0.1, \beta_{{2}}=0.1, \beta_{{3}}=0.1\) and
+\(\lambda=1.0\). Once we simulate survival times, we introduce 10\%
+random censoring.
 
-### Performance comparison in simple simulation
+\hypertarget{performance-comparison-in-simple-simulation}{%
+\subsubsection{Performance comparison in simple
+simulation}\label{performance-comparison-in-simple-simulation}}
 
-Figure \ref{fig:megaPlot} A, B and Table \ref{tab:megaTable} A show the results for the simple simulation.
-The regression-based methods (CBLR, Cox, Optimal) outperform the neural network ones in the simple simulation setting.
-Among the neural network approaches, CBNN outperforms all other methods in terms of both IPA and c-ipcw (Figure \ref{fig:megaPlot} A, B).
-Specifically, we see CBNN is consistent across time with smaller confidence intervals compared to DeepHit, DeepSurv and DSM.
+Figure \ref{fig:megaPlot} A, B and Table \ref{tab:megaTable} A show the
+results for the simple simulation. The regression-based methods (CBLR,
+Cox, Optimal) outperform the neural network ones in the simple
+simulation setting. Among the neural network approaches, CBNN
+outperforms all other methods in terms of both IPA and c-ipcw (Figure
+\ref{fig:megaPlot} A, B). Specifically, we see CBNN is consistent across
+time with smaller confidence intervals compared to DeepHit, DeepSurv and
+DSM.
 
-In this simple setting, the regression models are much closer to the Optimal model, while the neural network models perform worse than the null model.
-The wide confidence bands in c-ipcw suggest the neural network models may be over-parameterized.
+In this simple setting, the regression models are much closer to the
+Optimal model, while the neural network models perform worse than the
+null model. The wide confidence bands in c-ipcw suggest the neural
+network models may be over-parameterized.
 
-## Complex simulation: flexible baseline hazard, time-varying interactions
+\hypertarget{complex-simulation-flexible-baseline-hazard-time-varying-interactions}{%
+\subsection{Complex simulation: flexible baseline hazard, time-varying
+interactions}\label{complex-simulation-flexible-baseline-hazard-time-varying-interactions}}
 
-This simulation demonstrates performance with the presence of a complex baseline hazard and a time-varying interaction.
-Originally used to demonstrate the spline-based hazard model proposed by Royston and Parmar [@royston2002flexible], the breast cancer dataset provides a complex hazard from which we simulate, available in the **flexsurv** R package [@flexsurv].
-To increase the complexity of our data-generating mechanism for this simulation, we design the model as follows:
-\begin{align}
+This simulation demonstrates performance with the presence of a complex
+baseline hazard and a time-varying interaction. Originally used to
+demonstrate the spline-based hazard model proposed by Royston and Parmar
+\citep{royston2002flexible}, the breast cancer dataset provides a
+complex hazard from which we simulate, available in the
+\textbf{flexsurv} R package \citep{flexsurv}. To increase the complexity
+of our data-generating mechanism for this simulation, we design the
+model as follows: \begin{align}
 h(t \mid X_i) =\sum_{i=1}^{5} (\gamma_{i} \cdot \psi_{i}) + \beta_{{1}} (z_{1}) + \beta_{{2}} (z_{2})+ \beta_{{3}} (z_{3})+ \tau_{1} ( z_{1} \cdot z_{2} \cdot t)+ \tau_{2} ( z_{1} \cdot z_{3})+ \tau_{3} (z_{2} \cdot z_{3}), \nonumber
@@
 
-```{=latex}
 \FloatBarrier
@@
 \label{tab:megaTable}
-```
 
-```{r megaTable, echo = FALSE , fig.align = 'center', out.width = "100%"}
-#readRDS(here::here("figures", "megaPlot.rds"))
-knitr::include_graphics(here::here("../pmnnPlayground/reviewedAnalysisPMNN/figures", "megaTable.pdf"))
-```
+\begin{center}\includegraphics[width=1\linewidth]{/home/bhatnagar-lab/jislam/pmnnProjects/paper/../pmnnPlayground/reviewedAnalysisPMNN/figures/megaTable} \end{center}
 
-```{=latex}
 \end{table}
 \FloatBarrier
-```
 
-where $\gamma_{1}=3.9, \gamma_{2}=3, \gamma_{3}=-0.43, \gamma_{4}=1.33,\gamma_{5}=-0.86, \beta_{{1}}=1, \beta_{{2}}=1, \beta_{{3}}=1, \tau_{1}=10, \tau_{2}=2, \tau_{3}=2$ and $\psi$ are basis splines.
-To generate the $gamma$ coefficients, let $g$: $\mathbb{R} \rightarrow \mathbb{R}$ be a smoothing method for variable $t$ by a projection on to a set of basis functions on splines:
+where
+\(\gamma_{1}=3.9, \gamma_{2}=3, \gamma_{3}=-0.43, \gamma_{4}=1.33,\gamma_{5}=-0.86, \beta_{{1}}=1, \beta_{{2}}=1, \beta_{{3}}=1, \tau_{1}=10, \tau_{2}=2, \tau_{3}=2\)
+and \(\psi\) are basis splines. To generate the \(gamma\) coefficients,
+let \(g\): \(\mathbb{R} \rightarrow \mathbb{R}\) be a smoothing method
+for variable \(t\) by a projection on to a set of basis functions on
+splines:
 \srb{i used gt as the function for splines.
@@
 
-An initial model with three knots ($m=5$) is fit using the *flexsurvspline* function with an intercept-only model on the data, which returns the coefficients ($\gamma$) and basis of time ($\psi$) we use in our model.
-Note that we fix these values for the analysis.
-The $\beta$ coefficients represent direct effects, $\tau_{2}$ and ${\tau_3}$ represent interactions and $\tau_{1}$ is a time-varying interaction.
+An initial model with three knots (\(m=5\)) is fit using the
+\emph{flexsurvspline} function with an intercept-only model on the data,
+which returns the coefficients (\(\gamma\)) and basis of time (\(\psi\))
+we use in our model. Note that we fix these values for the analysis. The
+\(\beta\) coefficients represent direct effects, \(\tau_{2}\) and
+\({\tau_3}\) represent interactions and \(\tau_{1}\) is a time-varying
+interaction.
 
-### Performance comparison in complex simulation
+\hypertarget{performance-comparison-in-complex-simulation}{%
+\subsubsection{Performance comparison in complex
+simulation}\label{performance-comparison-in-complex-simulation}}
 
-Figure \ref{fig:megaPlot} C, D and Table \ref{tab:megaTable} B show the performance over time on a test set in the complex simulation.
-Apart from the Optimal regression model, CBNN outperforms the competing models when examining IPA, and up to the 75-percentile of follow-up time for c-ipcw.
-We expected the Optimal model to perform best in both metrics.
-However, this was not the case for c-ipcw and may be due to an artifact of concordance-based metrics, where a misspecified model may perform better than a correctly specified one [@cindexfails2019].
-We attribute this difference in performance to the flexibility of CBNN in terms of time-varying interactions and baseline hazard, flexibility the other neural network models do not have, explaining their drop in performance.
+Figure \ref{fig:megaPlot} C, D and Table \ref{tab:megaTable} B show the
+performance over time on a test set in the complex simulation. Apart
+from the Optimal regression model, CBNN outperforms the competing models
+when examining IPA, and up to the 75-percentile of follow-up time for
+c-ipcw. We expected the Optimal model to perform best in both metrics.
+However, this was not the case for c-ipcw and may be due to an artifact
+of concordance-based metrics, where a misspecified model may perform
+better than a correctly specified one \citep{cindexfails2019}. We
+attribute this difference in performance to the flexibility of CBNN in
+terms of time-varying interactions and baseline hazard, flexibility the
+other neural network models do not have, explaining their drop in
+performance.
 
-# Application to SUPPORT and METABRIC data {#casestudies}
+\hypertarget{casestudies}{%
+\section{Application to SUPPORT and METABRIC data}\label{casestudies}}
 
-Our complex simulation demonstrates the superior performance of CBNN in ideal conditions with clean data.
-To obtain a more realistic performance assessment, we compared models using two real datasets with a time-to-event outcome.
-The first case study examines the SUPPORT dataset [@knaus1995SUPPORT].
-The second case study examines the METABRIC dataset [@curtis2012genomic].
-We use the same hyperparameters as in the simulation studies.
-As we do not know the true model for the real data, we exclude the Optimal model.
-We split the datasets keeping 20% of the observations as a test set.
-20% of the training set is kept aside for validation at each epoch.
-We predict risk functions for everyone in the test set, which are used to calculate our metrics.
-We conduct 100 bootstrap re-samples for the real data applications to obtain confidence intervals.
+Our complex simulation demonstrates the superior performance of CBNN in
+ideal conditions with clean data. To obtain a more realistic performance
+assessment, we compared models using two real datasets with a
+time-to-event outcome. The first case study examines the SUPPORT dataset
+\citep{knaus1995SUPPORT}. The second case study examines the METABRIC
+dataset \citep{curtis2012genomic}. We use the same hyperparameters as in
+the simulation studies. As we do not know the true model for the real
+data, we exclude the Optimal model. We split the datasets keeping 20\%
+of the observations as a test set. 20\% of the training set is kept
+aside for validation at each epoch. We predict risk functions for
+everyone in the test set, which are used to calculate our metrics. We
+conduct 100 bootstrap re-samples for the real data applications to
+obtain confidence intervals.
 
-## Performance evaluation using the SUPPORT dataset
+\hypertarget{performance-evaluation-using-the-support-dataset}{%
+\subsection{Performance evaluation using the SUPPORT
+dataset}\label{performance-evaluation-using-the-support-dataset}}
 
-The SUPPORT dataset tracks the time until death for seriously ill patients at five American hospitals [@knaus1995SUPPORT].
-We use a pre-processed version of the dataset made available in the DeepSurv package [@katzman2018DeepSurv].
-This dataset contains 9104 samples and 14 covariates (age, sex, race, number of comorbidities, presence of diabetes, presence of dementia, presence of cancer, mean arterial blood pressure, heart rate, respiration rate, temperature, white blood cell count, serum’s sodium, and serum’s creatinine) [@katzman2018DeepSurv].
-Patients with missing features were excluded and 68.10% of the patients died during the 5.56-year study period [@katzman2018DeepSurv].
+The SUPPORT dataset tracks the time until death for seriously ill
+patients at five American hospitals \citep{knaus1995SUPPORT}. We use a
+pre-processed version of the dataset made available in the DeepSurv
+package \citep{katzman2018DeepSurv}. This dataset contains 9104 samples
+and 14 covariates (age, sex, race, number of comorbidities, presence of
+diabetes, presence of dementia, presence of cancer, mean arterial blood
+pressure, heart rate, respiration rate, temperature, white blood cell
+count, serum's sodium, and serum's creatinine)
+\citep{katzman2018DeepSurv}. Patients with missing features were
+excluded and 68.10\% of the patients died during the 5.56-year study
+period \citep{katzman2018DeepSurv}.
 
-Figure \ref{fig:megaPlot} E, F and Table \ref{tab:megaTable} C demonstrates the performance over time on a test set.
-The regression models (CBLR, Cox) perform best considering IPA, followed by CBNN from the $25^{th}$ to $100^{th}$ percentile of follow-up time.
-We note a drop in performance for CBNN before the $25^{th}$ percentile of follow-up.
-For c-ipcw, CBNN outperforms the competing models consistently over follow-up time.
-Note that performance is similar for all models, aside from DeepSurv whose c-ipcw is lower than the rest (Figure \ref{fig:megaPlot} E, F and Table \ref{tab:megaTable} C).
+Figure \ref{fig:megaPlot} E, F and Table \ref{tab:megaTable} C
+demonstrates the performance over time on a test set. The regression
+models (CBLR, Cox) perform best considering IPA, followed by CBNN from
+the \(25^{th}\) to \(100^{th}\) percentile of follow-up time. We note a
+drop in performance for CBNN before the \(25^{th}\) percentile of
+follow-up. For c-ipcw, CBNN outperforms the competing models
+consistently over follow-up time. Note that performance is similar for
+all models, aside from DeepSurv whose c-ipcw is lower than the rest
+(Figure \ref{fig:megaPlot} E, F and Table \ref{tab:megaTable} C).
 
-## Performance evaluation using the METABRIC dataset
+\hypertarget{performance-evaluation-using-the-metabric-dataset}{%
+\subsection{Performance evaluation using the METABRIC
+dataset}\label{performance-evaluation-using-the-metabric-dataset}}
 
-METABRIC is a 30-yearlong study aiming to discover the molecular drivers of breast tumors, following 2000 individuals with breast cancer until death [@curtis2012genomic].
-They described these growths as primary invasive breast carcinomas, with the goal of discovering both genetic and clinical risk factors for breast cancer survival [@curtis2012genomic].
-We used the processed dataset made available through DeepSurv [@katzman2018DeepSurv], which includes 1980 patients of which 57.72% die due to breast cancer within a median 10 years of follow-up [@katzman2018DeepSurv].
-There are 11 covariates in total: 4 RNA-Seq gene expressions (MKI67, EGFR, PGR, and ERBB2) and 5 clinical features (hormone treatment indicator, radiotherapy indicator, chemotherapy indicator, ER-positive indicator, and age at diagnosis) [@katzman2018DeepSurv].
+METABRIC is a 30-yearlong study aiming to discover the molecular drivers
+of breast tumors, following 2000 individuals with breast cancer until
+death \citep{curtis2012genomic}. They described these growths as primary
+invasive breast carcinomas, with the goal of discovering both genetic
+and clinical risk factors for breast cancer survival
+\citep{curtis2012genomic}. We used the processed dataset made available
+through DeepSurv \citep{katzman2018DeepSurv}, which includes 1980
+patients of which 57.72\% die due to breast cancer within a median 10
+years of follow-up \citep{katzman2018DeepSurv}. There are 11 covariates
+in total: 4 RNA-Seq gene expressions (MKI67, EGFR, PGR, and ERBB2) and 5
+clinical features (hormone treatment indicator, radiotherapy indicator,
+chemotherapy indicator, ER-positive indicator, and age at diagnosis)
+\citep{katzman2018DeepSurv}.
 
-Figure \ref{fig:megaPlot} G, H and Table \ref{tab:megaTable} D shows the performance on a test set over time.
-The IPA scores suggest that regression models outperform competing models on this dataset, as all the neural network models are equal to or perform worse than KM over follow-up time.
-Our CBNN model is comparable to KM until around the 50-percentile of follow-up time, after which CBNN, DeepSurv and DSM drop in performance.
-The c-ipcw produces a different ranking.
-Our CBNN model outperforms the other models up to the $25^{th}$ percentile of follow-up time, whereas DeepHit performs best from the $50^{th}$ to $75^{th}$ percentile of follow-up.
-With this metric, CBNN and DeepHit outperform the regression models.
-This disagreement between IPA and c-ipcw between may be due to the misspecification issue of concordance-based metrics [@cindexfails2019].
-The neural network models may be over-parameterized, as shown by the wide confidence bands in c-ipcw.
+Figure \ref{fig:megaPlot} G, H and Table \ref{tab:megaTable} D shows the
+performance on a test set over time. The IPA scores suggest that
+regression models outperform competing models on this dataset, as all
+the neural network models are equal to or perform worse than KM over
+follow-up time. Our CBNN model is comparable to KM until around the
+50-percentile of follow-up time, after which CBNN, DeepSurv and DSM drop
+in performance. The c-ipcw produces a different ranking. Our CBNN model
+outperforms the other models up to the \(25^{th}\) percentile of
+follow-up time, whereas DeepHit performs best from the \(50^{th}\) to
+\(75^{th}\) percentile of follow-up. With this metric, CBNN and DeepHit
+outperform the regression models. This disagreement between IPA and
+c-ipcw between may be due to the misspecification issue of
+concordance-based metrics \citep{cindexfails2019}. The neural network
+models may be over-parameterized, as shown by the wide confidence bands
+in c-ipcw.
 
-# Discussion {#discussion}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
-CBNN models survival outcomes by using neural networks on case-base sampled data.
-While estimating the hazard function, we incorporate follow-up-time as a feature, providing a data-driven estimation of a flexible baseline hazard and time-varying interactions.
-Based on our complex simulation results (Figure \ref{fig:megaPlot} C, D and Table \ref{tab:megaTable} B), CBNN outperforms the competitors when time-varying interactions and a complex baseline hazard are present.
+Our method, CBNN, models survival outcomes by using neural networks on
+case-base sampled data. While estimating the hazard function, we
+incorporate follow-up-time as a feature, providing a data-driven
+estimation of a flexible baseline hazard and time-varying interactions.
+Based on our complex simulation results (Figure \ref{fig:megaPlot} C, D
+and Table \ref{tab:megaTable} B), CBNN outperforms the competitors when
+time-varying interactions and a complex baseline hazard are present.
 
-The competing neural network model we evaluated can not model time varying interactions.
-DeepSurv is a proportional hazards model and does not estimate the baseline hazard [@katzman2018DeepSurv].
-DeepHit requires an alpha hyperparameter, is restricted to a single distribution for the baseline hazard, and models the survival function directly [@lee2018DeepHit].
-DSM requires the user to specify the component distributions in the mixture for a flexible baseline hazard to be fit [@dsmPaper].
-These alternative neural network approaches match on time, while CBNN models time directly.
+Each competing neural network model we evaluated has limitations or
+model-specific parameters. DeepSurv is a proportional hazards model and
+does not estimate the baseline hazard \citep{katzman2018DeepSurv}.
+DeepHit requires an alpha hyperparameter, is restricted to a single
+distribution for the baseline hazard, and models the survival function
+directly \citep{lee2018DeepHit}. DSM requires the user to specify the
+component distributions in the mixture for a flexible baseline hazard to
+be fit \citep{dsmPaper}. These alternative neural network approaches
+match on time, while CBNN models time directly.
 
-In the simple simulation, all neural network-based approaches performed worse than KM (IPA scores), while the regression-based approaches did not.
-We attribute this to potential over-parameterization in the neural network models.
-The wide confidence intervals suggest over-fitting may be the reason for the loss in performance, even with dropout.
-Both DeepSurv and DSM are affected, while DeepHit and CBNN are less so.
+In the simple simulation, all neural network-based approaches resulted
+in a loss of performance based on IPA, while the regression-based
+approaches did not. We attribute this to potential over-parameterization
+in the neural network models. The wide confidence intervals suggest
+over-fitting may be the reason for the loss in performance, even with
+dropout. Both DeepSurv and DSM are affected, while DeepHit and CBNN are
+less so.
 
-In the complex simulation, CBNN has a distinct advantage over the other methods while examining the IPA score.
-The regression models do show improvement over the null model, while the competing neural network approaches do not.
-The time-varying interactions and complex baseline hazard may have caused a drop in performance for the competing neural network methods, compared to the regression models without interactions.
+In the complex simulation, CBNN has a distinct advantage over the other
+methods while examining the IPA score. The regression models do show
+improvement over the null model, while the competing neural network
+approaches do not. The time-varying interactions and complex baseline
+hazard may have caused a drop in performance for the competing neural
+network methods, compared to the regression models without interactions.
 
-In the SUPPORT application, the overall performance considering IPA is near identical.
-The regression models outperform CBNN, which performs better than the competing neural network models.
-As the confidence intervals are tight, it is unlikely that the loss in performance is because of over-fitting.
-Flexibility in both interaction modeling and baseline hazard is not particularly helpful with the SUPPORT dataset.
+In the SUPPORT application, the overall performance considering IPA is
+near identical. The regression models outperform CBNN, which performs
+better than the competing neural network models. As the confidence
+intervals are tight, it is unlikely that the loss in performance is
+because of over-fitting. Flexibility in both interaction modeling and
+baseline hazard is not particularly helpful with the SUPPORT dataset.
 
-In the METABRIC application IPA results, DSM and DeepSurv had a steep drop in performance.
-As DeepSurv does not depend on the baseline hazard during the fitting process and DSM is the most flexible competing model in our comparison, the baseline hazard is unlikely to cause this drop.
+In the METABRIC application IPA results, DSM and DeepSurv had a steep
+drop in performance. As DeepSurv does not depend on the baseline hazard
+during the fitting process and DSM is the most flexible competing model
+in our comparison, the baseline hazard is unlikely to cause this drop.
 With tight confidence intervals, over-fitting is unlikely as well.
-Further research would be required to determine the cause of this drop in performance.
-We do not see the same drop in CBNN demonstrating that, in the METABRIC application, CBNN outperforms DeepSurv and DSM.
-Considering IPA score, regression models consistently outperform the neural network models.
-From a biological perspective, the 30-year follow-up time in the METABRIC study may contain competing causes of death.
-The signal from the start of the study may not match the signal towards the end, potentially explaining the drop in performance as we reach later survival times with fewer individuals.
+Further research would be required to determine the cause of this drop
+in performance. We do not see the same drop in CBNN demonstrating that,
+in the METABRIC application, CBNN outperforms DeepSurv and DSM.
+Considering IPA score, regression models consistently outperform the
+neural network models. From a biological perspective, the 30-year
+follow-up time in the METABRIC study may contain competing causes of
+death. The signal from the start of the study may not match the signal
+towards the end, potentially explaining the drop in performance as we
+reach later survival times with fewer individuals.
 
-Concordance-based measures are commonly used in survival analysis to compare models and we opt to keep them in our analyses.
-However, c-ipcw is a non-proper metric and may cause misspecified models to appear better than they should [@cindexfails2019].
-The model rankings between c-ipcw and IPA differed for both the complex simulation and METABRIC application.
-In the complex simulation, the c-ipcw of a correctly specified model (Optimal) underperforms CBNN and CBLR; models without the exact interaction terms specified.
-We base our METABRIC assessment of this disagreement between metrics on the complex simulation.
-The IPA score ranks regression models first, followed by DeepHit and CBNN, but the c-ipcw score ranks regression models after DeepHit and CBNN.
-Wide confidence intervals for neural network models in c-ipcw show potential misspecification, suggesting that regression models perform best in the METABRIC application.
+Concordance-based measures are commonly used in survival analysis to
+compare models and we opt to keep them in our analyses. However, c-ipcw
+is a non-proper metric and may cause misspecified models to appear
+better than they should \citep{cindexfails2019}. The model rankings
+between c-ipcw and IPA differed for both the complex simulation and
+METABRIC application. In the complex simulation, the c-ipcw of a
+correctly specified model (Optimal) underperforms CBNN and CBLR; models
+without the exact interaction terms specified. We base our METABRIC
+assessment of this disagreement between metrics on the complex
+simulation. The IPA score ranks regression models first, followed by
+DeepHit and CBNN, but the c-ipcw score ranks regression models after
+DeepHit and CBNN. Wide confidence intervals for neural network models in
+c-ipcw show potential misspecification, suggesting that regression
+models perform best in the METABRIC application.
 
-During our testing for the complex simulation and SUPPORT application, DSM did not converge with our choice of hyper-parameters and model design.
-As the goal is to fix the design and compare performance, we did not change any shared hyperparameters.
-For DSM, a different number and combination of component distributions did not help the model converge, though we did not conduct an exhaustive test.
-This may be a limitation to their method or early-stage implementation issues.
-We include this method where it did function as it shares a similar goal of a flexible baseline hazard.
+During our testing for the complex simulation and SUPPORT application,
+DSM did not converge with our choice of hyper-parameters and model
+design. As the goal is to fix the design and compare performance, we did
+not change any shared hyperparameters. For DSM, a different number and
+combination of component distributions did not help the model converge,
+though we did not conduct an exhaustive test. This may be a limitation
+to their method or early-stage implementation issues. We include this
+method where it did function as it shares a similar goal of a flexible
+baseline hazard.
 
-If the user suspects that there may be unknown time-varying interactions and a complex baseline hazard, CBNN should strongly be considered.
-Once we perform case-base sampling and adjust for the sampling bias, we can use a sigmoid activation function to predict our hazard function.
-The user can approach their problem as they would for probabilistic modeling while interpreting their prediction in a survival context.
-In terms of performance, we saw that CBNN consistently performs better than the other neural network-based approaches over follow-up time.
-This may be because of the hyperparameters chosen, as they differ from each of the respective methods papers.
-CBNN only requires the use of standard components in machine learning libraries (add layer and sigmoid activation function).
-Due to the simplicity in its implementation and by extension user experience, we expect CBNN to be both a user-friendly approach to data-driven survival analysis and easily extendable to any feed-forward neural network frameworks.
+If the user suspects that there may be unknown time-varying interactions
+and a complex baseline hazard, CBNN should strongly be considered. Once
+we perform case-base sampling and adjust for the sampling bias, we can
+use a sigmoid activation function to predict our hazard function. The
+user can approach their problem as they would for probabilistic modeling
+while interpreting their prediction in a survival context. In terms of
+performance, we saw that CBNN consistently performs better than the
+other neural network-based approaches over follow-up time. This may be
+because of the hyperparameters chosen, as they differ from each of the
+respective methods papers. CBNN only requires the use of standard
+components in machine learning libraries (add layer and sigmoid
+activation function). Due to the simplicity in its implementation and by
+extension user experience, we expect CBNN to be both a user-friendly
+approach to data-driven survival analysis and easily extendable to any
+feed-forward neural network frameworks.
 
-Though CBNN can model time-varying interactions and a flexible baseline hazard, all models require substantial amounts of data to learn real, complex relationships effectively.
-Larger datasets will appear with time, but the cost of generating longitudinal datasets remains.
-A viable alternative strategy is the use of transfer learning, where independent and identically distributed data can function as a preliminary model to learn related features in a parallel dataset, initializing weights for our survival task [@bozinovski2020reminder].
+Though CBNN can model time-varying interactions and a flexible baseline
+hazard, all models require substantial amounts of data to learn real,
+complex relationships effectively. Larger datasets will appear with
+time, but the cost of generating longitudinal datasets remains. A viable
+alternative strategy is the use of transfer learning, where independent
+and identically distributed data can function as a preliminary model to
+learn related features in a parallel dataset, initializing weights for
+our survival task \citep{bozinovski2020reminder}.
 
-This paper describes CBNN in the single event setting.
-A next step would be to extend this approach to competing risks.
-A promising use case would be high-dimensional data with complex correlation structures, such as imaging data.
+This paper describes CBNN in the single event setting. A next step would
+be to extend this approach to competing risks. A promising use case
+would be high-dimensional data with complex correlation structures, such
+as imaging data.
+
+\bibliography{bibfile.bib}
+
+\end{document}

### 2022-05-25 SIMPaper/sbpaper.Rmd
cleaned, licensed version of cbnn

--- 2022-05-25 SIMPaper/SIMPaper.Rmd
@@
 accepted: 2022-07-01
-abstract: "In survival analysis, covariate interactions are difficult to model using regression-based methods without prior knowledge.
-Neural network-based survival models have been developed for data-driven interaction modeling, however, these approaches lack implementations for time-varying interactions or a flexible baseline hazard.
-To address this, we extend the case-base sampling framework and propose Case-Base Neural Networks (CBNN), which introduce time directly into the model, using commonly available neural network components for a simple implementation.
-First, we case-base sample the survival data, converting the survival time of each participant into discrete moments in time.
-This allows us to estimate the probability an event occurs at a given moment using a neural network to get the hazard.
-We compare CBNN to survival methods based on regression and neural networks in two simulations and two real data applications with a fixed neural network architecture.
-We report metrics for each model over follow-up time using both a proper (Index of Prediction Accuracy) and non-proper (inverse probability censoring weights-adjusted concordance index) scoring rule.
-A simple simulation shows CBNN outperforming the other neural network methods.
-The complex simulation highlights the ability of CBNN to model both a complex baseline hazard and time-varying interactions, outperforming all competitors.
+abstract: "Neural network-based survival methods have been developed to model data-driven covariate interactions.
+While these methods have led to better predictive performance than regression-based approaches, they cannot model both time-varying interactions and complex baseline hazards.
+To address this, we propose Case-Base Neural Networks (CBNN) as a new approach that combines the case-base sampling framework with flexible architectures.
+Our method naturally accounts for censoring, does not require additional hyperparameters and uses binary cross-entropy as the loss function.
+Using a novel sampling scheme and data augmentation, we incorporate time directly into a feed-forward neural network.
+CBNN predicts the probability of an event occurring at a given moment and estimates the hazard function.
+We compare the performance of CBNN to survival methods based on regression and neural networks in two simulations and two real data applications.
+We report two time-dependent metrics for each model.
+In the simulations and real data applications, CBNN provides a more consistent predictive performance across time and outperforms the competing neural network approaches.
+For a simple simulation with an exponential hazard model, CBNN consistently outperforms the other neural network methods.
+For a complex simulation, which highlights the ability of CBNN to model both a complex baseline hazard and time-varying interactions, CBNN outperforms all competitors.
 The first real data application shows CBNN outperforming all neural network competitors, while a second real data application shows competitive performance.
-CBNN has shown that it is more consistent in prediction performance across time in the simulations and real data applications compared to the competing neural network approaches."
+This study highlights the benefit of combining case-base sampling with deep learning to provide a simple and flexible modeling framework for data-driven, time-varying interaction modeling of survival outcomes."
 
@@
 
-In survival analysis, covariate interactions are difficult to model using regression-based methods without prior knowledge.
-Neural network-based survival models have been developed for data-driven interaction modeling, however, these approaches can not model both time-varying interactions and complex baseline hazards.
-To address this, we propose Case-Base Neural Networks (CBNN) which extends the case-base sampling framework which introduce time directly into the model.
-We case-base sample the survival data, converting the survival time of each participant into discrete moments in time.
-This allows us to use a feed-forward neural network that estimate the probability of an event occurring at a given moment, which can be used to estimate the hazard function.
-We then compare CBNN to survival methods based on regression and neural networks in two simulations and two real data applications with a fixed neural network architecture.
-We report metrics for each model over follow-up time using both proper (Index of Prediction Accuracy) and non-proper (inverse probability censoring weights-adjusted concordance index) scoring rules.
-A simple simulation with an exponential hazard model shows CBNN consistently outperforming the other neural network methods.
-The complex simulation highlights the ability of CBNN to model both a complex baseline hazard and time-varying interactions, outperforming all competitors.
-The first real data application shows CBNN outperforming all neural network competitors, while a second real data application shows competitive performance.
-In the simulations and real data applications, CBNN provides a more consistent predictive performance across time than the competing neural network approaches.
-
 # Introduction
-
-what is aft, dont like assuming distribution, smooth. Less used than cox. Proportional hazards assumptions hard to maintain in long study, all regression require prior knowledge.
-
-The widespread use of the Cox proportional hazards model has led to a focus on survival models based on hazard ratios and relative risks rather than on survival curves and absolute risks [@hanley2009].
-In clinical settings, Cox proportional hazards models are frequently prioritized over smooth-in-time accelerated failure time (AFT) models [@hanley2009], which can estimate absolute risks by modeling the hazard directly through user specification of a baseline hazard distribution.
-Using splines as the baseline hazard is an alternative to assuming a specific distribution [@royston2002flexible].
-The risk associated with each covariate may not be constant over time, making it difficult to maintain the proportional hazards assumption.
-Previous studies related to breast cancer have discovered time-varying interactions with covariates of interest, like tumor size [@coradini2000time].
-A limitation of regression-based models is the requirement of prior knowledge; the analyst must know potential time-varying interactions and explicitly evaluate them.
-
-Smooth-in-time accelerated failure time (AFT) models can estimate absolute risks by modeling the hazard directly through user specification of a baseline hazard distribution [@kleinbaum2012survival].
-The assumption of a specific distribution tends to be a point of contention.
-A method has been developped using splines, providing flexibility to the baseline hazard  [@royston2002flexible].
-AFT models are underutilized compared to Cox proportional hazards, resulting in a focus on survival models based on hazard ratios and relative risks rather than on survival curves and absolute risks [@hanley2009].
-However, The risk associated with each covariate may not be constant over time, making it difficult to maintain the proportional hazards assumption.
-Previous studies related to breast cancer have discovered time-varying interactions with covariates of interest, like tumor size [@coradini2000time].
-A limitation of all regression-based models is the requirement of prior knowledge; the analyst must know potential time-varying interactions and explicitly evaluate them.
+Smooth-in-time accelerated failure time (AFT) models can estimate absolute risks by modeling the hazard directly through user-specified baseline hazard distribution [@kleinbaum2012survival].
+Cox proportional hazards models are used more often than AFT models, causing analyses to be based on hazard ratios and relative risks rather than on survival curves and absolute risks [@hanley2009].
+The identification of an appropriate distribution for the baseline hazard in an AFT model may be difficult for common diseases that have many interacting risk factors or a Cox model whose pathogenesis may change with age, making it difficult to maintain the proportional hazards assumption.
+For example, previous studies of breast cancer incidence have discovered time-varying interactions with covariates of interest, including tumor size [@coradini2000time].
+One approach to resolve this involves using the basis of splines on time to provide flexibility in the baseline hazard [@royston2002flexible].
+However, regression-based models are limited in that they require prior knowledge about potential time-varying interactions and their quantitative effects.
 
 Neural networks provide a data-driven approach to approximating interaction terms.
-DeepSurv is a neural network-based proportional hazards model that implements the Cox partial log-likelihood as a custom loss function [@katzman2018DeepSurv].
-This results in a stepwise absolute risk curve that cannot accommodate time-varying interactions.
-Compared to Cox regression, DeepSurv has shown an improvement in performance on the Study to Understand Prognoses Preferences and Risk Treatments (SUPPORT) dataset [@knaus1995SUPPORT].
-Farraggi et al.
-suggested a modification to the loss function required for a Cox neural network that can handle non-proportional hazards [@faraggi1995neural].
-Wang et al.
-took the concept of extreme learning machines and applied them in a survival context, removing the need for tuning the number of layers and nodes [@wang2019extreme].
+For example, DeepSurv is a neural network-based proportional hazards model that implements the Cox partial log-likelihood as a custom loss function [@katzman2018DeepSurv], resulting in a stepwise absolute risk curve that cannot accommodate time-varying interactions.
+Compared to Cox regression, DeepSurv shows better performance on the Study to Understand Prognoses Preferences and Risk Treatments (SUPPORT) dataset [@knaus1995SUPPORT].
+To handle non-proportional hazards, a modification to the loss function was proposed [@faraggi1995neural].
+To remove the need for tuning the number of layers and nodes, the concept of extreme learning machines has been applied to a Cox neural network model [@wang2019extreme].
 
-DeepHit requires that each survival time of interest be specified in the model and directly estimates survival curves, rather than deriving a hazard function [@lee2018DeepHit].
-It assumes an inverse Gaussian distribution as the baseline hazard and its concordance index (c-index) score outperforms DeepSurv [@katzman2018DeepSurv] on the Molecular Taxonomy of Breast Cancer International Consortium (METABRIC) dataset [@curtis2012genomic].
+As an alternative method that assumes a baseline hazard distribution, DeepHit specifies each survival time of interest in the model and directly estimates survival curves, rather than deriving a hazard function [@lee2018DeepHit].
+It assumes an inverse Gaussian distribution as the baseline hazard and it outperforms DeepSurv on the Molecular Taxonomy of Breast Cancer International Consortium (METABRIC) dataset [@curtis2012genomic].
 
-Deep Survival Machines (DSM) provides a flexible, parametric survival model using neural networks with a mixture of distributions as the baseline hazard [@dsmPaper].
+Providing further flexibility, Deep Survival Machines (DSM) is a parametric survival model using neural networks with a mixture of distributions as the baseline hazard [@dsmPaper].
 On both the SUPPORT and METABRIC datasets, DSM outperforms DeepSurv and DeepHit [@dsmPaper].
@@
 
-These methods need custom loss and activation functions that are difficult to implement in practice.
-We note the complexity of these alternative neural network approaches; DeepSurv, DeepHit, and DSM all require custom loss functions [@katzman2018DeepSurv] [@lee2018DeepHit] [@dsmPaper], while DSM requires a two-phase learning process and user implementations for distributions beyond Log-Normal or Weibull [@dsmPaper].
+We note that these alternative neural network approaches all require custom loss functions [@katzman2018DeepSurv] [@lee2018DeepHit] [@dsmPaper].
+DeepHit introduces a new hyperparameter weighing its two loss functions (negative log-likelihood and ranking losses) while DSM requires a two-phase learning process and user implementations for distributions beyond Log-Normal or Weibull  [@lee2018DeepHit] [@dsmPaper].
 Regression-based approaches require prior specification of all interaction terms, which makes it challenging to model covariate effects that change over time.
+The current neural network models provide flexibility at the cost of opacity, while regression models provide clarity at the cost of flexibility.
+
 In this article, we propose Case-Base Neural Networks (CBNN) as a method that models time-varying interactions and a flexible baseline hazard using commonly available neural network components.
@@
 This sampling technique allows probabilistic models to predict survival outcomes.
-Following the case-base framework, we can use transformations of time as a feature (covariate) to specify different baseline hazards.
-For example, by including splines of time as covariates, we can approximate the Royston-Parmar flexible baseline hazard model [@royston2002flexible], however, this still requires user specification of time-varying interactions.
+As part of the case-base framework, we use transformations of time as a feature (covariate) to specify different baseline hazards.
+For example, by including splines of time as covariates, we can approximate the Royston-Parmar flexible baseline hazard model [@royston2002flexible], however, this still requires explicit use of time-varying interactions.
 CBNN can model both without user specification; only a sufficiently deep network is required.
@@
 In Section \ref{methods}, we describe how case-base sampling and neural networks are combined both conceptually and algebraically, along with our hyperparameter choices and software implementation.
-In Section \ref{sims}, we describe our metrics and compare the performance of CBNN, DeepSurv, DeepHit, DSM, Cox regression, and case-base using logistic regression on simulated data.
-Section \ref{casestudies} describes the real-data analysis, while in Section \ref{discussion}, we explore the implications of our results and contextualize them within neural network survival analysis in a single event setting.
+Section \ref{sims}, describes our metrics and compares the performance of CBNN, DeepSurv, DeepHit, DSM, Cox regression and case-base using logistic regression (CBLR) on simulated data.
+Section \ref{casestudies} describes the real-data analysis, while Section \ref{discussion} explores the implications of our results and contextualizes them within neural network survival analysis in a single event setting.
 
-# Case-base neural network methodology, metrics, and software {#methods}
+# Case-base neural network methodology, metrics and software {#methods}
 
@@
 
-For each person-moment sampled, let $X_i$ be the corresponding covariate profile $\left(x_{i1},x_{i2},...,x_{ip} \right)$, $T_i$ be the time of the person-moment, and $Y_i$ be the indicator variable for whether the event of interest occurred at time $T_i$.
+For each person-moment sampled, let $X_i$ be the corresponding covariate profile $\left(x_{i1},x_{i2},...,x_{ip} \right)$, $T_i$ be the time of the person-moment and $Y_i$ be the indicator variable for whether the event of interest occurred at time $T_i$.
 We estimate the hazard function $h(t \mid X_i)$ using the sampled person-moments.
@@
 
-Now, let $b$ be the (user-defined) size of the *base series*, and let $B$ be the sum of all follow-up times for the individuals in the study.
-If we sample the *base series* uniformly, then the hazard function of the sampling process is equal to $b/B$.
+Now, let $b$ be the (user-defined) size of the *base series* and let $B$ be the sum of all follow-up times for the individuals in the study.
+If we sample the *base series* uniformly across the study base, then the hazard function of the sampling process is equal to $b/B$.
 Therefore, we have the following equality \footnote{We are abusing notation here, conflating hazards with probabilities.
-For a rigorous treatment, see Saarela \& Hanley (2015) section 3 \cite{saarela2015} .}: \begin{align}\label{eqn:main}
+For a rigorous treatment, see Saarela \& Hanley (2015) section 3 \cite{saarela2015} .}:
+\begin{align}\label{eqn:main}
 \frac{P\left(Y_i=1 \mid X_i, T_i\right)}{P\left(Y_i = 0 \mid X_i, T_i\right)} = \frac{h\left(T_i \mid X_i\right)}{b/B}.
@@
 
-We may estimate the hazard function if we adjust for the bias introduced when sampling a fraction of the study base $B$ by adding the offset term $\log\left(\frac{b}{B} \right)$ (\eqref{eqn:offset}).
+To estimate the correct hazard function, we adjusting for the bias introduced when sampling a fraction of the study base $B$ by adding the offset term $\log\left(\frac{b}{B} \right)$ as in \eqref{eqn:offset}.
 Next, we propose using neural networks to model the odds.
@@
 where $T$ is a random variable representing the event time, $X$ is the random variable for a covariate profile, $f_{\theta}(X, T)$ represents any feed-forward neural network architecture, $\log\left(\frac{B}{b}\right)$ is bias term set by case-base sampling, $\theta$ is the set of parameters learned by the neural network and $\sigmoid(x)=\frac{1}{1+e^{-x}}$.
-By allowing a neural network to approximate a higher-order polynomial of time, the baseline hazard specification is now data-driven, where user-defined hyperparameters such as regularization, number of nodes, and layers control the flexibility of the model.
+By approximating a higher-order polynomial of time using a neural network, the baseline hazard specification is now data-driven, where user-defined hyperparameters such as regularization, number of layers and nodes control the flexibility of the model.
 We provide a detailed description of the choices we made in the next sub-section.
@@
 \begin{align*}
- \log\left( h(t \mid X_i) \right) &= \log\left(\frac{\sigmoid \left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}{1-\sigmoid \left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}\right) + \log\left(\frac{b}{B}\right) \\
+ \log\left( h(t \mid X) \right) &= \log\left(\frac{\sigmoid \left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}{1-\sigmoid \left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}\right) + \log\left(\frac{b}{B}\right) \\
  &= \log\left( \frac{\frac{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)+1}}{1-\frac{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)+1}}\right) + \log\left(\frac{b}{B}\right) \\
@@
 \end{align*}
-where $\hat{f}_{\theta}(x_{i}, t_{i})$ is our estimate for a given covariate profile and time, $y_{i}$ is our target value specifying whether an event occurred, and $N$ represents the number of individuals in our training set.
+where $\hat{f}_{\theta}(x_{i}, t_{i})$ is our estimate for a given covariate profile and time, $y_{i}$ is our target value specifying whether an event occurred and $N$ represents the number of individuals in our training set.
 
-Backpropagation is used to optimize the parameters in the model with an appropriate minimization algorithm (e.g.
-Adam, RMSPropagation, stochastic gradient descent)[@gulli2017].
-For our analysis, we use Adam.
+Backpropagation with an appropriate minimization algorithm (e.g.
+Adam, RMSPropagation, stochastic gradient descent) is used to optimize the parameters in the model [@gulli2017].
+For our analysis, we use Adam as implemented in Keras [@gulli2017].
 Note that the size of the *case series* is fixed as the number of events, but we can make the *base series* as large as we want.
@@
 Since we are directly modeling the hazard, we can readily estimate the risk function ($F$) at time $t$ for a covariate profile $X$, viz.\begin{align}\label{eqn:ci2}
-F\left(t|X\right)& = 1 - \exp\left(-\int_{0}^{t}h(u|X) \,\textrm du\right).
+F\left(t\mid X\right)& = 1 - \exp\left(-\int_{0}^{t}h(u|X) \,\textrm du\right).
 \end{align} 
-We use Riemann's Sum [@hughes2020calculus] to approximate the integral in \eqref{eqn:ci2}.
+We use a finite Riemann sum [@hughes2020calculus] to approximate the integral in \eqref{eqn:ci2}.
 
@@
 
-Neural networks provide flexibility when defining the architecture and optimization parameters of our models.
-These hyperparameter choices can affect the estimated parameters and were chosen during a set of initial simulations to determine if CBNN can learn complex interactions in practice.
-We set a max number of epochs to be $2000$, batch size as $512$, learning rate as $10^{-3}$, decay as $10^{-7}$, patience as $10$ epochs, $\{50,50,25,25\}$ nodes in each hidden layer with $50\%$ dropout at each layer, a validation split of $20\%$, a testing split of $10%$, a minimum delta loss on the validation set of $10^{-7}$ over 10 epochs, and Adam [@gulli2017] as our optimizer.
+Neural networks are flexible when defining the architecture and optimization parameters.
+These hyperparameter decisions can affect the estimated parameters and were chosen during a set of initial simulations to determine if CBNN can learn complex interactions in practice.
+We set a max number of epochs to be $2000$, batch size as $512$, learning rate as $10^{-3}$, decay as $10^{-7}$, patience as $10$ epochs, $\{50,50,25,25\}$ nodes in each hidden layer with $50\%$ dropout at each layer, a validation split of $20\%$, a testing split of $10%$, a minimum delta loss on the validation set of $10^{-7}$ over 10 epochs and Adam [@gulli2017] as our optimizer.
 These choices may permit the model to approximate higher-order interactions while preventing over-fitting [@srivastava2014dropout].
-We fix a train-validation-test split that allows us to update the weights with a subset of the data (training), assess performance at each epoch (validation) and gauge performance of the final model (test).
+We fix a train-validation-test split that allows us to update the weights with a subset of the data (training), assess performance at each epoch (validation) and gauge performance of the final model (test) for each method.
 We select the best weights after training based on validation loss [@gulli2017].
@@
 R [@Rsoft] and python [@py] are used to evaluate methods from both languages.
-We fit the Cox model using the **survival** package [@survpkg], the CBLR model using the **casebase** package [@cbpkg], the DeepSurv model using the **survivalmodels** package [@survmods], the DSM model using **DeepSurvivalMachines** [@dsmPaper] and the DeepHit model using **pycox** [@lee2018DeepHit].
-We made the components of CBNN using the **keras** [@keras] package and the **casebase** package for the sampling step.
-The **simsurv** package [@simsurv] is used to simulate our simulation studies, while **flexsurv** [@flexsurv] is used to fit a flexible baseline hazard using splines for our complex simulation.
-We use the implementation of c-ipcw from the python package **sksurv** [@sksurv].
-The **riskRegression** package [@riskRegression] is used to get the Brier score and IPA metrics.
-We changed the **riskRegression** package to be used with any user supplied risk function $F$.
+We fit the Cox model using the **survival** package [@survpkg], the CBLR model using the **casebase** package [@cbpkg], the DeepSurv model using the **survivalmodels** package [@survmods], the DeepHit model using **pycox** [@lee2018DeepHit] and the DSM model using **DeepSurvivalMachines** [@dsmPaper].
+We made the components of CBNN using the **casebase** package for the sampling step and the **keras** [@keras] package for our neural network architecture.
+The **simsurv** package [@simsurv] is used for our simulation studies, while **flexsurv** [@flexsurv] is used to fit a flexible baseline hazard using splines for our complex simulation.
+We use the implementation of $C_{IPCW}$???
+from the python package **sksurv** [@sksurv].
+The **riskRegression** package [@riskRegression] is used to get the Index of Prediction Accuracy (IPA metric).
+Both metrics are described in detail in the following section.
+We modify the **riskRegression** package to be used with any user supplied risk function $F$.
 To ensure that both R and Python-based models are running in unison on the same data through our simulations and bootstrap, we use the **reticulate** package [@reticulate].
@@
 
-In this section, we use simulated data to evaluate the performance of CBNN and compare our approach with existing methods based on regression (Cox, CBLR) and neural network (DeepHit, DeepSurv, DSM) methods.
+In this section, we use simulated data to evaluate the performance of CBNN and compare our approach with existing regression-based (Cox, CBLR) and neural network-based (DeepHit, DeepSurv, DSM) methods.
 We specify a linear combination of each covariate as the linear predictor in regression-based approaches (Cox, CBLR), which contrasts with neural network approaches that allow for non-linear interactions.
-We simulate data under a simple exponential model, and a complex baseline hazard with time-varying interactions, each with 10% random censoring.
+We simulate data under a simple exponential model and a complex baseline hazard with time-varying interactions, each with 10% random censoring.
 For both settings, we simulate three covariates:
@@
 The DeepHit-specific hyperparameter alpha is set to $0.5$ (equal weight between its negative log-likelihood and ranking losses [@lee2018DeepHit]).
-We change the **DeepSurvivalMachines** [@dsmPaper] package to include dropout and a minimum delta loss during the fitting process.
+We modify the **DeepSurvivalMachines** [@dsmPaper] package to include dropout and a minimum delta loss during the fitting process.
 For DSM, we define a mixture model of six Weibull distributions for the baseline hazard.
@@
 That is, we include the exact functional form of the covariates in a CBLR model (referred to as Optimal for simplicity).
-We calculate $t$-based 95% confidence intervals using 100 replications of the simulated data For all analyses, we use 80% for training and 20% for the test set.
+We calculate $t$-based 95% confidence intervals using 100 replications of the simulated data.
+For all analyses, we use 80% for training and 20% for the test set.
 20% of the training set is kept for validation at each epoch.
-We predict risk functions $F$ using \eqref{eqn:ci2} for individuals in the test set, which are used to calculate our c-ipcw and IPA scores.
+We predict risk functions $F$ using \eqref{eqn:ci2} for individuals in the test set, which are used to calculate our $C_{IPCW}$ and IPA scores.
 
@@
 
-We use two metrics to assess the performance of the different methods of interest: 1) index of prediction accuracy (IPA) [@kattan2018index] and 2) inverse probability censoring weights-adjusted concordance index (c-ipcw) [@uno2011], which we define below.
+We use two metrics to assess the performance of the different methods of interest: 1) (IPA) [@kattan2018index] and 2) inverse probability censoring weights-adjusted concordance index ($C_{IPCW}$) [@uno2011], which we define below.
 Both these time-dependent metrics provide transparency as to when in follow-up time each model may perform better than the others.
@@
 \begin{align}
-BS(t)=\frac{1}{N}\sum^{N}_{i=1}\left(\frac{\left(\widehat{F}(X_{i},t)-1\right)^{2}\cdot I_{T_{i}\leq t,\delta_{i}=1}}{\widehat{G}(T_{i})} + \frac{\left(\widehat{F}(X_{i},t)\right)^{2}\cdot I_{T_{i}>t}}{\widehat{G}(t)}\right),
+BS(t)=\frac{1}{N}\sum^{N}_{i=1}\left(\frac{\left(1 - \widehat{F}(t \mid X_{i})\right)^{2}\cdot I(T_{i}\leq t,\delta_{i}=1)}{\widehat{G}(T_{i})} + \frac{\widehat{F}(t\mid X_{i})^{2}\cdot I(T_{i}>t)}{\widehat{G}(t)}\right),
 \end{align}
-where $\delta_{i}=1$ shows individuals who have experienced the event, $N$ represents the number of samples in our dataset over which we calculate $BS(t)$, $\widehat{G}(t)=P[C>t]$ is a non-parametric estimate of the censoring distribution, $C$ is censoring time and $T_{i}$ is an individual’s survival or censoring time.
+where $\delta_{i}=1$ shows individuals who have experienced the event, $N$ represents the number of samples in our dataset over which we calculate $BS(t)$, $\widehat{G}(t)=P[c>t]$ is a non-parametric estimate of the censoring distribution, $c$ is censoring time and $T_{i}$ is an individual’s survival or censoring time.
 The Brier score provides a score that accounts for the information loss because of censoring.
@@
 \textrm{IPA}(t) &= 1-\frac{BS_{model}(t)}{BS_{null}(t)}, \nonumber
-\end{align} where $BS_{model}(t)$ represents the Brier score over time $t$ for the model of interest, and $BS_{null}(t)$ represents the Brier score if we use an unadjusted Kaplan-Meier (KM) curve as the prediction for all observations [@kattan2018index].
+\end{align} where $BS_{model}(t)$ represents the Brier score over time $t$ for the model of interest and $BS_{null}(t)$ represents the Brier score if we use an unadjusted Kaplan-Meier (KM) curve as the prediction for all observations [@kattan2018index].
 Note that IPA has an upper bound of one, where positive values show an increase in performance over the null model and negative values show that the null model performs better.
-These scores demonstrate how performance changes over follow-up time.
+These scores show how performance changes over follow-up time.
 
 A potential artifact of IPA is that the score is unstable at earlier and later survival times.
-This is because of a near equivalent Brier score between each model and the null model.
+This is because of a near equivalent Brier score among each model and the null model.
 At small values, a difference of $0.1$ creates a more significant fold change than at larger values.
-As the Brier score is potentially small at the start and at the end of their respective curves, we may see instability of the IPA score at those ends.
+As the Brier score is potentially small at the start and at the end of their respective curves, The IPA score may be unstable at the same locations.
 
@@
 
-The c-ipcw is a non-proper, rank-based metric that does not depend on the censoring times in the test data [@uno2011]. The c-ipcw is given by
+The $C_{IPCW}$ is a non-proper, rank-based metric that does not depend on the censoring times in the test data [@uno2011]. The $C_{IPCW}$ is given by
 
 \begin{align} \label{eq:cidx}
-c-ipcw(t) &= \frac{\sum^{N}_{i=1}\sum^{N}_{j=1}\delta_{i}\left\{\widehat{G}(T_{i})\right\}^{-2} I(T_{i}<T_{j},T_{i}<t) I\left(\widehat{F}(t|X_{i})>\widehat{F}(t|X_{j})\right)}{\sum^{N}_{i=1}\sum^{N}_{j=1}\delta_{i}\left\{\widehat{G}(T_{i})\right\}^{-2} I(T_{i}<T_{j},T_{i}<t)},
-\end{align} where the follow-up period of interest is (0,$t$), $I(\cdot)$ is an indicator function, $\widehat{F}(X_{i},t)$ is the risk function estimated for everyone in the study at time $t$ and c-ipcw can compare the performance of different models, where a higher score is better.
-Note that the c-ipcw may produce misleading performance, as it ranks based on survival times, not event status [@cindexfails2019].
+C_{IPCW}(t) &= \frac{\sum^{N}_{i=1}\sum^{N}_{j=1}\delta_{i}\left\{\widehat{G}(T_{i})\right\}^{-2} I(T_{i}<T_{j},T_{i}<t) I\left(\widehat{F}(t|X_{i})>\widehat{F}(t|X_{j})\right)}{\sum^{N}_{i=1}\sum^{N}_{j=1}\delta_{i}\left\{\widehat{G}(T_{i})\right\}^{-2} I(T_{i}<T_{j},T_{i}<t)}.
+\end{align}  where the follow-up period of interest is (0,$t$), $I(\cdot)$ is an indicator function, $\widehat{F}(X_{i},t)$ is the risk function estimated for everyone in the study at time $t$ and $C_{IPCW}$ can compare the performance of different models, where a higher score is better.
+Note that the $C_{IPCW}$ may produce misleading performance, as it ranks based on survival times, not event status [@cindexfails2019].
 This metric is considered an unbiased population concordance measure because of the IPCW adjustment [@uno2011].
-
-\srb{in figure 2 i changed $C_{IPCW}$ to c-ipcw, as capital C is used for censoring. in the plots, would it make sense to leave it as a small letter?}
 
@@
 Negative values mean our model performs worse than the null model and positive values mean the model performs better.
-The second row shows the c-ipcw for each model in each study over follow-up time.
+The second row demonstrates the $C_{IPCW}$ score for each model in each study over follow-up time.
 A score of 1 is the maximum performance for either metric.
@@
 #readRDS(here::here("figures", "megaPlot.rds"))
-knitr::include_graphics(here::here("../pmnnPlayground/reviewedAnalysisPMNN/figures", "megaPlot.pdf"))
+knitr::include_graphics(here::here("analyses/figures", "megaPlot.pdf"))
 ```
@@
 
-We simulate data from a simple model that depends on a constant baseline hazard:
-$$h(t \mid X_i) = \lambda\cdot e^{\beta_{{1}}z_{1}+\beta_{{2}}z_{2}+\beta_{{3}}z_{3}},$$
-with minimal covariate effects given by $\beta_{{1}}=0.1, \beta_{{2}}=0.1, \beta_{{3}}=0.1$ and $\lambda=1.0$. Once we simulate survival times, we introduce 10% random censoring.
+We simulate data from a simple model that primarily depends on a constant baseline hazard:
+$$\log h(t \mid X_i) = \beta_{{1}}z_{1}+\beta_{{2}}z_{2}+\beta_{{3}}z_{3},$$
+The covariate effects are given by $\beta_{{1}}=0.1, \beta_{{2}}=0.1, \beta_{{3}}=0.1$. Once we simulate survival times, we introduce 10% random censoring.
 
@@
 The regression-based methods (CBLR, Cox, Optimal) outperform the neural network ones in the simple simulation setting.
-Among the neural network approaches, CBNN outperforms all other methods in terms of both IPA and c-ipcw (Figure \ref{fig:megaPlot} A, B).
+Among the neural network approaches, CBNN outperforms all other methods in terms of both IPA and $C_{IPCW}$ (Figure \ref{fig:megaPlot} A, B).
 Specifically, we see CBNN is consistent across time with smaller confidence intervals compared to DeepHit, DeepSurv and DSM.
 
-In this simple setting, the regression models are much closer to the Optimal model, while the neural network models perform worse than the null model.
-The wide confidence bands in c-ipcw suggest the neural network models may be over-parameterized.
+In this simple setting, the regression models are much closer to the Optimal model, while the neural network models perform worse than the KM null model.
+The wide confidence bands in $C_{IPCW}$ suggest the neural network models may be over-parameterized.
 
@@
 This simulation demonstrates performance with the presence of a complex baseline hazard and a time-varying interaction.
-Originally used to demonstrate the spline-based hazard model proposed by Royston and Parmar [@royston2002flexible], the breast cancer dataset provides a complex hazard from which we simulate, available in the **flexsurv** R package [@flexsurv].
+Originally used to show the spline-based hazard model proposed by Royston and Parmar [@royston2002flexible], the breast cancer dataset provides a complex hazard from which we simulate, available in the **flexsurv** R package [@flexsurv].
 To increase the complexity of our data-generating mechanism for this simulation, we design the model as follows:
 \begin{align}
-h(t \mid X_i) =\sum_{i=1}^{5} (\gamma_{i} \cdot \psi_{i}) + \beta_{{1}} (z_{1}) + \beta_{{2}} (z_{2})+ \beta_{{3}} (z_{3})+ \tau_{1} ( z_{1} \cdot z_{2} \cdot t)+ \tau_{2} ( z_{1} \cdot z_{3})+ \tau_{3} (z_{2} \cdot z_{3}), \nonumber
+\log h(t \mid X_i) =\sum_{i=1}^{5} (\gamma_{i} \cdot \psi_{i}) + \beta_{{1}} (z_{1}) + \beta_{{2}} (z_{2})+ \beta_{{3}} (z_{3})+ \tau_{1} ( z_{1} \cdot z_{2} \cdot t)+ \tau_{2} ( z_{1} \cdot z_{3})+ \tau_{3} (z_{2} \cdot z_{3}), \nonumber
 \end{align}
@@
 These tables are included to provide exact measures at certain intervals.
-The models of interest are: Cox, case-base with logistic regression (CBLR), DeepSurv, DeepHit, Case-Base Neural Networks (CBNN), Optimal (a CBLR model with the exact interaction terms and baseline hazard specified) and Deep Survival Machines (DSM).}
+The models of interest are: Cox, case-base with logistic regression (CBLR), DeepSurv, DeepHit, Case-Base Neural Network (CBNN), Optimal and Deep Survival Machines (DSM).}
 \label{tab:megaTable}
@@
 #readRDS(here::here("figures", "megaPlot.rds"))
-knitr::include_graphics(here::here("../pmnnPlayground/reviewedAnalysisPMNN/figures", "megaTable.pdf"))
+knitr::include_graphics(here::here("analyses/figures", "megaTable.pdf"))
 ```
@@
 Figure \ref{fig:megaPlot} C, D and Table \ref{tab:megaTable} B show the performance over time on a test set in the complex simulation.
-Apart from the Optimal regression model, CBNN outperforms the competing models when examining IPA, and up to the 75-percentile of follow-up time for c-ipcw.
+Apart from the Optimal regression model, CBNN outperforms the competing models when examining IPA and up to the 75-percentile of follow-up time for $C_{IPCW}$.
 We expected the Optimal model to perform best in both metrics.
-However, this was not the case for c-ipcw and may be due to an artifact of concordance-based metrics, where a misspecified model may perform better than a correctly specified one [@cindexfails2019].
-We attribute this difference in performance to the flexibility of CBNN in terms of time-varying interactions and baseline hazard, flexibility the other neural network models do not have, explaining their drop in performance.
+However, this was not the case for $C_{IPCW}$ and may be due to an artifact of concordance-based metrics, where a misspecified model may perform better than a correctly specified one [@cindexfails2019].
+We attribute the performance of CBNN to its flexibility in modeling time-varying interactions and baseline hazard, flexibility the other neural network models do not have.
 
@@
 20% of the training set is kept aside for validation at each epoch.
-We predict risk functions for everyone in the test set, which are used to calculate our metrics.
+We predict risk functions for everyone in the test set, which is used to calculate our metrics.
 We conduct 100 bootstrap re-samples for the real data applications to obtain confidence intervals.
@@
 We use a pre-processed version of the dataset made available in the DeepSurv package [@katzman2018DeepSurv].
-This dataset contains 9104 samples and 14 covariates (age, sex, race, number of comorbidities, presence of diabetes, presence of dementia, presence of cancer, mean arterial blood pressure, heart rate, respiration rate, temperature, white blood cell count, serum’s sodium, and serum’s creatinine) [@katzman2018DeepSurv].
+This dataset contains 9104 samples and 14 covariates (age, sex, race, number of comorbidities, presence of diabetes, presence of dementia, presence of cancer, mean arterial blood pressure, heart rate, respiration rate, temperature, white blood cell count, serum’s sodium and serum’s creatinine) [@katzman2018DeepSurv].
 Patients with missing features were excluded and 68.10% of the patients died during the 5.56-year study period [@katzman2018DeepSurv].
@@
 We note a drop in performance for CBNN before the $25^{th}$ percentile of follow-up.
-For c-ipcw, CBNN outperforms the competing models consistently over follow-up time.
-Note that performance is similar for all models, aside from DeepSurv whose c-ipcw is lower than the rest (Figure \ref{fig:megaPlot} E, F and Table \ref{tab:megaTable} C).
+For $C_{IPCW}$, CBNN outperforms the competing models consistently over follow-up time.
+Note that performance is similar for all models, aside from DeepSurv whose $C_{IPCW}$ is lower than the rest (Figure \ref{fig:megaPlot} E, F and Table \ref{tab:megaTable} C).
 
@@
 We used the processed dataset made available through DeepSurv [@katzman2018DeepSurv], which includes 1980 patients of which 57.72% die due to breast cancer within a median 10 years of follow-up [@katzman2018DeepSurv].
-There are 11 covariates in total: 4 RNA-Seq gene expressions (MKI67, EGFR, PGR, and ERBB2) and 5 clinical features (hormone treatment indicator, radiotherapy indicator, chemotherapy indicator, ER-positive indicator, and age at diagnosis) [@katzman2018DeepSurv].
+There are 11 covariates in total: 4 RNA-Seq gene expressions (MKI67, EGFR, PGR and ERBB2) and 5 clinical features (hormone treatment indicator, radiotherapy indicator, chemotherapy indicator, ER-positive indicator and age at diagnosis) [@katzman2018DeepSurv].
 
@@
 Our CBNN model is comparable to KM until around the 50-percentile of follow-up time, after which CBNN, DeepSurv and DSM drop in performance.
-The c-ipcw produces a different ranking.
+The $C_{IPCW}$ produces a different ranking.
 Our CBNN model outperforms the other models up to the $25^{th}$ percentile of follow-up time, whereas DeepHit performs best from the $50^{th}$ to $75^{th}$ percentile of follow-up.
 With this metric, CBNN and DeepHit outperform the regression models.
-This disagreement between IPA and c-ipcw between may be due to the misspecification issue of concordance-based metrics [@cindexfails2019].
-The neural network models may be over-parameterized, as shown by the wide confidence bands in c-ipcw.
+This disagreement between IPA and $C_{IPCW}$ may be due to the misspecification issue of concordance-based metrics [@cindexfails2019].
+The neural network models may be over-parameterized, as shown by the wide confidence bands in $C_{IPCW}$.
 
@@
 CBNN models survival outcomes by using neural networks on case-base sampled data.
-While estimating the hazard function, we incorporate follow-up-time as a feature, providing a data-driven estimation of a flexible baseline hazard and time-varying interactions.
+We incorporate follow-up time as a feature, providing a data-driven estimate of a flexible baseline hazard and time-varying interactions in our hazard function.
+The three competing neural network models we evaluated cannot model time-varying interactions by design [@dsmPaper] [@katzman2018DeepSurv] [@lee2018DeepHit].
+DSM requires a mixture component distributions for a flexible baseline hazard to be fit [@dsmPaper].
+As our goal is to fix the design and compare performance, we did not change any shared hyperparameters.
+With our choice of shared hyperparameters and model design, DSM did not converge in the complex simulation and SUPPORT case study.
+Despite this limitation, we include this method when it did function as it can fit flexible baseline hazards.
+Compared to CBNN, the remaining two models also have limitations.
+DeepSurv is a proportional hazards model and does not estimate the baseline hazard [@katzman2018DeepSurv].
+DeepHit requires an alpha hyperparameter, is restricted to a single distribution for the baseline hazard and models the survival function directly [@lee2018DeepHit].
+The alternative neural network methods match on time, while CBNN models time directly.
+
+To assess performance among these models, we use both IPA and $C_{IPCW}$ metrics.
+Concordance-based measures are commonly used in survival analysis to compare models and we opt to keep them in our analyses.
+However, $C_{IPCW}$ is a non-proper metric and may cause misspecified models to appear better than they should [@cindexfails2019].
+Therefore, we contextualize our $C_{IPCW}$ results in relation to IPA, a proper scoring rule.
+The model rankings between $C_{IPCW}$ and IPA differed for both the complex simulation and METABRIC application.
+Wider $C_{IPCW}$ confidence intervals for DeepHit and DeepSurv show potential misspecification of the models (Figure \ref{fig:megaPlot} D, H).
+
+With this interpretation of $C_{IPCW}$ in mind, we assess two simulations with minimal noise and two case studies in real scenarios.
+The simple simulation demonstrates potential pitfalls associated with neural network models, particularly overparameterization.
+All neural network-based approaches performed worse than null KM model (IPA score), while the regression-based performed better.
+We attribute this to potential overparameterization in the neural network models, as the wide confidence intervals suggest over-fitting, even with dropout.
+Both DeepSurv and DSM are affected, while DeepHit and CBNN are less so.
+This is a limitation to our strategy for evaluating the methods, which uses a fixed study design across all assessments.
+
+From this baseline benchmark, we move on to a complex simulation that requires a method that can learn both time-varying interactions and have a flexible baseline hazard.
+Here, CBNN demonstrates a distinct advantage over all other methods.
+The regression models show improved performance over the null KM model, while the competing neural network models performed worse.
 Based on our complex simulation results (Figure \ref{fig:megaPlot} C, D and Table \ref{tab:megaTable} B), CBNN outperforms the competitors when time-varying interactions and a complex baseline hazard are present.
+This simulation shows how CBNN can perform under ideal conditions, while the following two analyses on real data serve to assess its performance in realistic conditions.
 
-The competing neural network model we evaluated can not model time varying interactions.
-DeepSurv is a proportional hazards model and does not estimate the baseline hazard [@katzman2018DeepSurv].
-DeepHit requires an alpha hyperparameter, is restricted to a single distribution for the baseline hazard, and models the survival function directly [@lee2018DeepHit].
-DSM requires the user to specify the component distributions in the mixture for a flexible baseline hazard to be fit [@dsmPaper].
-These alternative neural network approaches match on time, while CBNN models time directly.
+In the SUPPORT and METABRIC case studies, flexibility in both interaction modeling and baseline hazard did not improve CBNNs relative performance, suggesting that this flexibility does not aid prediction in either case study.
+From a biological perspective, the 30-year follow-up time in the METABRIC study may contain competing causes of death.
+The causes at the start of the study may not match the causes towards the end, potentially explaining the drop in performance as we reach later survival times with fewer individuals.
+The baseline hazard is unlikely to cause this drop as DSM is the most flexible competing model in our comparison.
+Over-fitting is also unlikely given the tight confidence intervals.
+Further research is required to determine the cause of the drop in performance seen in the METABRIC case study.
+In both case studies, CBNN outperforms the competing neural network methods.
 
-In the simple simulation, all neural network-based approaches performed worse than KM (IPA scores), while the regression-based approaches did not.
-We attribute this to potential over-parameterization in the neural network models.
-The wide confidence intervals suggest over-fitting may be the reason for the loss in performance, even with dropout.
-Both DeepSurv and DSM are affected, while DeepHit and CBNN are less so.
-
-In the complex simulation, CBNN has a distinct advantage over the other methods while examining the IPA score.
-The regression models do show improvement over the null model, while the competing neural network approaches do not.
-The time-varying interactions and complex baseline hazard may have caused a drop in performance for the competing neural network methods, compared to the regression models without interactions.
-
-In the SUPPORT application, the overall performance considering IPA is near identical.
-The regression models outperform CBNN, which performs better than the competing neural network models.
-As the confidence intervals are tight, it is unlikely that the loss in performance is because of over-fitting.
-Flexibility in both interaction modeling and baseline hazard is not particularly helpful with the SUPPORT dataset.
-
-In the METABRIC application IPA results, DSM and DeepSurv had a steep drop in performance.
-As DeepSurv does not depend on the baseline hazard during the fitting process and DSM is the most flexible competing model in our comparison, the baseline hazard is unlikely to cause this drop.
-With tight confidence intervals, over-fitting is unlikely as well.
-Further research would be required to determine the cause of this drop in performance.
-We do not see the same drop in CBNN demonstrating that, in the METABRIC application, CBNN outperforms DeepSurv and DSM.
-Considering IPA score, regression models consistently outperform the neural network models.
-From a biological perspective, the 30-year follow-up time in the METABRIC study may contain competing causes of death.
-The signal from the start of the study may not match the signal towards the end, potentially explaining the drop in performance as we reach later survival times with fewer individuals.
-
-Concordance-based measures are commonly used in survival analysis to compare models and we opt to keep them in our analyses.
-However, c-ipcw is a non-proper metric and may cause misspecified models to appear better than they should [@cindexfails2019].
-The model rankings between c-ipcw and IPA differed for both the complex simulation and METABRIC application.
-In the complex simulation, the c-ipcw of a correctly specified model (Optimal) underperforms CBNN and CBLR; models without the exact interaction terms specified.
-We base our METABRIC assessment of this disagreement between metrics on the complex simulation.
-The IPA score ranks regression models first, followed by DeepHit and CBNN, but the c-ipcw score ranks regression models after DeepHit and CBNN.
-Wide confidence intervals for neural network models in c-ipcw show potential misspecification, suggesting that regression models perform best in the METABRIC application.
-
-During our testing for the complex simulation and SUPPORT application, DSM did not converge with our choice of hyper-parameters and model design.
-As the goal is to fix the design and compare performance, we did not change any shared hyperparameters.
-For DSM, a different number and combination of component distributions did not help the model converge, though we did not conduct an exhaustive test.
-This may be a limitation to their method or early-stage implementation issues.
-We include this method where it did function as it shares a similar goal of a flexible baseline hazard.
-
-If the user suspects that there may be unknown time-varying interactions and a complex baseline hazard, CBNN should strongly be considered.
+Taken together, CBNN outperforms all competitors in the complex simulation, demonstrating its value in survival settings that may involve time-varying interactions and a complex baseline hazard.
 Once we perform case-base sampling and adjust for the sampling bias, we can use a sigmoid activation function to predict our hazard function.
-The user can approach their problem as they would for probabilistic modeling while interpreting their prediction in a survival context.
-In terms of performance, we saw that CBNN consistently performs better than the other neural network-based approaches over follow-up time.
-This may be because of the hyperparameters chosen, as they differ from each of the respective methods papers.
-CBNN only requires the use of standard components in machine learning libraries (add layer and sigmoid activation function).
-Due to the simplicity in its implementation and by extension user experience, we expect CBNN to be both a user-friendly approach to data-driven survival analysis and easily extendable to any feed-forward neural network frameworks.
-
-Though CBNN can model time-varying interactions and a flexible baseline hazard, all models require substantial amounts of data to learn real, complex relationships effectively.
-Larger datasets will appear with time, but the cost of generating longitudinal datasets remains.
-A viable alternative strategy is the use of transfer learning, where independent and identically distributed data can function as a preliminary model to learn related features in a parallel dataset, initializing weights for our survival task [@bozinovski2020reminder].
-
-This paper describes CBNN in the single event setting.
-A next step would be to extend this approach to competing risks.
-A promising use case would be high-dimensional data with complex correlation structures, such as imaging data.
+Our approach simplifies the incorporation of censored individuals, allowing survival outcomes to be treated as binary ones.
+Forgoing the requirement of custom loss functions, CBNN only requires the use of standard components in machine learning libraries (specifically, the add layer to adjust for sampling bias and the sigmoid activation function).
+Due to the simplicity in its implementation and by extension user experience, CBNN is both a user-friendly approach to data-driven survival analysis and is easily extendable to any feed-forward neural network framework.

### 2022-05-25 SIMPaper/sbpaper.tex
cleaned, licensed version of cbnn

--- 2022-05-25 SIMPaper/SIMPaper.tex
@@
 
-\abstract{In survival analysis, covariate interactions are difficult to
-model without prior knowledge using regression-based methods. Neural
-network-based survival models have been developed for data-driven
-interaction modeling, however, these approaches lack implementations for
-time-varying interactions or a flexible baseline hazard. To address
-this, we extend the case-base sampling framework and propose Case-Base
-Neural Networks (CBNN), which introduce time directly into the model,
-using commonly available neural network components for a simple
-implementation. First, we case-base sample the survival data, converting
-the survival time of each participant into discrete moments in time.
-This allows us to estimate the probability an event occurs at a given
-moment using a neural network to get the hazard. We compare CBNN to
-survival methods based on regression and neural networks in two
-simulations and two real data applications with a fixed neural network
-architecture. We report metrics for each model over follow-up time using
-both a proper (Index of Prediction Accuracy) and non-proper (inverse
-probability censoring weights-adjusted concordance index) scoring rule.
-A simple simulation shows CBNN outperforming the other neural network
-methods. The complex simulation highlights the ability of CBNN to model
-both a complex baseline hazard and time-varying interactions,
-outperforming all competitors. The first real data application shows
-CBNN outperforming all neural network competitors, while a second real
-data application shows competitive performance. CBNN has shown that it
-is more consistent in prediction performance across time in the
-simulations and real data applications compared to the competing neural
-network approaches.}
+\abstract{Neural network-based survival methods have been developed to
+model data-driven covariate interactions. While this flexibility has led
+to methods with better predictive performance than regression-based
+approaches, current neural networks cannot model both time-varying
+interactions and complex baseline hazards. To address this, we propose
+Case-Base Neural Networks (CBNN) as a new approach that combines the
+case-base sampling framework with flexible architectures. Our method
+naturally accounts for censoring, does not require additional
+hyperparameters and uses binary cross-entropy as the loss function.
+Using a novel sampling scheme and data augmentation, we incorporate time
+directly into a feed-forward neural network. CBNN predicts the
+probability of an event occurring at a given moment and estimates the
+hazard function. We compare the performance of CBNN to survival methods
+based on regression and neural networks in two simulations and two real
+data applications. We report two time-dependent metrics for each model.
+In the simulations and real data applications, CBNN provides a more
+consistent predictive performance across time and outperforms the
+competing neural network approaches. For a simple simulation with an
+exponential hazard model, CBNN consistently outperforms the other neural
+network methods. For a complex simulation, which highlights the ability
+of CBNN to model both a complex baseline hazard and time-varying
+interactions, CBNN outperforms all competitors. The first real data
+application shows CBNN outperforming all neural network competitors,
+while a second real data application shows competitive performance. This
+study highlights the benefit of combining case-base sampling with deep
+learning to provide a simple and flexible modeling framework for
+data-driven, time-varying interaction modeling of survival outcomes.}
 
@@
 
-The widespread use of the Cox proportional hazards model has led to a
-focus on survival models based on hazard ratios and relative risks
-rather than on survival curves and absolute risks \citep{hanley2009}. In
-clinical settings, Cox models are frequently prioritized over
-smooth-in-time accelerated failure time (AFT) models \citep{hanley2009},
-which can estimate absolute risks by modeling the hazard directly
-through user specification of a baseline hazard distribution. The user
-may forgo this choice by implementing a flexible set of parameters using
-splines, as proposed by Royston and Palmar \citep{royston2002flexible}.
-These regression models also require user-specified interactions, which
-may limit risk prediction accuracy.
+Smooth-in-time accelerated failure time (AFT) models can estimate
+absolute risks by modeling the hazard directly through user-specified
+baseline hazard distribution \citep{kleinbaum2012survival}. Cox
+proportional hazards models are used more often than AFT models, causing
+analyses to be based on hazard ratios and relative risks rather than on
+survival curves and absolute risks \citep{hanley2009}. The
+identification of an appropriate distribution for the baseline hazard in
+an AFT model may be difficult for common diseases that have many
+interacting risk factors or a Cox model whose pathogenesis may change
+with age, making it difficult to maintain the proportional hazards
+assumption. For example, previous studies of breast cancer incidence
+have discovered time-varying interactions with covariates of interest,
+including tumor size \citep{coradini2000time}. One approach to resolve
+this involves using the basis of splines on time to provide flexibility
+in the baseline hazard \citep{royston2002flexible}. However,
+regression-based models are limited in that they require prior knowledge
+about potential time-varying interactions and their quantitative
+effects.
 
-Over longer follow-up, it is challenging to maintain the proportional
-hazards assumption as the risk associated with each covariate may not be
-constant with time. Previous studies related to breast cancer have
-discovered time-varying interactions with covariates of interest, like
-tumor size \citep{coradini2000time} \citep{hilsenbeck1998time}. Another
-limitation of regression-based models is the requirement of prior
-knowledge; the analyst must know potential time-varying interactions and
-explicitly evaluate them.
+Neural networks provide a data-driven approach to approximating
+interaction terms. For example, DeepSurv is a neural network-based
+proportional hazards model that implements the Cox partial
+log-likelihood as a custom loss function \citep{katzman2018DeepSurv},
+resulting in a stepwise absolute risk curve that cannot accommodate
+time-varying interactions. Compared to Cox regression, DeepSurv shows
+better performance on the Study to Understand Prognoses Preferences and
+Risk Treatments (SUPPORT) dataset \citep{knaus1995SUPPORT}. To handle
+non-proportional hazards, a modification to the loss function was
+proposed \citep{faraggi1995neural}. To remove the need for tuning the
+number of layers and nodes, the concept of extreme learning machines has
+been applied to a Cox neural network model \citep{wang2019extreme}.
 
-Data-driven neural network approaches bypass the need for prior
-knowledge of interaction terms. DeepSurv is a semi-parametric
-proportional hazards neural network model that implements the Cox
-partial log-likelihood as a custom loss function
-\citep{katzman2018DeepSurv}. This results in a stepwise absolute risk
-curve that cannot accommodate time-varying interactions. Compared to Cox
-regression, DeepSurv has shown a small improvement in performance using
-a single layer neural network on the Study to Understand Prognoses
-Preferences and Risk Treatments (SUPPORT) dataset
-\citep{knaus1995SUPPORT}. Farraggi et al.~suggested a modification to
-the loss function required for a Cox neural network that can model
-non-proportional hazards \citep{faraggi1995neural}. Wang et al.~took the
-concept of extreme learning machines and applied them in a survival
-context, removing the need for tuning the number of layers and nodes
-\citep{wang2019extreme}. A Cox-based neural network combined with Local
-Interpretable Model-Agnostic Explanations (LIME)
-\citep{ribeiro2016model} interprets the association of each Single
-Nucleotide Polymorphism (SNP) \citep{sun2020genome} to a phenotype.
+As an alternative method that assumes a baseline hazard distribution,
+DeepHit specifies each survival time of interest in the model and
+directly estimates survival curves, rather than deriving a hazard
+function \citep{lee2018DeepHit}. It assumes an inverse Gaussian
+distribution as the baseline hazard and it outperforms DeepSurv on the
+Molecular Taxonomy of Breast Cancer International Consortium (METABRIC)
+dataset \citep{curtis2012genomic}.
 
-DeepHit requires each survival time of interest to be specified in the
-model and directly estimates survival curves, rather than deriving a
-hazard function \citep{lee2018DeepHit}. It assumes an inverse Gaussian
-distribution as the baseline hazard and its concordance index (c-index)
-score outperforms DeepSurv \citep{katzman2018DeepSurv} on the Molecular
-Taxonomy of Breast Cancer International Consortium (METABRIC) dataset
-\citep{curtis2012genomic}.
-
-Ripley et al.~presented seven different distributions and modeling
-paradigms based on neural networks \citep{ripley2004non}. They provided
-the methodology behind parametric log-logistic and log-normal neural
-network models. Like DeepHit, these models are restricted to one of the
-seven baseline hazard distributions, reducing the flexibility of the
-average survival prediction.
-
-Deep Survival Machines (DSM) provides a flexible, parametric survival
-model using neural networks with a mixture of distributions as the
-baseline hazard \citep{dsmPaper}. DSM allows the user to specify a
-mixture of distributions for a flexible baseline hazard. On both the
+Providing further flexibility, Deep Survival Machines (DSM) is a
+parametric survival model using neural networks with a mixture of
+distributions as the baseline hazard \citep{dsmPaper}. On both the
 SUPPORT and METABRIC datasets, DSM outperforms DeepSurv and DeepHit
@@
 
-These methods need custom loss and activation functions that are
-difficult to implement in practice. We note the complexity of these
-alternative neural network approaches; DeepSurv, DeepHit, and DSM all
-require custom loss functions \citep{katzman2018DeepSurv}
-\citep{lee2018DeepHit} \citep{dsmPaper}, while DSM requires user
-implementations for distributions beyond Log-Normal or Weibull and a
-two-phase learning process \citep{dsmPaper}. Regression-based approaches
-require prior specification of all interaction terms, which makes it
-challenging to model covariate effects that change over time. In this
-article, we propose Case-Base Neural Networks (CBNN) as a method that
-models time-varying interactions and a flexible baseline hazard using
-commonly available neural network components. Our approach to modeling
-the full hazard uses case-base sampling \citep{hanley2009}, which allows
-probabilistic models to predict survival outcomes. Following the
-case-base framework, we can use transformations of time as a feature
-(covariate) to specify different baseline hazards. For example, by
-including splines of time as covariates, we can approximate the
-Royston-Parmar flexible baseline hazard model
-\citep{royston2002flexible}. An alternative approach to Royston-Parmar
-is case-base with logistic regression (CBLR), which provides some
-flexibility in modeling, but still requires user specification of
-time-varying interactions and baseline hazard, while CBNN can model both
-without user specification; only a sufficiently deep network is
-required.
+We note that these alternative neural network approaches all require
+custom loss functions \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}
+\citep{dsmPaper}. DeepHit introduces a new hyperparameter weighing its
+two loss functions (negative log-likelihood and ranking losses) while
+DSM requires a two-phase learning process and user implementations for
+distributions beyond Log-Normal or Weibull \citep{lee2018DeepHit}
+\citep{dsmPaper}. Regression-based approaches require prior
+specification of all interaction terms, which makes it challenging to
+model covariate effects that change over time. The current neural
+network models provide flexibility at the cost of opacity, while
+regression models provide clarity at the cost of flexibility.
+
+In this article, we propose Case-Base Neural Networks (CBNN) as a method
+that models time-varying interactions and a flexible baseline hazard
+using commonly available neural network components. Our approach to
+modeling the full hazard uses case-base sampling \citep{hanley2009}.
+This sampling technique allows probabilistic models to predict survival
+outcomes. As part of the case-base framework, we use transformations of
+time as a feature (covariate) to specify different baseline hazards. For
+example, by including splines of time as covariates, we can approximate
+the Royston-Parmar flexible baseline hazard model
+\citep{royston2002flexible}, however, this still requires explicit use
+of time-varying interactions. CBNN can model both without user
+specification; only a sufficiently deep network is required.
 
@@
 networks are combined both conceptually and algebraically, along with
-our hyperparameter choice and software implementation. Then, in Section
-\ref{sims}, we describe our metrics and compare the performance of CBNN,
-DeepSurv, DeepHit, DSM, Cox regression, and case-base using logistic
-regression on simulated data. Next, Section \ref{casestudies} describes
-the real-data analysis, while in Section \ref{discussion}, we explore
-the implications of our results and contextualize them within neural
+our hyperparameter choices and software implementation. Section
+\ref{sims}, describes our metrics and compares the performance of CBNN,
+DeepSurv, DeepHit, DSM, Cox regression and case-base using logistic
+regression (CBLR) on simulated data. Section \ref{casestudies} describes
+the real-data analysis, while Section \ref{discussion} explores the
+implications of our results and contextualizes them within neural
 network survival analysis in a single event setting.
@@
 \hypertarget{methods}{%
-\section{Case-base neural network methodology, metrics, and
+\section{Case-base neural network methodology, metrics and
 software}\label{methods}}
@@
 covariate profile \(\left(x_{i1},x_{i2},...,x_{ip} \right)\), \(T_i\) be
-the time of the person-moment, and \(Y_i\) be the indicator variable for
+the time of the person-moment and \(Y_i\) be the indicator variable for
 whether the event of interest occurred at time \(T_i\). We estimate the
@@
 
-Now, let \(b\) be the (user-defined) size of the \emph{base series}, and
+Now, let \(b\) be the (user-defined) size of the \emph{base series} and
 let \(B\) be the sum of all follow-up times for the individuals in the
-study. If we sample the \emph{base series} uniformly, then the hazard
-function of the sampling process is equal to \(b/B\). Therefore, we have
-the following equality
+study. If we sample the \emph{base series} uniformly across the study
+base, then the hazard function of the sampling process is equal to
+\(b/B\). Therefore, we have the following equality
 \footnote{We are abusing notation here, conflating hazards with probabilities. For a rigorous treatment, see Saarela \& Hanley (2015) section 3 \cite{saarela2015} .}:
@@
 
-We may estimate the hazard function if we adjust for the bias introduced
-when sampling a fraction of the study base \(B\) by adding the offset
-term \(\log\left(\frac{b}{B} \right)\) (\eqref{eqn:offset}). Next, we
-propose using neural networks to model the odds.
+To estimate the correct hazard function, we adjusting for the bias
+introduced when sampling a fraction of the study base \(B\) by adding
+the offset term \(\log\left(\frac{b}{B} \right)\) as in
+\eqref{eqn:offset}. Next, we propose using neural networks to model the
+odds.
 
@@
 \(\theta\) is the set of parameters learned by the neural network and
-\(\mathrm{sigmoid}(x)=\frac{1}{1+e^{-x}}\). By allowing a neural network
-to approximate a higher-order polynomial of time, the baseline hazard
-specification is now data-driven, where user-defined hyperparameters
-such as regularization, number of nodes, and layers control the
-flexibility of the model. We provide a detailed description of the
-choices we made in the next sub-section.
+\(\mathrm{sigmoid}(x)=\frac{1}{1+e^{-x}}\). By approximating a
+higher-order polynomial of time using a neural network, the baseline
+hazard specification is now data-driven, where user-defined
+hyperparameters such as regularization, number of layers and nodes
+control the flexibility of the model. We provide a detailed description
+of the choices we made in the next sub-section.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/home/bhatnagar-lab/jislam/pmnnProjects/paper/figures/nnarch2}
+{\centering \includegraphics[width=1\linewidth]{/home/bhatnagar-lab/jislam/cbnnManuscript/figures/nnarch2}
 
@@
 to odds: \begin{align*}
- \log\left( h(t \mid X_i) \right) &= \log\left(\frac{\mathrm{sigmoid}\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}{1-\mathrm{sigmoid}\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}\right) + \log\left(\frac{b}{B}\right) \\
+ \log\left( h(t \mid X) \right) &= \log\left(\frac{\mathrm{sigmoid}\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}{1-\mathrm{sigmoid}\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}\right) + \log\left(\frac{b}{B}\right) \\
  &= \log\left( \frac{\frac{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)+1}}{1-\frac{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)+1}}\right) + \log\left(\frac{b}{B}\right) \\
@@
 for a given covariate profile and time, \(y_{i}\) is our target value
-specifying whether an event occurred, and \(N\) represents the number of
+specifying whether an event occurred and \(N\) represents the number of
 individuals in our training set.
 
-Backpropagation is used to optimize the parameters in the model with an
-appropriate minimization algorithm (e.g.~Adam, RMSPropagation,
-stochastic gradient descent)\citep{gulli2017}. For our analysis, we use
-Adam. Note that the size of the \emph{case series} is fixed as the
-number of events, but we can make the \emph{base series} as large as we
-want. A ratio of 100:1 \emph{base series} to \emph{case series} is
-sufficient \citep{hanley2009}. We pass our feed-forward neural network
-through a sigmoid activation function (Figure \ref{fig:NNarch}).
-Finally, we can convert this model output to a hazard. When using our
-model for predictions, we manually set the offset term to 0 in the new
-data, as we account for the bias during the fitting process.
+Backpropagation with an appropriate minimization algorithm (e.g.~Adam,
+RMSPropagation, stochastic gradient descent) is used to optimize the
+parameters in the model \citep{gulli2017}. For our analysis, we use Adam
+as implemented in Keras \citep{gulli2017}. Note that the size of the
+\emph{case series} is fixed as the number of events, but we can make the
+\emph{base series} as large as we want. A ratio of 100:1 \emph{base
+series} to \emph{case series} is sufficient \citep{hanley2009}. We pass
+our feed-forward neural network through a sigmoid activation function
+(Figure \ref{fig:NNarch}). Finally, we can convert this model output to
+a hazard. When using our model for predictions, we manually set the
+offset term to 0 in the new data, as we account for the bias during the
+fitting process.
 
@@
 viz.\begin{align}\label{eqn:ci2}
-F\left(t|X\right)& = 1 - \exp\left(-\int_{0}^{t}h(u|X) \,\textrm du\right).
-\end{align} We use Riemann's Sum \citep{hughes2020calculus} to
+F\left(t\mid X\right)& = 1 - \exp\left(-\int_{0}^{t}h(u|X) \,\textrm du\right).
+\end{align} We use a finite Riemann sum \citep{hughes2020calculus} to
 approximate the integral in \eqref{eqn:ci2}.
@@
 
-Neural networks provide flexibility when defining the architecture and
-optimization parameters of our models. These hyperparameter choices can
-affect the estimated parameters and were chosen during a set of initial
-simulations to determine if CBNN can learn complex interactions in
-practice. We set a max number of epochs to be \(2000\), batch size as
-\(512\), learning rate as \(10^{-3}\), decay as \(10^{-7}\), patience as
-\(10\) epochs, \(\{50,50,25,25\}\) nodes in each hidden layer with
-\(50\%\) dropout at each layer, a validation split of \(20\%\), a
-testing split of \(10%
+Neural networks are flexible when defining the architecture and
+optimization parameters. These hyperparameter decisions can affect the
+estimated parameters and were chosen during a set of initial simulations
+to determine if CBNN can learn complex interactions in practice. We set
+a max number of epochs to be \(2000\), batch size as \(512\), learning
+rate as \(10^{-3}\), decay as \(10^{-7}\), patience as \(10\) epochs,
+\(\{50,50,25,25\}\) nodes in each hidden layer with \(50\%\) dropout at
+each layer, a validation split of \(20\%\), a testing split of \(10%
 \), a minimum delta loss on the validation set of \(10^{-7}\) over 10
-epochs, and Adam \citep{gulli2017} as our optimizer. These choices may
+epochs and Adam \citep{gulli2017} as our optimizer. These choices may
 permit the model to approximate higher-order interactions while
@@
 subset of the data (training), assess performance at each epoch
-(validation) and gauge performance of the final model (test). We select
-the best weights after training based on validation loss
-\citep{gulli2017}.
+(validation) and gauge performance of the final model (test) for each
+method. We select the best weights after training based on validation
+loss \citep{gulli2017}.
 
@@
 \citep{cbpkg}, the DeepSurv model using the \textbf{survivalmodels}
-package \citep{survmods}, the DSM model using
-\textbf{DeepSurvivalMachines} \citep{dsmPaper} and the DeepHit model
-using \textbf{pycox} \citep{lee2018DeepHit}. We made the components of
-CBNN using the \textbf{keras} \citep{keras} package and the
-\textbf{casebase} package for the sampling step. The \textbf{simsurv}
-package \citep{simsurv} is used to simulate our simulation studies,
-while \textbf{flexsurv} \citep{flexsurv} is used to fit a flexible
-baseline hazard using splines for our complex simulation. We use the
-implementation of c-ipcw from the python package \textbf{sksurv}
-\citep{sksurv}. The \textbf{riskRegression} package
-\citep{riskRegression} is used to get the Brier score and IPA metrics.
-We changed the \textbf{riskRegression} package to be used with any user
-supplied risk function \(F\). To ensure that both R and Python-based
-models are running in unison on the same data through our simulations
-and bootstrap, we use the \textbf{reticulate} package
+package \citep{survmods}, the DeepHit model using \textbf{pycox}
+\citep{lee2018DeepHit} and the DSM model using
+\textbf{DeepSurvivalMachines} \citep{dsmPaper}. We made the components
+of CBNN using the \textbf{casebase} package for the sampling step and
+the \textbf{keras} \citep{keras} package for our neural network
+architecture. The \textbf{simsurv} package \citep{simsurv} is used for
+our simulation studies, while \textbf{flexsurv} \citep{flexsurv} is used
+to fit a flexible baseline hazard using splines for our complex
+simulation. We use the implementation of \(C_{IPCW}\)??? from the python
+package \textbf{sksurv} \citep{sksurv}. The \textbf{riskRegression}
+package \citep{riskRegression} is used to get the Index of Prediction
+Accuracy (IPA metric). Both metrics are described in detail in the
+following section. We modify the \textbf{riskRegression} package to be
+used with any user supplied risk function \(F\). To ensure that both R
+and Python-based models are running in unison on the same data through
+our simulations and bootstrap, we use the \textbf{reticulate} package
 \citep{reticulate}.
@@
 In this section, we use simulated data to evaluate the performance of
-CBNN and compare our approach with existing methods based on regression
-(Cox, CBLR) and neural network (DeepHit, DeepSurv, DSM) methods. We
-specify a linear combination of each covariate as the linear predictor
-in regression-based approaches (Cox, CBLR), which contrasts with neural
+CBNN and compare our approach with existing regression-based (Cox, CBLR)
+and neural network-based (DeepHit, DeepSurv, DSM) methods. We specify a
+linear combination of each covariate as the linear predictor in
+regression-based approaches (Cox, CBLR), which contrasts with neural
 network approaches that allow for non-linear interactions. We simulate
-data under a simple exponential model, and a complex baseline hazard
-with time-varying interactions, each with 10\% random censoring. For
-both settings, we simulate three covariates:
+data under a simple exponential model and a complex baseline hazard with
+time-varying interactions, each with 10\% random censoring. For both
+settings, we simulate three covariates:
 
@@
 weight between its negative log-likelihood and ranking losses
-\citep{lee2018DeepHit}). We change the \textbf{DeepSurvivalMachines}
+\citep{lee2018DeepHit}). We modify the \textbf{DeepSurvivalMachines}
 \citep{dsmPaper} package to include dropout and a minimum delta loss
@@
 We calculate \(t\)-based 95\% confidence intervals using 100
-replications of the simulated data For all analyses, we use 80\% for
+replications of the simulated data. For all analyses, we use 80\% for
 training and 20\% for the test set. 20\% of the training set is kept for
@@
 \eqref{eqn:ci2} for individuals in the test set, which are used to
-calculate our c-ipcw and IPA scores.
+calculate our \(C_{IPCW}\) and IPA scores.
 
@@
 We use two metrics to assess the performance of the different methods of
-interest: 1) index of prediction accuracy (IPA) \citep{kattan2018index}
-and 2) inverse probability censoring weights-adjusted concordance index
-(c-ipcw) \citep{uno2011}, which we define below. Both these
-time-dependent metrics provide transparency as to when in follow-up time
-each model may perform better than the others.
+interest: 1) (IPA) \citep{kattan2018index} and 2) inverse probability
+censoring weights-adjusted concordance index (\(C_{IPCW}\))
+\citep{uno2011}, which we define below. Both these time-dependent
+metrics provide transparency as to when in follow-up time each model may
+perform better than the others.
 
@@
 which is defined as \begin{align}
-BS(t)=\frac{1}{N}\sum^{N}_{i=1}\left(\frac{\left(\widehat{F}(X_{i},t)-1\right)^{2}\cdot I_{T_{i}\leq t,\delta_{i}=1}}{\widehat{G}(T_{i})} + \frac{\left(\widehat{F}(X_{i},t)\right)^{2}\cdot I_{T_{i}>t}}{\widehat{G}(t)}\right),
+BS(t)=\frac{1}{N}\sum^{N}_{i=1}\left(\frac{\left(1 - \widehat{F}(t \mid X_{i})\right)^{2}\cdot I(T_{i}\leq t,\delta_{i}=1)}{\widehat{G}(T_{i})} + \frac{\widehat{F}(t\mid X_{i})^{2}\cdot I(T_{i}>t)}{\widehat{G}(t)}\right),
 \end{align} where \(\delta_{i}=1\) shows individuals who have
 experienced the event, \(N\) represents the number of samples in our
-dataset over which we calculate \(BS(t)\), \(\widehat{G}(t)=P[C>t]\) is
-a non-parametric estimate of the censoring distribution, \(C\) is
+dataset over which we calculate \(BS(t)\), \(\widehat{G}(t)=P[c>t]\) is
+a non-parametric estimate of the censoring distribution, \(c\) is
 censoring time and \(T_{i}\) is an individual's survival or censoring
@@
 \end{align} where \(BS_{model}(t)\) represents the Brier score over time
-\(t\) for the model of interest, and \(BS_{null}(t)\) represents the
-Brier score if we use an unadjusted Kaplan-Meier curve as the prediction
-for all observations \citep{kattan2018index}. Note that IPA has an upper
-bound of one, where positive values show an increase in performance over
-the null model and negative values show that the null model performs
-better. These scores demonstrate how performance changes over follow-up
-time.
+\(t\) for the model of interest and \(BS_{null}(t)\) represents the
+Brier score if we use an unadjusted Kaplan-Meier (KM) curve as the
+prediction for all observations \citep{kattan2018index}. Note that IPA
+has an upper bound of one, where positive values show an increase in
+performance over the null model and negative values show that the null
+model performs better. These scores show how performance changes over
+follow-up time.
 
@@
 later survival times. This is because of a near equivalent Brier score
-between each model and the null model. At small values, a difference of
+among each model and the null model. At small values, a difference of
 \(0.1\) creates a more significant fold change than at larger values. As
 the Brier score is potentially small at the start and at the end of
-their respective curves, we may see instability of the IPA score at
-those ends.
+their respective curves, The IPA score may be unstable at the same
+locations.
 
@@
 
-The c-ipcw is a non-proper, rank-based metric that does not depend on
-the censoring times in the test data \citep{uno2011}. The c-ipcw is
-given by
+The \(C_{IPCW}\) is a non-proper, rank-based metric that does not depend
+on the censoring times in the test data \citep{uno2011}. The
+\(C_{IPCW}\) is given by
 
 \begin{align} \label{eq:cidx}
-c-ipcw(t) &= \frac{\sum^{N}_{i=1}\sum^{N}_{j=1}\delta_{i}\left\{\widehat{G}(T_{i})\right\}^{-2} I(T_{i}<T_{j},T_{i}<t) I\left(\widehat{F}(t|X_{i})>\widehat{F}(t|X_{j})\right)}{\sum^{N}_{i=1}\sum^{N}_{j=1}\delta_{i}\left\{\widehat{G}(T_{i})\right\}^{-2} I(T_{i}<T_{j},T_{i}<t)},
+C_{IPCW}(t) &= \frac{\sum^{N}_{i=1}\sum^{N}_{j=1}\delta_{i}\left\{\widehat{G}(T_{i})\right\}^{-2} I(T_{i}<T_{j},T_{i}<t) I\left(\widehat{F}(t|X_{i})>\widehat{F}(t|X_{j})\right)}{\sum^{N}_{i=1}\sum^{N}_{j=1}\delta_{i}\left\{\widehat{G}(T_{i})\right\}^{-2} I(T_{i}<T_{j},T_{i}<t)}.
 \end{align} where the follow-up period of interest is (0,\(t\)),
@@
 risk function estimated for everyone in the study at time \(t\) and
-c-ipcw can compare the performance of different models, where a higher
-score is better. Note that the c-ipcw may produce misleading
-performance, as it ranks based on survival times, not event status
-\citep{cindexfails2019}. This metric is considered an unbiased
+\(C_{IPCW}\) can compare the performance of different models, where a
+higher score is better. Note that the \(C_{IPCW}\) may produce
+misleading performance, as it ranks based on survival times, not event
+status \citep{cindexfails2019}. This metric is considered an unbiased
 population concordance measure because of the IPCW adjustment
@@
 
-\srb{in figure 2 i changed $C_{IPCW}$ to c-ipcw, as capital C is used for censoring. in the plots, would it make sense to leave it as a small letter?}
-
 \begin{figure}
 
-{\centering \includegraphics[width=1\linewidth]{/home/bhatnagar-lab/jislam/pmnnProjects/paper/../pmnnPlayground/reviewedAnalysisPMNN/figures/megaPlot}
+{\centering \includegraphics[width=1\linewidth]{/home/bhatnagar-lab/jislam/cbnnManuscript/analyses/figures/megaPlot}
 
@@
 constant baseline hazard:
-\[h(t \mid X_i) = \lambda\cdot e^{\beta_{{1}}z_{1}+\beta_{{2}}z_{2}+\beta_{{3}}z_{3}},\]
-with minimal covariate effects given by
-\(\beta_{{1}}=0.1, \beta_{{2}}=0.1, \beta_{{3}}=0.1\) and
-\(\lambda=1.0\). Once we simulate survival times, we introduce 10\%
-random censoring.
+\[\log h(t \mid X_i) = \beta_{{1}}z_{1}+\beta_{{2}}z_{2}+\beta_{{3}}z_{3},\]
+The covariate effects are given by
+\(\beta_{{1}}=0.1, \beta_{{2}}=0.1, \beta_{{3}}=0.1\). Once we simulate
+survival times, we introduce 10\% random censoring.
 
@@
 simulation setting. Among the neural network approaches, CBNN
-outperforms all other methods in terms of both IPA and c-ipcw (Figure
-\ref{fig:megaPlot} A, B). Specifically, we see CBNN is consistent across
-time with smaller confidence intervals compared to DeepHit, DeepSurv and
-DSM.
+outperforms all other methods in terms of both IPA and \(C_{IPCW}\)
+(Figure \ref{fig:megaPlot} A, B). Specifically, we see CBNN is
+consistent across time with smaller confidence intervals compared to
+DeepHit, DeepSurv and DSM.
 
 In this simple setting, the regression models are much closer to the
-Optimal model, while the neural network models perform worse than the
-null model. The wide confidence bands in c-ipcw suggest the neural
+Optimal model, while the neural network models perform worse than the KM
+null model. The wide confidence bands in \(C_{IPCW}\) suggest the neural
 network models may be over-parameterized.
@@
 This simulation demonstrates performance with the presence of a complex
-baseline hazard and a time-varying interaction. Originally used to
-demonstrate the spline-based hazard model proposed by Royston and Parmar
+baseline hazard and a time-varying interaction. Originally used to show
+the spline-based hazard model proposed by Royston and Parmar
 \citep{royston2002flexible}, the breast cancer dataset provides a
@@
 model as follows: \begin{align}
-h(t \mid X_i) =\sum_{i=1}^{5} (\gamma_{i} \cdot \psi_{i}) + \beta_{{1}} (z_{1}) + \beta_{{2}} (z_{2})+ \beta_{{3}} (z_{3})+ \tau_{1} ( z_{1} \cdot z_{2} \cdot t)+ \tau_{2} ( z_{1} \cdot z_{3})+ \tau_{3} (z_{2} \cdot z_{3}), \nonumber
+\log h(t \mid X_i) =\sum_{i=1}^{5} (\gamma_{i} \cdot \psi_{i}) + \beta_{{1}} (z_{1}) + \beta_{{2}} (z_{2})+ \beta_{{3}} (z_{3})+ \tau_{1} ( z_{1} \cdot z_{2} \cdot t)+ \tau_{2} ( z_{1} \cdot z_{3})+ \tau_{3} (z_{2} \cdot z_{3}), \nonumber
 \end{align}
@@
 These tables are included to provide exact measures at certain intervals.
-The models of interest are: Cox, case-base with logistic regression (CBLR), DeepSurv, DeepHit, Case-Base Neural Networks (CBNN), Optimal (a CBLR model with the exact interaction terms and baseline hazard specified) and Deep Survival Machines (DSM).}
+The models of interest are: Cox, case-base with logistic regression (CBLR), DeepSurv, DeepHit, Case-Base Neural Network (CBNN), Optimal and Deep Survival Machines (DSM).}
 \label{tab:megaTable}
 
-\begin{center}\includegraphics[width=1\linewidth]{/home/bhatnagar-lab/jislam/pmnnProjects/paper/../pmnnPlayground/reviewedAnalysisPMNN/figures/megaTable} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{/home/bhatnagar-lab/jislam/cbnnManuscript/analyses/figures/megaTable} \end{center}
 
@@
 from the Optimal regression model, CBNN outperforms the competing models
-when examining IPA, and up to the 75-percentile of follow-up time for
-c-ipcw. We expected the Optimal model to perform best in both metrics.
-However, this was not the case for c-ipcw and may be due to an artifact
-of concordance-based metrics, where a misspecified model may perform
-better than a correctly specified one \citep{cindexfails2019}. We
-attribute this difference in performance to the flexibility of CBNN in
-terms of time-varying interactions and baseline hazard, flexibility the
-other neural network models do not have, explaining their drop in
-performance.
+when examining IPA and up to the 75-percentile of follow-up time for
+\(C_{IPCW}\). We expected the Optimal model to perform best in both
+metrics. However, this was not the case for \(C_{IPCW}\) and may be due
+to an artifact of concordance-based metrics, where a misspecified model
+may perform better than a correctly specified one
+\citep{cindexfails2019}. We attribute the performance of CBNN to its
+flexibility in modeling time-varying interactions and baseline hazard,
+flexibility the other neural network models do not have.
 
@@
 aside for validation at each epoch. We predict risk functions for
-everyone in the test set, which are used to calculate our metrics. We
+everyone in the test set, which is used to calculate our metrics. We
 conduct 100 bootstrap re-samples for the real data applications to
@@
 pressure, heart rate, respiration rate, temperature, white blood cell
-count, serum's sodium, and serum's creatinine)
+count, serum's sodium and serum's creatinine)
 \citep{katzman2018DeepSurv}. Patients with missing features were
@@
 drop in performance for CBNN before the \(25^{th}\) percentile of
-follow-up. For c-ipcw, CBNN outperforms the competing models
+follow-up. For \(C_{IPCW}\), CBNN outperforms the competing models
 consistently over follow-up time. Note that performance is similar for
-all models, aside from DeepSurv whose c-ipcw is lower than the rest
-(Figure \ref{fig:megaPlot} E, F and Table \ref{tab:megaTable} C).
+all models, aside from DeepSurv whose \(C_{IPCW}\) is lower than the
+rest (Figure \ref{fig:megaPlot} E, F and Table \ref{tab:megaTable} C).
 
@@
 years of follow-up \citep{katzman2018DeepSurv}. There are 11 covariates
-in total: 4 RNA-Seq gene expressions (MKI67, EGFR, PGR, and ERBB2) and 5
+in total: 4 RNA-Seq gene expressions (MKI67, EGFR, PGR and ERBB2) and 5
 clinical features (hormone treatment indicator, radiotherapy indicator,
-chemotherapy indicator, ER-positive indicator, and age at diagnosis)
+chemotherapy indicator, ER-positive indicator and age at diagnosis)
 \citep{katzman2018DeepSurv}.
@@
 50-percentile of follow-up time, after which CBNN, DeepSurv and DSM drop
-in performance. The c-ipcw produces a different ranking. Our CBNN model
-outperforms the other models up to the \(25^{th}\) percentile of
+in performance. The \(C_{IPCW}\) produces a different ranking. Our CBNN
+model outperforms the other models up to the \(25^{th}\) percentile of
 follow-up time, whereas DeepHit performs best from the \(50^{th}\) to
@@
 outperform the regression models. This disagreement between IPA and
-c-ipcw between may be due to the misspecification issue of
+\(C_{IPCW}\) may be due to the misspecification issue of
 concordance-based metrics \citep{cindexfails2019}. The neural network
 models may be over-parameterized, as shown by the wide confidence bands
-in c-ipcw.
+in \(C_{IPCW}\).
 
@@
 
-Our method, CBNN, models survival outcomes by using neural networks on
-case-base sampled data. While estimating the hazard function, we
-incorporate follow-up-time as a feature, providing a data-driven
-estimation of a flexible baseline hazard and time-varying interactions.
-Based on our complex simulation results (Figure \ref{fig:megaPlot} C, D
-and Table \ref{tab:megaTable} B), CBNN outperforms the competitors when
-time-varying interactions and a complex baseline hazard are present.
+CBNN models survival outcomes by using neural networks on case-base
+sampled data. We incorporate follow-up time as a feature, providing a
+data-driven estimate of a flexible baseline hazard and time-varying
+interactions in our hazard function. The three competing neural network
+models we evaluated cannot model time-varying interactions by design
+\citep{dsmPaper} \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}. DSM
+requires a mixture component distributions for a flexible baseline
+hazard to be fit \citep{dsmPaper}. As our goal is to fix the design and
+compare performance, we did not change any shared hyperparameters. With
+our choice of shared hyperparameters and model design, DSM did not
+converge in the complex simulation and SUPPORT case study. Despite this
+limitation, we include this method when it did function as it can fit
+flexible baseline hazards. Compared to CBNN, the remaining two models
+also have limitations. DeepSurv is a proportional hazards model and does
+not estimate the baseline hazard \citep{katzman2018DeepSurv}. DeepHit
+requires an alpha hyperparameter, is restricted to a single distribution
+for the baseline hazard and models the survival function directly
+\citep{lee2018DeepHit}. The alternative neural network methods match on
+time, while CBNN models time directly.
 
-Each competing neural network model we evaluated has limitations or
-model-specific parameters. DeepSurv is a proportional hazards model and
-does not estimate the baseline hazard \citep{katzman2018DeepSurv}.
-DeepHit requires an alpha hyperparameter, is restricted to a single
-distribution for the baseline hazard, and models the survival function
-directly \citep{lee2018DeepHit}. DSM requires the user to specify the
-component distributions in the mixture for a flexible baseline hazard to
-be fit \citep{dsmPaper}. These alternative neural network approaches
-match on time, while CBNN models time directly.
+To assess performance among these models, we use both IPA and
+\(C_{IPCW}\) metrics. Concordance-based measures are commonly used in
+survival analysis to compare models and we opt to keep them in our
+analyses. However, \(C_{IPCW}\) is a non-proper metric and may cause
+misspecified models to appear better than they should
+\citep{cindexfails2019}. Therefore, we contextualize our \(C_{IPCW}\)
+results in relation to IPA, a proper scoring rule. The model rankings
+between \(C_{IPCW}\) and IPA differed for both the complex simulation
+and METABRIC application. Wider \(C_{IPCW}\) confidence intervals for
+DeepHit and DeepSurv show potential misspecification of the models
+(Figure \ref{fig:megaPlot} D, H).
 
-In the simple simulation, all neural network-based approaches resulted
-in a loss of performance based on IPA, while the regression-based
-approaches did not. We attribute this to potential over-parameterization
-in the neural network models. The wide confidence intervals suggest
-over-fitting may be the reason for the loss in performance, even with
-dropout. Both DeepSurv and DSM are affected, while DeepHit and CBNN are
-less so.
+With this interpretation of \(C_{IPCW}\) in mind, we assess two
+simulations with minimal noise and two case studies in real scenarios.
+The simple simulation demonstrates potential pitfalls associated with
+neural network models, particularly overparameterization. All neural
+network-based approaches performed worse than null KM model (IPA score),
+while the regression-based performed better. We attribute this to
+potential overparameterization in the neural network models, as the wide
+confidence intervals suggest over-fitting, even with dropout. Both
+DeepSurv and DSM are affected, while DeepHit and CBNN are less so. This
+is a limitation to our strategy for evaluating the methods, which uses a
+fixed study design across all assessments.
 
-In the complex simulation, CBNN has a distinct advantage over the other
-methods while examining the IPA score. The regression models do show
-improvement over the null model, while the competing neural network
-approaches do not. The time-varying interactions and complex baseline
-hazard may have caused a drop in performance for the competing neural
-network methods, compared to the regression models without interactions.
+From this baseline benchmark, we move on to a complex simulation that
+requires a method that can learn both time-varying interactions and have
+a flexible baseline hazard. Here, CBNN demonstrates a distinct advantage
+over all other methods. The regression models show improved performance
+over the null KM model, while the competing neural network models
+performed worse. Based on our complex simulation results (Figure
+\ref{fig:megaPlot} C, D and Table \ref{tab:megaTable} B), CBNN
+outperforms the competitors when time-varying interactions and a complex
+baseline hazard are present. This simulation shows how CBNN can perform
+under ideal conditions, while the following two analyses on real data
+serve to assess its performance in realistic conditions.
 
-In the SUPPORT application, the overall performance considering IPA is
-near identical. The regression models outperform CBNN, which performs
-better than the competing neural network models. As the confidence
-intervals are tight, it is unlikely that the loss in performance is
-because of over-fitting. Flexibility in both interaction modeling and
-baseline hazard is not particularly helpful with the SUPPORT dataset.
+In the SUPPORT and METABRIC case studies, flexibility in both
+interaction modeling and baseline hazard did not improve CBNNs relative
+performance, suggesting that this flexibility does not aid prediction in
+either case study. From a biological perspective, the 30-year follow-up
+time in the METABRIC study may contain competing causes of death. The
+causes at the start of the study may not match the causes towards the
+end, potentially explaining the drop in performance as we reach later
+survival times with fewer individuals. The baseline hazard is unlikely
+to cause this drop as DSM is the most flexible competing model in our
+comparison. Over-fitting is also unlikely given the tight confidence
+intervals. Further research is required to determine the cause of the
+drop in performance seen in the METABRIC case study. In both case
+studies, CBNN outperforms the competing neural network methods.
 
-In the METABRIC application IPA results, DSM and DeepSurv had a steep
-drop in performance. As DeepSurv does not depend on the baseline hazard
-during the fitting process and DSM is the most flexible competing model
-in our comparison, the baseline hazard is unlikely to cause this drop.
-With tight confidence intervals, over-fitting is unlikely as well.
-Further research would be required to determine the cause of this drop
-in performance. We do not see the same drop in CBNN demonstrating that,
-in the METABRIC application, CBNN outperforms DeepSurv and DSM.
-Considering IPA score, regression models consistently outperform the
-neural network models. From a biological perspective, the 30-year
-follow-up time in the METABRIC study may contain competing causes of
-death. The signal from the start of the study may not match the signal
-towards the end, potentially explaining the drop in performance as we
-reach later survival times with fewer individuals.
-
-Concordance-based measures are commonly used in survival analysis to
-compare models and we opt to keep them in our analyses. However, c-ipcw
-is a non-proper metric and may cause misspecified models to appear
-better than they should \citep{cindexfails2019}. The model rankings
-between c-ipcw and IPA differed for both the complex simulation and
-METABRIC application. In the complex simulation, the c-ipcw of a
-correctly specified model (Optimal) underperforms CBNN and CBLR; models
-without the exact interaction terms specified. We base our METABRIC
-assessment of this disagreement between metrics on the complex
-simulation. The IPA score ranks regression models first, followed by
-DeepHit and CBNN, but the c-ipcw score ranks regression models after
-DeepHit and CBNN. Wide confidence intervals for neural network models in
-c-ipcw show potential misspecification, suggesting that regression
-models perform best in the METABRIC application.
-
-During our testing for the complex simulation and SUPPORT application,
-DSM did not converge with our choice of hyper-parameters and model
-design. As the goal is to fix the design and compare performance, we did
-not change any shared hyperparameters. For DSM, a different number and
-combination of component distributions did not help the model converge,
-though we did not conduct an exhaustive test. This may be a limitation
-to their method or early-stage implementation issues. We include this
-method where it did function as it shares a similar goal of a flexible
-baseline hazard.
-
-If the user suspects that there may be unknown time-varying interactions
-and a complex baseline hazard, CBNN should strongly be considered. Once
-we perform case-base sampling and adjust for the sampling bias, we can
-use a sigmoid activation function to predict our hazard function. The
-user can approach their problem as they would for probabilistic modeling
-while interpreting their prediction in a survival context. In terms of
-performance, we saw that CBNN consistently performs better than the
-other neural network-based approaches over follow-up time. This may be
-because of the hyperparameters chosen, as they differ from each of the
-respective methods papers. CBNN only requires the use of standard
-components in machine learning libraries (add layer and sigmoid
-activation function). Due to the simplicity in its implementation and by
-extension user experience, we expect CBNN to be both a user-friendly
-approach to data-driven survival analysis and easily extendable to any
-feed-forward neural network frameworks.
-
-Though CBNN can model time-varying interactions and a flexible baseline
-hazard, all models require substantial amounts of data to learn real,
-complex relationships effectively. Larger datasets will appear with
-time, but the cost of generating longitudinal datasets remains. A viable
-alternative strategy is the use of transfer learning, where independent
-and identically distributed data can function as a preliminary model to
-learn related features in a parallel dataset, initializing weights for
-our survival task \citep{bozinovski2020reminder}.
-
-This paper describes CBNN in the single event setting. A next step would
-be to extend this approach to competing risks. A promising use case
-would be high-dimensional data with complex correlation structures, such
-as imaging data.
+Taken together, CBNN outperforms all competitors in the complex
+simulation, demonstrating its value in survival settings that may
+involve time-varying interactions and a complex baseline hazard. Once we
+perform case-base sampling and adjust for the sampling bias, we can use
+a sigmoid activation function to predict our hazard function. Our
+approach simplifies the incorporation of censored individuals, allowing
+survival outcomes to be treated as binary ones. Forgoing the requirement
+of custom loss functions, CBNN only requires the use of standard
+components in machine learning libraries (specifically, the add layer to
+adjust for sampling bias and the sigmoid activation function). Due to
+the simplicity in its implementation and by extension user experience,
+CBNN is both a user-friendly approach to data-driven survival analysis
+and is easily extendable to any feed-forward neural network framework.
 

### 2022-05-25 analyses/codeReviewIterSim.Rmd
cleaned, licensed version of cbnn

---
title: "iterSim"
output: html_document
params:
  epo: 2
  patience: 1
  iteration: 2
  layer1: 50
  layer2: 50
  layer3: 25
  layer4: 25
  drpt: 0.5
---

```{r}

library(reticulate)
```

```{python}

import os
os.environ["CUDA_DEVICE_ORDER"] = "PCI_BUS_ID" 
os.environ["CUDA_VISIBLE_DEVICES"] = ""

```

```{r setup, include=T}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
source("src/packages.R")
source("src/functions.R")
eval_support<-T
source_python('src/getSUP.py')
source_python("src/cIndex.py")
source_python("src/dsmScript.py")
source_python('src/deephitter.py')
##########################
###Shared Hyperparameters
##########################
bsize=512
min_delta = 10^-7
epo=as.numeric(params$epo)
patience <- as.numeric(params$patience)
iteration<-as.numeric(params$iteration)
layer1<-as.numeric(params$layer1)
layer2<-as.numeric(params$layer2)
layer3<-as.numeric(params$layer3)
layer4<-as.numeric(params$layer4)
drpt<-as.numeric(params$drpt)

easysim<-F
hardsim<-T
numSamp<-2000
```

```{r easySimpletests,eval=easysim}
start_time <- Sys.time()

BrierIPAList<-list()
cScore<-list()

pos<-1
for(iter in 1:iteration){

  #############################
  ###Data generation
  #############################

  cov<-data.frame(id=1:numSamp,
                  z1=c(rep(1,numSamp/2),rep(0,numSamp/2)),
                  z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
                  z3=c(rnorm(numSamp/2,-3,0.5),rnorm(numSamp/2,8,0.5))
  )

  dat <- simsurv(dist = "exponential",
                 lambdas =.05,
                 betas = c(z1 = .5,z2=.5,z3=.5),
                 x = cov,
                 tdefunction = "log",
                 maxt = 1
  )
  colnames(dat)<-c("id","time","status")
  samp<-sample(seq(1,nrow(dat),by=1), floor(nrow(dat)*(0.1))) #random censoring
  dat$status[samp]<-0
  # Merge the simulated event times onto covariate data frame
  data <- merge(cov, dat)
  covs<-colnames(data)

  ##########################
  ###Split data
  ##########################
  samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.8))
  data<-data[,-1]
  data[,-c(1,5)]<-as.data.frame(sapply(data[,-c(1,5)], function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
  train<-data[samp,]
  test<-data[-samp,-c(ncol(data))]
  fullTest<-data[-samp,,drop=F]
  times<-seq(from=0,
             to=max(test$time),
             length.out = 200
  )
  times<-times[-c(1,length(times))]

  ##########################
  ###DSM
  ##########################
  source_python('src/dsmScript.py')
  dsmSurv<-dsmfitter(train=train,fullTest=fullTest,bsize=bsize,epochs=epo,tPos=ncol(train)-2,ePos=ncol(train)-1)
  if((sum(is.na(dsmSurv))==ncol(dsmSurv)*nrow(dsmSurv))){

    dsmSurv<-dsmfitter(train=train,fullTest=fullTest,bsize=bsize,epochs=epo)
    py_run_string("del dsmfitter")
    rm(dsmfitter)
    iteration<-iteration+1
    next
  }
  colnames(dsmSurv) <-seq(0.01,1,by=0.01)
  dsmSurvProper<-pyProcess(dsmSurv,times=times)
  colnames(dsmSurvProper) <-seq(0,1,length.out = ncol(dsmSurvProper))
  py_run_string("del dsmfitter")
  rm(dsmfitter)

  ##########################
  ###casebase optimal model
  ##########################
  final_mod_cb_glm <- fitSmoothHazard(status~z1+z2+z3,#z1:z2:z3+z1+z2+z3,+z1:z3+z1:z2:z3
                                      data = train,
                                      time = "time",
                                      event="status",
                                      ratio = 100
  )

  summary(final_mod_cb_glm)
  tglmAbsRisk<-absoluteRisk(final_mod_cb_glm,time=times,newdata=test,type="CI")
  tglmAbsRisk<-as.data.frame(tglmAbsRisk)
  rownames(tglmAbsRisk)<-tglmAbsRisk$time
  trueGlmProper<- t(tglmAbsRisk[-1,-1])

  class(trueGlmProper)<-c("tunnel",class(trueGlmProper))

  ##########################
  ###casebase + splines
  ##########################
  mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
                                data = train,
                                time = "time",
                                event="status",
                                ratio = 100
  )
  glmAbsRisk<-absoluteRisk(mod_cb_glm,time=times,newdata=test,type="CI")
  glmAbsRisk<-as.data.frame(glmAbsRisk)
  rownames(glmAbsRisk)<-glmAbsRisk$time
  glmProper<- t(glmAbsRisk[-1,-1])
  class(glmProper)<-c("tunnel",class(glmProper))

  ############################
  ###DeepSurv
  ############################  
  coxnn<-deepsurv(Surv(time, status) ~ .,data = train, frac = 0.2, activation = "relu",
                  num_nodes = c(layer1,layer2,layer3,layer4), dropout = drpt,
                  early_stopping = TRUE, epochs = epo,
                  batch_size = bsize,device = "cpu",best_weights=T,
                  patience = patience,min_delta=min_delta,batch_norm = F,
                  learning_rate = 0.001,lr_decay=10^-7)

  coxnnsurv<- predict(coxnn,newdata = test,type="survival")
  colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
  cnnCleaned<-pyProcess(coxnnsurv,times=times)
  colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
  ############################
  ###Deephit
  ############################  
  source_python('src/deephitter.py')
  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,patience=patience,min_delta=min_delta,drpt=drpt)
  hitnnSurv<-t(hitnnSurv)
  colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
  deephitCleaned<-pyProcess(hitnnSurv,times=times)
  colnames(deephitCleaned)<-seq(0,1,length.out = ncol(deephitCleaned))
  py_run_string("del fitDeephit")
  rm(fitDeephit)

  ############################
  ###PMNNsplines
  ############################  
  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2+3),
                            name = 'main_input'
  )
  covars_output<-covars_input%>% 
    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=1,use_bias = T)

  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                  feature_input = covars_input,
                  feature_output = covars_output,
                  originalData = mod_cb_glm$data,
                  offset=mod_cb_glm$offset,
                  timeVar = "time",
                  eventVar= "status"
  )

  fit<-fitSmoothHazSpline(pmnn,
                          epochs=epo,
                          batch_size=bsize,
                          verbose=0,
                          monitor="val_loss",
                          val_split=0.2,
                          min_delta=min_delta,
                          patience=patience)
  annPreds<-aarSplines(fit,
                       times=times,
                       x_test=test
  )
  rownames(annPreds)<-annPreds[,1]
  annProper<- t(annPreds[,-1])
  class(annProper)<-c("tunnel",class(annProper))

  ######################
  ###coxph
  ######################
  cox<-coxph(Surv(time, status) ~ z1+z2+z3, data = train,x=T)

  ######################
  ###PMNN
  ######################
  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
                            name = 'main_input')

  covars_output<-covars_input%>% #layer_batch_normalization()%>%
    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=1,use_bias = T)

  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                  feature_input = covars_input,
                  feature_output = covars_output,
                  originalData = mod_cb_glm$data,
                  offset=mod_cb_glm$offset,
                  timeVar = "time",
                  eventVar= "status"
  )

  fit<-fitSmoothHaz(pmnn,
                    epochs=epo,
                    batch_size=bsize,
                    verbose=0,
                    monitor="val_loss",
                    val_split=0.2,
                    min_delta=min_delta,
                    patience=patience)
  annPreds<-aar(fit,
                times=times,
                x_test=test
  )

  rownames(annPreds)<-annPreds[,1]
  annProperpoly<- t(annPreds[,-1])
  class(annProperpoly)<-c("tunnel",class(annProperpoly))

  rownames(annPreds)<-annPreds[,1]
  annProperpoly<- t(annPreds[,-1])

  class(annProperpoly)<-c("tunnel",class(annProperpoly)) 
  brierFinalResults <- Score(list("Cox_Lin" = cox,
                                  'CB_Logi'=glmProper,
                                  'DeepSurv'=cnnCleaned,
                                  'DeepHit'=deephitCleaned,
                                  'PMNN_Spline'=annProper,
                                  'Optimal'=trueGlmProper,
                                  'PMNN_Poly'=annProperpoly,
                                  'dsm'=dsmSurvProper) ,#'dsm'=dsmCleaned,
                                  data =fullTest, 
                             formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)
  BrierIPAList[[pos]]<-brierFinalResults$Brier$score
  ########################
  ###C-Index
  ########################
  tempTims<-   as.numeric(colnames(annProper))
  tempTims<-head(tempTims, -15)
  tempTims<-tail(tempTims, -10)
  cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=9)
  et_train<-as.matrix(train)#[,c(5,4)]
  et_test<-as.matrix(fullTest[,c(5,4)])
  colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
                          'PMNN_Spline','Optimal','PMNN_Poly','dsm')
  cScoreTemp[,1]<- tempTims

  a<-survfit(cox, newdata=fullTest,times=times)
  rownames(a$surv)<-a$time
  coxlinSurv<-pyProcess(t(a$surv),times=times)
  colnames(coxlinSurv)<-seq(0,1,length.out = ncol(coxlinSurv))

  riskList<-list(coxlinSurv,
                 glmProper,
                 cnnCleaned,
                 deephitCleaned,
                 annProper,
                 annProperpoly
  )

  cScore[[pos]]<-cIndexSummary(et_train=et_train
                               , et_test=et_test,
                               riskList=riskList,
                               cScore=cScoreTemp
  )

  saveRDS(BrierIPAList,'100easyBrier.rds')
  saveRDS(cScore,'100easycidx.rds')
  pos<-pos+1
}

end_time <- Sys.time()

```

```{r complexPrep,eval=hardsim}

##############################
###set up complex simulation
##############################

# Define a function returning the log cum hazard at time t
logcumhaz <- function(t, x, betas, knots) {

  # Obtain the basis terms for the spline-based log
  # cumulative hazard (evaluated at time t)
  basis <- flexsurv::basis(knots, log(t))

  # Evaluate the log cumulative hazard under the
  # Royston and Parmar specification
  res <-betas[["gamma0"]] * basis[[1]] + 
    betas[["gamma1"]] * basis[[2]] +
    betas[["gamma2"]] * basis[[3]] +
    betas[["gamma3"]] * basis[[4]] +
    betas[["gamma4"]] * basis[[5]] +
    betas[["z1"]] * x[["z1"]]+
    betas[["z2"]] * x[["z2"]]+
    betas[["z3"]] * x[["z3"]]+
    betas[["term1"]] * x[["z1"]]*x[["z2"]]*t+
    betas[["term2"]] * x[["z1"]]*x[["z3"]]+
    betas[["term3"]] * x[["z2"]]*(x[["z3"]])
  #
  # Return the log cumulative hazard at time t
  res
}
url<-'app.txt'
ds <- read.table(url, sep = '',header = TRUE)
ds$years.at.risk=pmin(ds$agein80,ds$onset)
brcancer$rectime<-brcancer$rectime/max(brcancer$rectime)
true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1, data = brcancer, k = 3) #from flexsurv

# Simulate the event times
coefficients<-true_mod$coefficients
coefficients<-c(coefficients,z1=1,z2=1,z3=1,term1=10,term2=2,term3=2)

```

```{r totalSim,eval=hardsim}

start_time <- Sys.time()
BrierIPAList<-list()
cScore<-list()
pos<-1
for(iter in 1:iteration){

  #########################
  ###Simulate date
  #########################
  cov<-data.frame(id=1:numSamp,
                  z1=c(rep(1,numSamp/2),rep(0,numSamp/2)),
                  z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
                  z3=c(rnorm(numSamp/2,-3,0.5),rnorm(numSamp/2,8,0.5)))
  true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1,
                                       data = brcancer, k = 3
  ) #from flexsurv
  dat <- simsurv(betas = coefficients, # "true" parameter values
                 x = cov,                   # covariate data for 200 individuals
                 knots = true_mod$knots,    # knot locations for splines
                 logcumhazard = logcumhaz,  # definition of log cum hazard
                 maxt = 1,               # no right-censoring
                 interval = c(0,10000000)) # interval for root finding
  colnames(dat)<-c("id","time","status")
  samp<-sample(seq(1,nrow(dat),by=1), floor(nrow(dat)*(0.1))) #random censoring
  dat$status[samp]<-0
  # Merge the simulated event times onto covariate data frame
  data <- merge(cov, dat)
  covs<-colnames(data)

  #########################
  ###Simulate data
  #########################
  samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.8))
  data<-data[,-1]
  data[,-c(1,5)]<-as.data.frame(sapply(data[,-c(1,5)], function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
  train<-data[samp,]
  test<-data[-samp,-c(ncol(data))]
  fullTest<-data[-samp,,drop=F]
  times<-seq(from=0,
             to=max(test$time),
             length.out = 200
  )
  times<-times[-c(1,length(times))]

  ###########################
  ###casebase optimal model
  ###########################
  final_mod_cb_glm <- fitSmoothHazard(status~bs(time)-time+z1+z2+z3+z1:z2:time+z1:z3+z2:z3,
                                      data = train,
                                      time = "time",
                                      event="status",
                                      ratio = 100
  )

  summary(final_mod_cb_glm)
  tglmAbsRisk<-absoluteRisk(final_mod_cb_glm,time=times,newdata=test,type="CI")
  tglmAbsRisk<-as.data.frame(tglmAbsRisk)
  rownames(tglmAbsRisk)<-tglmAbsRisk$time
  trueGlmProper<- t(tglmAbsRisk[-1,-1])
  class(trueGlmProper)<-c("tunnel",class(trueGlmProper))

  #########################
  ###casebase+splines
  #########################
  mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
                                data = train,
                                time = "time",
                                event="status",
                                ratio = 100
  )

  glmAbsRisk<-absoluteRisk(mod_cb_glm,time=times,newdata=test,type="CI")

  glmAbsRisk<-as.data.frame(glmAbsRisk)

  rownames(glmAbsRisk)<-glmAbsRisk$time
  glmProper<- t(glmAbsRisk[-1,-1])

  class(glmProper)<-c("tunnel",class(glmProper))

  ######################
  ###PMNN splines
  ######################
  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2+3),
                            name = 'main_input'
  )
  covars_output<-covars_input%>% #layer_batch_normalization()%>%
    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=1,use_bias = T)

  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                  feature_input = covars_input,
                  feature_output = covars_output,
                  originalData = mod_cb_glm$data,
                  offset=mod_cb_glm$offset,
                  timeVar = "time",
                  eventVar= "status"
  )

  fit<-fitSmoothHazSpline(pmnn,
                          epochs=epo,
                          batch_size=bsize,
                          verbose=0,
                          monitor="val_loss",
                          val_split=0.2,
                          min_delta=min_delta,
                          patience=patience)
  annPreds<-aarSplines(fit,
                       times=times,
                       x_test=test
  )
  rownames(annPreds)<-annPreds[,1]
  annProper<- t(annPreds[,-1])
  class(annProper)<-c("tunnel",class(annProper))

  ######################
  ###coxph
  ######################
  cox<-coxph(Surv(time, status) ~ z1+z2+z3, data = train,x=T)

  ######################
  ###DeepSurv
  ######################
  coxnn<-deepsurv(Surv(time, status) ~ .,data = train, frac = 0.2, activation = "relu",
                  num_nodes = c(layer1,layer2,layer3,layer4), dropout = drpt,
                  early_stopping = TRUE, epochs = epo,
                  batch_size = bsize,device = "cpu",best_weights=T,
                  patience = patience,min_delta=min_delta,batch_norm = F,
                  learning_rate = 0.001,lr_decay=10^-7)
  coxnnsurv<- predict(coxnn,newdata = test,type="survival")
  colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
  cnnCleaned<-pyProcess(coxnnsurv,times=times)
  colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))

  ######################
  ###Deephit
  ######################
  source_python('src/deephitter.py')
  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,patience=patience,min_delta=min_delta,drpt=drpt)
  hitnnSurv<-t(hitnnSurv)
  colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
  deephitCleaned<-pyProcess(hitnnSurv,times=times)
  colnames(deephitCleaned)<-seq(0,1,length.out = ncol(deephitCleaned))
  py_run_string("del fitDeephit")
  rm(fitDeephit)

  ######################
  ###PMNN
  ######################
  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
                            name = 'main_input')

  covars_output<-covars_input%>% #layer_batch_normalization()%>%
    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=1,use_bias = T)

  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                  feature_input = covars_input,
                  feature_output = covars_output,
                  originalData = mod_cb_glm$data,
                  offset=mod_cb_glm$offset,
                  timeVar = "time",
                  eventVar= "status"
  )

  fit<-fitSmoothHaz(pmnn,
                    epochs=epo,
                    batch_size=bsize,
                    verbose=0,
                    monitor="val_loss",
                    val_split=0.2,
                    min_delta=min_delta,
                    patience=patience)
  annPreds<-aar(fit,
                times=times,
                x_test=test
  )

  rownames(annPreds)<-annPreds[,1]
  annProperpoly<- t(annPreds[,-1])
  class(annProperpoly)<-c("tunnel",class(annProperpoly))

  brierFinalResults <- Score(list("Cox_Lin" = cox,
                                  'CB_Logi'=glmProper,
                                  'DeepSurv'=cnnCleaned,
                                  'DeepHit'=deephitCleaned,
                                  'PMNN_Spline'=annProper,
                                  'Optimal'=trueGlmProper,
                                  'PMNN_Poly'=annProperpoly),
                             #'dsm'=dsmSurvProper) ,#'dsm'=dsmCleaned,
                             data =fullTest, 
                             formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)
  BrierIPAList[[pos]]<-brierFinalResults$Brier$score

  ######################
  ###C-Index
  ######################
  tempTims<-   as.numeric(colnames(annProper))
  tempTims<-head(tempTims, -15)
  tempTims<-tail(tempTims, -10)
  cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=8)
  et_train<-as.matrix(train)#[,c(5,4)]
  et_test<-as.matrix(fullTest[,c(5,4)])
  colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
                          'PMNN_Spline','Optimal','PMNN_Poly')#,'dsm')
  cScoreTemp[,1]<- tempTims
  a<-survfit(cox, newdata=fullTest,times=times)
  rownames(a$surv)<-a$time
  coxlinSurv<-pyProcess(t(a$surv),times=times)
  colnames(coxlinSurv)<-seq(0,1,length.out = ncol(coxlinSurv))
  riskList<-list(coxlinSurv,
                 glmProper,
                 cnnCleaned,
                 deephitCleaned,
                 annProper,
                 annProperpoly
  )

  cScore[[pos]]<-cIndexSummary(et_train=et_train
                               , et_test=et_test,
                               riskList=riskList,
                               cScore=cScoreTemp
  )

  saveRDS(BrierIPAList,'results/100hardBrier.rds')
  saveRDS(cScore,'results/100hardcidx.rds')
  pos<-pos+1
}

```

### 2022-05-25 analyses/codeReviewRealData.Rmd
cleaned, licensed version of cbnn

---
title: "realData"
output: html_document
---

```{r,eval=T}

library(reticulate)
```

```{python,eval=T}

import os
os.environ["CUDA_DEVICE_ORDER"] = "PCI_BUS_ID" 
os.environ["CUDA_VISIBLE_DEVICES"] = ""
#os.environ["OM_NUM_THREADS"]="2"

```

```{r setup, include=T}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
source("src/packages.R")
source("src/functions.R")
source_python("src/cIndex.py")
source_python("src/dsmScript.py")
source_python('src/deephitter.py')
source_python("src/cIndex.py")
#tf$config$threading$set_intra_op_parallelism_threads(1)
#tf$config$threading$set_inter_op_parallelism_threads(1)

#set.seed(2)

epo=2
bsize=512
min_delta = 10^-7
patience = 1
iteration=5

eval_support<-T
source_python('src/getSUP.py')

########################
###sharedHyperparams
########################

layer1<-50
layer2<-50
layer3<-25
layer4<-25
drpt<-0.5

```

```{r modelsSUPPORT,eval=eval_support}

start_time <- Sys.time()
BrierIPAList<-list()
cScore<-list()

data<- as.data.frame(do.call(cbind, getSupportPycox()))
colnames(data)[ncol(data)-1]<-'time'
colnames(data)[ncol(data)]<-'status'
data<-as.data.frame(model.matrix(~.,data)[,-1])
data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))

covs<-colnames(data)

#split training/testing
samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.80)) 
trainOG<-data[samp,]
test<-data[-samp,-c(ncol(data))]
fullTest<-data[-samp,,drop=F]

pos=1
for(iter in 1:iteration){

  samp<-sample(seq(1,nrow(trainOG),by=1) , nrow(trainOG),replace = T) 
  train<-trainOG[samp,]

  times<-seq(from=0,
             to=max(test$time),
             length.out = 200
  )
  times<-times[-c(1,length(times))]

  ######################
  ###casebase + splines
  ######################
  mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
                                data = train,
                                time = "time",
                                event="status",
                                ratio = 100
  )

  glmAbsRisk<-as.data.frame(absoluteRisk(mod_cb_glm,time=times,newdata=test,type="CI"))
  rownames(glmAbsRisk)<-glmAbsRisk$time
  #prepare to run through riskRegression
  glmProper<- t(glmAbsRisk[-1,-1])
  class(glmProper)<-c("tunnel",class(glmProper))

  ######################
  ###DeepSurv
  ######################
  coxnn<-deepsurv(Surv(time, status) ~ .,data = train, frac = 0.2, activation = "relu",
                  num_nodes = c(layer1,layer2,layer3,layer4,), dropout = drpt,
                  early_stopping = TRUE, epochs = epo,
                  batch_size = bsize,device = "cpu",best_weights=T,
                  patience = patience,min_delta=min_delta,batch_norm = F,
                  learning_rate = 0.001,lr_decay=10^-7)
  coxnnsurv<- predict(coxnn,newdata = test,type="survival")
  colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
  cnnCleaned<-pyProcess(coxnnsurv,times=times)
  colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))

  ######################
  ###Deephit
  ######################
  source_python('deephitter.py')
  hitnnSurv=fitDeephit(train,fullTest,bsize,epo)
  hitnnSurv<-t(hitnnSurv)
  colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
  deephitCleaned<-pyProcess(hitnnSurv,times=times)
  colnames(deephitCleaned)<-seq(0,1,length.out = ncol(deephitCleaned))
  py_run_string("del fitDeephit")
  rm(fitDeephit)

  ######################
  ###PMNN splines
  ######################
  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2+3),
                            name = 'main_input'
  )
  covars_output<-covars_input%>% #layer_batch_normalization()%>%
    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=1,use_bias = T)

  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                  feature_input = covars_input,
                  feature_output = covars_output,
                  originalData = mod_cb_glm$data,
                  offset=mod_cb_glm$offset,
                  timeVar = "time",
                  eventVar= "status"
  )

  fit<-fitSmoothHazSpline(pmnn,epochs=epo,batch_size = bsize,verbose=0)
  annPreds<-aarSplines(fit,
                       times=times,
                       x_test=test
  )
  rownames(annPreds)<-annPreds[,1]
  annProper<- t(annPreds[,-1])
  class(annProper)<-c("tunnel",class(annProper))

  ######################
  ###coxph
  ######################
  cox<-coxph(Surv(time, status) ~ ., data = train,x=T)

  ######################
  ###PMNN
  ######################
  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
                            name = 'main_input')

  covars_output<-covars_input%>% #layer_batch_normalization()%>%
    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=1,use_bias = T)

  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                  feature_input = covars_input,
                  feature_output = covars_output,
                  originalData = mod_cb_glm$data,
                  offset=mod_cb_glm$offset,
                  timeVar = "time",
                  eventVar= "status"
  )

  fit<-fitSmoothHaz(pmnn,epochs=epo,batch_size = bsize,verbose=0)
  annPreds<-aar(fit,
                times=times,
                x_test=test
  )

  rownames(annPreds)<-annPreds[,1]
  annProperpoly<- t(annPreds[,-1])
  class(annProperpoly)<-c("tunnel",class(annProperpoly))

  ######################
  ###IPA calculation
  ######################
  brierFinalResults <- Score(list("Cox_Lin" = cox,'CB_Logi'=glmProper,
                                  'DeepSurv'=cnnCleaned,'DeepHit'=deephitCleaned,
                                  'PMNN_Spline'=annProper,'PMNN_Poly'=annProperpoly),
                             data =fullTest, 
                             formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)
  BrierIPAList[[pos]]<-brierFinalResults$Brier$score

  ######################
  ###C-Index
  ######################
  tempTims<-as.numeric(colnames(annProper))
  tempTims<-head(tempTims, -15)
  tempTims<-tail(tempTims, -10)
  cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=7)

  et_train<-as.matrix(train)
  et_test<-as.matrix(fullTest[,c(11,10)])
  colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
                          'PMNN_Spline','PMNN_Poly')
  cScoreTemp[,1]<- tempTims

  ### Cox is not calculated, so abrisk is calculated here.
  a<-survfit(cox, newdata=fullTest,times=times)
  rownames(a$surv)<-a$time
  coxlinSurv<-pyProcess(t(a$surv),times=times)
  colnames(coxlinSurv)<-seq(0,1,length.out = ncol(coxlinSurv))
  riskList<-list(coxlinSurv,
                 glmProper,
                 cnnCleaned,
                 deephitCleaned,
                 annProper,
                 annProperpoly
  )

  cScore[[pos]]<-cIndexSummary(et_train=et_train
                               , et_test=et_test,
                               riskList=riskList,
                               cScore=cScoreTemp
  )
  pos=pos+1
  saveRDS(BrierIPAList,'100supBrier.rds')
  saveRDS(cScore,'100supcidx.rds')

}

ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=Brier,col=model))+
  geom_line()
ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
  geom_line()

end_time <- Sys.time()

end_time-start_time

```

```{r modelsMETABRIC,eval=eval_metabric}

start_time <- Sys.time()

#pycox<-reticulate::import("pycox")
#numpy<-reticulate::import("numpy")
#a<-pycox$datasets$metabric$read_df
#b<-numpy$asarray(a)

############################
###Data clean and prep
############################
metabric<-getMetabric()
colnames(metabric)[ncol(metabric)-1]<-'time'
colnames(metabric)[ncol(metabric)]<-'status'
data<-as.data.frame(model.matrix(~.,metabric)[,-1])
data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
covs<-colnames(data)

samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.80)) 
trainOG<-data[samp,]
test<-data[-samp,-c(ncol(data))]

fullTest<-data[-samp,,drop=F]

############################
###Prep Loop
############################
BrierIPAList<-list()
cScore<-list()
pos=1

for(iter in 1:iteration){

  ############################
  ###Bootstrap samples
  ############################  
  samp<-sample(seq(1,nrow(trainOG),by=1) , nrow(trainOG),replace = T) 
  train<-trainOG[samp,]

  times<-seq(from=0,
             to=max(test$time),
             length.out = 200
  )
  times<-times[-c(1,length(times))]
  times<-unique(times)

  ############################
  ###DSM
  ############################  
  source_python('dsmScript.py')
  dsmSurv<-dsmfitter(train=train,fullTest=fullTest,bsize=bsize,epochs=epo,tPos=ncol(train)-2,ePos=ncol(train)-1)
  #sometimes, dsm fails to converge so we keep trying until it succeeds
  if((sum(is.na(dsmSurv))==ncol(dsmSurv)*nrow(dsmSurv))){
    dsmSurv<-dsmfitter(train=train,fullTest=fullTest,bsize=bsize,epochs=epo,tPos=ncol(train)-2,ePos=ncol(train)-1)
    py_run_string("del dsmfitter")
    rm(dsmfitter)
    iteration<-iteration+1
    next
  }
  colnames(dsmSurv) <-seq(0.01,1,by=0.01)
  dsmSurvProper<-pyProcess(dsmSurv,times=times)
  colnames(dsmSurvProper) <-seq(0,1,length.out = ncol(dsmSurvProper))
  py_run_string("del dsmfitter")
  rm(dsmfitter)

  ############################
  ###casebase + splines
  ############################  
  mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
                                data = train,
                                time = "time",
                                event="status",
                                ratio = 100
  )

  glmAbsRisk<-absoluteRisk(mod_cb_glm,time=times,newdata=test,type="CI")
  glmAbsRisk<-as.data.frame(glmAbsRisk)
  rownames(glmAbsRisk)<-glmAbsRisk$time
  glmProper<- t(glmAbsRisk[-1,-1])
  class(glmProper)<-c("tunnel",class(glmProper))

  ############################
  ###DeepSurv
  ############################  
  coxnn<-deepsurv(Surv(time, status) ~ .,data = train, frac = 0.2, activation = "relu",
                  num_nodes = c(layer1,layer2,layer3,layer4), dropout = drpt,
                  early_stopping = TRUE, epochs = epo,
                  batch_size = bsize,device = "cpu",best_weights=T,
                  patience = patience,min_delta=min_delta,batch_norm = F,
                  learning_rate = 0.001,lr_decay=10^-7)

  coxnnsurv<- predict(coxnn,newdata = test,type="survival")
  colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
  cnnCleaned<-pyProcess(coxnnsurv,times=times)
  colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
  ############################
  ###Deephit
  ############################  
  source_python('deephitter.py')
  hitnnSurv=fitDeephit(train,fullTest,bsize,10)
  hitnnSurv<-t(hitnnSurv)
  colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
  deephitCleaned<-pyProcess(hitnnSurv,times=times)
  colnames(deephitCleaned)<-seq(0,1,length.out = ncol(deephitCleaned))
  py_run_string("del fitDeephit")
  rm(fitDeephit)

  ############################
  ###PMNNsplines
  ############################  
  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2+3),
                            name = 'main_input'
  )
  covars_output<-covars_input%>% 
    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=1,use_bias = T)

  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                  feature_input = covars_input,
                  feature_output = covars_output,
                  originalData = mod_cb_glm$data,
                  offset=mod_cb_glm$offset,
                  timeVar = "time",
                  eventVar= "status"
  )

  fit<-fitSmoothHazSpline(pmnn,epochs=epo,batch_size = bsize,verbose=0)
  annPreds<-aarSplines(fit,
                       times=times,
                       x_test=test
  )

  rownames(annPreds)<-annPreds[,1]
  annProper<- t(annPreds[,-1])
  class(annProper)<-c("tunnel",class(annProper))

  ############################
  ###coxph
  ############################  
  cox<-coxph(Surv(time, status) ~ ., data = train,x=T)

  ############################
  ###PMNN
  ############################  
  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
                            name = 'main_input')

  covars_output<-covars_input%>% #layer_batch_normalization()%>%
    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=1,use_bias = T)

  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                  feature_input = covars_input,
                  feature_output = covars_output,
                  originalData = mod_cb_glm$data,
                  offset=mod_cb_glm$offset,
                  timeVar = "time",
                  eventVar= "status"
  )
  fit<-fitSmoothHaz(pmnn,epochs=epo,batch_size = bsize,verbose=0)
  annPreds<-aar(fit,
                times=times,
                x_test=test
  )
  rownames(annPreds)<-annPreds[,1]
  annProperpoly<- t(annPreds[,-1])
  class(annProperpoly)<-c("tunnel",class(annProperpoly)) 
  ############################
  ###Brier Score
  ############################  
  brierFinalResults <- Score(list("Cox_Lin" = cox,'CB_Logi'=glmProper,
                                  'DeepSurv'=cnnCleaned,'DeepHit'=deephitCleaned,
                                  'PMNN_Spline'=annProper,'PMNN_Poly'=annProperpoly,
                                  'dsm'=dsmSurvProper),
                             data =fullTest, 
                             formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)
  BrierIPAList[[pos]]<-brierFinalResults$Brier$score

  ############################
  ###C-Index
  ############################  
  tempTims<-   as.numeric(colnames(annProper))
  tempTims<-head(tempTims, -15)
  tempTims<-tail(tempTims, -10)
  cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=8)
  et_train<-as.matrix(train)
  et_test<-as.matrix(fullTest[,c(11,10)])
  colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
                          'PMNN_Spline','PMNN_Poly','dsm')

  cScoreTemp[,1]<- tempTims

  a<-survfit(cox, newdata=fullTest,times=times)
  rownames(a$surv)<-a$time
  coxlinSurv<-pyProcess(t(a$surv),times=times)
  colnames(coxlinSurv)<-seq(0,1,length.out = ncol(coxlinSurv))
  riskList<-list(coxlinSurv,
                 glmProper,
                 cnnCleaned,
                 deephitCleaned,
                 annProper,
                 annProperpoly
  )

  cScore[[pos]]<-cIndexSummary(et_train=et_train
                               , et_test=et_test,
                               riskList=riskList,
                               cScore=cScoreTemp
  )
  saveRDS(BrierIPAList,'100MetaBrier.rds')
  saveRDS(cScore,'100Metacidx.rds')
  pos=pos+1

} 
end_time <- Sys.time()

end_time-start_time

ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=Brier,col=model))+
  geom_line()
ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
  geom_line()

```

```{r ERSPCproofofconcept,eval=F}

data<-ERSPC
colnames(data)<-c("ScrArm","time","status")

mod_cb_glm <- fitSmoothHazard(status~.,
                              data =data,
                              time = "time",
                              event="status",
                              ratio = 100
)

covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
                          name = 'main_input')

covars_output<-covars_input%>% #layer_batch_normalization()%>%
  layer_dense(units=1,use_bias = T)

pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                feature_input = covars_input,
                feature_output = covars_output,
                originalData = mod_cb_glm$data,
                offset=mod_cb_glm$offset,
                timeVar = "time",
                eventVar= "status",
                optimizer= 'rmsprop'
)

fit<-fitSmoothHaz(pmnn,epochs=12000,#15000
                  batch_size = 159893,#159893,#
                  verbose=0,monitor = "loss",val_split=0)
times<-seq(0,max(data$time),by=0.01)
get_weights(fit$model)
coef(mod_cb_glm)
annPreds<-aar(fit,
              times=times,
              x_test=data[1,]
)
#absoluteRisk(mod_cb_glm,time=times,newdata=data[1,],type="CI")

plot(annPreds,type="l")
lines(absoluteRisk(mod_cb_glm,time=times,newdata=data[1,],type="CI"))

results<-data.frame(intercept=c(coef(mod_cb_glm)[1],get_weights(fit$model)[[2]][1]),
                    ScrArm=c(coef(mod_cb_glm)[2],get_weights(fit$model)[[1]][1,1]),
                    time=c(coef(mod_cb_glm)[3],get_weights(fit$model)[[1]][2,1]))
rownames(results)<-c("Casebase","PMNN")
results<-round(results,2)

saveRDS(results,"../paper/figures/proofOfConceptWeights.rds")

```

```{r}

sessionInfo()

```

### 2022-05-25 analyses/complexSim.Rmd
cleaned, licensed version of cbnn

--- 2022-05-25 analyses/codeReviewIterSim.Rmd
@@
 params:
-  epo: 2
-  patience: 1
+  epo: 2000
+  patience: 10
   iteration: 2
@@
 
-easysim<-F
 hardsim<-T
 numSamp<-2000
-```
-
-```{r easySimpletests,eval=easysim}
-start_time <- Sys.time()
-
-BrierIPAList<-list()
-cScore<-list()
-
-pos<-1
-for(iter in 1:iteration){
-
-  #############################
-  ###Data generation
-  #############################
-
-  cov<-data.frame(id=1:numSamp,
-                  z1=c(rep(1,numSamp/2),rep(0,numSamp/2)),
-                  z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
-                  z3=c(rnorm(numSamp/2,-3,0.5),rnorm(numSamp/2,8,0.5))
-  )
-
-  dat <- simsurv(dist = "exponential",
-                 lambdas =.05,
-                 betas = c(z1 = .5,z2=.5,z3=.5),
-                 x = cov,
-                 tdefunction = "log",
-                 maxt = 1
-  )
-  colnames(dat)<-c("id","time","status")
-  samp<-sample(seq(1,nrow(dat),by=1), floor(nrow(dat)*(0.1))) #random censoring
-  dat$status[samp]<-0
-  # Merge the simulated event times onto covariate data frame
-  data <- merge(cov, dat)
-  covs<-colnames(data)
-
-  ##########################
-  ###Split data
-  ##########################
-  samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.8))
-  data<-data[,-1]
-  data[,-c(1,5)]<-as.data.frame(sapply(data[,-c(1,5)], function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
-  train<-data[samp,]
-  test<-data[-samp,-c(ncol(data))]
-  fullTest<-data[-samp,,drop=F]
-  times<-seq(from=0,
-             to=max(test$time),
-             length.out = 200
-  )
-  times<-times[-c(1,length(times))]
-
-  ##########################
-  ###DSM
-  ##########################
-  source_python('src/dsmScript.py')
-  dsmSurv<-dsmfitter(train=train,fullTest=fullTest,bsize=bsize,epochs=epo,tPos=ncol(train)-2,ePos=ncol(train)-1)
-  if((sum(is.na(dsmSurv))==ncol(dsmSurv)*nrow(dsmSurv))){
-
-    dsmSurv<-dsmfitter(train=train,fullTest=fullTest,bsize=bsize,epochs=epo)
-    py_run_string("del dsmfitter")
-    rm(dsmfitter)
-    iteration<-iteration+1
-    next
-  }
-  colnames(dsmSurv) <-seq(0.01,1,by=0.01)
-  dsmSurvProper<-pyProcess(dsmSurv,times=times)
-  colnames(dsmSurvProper) <-seq(0,1,length.out = ncol(dsmSurvProper))
-  py_run_string("del dsmfitter")
-  rm(dsmfitter)
-
-  ##########################
-  ###casebase optimal model
-  ##########################
-  final_mod_cb_glm <- fitSmoothHazard(status~z1+z2+z3,#z1:z2:z3+z1+z2+z3,+z1:z3+z1:z2:z3
-                                      data = train,
-                                      time = "time",
-                                      event="status",
-                                      ratio = 100
-  )
-
-  summary(final_mod_cb_glm)
-  tglmAbsRisk<-absoluteRisk(final_mod_cb_glm,time=times,newdata=test,type="CI")
-  tglmAbsRisk<-as.data.frame(tglmAbsRisk)
-  rownames(tglmAbsRisk)<-tglmAbsRisk$time
-  trueGlmProper<- t(tglmAbsRisk[-1,-1])
-
-  class(trueGlmProper)<-c("tunnel",class(trueGlmProper))
-
-  ##########################
-  ###casebase + splines
-  ##########################
-  mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
-                                data = train,
-                                time = "time",
-                                event="status",
-                                ratio = 100
-  )
-  glmAbsRisk<-absoluteRisk(mod_cb_glm,time=times,newdata=test,type="CI")
-  glmAbsRisk<-as.data.frame(glmAbsRisk)
-  rownames(glmAbsRisk)<-glmAbsRisk$time
-  glmProper<- t(glmAbsRisk[-1,-1])
-  class(glmProper)<-c("tunnel",class(glmProper))
-
-  ############################
-  ###DeepSurv
-  ############################  
-  coxnn<-deepsurv(Surv(time, status) ~ .,data = train, frac = 0.2, activation = "relu",
-                  num_nodes = c(layer1,layer2,layer3,layer4), dropout = drpt,
-                  early_stopping = TRUE, epochs = epo,
-                  batch_size = bsize,device = "cpu",best_weights=T,
-                  patience = patience,min_delta=min_delta,batch_norm = F,
-                  learning_rate = 0.001,lr_decay=10^-7)
-
-  coxnnsurv<- predict(coxnn,newdata = test,type="survival")
-  colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
-  cnnCleaned<-pyProcess(coxnnsurv,times=times)
-  colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
-  ############################
-  ###Deephit
-  ############################  
-  source_python('src/deephitter.py')
-  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,patience=patience,min_delta=min_delta,drpt=drpt)
-  hitnnSurv<-t(hitnnSurv)
-  colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
-  deephitCleaned<-pyProcess(hitnnSurv,times=times)
-  colnames(deephitCleaned)<-seq(0,1,length.out = ncol(deephitCleaned))
-  py_run_string("del fitDeephit")
-  rm(fitDeephit)
-
-  ############################
-  ###PMNNsplines
-  ############################  
-  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2+3),
-                            name = 'main_input'
-  )
-  covars_output<-covars_input%>% 
-    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=1,use_bias = T)
-
-  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
-                  feature_input = covars_input,
-                  feature_output = covars_output,
-                  originalData = mod_cb_glm$data,
-                  offset=mod_cb_glm$offset,
-                  timeVar = "time",
-                  eventVar= "status"
-  )
-
-  fit<-fitSmoothHazSpline(pmnn,
-                          epochs=epo,
-                          batch_size=bsize,
-                          verbose=0,
-                          monitor="val_loss",
-                          val_split=0.2,
-                          min_delta=min_delta,
-                          patience=patience)
-  annPreds<-aarSplines(fit,
-                       times=times,
-                       x_test=test
-  )
-  rownames(annPreds)<-annPreds[,1]
-  annProper<- t(annPreds[,-1])
-  class(annProper)<-c("tunnel",class(annProper))
-
-  ######################
-  ###coxph
-  ######################
-  cox<-coxph(Surv(time, status) ~ z1+z2+z3, data = train,x=T)
-
-  ######################
-  ###PMNN
-  ######################
-  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
-                            name = 'main_input')
-
-  covars_output<-covars_input%>% #layer_batch_normalization()%>%
-    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=1,use_bias = T)
-
-  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
-                  feature_input = covars_input,
-                  feature_output = covars_output,
-                  originalData = mod_cb_glm$data,
-                  offset=mod_cb_glm$offset,
-                  timeVar = "time",
-                  eventVar= "status"
-  )
-
-  fit<-fitSmoothHaz(pmnn,
-                    epochs=epo,
-                    batch_size=bsize,
-                    verbose=0,
-                    monitor="val_loss",
-                    val_split=0.2,
-                    min_delta=min_delta,
-                    patience=patience)
-  annPreds<-aar(fit,
-                times=times,
-                x_test=test
-  )
-
-  rownames(annPreds)<-annPreds[,1]
-  annProperpoly<- t(annPreds[,-1])
-  class(annProperpoly)<-c("tunnel",class(annProperpoly))
-
-  rownames(annPreds)<-annPreds[,1]
-  annProperpoly<- t(annPreds[,-1])
-
-  class(annProperpoly)<-c("tunnel",class(annProperpoly)) 
-  brierFinalResults <- Score(list("Cox_Lin" = cox,
-                                  'CB_Logi'=glmProper,
-                                  'DeepSurv'=cnnCleaned,
-                                  'DeepHit'=deephitCleaned,
-                                  'PMNN_Spline'=annProper,
-                                  'Optimal'=trueGlmProper,
-                                  'PMNN_Poly'=annProperpoly,
-                                  'dsm'=dsmSurvProper) ,#'dsm'=dsmCleaned,
-                                  data =fullTest, 
-                             formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)
-  BrierIPAList[[pos]]<-brierFinalResults$Brier$score
-  ########################
-  ###C-Index
-  ########################
-  tempTims<-   as.numeric(colnames(annProper))
-  tempTims<-head(tempTims, -15)
-  tempTims<-tail(tempTims, -10)
-  cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=9)
-  et_train<-as.matrix(train)#[,c(5,4)]
-  et_test<-as.matrix(fullTest[,c(5,4)])
-  colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
-                          'PMNN_Spline','Optimal','PMNN_Poly','dsm')
-  cScoreTemp[,1]<- tempTims
-
-  a<-survfit(cox, newdata=fullTest,times=times)
-  rownames(a$surv)<-a$time
-  coxlinSurv<-pyProcess(t(a$surv),times=times)
-  colnames(coxlinSurv)<-seq(0,1,length.out = ncol(coxlinSurv))
-
-  riskList<-list(coxlinSurv,
-                 glmProper,
-                 cnnCleaned,
-                 deephitCleaned,
-                 annProper,
-                 annProperpoly
-  )
-
-  cScore[[pos]]<-cIndexSummary(et_train=et_train
-                               , et_test=et_test,
-                               riskList=riskList,
-                               cScore=cScoreTemp
-  )
-
-  saveRDS(BrierIPAList,'100easyBrier.rds')
-  saveRDS(cScore,'100easycidx.rds')
-  pos<-pos+1
-}
-
-end_time <- Sys.time()
-
 ```
@@
 pos<-1
-for(iter in 1:iteration){
+
+while (pos <= iteration) {
 
@@
   fullTest<-data[-samp,,drop=F]
-  times<-seq(from=0,
+  times<-seq(from=min(test$time),
              to=max(test$time),
@@
 
+  #######################
+  ###Cox 
   ######################
-  ###C-Index
-  ######################
-  tempTims<-   as.numeric(colnames(annProper))
-  tempTims<-head(tempTims, -15)
-  tempTims<-tail(tempTims, -10)
-  cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=8)
-  et_train<-as.matrix(train)#[,c(5,4)]
-  et_test<-as.matrix(fullTest[,c(5,4)])
-  colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
-                          'PMNN_Spline','Optimal','PMNN_Poly')#,'dsm')
-  cScoreTemp[,1]<- tempTims
   a<-survfit(cox, newdata=fullTest,times=times)
@@
   colnames(coxlinSurv)<-seq(0,1,length.out = ncol(coxlinSurv))
+
+  ######################
+  ###C-Index
+  ######################
+
   riskList<-list(coxlinSurv,
@@
                  annProper,
+                 trueGlmProper,
                  annProperpoly
@@
 
-  cScore[[pos]]<-cIndexSummary(et_train=et_train
-                               , et_test=et_test,
+  tempTims<-   as.numeric(colnames(annProper))
+  tempTims<-head(tempTims, -1)
+  tempTims<-tail(tempTims, -1)
+  cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=length(riskList)+1)
+  et_train<-as.matrix(train)[,c(5,4)]
+  et_test<-as.matrix(fullTest[,c(5,4)])
+  colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
+                          'PMNN_Spline','Optimal','PMNN_Poly')#,'dsm')
+  cScoreTemp[,1]<- tempTims
+
+  cScore[[pos]]<-cIndexSummary(et_train=et_train,
+                               et_test=et_test,
                                riskList=riskList,

### 2022-05-25 analyses/plotterScript.Rmd
cleaned, licensed version of cbnn

---
title: "plottingScript"
output: html_document
---

```{r setup, include=T}
knitr::opts_chunk$set(echo = TRUE)
library(casebase) #core
library(keras) #core
source("src/functions.R") #core
library(data.table)#corecr
library(mltools)#corecr

library(survival)#compare
library(survminer)#compare
library(ggplot2)#compare
library(colorspace)#compare
library(DT)
library(splines)
library(reticulate)

library(simsurv)
library(flexsurv)
library(survivalmodels)
library(visreg)

#library(tidyverse)
library(riskRegression)
library(pseudo)

library(tensorflow)
library(ggrepel)

library(dplyr)
library(flextable)

library(cowplot)
library(matrixStats)
library(tableone)
library(gridExtra)
library(grid)
library(gtable)
#tf$config$threading$set_intra_op_parallelism_threads(1)
#tf$config$threading$set_inter_op_parallelism_threads(1)

#set.seed(2)
set.seed(1)
epo=2000
bsize=512
min_delta = 10^-7
patience = 10
iteration=50
e_sup=F
eval_firstSim<-F
numSamp<-2000

 pal <- function(col, border = "light gray", ...)  {
   n <- length(col)
   plot(0, 0, type="n", xlim = c(0, 1), ylim = c(0, 1),
   axes = FALSE, xlab = "", ylab = "", ...)
   rect(0:(n-1)/n, 0, 1:n/n, 1, col = col, border = border)
   }
statCalcSim<-function(metricSims){

toRemove <- which(apply(metricSims, 1, function(x){any(is.na(x))})==T)
if(length(toRemove)>0){
metricSims<-metricSims[-c(toRemove),]
}
  metricSummary<-data.frame(model=metricSims$model,times=metricSims$times)
  metricSummary<-cbind(metricSummary,sds=rowSds(as.matrix(metricSims[,-c(1,2)])),means=rowMeans(as.matrix(metricSims[,-c(1,2)])))

  ci<-apply(as.matrix(metricSims[,-c(1,2)]), 1, function(x){mean(x)+c(-1.96,1.96)*sd(x)/sqrt(length(x))})
  rownames(ci)<-c('cilower95','ciupper95')
  pi<-apply(as.matrix(metricSims[,-c(1,2)]), 1, function(x){quantile(x,c(.025, .975))})
  rownames(pi)<-c('pilower95','piupper95')

  metricSummary<-cbind(metricSummary,t(ci),t(pi))

  return(metricSummary)
}

plotPerformance<-function(metricSummary,xlab="Follow-up time",method='Metric',iterations='NA'){

  allModels<-data.frame(model=c("Null model","Cox_Lin",'CB_Logi','DeepSurv',
                                'DeepHit','PMNN_Spline','PMNN_Poly','Optimal',
                                "dsm"),
                        color=qualitative_hcl(9,palette='Dark 3'),
                        trueName=c("KM","Cox",'CBLR','DeepSurv',
                                'DeepHit','PMNN_Spline','CBNN','Optimal',
                                "DSM"))
  allModels$color[1]="#bfc1c2"
  allModels$color[2]="#FFED11"#"#7C4C74"
  allModels$color[3]= "#80471C" #"#F9A602"
  allModels$color[5]="#FF00AB"
  allModels$color[8]="#000000"
  allModels$color[9]="#8077D5"
 allModels<- allModels[order(allModels$trueName),]
    i <- sapply(metricSummary, is.factor)
    metricSummary[i] <- lapply(metricSummary[i], as.character)
    metricSummary<-metricSummary[complete.cases(metricSummary), ]
    mainTitle<-""
    metricSummary$color<-NA
    for(i in 1:nrow(metricSummary)){
      modelPos<-which(metricSummary$model[i] ==allModels$model)
      metricSummary$model[i]<-allModels$trueName[modelPos]
      metricSummary$color[i]<-allModels$color[modelPos]

    }

   colorsToUse<- sort(which(allModels$trueName %in% metricSummary$model))

    if(method=="C-Index"){
    method=expression("c-ipcw")
    }   
 metricSummary<- metricSummary[-which(metricSummary$model=="PMNN_Spline"),]
  #mainTitle<-paste(method,' : 95% P.I. over ', iterations, ' iterations',sep='')
  colnames(metricSummary)[1]<-"Models"
  a<-ggplot(data=metricSummary, aes(x=times,y=means,col=Models))+
    geom_line() +
    geom_ribbon(aes(ymin=cilower95,ymax=ciupper95,fill=Models),linetype=0,alpha=0.3) +
    ggtitle(mainTitle) + 
    xlab(xlab) +
    ylab(method)+
    theme_minimal()+
    scale_color_manual(values=allModels$color[colorsToUse])+
    scale_fill_manual(values=allModels$color[colorsToUse])
   # scale_color_manual(labels = metricSummary$model, values = metricSummary$color)

  return(a)
}

```

```{r plotsupport,eval=T}

#age, sex, race, number of comorbidities, presence of diabetes, presence of dementia, presence of cancer, mean arterial blood pressure, heart rate, respiration rate, temperature, white blood cell count, serum’s sodium, and serum’s creatinine
BrierIPAList<-readRDS('results/100supBrier.rds')
BrierIPAList[[1]]$times<-BrierIPAList[[1]]$times*(max(support$d.time, na.rm = T)-min(support$d.time, na.rm = T)) + min(support$d.time, na.rm = T)
ipaSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
brierSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
ibsSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)

ibsTable<-as.data.frame(matrix(data=NA,nrow=100,ncol=length(unique(BrierIPAList[[1]]$model))))
colnames(ibsTable)<-unique(BrierIPAList[[1]]$model)
count=1
for(sim in BrierIPAList){
  ipaSims<-cbind(ipaSims, sim$IPA)
  brierSims<-cbind(brierSims,sim$Brier)
  ibsSims<-cbind(ibsSims,sim$IBS)
  for(mod in colnames(ibsTable)){

    ibsTable[count,which(mod ==colnames(ibsTable))]<-sim$IBS[sim$model==mod & sim$time==max(sim$time)]

  }
  count<-count+1
}

ibsOverall<-data.frame(mean=apply(ibsTable,2,mean),sd = apply(ibsTable,2,sd))
ibsOverall

ipaSummary<-statCalcSim(ipaSims)
brierSummary<-statCalcSim(brierSims)
ibsSummary<-statCalcSim(ibsSims)
ipa<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
brier<-plotPerformance(brierSummary,method="BS",iterations=iteration )
ibs<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )
data<-casebase::support
covs<-colnames(data)
catVars<-c('death','sex','dzgroup','dzclass','race','ca')

data[,-which( covs%in%catVars )]<-as.data.frame(sapply(data[,-which( covs%in%catVars )], function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))

colnames(data)[5]<-'time'
colnames(data)[2]<-'status'
ptPlot<-plot(popTime(data,time="time",event="status"),percentile_number=0.5)#+  theme(axis.text=element_text(size=8))
#end_time-start_time

pt <- plot_grid(ptPlot,
          ipa+ theme(legend.position="none"),
          brier+ theme(legend.position="none"),
          ibs+ theme(legend.position="none"),
          labels = c('A', 'B','C','D'))
#https://wilkelab.org/cowplot/articles/shared_legends.html
legend <- get_legend(
  # create some space to the left of the legend
  ipa  +
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "bottom")+
   theme(legend.text=element_text(size=8))
)

pdf(file="figures/supportResult.pdf")
plot_grid(pt, legend,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))
dev.off()

myVars<-colnames(data)
#catVars<-c('x4','x5','x6','x7','status')

tab2 <- as.data.frame(print(CreateTableOne(vars = myVars, data = data, factorVars = catVars)))
saveRDS(tab2,"figures/t1Support.rds")

pdf("figures/t1Support.pdf", height = 50*nrow(tab2), width = 200*ncol(tab2))
grid.table(tab2)
dev.off()

rm(pt)
#min(support$d.time, na.rm = T)
#max(support$d.time, na.rm = T)

ipaSu<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
brierSu<-plotPerformance(brierSummary,method="BS",iterations=iteration )
ibsSu<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )

cidxList<-readRDS('results/100supcidx.rds')

temp<-melt(setDT(as.data.frame(cidxList[[1]])), id.vars = c("times"), variable.name = "model")
cidxSims=data.frame(model=temp$model,times=temp$times)

count=1
for(sim in cidxList){
  cidxSims<-cbind(cidxSims, melt(setDT(as.data.frame(sim)), id.vars = c("times"), variable.name = "model")$value)

  count<-count+1
}

cidxSummary<-statCalcSim(cidxSims)
cidxSu<-plotPerformance(cidxSummary,method="C-Index",iterations=iteration  )
cidxSu

```

```{r plotMetabric,fig.height=6,fig.width=6}

BrierIPAList<-readRDS('results/100MetaBrier.rds')
source_python("src/getMetabricData.py")
metabric<-getMetabric()

colnames(metabric)<-c('MK167','EGFR','PGR','ERBB2','Hormone treatment','radiotherapy','Chemotherapy','ER_positive','age at diagnosis','time','status')
BrierIPAList[[1]]$times<-BrierIPAList[[1]]$times*(max(metabric$time, na.rm = T)-min(metabric$time, na.rm = T)) + min(metabric$time, na.rm = T)
ipaSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
brierSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
ibsSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
ibsTable<-as.data.frame(matrix(data=NA,nrow=100,ncol=length(unique(BrierIPAList[[1]]$model))))
colnames(ibsTable)<-unique(BrierIPAList[[1]]$model)
count=1
for(sim in BrierIPAList){
  ipaSims<-cbind(ipaSims, sim$IPA)
  brierSims<-cbind(brierSims,sim$Brier)
  ibsSims<-cbind(ibsSims,sim$IBS)
  for(mod in colnames(ibsTable)){

    ibsTable[count,which(mod ==colnames(ibsTable))]<-sim$IBS[sim$model==mod & sim$time==max(sim$time)]

  }
  count<-count+1
}
ibsTable<-na.omit(ibsTable)
ibsOverall<-data.frame(mean=apply(ibsTable,2,mean),sd = apply(ibsTable,2,sd))
ibsOverall
ipaSummary<-statCalcSim(ipaSims)
brierSummary<-statCalcSim(brierSims)
ibsSummary<-statCalcSim(ibsSims)
ipa<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
brier<-plotPerformance(brierSummary,method="BS",iterations=iteration )
ibs<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )

colnames(metabric)[ncol(metabric)-1]<-'time'
colnames(metabric)[ncol(metabric)]<-'status'
data<-metabric
#data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
covs<-colnames(data)

ptPlot<-plot(popTime(data,time="time",event="status"),percentile_number=0.5)#+  theme(axis.text=element_text(size=8))
#end_time-start_time

pt <- plot_grid(ptPlot,
          ipa+ theme(legend.position="none"),
          brier+ theme(legend.position="none"),
          ibs+ theme(legend.position="none"),
          labels = c('A', 'B','C','D'))
#https://wilkelab.org/cowplot/articles/shared_legends.html
legend <- get_legend(
  # create some space to the left of the legend
  ipa  +
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "bottom")+
   theme(legend.text=element_text(size=8))
)

pdf(file="figures/metabricResult.pdf")
plot_grid(pt,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))#, legend
dev.off()

myVars<-c('MK167','EGFR','PGR','ERBB2','Hormone treatment','radiotherapy','Chemotherapy','ER_positive','age at diagnosis','time','status')
catVars<-c('Hormone treatment','radiotherapy','Chemotherapy','ER_positive','status')

tab2 <- as.data.frame(print(CreateTableOne(vars = myVars, data = data, factorVars = catVars)))
saveRDS(tab2,"figures/t1Metabric.rds")

pdf("figures/t1metabric.pdf", height = 50*nrow(tab2), width = 200*ncol(tab2))
grid.table(tab2)
dev.off()
rm(pt)

ipaM<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
brierM<-plotPerformance(brierSummary,method="BS",iterations=iteration )
ibsM<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )

ipa
ibs
brier

cidxList<-readRDS('results/100Metacidx.rds')

temp<-melt(setDT(as.data.frame(cidxList[[1]])), id.vars = c("times"), variable.name = "model")
cidxSims=data.frame(model=temp$model,times=temp$times)

count=1
for(sim in cidxList){
  cidxSims<-cbind(cidxSims, melt(setDT(as.data.frame(sim)), id.vars = c("times"), variable.name = "model")$value)

  count<-count+1
}

cidxSummary<-statCalcSim(cidxSims)
cidxM<-plotPerformance(cidxSummary,method='C-Index',iterations=iteration  )
cidxM
```

```{r plotsimple,eval=T}

BrierIPAList<-readRDS('results/100easyBrier.rds')

ipaSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
brierSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
ibsSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
ibsTable<-as.data.frame(matrix(data=NA,nrow=100,ncol=length(unique(BrierIPAList[[1]]$model))))
colnames(ibsTable)<-unique(BrierIPAList[[1]]$model)
count=1
for(sim in BrierIPAList){
  ipaSims<-cbind(ipaSims, sim$IPA)
  brierSims<-cbind(brierSims,sim$Brier)
  ibsSims<-cbind(ibsSims,sim$IBS)
  for(mod in colnames(ibsTable)){

    ibsTable[count,which(mod ==colnames(ibsTable))]<-sim$IBS[sim$model==mod & sim$time==max(sim$time)]

  }
  count<-count+1
}
ibsTable<-na.omit(ibsTable)
ibsOverall<-data.frame(mean=apply(ibsTable,2,mean),sd = apply(ibsTable,2,sd))
ibsOverall
ipaSummary<-statCalcSim(ipaSims)
brierSummary<-statCalcSim(brierSims)
ibsSummary<-statCalcSim(ibsSims)
ipa<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
brier<-plotPerformance(brierSummary,method="BS",iterations=iteration )
ibs<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )

  cov<-data.frame(id=1:numSamp,
                z1=c(rep(1,numSamp/2),rep(0,numSamp/2)),
                z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
                z3=c(rnorm(numSamp/2,-3,0.5),rnorm(numSamp/2,8,0.5)))

dat <- simsurv(dist = "exponential", lambdas = 1, betas = c(z1 = .1,z2=.1,z3=.1),
                  x = cov, tdefunction = "log", maxt = 1)

  colnames(dat)<-c("id","time","status")
  samp<-sample(seq(1,nrow(dat),by=1), floor(nrow(dat)*(0.1))) #random censoring
  dat$status[samp]<-0
  # Merge the simulated event times onto covariate data frame
  data <- merge(cov, dat)

ptPlot<-plot(popTime(data,time="time",event="status"),percentile_number=.01)#+  theme(axis.text=element_text(size=8))
#end_time-start_time

pt <- plot_grid(ptPlot,
          ipa,
          brier,#+ theme(legend.position="none"),
          ibs,#+ theme(legend.position="none"),
          labels = c('A', 'B','C','D'))
#https://wilkelab.org/cowplot/articles/shared_legends.html
legend <- get_legend(
  # create some space to the left of the legend
  ipa  +
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "bottom")+
   theme(legend.text=element_text(size=8))
)

pdf(file="figures/simpleResult.pdf")
plot_grid(pt, legend,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))
dev.off()

myVars<-colnames(data)[-1]
catVars<-c('z1','status')

tab2 <- as.data.frame(print(CreateTableOne(vars = myVars, data = data, factorVars = catVars,strata='z1')))
tab2<-tab2[,-c(3,4)]
tab2[2,1]<-"0 (0.0)"
tab2[2,2]<-"1000 (100.0)"
tab2[3,1]<-"0 (0.50)"
tab2[3,2]<-"10 (0.50)"
tab2[4,1]<-"8 (0.50)"
tab2[4,2]<-"-3 (0.50)"
tab2<-tab2[-c(5,6),]
tab2
saveRDS(tab2,"figures/covariatesTable.rds")

pdf("figures/t1Simple.pdf", height = 50*nrow(tab2), width = 200*ncol(tab2))
grid.table(tab2[,-ncol(tab2)])
dev.off()

rm(pt)

ipaS<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
brierS<-plotPerformance(brierSummary,method="BS",iterations=iteration )
ibsS<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )

cidxList<-readRDS('results/100easycidx.rds')

temp<-melt(setDT(as.data.frame(cidxList[[1]])), id.vars = c("times"), variable.name = "model")
cidxSims=data.frame(model=temp$model,times=temp$times)

count=1
for(sim in cidxList){
  cidxSims<-cbind(cidxSims, melt(setDT(as.data.frame(sim)), id.vars = c("times"), variable.name = "model")$value)

  count<-count+1
}

cidxSummary<-statCalcSim(cidxSims)
cidxS<-plotPerformance(cidxSummary,method="C-Index",iterations=iteration  )

```

```{r complexPrep}

# Define a function returning the log cum hazard at time t
logcumhaz <- function(t, x, betas, knots) {

  # Obtain the basis terms for the spline-based log
  # cumulative hazard (evaluated at time t)
  basis <- flexsurv::basis(knots, log(t))

  # Evaluate the log cumulative hazard under the
  # Royston and Parmar specification
  res <-betas[["gamma0"]] * basis[[1]] + 
    betas[["gamma1"]] * basis[[2]] +
    betas[["gamma2"]] * basis[[3]] +
    betas[["gamma3"]] * basis[[4]] +
    betas[["gamma4"]] * basis[[5]] +
    betas[["z1"]] * x[["z1"]]+
    betas[["z2"]] * x[["z2"]]+
    betas[["z3"]] * x[["z3"]]+
    betas[["term1"]] * x[["z1"]]*x[["z2"]]*t+
    betas[["term2"]] * x[["z1"]]*x[["z3"]]+
    betas[["term3"]] * x[["z2"]]*(x[["z3"]])
  #
  # Return the log cumulative hazard at time t
  res
}

pyProcess<-function(predictions,times=times){
  # test<-t(data.frame(a=seq(0,100,length.out = 200),b=seq(0,100,length.out = 200),c=seq(0,100,length.out = 200)))
  #predictions[is.na(predictions)] <- 0
  interpolated <- apply(predictions, 1, function(x) approx(as.numeric(colnames(predictions)), y=x, method = "linear", xout=times)$y)

  class(interpolated)<-c("tunnel",class(interpolated))
  interpolated<-1-interpolated

  interpolated[is.na(interpolated)] <- 0
  return(t(interpolated))
}

url<-'app.txt'
ds <- read.table(url, sep = '',header = TRUE)

ds$years.at.risk=pmin(ds$agein80,ds$onset)

#cov <- data.frame(id = 1:2000, hormon = rbinom(2000, 1, 0.5))

#ds$years.at.risk<-ds$years.at.risk/max(ds$years.at.risk)
#true_mod <- flexsurv::flexsurvspline(Surv(years.at.risk, app) ~ 1, data = ds, k = 3) #appendicitis from hanley

  cov<-data.frame(id=1:numSamp,
                z1=c(rep(1,numSamp/2),rep(0,numSamp/2)),
                z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
                z3=c(rnorm(numSamp/2,-3,0.5),rnorm(numSamp/2,8,0.5)))
  brcancer$rectime<-brcancer$rectime/max(brcancer$rectime)
  true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1, data = brcancer, k = 3) #from flexsurv

  # Simulate the event times
  coefficients<-true_mod$coefficients
  coefficients<-c(coefficients,z1=1,z2=1,z3=1,term1=10,term2=2,term3=2)
  #names(coefficients)<-c("shape",names(coefficients[-1]))

```

```{r plotComplex}

BrierIPAList<-readRDS('results/100hardBrier.rds')

ipaSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
brierSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
ibsSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
ibsTable<-as.data.frame(matrix(data=NA,nrow=100,ncol=length(unique(BrierIPAList[[1]]$model))))
colnames(ibsTable)<-unique(BrierIPAList[[1]]$model)
count=1
for(sim in BrierIPAList){
  ipaSims<-cbind(ipaSims, sim$IPA)
  brierSims<-cbind(brierSims,sim$Brier)
  ibsSims<-cbind(ibsSims,sim$IBS)
  for(mod in colnames(ibsTable)){

    ibsTable[count,which(mod ==colnames(ibsTable))]<-sim$IBS[sim$model==mod & sim$time==max(sim$time)]

  }
  count<-count+1
}
ibsTable<-na.omit(ibsTable)
ibsOverall<-data.frame(mean=apply(ibsTable,2,mean),sd = apply(ibsTable,2,sd))
ibsOverall
ipaSummary<-statCalcSim(ipaSims)
brierSummary<-statCalcSim(brierSims)
ibsSummary<-statCalcSim(ibsSims)
ipa<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
brier<-plotPerformance(brierSummary,method="BS",iterations=iteration )
ibs<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )

#end_time-start_time

  cov<-data.frame(id=1:numSamp,
                z1=c(rep(1,numSamp/2),rep(0,numSamp/2)),
                z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
                z3=c(rnorm(numSamp/2,-3,0.5),rnorm(numSamp/2,8,0.5)))

  dat <- simsurv(betas = coefficients, # "true" parameter values
                 x = cov,                   # covariate data for 200 individuals
                 knots = true_mod$knots,    # knot locations for splines
                 logcumhazard = logcumhaz,  # definition of log cum hazard
                 maxt = 1,               # no right-censoring
                 interval = c(0,10000000)) # interval for root finding
  colnames(dat)<-c("id","time","status")
  samp<-sample(seq(1,nrow(dat),by=1), floor(nrow(dat)*(0.1))) #random censoring
  dat$status[samp]<-0
  # Merge the simulated event times onto covariate data frame
  data <- merge(cov, dat)

ptPlot<-plot(popTime(data,time="time",event="status"),percentile_number=.01)#+  theme(axis.text=element_text(size=8))
#end_time-start_time

pt <- plot_grid(ptPlot,
          ipa+ theme(legend.position="none"),
          brier+ theme(legend.position="none"),
          ibs+ theme(legend.position="none"),
          labels = c('A', 'B','C','D'))
#https://wilkelab.org/cowplot/articles/shared_legends.html
legend <- get_legend(
  # create some space to the left of the legend
  ipa  +
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "bottom")+
   theme(legend.text=element_text(size=8))
)

pdf(file="figures/complexResult.pdf")
plot_grid(pt, legend,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))
dev.off()

myVars<-colnames(data)[-1]
catVars<-c('z1','status')

tab2 <- as.data.frame(print(CreateTableOne(vars = myVars, data = data, factorVars = catVars,strata='z1')))
saveRDS(tab2,"figures/t1Complex.rds")

pdf("figures/t1Complex.pdf", height = 50*nrow(tab2), width = 200*ncol(tab2))
grid.table(tab2[,-ncol(tab2)])
dev.off()
rm(pt)

ipaC<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
brierC<-plotPerformance(brierSummary,method="BS",iterations=iteration )
ibsC<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )

cidxList<-readRDS('results/100hardcidx.rds')

temp<-melt(setDT(as.data.frame(cidxList[[1]])), id.vars = c("times"), variable.name = "model")
cidxSims=data.frame(model=temp$model,times=temp$times)

count=1
for(sim in cidxList){
  cidxSims<-cbind(cidxSims, melt(setDT(as.data.frame(sim)), id.vars = c("times"), variable.name = "model")$value)

  count<-count+1
}

cidxSummary<-statCalcSim(cidxSims)
cidxC<-plotPerformance(cidxSummary,method="C-Index",iterations=iteration  )
cidxC

```

```{r megaplot,eval=T,fig.height=12,fig.width=12}

pt <- plot_grid(
          ipaS+ theme(legend.position="none")+ ggtitle("Simple simulation")+ labs(x = "")+coord_cartesian(ylim=c(-0.5,0.1)),
          ipaC+ theme(legend.position="none")+ ggtitle("Complex simulation")+ labs(x = "",y="")+coord_cartesian(ylim=c(-.6,1)),
          ipaSu+ theme(legend.position="none")+ ggtitle("SUPPORT")+ labs(x = "",y="")+coord_cartesian(ylim=c(-0.20,.1)),
          ipaM+ theme(legend.position="none")+ ggtitle("METABRIC")+ labs(x = "",y="")+coord_cartesian(ylim=c(-0.8,.25)),
          #brierS+ theme(legend.position="none")+ labs(x = ""),
          #brierC+ theme(legend.position="none")+ labs(x = "",y=""),
          #brierSu+ theme(legend.position="none")+ labs(x = "",y=""),
          #brierM+ theme(legend.position="none")+ labs(x = "",y=""),
          #ibsS+ theme(legend.position="none")+ labs(x = ""),
          #ibsC+ theme(legend.position="none")+ labs(x = "",y=""),
          #ibsSu+ theme(legend.position="none")+ labs(x = "",y=""),
          #ibsM+ theme(legend.position="none")+ labs(x = "",y=""), 
          cidxS+ theme(legend.position="none")+coord_cartesian(ylim=c(.45,.6)),
          cidxC+ theme(legend.position="none")+ labs(y="")+coord_cartesian(ylim=c(0.3,.9)),
          cidxSu+ theme(legend.position="none")+ labs(y="")+coord_cartesian(ylim=c(0.48,.6)),
          cidxM+ theme(legend.position="none")+ labs(y="")+coord_cartesian(ylim=c(0.3,.65)),
          #labels = c('A', 'E','I','M',"B",'F','J','N','C','G','K','O','D','H','L','P'))
          labels = c('A',"C","E","G",'B','D','F','H'),
          ncol=4)
#https://wilkelab.org/cowplot/articles/shared_legends.html
legend <- get_legend(
  # create some space to the left of the legend
  ipaS  +
    guides(color = guide_legend(nrow = 2)) +
    theme(legend.position = "bottom")+
   theme(legend.text=element_text(size=8))
)

g<-plot_grid(pt, legend,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))

pdf(file="figures/megaPlot.pdf",width=12,height=6)
g
dev.off()

saveRDS(g,"figures/megaPlot.rds")

```

```{r megaplotSupplemental,fig.height=12,fig.width=12}
pt <- plot_grid(
          #ipaS+ theme(legend.position="none")+ ggtitle("Simple simulation")+ labs(x = ""),
          #ipaC+ theme(legend.position="none")+ ggtitle("Complex simulation")+ labs(x = "",y=""),
          #ipaSu+ theme(legend.position="none")+ ggtitle("SUPPORT")+ labs(x = "",y=""),
          #ipaM+ theme(legend.position="none")+ ggtitle("METABRIC")+ labs(x = "",y=""),
          brierS+ theme(legend.position="none")+ labs(x = "")+coord_cartesian(ylim=c(0,.4))+ ggtitle("Simple simulation"),
          brierC+ theme(legend.position="none")+ labs(x = "",y="")+coord_cartesian(ylim=c(0,.3))+ ggtitle("Complex simulation"),
          brierSu+ theme(legend.position="none")+ labs(x = "",y="")+coord_cartesian(ylim=c(0.1,.3))+ ggtitle("SUPPORT"),
          brierM+ theme(legend.position="none")+ labs(x = "",y="")+coord_cartesian(ylim=c(0,.4))+ ggtitle("METABRIC"),
          ibsS+ theme(legend.position="none")+ labs(x = "")+coord_cartesian(ylim=c(0,.3)),
          ibsC+ theme(legend.position="none")+ labs(x = "",y="")+coord_cartesian(ylim=c(0,.3)),
          ibsSu+ theme(legend.position="none")+ labs(x = "",y="")+coord_cartesian(ylim=c(0,.3)),
          ibsM+ theme(legend.position="none")+ labs(x = "",y="")+coord_cartesian(ylim=c(0,.3)), 
          #cidxS+ theme(legend.position="none"),
          #cidxC+ theme(legend.position="none")+ labs(y=""),
          #cidxSu+ theme(legend.position="none")+ labs(y=""),
          #cidxM+ theme(legend.position="none")+ labs(y=""),
          #labels = c('A', 'E','I','M',"B",'F','J','N','C','G','K','O','D','H','L','P'))
          labels = c('A',"C","E","G",'B','D','F','H'),
          ncol=4)
#https://wilkelab.org/cowplot/articles/shared_legends.html
legend <- get_legend(
  # create some space to the left of the legend
  ipaS  +
    guides(color = guide_legend(nrow = 2)) +
    theme(legend.position = "bottom")+
   theme(legend.text=element_text(size=8))
)

g<-plot_grid(pt, legend,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))

pdf(file="figures/megaPlotSupplement.pdf",width=12,height=6)
g
dev.off()

saveRDS(g,"figures/megaPlotSupplement.rds")

```

```{r flextableTablePrep,fig.height=46,fig.width=26}

edit_cell <- function(table, row, col, name="core-fg", ...){
  l <- table$layout
  ids <- which(l$t %in% row & l$l %in% col & l$name==name)
  for (id in ids){
  newgrob <- editGrob(table$grobs[id][[1]], ...)
  table$grobs[id][[1]] <- newgrob
  }
  table
}

#SUMMARY DATA ORDER, IPA BS IBS CIDX
makeTable<-function(summaryData,study="study"){
   allModels<-data.frame(model=c("Null model","Cox_Lin",'CB_Logi','DeepSurv',
                                'DeepHit','PMNN_Spline','PMNN_Poly','Optimal',
                                "dsm"),
                        color=rainbow_hcl(9, start = 30, end = 300),
                        trueName=c("KM","Cox",'CBLR','DeepSurv',
                                'DeepHit','PMNN_Spline','CBNN','Optimal',
                                "DSM"))

  wantedScores<-c(0.25,0.5,0.75,1)

  tableResults<-as.data.frame(matrix(NA,nrow=nrow(allModels)+2,ncol=16))
  tableResults[1,]<-c("IPA","IPA","IPA","IPA",
                      "BS","BS","BS","BS",
                      "IBS","IBS","IBS","IBS",
                      "C-Index","C-Index","C-Index","C-Index"
                      )
  tableResults[2,]<-c(0.25,0.5,0.75,1,0.25,0.5,0.75,1,0.25,0.5,0.75,1,0.25,0.5,0.75,1)*100
  indexer<-1
  rownames(tableResults)<-c(study,"% of time",allModels$trueName)
  for (i in 1:length(summaryData)){ #IPA BRIER IBS CIDX is indexed by i

    tempPos<-which(rownames(tableResults) %in% unique(summaryData[[i]]$Models)  )
    tempModels<-rownames(tableResults)[ tempPos]
    for(j in 1:length(tempModels)){ #indexes model
      offset<-(i-1)*4
      indexer=1+offset
      for(k in 1:length(wantedScores)){ #indexes .25 .5 .75 1.0
        tempTime<-unique(summaryData[[i]]$times)[(floor(length(unique(summaryData[[i]]$times))*wantedScores[k]))]
        tempResRow<-which(summaryData[[i]]$Models==rownames(tableResults)[ tempPos[j]] & summaryData[[i]]$times==tempTime)
        if(length(round(summaryData[[i]]$means[tempResRow],2))<1){
          tableResults[tempPos[j],indexer]<-"-0.23 (-0.25,-0.23)" #note that this is grabbed directly from the last estimate of Cox, as its the only time we don't get a value. So we extrapolate from the last available survival time.
        }
        else{tableResults[tempPos[j],indexer]<-paste(round(summaryData[[i]]$means[tempResRow],2),
              " (",round(summaryData[[i]]$cilower95[tempResRow],2),
              ",",round(summaryData[[i]]$ciupper95[tempResRow],2),")",sep="")
        }
        indexer<-indexer+1

      }
    }

  }

  return(tableResults)
}

cleantables<-function(tab){
  toRemove<-which(rowSums(is.na(tab))==ncol(tab))
      tab[is.na(tab)]<-"-"
  if(length(toRemove)>0){

    return(tab[-toRemove,])
  }
  return(tab)
}

quickEdit<-function(grob,rows,columns){
  return(edit_cell(grob, rows, columns, "core-fg", 
               gp=gpar(fontsize=12, fontface="bold")))
}

addBorders<-function(grob){
grob$layout[1:2 , c("l", "r")] <- list(c(5, 9),c(2,6))
grob<- gtable_add_grob(grob,
        grobs = rectGrob(gp = gpar(fill = NA, lwd = 2)),
        t = 2, b = nrow(grob), l = 1, r = ncol(grob))
grob <- gtable_add_grob(grob,
        grobs = rectGrob(gp = gpar(fill = NA, lwd = 2)),
        t = 1, l = 1, r = ncol(grob))
grob <- gtable_add_grob(grob,
        grobs = rectGrob(gp = gpar(fill = NA, lwd = 2)),
        t = 1, l = 2,r=5,b= nrow(grob))
return(grob)
}

flexMaker<-function(tab,boldPicks=NA,title){
  boldPicks$col=boldPicks$col+1 #makes finding the correct ones easier.
bord<-fp_border_default(color="black",style="solid")
bord2<-fp_border_default(color="black",style="solid",width = 2)
tab<-cbind(rownames(tab),tab)
ft<-delete_part(flextable(tab[-c(1,2),],cwidth=5,cheight=1), part = "header")

ft <- add_header_row(x = ft, values = c(title,"IPA", "c-ipcw"),
  colwidths = c(1,4,4))
#ft<-bold(ft, bold = TRUE, part = "header")

#ft$header$styles$text$bold$data
ft <- add_header_row(x = ft,top=F, values = c("Method", 25,50,75,100,25,50,75,100),
  colwidths = c(1,1,1,1,1,1,1,1,1))
ft$header$styles$text$bold$data[1,1]<-T
#ft <- add_header_row(
#  x = ft, values = c(title),
#  colwidths = c(9))
ft <- align(ft, i = c(1,2), part = "header", align = "center")
ft<-vline(ft, j = c(1,5), border = bord2, part = "body")
ft<-hline(ft,i=c(1,2),border=bord2,part="header")
#ft<-hline(ft,i=c(1),border=bord,part="body")
ft<-fontsize(ft, i = NULL, j = NULL, size =7.5, part = "all")
ft<-fontsize(ft, i = NULL, j = NULL, size =10, part = "header")
ft <- align(ft, i = NULL, part = "body", align = "center")
ft$body$styles$text$font.size$data[,1]<-10

if(is.data.frame(boldPicks)){
  for(i in 1:nrow(boldPicks)){
    ft$body$styles$text$bold$data[boldPicks$row[i],boldPicks$col[i]]<-T
  }
}
return(rasterGrob(as_raster(autofit(ft))))
}

```

```{r flexTableMain,fig.height=46,fig.width=26}

sTab<-makeTable(list(ipaS$data,brierS$data,ibsS$data,cidxS$data),study="Simple")[-3,-c(5:12)]
#[-3,-c(1,2,3,4,13,14,15,16)] for the supplemental
cTab<-makeTable(list(ipaC$data,brierC$data,ibsC$data,cidxC$data),study="Complex")[-3,-c(5:12)]
suTab<-makeTable(list(ipaSu$data,brierSu$data,ibsSu$data,cidxSu$data),study="SUPPORT")[-3,-c(5:12)]
mTab<-makeTable(list(ipaM$data,brierM$data,ibsM$data,cidxM$data),study="METABRIC")[-3,-c(5:12)]
sTab<-cleantables(sTab)
cTab<-cleantables(cTab)
suTab<-cleantables(suTab)
mTab<-cleantables(mTab)
tempNames<-rownames(mTab)
mTab<-as.data.frame(lapply(mTab, function(x){gsub("0 ","0.00 ", x,fixed=T)}))
mTab<-as.data.frame(lapply(mTab, function(x){gsub('(0,',"(0.00,", x,fixed=T)}))
mTab<-as.data.frame(lapply(mTab, function(x) gsub(",0)",",0.00)", x,fixed=T)))
rownames(mTab)<-tempNames
sPicks<-data.frame(row=c(1,2,6,1,2,6,1,2,6,1,2,6,1,2,6,1,2,6,1,2,6,1,2,6),
                   col=c(1,2,3,4,5,6,7,8,1,2,3,4,5,6,7,8,1,2,3,4,5,6,7,8))
cPicks<-data.frame(row=c(6,6,6,6,5,5,5,2),
                   col=c(1,2,3,4,5,6,7,8))
suPicks<-data.frame(row=c(1,2,1,2,5,1,2,3,1,2,5,1,2,5,1,2,5,1,2,5),
                    col=c(1,1,2,2,2,3,3,4,5,5,5,6,6,6,7,7,7,8,8,8))
mPicks<-data.frame(row=c(2,4,5,1,2,1,2,4,5,4,4,5),
                   col=c(1,1,1,2,2,3,3,4,5,6,7,8))

sFlex<-flexMaker(tab=sTab,boldPicks = sPicks,title="A: Simple")
suFlex<-flexMaker(tab=suTab,boldPicks = suPicks,title="C: SUPPORT")
cFlex<-flexMaker(tab=cTab,boldPicks = cPicks,title="B: Complex")
mFlex<-flexMaker(tab=mTab,boldPicks = mPicks,title="D: METABRIC")

pdf(file="figures/megaTable.pdf",width=18,height=19.5)
cowplot::plot_grid(sFlex,cFlex, suFlex,mFlex, nrow = 4, ncol = 1)
dev.off()

```

```{r getMeans,fig.height=8,fig.width=16}

meanAndSort<-function(scores){
  meanScores<-aggregate(means~Models, data=scores, mean)
  return(tableGrob(meanScores[order(meanScores$means,decreasing=T),]))
}

simpleIPAMean<-meanAndSort(ipaS$data)
simplecidxMean<-meanAndSort(cidxS$data)

supportIPAMean<-meanAndSort(ipaSu$data)
supportcidxMean<-meanAndSort(cidxSu$data)

complexIPAMean<-meanAndSort(ipaC$data)
complexcidxMean<-meanAndSort(cidxC$data)

metaIPAMean<-meanAndSort(ipaM$data)
metacidxMean<-meanAndSort(cidxM$data)

pdf(file="figures/averageScores.pdf",width=16,height=8)
grid.arrange(simpleIPAMean,complexIPAMean,supportIPAMean,metaIPAMean,simplecidxMean,complexcidxMean,supportcidxMean,metacidxMean,nrow=2,ncol=4,top = "simple,complex,support,metabric")
dev.off()

```

### 2022-05-25 analyses/proofOfConceptWeights.Rmd
cleaned, licensed version of cbnn

--- 2022-05-25 analyses/codeReviewRealData.Rmd
@@
 iteration=5
-
-eval_support<-T
-source_python('src/getSUP.py')
-
 ########################
@@
 
-```{r modelsSUPPORT,eval=eval_support}
-
-start_time <- Sys.time()
-BrierIPAList<-list()
-cScore<-list()
-
-data<- as.data.frame(do.call(cbind, getSupportPycox()))
-colnames(data)[ncol(data)-1]<-'time'
-colnames(data)[ncol(data)]<-'status'
-data<-as.data.frame(model.matrix(~.,data)[,-1])
-data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
-
-covs<-colnames(data)
-
-#split training/testing
-samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.80)) 
-trainOG<-data[samp,]
-test<-data[-samp,-c(ncol(data))]
-fullTest<-data[-samp,,drop=F]
-
-pos=1
-for(iter in 1:iteration){
-
-  samp<-sample(seq(1,nrow(trainOG),by=1) , nrow(trainOG),replace = T) 
-  train<-trainOG[samp,]
-
-  times<-seq(from=0,
-             to=max(test$time),
-             length.out = 200
-  )
-  times<-times[-c(1,length(times))]
-
-  ######################
-  ###casebase + splines
-  ######################
-  mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
-                                data = train,
-                                time = "time",
-                                event="status",
-                                ratio = 100
-  )
-
-  glmAbsRisk<-as.data.frame(absoluteRisk(mod_cb_glm,time=times,newdata=test,type="CI"))
-  rownames(glmAbsRisk)<-glmAbsRisk$time
-  #prepare to run through riskRegression
-  glmProper<- t(glmAbsRisk[-1,-1])
-  class(glmProper)<-c("tunnel",class(glmProper))
-
-  ######################
-  ###DeepSurv
-  ######################
-  coxnn<-deepsurv(Surv(time, status) ~ .,data = train, frac = 0.2, activation = "relu",
-                  num_nodes = c(layer1,layer2,layer3,layer4,), dropout = drpt,
-                  early_stopping = TRUE, epochs = epo,
-                  batch_size = bsize,device = "cpu",best_weights=T,
-                  patience = patience,min_delta=min_delta,batch_norm = F,
-                  learning_rate = 0.001,lr_decay=10^-7)
-  coxnnsurv<- predict(coxnn,newdata = test,type="survival")
-  colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
-  cnnCleaned<-pyProcess(coxnnsurv,times=times)
-  colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
-
-  ######################
-  ###Deephit
-  ######################
-  source_python('deephitter.py')
-  hitnnSurv=fitDeephit(train,fullTest,bsize,epo)
-  hitnnSurv<-t(hitnnSurv)
-  colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
-  deephitCleaned<-pyProcess(hitnnSurv,times=times)
-  colnames(deephitCleaned)<-seq(0,1,length.out = ncol(deephitCleaned))
-  py_run_string("del fitDeephit")
-  rm(fitDeephit)
-
-  ######################
-  ###PMNN splines
-  ######################
-  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2+3),
-                            name = 'main_input'
-  )
-  covars_output<-covars_input%>% #layer_batch_normalization()%>%
-    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=1,use_bias = T)
-
-  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
-                  feature_input = covars_input,
-                  feature_output = covars_output,
-                  originalData = mod_cb_glm$data,
-                  offset=mod_cb_glm$offset,
-                  timeVar = "time",
-                  eventVar= "status"
-  )
-
-  fit<-fitSmoothHazSpline(pmnn,epochs=epo,batch_size = bsize,verbose=0)
-  annPreds<-aarSplines(fit,
-                       times=times,
-                       x_test=test
-  )
-  rownames(annPreds)<-annPreds[,1]
-  annProper<- t(annPreds[,-1])
-  class(annProper)<-c("tunnel",class(annProper))
-
-  ######################
-  ###coxph
-  ######################
-  cox<-coxph(Surv(time, status) ~ ., data = train,x=T)
-
-  ######################
-  ###PMNN
-  ######################
-  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
-                            name = 'main_input')
-
-  covars_output<-covars_input%>% #layer_batch_normalization()%>%
-    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=1,use_bias = T)
-
-  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
-                  feature_input = covars_input,
-                  feature_output = covars_output,
-                  originalData = mod_cb_glm$data,
-                  offset=mod_cb_glm$offset,
-                  timeVar = "time",
-                  eventVar= "status"
-  )
-
-  fit<-fitSmoothHaz(pmnn,epochs=epo,batch_size = bsize,verbose=0)
-  annPreds<-aar(fit,
-                times=times,
-                x_test=test
-  )
-
-  rownames(annPreds)<-annPreds[,1]
-  annProperpoly<- t(annPreds[,-1])
-  class(annProperpoly)<-c("tunnel",class(annProperpoly))
-
-  ######################
-  ###IPA calculation
-  ######################
-  brierFinalResults <- Score(list("Cox_Lin" = cox,'CB_Logi'=glmProper,
-                                  'DeepSurv'=cnnCleaned,'DeepHit'=deephitCleaned,
-                                  'PMNN_Spline'=annProper,'PMNN_Poly'=annProperpoly),
-                             data =fullTest, 
-                             formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)
-  BrierIPAList[[pos]]<-brierFinalResults$Brier$score
-
-  ######################
-  ###C-Index
-  ######################
-  tempTims<-as.numeric(colnames(annProper))
-  tempTims<-head(tempTims, -15)
-  tempTims<-tail(tempTims, -10)
-  cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=7)
-
-  et_train<-as.matrix(train)
-  et_test<-as.matrix(fullTest[,c(11,10)])
-  colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
-                          'PMNN_Spline','PMNN_Poly')
-  cScoreTemp[,1]<- tempTims
-
-  ### Cox is not calculated, so abrisk is calculated here.
-  a<-survfit(cox, newdata=fullTest,times=times)
-  rownames(a$surv)<-a$time
-  coxlinSurv<-pyProcess(t(a$surv),times=times)
-  colnames(coxlinSurv)<-seq(0,1,length.out = ncol(coxlinSurv))
-  riskList<-list(coxlinSurv,
-                 glmProper,
-                 cnnCleaned,
-                 deephitCleaned,
-                 annProper,
-                 annProperpoly
-  )
-
-  cScore[[pos]]<-cIndexSummary(et_train=et_train
-                               , et_test=et_test,
-                               riskList=riskList,
-                               cScore=cScoreTemp
-  )
-  pos=pos+1
-  saveRDS(BrierIPAList,'100supBrier.rds')
-  saveRDS(cScore,'100supcidx.rds')
-
-}
-
-ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=Brier,col=model))+
-  geom_line()
-ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
-  geom_line()
-
-end_time <- Sys.time()
-
-end_time-start_time
-
-```
-
-```{r modelsMETABRIC,eval=eval_metabric}
-
-start_time <- Sys.time()
-
-#pycox<-reticulate::import("pycox")
-#numpy<-reticulate::import("numpy")
-#a<-pycox$datasets$metabric$read_df
-#b<-numpy$asarray(a)
-
-############################
-###Data clean and prep
-############################
-metabric<-getMetabric()
-colnames(metabric)[ncol(metabric)-1]<-'time'
-colnames(metabric)[ncol(metabric)]<-'status'
-data<-as.data.frame(model.matrix(~.,metabric)[,-1])
-data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
-covs<-colnames(data)
-
-samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.80)) 
-trainOG<-data[samp,]
-test<-data[-samp,-c(ncol(data))]
-
-fullTest<-data[-samp,,drop=F]
-
-############################
-###Prep Loop
-############################
-BrierIPAList<-list()
-cScore<-list()
-pos=1
-
-for(iter in 1:iteration){
-
-  ############################
-  ###Bootstrap samples
-  ############################  
-  samp<-sample(seq(1,nrow(trainOG),by=1) , nrow(trainOG),replace = T) 
-  train<-trainOG[samp,]
-
-  times<-seq(from=0,
-             to=max(test$time),
-             length.out = 200
-  )
-  times<-times[-c(1,length(times))]
-  times<-unique(times)
-
-  ############################
-  ###DSM
-  ############################  
-  source_python('dsmScript.py')
-  dsmSurv<-dsmfitter(train=train,fullTest=fullTest,bsize=bsize,epochs=epo,tPos=ncol(train)-2,ePos=ncol(train)-1)
-  #sometimes, dsm fails to converge so we keep trying until it succeeds
-  if((sum(is.na(dsmSurv))==ncol(dsmSurv)*nrow(dsmSurv))){
-    dsmSurv<-dsmfitter(train=train,fullTest=fullTest,bsize=bsize,epochs=epo,tPos=ncol(train)-2,ePos=ncol(train)-1)
-    py_run_string("del dsmfitter")
-    rm(dsmfitter)
-    iteration<-iteration+1
-    next
-  }
-  colnames(dsmSurv) <-seq(0.01,1,by=0.01)
-  dsmSurvProper<-pyProcess(dsmSurv,times=times)
-  colnames(dsmSurvProper) <-seq(0,1,length.out = ncol(dsmSurvProper))
-  py_run_string("del dsmfitter")
-  rm(dsmfitter)
-
-  ############################
-  ###casebase + splines
-  ############################  
-  mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
-                                data = train,
-                                time = "time",
-                                event="status",
-                                ratio = 100
-  )
-
-  glmAbsRisk<-absoluteRisk(mod_cb_glm,time=times,newdata=test,type="CI")
-  glmAbsRisk<-as.data.frame(glmAbsRisk)
-  rownames(glmAbsRisk)<-glmAbsRisk$time
-  glmProper<- t(glmAbsRisk[-1,-1])
-  class(glmProper)<-c("tunnel",class(glmProper))
-
-  ############################
-  ###DeepSurv
-  ############################  
-  coxnn<-deepsurv(Surv(time, status) ~ .,data = train, frac = 0.2, activation = "relu",
-                  num_nodes = c(layer1,layer2,layer3,layer4), dropout = drpt,
-                  early_stopping = TRUE, epochs = epo,
-                  batch_size = bsize,device = "cpu",best_weights=T,
-                  patience = patience,min_delta=min_delta,batch_norm = F,
-                  learning_rate = 0.001,lr_decay=10^-7)
-
-  coxnnsurv<- predict(coxnn,newdata = test,type="survival")
-  colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
-  cnnCleaned<-pyProcess(coxnnsurv,times=times)
-  colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
-  ############################
-  ###Deephit
-  ############################  
-  source_python('deephitter.py')
-  hitnnSurv=fitDeephit(train,fullTest,bsize,10)
-  hitnnSurv<-t(hitnnSurv)
-  colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
-  deephitCleaned<-pyProcess(hitnnSurv,times=times)
-  colnames(deephitCleaned)<-seq(0,1,length.out = ncol(deephitCleaned))
-  py_run_string("del fitDeephit")
-  rm(fitDeephit)
-
-  ############################
-  ###PMNNsplines
-  ############################  
-  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2+3),
-                            name = 'main_input'
-  )
-  covars_output<-covars_input%>% 
-    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=1,use_bias = T)
-
-  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
-                  feature_input = covars_input,
-                  feature_output = covars_output,
-                  originalData = mod_cb_glm$data,
-                  offset=mod_cb_glm$offset,
-                  timeVar = "time",
-                  eventVar= "status"
-  )
-
-  fit<-fitSmoothHazSpline(pmnn,epochs=epo,batch_size = bsize,verbose=0)
-  annPreds<-aarSplines(fit,
-                       times=times,
-                       x_test=test
-  )
-
-  rownames(annPreds)<-annPreds[,1]
-  annProper<- t(annPreds[,-1])
-  class(annProper)<-c("tunnel",class(annProper))
-
-  ############################
-  ###coxph
-  ############################  
-  cox<-coxph(Surv(time, status) ~ ., data = train,x=T)
-
-  ############################
-  ###PMNN
-  ############################  
-  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
-                            name = 'main_input')
-
-  covars_output<-covars_input%>% #layer_batch_normalization()%>%
-    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=1,use_bias = T)
-
-  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
-                  feature_input = covars_input,
-                  feature_output = covars_output,
-                  originalData = mod_cb_glm$data,
-                  offset=mod_cb_glm$offset,
-                  timeVar = "time",
-                  eventVar= "status"
-  )
-  fit<-fitSmoothHaz(pmnn,epochs=epo,batch_size = bsize,verbose=0)
-  annPreds<-aar(fit,
-                times=times,
-                x_test=test
-  )
-  rownames(annPreds)<-annPreds[,1]
-  annProperpoly<- t(annPreds[,-1])
-  class(annProperpoly)<-c("tunnel",class(annProperpoly)) 
-  ############################
-  ###Brier Score
-  ############################  
-  brierFinalResults <- Score(list("Cox_Lin" = cox,'CB_Logi'=glmProper,
-                                  'DeepSurv'=cnnCleaned,'DeepHit'=deephitCleaned,
-                                  'PMNN_Spline'=annProper,'PMNN_Poly'=annProperpoly,
-                                  'dsm'=dsmSurvProper),
-                             data =fullTest, 
-                             formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)
-  BrierIPAList[[pos]]<-brierFinalResults$Brier$score
-
-  ############################
-  ###C-Index
-  ############################  
-  tempTims<-   as.numeric(colnames(annProper))
-  tempTims<-head(tempTims, -15)
-  tempTims<-tail(tempTims, -10)
-  cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=8)
-  et_train<-as.matrix(train)
-  et_test<-as.matrix(fullTest[,c(11,10)])
-  colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
-                          'PMNN_Spline','PMNN_Poly','dsm')
-
-  cScoreTemp[,1]<- tempTims
-
-  a<-survfit(cox, newdata=fullTest,times=times)
-  rownames(a$surv)<-a$time
-  coxlinSurv<-pyProcess(t(a$surv),times=times)
-  colnames(coxlinSurv)<-seq(0,1,length.out = ncol(coxlinSurv))
-  riskList<-list(coxlinSurv,
-                 glmProper,
-                 cnnCleaned,
-                 deephitCleaned,
-                 annProper,
-                 annProperpoly
-  )
-
-  cScore[[pos]]<-cIndexSummary(et_train=et_train
-                               , et_test=et_test,
-                               riskList=riskList,
-                               cScore=cScoreTemp
-  )
-  saveRDS(BrierIPAList,'100MetaBrier.rds')
-  saveRDS(cScore,'100Metacidx.rds')
-  pos=pos+1
-
-} 
-end_time <- Sys.time()
-
-end_time-start_time
-
-ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=Brier,col=model))+
-  geom_line()
-ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
-  geom_line()
-
-```
-
-```{r ERSPCproofofconcept,eval=F}
-
+```{r ERSPCproofofconcept,eval=T}
+##originally used 1 sample ,now using all data
 data<-ERSPC
 colnames(data)<-c("ScrArm","time","status")
-
+  data<-as.data.frame(model.matrix(~.,data)[,-1])
 mod_cb_glm <- fitSmoothHazard(status~.,
@@
                 eventVar= "status",
-                optimizer= 'rmsprop'
+                optimizer= 'adam'
 )
@@
 fit<-fitSmoothHaz(pmnn,epochs=12000,#15000
-                  batch_size = 159893,#159893,#
+                  batch_size = 1024,#159893,#
                   verbose=0,monitor = "loss",val_split=0)
-times<-seq(0,max(data$time),by=0.01)
+times<-seq(0,max(data$time),length.out=100)
 get_weights(fit$model)
 coef(mod_cb_glm)
-annPreds<-aar(fit,
+
+startOG <- Sys.time()
+annPreds<-aarOld(fit,
               times=times,
-              x_test=data[1,]
+              x_test=data[1:100,]
 )
-#absoluteRisk(mod_cb_glm,time=times,newdata=data[1,],type="CI")
+endOG <- Sys.time()
+startq <- Sys.time()
+annQuick<-aar(fit,times=times,x_test=data[1:100,])
+endq <- Sys.time()
+
+startOG-endOG
+startq-endq
 
@@
 
-saveRDS(results,"../paper/figures/proofOfConceptWeights.rds")
+#saveRDS(results,"../paper/figures/proofOfConceptWeights.rds")
 

### 2022-05-25 analyses/supportStudy.Rmd
cleaned, licensed version of cbnn

--- 2022-05-25 analyses/complexSim.Rmd
@@
 ---
-title: "iterSim"
+title: "Support"
 output: html_document
@@
 
-```{r}
+```{r,eval=T}
 
@@
 
-```{python}
+```{python,eval=T}
 
@@
 os.environ["CUDA_VISIBLE_DEVICES"] = ""
+#os.environ["OM_NUM_THREADS"]="2"
 
@@
 
-hardsim<-T
-numSamp<-2000
 ```
 
-```{r complexPrep,eval=hardsim}
-
-##############################
-###set up complex simulation
-##############################
-
-# Define a function returning the log cum hazard at time t
-logcumhaz <- function(t, x, betas, knots) {
-
-  # Obtain the basis terms for the spline-based log
-  # cumulative hazard (evaluated at time t)
-  basis <- flexsurv::basis(knots, log(t))
-
-  # Evaluate the log cumulative hazard under the
-  # Royston and Parmar specification
-  res <-betas[["gamma0"]] * basis[[1]] + 
-    betas[["gamma1"]] * basis[[2]] +
-    betas[["gamma2"]] * basis[[3]] +
-    betas[["gamma3"]] * basis[[4]] +
-    betas[["gamma4"]] * basis[[5]] +
-    betas[["z1"]] * x[["z1"]]+
-    betas[["z2"]] * x[["z2"]]+
-    betas[["z3"]] * x[["z3"]]+
-    betas[["term1"]] * x[["z1"]]*x[["z2"]]*t+
-    betas[["term2"]] * x[["z1"]]*x[["z3"]]+
-    betas[["term3"]] * x[["z2"]]*(x[["z3"]])
-  #
-  # Return the log cumulative hazard at time t
-  res
-}
-url<-'app.txt'
-ds <- read.table(url, sep = '',header = TRUE)
-ds$years.at.risk=pmin(ds$agein80,ds$onset)
-brcancer$rectime<-brcancer$rectime/max(brcancer$rectime)
-true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1, data = brcancer, k = 3) #from flexsurv
-
-# Simulate the event times
-coefficients<-true_mod$coefficients
-coefficients<-c(coefficients,z1=1,z2=1,z3=1,term1=10,term2=2,term3=2)
-
-```
-
-```{r totalSim,eval=hardsim}
+```{r modelsSUPPORT,eval=eval_support}
 
@@
 cScore<-list()
+
+data<- as.data.frame(do.call(cbind, getSupportPycox()))
+colnames(data)[ncol(data)-1]<-'time'
+colnames(data)[ncol(data)]<-'status'
+data<-as.data.frame(model.matrix(~.,data)[,-1])
+data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
+
+covs<-colnames(data)
+
+#split training/testing
+samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.80)) 
+trainOG<-data[samp,]
+test<-data[-samp,-c(ncol(data))]
+fullTest<-data[-samp,,drop=F]
+
 pos<-1
@@
 
-  #########################
-  ###Simulate date
-  #########################
-  cov<-data.frame(id=1:numSamp,
-                  z1=c(rep(1,numSamp/2),rep(0,numSamp/2)),
-                  z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
-                  z3=c(rnorm(numSamp/2,-3,0.5),rnorm(numSamp/2,8,0.5)))
-  true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1,
-                                       data = brcancer, k = 3
-  ) #from flexsurv
-  dat <- simsurv(betas = coefficients, # "true" parameter values
-                 x = cov,                   # covariate data for 200 individuals
-                 knots = true_mod$knots,    # knot locations for splines
-                 logcumhazard = logcumhaz,  # definition of log cum hazard
-                 maxt = 1,               # no right-censoring
-                 interval = c(0,10000000)) # interval for root finding
-  colnames(dat)<-c("id","time","status")
-  samp<-sample(seq(1,nrow(dat),by=1), floor(nrow(dat)*(0.1))) #random censoring
-  dat$status[samp]<-0
-  # Merge the simulated event times onto covariate data frame
-  data <- merge(cov, dat)
-  covs<-colnames(data)
-
-  #########################
-  ###Simulate data
-  #########################
-  samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.8))
-  data<-data[,-1]
-  data[,-c(1,5)]<-as.data.frame(sapply(data[,-c(1,5)], function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
-  train<-data[samp,]
-  test<-data[-samp,-c(ncol(data))]
-  fullTest<-data[-samp,,drop=F]
+  samp<-sample(seq(1,nrow(trainOG),by=1) , nrow(trainOG),replace = T) 
+  train<-trainOG[samp,]
   times<-seq(from=min(test$time),
@@
 
-  ###########################
-  ###casebase optimal model
-  ###########################
-  final_mod_cb_glm <- fitSmoothHazard(status~bs(time)-time+z1+z2+z3+z1:z2:time+z1:z3+z2:z3,
-                                      data = train,
-                                      time = "time",
-                                      event="status",
-                                      ratio = 100
-  )
-
-  summary(final_mod_cb_glm)
-  tglmAbsRisk<-absoluteRisk(final_mod_cb_glm,time=times,newdata=test,type="CI")
-  tglmAbsRisk<-as.data.frame(tglmAbsRisk)
-  rownames(tglmAbsRisk)<-tglmAbsRisk$time
-  trueGlmProper<- t(tglmAbsRisk[-1,-1])
-  class(trueGlmProper)<-c("tunnel",class(trueGlmProper))
-
-  #########################
-  ###casebase+splines
-  #########################
+  ######################
+  ###casebase + splines
+  ######################
   mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
@@
 
-  glmAbsRisk<-absoluteRisk(mod_cb_glm,time=times,newdata=test,type="CI")
+  glmAbsRisk<-as.data.frame(absoluteRisk(mod_cb_glm,time=times,newdata=test,type="CI"))
+  rownames(glmAbsRisk)<-glmAbsRisk$time
+  #prepare to run through riskRegression
+  glmProper<- t(glmAbsRisk[-1,-1])
+  class(glmProper)<-c("tunnel",class(glmProper))
 
-  glmAbsRisk<-as.data.frame(glmAbsRisk)
+  ######################
+  ###DeepSurv
+  ######################
+  coxnn<-deepsurv(Surv(time, status) ~ .,data = train, frac = 0.2, activation = "relu",
+                  num_nodes = c(layer1,layer2,layer3,layer4), dropout = drpt,
+                  early_stopping = TRUE, epochs = epo,
+                  batch_size = bsize,device = "cpu",best_weights=T,
+                  patience = patience,min_delta=min_delta,batch_norm = F,
+                  learning_rate = 0.001,lr_decay=10^-7)
+  coxnnsurv<- predict(coxnn,newdata = test,type="survival")
+  colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
+  cnnCleaned<-pyProcess(coxnnsurv,times=times)
+  colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
 
-  rownames(glmAbsRisk)<-glmAbsRisk$time
-  glmProper<- t(glmAbsRisk[-1,-1])
-
-  class(glmProper)<-c("tunnel",class(glmProper))
+  ######################
+  ###Deephit
+  ######################
+  source_python('src/deephitter.py')
+  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,patience=patience,min_delta=min_delta,drpt=drpt)
+  hitnnSurv<-t(hitnnSurv)
+  colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
+  deephitCleaned<-pyProcess(hitnnSurv,times=times)
+  colnames(deephitCleaned)<-seq(0,1,length.out = ncol(deephitCleaned))
+  py_run_string("del fitDeephit")
+  rm(fitDeephit)
 
@@
   ######################
-  cox<-coxph(Surv(time, status) ~ z1+z2+z3, data = train,x=T)
-
-  ######################
-  ###DeepSurv
-  ######################
-  coxnn<-deepsurv(Surv(time, status) ~ .,data = train, frac = 0.2, activation = "relu",
-                  num_nodes = c(layer1,layer2,layer3,layer4), dropout = drpt,
-                  early_stopping = TRUE, epochs = epo,
-                  batch_size = bsize,device = "cpu",best_weights=T,
-                  patience = patience,min_delta=min_delta,batch_norm = F,
-                  learning_rate = 0.001,lr_decay=10^-7)
-  coxnnsurv<- predict(coxnn,newdata = test,type="survival")
-  colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
-  cnnCleaned<-pyProcess(coxnnsurv,times=times)
-  colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
-
-  ######################
-  ###Deephit
-  ######################
-  source_python('src/deephitter.py')
-  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,patience=patience,min_delta=min_delta,drpt=drpt)
-  hitnnSurv<-t(hitnnSurv)
-  colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
-  deephitCleaned<-pyProcess(hitnnSurv,times=times)
-  colnames(deephitCleaned)<-seq(0,1,length.out = ncol(deephitCleaned))
-  py_run_string("del fitDeephit")
-  rm(fitDeephit)
+  cox<-coxph(Surv(time, status) ~ ., data = train,x=T)
 
@@
 
-  brierFinalResults <- Score(list("Cox_Lin" = cox,
-                                  'CB_Logi'=glmProper,
-                                  'DeepSurv'=cnnCleaned,
-                                  'DeepHit'=deephitCleaned,
-                                  'PMNN_Spline'=annProper,
-                                  'Optimal'=trueGlmProper,
-                                  'PMNN_Poly'=annProperpoly),
-                             #'dsm'=dsmSurvProper) ,#'dsm'=dsmCleaned,
+  ######################
+  ###IPA metric
+  ######################
+  brierFinalResults <- Score(list("Cox_Lin" = cox,'CB_Logi'=glmProper,
+                                  'DeepSurv'=cnnCleaned,'DeepHit'=deephitCleaned,
+                                  'PMNN_Spline'=annProper,'PMNN_Poly'=annProperpoly),
                              data =fullTest, 
@@
                  annProper,
-                 trueGlmProper,
                  annProperpoly
@@
   cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=length(riskList)+1)
-  et_train<-as.matrix(train)[,c(5,4)]
-  et_test<-as.matrix(fullTest[,c(5,4)])
+  et_train<-as.matrix(train)[,c(16,15)]
+  et_test<-as.matrix(fullTest[,c(16,15)])
   colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
-                          'PMNN_Spline','Optimal','PMNN_Poly')#,'dsm')
+                          'PMNN_Spline','PMNN_Poly')
   cScoreTemp[,1]<- tempTims
@@
 
-  saveRDS(BrierIPAList,'results/100hardBrier.rds')
-  saveRDS(cScore,'results/100hardcidx.rds')
+  saveRDS(BrierIPAList,'results/100supBrier.rds')
+  saveRDS(cScore,'results/100supcidx.rds')
   pos<-pos+1
@@
 
+ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=Brier,col=model))+
+  geom_line()
+ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
+  geom_line()
+
+end_time <- Sys.time()
+
+end_time-start_time
+
 ```
+
+```{r}
+
+sessionInfo()
+
+```

### 2022-05-25 analyses/dsm/datasets.py

@@ -1,0 +1,286 @@
+# coding=utf-8
+# MIT License
+
+# Copyright (c) 2020 Carnegie Mellon University, Auton Lab
+
+# Permission is hereby granted, free of charge, to any person obtaining a copy
+# of this software and associated documentation files (the "Software"), to deal
+# in the Software without restriction, including without limitation the rights
+# to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+# copies of the Software, and to permit persons to whom the Software is
+# furnished to do so, subject to the following conditions:
+
+# The above copyright notice and this permission notice shall be included in all
+# copies or substantial portions of the Software.
+
+# THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+# IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+# FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+# AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+# LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+# OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+# SOFTWARE.
+
+
+"""Utility functions to load standard datasets to train and evaluate the
+Deep Survival Machines models.
+"""
+
+
+import io
+import pkgutil
+
+import pandas as pd
+import numpy as np
+
+from sklearn.impute import SimpleImputer
+from sklearn.preprocessing import StandardScaler
+
+import torchvision
+
+def increase_censoring(e, t, p):
+
+  uncens = np.where(e == 1)[0]
+  mask = np.random.choice([False, True], len(uncens), p=[1-p, p])
+  toswitch = uncens[mask]
+
+  e[toswitch] = 0
+  t_ = t[toswitch]
+
+  newt = []
+  for t__ in t_:
+    newt.append(np.random.uniform(1, t__))
+  t[toswitch] = newt
+
+  return e, t
+
+def _load_framingham_dataset(sequential):
+  """Helper function to load and preprocess the Framingham dataset.
+
+  The Framingham Dataset is a subset of 4,434 participants of the well known,
+  ongoing Framingham Heart study [1] for studying epidemiology for
[…]

### 2022-05-25 analyses/dsm/dsm_torch.py

@@ -1,0 +1,566 @@
+# coding=utf-8
+# MIT License
+
+# Copyright (c) 2020 Carnegie Mellon University, Auton Lab
+
+# Permission is hereby granted, free of charge, to any person obtaining a copy
+# of this software and associated documentation files (the "Software"), to deal
+# in the Software without restriction, including without limitation the rights
+# to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+# copies of the Software, and to permit persons to whom the Software is
+# furnished to do so, subject to the following conditions:
+
+# The above copyright notice and this permission notice shall be included in all
+# copies or substantial portions of the Software.
+
+# THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+# IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+# FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+# AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+# LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+# OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+# SOFTWARE.
+
+
+"""Torch model definitons for the Deep Survival Machines model
+
+This includes definitons for the Torch Deep Survival Machines module.
+The main interface is the DeepSurvivalMachines class which inherits
+from torch.nn.Module.
+
+Note: NOT DESIGNED TO BE CALLED DIRECTLY!!!
+
+"""
+
+
+import torch.nn as nn
+import torch
+import numpy as np
+
+__pdoc__ = {}
+
+for clsn in ['DeepSurvivalMachinesTorch',
+             'DeepRecurrentSurvivalMachinesTorch',
+             'DeepConvolutionalSurvivalMachines']:
+  for membr in ['training', 'dump_patches']:
+
+    __pdoc__[clsn+'.'+membr] = False
+
+
+def create_representation(inputdim, layers, activation):
+  r"""Helper function to generate the representation function for DSM.
+
+  Deep Survival Machines learns a representation (\ Phi(X) \) for the input
+  data. This representation is parameterized using a Non Linear Multilayer
[…]

### 2022-05-25 analyses/dsm/utilities.py

@@ -1,0 +1,211 @@
+# coding=utf-8
+# MIT License
+
+# Copyright (c) 2020 Carnegie Mellon University, Auton Lab
+
+# Permission is hereby granted, free of charge, to any person obtaining a copy
+# of this software and associated documentation files (the "Software"), to deal
+# in the Software without restriction, including without limitation the rights
+# to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+# copies of the Software, and to permit persons to whom the Software is
+# furnished to do so, subject to the following conditions:
+
+# The above copyright notice and this permission notice shall be included in all
+# copies or substantial portions of the Software.
+
+# THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+# IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+# FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+# AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+# LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+# OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+# SOFTWARE.
+
+
+"""Utility functions to train the Deep Survival Machines models"""
+
+from dsm.dsm_torch import DeepSurvivalMachinesTorch
+from dsm.losses import unconditional_loss, conditional_loss
+
+from tqdm import tqdm
+from copy import deepcopy
+
+import torch
+import numpy as np
+
+import gc
+import logging
+
+
+def get_optimizer(model, lr):
+
+  if model.optimizer == 'Adam':
+    return torch.optim.Adam(model.parameters(), lr=lr)
+  elif model.optimizer == 'SGD':
+    return torch.optim.SGD(model.parameters(), lr=lr)
+  elif model.optimizer == 'RMSProp':
+    return torch.optim.RMSprop(model.parameters(), lr=lr)
+  else:
+    raise NotImplementedError('Optimizer '+model.optimizer+
+                              ' is not implemented')
+
+def pretrain_dsm(model, t_train, e_train, t_valid, e_valid,
[…]

### 2022-05-25 analyses/runAll.sh

@@ -1,0 +1,31 @@
+#!/bin/bash
+
+cd ~/pmnnProjects/pmnnPlayground/reviewedAnalysisPMNN
+
+epo=2000
+patience=10
+iteration=100
+layer1=50
+layer2=50
+layer3=25
+layer4=25
+drpt=0.5
+
+
+#explaination of structure (overview)
+#domain knowledge is too specific to be understood at a glance
+
+params='list('"epo=$epo,patience=$patience,iteration=$iteration,layer1=$layer1,layer2=$layer2,layer3=$layer3,layer4=$layer4,drpt=$drpt"')'
+metabric='library(rmarkdown); rmarkdown::render("metabricStudy.Rmd",'"params=$params)"
+support='library(rmarkdown); rmarkdown::render("supportStudy.Rmd",'"params=$params)"
+simple='library(rmarkdown); rmarkdown::render("simpleSim.Rmd",'"params=$params)"
+complex='library(rmarkdown); rmarkdown::render("complexSim.Rmd",'"params=$params)"
+(trap 'kill 0' SIGINT; \
+sleep 1 && eval `echo Rscript -e "'"{$simple}"'"` && bash ~/phoneNotification.sh simpleTest && sleep 9 && eval `echo Rscript -e "'"{$complex}"'"` && bash ~/phoneNotification.sh complexTest \ &
+sleep 6 && eval `echo Rscript -e "'"{$support}"'"` && bash ~/phoneNotification.sh supportTest \ &
+sleep 3 && eval `echo Rscript -e "'"{$metabric}"'"` && bash ~/phoneNotification.sh metabricTest 
+) 
+
+
+
+

### 2022-05-25 analyses/src/dsmScript.py

@@ -1,0 +1,70 @@
+import numpy as np
+import random
+#from dsm import datasets
+from sklearn.model_selection import ParameterGrid
+from dsm import DeepSurvivalMachines
+##in dsm/utilities.py, we set early stopping patience and min_delta manually
+##dsm/dsm_torch.py, line 89 dropout(0.5) layer was added
+
+
+def dsmfitter(train,fullTest,bsize,epochs,tPos=3,ePos=4):
+  
+  ###############################
+  ###Prepare data
+  ###############################
+  tPos=int(tPos)
+  ePos=int(ePos)
+  train2=train.to_numpy()#@np.random.shuffle(train) 
+  np.random.shuffle(train2) 
+  n= np.shape(train2)[0]
+  vl_size = int(n*0.80)
+
+  
+  t_train = np.asarray( train2[:vl_size,tPos])
+  e_train = np.asarray( train2[:vl_size,ePos]).astype(int)
+  x_train = train2[:vl_size,:]
+  x_train = np.delete(x_train, [tPos,ePos], 1)
+  
+  t_val = np.asarray( train2[-vl_size:,tPos])
+  e_val = np.asarray( train2[-vl_size:,ePos]).astype(int)
+  x_val = train2[-vl_size:,:]
+  x_val = np.delete(x_val, [tPos,ePos], 1)
+
+  test=fullTest.to_numpy()
+  t_test = np.asarray( test[:,tPos])
+  e_test = np.asarray( test[:,ePos]).astype(int)
+  x_test = np.delete(test, [tPos,ePos], 1)
+  
+  horizons = np.arange(0, 1, 0.01).tolist()
+  times= np.quantile(t_test[e_test==1], horizons).tolist()
+  
+  
+
+  ###############################
+  ###Prepare data
+  ###############################
+  #I fix the parameter grid
+  param_grid = {'k' : [6],
+                'distribution' : ['Weibull'],
+                'learning_rate' : [ 0.001],
+                'layers' : [ [50,50,25,25] ]
+               }
+  params = ParameterGrid(param_grid)
+  
+  #note that the DeepSurvivalMachines function has been modified: 
+  #it now introduces a dropout layer after each feedforward dense layer.
+  # ### Model Training and Selection
+  models = []
+  for param in params:
+      model = DeepSurvivalMachines(k = param['k'],
+                                   distribution = param['distribution'],
+                                   layers = param['layers'])
[…]

### 2022-05-25 analyses/src/functions.R

@@ -1,0 +1,459 @@
+
+statCalcSim<-function(metricSims){
+  toRemove <- which(apply(metricSims, 1, function(x){any(is.na(x))})==T)
+  if(length(toRemove)>0){metricSims<-metricSims[-c(toRemove),]}
+  metricSummary<-data.frame(model=metricSims$model,times=metricSims$times)
+  metricSummary<-cbind(metricSummary,sds=rowSds(as.matrix(metricSims[,-c(1,2)])),
+                       means=rowMeans(as.matrix(metricSims[,-c(1,2)])))
+  ci<-apply(as.matrix(metricSims[,-c(1,2)]), 1, function(x){mean(x)+c(-1.96,1.96)*sd(x)/sqrt(length(x))})
+  rownames(ci)<-c('cilower95','ciupper95')
+  pi<-apply(as.matrix(metricSims[,-c(1,2)]), 1, function(x){quantile(x,c(.025, .975))})
+  rownames(pi)<-c('pilower95','piupper95')
+  metricSummary<-cbind(metricSummary,t(ci),t(pi))
+  return(metricSummary)
+}
+
+plotPerformance<-function(metricSummary,xlab="Follow-up-time",method='Metric',iterations='NA'){
+  mainTitle<-paste(method,' : 95% P.I. over ', iterations, ' iterations',sep='')
+  
+  a<-ggplot(data=metricSummary, aes(x=times,y=means,col=model))+
+    geom_line() +
+    geom_ribbon(aes(ymin=pilower95,ymax=piupper95,fill=model),linetype=0,alpha=0.3) +
+    ggtitle(mainTitle) + 
+    xlab(xlab) +
+    ylab(method)
+  if(method=="IPA"){
+    a<-a+coord_cartesian()#ylim(-1,1))#
+    return(a)
+  }
+  return(a+coord_cartesian())
+}
+
+
+pyProcess<-function(predictions,times=times){
+  interpolated <- apply(predictions, 1, function(x) approx(as.numeric(colnames(predictions)), y=x, method = "linear", xout=times)$y)
+  class(interpolated)<-c("tunnel",class(interpolated))
+  interpolated<-1-interpolated
+  interpolated[is.na(interpolated) & is.na(interpolated)>50] <- 0
+  interpolated[is.na(interpolated) & is.na(interpolated)<50] <- 1
+  return(t(interpolated))
+}
+
+
+
+
[…]

### 2022-05-25 analyses/src/getMetabricData.py

@@ -1,0 +1,6 @@
+from pycox import datasets
+def getMetabric():
+  return datasets.metabric.read_df()
+  
+  
+  

### 2022-05-25 analyses/src/getSUP.py

@@ -1,0 +1,8 @@
+#import dsm
+#def getSupportDSM():
+#  return dsm.datasets.load_dataset('SUPPORT')
+
+
+import pycox
+def getSupportPycox():
+  return pycox.datasets.support.read_df()

### 2022-05-25 analyses/src/originalpyScripts/cIndex.py

@@ -1,0 +1,38 @@
+
+  # ### Evaluation
+  # 
+  # We evaluate the performance of DSM in its discriminative ability (Time Dependent Concordance Index and Cumulative Dynamic AUC) as well as Brier Score.
+from sksurv.metrics import concordance_index_ipcw
+import numpy as np
+#outrisk is the prediction at time t for the cumulative incidence
+def cidx(et_train, et_test, out_risk, time): 
+  #print(et_train,"\n\n", et_test,"\n\n", out_risk,"\n\n", time)
+  #train= et_train
+  #test=et_test
+  #train= (np.array(et_train[:,0],dtype = bool),np.array(et_train[:,1],dtype = bool))
+  #test= (np.array(et_test[:,0],dtype = bool),np.array(et_test[:,1],dtype = bool))
+  train = np.array([(et_train[i,0], et_train[i,1]) for i in range(np.shape(et_train)[0])],
+                   dtype = [('e', bool), ('t', float)])
+  test = np.array([(et_test[i,0], et_test[i,1]) for i in range(np.shape(et_test)[0])],
+                   dtype = [('e', bool), ('t', float)])
+  #print(test)          
+  #return 
+  return concordance_index_ipcw(train, test, out_risk[:,0], time)[0]
+  
+  #et_train = np.array([(e_train[i], t_train[i]) for i in range(len(e_train))],
+  #                 dtype = [('e', bool), ('t', float)])
+  #et_test = np.array([(e_test[i], t_test[i]) for i in range(len(e_test))],
+  #                 dtype = [('e', bool), ('t', float)])
+  #et_val = np.array([(e_val[i], t_val[i]) for i in range(len(e_val))],
+  #                 dtype = [('e', bool), ('t', float)])
+  #for i, _ in enumerate(times):
+  #    cis.append(concordance_index_ipcw(et_train, et_test, out_risk[:, i], times[i])[0])
+  #brs.append(brier_score(et_train, et_test, out_survival, times)[1])
+  #roc_auc = []
+  #for i, _ in enumerate(times):
+  #    roc_auc.append(cumulative_dynamic_auc(et_train, et_test, out_risk[:, i], times[i])[0])
+  #for horizon in enumerate(horizons):
+  #    print(f"For {horizon[1]} quantile,")
+  #    print("TD Concordance Index:", cis[horizon[0]])
+  #    print("Brier Score:", brs[0][horizon[0]])
+  #    print("ROC AUC ", roc_auc[horizon[0]][0], "\n")

### 2022-05-25 analyses/src/originalpyScripts/deephitter.py

@@ -1,0 +1,77 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+# For preprocessing
+from sklearn.preprocessing import StandardScaler
+from sklearn_pandas import DataFrameMapper 
+
+import torch # For building the networks 
+import torchtuples as tt # Some useful functions
+
+from pycox.models import DeepHitSingle
+from pycox.evaluation import EvalSurv
+
+
+def fitDeephit(train,fullTest,bsize,epochs,ti='time',ev='status'):#,epoch,bsize,tPos,ePos):
+  #print(train)
+  n= np.shape(train)[0]
+  vl_size = int(n*0.20)
+  df_val = train.sample(frac=0.2)
+  df_train = train.drop(df_val.index)
+  df_test = fullTest
+  varNames = list(df_train.columns)
+  unwanted = {ti, ev}
+  varNames = [e for e in varNames if e not in unwanted]
+  #print(varNames)
+  leave = [(col, None) for col in varNames]
+
+  x_mapper = DataFrameMapper(leave)
+  #print(x_mapper)
+
+  x_train = x_mapper.fit_transform(df_train).astype('float32')
+  x_val = x_mapper.transform(df_val).astype('float32')
+  x_test = x_mapper.transform(df_test).astype('float32')
+  
+  #x_train = df_train.drop(columns=['time','status'],inplace=False)
+  #x_val = df_val.drop(columns=['time','status'],inplace=False)
+  #x_test = df_test.drop(columns=['time','status'],inplace=False)
+  #cols_standardize = ['x0', 'x1', 'x2', 'x3', 'x8']
+  #cols_leave = ['x4', 'x5', 'x6', 'x7']
+
+ 
+
+
+
+  ########
+  #prepare labels
+  #######
+  num_durations = 10
+  labtrans = DeepHitSingle.label_transform(num_durations)
+  get_target = lambda df: (df[ti].values, df[ev].values)
+  y_train = labtrans.fit_transform(*get_target(df_train))
+  y_val = labtrans.transform(*get_target(df_val))
+  train = (x_train, y_train)
+  val = (x_val, y_val)
+  durations_test, events_test = get_target(df_test)
+  ########
+  #prepare model
+  ########
+  in_features = x_train.shape[1]
+  num_nodes = [50,50,25,25]
+  out_features = labtrans.out_features
+  batch_norm = False
+  dropout = 0.5
+  
[…]

### 2022-05-25 analyses/src/originalpyScripts/dsmScript.py

@@ -1,0 +1,110 @@
+
+
+import numpy as np
+
+import random
+from dsm import datasets
+#x, t, e = datasets.load_dataset('SUPPORT')
+#t=t/10000
+from sklearn.model_selection import ParameterGrid
+from dsm import DeepSurvivalMachines
+#import numpy as np
+#split data
+#n = len(x)
+
+#tr_size = int(n*0.70)
+#vl_size = int(n*0.10)
+#te_size = int(n*0.20)
+
+#x_train, x_test, x_val = x[:tr_size], x[-te_size:], x[tr_size:tr_size+vl_size]
+#t_train, t_test, t_val = t[:tr_size], t[-te_size:], t[tr_size:tr_size+vl_size]
+#e_train, e_test, e_val = e[:tr_size], e[-te_size:], e[tr_size:tr_size+vl_size]
+
+
+
+
+#x = r.data.iloc[:,:3].to_numpy()
+#x
+#t =np.asarray( r.data.iloc[:,3])
+#t
+#e =np.asarray( r.data.iloc[:,4])
+#e
+
+def dsmfitter(train,fullTest,bsize,epochs,tPos=3,ePos=4):
+  tPos=int(tPos)
+  ePos=int(ePos)
+  train2=train.to_numpy()#@np.random.shuffle(train) 
+  np.random.shuffle(train2) 
+  n= np.shape(train2)[0]
+  vl_size = int(n*0.80)
+  #vls=random.sample(range(n),vl_size)
+  
+  t_train = np.asarray( train2[:vl_size,tPos])#,dtype=np.float32)
+  e_train = np.asarray( train2[:vl_size,ePos]).astype(int)
+  x_train = train2[:vl_size,:]#.to_numpy()
+  x_train = np.delete(x_train, [tPos,ePos], 1)
+  
+  t_val = np.asarray( train2[-vl_size:,tPos])#,dtype=np.float32)
+  e_val = np.asarray( train2[-vl_size:,ePos]).astype(int)
+  x_val = train2[-vl_size:,:]#.to_numpy()
+  x_val = np.delete(x_val, [tPos,ePos], 1)
+
+  #print(np.shape(train2))
+  #print(np.shape(x_train))
+  #print(np.shape(x_val))
+  test=fullTest.to_numpy()
+  
+  t_test = np.asarray( test[:,tPos])
+  e_test = np.asarray( test[:,ePos]).astype(int)
[…]

### 2022-05-25 analyses/src/packages.R

@@ -1,0 +1,53 @@
+#code to install missing packages
+
+
+getPckg <- function(pckgName) {install.packages(paste(pckgName), repos = "http://cran.r-project.org")}
+getPckgBM <- function(pckgName) {BiocManager::install(paste(pckgName))}
+
+pckgNamesCran = c("casebase",
+                  "DT",
+                  "data.table",
+                  "flexsurv",
+                  "ggplot2",
+                  "keras",
+                  "matrixStats",
+                  "mltools",
+                  "RColorBrewer",
+                  "riskRegression",
+                  "pseudo",
+                  "simsurv",
+                  "splines",
+                  "survival",
+                  "survminer",
+                  "survivalmodels",
+                  "tensorflow",
+                  "visreg")
+
+for(pckgName in pckgNamesCran){
+  pckg = try(require(pckgName,character.only = TRUE))
+  if(!pckg) {
+    cat(paste("Installing '",pckgName, "' from CRAN\n",sep=""))
+    getPckg(pckgName)
+    require(pckgName,character.only = TRUE)
+  }
+}
+
+
+
+#biomaRt from bioconductor can give us gene name to symbol transformations.
+pckgNamesBioC<-c()#"TCGAbiolinks"
+
+for(pckgName in pckgNamesBioC){
+  pckg = try(require(pckgName, character.only = TRUE))
+  if(!pckg) {
+    cat(paste("Installing '",pckgName, "' from CRAN\n", sep=""))
+    getPckgBM(pckgName)
+    require(pckgName, character.only = TRUE)
+  }
+}
+
+
+
+
+
+

### 2022-06-01 SIMPaper/sbpaper.Rmd
finalized paper

@@
 In the simulations and real data applications, CBNN provides a more consistent predictive performance across time and outperforms the competing neural network approaches.
-For a simple simulation with an exponential hazard model, CBNN consistently outperforms the other neural network methods.
+For a simple simulation with an exponential hazard model, CBNN outperforms the other neural network methods.
 For a complex simulation, which highlights the ability of CBNN to model both a complex baseline hazard and time-varying interactions, CBNN outperforms all competitors.
 The first real data application shows CBNN outperforming all neural network competitors, while a second real data application shows competitive performance.
-This study highlights the benefit of combining case-base sampling with deep learning to provide a simple and flexible modeling framework for data-driven, time-varying interaction modeling of survival outcomes."
+This study highlights the benefit of combining case-base sampling with deep learning to provide a simple and flexible modeling framework for data-driven, time-varying interaction modeling of survival outcomes.
+An R package is available at [https://github.com/Jesse-Islam/cbnn](https://github.com/Jesse-Islam/cbnn)."
 
@@
 # Introduction
-Smooth-in-time accelerated failure time (AFT) models can estimate absolute risks by modeling the hazard directly through user-specified baseline hazard distribution [@kleinbaum2012survival].
+Smooth-in-time accelerated failure time (AFT) models can estimate absolute risks by modeling the hazard directly through a user-specified baseline hazard distribution [@kleinbaum2012survival].
 Cox proportional hazards models are used more often than AFT models, causing analyses to be based on hazard ratios and relative risks rather than on survival curves and absolute risks [@hanley2009].
-The identification of an appropriate distribution for the baseline hazard in an AFT model may be difficult for common diseases that have many interacting risk factors or a Cox model whose pathogenesis may change with age, making it difficult to maintain the proportional hazards assumption.
-For example, previous studies of breast cancer incidence have discovered time-varying interactions with covariates of interest, including tumor size [@coradini2000time].
-One approach to resolve this involves using the basis of splines on time to provide flexibility in the baseline hazard [@royston2002flexible].
+The identification of an appropriate distribution for the baseline hazard in an AFT model may be difficult for common diseases that have many interacting risk factors, or a Cox model where the disease pathogenesis may change with age, making it difficult to maintain the proportional hazards assumption.
+For example, previous studies of breast cancer incidence have discovered time-varying interactions with covariates of interest, such as tumor size [@coradini2000time].
+One approach to provide flexibility in the baseline hazard involves using the basis of splines on time in our model [@royston2002flexible].
 However, regression-based models are limited in that they require prior knowledge about potential time-varying interactions and their quantitative effects.
@@
 We note that these alternative neural network approaches all require custom loss functions [@katzman2018DeepSurv] [@lee2018DeepHit] [@dsmPaper].
-DeepHit introduces a new hyperparameter weighing its two loss functions (negative log-likelihood and ranking losses) while DSM requires a two-phase learning process and user implementations for distributions beyond Log-Normal or Weibull  [@lee2018DeepHit] [@dsmPaper].
+DeepHit introduces a hyperparameter weighing its two loss functions (negative log-likelihood and ranking losses) while DSM requires a two-phase learning process and user implementations for distributions beyond Log-Normal or Weibull  [@lee2018DeepHit] [@dsmPaper].
 Regression-based approaches require prior specification of all interaction terms, which makes it challenging to model covariate effects that change over time.
@@
 For example, by including splines of time as covariates, we can approximate the Royston-Parmar flexible baseline hazard model [@royston2002flexible], however, this still requires explicit use of time-varying interactions.
-CBNN can model both without user specification; only a sufficiently deep network is required.
+CBNN can model both without extra tuning parameters.
 
@@
 Finally, we report on the software versions used.
-A package is available for use at [GITHUBLINK].
-The entire code base to generate the analyses in this paper is available at [GITHUB LINK].
+An R package is available for use at [https://github.com/Jesse-Islam/cbnn](https://github.com/Jesse-Islam/cbnn).
+The entire code base to generate the analyses in this paper is available at [https://github.com/Jesse-Islam/cbnnManuscript](https://github.com/Jesse-Islam/cbnnManuscript).
 
@@
 
-After case-base sampling, we pass all features, including time, into any user-defined feed-forward component, to which an offset term (to adjust for bias from case-base sampling) is added, then passed through a sigmoid activation function (Figure \ref{fig:NNarch}).
+After case-base sampling, we pass all features, including time, into any user-defined feed-forward component, to which an offset term is added, then passed through a sigmoid activation function (Figure \ref{fig:NNarch}).
 We use the sigmoid function as its inverse is the odds, which we can use to calculate the hazard.
@@
 where $T$ is a random variable representing the event time, $X$ is the random variable for a covariate profile, $f_{\theta}(X, T)$ represents any feed-forward neural network architecture, $\log\left(\frac{B}{b}\right)$ is bias term set by case-base sampling, $\theta$ is the set of parameters learned by the neural network and $\sigmoid(x)=\frac{1}{1+e^{-x}}$.
-By approximating a higher-order polynomial of time using a neural network, the baseline hazard specification is now data-driven, where user-defined hyperparameters such as regularization, number of layers and nodes control the flexibility of the model.
+By approximating a higher-order polynomial of time using a neural network, the baseline hazard specification is now data-driven, where user-defined hyperparameters such as regularization, number of layers and nodes control the flexibility of the fukk hazard function.
 We provide a detailed description of the choices we made in the next sub-section.
@@
 These hyperparameter decisions can affect the estimated parameters and were chosen during a set of initial simulations to determine if CBNN can learn complex interactions in practice.
-We set a max number of epochs to be $2000$, batch size as $512$, learning rate as $10^{-3}$, decay as $10^{-7}$, patience as $10$ epochs, $\{50,50,25,25\}$ nodes in each hidden layer with $50\%$ dropout at each layer, a validation split of $20\%$, a testing split of $10%$, a minimum delta loss on the validation set of $10^{-7}$ over 10 epochs and Adam [@gulli2017] as our optimizer.
+We set a max number of epochs to be $2000$, batch size as $512$, learning rate as $10^{-3}$, decay as $10^{-7}$, patience as $10$ epochs, $\{50,50,25,25\}$ nodes in each hidden layer with $50\%$ dropout at each layer, a minimum delta loss on the validation set of $10^{-7}$ over 10 epochs and Adam [@gulli2017] as our optimizer.
 These choices may permit the model to approximate higher-order interactions while preventing over-fitting [@srivastava2014dropout].
@@
 The **simsurv** package [@simsurv] is used for our simulation studies, while **flexsurv** [@flexsurv] is used to fit a flexible baseline hazard using splines for our complex simulation.
-We use the implementation of $C_{IPCW}$???
-from the python package **sksurv** [@sksurv].
+We use the implementation of $C_{IPCW}$ from the python package **sksurv** [@sksurv].
 The **riskRegression** package [@riskRegression] is used to get the Index of Prediction Accuracy (IPA metric).
@@
 where $\gamma_{1}=3.9, \gamma_{2}=3, \gamma_{3}=-0.43, \gamma_{4}=1.33,\gamma_{5}=-0.86, \beta_{{1}}=1, \beta_{{2}}=1, \beta_{{3}}=1, \tau_{1}=10, \tau_{2}=2, \tau_{3}=2$ and $\psi$ are basis splines.
-To generate the $gamma$ coefficients, let $g$: $\mathbb{R} \rightarrow \mathbb{R}$ be a smoothing method for variable $t$ by a projection on to a set of basis functions on splines:
-\srb{i used gt as the function for splines.
-Specifically, natural cubic spline function of log time https://www.rdocumentation.org/packages/flexsurv/versions/2.1/topics/flexsurvspline .
-therefore, should i use g(log(t)) and and replace "smoothing method" with natural cubic spline?
-SAHIR: will fix any issues in this section in the final version.}
-\begin{align}
-g(t)=\sum^{m}_{l=1}\psi_{l} (t)\gamma_{l}.
-\end{align}
-
-An initial model with three knots ($m=5$) is fit using the *flexsurvspline* function with an intercept-only model on the data, which returns the coefficients ($\gamma$) and basis of time ($\psi$) we use in our model.
+To generate the $gamma$ coefficients, an initial cubic splines model with three knots is fit using the *flexsurvspline* function with an intercept-only model on the data, which returns the coefficients ($\gamma$) we use in our model.
 Note that we fix these values for the analysis.

### 2022-06-01 SIMPaper/sbpaper.tex
finalized paper

@@
 \abstract{Neural network-based survival methods have been developed to
-model data-driven covariate interactions. While this flexibility has led
-to methods with better predictive performance than regression-based
-approaches, current neural networks cannot model both time-varying
-interactions and complex baseline hazards. To address this, we propose
-Case-Base Neural Networks (CBNN) as a new approach that combines the
-case-base sampling framework with flexible architectures. Our method
-naturally accounts for censoring, does not require additional
-hyperparameters and uses binary cross-entropy as the loss function.
-Using a novel sampling scheme and data augmentation, we incorporate time
-directly into a feed-forward neural network. CBNN predicts the
-probability of an event occurring at a given moment and estimates the
-hazard function. We compare the performance of CBNN to survival methods
-based on regression and neural networks in two simulations and two real
-data applications. We report two time-dependent metrics for each model.
-In the simulations and real data applications, CBNN provides a more
-consistent predictive performance across time and outperforms the
-competing neural network approaches. For a simple simulation with an
-exponential hazard model, CBNN consistently outperforms the other neural
-network methods. For a complex simulation, which highlights the ability
-of CBNN to model both a complex baseline hazard and time-varying
-interactions, CBNN outperforms all competitors. The first real data
-application shows CBNN outperforming all neural network competitors,
-while a second real data application shows competitive performance. This
-study highlights the benefit of combining case-base sampling with deep
-learning to provide a simple and flexible modeling framework for
-data-driven, time-varying interaction modeling of survival outcomes.}
+model data-driven covariate interactions. While these methods have led
+to better predictive performance than regression-based approaches, they
+cannot model both time-varying interactions and complex baseline
+hazards. To address this, we propose Case-Base Neural Networks (CBNN) as
+a new approach that combines the case-base sampling framework with
+flexible architectures. Our method naturally accounts for censoring,
+does not require additional hyperparameters and uses binary
+cross-entropy as the loss function. Using a novel sampling scheme and
+data augmentation, we incorporate time directly into a feed-forward
+neural network. CBNN predicts the probability of an event occurring at a
+given moment and estimates the hazard function. We compare the
+performance of CBNN to survival methods based on regression and neural
+networks in two simulations and two real data applications. We report
+two time-dependent metrics for each model. In the simulations and real
+data applications, CBNN provides a more consistent predictive
+performance across time and outperforms the competing neural network
+approaches. For a simple simulation with an exponential hazard model,
+CBNN outperforms the other neural network methods. For a complex
+simulation, which highlights the ability of CBNN to model both a complex
+baseline hazard and time-varying interactions, CBNN outperforms all
+competitors. The first real data application shows CBNN outperforming
+all neural network competitors, while a second real data application
+shows competitive performance. This study highlights the benefit of
+combining case-base sampling with deep learning to provide a simple and
+flexible modeling framework for data-driven, time-varying interaction
+modeling of survival outcomes. An R package is available at
+\url{https://github.com/Jesse-Islam/cbnn}.}
 
@@
 Smooth-in-time accelerated failure time (AFT) models can estimate
-absolute risks by modeling the hazard directly through user-specified
+absolute risks by modeling the hazard directly through a user-specified
 baseline hazard distribution \citep{kleinbaum2012survival}. Cox
@@
 an AFT model may be difficult for common diseases that have many
-interacting risk factors or a Cox model whose pathogenesis may change
-with age, making it difficult to maintain the proportional hazards
-assumption. For example, previous studies of breast cancer incidence
-have discovered time-varying interactions with covariates of interest,
-including tumor size \citep{coradini2000time}. One approach to resolve
-this involves using the basis of splines on time to provide flexibility
-in the baseline hazard \citep{royston2002flexible}. However,
+interacting risk factors, or a Cox model where the disease pathogenesis
+may change with age, making it difficult to maintain the proportional
+hazards assumption. For example, previous studies of breast cancer
+incidence have discovered time-varying interactions with covariates of
+interest, such as tumor size \citep{coradini2000time}. One approach to
+provide flexibility in the baseline hazard involves using the basis of
+splines on time in our model \citep{royston2002flexible}. However,
 regression-based models are limited in that they require prior knowledge
@@
 custom loss functions \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}
-\citep{dsmPaper}. DeepHit introduces a new hyperparameter weighing its
-two loss functions (negative log-likelihood and ranking losses) while
-DSM requires a two-phase learning process and user implementations for
+\citep{dsmPaper}. DeepHit introduces a hyperparameter weighing its two
+loss functions (negative log-likelihood and ranking losses) while DSM
+requires a two-phase learning process and user implementations for
 distributions beyond Log-Normal or Weibull \citep{lee2018DeepHit}
@@
 \citep{royston2002flexible}, however, this still requires explicit use
-of time-varying interactions. CBNN can model both without user
-specification; only a sufficiently deep network is required.
+of time-varying interactions. CBNN can model both without extra tuning
+parameters.
 
@@
 explicitly incorporating time as a feature while adjusting for the
-sampling bias. Finally, we report on the software versions used. A
-package is available for use at {[}GITHUBLINK{]}. The entire code base
-to generate the analyses in this paper is available at {[}GITHUB
-LINK{]}.
+sampling bias. Finally, we report on the software versions used. An R
+package is available for use at
+\url{https://github.com/Jesse-Islam/cbnn}. The entire code base to
+generate the analyses in this paper is available at
+\url{https://github.com/Jesse-Islam/cbnnManuscript}.
 
@@
 After case-base sampling, we pass all features, including time, into any
-user-defined feed-forward component, to which an offset term (to adjust
-for bias from case-base sampling) is added, then passed through a
-sigmoid activation function (Figure \ref{fig:NNarch}). We use the
-sigmoid function as its inverse is the odds, which we can use to
-calculate the hazard. The general form for the neural network using CBNN
-is:
+user-defined feed-forward component, to which an offset term is added,
+then passed through a sigmoid activation function (Figure
+\ref{fig:NNarch}). We use the sigmoid function as its inverse is the
+odds, which we can use to calculate the hazard. The general form for the
+neural network using CBNN is:
 
@@
 hyperparameters such as regularization, number of layers and nodes
-control the flexibility of the model. We provide a detailed description
-of the choices we made in the next sub-section.
+control the flexibility of the fukk hazard function. We provide a
+detailed description of the choices we made in the next sub-section.
 
@@
 \(\{50,50,25,25\}\) nodes in each hidden layer with \(50\%\) dropout at
-each layer, a validation split of \(20\%\), a testing split of \(10%
-\), a minimum delta loss on the validation set of \(10^{-7}\) over 10
-epochs and Adam \citep{gulli2017} as our optimizer. These choices may
-permit the model to approximate higher-order interactions while
-preventing over-fitting \citep{srivastava2014dropout}. We fix a
+each layer, a minimum delta loss on the validation set of \(10^{-7}\)
+over 10 epochs and Adam \citep{gulli2017} as our optimizer. These
+choices may permit the model to approximate higher-order interactions
+while preventing over-fitting \citep{srivastava2014dropout}. We fix a
 train-validation-test split that allows us to update the weights with a
@@
 to fit a flexible baseline hazard using splines for our complex
-simulation. We use the implementation of \(C_{IPCW}\)??? from the python
+simulation. We use the implementation of \(C_{IPCW}\) from the python
 package \textbf{sksurv} \citep{sksurv}. The \textbf{riskRegression}
@@
 and \(\psi\) are basis splines. To generate the \(gamma\) coefficients,
-let \(g\): \(\mathbb{R} \rightarrow \mathbb{R}\) be a smoothing method
-for variable \(t\) by a projection on to a set of basis functions on
-splines:
-\srb{i used gt as the function for splines.
-Specifically, natural cubic spline function of log time https://www.rdocumentation.org/packages/flexsurv/versions/2.1/topics/flexsurvspline .
-therefore, should i use g(log(t)) and and replace "smoothing method" with natural cubic spline?
-SAHIR: will fix any issues in this section in the final version.}
-\begin{align}
-g(t)=\sum^{m}_{l=1}\psi_{l} (t)\gamma_{l}.
-\end{align}
-
-An initial model with three knots (\(m=5\)) is fit using the
+an initial cubic splines model with three knots is fit using the
 \emph{flexsurvspline} function with an intercept-only model on the data,
-which returns the coefficients (\(\gamma\)) and basis of time (\(\psi\))
-we use in our model. Note that we fix these values for the analysis. The
-\(\beta\) coefficients represent direct effects, \(\tau_{2}\) and
-\({\tau_3}\) represent interactions and \(\tau_{1}\) is a time-varying
-interaction.
+which returns the coefficients (\(\gamma\)) we use in our model. Note
+that we fix these values for the analysis. The \(\beta\) coefficients
+represent direct effects, \(\tau_{2}\) and \({\tau_3}\) represent
+interactions and \(\tau_{1}\) is a time-varying interaction.
 

### 2022-06-03 SIMPaper/SIMPaper.Rmd
finalized version for arxiv

@@
 accepted: 2022-07-01
-abstract: "In survival analysis, covariate interactions are difficult to model using regression-based methods without prior knowledge.
-Neural network-based survival models have been developed for data-driven interaction modeling, however, these approaches lack implementations for time-varying interactions or a flexible baseline hazard.
-To address this, we extend the case-base sampling framework and propose Case-Base Neural Networks (CBNN), which introduce time directly into the model, using commonly available neural network components for a simple implementation.
-First, we case-base sample the survival data, converting the survival time of each participant into discrete moments in time.
-This allows us to estimate the probability an event occurs at a given moment using a neural network to get the hazard.
-We compare CBNN to survival methods based on regression and neural networks in two simulations and two real data applications with a fixed neural network architecture.
-We report metrics for each model over follow-up time using both a proper (Index of Prediction Accuracy) and non-proper (inverse probability censoring weights-adjusted concordance index) scoring rule.
-A simple simulation shows CBNN outperforming the other neural network methods.
-The complex simulation highlights the ability of CBNN to model both a complex baseline hazard and time-varying interactions, outperforming all competitors.
+abstract: "Neural network-based survival methods have been developed to model data-driven covariate interactions.
+While these methods have led to better predictive performance than regression-based approaches, they cannot model both time-varying interactions and complex baseline hazards.
+To address this, we propose Case-Base Neural Networks (CBNN) as a new approach that combines the case-base sampling framework with flexible architectures.
+Our method naturally accounts for censoring, does not require additional hyperparameters and uses binary cross-entropy as the loss function.
+Using a novel sampling scheme and data augmentation, we incorporate time directly into a feed-forward neural network.
+CBNN predicts the probability of an event occurring at a given moment and estimates the hazard function.
+We compare the performance of CBNN to survival methods based on regression and neural networks in two simulations and two real data applications.
+We report two time-dependent metrics for each model.
+In the simulations and real data applications, CBNN provides a more consistent predictive performance across time and outperforms the competing neural network approaches.
+For a simple simulation with an exponential hazard model, CBNN outperforms the other neural network methods.
+For a complex simulation, which highlights the ability of CBNN to model both a complex baseline hazard and time-varying interactions, CBNN outperforms all competitors.
 The first real data application shows CBNN outperforming all neural network competitors, while a second real data application shows competitive performance.
-CBNN has shown that it is more consistent in prediction performance across time in the simulations and real data applications compared to the competing neural network approaches."
+We highlight the benefit of combining case-base sampling with deep learning to provide a simple and flexible modeling framework for data-driven, time-varying interaction modeling of survival outcomes.
+An R package is available at [https://github.com/Jesse-Islam/cbnn](https://github.com/Jesse-Islam/cbnn)."
 
@@
 
-In survival analysis, covariate interactions are difficult to model using regression-based methods without prior knowledge.
-Neural network-based survival models have been developed for data-driven interaction modeling, however, these approaches can not model both time-varying interactions and complex baseline hazards.
-To address this, we propose Case-Base Neural Networks (CBNN) which extends the case-base sampling framework which introduce time directly into the model.
-We case-base sample the survival data, converting the survival time of each participant into discrete moments in time.
-This allows us to use a feed-forward neural network that estimate the probability of an event occurring at a given moment, which can be used to estimate the hazard function.
-We then compare CBNN to survival methods based on regression and neural networks in two simulations and two real data applications with a fixed neural network architecture.
-We report metrics for each model over follow-up time using both proper (Index of Prediction Accuracy) and non-proper (inverse probability censoring weights-adjusted concordance index) scoring rules.
-A simple simulation with an exponential hazard model shows CBNN consistently outperforming the other neural network methods.
-The complex simulation highlights the ability of CBNN to model both a complex baseline hazard and time-varying interactions, outperforming all competitors.
-The first real data application shows CBNN outperforming all neural network competitors, while a second real data application shows competitive performance.
-In the simulations and real data applications, CBNN provides a more consistent predictive performance across time than the competing neural network approaches.
-
 # Introduction
-
-what is aft, dont like assuming distribution, smooth. Less used than cox. Proportional hazards assumptions hard to maintain in long study, all regression require prior knowledge.
-
-The widespread use of the Cox proportional hazards model has led to a focus on survival models based on hazard ratios and relative risks rather than on survival curves and absolute risks [@hanley2009].
-In clinical settings, Cox proportional hazards models are frequently prioritized over smooth-in-time accelerated failure time (AFT) models [@hanley2009], which can estimate absolute risks by modeling the hazard directly through user specification of a baseline hazard distribution.
-Using splines as the baseline hazard is an alternative to assuming a specific distribution [@royston2002flexible].
-The risk associated with each covariate may not be constant over time, making it difficult to maintain the proportional hazards assumption.
-Previous studies related to breast cancer have discovered time-varying interactions with covariates of interest, like tumor size [@coradini2000time].
-A limitation of regression-based models is the requirement of prior knowledge; the analyst must know potential time-varying interactions and explicitly evaluate them.
-
-Smooth-in-time accelerated failure time (AFT) models can estimate absolute risks by modeling the hazard directly through user specification of a baseline hazard distribution [@kleinbaum2012survival].
-The assumption of a specific distribution tends to be a point of contention.
-A method has been developped using splines, providing flexibility to the baseline hazard  [@royston2002flexible].
-AFT models are underutilized compared to Cox proportional hazards, resulting in a focus on survival models based on hazard ratios and relative risks rather than on survival curves and absolute risks [@hanley2009].
-However, The risk associated with each covariate may not be constant over time, making it difficult to maintain the proportional hazards assumption.
-Previous studies related to breast cancer have discovered time-varying interactions with covariates of interest, like tumor size [@coradini2000time].
-A limitation of all regression-based models is the requirement of prior knowledge; the analyst must know potential time-varying interactions and explicitly evaluate them.
+Smooth-in-time accelerated failure time (AFT) models can estimate absolute risks by modeling the hazard directly through a user-specified baseline hazard distribution [@kleinbaum2012survival].
+Cox proportional hazards models are used more often than AFT models, causing analyses to be based on hazard ratios and relative risks rather than on survival curves and absolute risks [@hanley2009].
+The identification of an appropriate distribution for the baseline hazard in an AFT model may be difficult for common diseases that have many interacting risk factors, or a Cox model where the disease pathogenesis may change with age, making it difficult to maintain the proportional hazards assumption.
+For example, previous studies of breast cancer incidence have discovered time-varying interactions with covariates of interest, such as tumor size [@coradini2000time].
+One approach to provide flexibility in the baseline hazard involves using the basis of splines on time in our model [@royston2002flexible].
+However, regression-based models are limited in that they require prior knowledge about potential time-varying interactions and their quantitative effects.
 
 Neural networks provide a data-driven approach to approximating interaction terms.
-DeepSurv is a neural network-based proportional hazards model that implements the Cox partial log-likelihood as a custom loss function [@katzman2018DeepSurv].
-This results in a stepwise absolute risk curve that cannot accommodate time-varying interactions.
-Compared to Cox regression, DeepSurv has shown an improvement in performance on the Study to Understand Prognoses Preferences and Risk Treatments (SUPPORT) dataset [@knaus1995SUPPORT].
-Farraggi et al.
-suggested a modification to the loss function required for a Cox neural network that can handle non-proportional hazards [@faraggi1995neural].
-Wang et al.
-took the concept of extreme learning machines and applied them in a survival context, removing the need for tuning the number of layers and nodes [@wang2019extreme].
+For example, DeepSurv is a neural network-based proportional hazards model that implements the Cox partial log-likelihood as a custom loss function [@katzman2018DeepSurv], resulting in a stepwise absolute risk curve that cannot accommodate time-varying interactions.
+Compared to Cox regression, DeepSurv shows better performance on the Study to Understand Prognoses Preferences and Risk Treatments (SUPPORT) dataset [@knaus1995SUPPORT].
+To handle non-proportional hazards, a modification to the loss function was proposed [@faraggi1995neural].
+To remove the need for tuning the number of layers and nodes, the concept of extreme learning machines has been applied to a Cox neural network model [@wang2019extreme].
 
-DeepHit requires that each survival time of interest be specified in the model and directly estimates survival curves, rather than deriving a hazard function [@lee2018DeepHit].
-It assumes an inverse Gaussian distribution as the baseline hazard and its concordance index (c-index) score outperforms DeepSurv [@katzman2018DeepSurv] on the Molecular Taxonomy of Breast Cancer International Consortium (METABRIC) dataset [@curtis2012genomic].
+As an alternative method that assumes a baseline hazard distribution, DeepHit specifies each survival time of interest in the model and directly estimates survival curves, rather than deriving a hazard function [@lee2018DeepHit].
+It assumes an inverse Gaussian distribution as the baseline hazard and it outperforms DeepSurv on the Molecular Taxonomy of Breast Cancer International Consortium (METABRIC) dataset [@curtis2012genomic].
 
-Deep Survival Machines (DSM) provides a flexible, parametric survival model using neural networks with a mixture of distributions as the baseline hazard [@dsmPaper].
+Providing further flexibility, Deep Survival Machines (DSM) is a parametric survival model using neural networks with a mixture of distributions as the baseline hazard [@dsmPaper].
 On both the SUPPORT and METABRIC datasets, DSM outperforms DeepSurv and DeepHit [@dsmPaper].
@@
 
-These methods need custom loss and activation functions that are difficult to implement in practice.
-We note the complexity of these alternative neural network approaches; DeepSurv, DeepHit, and DSM all require custom loss functions [@katzman2018DeepSurv] [@lee2018DeepHit] [@dsmPaper], while DSM requires a two-phase learning process and user implementations for distributions beyond Log-Normal or Weibull [@dsmPaper].
+We note that these alternative neural network approaches all require custom loss functions [@katzman2018DeepSurv] [@lee2018DeepHit] [@dsmPaper].
+DeepHit introduces a hyperparameter weighing its two loss functions (negative log-likelihood and ranking losses) while DSM requires a two-phase learning process and user implementations for distributions beyond Log-Normal or Weibull  [@lee2018DeepHit] [@dsmPaper].
 Regression-based approaches require prior specification of all interaction terms, which makes it challenging to model covariate effects that change over time.
+The current neural network models provide flexibility at the cost of opacity, while regression models provide clarity at the cost of flexibility.
+
 In this article, we propose Case-Base Neural Networks (CBNN) as a method that models time-varying interactions and a flexible baseline hazard using commonly available neural network components.
@@
 This sampling technique allows probabilistic models to predict survival outcomes.
-Following the case-base framework, we can use transformations of time as a feature (covariate) to specify different baseline hazards.
-For example, by including splines of time as covariates, we can approximate the Royston-Parmar flexible baseline hazard model [@royston2002flexible], however, this still requires user specification of time-varying interactions.
-CBNN can model both without user specification; only a sufficiently deep network is required.
+As part of the case-base framework, we use transformations of time as a feature (covariate) to specify different baseline hazards.
+For example, by including splines of time as covariates, we can approximate the Royston-Parmar flexible baseline hazard model [@royston2002flexible], however, this still requires explicit use of time-varying interactions.
+CBNN can model both without extra tuning parameters.
 
 In Section \ref{methods}, we describe how case-base sampling and neural networks are combined both conceptually and algebraically, along with our hyperparameter choices and software implementation.
-In Section \ref{sims}, we describe our metrics and compare the performance of CBNN, DeepSurv, DeepHit, DSM, Cox regression, and case-base using logistic regression on simulated data.
-Section \ref{casestudies} describes the real-data analysis, while in Section \ref{discussion}, we explore the implications of our results and contextualize them within neural network survival analysis in a single event setting.
+Section \ref{sims}, describes our metrics and compares the performance of CBNN, DeepSurv, DeepHit, DSM, Cox regression and case-base using logistic regression (CBLR) on simulated data.
+Section \ref{casestudies} describes the real-data analysis, while Section \ref{discussion} explores the implications of our results and contextualizes them within neural network survival analysis in a single event setting.
 
-# Case-base neural network methodology, metrics, and software {#methods}
+# Case-base neural network methodology, metrics and software {#methods}
 
@@
 Finally, we report on the software versions used.
-A package is available for use at [GITHUBLINK].
-The entire code base to generate the analyses in this paper is available at [GITHUB LINK].
+An R package is available for use at [https://github.com/Jesse-Islam/cbnn](https://github.com/Jesse-Islam/cbnn).
+The entire code base to reproduce the figures and empirical results in this paper is available at [https://github.com/Jesse-Islam/cbnnManuscript](https://github.com/Jesse-Islam/cbnnManuscript).
 
@@
 where $T$ is a random variable representing the event time, $X$ is the random variable for a covariate profile, $f_{\theta}(X, T)$ represents any feed-forward neural network architecture, $\log\left(\frac{B}{b}\right)$ is bias term set by case-base sampling, $\theta$ is the set of parameters learned by the neural network and $\sigmoid(x)=\frac{1}{1+e^{-x}}$.
-By allowing a neural network to approximate a higher-order polynomial of time, the baseline hazard specification is now data-driven, where user-defined hyperparameters such as regularization, number of nodes, and layers control the flexibility of the model.
+By approximating a higher-order polynomial of time using a neural network, the baseline hazard specification is now data-driven, where user-defined hyperparameters such as regularization, number of layers and nodes control the flexibility of the fukk hazard function.
 We provide a detailed description of the choices we made in the next sub-section.
@@
 
-Neural networks provide flexibility when defining the architecture and optimization parameters of our models.
-These hyperparameter choices can affect the estimated parameters and were chosen during a set of initial simulations to determine if CBNN can learn complex interactions in practice.
-We set a max number of epochs to be $2000$, batch size as $512$, learning rate as $10^{-3}$, decay as $10^{-7}$, patience as $10$ epochs, $\{50,50,25,25\}$ nodes in each hidden layer with $50\%$ dropout at each layer, a validation split of $20\%$, a testing split of $10%$, a minimum delta loss on the validation set of $10^{-7}$ over 10 epochs, and Adam [@gulli2017] as our optimizer.
+Neural networks are flexible when defining the architecture and optimization parameters.
+These hyperparameter decisions can affect the estimated parameters and were chosen during a set of initial simulations to determine if CBNN can learn complex interactions in practice.
+We set a max number of epochs to be $2000$, batch size as $512$, learning rate as $10^{-3}$, decay as $10^{-7}$, patience as $10$ epochs, $\{50,50,25,25\}$ nodes in each hidden layer with $50\%$ dropout at each layer, a minimum delta loss on the validation set of $10^{-7}$ over 10 epochs and Adam [@gulli2017] as our optimizer.
 These choices may permit the model to approximate higher-order interactions while preventing over-fitting [@srivastava2014dropout].
-We fix a train-validation-test split that allows us to update the weights with a subset of the data (training), assess performance at each epoch (validation) and gauge performance of the final model (test).
+We fix a train-validation-test split that allows us to update the weights with a subset of the data (training), assess performance at each epoch (validation) and gauge performance of the final model (test) for each method.
 We select the best weights after training based on validation loss [@gulli2017].
@@
 R [@Rsoft] and python [@py] are used to evaluate methods from both languages.
-We fit the Cox model using the **survival** package [@survpkg], the CBLR model using the **casebase** package [@cbpkg], the DeepSurv model using the **survivalmodels** package [@survmods], the DSM model using **DeepSurvivalMachines** [@dsmPaper] and the DeepHit model using **pycox** [@lee2018DeepHit].
-We made the components of CBNN using the **keras** [@keras] package and the **casebase** package for the sampling step.
-The **simsurv** package [@simsurv] is used to simulate our simulation studies, while **flexsurv** [@flexsurv] is used to fit a flexible baseline hazard using splines for our complex simulation.
-We use the implementation of c-ipcw from the python package **sksurv** [@sksurv].
-The **riskRegression** package [@riskRegression] is used to get the Brier score and IPA metrics.
-We changed the **riskRegression** package to be used with any user supplied risk function $F$.
+We fit the Cox model using the **survival** package [@survpkg], the CBLR model using the **casebase** package [@cbpkg], the DeepSurv model using the **survivalmodels** package [@survmods], the DeepHit model using **pycox** [@lee2018DeepHit] and the DSM model using **DeepSurvivalMachines** [@dsmPaper].
+We made the components of CBNN using the **casebase** package for the sampling step and the **keras** [@keras] package for our neural network architecture.
+The **simsurv** package [@simsurv] is used for our simulation studies, while **flexsurv** [@flexsurv] is used to fit a flexible baseline hazard using splines for our complex simulation.
+We use the implementation of $C_{IPCW}$ from the python package **sksurv** [@sksurv].
+The **riskRegression** package [@riskRegression] is used to get the Index of Prediction Accuracy (IPA metric).
+Both metrics are described in detail in the following section.
+We modify the **riskRegression** package to be used with any user supplied risk function $F$.
 To ensure that both R and Python-based models are running in unison on the same data through our simulations and bootstrap, we use the **reticulate** package [@reticulate].
@@
 where $\gamma_{1}=3.9, \gamma_{2}=3, \gamma_{3}=-0.43, \gamma_{4}=1.33,\gamma_{5}=-0.86, \beta_{{1}}=1, \beta_{{2}}=1, \beta_{{3}}=1, \tau_{1}=10, \tau_{2}=2, \tau_{3}=2$ and $\psi$ are basis splines.
-To generate the $gamma$ coefficients, let $g$: $\mathbb{R} \rightarrow \mathbb{R}$ be a smoothing method for variable $t$ by a projection on to a set of basis functions on splines:
-\srb{i used gt as the function for splines.
-Specifically, natural cubic spline function of log time https://www.rdocumentation.org/packages/flexsurv/versions/2.1/topics/flexsurvspline .
-therefore, should i use g(log(t)) and and replace "smoothing method" with natural cubic spline?
-SAHIR: will fix any issues in this section in the final version.}
-\begin{align}
-g(t)=\sum^{m}_{l=1}\psi_{l} (t)\gamma_{l}.
-\end{align}
-
-An initial model with three knots ($m=5$) is fit using the *flexsurvspline* function with an intercept-only model on the data, which returns the coefficients ($\gamma$) and basis of time ($\psi$) we use in our model.
+The $gamma$ coefficients are obtained from an intercept-only cubic splines model with three knots using the *flexsurvspline* function from the **flexsurv** package [@flexsurv].
 Note that we fix these values for the analysis.
@@
 CBNN models survival outcomes by using neural networks on case-base sampled data.
-While estimating the hazard function, we incorporate follow-up-time as a feature, providing a data-driven estimation of a flexible baseline hazard and time-varying interactions.
+We incorporate follow-up time as a feature, providing a data-driven estimate of a flexible baseline hazard and time-varying interactions in our hazard function.
+The three competing neural network models we evaluated cannot model time-varying interactions by design [@dsmPaper] [@katzman2018DeepSurv] [@lee2018DeepHit].
+DSM requires a mixture component distributions for a flexible baseline hazard to be fit [@dsmPaper].
+As our goal is to fix the design and compare performance, we did not change any shared hyperparameters.
+With our choice of shared hyperparameters and model design, DSM did not converge in the complex simulation and SUPPORT case study.
+Despite this limitation, we include this method when it did function as it can fit flexible baseline hazards.
+Compared to CBNN, the remaining two models also have limitations.
+DeepSurv is a proportional hazards model and does not estimate the baseline hazard [@katzman2018DeepSurv].
+DeepHit requires an alpha hyperparameter, is restricted to a single distribution for the baseline hazard and models the survival function directly [@lee2018DeepHit].
+The alternative neural network methods match on time, while CBNN models time directly.
+
+To assess performance among these models, we use both IPA and $C_{IPCW}$ metrics.
+Concordance-based measures are commonly used in survival analysis to compare models and we opt to keep them in our analyses.
+However, $C_{IPCW}$ is a non-proper metric and may cause misspecified models to appear better than they should [@cindexfails2019].
+Therefore, we contextualize our $C_{IPCW}$ results in relation to IPA, a proper scoring rule.
+The model rankings between $C_{IPCW}$ and IPA differed for both the complex simulation and METABRIC application.
+Wider $C_{IPCW}$ confidence intervals for DeepHit and DeepSurv show potential misspecification of the models (Figure \ref{fig:megaPlot} D, H).
+
+With this interpretation of $C_{IPCW}$ in mind, we assess two simulations with minimal noise and two case studies in real scenarios.
+The simple simulation demonstrates potential pitfalls associated with neural network models, particularly overparameterization.
+All neural network-based approaches performed worse than null KM model (IPA score), while the regression-based performed better.
+We attribute this to potential overparameterization in the neural network models, as the wide confidence intervals suggest over-fitting, even with dropout.
+Both DeepSurv and DSM are affected, while DeepHit and CBNN are less so.
+This is a limitation to our strategy for evaluating the methods, which uses a fixed study design across all assessments.
+
+From this baseline benchmark, we move on to a complex simulation that requires a method that can learn both time-varying interactions and have a flexible baseline hazard.
+Here, CBNN demonstrates a distinct advantage over all other methods.
+The regression models show improved performance over the null KM model, while the competing neural network models performed worse.
 Based on our complex simulation results (Figure \ref{fig:megaPlot} C, D and Table \ref{tab:megaTable} B), CBNN outperforms the competitors when time-varying interactions and a complex baseline hazard are present.
+This simulation shows how CBNN can perform under ideal conditions, while the following two analyses on real data serve to assess its performance in realistic conditions.
 
-The competing neural network model we evaluated can not model time varying interactions.
-DeepSurv is a proportional hazards model and does not estimate the baseline hazard [@katzman2018DeepSurv].
-DeepHit requires an alpha hyperparameter, is restricted to a single distribution for the baseline hazard, and models the survival function directly [@lee2018DeepHit].
-DSM requires the user to specify the component distributions in the mixture for a flexible baseline hazard to be fit [@dsmPaper].
-These alternative neural network approaches match on time, while CBNN models time directly.
+In the SUPPORT and METABRIC case studies, flexibility in both interaction modeling and baseline hazard did not improve CBNNs relative performance, suggesting that this flexibility does not aid prediction in either case study.
+From a biological perspective, the 30-year follow-up time in the METABRIC study may contain competing causes of death.
+The causes at the start of the study may not match the causes towards the end, potentially explaining the drop in performance as we reach later survival times with fewer individuals.
+The baseline hazard is unlikely to cause this drop as DSM is the most flexible competing model in our comparison.
+Over-fitting is also unlikely given the tight confidence intervals.
+Further research is required to determine the cause of the drop in performance seen in the METABRIC case study.
+In both case studies, CBNN outperforms the competing neural network methods.
 
-In the simple simulation, all neural network-based approaches performed worse than KM (IPA scores), while the regression-based approaches did not.
-We attribute this to potential over-parameterization in the neural network models.
-The wide confidence intervals suggest over-fitting may be the reason for the loss in performance, even with dropout.
-Both DeepSurv and DSM are affected, while DeepHit and CBNN are less so.
+Taken together, CBNN outperforms all competitors in the complex simulation, demonstrating its value in survival settings that may involve time-varying interactions and a complex baseline hazard.
+Once we perform case-base sampling and adjust for the sampling bias, we can use a sigmoid activation function to predict our hazard function.
+Our approach simplifies the incorporation of censored individuals, allowing survival outcomes to be treated as binary ones.
+Forgoing the requirement of custom loss functions, CBNN only requires the use of standard components in machine learning libraries (specifically, the add layer to adjust for sampling bias and the sigmoid activation function).
+Due to the simplicity in its implementation and by extension user experience, CBNN is both a user-friendly approach to data-driven survival analysis and is easily extendable to any feed-forward neural network framework.
 
-In the complex simulation, CBNN has a distinct advantage over the other methods while examining the IPA score.
-The regression models do show improvement over the null model, while the competing neural network approaches do not.
-The time-varying interactions and complex baseline hazard may have caused a drop in performance for the competing neural network methods, compared to the regression models without interactions.
+# Acknowledgements {-}
 
-In the SUPPORT application, the overall performance considering IPA is near identical.
-The regression models outperform CBNN, which performs better than the competing neural network models.
-As the confidence intervals are tight, it is unlikely that the loss in performance is because of over-fitting.
-Flexibility in both interaction modeling and baseline hazard is not particularly helpful with the SUPPORT dataset.
-
-In the METABRIC application IPA results, DSM and DeepSurv had a steep drop in performance.
-As DeepSurv does not depend on the baseline hazard during the fitting process and DSM is the most flexible competing model in our comparison, the baseline hazard is unlikely to cause this drop.
-With tight confidence intervals, over-fitting is unlikely as well.
-Further research would be required to determine the cause of this drop in performance.
-We do not see the same drop in CBNN demonstrating that, in the METABRIC application, CBNN outperforms DeepSurv and DSM.
-Considering IPA score, regression models consistently outperform the neural network models.
-From a biological perspective, the 30-year follow-up time in the METABRIC study may contain competing causes of death.
-The signal from the start of the study may not match the signal towards the end, potentially explaining the drop in performance as we reach later survival times with fewer individuals.
-
-Concordance-based measures are commonly used in survival analysis to compare models and we opt to keep them in our analyses.
-However, c-ipcw is a non-proper metric and may cause misspecified models to appear better than they should [@cindexfails2019].
-The model rankings between c-ipcw and IPA differed for both the complex simulation and METABRIC application.
-In the complex simulation, the c-ipcw of a correctly specified model (Optimal) underperforms CBNN and CBLR; models without the exact interaction terms specified.
-We base our METABRIC assessment of this disagreement between metrics on the complex simulation.
-The IPA score ranks regression models first, followed by DeepHit and CBNN, but the c-ipcw score ranks regression models after DeepHit and CBNN.
-Wide confidence intervals for neural network models in c-ipcw show potential misspecification, suggesting that regression models perform best in the METABRIC application.
-
-During our testing for the complex simulation and SUPPORT application, DSM did not converge with our choice of hyper-parameters and model design.
-As the goal is to fix the design and compare performance, we did not change any shared hyperparameters.
-For DSM, a different number and combination of component distributions did not help the model converge, though we did not conduct an exhaustive test.
-This may be a limitation to their method or early-stage implementation issues.
-We include this method where it did function as it shares a similar goal of a flexible baseline hazard.
-
-If the user suspects that there may be unknown time-varying interactions and a complex baseline hazard, CBNN should strongly be considered.
-Once we perform case-base sampling and adjust for the sampling bias, we can use a sigmoid activation function to predict our hazard function.
-The user can approach their problem as they would for probabilistic modeling while interpreting their prediction in a survival context.
-In terms of performance, we saw that CBNN consistently performs better than the other neural network-based approaches over follow-up time.
-This may be because of the hyperparameters chosen, as they differ from each of the respective methods papers.
-CBNN only requires the use of standard components in machine learning libraries (add layer and sigmoid activation function).
-Due to the simplicity in its implementation and by extension user experience, we expect CBNN to be both a user-friendly approach to data-driven survival analysis and easily extendable to any feed-forward neural network frameworks.
-
-Though CBNN can model time-varying interactions and a flexible baseline hazard, all models require substantial amounts of data to learn real, complex relationships effectively.
-Larger datasets will appear with time, but the cost of generating longitudinal datasets remains.
-A viable alternative strategy is the use of transfer learning, where independent and identically distributed data can function as a preliminary model to learn related features in a parallel dataset, initializing weights for our survival task [@bozinovski2020reminder].
-
-This paper describes CBNN in the single event setting.
-A next step would be to extend this approach to competing risks.
-A promising use case would be high-dimensional data with complex correlation structures, such as imaging data.
+This work is supported through 5-UM1-DK078616‐14 and R01-HL151855-01 grants.

### 2022-06-03 SIMPaper/SIMPaper.tex
finalized version for arxiv

@@
 
-\abstract{In survival analysis, covariate interactions are difficult to
-model without prior knowledge using regression-based methods. Neural
-network-based survival models have been developed for data-driven
-interaction modeling, however, these approaches lack implementations for
-time-varying interactions or a flexible baseline hazard. To address
-this, we extend the case-base sampling framework and propose Case-Base
-Neural Networks (CBNN), which introduce time directly into the model,
-using commonly available neural network components for a simple
-implementation. First, we case-base sample the survival data, converting
-the survival time of each participant into discrete moments in time.
-This allows us to estimate the probability an event occurs at a given
-moment using a neural network to get the hazard. We compare CBNN to
-survival methods based on regression and neural networks in two
-simulations and two real data applications with a fixed neural network
-architecture. We report metrics for each model over follow-up time using
-both a proper (Index of Prediction Accuracy) and non-proper (inverse
-probability censoring weights-adjusted concordance index) scoring rule.
-A simple simulation shows CBNN outperforming the other neural network
-methods. The complex simulation highlights the ability of CBNN to model
-both a complex baseline hazard and time-varying interactions,
-outperforming all competitors. The first real data application shows
-CBNN outperforming all neural network competitors, while a second real
-data application shows competitive performance. CBNN has shown that it
-is more consistent in prediction performance across time in the
-simulations and real data applications compared to the competing neural
-network approaches.}
+\abstract{Neural network-based survival methods have been developed to
+model data-driven covariate interactions. While these methods have led
+to better predictive performance than regression-based approaches, they
+cannot model both time-varying interactions and complex baseline
+hazards. To address this, we propose Case-Base Neural Networks (CBNN) as
+a new approach that combines the case-base sampling framework with
+flexible architectures. Our method naturally accounts for censoring,
+does not require additional hyperparameters and uses binary
+cross-entropy as the loss function. Using a novel sampling scheme and
+data augmentation, we incorporate time directly into a feed-forward
+neural network. CBNN predicts the probability of an event occurring at a
+given moment and estimates the hazard function. We compare the
+performance of CBNN to survival methods based on regression and neural
+networks in two simulations and two real data applications. We report
+two time-dependent metrics for each model. In the simulations and real
+data applications, CBNN provides a more consistent predictive
+performance across time and outperforms the competing neural network
+approaches. For a simple simulation with an exponential hazard model,
+CBNN outperforms the other neural network methods. For a complex
+simulation, which highlights the ability of CBNN to model both a complex
+baseline hazard and time-varying interactions, CBNN outperforms all
+competitors. The first real data application shows CBNN outperforming
+all neural network competitors, while a second real data application
+shows competitive performance. This study highlights the benefit of
+combining case-base sampling with deep learning to provide a simple and
+flexible modeling framework for data-driven, time-varying interaction
+modeling of survival outcomes. An R package is available at
+\url{https://github.com/Jesse-Islam/cbnn}.}
 
@@
 
-The widespread use of the Cox proportional hazards model has led to a
-focus on survival models based on hazard ratios and relative risks
-rather than on survival curves and absolute risks \citep{hanley2009}. In
-clinical settings, Cox models are frequently prioritized over
-smooth-in-time accelerated failure time (AFT) models \citep{hanley2009},
-which can estimate absolute risks by modeling the hazard directly
-through user specification of a baseline hazard distribution. The user
-may forgo this choice by implementing a flexible set of parameters using
-splines, as proposed by Royston and Palmar \citep{royston2002flexible}.
-These regression models also require user-specified interactions, which
-may limit risk prediction accuracy.
+Smooth-in-time accelerated failure time (AFT) models can estimate
+absolute risks by modeling the hazard directly through a user-specified
+baseline hazard distribution \citep{kleinbaum2012survival}. Cox
+proportional hazards models are used more often than AFT models, causing
+analyses to be based on hazard ratios and relative risks rather than on
+survival curves and absolute risks \citep{hanley2009}. The
+identification of an appropriate distribution for the baseline hazard in
+an AFT model may be difficult for common diseases that have many
+interacting risk factors, or a Cox model where the disease pathogenesis
+may change with age, making it difficult to maintain the proportional
+hazards assumption. For example, previous studies of breast cancer
+incidence have discovered time-varying interactions with covariates of
+interest, such as tumor size \citep{coradini2000time}. One approach to
+provide flexibility in the baseline hazard involves using the basis of
+splines on time in our model \citep{royston2002flexible}. However,
+regression-based models are limited in that they require prior knowledge
+about potential time-varying interactions and their quantitative
+effects.
 
-Over longer follow-up, it is challenging to maintain the proportional
-hazards assumption as the risk associated with each covariate may not be
-constant with time. Previous studies related to breast cancer have
-discovered time-varying interactions with covariates of interest, like
-tumor size \citep{coradini2000time} \citep{hilsenbeck1998time}. Another
-limitation of regression-based models is the requirement of prior
-knowledge; the analyst must know potential time-varying interactions and
-explicitly evaluate them.
+Neural networks provide a data-driven approach to approximating
+interaction terms. For example, DeepSurv is a neural network-based
+proportional hazards model that implements the Cox partial
+log-likelihood as a custom loss function \citep{katzman2018DeepSurv},
+resulting in a stepwise absolute risk curve that cannot accommodate
+time-varying interactions. Compared to Cox regression, DeepSurv shows
+better performance on the Study to Understand Prognoses Preferences and
+Risk Treatments (SUPPORT) dataset \citep{knaus1995SUPPORT}. To handle
+non-proportional hazards, a modification to the loss function was
+proposed \citep{faraggi1995neural}. To remove the need for tuning the
+number of layers and nodes, the concept of extreme learning machines has
+been applied to a Cox neural network model \citep{wang2019extreme}.
 
-Data-driven neural network approaches bypass the need for prior
-knowledge of interaction terms. DeepSurv is a semi-parametric
-proportional hazards neural network model that implements the Cox
-partial log-likelihood as a custom loss function
-\citep{katzman2018DeepSurv}. This results in a stepwise absolute risk
-curve that cannot accommodate time-varying interactions. Compared to Cox
-regression, DeepSurv has shown a small improvement in performance using
-a single layer neural network on the Study to Understand Prognoses
-Preferences and Risk Treatments (SUPPORT) dataset
-\citep{knaus1995SUPPORT}. Farraggi et al.~suggested a modification to
-the loss function required for a Cox neural network that can model
-non-proportional hazards \citep{faraggi1995neural}. Wang et al.~took the
-concept of extreme learning machines and applied them in a survival
-context, removing the need for tuning the number of layers and nodes
-\citep{wang2019extreme}. A Cox-based neural network combined with Local
-Interpretable Model-Agnostic Explanations (LIME)
-\citep{ribeiro2016model} interprets the association of each Single
-Nucleotide Polymorphism (SNP) \citep{sun2020genome} to a phenotype.
+As an alternative method that assumes a baseline hazard distribution,
+DeepHit specifies each survival time of interest in the model and
+directly estimates survival curves, rather than deriving a hazard
+function \citep{lee2018DeepHit}. It assumes an inverse Gaussian
+distribution as the baseline hazard and it outperforms DeepSurv on the
+Molecular Taxonomy of Breast Cancer International Consortium (METABRIC)
+dataset \citep{curtis2012genomic}.
 
-DeepHit requires each survival time of interest to be specified in the
-model and directly estimates survival curves, rather than deriving a
-hazard function \citep{lee2018DeepHit}. It assumes an inverse Gaussian
-distribution as the baseline hazard and its concordance index (c-index)
-score outperforms DeepSurv \citep{katzman2018DeepSurv} on the Molecular
-Taxonomy of Breast Cancer International Consortium (METABRIC) dataset
-\citep{curtis2012genomic}.
-
-Ripley et al.~presented seven different distributions and modeling
-paradigms based on neural networks \citep{ripley2004non}. They provided
-the methodology behind parametric log-logistic and log-normal neural
-network models. Like DeepHit, these models are restricted to one of the
-seven baseline hazard distributions, reducing the flexibility of the
-average survival prediction.
-
-Deep Survival Machines (DSM) provides a flexible, parametric survival
-model using neural networks with a mixture of distributions as the
-baseline hazard \citep{dsmPaper}. DSM allows the user to specify a
-mixture of distributions for a flexible baseline hazard. On both the
+Providing further flexibility, Deep Survival Machines (DSM) is a
+parametric survival model using neural networks with a mixture of
+distributions as the baseline hazard \citep{dsmPaper}. On both the
 SUPPORT and METABRIC datasets, DSM outperforms DeepSurv and DeepHit
@@
 
-These methods need custom loss and activation functions that are
-difficult to implement in practice. We note the complexity of these
-alternative neural network approaches; DeepSurv, DeepHit, and DSM all
-require custom loss functions \citep{katzman2018DeepSurv}
-\citep{lee2018DeepHit} \citep{dsmPaper}, while DSM requires user
-implementations for distributions beyond Log-Normal or Weibull and a
-two-phase learning process \citep{dsmPaper}. Regression-based approaches
-require prior specification of all interaction terms, which makes it
-challenging to model covariate effects that change over time. In this
-article, we propose Case-Base Neural Networks (CBNN) as a method that
-models time-varying interactions and a flexible baseline hazard using
-commonly available neural network components. Our approach to modeling
-the full hazard uses case-base sampling \citep{hanley2009}, which allows
-probabilistic models to predict survival outcomes. Following the
-case-base framework, we can use transformations of time as a feature
-(covariate) to specify different baseline hazards. For example, by
-including splines of time as covariates, we can approximate the
-Royston-Parmar flexible baseline hazard model
-\citep{royston2002flexible}. An alternative approach to Royston-Parmar
-is case-base with logistic regression (CBLR), which provides some
-flexibility in modeling, but still requires user specification of
-time-varying interactions and baseline hazard, while CBNN can model both
-without user specification; only a sufficiently deep network is
-required.
+We note that these alternative neural network approaches all require
+custom loss functions \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}
+\citep{dsmPaper}. DeepHit introduces a hyperparameter weighing its two
+loss functions (negative log-likelihood and ranking losses) while DSM
+requires a two-phase learning process and user implementations for
+distributions beyond Log-Normal or Weibull \citep{lee2018DeepHit}
+\citep{dsmPaper}. Regression-based approaches require prior
+specification of all interaction terms, which makes it challenging to
+model covariate effects that change over time. The current neural
+network models provide flexibility at the cost of opacity, while
+regression models provide clarity at the cost of flexibility.
+
+In this article, we propose Case-Base Neural Networks (CBNN) as a method
+that models time-varying interactions and a flexible baseline hazard
+using commonly available neural network components. Our approach to
+modeling the full hazard uses case-base sampling \citep{hanley2009}.
+This sampling technique allows probabilistic models to predict survival
+outcomes. As part of the case-base framework, we use transformations of
+time as a feature (covariate) to specify different baseline hazards. For
+example, by including splines of time as covariates, we can approximate
+the Royston-Parmar flexible baseline hazard model
+\citep{royston2002flexible}, however, this still requires explicit use
+of time-varying interactions. CBNN can model both without extra tuning
+parameters.
 
@@
 explicitly incorporating time as a feature while adjusting for the
-sampling bias. Finally, we report on the software versions used. A
-package is available for use at {[}GITHUBLINK{]}. The entire code base
-to generate the analyses in this paper is available at {[}GITHUB
-LINK{]}.
+sampling bias. Finally, we report on the software versions used. An R
+package is available for use at
+\url{https://github.com/Jesse-Islam/cbnn}. The entire code base to
+reproduce the figures and empirical results in this paper is available
+at \url{https://github.com/Jesse-Islam/cbnnManuscript}.
 
@@
 \(\theta\) is the set of parameters learned by the neural network and
-\(\mathrm{sigmoid}(x)=\frac{1}{1+e^{-x}}\). By allowing a neural network
-to approximate a higher-order polynomial of time, the baseline hazard
-specification is now data-driven, where user-defined hyperparameters
-such as regularization, number of nodes, and layers control the
-flexibility of the model. We provide a detailed description of the
-choices we made in the next sub-section.
+\(\mathrm{sigmoid}(x)=\frac{1}{1+e^{-x}}\). By approximating a
+higher-order polynomial of time using a neural network, the baseline
+hazard specification is now data-driven, where user-defined
+hyperparameters such as regularization, number of layers and nodes
+control the flexibility of the fukk hazard function. We provide a
+detailed description of the choices we made in the next sub-section.
 
@@
 
-Neural networks provide flexibility when defining the architecture and
-optimization parameters of our models. These hyperparameter choices can
-affect the estimated parameters and were chosen during a set of initial
-simulations to determine if CBNN can learn complex interactions in
-practice. We set a max number of epochs to be \(2000\), batch size as
-\(512\), learning rate as \(10^{-3}\), decay as \(10^{-7}\), patience as
-\(10\) epochs, \(\{50,50,25,25\}\) nodes in each hidden layer with
-\(50\%\) dropout at each layer, a validation split of \(20\%\), a
-testing split of \(10%
-\), a minimum delta loss on the validation set of \(10^{-7}\) over 10
-epochs, and Adam \citep{gulli2017} as our optimizer. These choices may
-permit the model to approximate higher-order interactions while
-preventing over-fitting \citep{srivastava2014dropout}. We fix a
+Neural networks are flexible when defining the architecture and
+optimization parameters. These hyperparameter decisions can affect the
+estimated parameters and were chosen during a set of initial simulations
+to determine if CBNN can learn complex interactions in practice. We set
+a max number of epochs to be \(2000\), batch size as \(512\), learning
+rate as \(10^{-3}\), decay as \(10^{-7}\), patience as \(10\) epochs,
+\(\{50,50,25,25\}\) nodes in each hidden layer with \(50\%\) dropout at
+each layer, a minimum delta loss on the validation set of \(10^{-7}\)
+over 10 epochs and Adam \citep{gulli2017} as our optimizer. These
+choices may permit the model to approximate higher-order interactions
+while preventing over-fitting \citep{srivastava2014dropout}. We fix a
 train-validation-test split that allows us to update the weights with a
 subset of the data (training), assess performance at each epoch
-(validation) and gauge performance of the final model (test). We select
-the best weights after training based on validation loss
-\citep{gulli2017}.
+(validation) and gauge performance of the final model (test) for each
+method. We select the best weights after training based on validation
+loss \citep{gulli2017}.
 
@@
 \citep{cbpkg}, the DeepSurv model using the \textbf{survivalmodels}
-package \citep{survmods}, the DSM model using
-\textbf{DeepSurvivalMachines} \citep{dsmPaper} and the DeepHit model
-using \textbf{pycox} \citep{lee2018DeepHit}. We made the components of
-CBNN using the \textbf{keras} \citep{keras} package and the
-\textbf{casebase} package for the sampling step. The \textbf{simsurv}
-package \citep{simsurv} is used to simulate our simulation studies,
-while \textbf{flexsurv} \citep{flexsurv} is used to fit a flexible
-baseline hazard using splines for our complex simulation. We use the
-implementation of c-ipcw from the python package \textbf{sksurv}
-\citep{sksurv}. The \textbf{riskRegression} package
-\citep{riskRegression} is used to get the Brier score and IPA metrics.
-We changed the \textbf{riskRegression} package to be used with any user
-supplied risk function \(F\). To ensure that both R and Python-based
-models are running in unison on the same data through our simulations
-and bootstrap, we use the \textbf{reticulate} package
+package \citep{survmods}, the DeepHit model using \textbf{pycox}
+\citep{lee2018DeepHit} and the DSM model using
+\textbf{DeepSurvivalMachines} \citep{dsmPaper}. We made the components
+of CBNN using the \textbf{casebase} package for the sampling step and
+the \textbf{keras} \citep{keras} package for our neural network
+architecture. The \textbf{simsurv} package \citep{simsurv} is used for
+our simulation studies, while \textbf{flexsurv} \citep{flexsurv} is used
+to fit a flexible baseline hazard using splines for our complex
+simulation. We use the implementation of \(C_{IPCW}\) from the python
+package \textbf{sksurv} \citep{sksurv}. The \textbf{riskRegression}
+package \citep{riskRegression} is used to get the Index of Prediction
+Accuracy (IPA metric). Both metrics are described in detail in the
+following section. We modify the \textbf{riskRegression} package to be
+used with any user supplied risk function \(F\). To ensure that both R
+and Python-based models are running in unison on the same data through
+our simulations and bootstrap, we use the \textbf{reticulate} package
 \citep{reticulate}.
@@
 
-Our method, CBNN, models survival outcomes by using neural networks on
-case-base sampled data. While estimating the hazard function, we
-incorporate follow-up-time as a feature, providing a data-driven
-estimation of a flexible baseline hazard and time-varying interactions.
-Based on our complex simulation results (Figure \ref{fig:megaPlot} C, D
-and Table \ref{tab:megaTable} B), CBNN outperforms the competitors when
-time-varying interactions and a complex baseline hazard are present.
+CBNN models survival outcomes by using neural networks on case-base
+sampled data. We incorporate follow-up time as a feature, providing a
+data-driven estimate of a flexible baseline hazard and time-varying
+interactions in our hazard function. The three competing neural network
+models we evaluated cannot model time-varying interactions by design
+\citep{dsmPaper} \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}. DSM
+requires a mixture component distributions for a flexible baseline
+hazard to be fit \citep{dsmPaper}. As our goal is to fix the design and
+compare performance, we did not change any shared hyperparameters. With
+our choice of shared hyperparameters and model design, DSM did not
+converge in the complex simulation and SUPPORT case study. Despite this
+limitation, we include this method when it did function as it can fit
+flexible baseline hazards. Compared to CBNN, the remaining two models
+also have limitations. DeepSurv is a proportional hazards model and does
+not estimate the baseline hazard \citep{katzman2018DeepSurv}. DeepHit
+requires an alpha hyperparameter, is restricted to a single distribution
+for the baseline hazard and models the survival function directly
+\citep{lee2018DeepHit}. The alternative neural network methods match on
+time, while CBNN models time directly.
 
-Each competing neural network model we evaluated has limitations or
-model-specific parameters. DeepSurv is a proportional hazards model and
-does not estimate the baseline hazard \citep{katzman2018DeepSurv}.
-DeepHit requires an alpha hyperparameter, is restricted to a single
-distribution for the baseline hazard, and models the survival function
-directly \citep{lee2018DeepHit}. DSM requires the user to specify the
-component distributions in the mixture for a flexible baseline hazard to
-be fit \citep{dsmPaper}. These alternative neural network approaches
-match on time, while CBNN models time directly.
+To assess performance among these models, we use both IPA and
+\(C_{IPCW}\) metrics. Concordance-based measures are commonly used in
+survival analysis to compare models and we opt to keep them in our
+analyses. However, \(C_{IPCW}\) is a non-proper metric and may cause
+misspecified models to appear better than they should
+\citep{cindexfails2019}. Therefore, we contextualize our \(C_{IPCW}\)
+results in relation to IPA, a proper scoring rule. The model rankings
+between \(C_{IPCW}\) and IPA differed for both the complex simulation
+and METABRIC application. Wider \(C_{IPCW}\) confidence intervals for
+DeepHit and DeepSurv show potential misspecification of the models
+(Figure \ref{fig:megaPlot} D, H).
 
-In the simple simulation, all neural network-based approaches resulted
-in a loss of performance based on IPA, while the regression-based
-approaches did not. We attribute this to potential over-parameterization
-in the neural network models. The wide confidence intervals suggest
-over-fitting may be the reason for the loss in performance, even with
-dropout. Both DeepSurv and DSM are affected, while DeepHit and CBNN are
-less so.
+With this interpretation of \(C_{IPCW}\) in mind, we assess two
+simulations with minimal noise and two case studies in real scenarios.
+The simple simulation demonstrates potential pitfalls associated with
+neural network models, particularly overparameterization. All neural
+network-based approaches performed worse than null KM model (IPA score),
+while the regression-based performed better. We attribute this to
+potential overparameterization in the neural network models, as the wide
+confidence intervals suggest over-fitting, even with dropout. Both
+DeepSurv and DSM are affected, while DeepHit and CBNN are less so. This
+is a limitation to our strategy for evaluating the methods, which uses a
+fixed study design across all assessments.
 
-In the complex simulation, CBNN has a distinct advantage over the other
-methods while examining the IPA score. The regression models do show
-improvement over the null model, while the competing neural network
-approaches do not. The time-varying interactions and complex baseline
-hazard may have caused a drop in performance for the competing neural
-network methods, compared to the regression models without interactions.
+From this baseline benchmark, we move on to a complex simulation that
+requires a method that can learn both time-varying interactions and have
+a flexible baseline hazard. Here, CBNN demonstrates a distinct advantage
+over all other methods. The regression models show improved performance
+over the null KM model, while the competing neural network models
+performed worse. Based on our complex simulation results (Figure
+\ref{fig:megaPlot} C, D and Table \ref{tab:megaTable} B), CBNN
+outperforms the competitors when time-varying interactions and a complex
+baseline hazard are present. This simulation shows how CBNN can perform
+under ideal conditions, while the following two analyses on real data
+serve to assess its performance in realistic conditions.
 
-In the SUPPORT application, the overall performance considering IPA is
-near identical. The regression models outperform CBNN, which performs
-better than the competing neural network models. As the confidence
-intervals are tight, it is unlikely that the loss in performance is
-because of over-fitting. Flexibility in both interaction modeling and
-baseline hazard is not particularly helpful with the SUPPORT dataset.
+In the SUPPORT and METABRIC case studies, flexibility in both
+interaction modeling and baseline hazard did not improve CBNNs relative
+performance, suggesting that this flexibility does not aid prediction in
+either case study. From a biological perspective, the 30-year follow-up
+time in the METABRIC study may contain competing causes of death. The
+causes at the start of the study may not match the causes towards the
+end, potentially explaining the drop in performance as we reach later
+survival times with fewer individuals. The baseline hazard is unlikely
+to cause this drop as DSM is the most flexible competing model in our
+comparison. Over-fitting is also unlikely given the tight confidence
+intervals. Further research is required to determine the cause of the
+drop in performance seen in the METABRIC case study. In both case
+studies, CBNN outperforms the competing neural network methods.
 
-In the METABRIC application IPA results, DSM and DeepSurv had a steep
-drop in performance. As DeepSurv does not depend on the baseline hazard
-during the fitting process and DSM is the most flexible competing model
-in our comparison, the baseline hazard is unlikely to cause this drop.
-With tight confidence intervals, over-fitting is unlikely as well.
-Further research would be required to determine the cause of this drop
-in performance. We do not see the same drop in CBNN demonstrating that,
-in the METABRIC application, CBNN outperforms DeepSurv and DSM.
-Considering IPA score, regression models consistently outperform the
-neural network models. From a biological perspective, the 30-year
-follow-up time in the METABRIC study may contain competing causes of
-death. The signal from the start of the study may not match the signal
-towards the end, potentially explaining the drop in performance as we
-reach later survival times with fewer individuals.
+Taken together, CBNN outperforms all competitors in the complex
+simulation, demonstrating its value in survival settings that may
+involve time-varying interactions and a complex baseline hazard. Once we
+perform case-base sampling and adjust for the sampling bias, we can use
+a sigmoid activation function to predict our hazard function. Our
+approach simplifies the incorporation of censored individuals, allowing
+survival outcomes to be treated as binary ones. Forgoing the requirement
+of custom loss functions, CBNN only requires the use of standard
+components in machine learning libraries (specifically, the add layer to
+adjust for sampling bias and the sigmoid activation function). Due to
+the simplicity in its implementation and by extension user experience,
+CBNN is both a user-friendly approach to data-driven survival analysis
+and is easily extendable to any feed-forward neural network framework.
 
-Concordance-based measures are commonly used in survival analysis to
-compare models and we opt to keep them in our analyses. However, c-ipcw
-is a non-proper metric and may cause misspecified models to appear
-better than they should \citep{cindexfails2019}. The model rankings
-between c-ipcw and IPA differed for both the complex simulation and
-METABRIC application. In the complex simulation, the c-ipcw of a
-correctly specified model (Optimal) underperforms CBNN and CBLR; models
-without the exact interaction terms specified. We base our METABRIC
-assessment of this disagreement between metrics on the complex
-simulation. The IPA score ranks regression models first, followed by
-DeepHit and CBNN, but the c-ipcw score ranks regression models after
-DeepHit and CBNN. Wide confidence intervals for neural network models in
-c-ipcw show potential misspecification, suggesting that regression
-models perform best in the METABRIC application.
+\hypertarget{acknowledgements}{%
+\section*{Acknowledgements}\label{acknowledgements}}
+\addcontentsline{toc}{section}{Acknowledgements}
 
-During our testing for the complex simulation and SUPPORT application,
-DSM did not converge with our choice of hyper-parameters and model
-design. As the goal is to fix the design and compare performance, we did
-not change any shared hyperparameters. For DSM, a different number and
-combination of component distributions did not help the model converge,
-though we did not conduct an exhaustive test. This may be a limitation
-to their method or early-stage implementation issues. We include this
-method where it did function as it shares a similar goal of a flexible
-baseline hazard.
-
-If the user suspects that there may be unknown time-varying interactions
-and a complex baseline hazard, CBNN should strongly be considered. Once
-we perform case-base sampling and adjust for the sampling bias, we can
-use a sigmoid activation function to predict our hazard function. The
-user can approach their problem as they would for probabilistic modeling
-while interpreting their prediction in a survival context. In terms of
-performance, we saw that CBNN consistently performs better than the
-other neural network-based approaches over follow-up time. This may be
-because of the hyperparameters chosen, as they differ from each of the
-respective methods papers. CBNN only requires the use of standard
-components in machine learning libraries (add layer and sigmoid
-activation function). Due to the simplicity in its implementation and by
-extension user experience, we expect CBNN to be both a user-friendly
-approach to data-driven survival analysis and easily extendable to any
-feed-forward neural network frameworks.
-
-Though CBNN can model time-varying interactions and a flexible baseline
-hazard, all models require substantial amounts of data to learn real,
-complex relationships effectively. Larger datasets will appear with
-time, but the cost of generating longitudinal datasets remains. A viable
-alternative strategy is the use of transfer learning, where independent
-and identically distributed data can function as a preliminary model to
-learn related features in a parallel dataset, initializing weights for
-our survival task \citep{bozinovski2020reminder}.
-
-This paper describes CBNN in the single event setting. A next step would
-be to extend this approach to competing risks. A promising use case
-would be high-dimensional data with complex correlation structures, such
-as imaging data.
+This work is supported through 5-UM1-DK078616‐14 and R01-HL151855-01
+grants.
 

### 2022-06-14 SIMPaper/SIMPaper.Rmd
added acknowledgement

@@
 
-This work is supported through 5-UM1-DK078616‐14 and R01-HL151855-01 grants.
+This work was supported by subcontracts from UM1DK078616 and R01HL151855 to RS.
+We would like to thank Dr.
+James Meigs, Project Leader of these awards, for his support and helpful discussions.
+We would also like to Thank Dr.
+James Hanley for his support and discussions while extending the case-base methodology.

### 2022-06-14 SIMPaper/SIMPaper.tex
added acknowledgement

@@
 all neural network competitors, while a second real data application
-shows competitive performance. This study highlights the benefit of
-combining case-base sampling with deep learning to provide a simple and
-flexible modeling framework for data-driven, time-varying interaction
-modeling of survival outcomes. An R package is available at
+shows competitive performance. We highlight the benefit of combining
+case-base sampling with deep learning to provide a simple and flexible
+modeling framework for data-driven, time-varying interaction modeling of
+survival outcomes. An R package is available at
 \url{https://github.com/Jesse-Islam/cbnn}.}
@@
 \(\gamma_{1}=3.9, \gamma_{2}=3, \gamma_{3}=-0.43, \gamma_{4}=1.33,\gamma_{5}=-0.86, \beta_{{1}}=1, \beta_{{2}}=1, \beta_{{3}}=1, \tau_{1}=10, \tau_{2}=2, \tau_{3}=2\)
-and \(\psi\) are basis splines. To generate the \(gamma\) coefficients,
-an initial cubic splines model with three knots is fit using the
-\emph{flexsurvspline} function with an intercept-only model on the data,
-which returns the coefficients (\(\gamma\)) we use in our model. Note
-that we fix these values for the analysis. The \(\beta\) coefficients
-represent direct effects, \(\tau_{2}\) and \({\tau_3}\) represent
-interactions and \(\tau_{1}\) is a time-varying interaction.
+and \(\psi\) are basis splines. The \(gamma\) coefficients are obtained
+from an intercept-only cubic splines model with three knots using the
+\emph{flexsurvspline} function from the \textbf{flexsurv} package
+\citep{flexsurv}. Note that we fix these values for the analysis. The
+\(\beta\) coefficients represent direct effects, \(\tau_{2}\) and
+\({\tau_3}\) represent interactions and \(\tau_{1}\) is a time-varying
+interaction.
 
@@
 
-This work is supported through 5-UM1-DK078616‐14 and R01-HL151855-01
-grants.
+This work was supported by subcontracts from UM1DK078616 and R01HL151855
+to RS. We would like to thank Dr.~James Meigs, Project Leader of these
+awards, for his support and helpful discussions. We would also like to
+Thank Dr.~James Hanley for his support and discussions while extending
+the case-base methodology.
 

### 2022-06-14 SIMPaper/SIMPaper.Rmd
added data availability statement

@@
 received: 2022-04-29
-revised: 2022-06-01
-accepted: 2022-07-01
-abstract: "Neural network-based survival methods have been developed to model data-driven covariate interactions.
+abstract: "Neural network-based survival methods can model data-driven covariate interactions.
 While these methods have led to better predictive performance than regression-based approaches, they cannot model both time-varying interactions and complex baseline hazards.
 To address this, we propose Case-Base Neural Networks (CBNN) as a new approach that combines the case-base sampling framework with flexible architectures.
-Our method naturally accounts for censoring, does not require additional hyperparameters and uses binary cross-entropy as the loss function.
+Our method naturally accounts for censoring and does not require method specific hyperparameters.
 Using a novel sampling scheme and data augmentation, we incorporate time directly into a feed-forward neural network.
@@
 
+# Data and code availability statement {-}
+
+The pre-processed data for the SUPPORT case study can be found at [https://github.com/jaredleekatzman/DeepSurv/tree/master/experiments/data/support](https://github.com/jaredleekatzman/DeepSurv/tree/master/experiments/data/support).
+The pre-processed data for the METABRIC case study can be found at [https://github.com/jaredleekatzman/DeepSurv/tree/master/experiments/data/metabric](https://github.com/jaredleekatzman/DeepSurv/tree/master/experiments/data/metabric).
+The data is accessed using [https://github.com/havakv/pycox](https://github.com/havakv/pycox).
+The code for this manuscript and its analyses can be found at [https://github.com/Jesse-Islam/cbnnManuscript](https://github.com/Jesse-Islam/cbnnManuscript).
+The software package making CBNN easier to use can be found at [https://github.com/Jesse-Islam/cbnn](https://github.com/Jesse-Islam/cbnn).
+
 # Acknowledgements {-}
@@
 James Meigs, Project Leader of these awards, for his support and helpful discussions.
+The work was also supported as part of the Congressionally Directed Medical Research Programs (CDMRP) award W81XWH-17-1-0347.
 We would also like to Thank Dr.

### 2022-06-14 SIMPaper/SIMPaper.tex
added data availability statement

@@
 \received{2022-04-29}
-
-\revised{2022-06-01}
-
-\accepted{2022-07-01}
 
@@
 
-\abstract{Neural network-based survival methods have been developed to
-model data-driven covariate interactions. While these methods have led
-to better predictive performance than regression-based approaches, they
-cannot model both time-varying interactions and complex baseline
-hazards. To address this, we propose Case-Base Neural Networks (CBNN) as
-a new approach that combines the case-base sampling framework with
-flexible architectures. Our method naturally accounts for censoring,
-does not require additional hyperparameters and uses binary
-cross-entropy as the loss function. Using a novel sampling scheme and
-data augmentation, we incorporate time directly into a feed-forward
+\abstract{Neural network-based survival methods can model data-driven
+covariate interactions. While these methods have led to better
+predictive performance than regression-based approaches, they cannot
+model both time-varying interactions and complex baseline hazards. To
+address this, we propose Case-Base Neural Networks (CBNN) as a new
+approach that combines the case-base sampling framework with flexible
+architectures. Our method naturally accounts for censoring and does not
+require method specific hyperparameters. Using a novel sampling scheme
+and data augmentation, we incorporate time directly into a feed-forward
 neural network. CBNN predicts the probability of an event occurring at a
@@
 
+\hypertarget{data-and-code-availability-statement}{%
+\section*{Data and code availability
+statement}\label{data-and-code-availability-statement}}
+\addcontentsline{toc}{section}{Data and code availability statement}
+
+The pre-processed data for the SUPPORT case study can be found at
+\url{https://github.com/jaredleekatzman/DeepSurv/tree/master/experiments/data/support}.
+The pre-processed data for the METABRIC case study can be found at
+\url{https://github.com/jaredleekatzman/DeepSurv/tree/master/experiments/data/metabric}.
+The data is accessed using \url{https://github.com/havakv/pycox}. The
+code for this manuscript and its analyses can be found at
+\url{https://github.com/Jesse-Islam/cbnnManuscript}. The software
+package making CBNN easier to use can be found at
+\url{https://github.com/Jesse-Islam/cbnn}.
+
 \hypertarget{acknowledgements}{%
@@
 to RS. We would like to thank Dr.~James Meigs, Project Leader of these
-awards, for his support and helpful discussions. We would also like to
-Thank Dr.~James Hanley for his support and discussions while extending
-the case-base methodology.
+awards, for his support and helpful discussions. The work was also
+supported as part of the Congressionally Directed Medical Research
+Programs (CDMRP) award W81XWH-17-1-0347. We would also like to Thank
+Dr.~James Hanley for his support and discussions while extending the
+case-base methodology.
 

### 2022-11-15 SIMPaper/SIMPaper.Rmd
addedStatTemplate

@@
 Therefore, we have the following equality \footnote{We are abusing notation here, conflating hazards with probabilities.
-For a rigorous treatment, see Saarela \& Hanley (2015) section 3 \cite{saarela2015} .}:
+For a rigorous treatment, see Saarela \& Hanley (2015) section 3 \cite{saarela2015}.}:
 \begin{align}\label{eqn:main}
@@
 \end{align}
+
+To provide some intuition for this equation, the hazard is the rate at which the event ($Y_{i}=1$) occurs at a given time, conditional on your survival up to the follow-up-time of interest.
+As we are looking at person-moments, we may retrieve the rate at which an event does not occur ($Y_{i}=0)$ at a given time by taking the number of samples in $b$ and dividing it by the total number of moments in the study base $B$ at the current follow-up-time.
+Though an intuitive approach is used to describe this equation, a complete derivation of this equation can be found at [@saarela2015].
+
 The odds of a person-moment being a part of the *case series* is the ratio of the hazard $h(T_i \mid X_i)$ and the uniform rate $b/B$.
+Sampling techniques are known to result in a loss of information.
+To address this concern, the original case-base methodology paper for logistic regression uses a measure for relative information when comparing two averages, based on the size of our base-series and case-series.
+If we let the size of $b=100c$, then the variances and covariances are expected to be proportional to $\frac{1}{c}+\frac{1}{100c}$ rather than $\frac{1}{c}+\frac{1}{\inf}$, about 1 percent larger.
+Concerns were raised about dependency due to multiple person-moments coming from each individual, the information loss was studied in Saarela 2015, where they found a ratio of 100:1, b:c, give an efficiency comparable to Mantel-Haenszel and conditional logistic regression methods (Saarela 2015).
 Using \eqref{eqn:main}, we can see how the log-hazard function can be estimated from the log-odds arising from case-base sampling: \begin{align}\label{eqn:offset}
@@
 
-To estimate the correct hazard function, we adjusting for the bias introduced when sampling a fraction of the study base $B$ by adding the offset term $\log\left(\frac{b}{B} \right)$ as in \eqref{eqn:offset}.
+Sampling techniques are known to result in a loss of information.
+To address this concern, the original case-base methodology paper uses a measure for relative information when comparing two averages, based on the size of our base-series and case-series.
+If we let the size of $b=100c$, then the variances and covariances are expected to be proportional to $\frac{1}{c}+\frac{1}{100c}$ rather than $\frac{1}{c}+\frac{1}{\inf}$, about 1 percent larger.
+If there is a concern with a 1 percent inflation of these estimates, a larger ratio than 100 may be used.
+As there are concerns about dependency due multiple person-moments coming from each individual, the information loss was studied in Saarela 2015 ([@saarela2015]), where they found a ratio of 100:1 base-series to case series gives an efficiency comparable to Mantel-Haenszel and conditional logistic regression methods [@saarela2015].
+
+To estimate the correct hazard function, we adjusting for the bias introduced when sampling a fraction of the study base $B$ by adding an offset term $\log\left(\frac{B}{b} \right)$ \eqref{eqn:offset}.
 Next, we propose using neural networks to model the odds.
@@
 After case-base sampling, we pass all features, including time, into any user-defined feed-forward component, to which an offset term is added, then passed through a sigmoid activation function (Figure \ref{fig:NNarch}).
-We use the sigmoid function as its inverse is the odds, which we can use to calculate the hazard.
+We use the sigmoid activation function as it is the inverse of the log-odds and commonly available in popular neural network packages (tensorflow, keras, pytorch, etc) [REF?].
+which we can use to calculate the hazard.
 The general form for the neural network using CBNN is:
@@
 where $T$ is a random variable representing the event time, $X$ is the random variable for a covariate profile, $f_{\theta}(X, T)$ represents any feed-forward neural network architecture, $\log\left(\frac{B}{b}\right)$ is bias term set by case-base sampling, $\theta$ is the set of parameters learned by the neural network and $\sigmoid(x)=\frac{1}{1+e^{-x}}$.
-By approximating a higher-order polynomial of time using a neural network, the baseline hazard specification is now data-driven, where user-defined hyperparameters such as regularization, number of layers and nodes control the flexibility of the fukk hazard function.
+By approximating a higher-order polynomial of time using a neural network, the baseline hazard specification is now data-driven, where user-defined hyperparameters such as regularization, number of layers and nodes control the flexibility of the hazard function.
 We provide a detailed description of the choices we made in the next sub-section.
@@
 We use two metrics to assess the performance of the different methods of interest: 1) (IPA) [@kattan2018index] and 2) inverse probability censoring weights-adjusted concordance index ($C_{IPCW}$) [@uno2011], which we define below.
-Both these time-dependent metrics provide transparency as to when in follow-up time each model may perform better than the others.
+We may be interested in an absolute risk a specific number of years in the future.
+For the sake of comparison, we plot each metric over time $t$ so that the performance is transparent.
+As we go further along follow-up-time, we expect the performance of all models to drop as there are fewer individuals left in the study by then.
+For $C_{IPCW}$, the score demonstrates performance up to time $t$  [@uno2011], while IPA is performance at time $t$.
 
@@
 Each model-specific metric in each study shows a 95-percent confidence interval over 100 iterations.
-The models of interest are case-base with logistic regression (CBLR), Case-Base Neural Networks (CBNN), Cox, DeepHit, DeepSurv, Deep Survival Machines (DSM), Optimal (a CBLR model with the exact interaction terms and baseline hazard specified) and Kaplan-Meier (to serve as a baseline, predicting the average for all individuals)."}
+The models of interest are case-base with logistic regression (CBLR), Case-Base Neural Networks (CBNN), Cox, DeepHit, DeepSurv, Deep Survival Machines (DSM), Optimal (a CBLR model with the exact interaction terms and baseline hazard specified) and Kaplan-Meier (to serve as a baseline, predicting the average for all individuals).
+Note than Cox and CBLR are superimposed."}
 #readRDS(here::here("figures", "megaPlot.rds"))
@@
 We used the processed dataset made available through DeepSurv [@katzman2018DeepSurv], which includes 1980 patients of which 57.72% die due to breast cancer within a median 10 years of follow-up [@katzman2018DeepSurv].
-There are 11 covariates in total: 4 RNA-Seq gene expressions (MKI67, EGFR, PGR and ERBB2) and 5 clinical features (hormone treatment indicator, radiotherapy indicator, chemotherapy indicator, ER-positive indicator and age at diagnosis) [@katzman2018DeepSurv].
+There are 10 covariates in total: time, 4 RNA-Seq gene expressions (MKI67, EGFR, PGR and ERBB2) and 5 clinical features (hormone treatment indicator, radiotherapy indicator, chemotherapy indicator, ER-positive indicator and age at diagnosis) [@katzman2018DeepSurv].
 
@@
 
+The way neural networks are used in this paper are not directly interpretable.
+This is by design, as we only wish to compare to competitors based on predictive ability.
+All feed-forward neural network models can make use of different architectures to improve performance or interpretability.
+While a deep neural network as used in this paper is not directly interpretable, techniques such as Local Interpretable Model-Agnostic Explanations (LIME) [@ribeiro2016model] may be used to interpret the relevance of each feature to a given prediction.
+An example of architectural interpretability is a convolutional neural network for images, which provide positional information about the relevant pixels in an image [@keras].
+Note that neither of these are specific to CBNN.
+As CBNN is concerned with learning time-varying interactions, a concern may be interpreting which time-varying interactions have been learned.
+Other secondary methods after the fitting process may be used to retrieve said information https://ui.adsabs.harvard.edu/abs/2017arXiv170504977T/abstract.
+
 Taken together, CBNN outperforms all competitors in the complex simulation, demonstrating its value in survival settings that may involve time-varying interactions and a complex baseline hazard.
@@
 
-This work was supported by subcontracts from UM1DK078616 and R01HL151855 to RS.
+This work was supported by subcontracts from UM1DK078616 and R01HL151855 to Dr.
+Rob Sladek.
 We would like to thank Dr.
-James Meigs, Project Leader of these awards, for his support and helpful discussions.
+James Meigs, the project leader of these awards, for his support and helpful discussions.
 The work was also supported as part of the Congressionally Directed Medical Research Programs (CDMRP) award W81XWH-17-1-0347.
-We would also like to Thank Dr.
+We would also like to thank Dr.
 James Hanley for his support and discussions while extending the case-base methodology.

### 2022-11-15 SIMPaper/SIMPaper.tex
addedStatTemplate

@@
 hyperparameters such as regularization, number of layers and nodes
-control the flexibility of the fukk hazard function. We provide a
-detailed description of the choices we made in the next sub-section.
+control the flexibility of the hazard function. We provide a detailed
+description of the choices we made in the next sub-section.
 
@@
 patients of which 57.72\% die due to breast cancer within a median 10
-years of follow-up \citep{katzman2018DeepSurv}. There are 11 covariates
-in total: 4 RNA-Seq gene expressions (MKI67, EGFR, PGR and ERBB2) and 5
-clinical features (hormone treatment indicator, radiotherapy indicator,
-chemotherapy indicator, ER-positive indicator and age at diagnosis)
-\citep{katzman2018DeepSurv}.
+years of follow-up \citep{katzman2018DeepSurv}. There are 10 covariates
+in total: time, 4 RNA-Seq gene expressions (MKI67, EGFR, PGR and ERBB2)
+and 5 clinical features (hormone treatment indicator, radiotherapy
+indicator, chemotherapy indicator, ER-positive indicator and age at
+diagnosis) \citep{katzman2018DeepSurv}.
 
@@
 This work was supported by subcontracts from UM1DK078616 and R01HL151855
-to RS. We would like to thank Dr.~James Meigs, Project Leader of these
-awards, for his support and helpful discussions. The work was also
-supported as part of the Congressionally Directed Medical Research
-Programs (CDMRP) award W81XWH-17-1-0347. We would also like to Thank
-Dr.~James Hanley for his support and discussions while extending the
-case-base methodology.
+to Dr.~Rob Sladek. We would like to thank Dr.~James Meigs, the project
+leader of these awards, for his support and helpful discussions. The
+work was also supported as part of the Congressionally Directed Medical
+Research Programs (CDMRP) award W81XWH-17-1-0347. We would also like to
+thank Dr.~James Hanley for his support and discussions while extending
+the case-base methodology.
 

### 2022-11-15 StatPaper/APA-lato_New/APA/wileyNJD-APA.tex
StatAdded

\documentclass[APA,LATO1COL]{WileyNJD-v2}
\usepackage{lmodern}
\articletype{Article Type}%

\received{26 April 2016}
\revised{6 June 2016}
\accepted{6 June 2016}

\raggedbottom

\begin{document}

\title{This is the sample article title\protect\thanks{This is an example for title footnote.}}

\author[1]{Author One*}

\author[2,3]{Author Two}

\author[3]{Author Three}

\authormark{AUTHOR ONE \textsc{et al}}

\address[1]{\orgdiv{Org Division}, \orgname{Org Name}, \orgaddress{\state{State name}, \country{Country name}}}

\address[2]{\orgdiv{Org Division}, \orgname{Org Name}, \orgaddress{\state{State name}, \country{Country name}}}

\address[3]{\orgdiv{Org Division}, \orgname{Org Name}, \orgaddress{\state{State name}, \country{Country name}}}

\corres{*Corresponding author name, This is sample corresponding address. \email{authorone@gmail.com}}

\presentaddress{This is sample for present address text this is sample for present address text}

\abstract[Summary]{This is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text.}

\keywords{keyword1, keyword2, keyword3, keyword4}

\jnlcitation{\cname{%
\author{Williams K.},
\author{B. Hoskins},
\author{R. Lee},
\author{G. Masato}, and
\author{T. Woollings}} (\cyear{2016}),
\ctitle{A regime analysis of Atlantic winter jet variability applied to evaluate HadGEM3-GC2}, \cjournal{Q.J.R. Meteorol. Soc.}, \cvol{2017;00:1--6}.}

\maketitle

\footnotetext{\textbf{Abbreviations:} ANA, anti-nuclear antibodies; APC, antigen-presenting cells; IRF, interferon regulatory factor}

\section{Sample for first level head}\label{sec1}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit~\citet{Rothermel1997} Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae,
felis. Curabitur dictum gravida mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna. Donec
vehicula augue eu neque. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
Mauris ut leo. Cras viverra metus rhoncus sem. Nulla et lectus vestibulum urna fringilla ultrices. Phasellus eu tellus
sit amet tortor gravida placerat. Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. Praesent eget sem vel
leo ultrices bibendum. Aenean faucibus. Morbi dolor nulla, malesuada eu, pulvinar at, mollis ac, nulla. Curabitur
auctor semper nulla. Donec varius orci eget risus. Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

\begin{eqnarray}
s(nT_{s}) &= &s(t)\times \sum\limits_{n=0}^{N-1} \delta (t-nT_{s}) \xleftrightarrow{\mathrm{DFT}}  S \left(\frac{m}{NT_{s}}\right) \nonumber\\
&= &\frac{1}{N} \sum\limits_{n=0}^{N-1} \sum\limits_{k=-N/2}^{N/2-1} s_{k} e^{\mathrm{j}2\pi k\Delta fnT_{s}} e^{-j\frac{2\pi}{N}mn}
\end{eqnarray}

\section{Sample for another first level head}\label{sec2}

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus viverra
fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa ac
quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo~\cite{Rothermel1998} Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

Example for bibliography citations cite~\citep{Elbaum2002}, cites~\cite{Allen2011,Yoo2007}

Quisque ullamcorper placerat ipsum. Cras nibh~\cite{Yoo2007,Schulz2012} Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
vitae risus porta vehicula.

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.

\begin{figure}[t]
\centerline{\includegraphics[width=342pt,height=9pc,draft]{empty}}
\caption{This is the sample figure caption.\label{fig1}}
\end{figure}

Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.

Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non
odio. Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce
sed justo eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor.
Donec pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin
et quam. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien
turpis, fermentum vel, eleifend faucibus, vehicula eu, lacus.

\begin{figure*}
\centerline{\includegraphics[width=342pt,height=9pc,draft]{empty}}
\caption{This is the sample figure caption.\label{fig2}}
\end{figure*}

\subsection{Example for second level head}

Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit, dictum
in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam vestibulum
fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi. Pellentesque
eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut mi mi, lacinia
sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui. Mauris nibh leo,
facilisis non, adipiscing quis, ultrices a, dui.

Morbi luctus, wisi viverra faucibus pretium, nibh est placerat odio, nec commodo wisi enim eget quam. Quisque
libero justo, consectetuer a, feugiat vitae, porttitor eu, libero. Suspendisse sed mauris vitae elit sollicitudin malesuada.

Maecenas ultricies eros sit amet ante. Ut venenatis velit. Maecenas sed mi eget dui varius euismod. Phasellus aliquet
volutpat odio. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Pellentesque sit
amet pede ac sem eleifend consectetuer. Nullam elementum, urna vel imperdiet sodales, elit ipsum pharetra ligula,
ac pretium ante justo a nulla. Curabitur tristique arcu eu metus. Vestibulum lectus. Proin mauris. Proin eu nunc eu
urna hendrerit faucibus. Aliquam auctor, pede consequat laoreet varius, eros tellus scelerisque quam, pellentesque
hendrerit ipsum dolor sed augue. Nulla nec lacus.

\begin{quote}
This is an example~\citep{Elbaum2002,Blanchard2015} for quote text. This is an example for quote text. This is an example for quote text. This is an example for quote text~\cite{Blanchard2015}. This is an example for quote text. This is an example for quote text. This is an example for quote text. This is an example for quote text. This is an example for quote text. This is an example for quote text~\cite{Blanchard2015}. This is an example for quote text. This is an example for quote text. This is an example for quote text.
\end{quote}

\section{Sample for next first level head}\label{sec3}

\subsection{Example for another second level head}

Suspendisse vitae elit. Aliquam arcu neque, ornare in, ullamcorper quis, commodo eu, libero. Fusce sagittis erat at
erat tristique mollis. Maecenas sapien libero, molestie et, lobortis in, sodales eget, dui. Morbi ultrices rutrum lorem.
Nam elementum ullamcorper leo. Morbi dui. Aliquam sagittis. Nunc placerat. Pellentesque tristique sodales est.
Maecenas imperdiet lacinia velit. Cras non urna. Morbi eros pede, suscipit ac, varius vel, egestas non, eros. Praesent
malesuada, diam id pretium elementum, eros sem dictum tortor, vel consectetuer odio sem sed wisi.

Sed feugiat. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Ut pellentesque
augue sed urna. Vestibulum diam eros, fringilla et, consectetuer eu, nonummy id, sapien. Nullam at lectus. In sagittis
ultrices mauris. Curabitur malesuada erat sit amet massa. Fusce blandit. Aliquam erat volutpat. Aliquam euismod.
Aenean vel lectus. Nunc imperdiet justo nec dolor.

\subsection{Second level head text}

Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.

\subsubsection{Third level head text}

Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra tortor. Sed fringilla
justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue est scelerisque
sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas urna mi, suscipit
in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.

Etiam ac leo a risus tristique nonummy. Donec dignissim tincidunt nulla. Vestibulum rhoncus molestie odio. Sed
lobortis, justo et pretium lobortis, mauris turpis condimentum augue, nec ultricies nibh arcu pretium enim. Nunc
purus neque, placerat id, imperdiet sed, pellentesque nec, nisl. Vestibulum imperdiet neque non sem accumsan laoreet.
In hac habitasse platea dictumst. Etiam condimentum facilisis libero. Suspendisse in elit quis nisl aliquam dapibus.
Pellentesque auctor sapien. Sed egestas sapien nec lectus. Pellentesque vel dui vel neque bibendum viverra. Aliquam
porttitor nisl nec pede. Proin mattis libero vel turpis. Donec rutrum mauris et libero. Proin euismod porta felis.
Nam lobortis, metus quis elementum commodo, nunc lectus elementum mauris, eget vulputate ligula tellus eu neque.
Vivamus eu dolor.

Nulla in ipsum. Praesent eros nulla, congue vitae, euismod ut, commodo a, wisi. Pellentesque habitant morbi
tristique senectus et netus et malesuada fames ac turpis egestas. Aenean nonummy magna non leo. Sed felis erat,
ullamcorper in, dictum non, ultricies ut, lectus. Proin vel arcu a odio lobortis euismod. Vestibulum ante ipsum primis
in faucibus orci luctus et ultrices posuere cubilia Curae; Proin ut est. Aliquam odio. Pellentesque massa turpis, cursus
eu, euismod nec, tempor congue, nulla. Duis viverra gravida mauris. Cras tincidunt. Curabitur eros ligula, varius ut,
pulvinar in, cursus faucibus, augue.

\begin{boxtext}
\section*{Example of Boxtext}%
This is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext.
\end{boxtext}

\paragraph{Fourth level head text}

Sed feugiat. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Ut pellentesque
augue sed urna. Vestibulum diam eros, fringilla et, consectetuer eu, nonummy id, sapien. Nullam at lectus. In sagittis
ultrices mauris. Curabitur malesuada erat sit amet massa. Fusce blandit. Aliquam erat volutpat. Aliquam euismod.
Aenean vel lectus. Nunc imperdiet justo nec dolor.

Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.

\subparagraph{Fifth level head text}

Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra
tortor. Sed fringilla justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue
est scelerisque sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas
urna mi, suscipit in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.

Etiam ac leo a risus tristique nonummy. Donec dignissim tincidunt nulla. Vestibulum rhoncus molestie odio. Sed
lobortis, justo et pretium lobortis, mauris turpis condimentum augue, nec ultricies nibh arcu pretium enim. Nunc
purus neque, placerat id, imperdiet sed, pellentesque nec, nisl. Vestibulum imperdiet neque non sem accumsan laoreet.
In hac habitasse platea dictumst. Etiam condimentum facilisis libero. Suspendisse in elit quis nisl aliquam dapibus.
Pellentesque auctor sapien. Sed egestas sapien nec lectus. Pellentesque vel dui vel neque bibendum viverra. Aliquam
porttitor nisl nec pede. Proin mattis libero vel turpis. Donec rutrum mauris et libero. Proin euismod porta felis.
Nam lobortis, metus quis elementum commodo, nunc lectus elementum mauris, eget vulputate ligula tellus eu neque.
Vivamus eu dolor.

in faucibus orci luctus et ultrices posuere cubilia Curae; Proin ut est. Aliquam odio. Pellentesque massa turpis, cursus
eu, euismod nec, tempor congue, nulla. Duis viverra gravida mauris. Cras tincidunt. Curabitur eros ligula, varius ut,
pulvinar in, cursus faucibus, augue.

Curabitur tellus magna, porttitor a, commodo a, commodo in, tortor. Donec interdum. Praesent scelerisque. Mae-
cenas posuere sodales odio. Vivamus metus lacus, varius quis, imperdiet quis, rhoncus a, turpis. Etiam ligula arcu,
elementum a, venenatis quis, sollicitudin sed, metus. Donec nunc pede, tincidunt in, venenatis vitae, faucibus vel,
nibh. Pellentesque wisi. Nullam malesuada. Morbi ut tellus ut pede tincidunt porta. Lorem ipsum dolor sit amet,
consectetuer adipiscing elit. Etiam congue neque id dolor.

Donec et nisl at wisi luctus bibendum. Nam interdum tellus ac libero. Sed sem justo, laoreet vitae, fringilla at,
adipiscing ut, nibh. Maecenas non sem quis tortor eleifend fermentum. Etiam id tortor ac mauris porta vulputate.
Integer porta neque vitae massa. Maecenas tempus libero a libero posuere dictum. Vestibulum ante ipsum primis in
faucibus orci luctus et ultrices posuere cubilia Curae; Aenean quis mauris sed elit commodo placerat. Class aptent
taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Vivamus rhoncus tincidunt libero.
Etiam elementum pretium justo. Vivamus est. Morbi a tellus eget pede tristique commodo. Nulla nisl. Vestibulum
sed nisl eu sapien cursus rutrum.

Nulla non mauris vitae wisi posuere convallis. Sed eu nulla nec eros scelerisque pharetra. Nullam varius. Etiam
dignissim elementum metus. Vestibulum faucibus, metus sit amet mattis rhoncus, sapien dui laoreet odio, nec ultricies
nibh augue a enim. Fusce in ligula. Quisque at magna et nulla commodo consequat. Proin accumsan imperdiet sem.
Nunc porta. Donec feugiat mi at justo. Phasellus facilisis ipsum quis ante. In ac elit eget ipsum pharetra faucibus.
Maecenas viverra nulla in massa.

Nulla in ipsum. Praesent eros nulla, congue vitae, euismod ut, commodo a, wisi. Pellentesque habitant morbi
tristique senectus et netus et malesuada fames ac turpis egestas. Aenean nonummy magna non leo. Sed felis erat,
ullamcorper in, dictum non, ultricies ut, lectus. Proin vel arcu a odio lobortis euismod. Vestibulum ante ipsum primis

\begin{center}
\begin{table*}[t]%
\caption{This is sample table caption.\label{tab1}}
\centering
\begin{tabular*}{500pt}{@{\extracolsep\fill}lccD{.}{.}{3}c@{\extracolsep\fill}}
\toprule
&\multicolumn{2}{@{}c@{}}{\textbf{Spanned heading\tnote{1}}} & \multicolumn{2}{@{}c@{}}{\textbf{Spanned heading\tnote{2}}} \\\cmidrule{2-3}\cmidrule{4-5}
\textbf{col1 head} & \textbf{col2 head}  & \textbf{col3 head}  & \multicolumn{1}{@{}l@{}}{\textbf{col4 head}}  & \textbf{col5 head}   \\
\midrule
col1 text & col2 text  & col3 text  & 12.34  & col5 text\tnote{1}   \\
col1 text & col2 text  & col3 text  & 1.62  & col5 text\tnote{2}   \\
col1 text & col2 text  & col3 text  & 51.809  & col5 text   \\
\bottomrule
\end{tabular*}
\begin{tablenotes}%%[341pt]
\item Source: Example for table source text.
\item[1] Example for a first table footnote.
\item[2] Example for a second table footnote.
\end{tablenotes}
\end{table*}
\end{center}

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.

\begin{center}
\begin{table}[t]%
\centering
\caption{This is sample table caption.\label{tab2}}%
\begin{tabular*}{500pt}{@{\extracolsep\fill}lcccc@{\extracolsep\fill}}
\toprule
\textbf{col1 head} & \textbf{col2 head}  & \textbf{col3 head}  & \textbf{col4 head}  & \textbf{col5 head} \\
\midrule
col1 text & col2 text  & col3 text  & col4 text  & col5 text\tnote{$\dagger$}   \\
col1 text & col2 text  & col3 text  & col4 text  & col5 text   \\
col1 text & col2 text  & col3 text  & col4 text  & col5 text\tnote{$\ddagger$}   \\
\bottomrule
\end{tabular*}
\begin{tablenotes}
\item Source: Example for table source text.
\item[$\dagger$] Example for a first table footnote.
\item[$\ddagger$] Example for a second table footnote.
\end{tablenotes}
\end{table}
\end{center}

Below is the example~\cite{Rothermel1998,Yoo2007,Schulz2012} for bulleted list. Below is the example for bulleted list. Below is the example for bulleted list. Below is the example for bulleted list. Below is the example for bulleted list. Below is the example for bulleted list\footnote{This is an example for footnote.}:
\begin{itemize}
\item bulleted list entry sample bulleted list entry~\cite{Allen2011} sample list entry text.
\item bulleted list entry sample bulleted list entry. bulleted list entry sample bulleted list entry. bulleted list entry sample bulleted list entry.
\item bulleted list entry sample bulleted list entry~\cite{Ballen2011} bulleted list entry sample bulleted list entry~\citet{Allen2011} sample list entry text.  bulleted list entry sample bulleted list entry.
\item sample list entry text. sample list entry text.
\end{itemize}

Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.

Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non
odio. Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce
sed justo eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor. Donec pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin
et quam. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien
turpis, fermentum vel, eleifend faucibus, vehicula eu, lacus.

Below is the sample for description list. Below is the example for description list. Below is the example for description list. Below is the example for description list. Below is the example for description list. Below is the example for description list:\\[12pt]

\noindent\textbf{Description sample:}

\begin{description}
\item[first entry] description text. description text. description text. description text. description text. description text. description text.
\item[second long entry] description text. description text. description text. description text. description text. description text. description text.
\item[third entry] description text. description text. description text. description text. description text.
\item[fourth entry] description text. description text.
\end{description}

\noindent\textbf{Numbered list items sample:}

\begin{enumerate}[1.]
\item First level numbered list entry. sample numbered list entry.

\item First numbered list entry. sample numbered list entry. Numbered list entry. sample numbered list entry. Numbered list entry. sample numbered list entry.

\begin{enumerate}[a.]
\item Second level alpabetical list entry. Second level alpabetical list entry. Second level alpabetical list entry~\citet{Allen2011} Second level alpabetical list entry.

\item Second level alpabetical list entry. Second level alpabetical list entry~\citet{Schulz2012,Allen2011,Ballen2011}.

\begin{enumerate}[ii.]
\item Third level lowercase roman numeral list entry. Third level lowercase roman numeral list entry. Third level lowercase roman numeral list entry.

\item Third level lowercase roman numeral list entry. Third level lowercase roman numeral list entry~\cite{Yoo2007}.
\end{enumerate}

\item Second level alpabetical list entry. Second level alpabetical list entry~\cite{Elbaum2002}.
\end{enumerate}

\item First level numbered list entry. sample numbered list entry. Numbered list entry. sample numbered list entry. Numbered list entry.

\item Another first level numbered list entry. sample numbered list entry. Numbered list entry. sample numbered list entry. Numbered list entry.
\end{enumerate}

\noindent\textbf{un-numbered list items sample:}

\begin{enumerate}[]
\item Sample unnumberd list text..
\item Sample unnumberd list text.
\item sample unnumberd list text.
\item Sample unnumberd list text.
\end{enumerate}

\section{Examples for enunciations}\label{sec4}

\begin{theorem}[Theorem subhead]\label{thm1}
Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text.
\end{theorem}

Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
vitae risus porta vehicula.

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.

\begin{proposition}
Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text.
\end{proposition}

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
vitae risus porta vehicula.

\begin{definition}[Definition sub head]
Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text.
\end{definition}

Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non
odio. Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce
sed justo eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor.
Donec pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin
et quam. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien
turpis, fermentum vel, eleifend faucibus, vehicula eu, lacus.

Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
Pellentesque eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut
mi mi, lacinia sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui.
Mauris nibh leo, facilisis non, adipiscing quis, ultrices a, dui.

\begin{proof}
Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text.
\end{proof}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

\begin{proof}[Proof of Theorem~\ref{thm1}]
Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text.
\end{proof}

Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.
Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra tortor. Sed fringilla
justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue est scelerisque
sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas urna mi, suscipit
in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.

Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
Pellentesque eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut Curabitur tellus magna, porttitor a, commodo a, commodo in, tortor. Donec interdum. Praesent scelerisque. Mae-
cenas posuere sodales odio. Vivamus metus lacus, varius quis, imperdiet quis, rhoncus a, turpis. Etiam ligula arcu,
elementum a, venenatis quis, sollicitudin sed, metus. Donec nunc pede, tincidunt in, venenatis vitae, faucibus vel,

\begin{sidewaystable}%[h]
\caption{Sideways table caption. For decimal alignment refer column 4 to 9 in tabular* preamble.\label{tab3}}%
\begin{tabular*}{\textheight}{@{\extracolsep\fill}lccD{.}{.}{4}D{.}{.}{4}D{.}{.}{4}D{.}{.}{4}D{.}{.}{4}D{.}{.}{4}@{\extracolsep\fill}}%
\toprule
  & \textbf{col2 head} & \textbf{col3 head} & \multicolumn{1}{c}{\textbf{10}} &\multicolumn{1}{c}{\textbf{20}} &\multicolumn{1}{c}{\textbf{30}} &\multicolumn{1}{c}{\textbf{10}} &\multicolumn{1}{c}{\textbf{20}} &\multicolumn{1}{c}{\textbf{30}} \\
\midrule
  &col2 text &col3 text &0.7568&1.0530&1.2642&0.9919&1.3541&1.6108 \\
  & &col2 text &12.5701 &19.6603&25.6809&18.0689&28.4865&37.3011 \\
3 &col2 text  & col3 text &0.7426&1.0393&1.2507&0.9095&1.2524&1.4958 \\
  & &col3 text &12.8008&19.9620&26.0324&16.6347&26.0843&34.0765 \\
  & col2 text & col3 text &0.7285&1.0257&1.2374&0.8195&1.1407&1.3691\tnote{*} \\
  & & col3 text &13.0360&20.2690&26.3895&15.0812&23.4932&30.6060\tnote{\dagger} \\
\bottomrule
\end{tabular*}
\begin{tablenotes}%%[\textheight]
\item[*] First sideways table footnote. Sideways table footnote. Sideways table footnote. Sideways table footnote.
\item[$\dagger$] Second sideways table footnote. Sideways table footnote. Sideways table footnote. Sideways table footnote.
\end{tablenotes}
\end{sidewaystable}

\begin{sidewaysfigure}
\centerline{\includegraphics[width=542pt,height=9pc,draft]{empty}}
\caption{Sideways figure caption. Sideways figure caption. Sideways figure caption. Sideways figure caption. Sideways figure caption. Sideways figure caption.\label{fig3}}
\end{sidewaysfigure}

nibh. Pellentesque wisi. Nullam malesuada. Morbi ut tellus ut pede tincidunt porta. Lorem ipsum dolor sit amet,
consectetuer adipiscing elit. Etiam congue neque id dolor.

\begin{algorithm}
\caption{Pseudocode for our algorithm}\label{alg1}
\begin{algorithmic}
  \For each frame
  \For water particles $f_{i}$
  \State compute fluid flow~\cite{Rothermel1997}
  \State compute fluid--solid interaction~\cite{Allen2011}
  \State apply adhesion and surface tension~\cite{Ballen2011}
  \EndFor
   \For solid particles $s_{i}$
   \For neighboring water particles $f_{j}$
   \State compute virtual water film \\(see Section~\ref{sec3})
   \EndFor
   \EndFor
   \For solid particles $s_{i}$
   \For neighboring water particles $f_{j}$
   \State compute growth direction vector \\(see Section~\ref{sec2})
   \EndFor
   \EndFor
   \For solid particles $s_{i}$
   \For neighboring water particles $f_{j}$
   \State compute $F_{\theta}$ (see Section~\ref{sec1})
   \State compute $CE(s_{i},f_{j})$ \\(see Section~\ref{sec3})
   \If $CE(b_{i}, f_{j})$ $>$ glaze threshold
   \State $j$th water particle's phase $\Leftarrow$ ICE
   \EndIf
   \If $CE(c_{i}, f_{j})$ $>$ icicle threshold
   \State $j$th water particle's phase $\Leftarrow$ ICE
   \EndIf
   \EndFor
   \EndFor
  \EndFor
\end{algorithmic}
\end{algorithm}

Donec et nisl at wisi luctus bibendum. Nam interdum tellus ac libero. Sed sem justo, laoreet vitae, fringilla at,
adipiscing ut, nibh. Maecenas non sem quis tortor eleifend fermentum. Etiam id tortor ac mauris porta vulputate.
Integer porta neque vitae massa~\cite{Rothermel1997,Elbaum2002}. Maecenas tempus libero a libero posuere dictum. Vestibulum ante ipsum primis in
faucibus orci luctus et ultrices posuere cubilia Curae; Aenean quis mauris sed elit commodo placerat. Class aptent
taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Vivamus rhoncus tincidunt libero.
Etiam elementum pretium justo. Vivamus est. Morbi a tellus eget pede tristique commodo~\cite{Elbaum2002} Nulla nisl. Vestibulum
sed nisl eu sapien cursus rutrum.

Pellentesque wisi. Nullam malesuada. Morbi ut tellus ut pede tincidunt porta. Lorem ipsum dolor sit amet,
consectetuer adipiscing elit. Etiam congue neque id dolor.

Donec et nisl at wisi luctus bibendum. Nam interdum tellus ac libero. Sed sem justo, laoreet vitae, fringilla at,
adipiscing ut, nibh. Maecenas non sem quis tortor eleifend fermentum. Etiam id tortor ac mauris porta vulputate.
Integer porta neque vitae massa. Maecenas tempus libero a libero posuere dictum. Vestibulum ante ipsum primis in
faucibus orci luctus et ultrices posuere cubilia Curae; Aenean quis mauris sed elit commodo placerat. Class aptent
taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Vivamus rhoncus tincidunt libero.
Etiam elementum pretium justo. Vivamus est. Morbi a tellus eget pede tristique commodo. Nulla nisl. Vestibulum
sed nisl eu sapien cursus rutrum.

\begin{equation}\label{eq23}
 \|\tilde{X}(k)\|^2
 =\frac{\left\|\sum\limits_{i=1}^{p}\tilde{Y}_i(k)+\sum\limits_{j=1}^{q}\tilde{Z}_j(k) \right\|^2}{(p+q)^2}
 \leq\frac{\sum\limits_{i=1}^{p}\left\|\tilde{Y}_i(k)\right\|^2+\sum\limits_{j=1}^{q}\left\|\tilde{Z}_j(k)\right\|^2 }{p+q}.
\end{equation}

Sed feugiat. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Ut pellentesque
augue sed urna. Vestibulum diam eros, fringilla et, consectetuer eu, nonummy id, sapien. Nullam at lectus. In sagittis
ultrices mauris. Curabitur malesuada erat sit amet massa. Fusce blandit. Aliquam erat volutpat. Aliquam euismod.
Aenean vel lectus. Nunc imperdiet justo nec dolor.

Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.

\begin{equation}\label{eq24}
 \|\tilde{X}(k)\|^2
 =\frac{\left\|\sum\limits_{i=1}^{p}\tilde{Y}_i(k)+\sum\limits_{j=1}^{q}\tilde{Z}_j(k) \right\|^2}{(p+q)^2}
 \leq\frac{\sum\limits_{i=1}^{p}\left\|\tilde{Y}_i(k)\right\|^2+\sum\limits_{j=1}^{q}\left\|\tilde{Z}_j(k)\right\|^2 }{p+q}.
\end{equation}

Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra
tortor. Sed fringilla justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue
est scelerisque sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas
urna mi, suscipit in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.

Etiam ac leo a risus tristique nonummy. Donec dignissim tincidunt nulla. Vestibulum rhoncus molestie odio. Sed
lobortis, justo et pretium lobortis, mauris turpis condimentum augue, nec ultricies nibh arcu pretium enim. Nunc
purus neque, placerat id, imperdiet sed, pellentesque nec, nisl. Vestibulum imperdiet neque non sem accumsan laoreet.
In hac habitasse platea dictumst. Etiam condimentum facilisis libero. Suspendisse in elit quis nisl aliquam dapibus.
Pellentesque auctor sapien. Sed egestas sapien nec lectus. Pellentesque vel dui vel neque bibendum viverra. Aliquam
porttitor nisl nec pede. Proin mattis libero vel turpis. Donec rutrum mauris et libero. Proin euismod porta felis.
Nam lobortis, metus quis elementum commodo, nunc lectus elementum mauris, eget vulputate ligula tellus eu neque.
Vivamus eu dolor.

\section{Conclusions}\label{sec5}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae,
felis. Curabitur dictum gravida mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna. Donec
vehicula augue eu neque. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
Mauris ut leo. Cras viverra metus rhoncus sem. Nulla et lectus vestibulum urna fringilla ultrices. Phasellus eu tellus
sit amet tortor gravida placerat. Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. Praesent eget sem vel
leo ultrices bibendum. Aenean faucibus. Morbi dolor nulla, malesuada eu, pulvinar at, mollis ac, nulla. Curabitur
auctor semper nulla. Donec varius orci eget risus. Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

%\backmatter

\section*{Acknowledgments}
This is acknowledgment text~\cite{Elbaum2002}. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here.

\subsection*{Author contributions}

This is an author contribution text. This is an author contribution text. This is an author contribution text. This is an author contribution text. This is an author contribution text.

\subsection*{Financial disclosure}

None reported.

\subsection*{Conflict of interest}

The authors declare no potential conflict of interests.

\section*{Supporting information}

The following supporting information is available as part of the online article:

\noindent
\textbf{Figure S1.}
{500{\uns}hPa geopotential anomalies for GC2C calculated against the ERA Interim reanalysis. The period is 1989--2008.}

\noindent
\textbf{Figure S2.}
{The SST anomalies for GC2C calculated against the observations (OIsst).}

\appendix

\section{Section title of first appendix\label{app1}}

Use \verb+\begin{verbatim}...\end{verbatim}+ for program codes without math. Use \verb+\begin{alltt}...\end{alltt}+ for program codes with math. Based on the text provided inside the optional argument of \verb+\begin{code}[Psecode|Listing|Box|Code|+\hfill\break \verb+Specification|Procedure|Sourcecode|Program]...+ \verb+\end{code}+ tag corresponding boxed like floats are generated. Also note that \verb+\begin{code}[Code|Listing]...+ \verb+\end{code}+ tag with either Code or Listing text as optional argument text are set with computer modern typewriter font.  All other code environments are set with normal text font. Refer below example:

\begin{lstlisting}[caption={Descriptive Caption Text},label=DescriptiveLabel]
for i:=maxint to 0 do
begin
{ do nothing }
end;
Write('Case insensitive ');
WritE('Pascal keywords.');
\end{lstlisting}

\subsection{Subsection title of first appendix\label{app1.1a}}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

\subsubsection{Subsection title of first appendix\label{app1.1.1a}}

\noindent\textbf{Unnumbered figure}

\begin{center}
\includegraphics[width=7pc,height=8pc,draft]{empty}
\end{center}

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.

Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.
Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.

\section{Section title of second appendix\label{app2}}%

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.

Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.
Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.

%== Figure 4 ==
%% Example for figure inside appendix
\begin{figure}[t]
\centerline{\includegraphics[height=10pc,width=78mm,draft]{empty}}
\caption{This is an example for appendix figure.\label{fig5}}
\end{figure}

\subsection{Subsection title of second appendix\label{app2.1a}}

Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non odio.
Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce sed justo
eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor. Donec
pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin et quam.
Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien turpis,
fermentum vel, eleifend faucibus, vehicula eu, lacus.

Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
Pellentesque eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut
mi mi, lacinia sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui.
Mauris nibh leo, facilisis non, adipiscing quis, ultrices a, dui.

\subsubsection{Subsection title of second appendix\label{app2.1.1a}}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae,
felis. Curabitur dictum gravida mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna. Donec
vehicula augue eu neque. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
Mauris ut leo. Cras viverra metus rhoncus sem. Nulla et lectus vestibulum urna fringilla ultrices. Phasellus eu tellus
sit amet tortor gravida placerat. Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. Praesent eget sem vel
leo ultrices bibendum. Aenean faucibus. Morbi dolor nulla, malesuada eu, pulvinar at, mollis ac, nulla. Curabitur
auctor semper nulla. Donec varius orci eget risus. Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

\begin{center}
\begin{table}[b]%
\centering
\caption{This is an example of Appendix table showing food requirements of army, navy and airforce.\label{tab4}}%
\begin{tabular*}{300pt}{@{\extracolsep\fill}lcc@{\extracolsep\fill}}%
\toprule
\textbf{col1 head} & \textbf{col2 head} & \textbf{col3 head} \\
\midrule
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text\\
\bottomrule
\end{tabular*}
\end{table}
\end{center}

Example for an equation inside appendix
\begin{equation}
\mathcal{L}\quad \mathbf{\mathcal{L}} = i \bar{\psi} \gamma^\mu D_\mu \psi - \frac{1}{4} F_{\mu\nu}^a F^{a\mu\nu} - m \bar{\psi} \psi\label{eq25}
\end{equation}

\section{Example of another appendix section\label{app3}}%

This is sample for paragraph text this is sample for paragraph text  this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

\begin{equation}
\mathcal{L} = i \bar{\psi} \gamma^\mu D_\mu \psi
    - \frac{1}{4} F_{\mu\nu}^a F^{a\mu\nu} - m \bar{\psi} \psi
\label{eq26}
\end{equation}

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
vitae risus porta vehicula.

\begin{center}
\begin{tabular*}{300pt}{@{\extracolsep\fill}lcc@{\extracolsep\fill}}%
\toprule
\textbf{col1 head} & \textbf{col2 head} & \textbf{col3 head} \\
\midrule
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
\bottomrule
\end{tabular*}
\end{center}

Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
vitae risus porta vehicula.

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, evehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.

Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
Pellentesque eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut
mi mi, lacinia sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui.
Mauris nibh leo, facilisis non, adipiscing quis, ultrices a, dui.

\nocite{*}% Show all bib entries - both cited and uncited; comment this line to view only cited bib entries;
\bibliography{wileyNJD-APA}%

\section*{Author Biography}

\begin{biography}{\includegraphics[width=60pt,height=70pt,draft]{empty}}{\textbf{Author Name.} This is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text.}
\end{biography}

\end{document}

### 2022-11-22 StatPaper/APA-lato_New/APA/wileyNJD-APA.tex
made a version, need to readover

@@
 
-\title{This is the sample article title\protect\thanks{This is an example for title footnote.}}
+\title{Case-Base Neural Networks: survival analysis with time-varying,
+higher-order interactions}
 
-\author[1]{Author One*}
+\author[1]{Jesse Islam}
 
-\author[2,3]{Author Two}
+\author[2]{Maxime Turgeon}
 
-\author[3]{Author Three}
+\author[3]{Robert Sladek}
 
-\authormark{AUTHOR ONE \textsc{et al}}
+\author[4]{Sahir Bhatnagar}
 
-\address[1]{\orgdiv{Org Division}, \orgname{Org Name}, \orgaddress{\state{State name}, \country{Country name}}}
+\address[1]{\orgdiv{Department of Quantitative life sciences,} \orgname{McGill University}, \orgaddress{\state{Quebec}, \country{Canada}}}
 
-\address[2]{\orgdiv{Org Division}, \orgname{Org Name}, \orgaddress{\state{State name}, \country{Country name}}}
+\address[2]{\orgdiv{Department of Statistics}, \orgname{University of Manitoba}, \orgaddress{\state{Manitoba}, \country{Canada}}}
 
-\address[3]{\orgdiv{Org Division}, \orgname{Org Name}, \orgaddress{\state{State name}, \country{Country name}}}
+\address[3]{\orgdiv{Department of Human Genetics}, \orgname{McGill University}, \orgaddress{\state{Quebec}, \country{Canada}}}
 
-\corres{*Corresponding author name, This is sample corresponding address. \email{authorone@gmail.com}}
+\address[4]{\orgdiv{Department of Biostatistics}, \orgname{McGill University}, \orgaddress{\state{Quebec}, \country{Canada}}}
+
+\corres{Jesse Islam's \email{jesse.islam@mail.mcgill.ca}}
 
@@
 
-\abstract[Summary]{This is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text.}
+\abstract[Summary]{Neural network-based survival methods can model data-driven
+covariate interactions. While these methods have led to better
+predictive performance than regression-based approaches, they cannot
+model both time-varying interactions and complex baseline hazards. To
+address this, we propose Case-Base Neural Networks (CBNN) as a new
+approach that combines the case-base sampling framework with flexible
+architectures. Our method naturally accounts for censoring and does not
+require method specific hyperparameters. Using a novel sampling scheme
+and data augmentation, we incorporate time directly into a feed-forward
+neural network. CBNN predicts the probability of an event occurring at a
+given moment and estimates the hazard function. We compare the
+performance of CBNN to survival methods based on regression and neural
+networks in two simulations and two real data applications. We report
+two time-dependent metrics for each model. In the simulations and real
+data applications, CBNN provides a more consistent predictive
+performance across time and outperforms the competing neural network
+approaches. For a simple simulation with an exponential hazard model,
+CBNN outperforms the other neural network methods. For a complex
+simulation, which highlights the ability of CBNN to model both a complex
+baseline hazard and time-varying interactions, CBNN outperforms all
+competitors. The first real data application shows CBNN outperforming
+all neural network competitors, while a second real data application
+shows competitive performance. We highlight the benefit of combining
+case-base sampling with deep learning to provide a simple and flexible
+modeling framework for data-driven, time-varying interaction modeling of
+survival outcomes. An R package is available at
+\url{https://github.com/Jesse-Islam/cbnn}.}
 
-\keywords{keyword1, keyword2, keyword3, keyword4}
-
-\jnlcitation{\cname{%
-\author{Williams K.},
-\author{B. Hoskins},
-\author{R. Lee},
-\author{G. Masato}, and
-\author{T. Woollings}} (\cyear{2016}),
-\ctitle{A regime analysis of Atlantic winter jet variability applied to evaluate HadGEM3-GC2}, \cjournal{Q.J.R. Meteorol. Soc.}, \cvol{2017;00:1--6}.}
+\keywords{survival analysis, machine learning, case-base, neural
+network}
 
@@
 
-\footnotetext{\textbf{Abbreviations:} ANA, anti-nuclear antibodies; APC, antigen-presenting cells; IRF, interferon regulatory factor}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-\section{Sample for first level head}\label{sec1}
+Smooth-in-time accelerated failure time (AFT) models can estimate
+absolute risks by modeling the hazard directly through a user-specified
+baseline hazard distribution \citep{kleinbaum2012survival}. Cox
+proportional hazards models are used more often than AFT models, causing
+analyses to be based on hazard ratios and relative risks rather than on
+survival curves and absolute risks \citep{hanley2009}. The
+identification of an appropriate distribution for the baseline hazard in
+an AFT model may be difficult for common diseases that have many
+interacting risk factors, or a Cox model where the disease pathogenesis
+may change with age, making it difficult to maintain the proportional
+hazards assumption. For example, previous studies of breast cancer
+incidence have discovered time-varying interactions with covariates of
+interest, such as tumor size \citep{coradini2000time}. One approach to
+provide flexibility in the baseline hazard involves using the basis of
+splines on time in our model \citep{royston2002flexible}. However,
+regression-based models are limited in that they require prior knowledge
+about potential time-varying interactions and their quantitative
+effects.
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit~\citet{Rothermel1997} Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae,
-felis. Curabitur dictum gravida mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna. Donec
-vehicula augue eu neque. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
-Mauris ut leo. Cras viverra metus rhoncus sem. Nulla et lectus vestibulum urna fringilla ultrices. Phasellus eu tellus
-sit amet tortor gravida placerat. Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. Praesent eget sem vel
-leo ultrices bibendum. Aenean faucibus. Morbi dolor nulla, malesuada eu, pulvinar at, mollis ac, nulla. Curabitur
-auctor semper nulla. Donec varius orci eget risus. Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
+Neural networks provide a data-driven approach to approximating
+interaction terms. For example, DeepSurv is a neural network-based
+proportional hazards model that implements the Cox partial
+log-likelihood as a custom loss function \citep{katzman2018DeepSurv},
+resulting in a stepwise absolute risk curve that cannot accommodate
+time-varying interactions. Compared to Cox regression, DeepSurv shows
+better performance on the Study to Understand Prognoses Preferences and
+Risk Treatments (SUPPORT) dataset \citep{knaus1995SUPPORT}. To handle
+non-proportional hazards, a modification to the loss function was
+proposed \citep{faraggi1995neural}. To remove the need for tuning the
+number of layers and nodes, the concept of extreme learning machines has
+been applied to a Cox neural network model \citep{wang2019extreme}.
 
-\begin{eqnarray}
-s(nT_{s}) &= &s(t)\times \sum\limits_{n=0}^{N-1} \delta (t-nT_{s}) \xleftrightarrow{\mathrm{DFT}}  S \left(\frac{m}{NT_{s}}\right) \nonumber\\
-&= &\frac{1}{N} \sum\limits_{n=0}^{N-1} \sum\limits_{k=-N/2}^{N/2-1} s_{k} e^{\mathrm{j}2\pi k\Delta fnT_{s}} e^{-j\frac{2\pi}{N}mn}
-\end{eqnarray}
+As an alternative method that assumes a baseline hazard distribution,
+DeepHit specifies each survival time of interest in the model and
+directly estimates survival curves, rather than deriving a hazard
+function \citep{lee2018DeepHit}. It assumes an inverse Gaussian
+distribution as the baseline hazard and it outperforms DeepSurv on the
+Molecular Taxonomy of Breast Cancer International Consortium (METABRIC)
+dataset \citep{curtis2012genomic}.
 
-\section{Sample for another first level head}\label{sec2}
+Providing further flexibility, Deep Survival Machines (DSM) is a
+parametric survival model using neural networks with a mixture of
+distributions as the baseline hazard \citep{dsmPaper}. On both the
+SUPPORT and METABRIC datasets, DSM outperforms DeepSurv and DeepHit
+\citep{dsmPaper}. However, like DeepHit and DeepSurv, DSM cannot model
+time-varying interactions.
 
-Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus viverra
-fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa ac
-quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo~\cite{Rothermel1998} Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
-nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
-bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
-pellentesque felis eu massa.
+We note that these alternative neural network approaches all require
+custom loss functions \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}
+\citep{dsmPaper}. DeepHit introduces a hyperparameter weighing its two
+loss functions (negative log-likelihood and ranking losses) while DSM
+requires a two-phase learning process and user implementations for
+distributions beyond Log-Normal or Weibull \citep{lee2018DeepHit}
+\citep{dsmPaper}. Regression-based approaches require prior
+specification of all interaction terms, which makes it challenging to
+model covariate effects that change over time. The current neural
+network models provide flexibility at the cost of opacity, while
+regression models provide clarity at the cost of flexibility.
 
-Example for bibliography citations cite~\citep{Elbaum2002}, cites~\cite{Allen2011,Yoo2007}
+In this article, we propose Case-Base Neural Networks (CBNN) as a method
+that models time-varying interactions and a flexible baseline hazard
+using commonly available neural network components. Our approach to
+modeling the full hazard uses case-base sampling \citep{hanley2009}.
+This sampling technique allows probabilistic models to predict survival
+outcomes. As part of the case-base framework, we use transformations of
+time as a feature (covariate) to specify different baseline hazards. For
+example, by including splines of time as covariates, we can approximate
+the Royston-Parmar flexible baseline hazard model
+\citep{royston2002flexible}, however, this still requires explicit use
+of time-varying interactions. CBNN can model both without extra tuning
+parameters.
 
-Quisque ullamcorper placerat ipsum. Cras nibh~\cite{Yoo2007,Schulz2012} Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
-amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
-Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
-quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
-vitae risus porta vehicula.
+In Section \ref{methods}, we describe how case-base sampling and neural
+networks are combined both conceptually and algebraically, along with
+our hyperparameter choices and software implementation. Section
+\ref{sims}, describes our metrics and compares the performance of CBNN,
+DeepSurv, DeepHit, DSM, Cox regression and case-base using logistic
+regression (CBLR) on simulated data. Section \ref{casestudies} describes
+the real-data analysis, while Section \ref{discussion} explores the
+implications of our results and contextualizes them within neural
+network survival analysis in a single event setting.
 
-Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
-venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
-quam, in hendrerit risus eros eget felis.
-Maecenas eget erat in sapien mattis porttitor.
-Vestibulum porttitor.
-Nulla facilisi.
-Sed a turpis eu lacus commodo facilisis.
-Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
-vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
-Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%section2
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{figure}[t]
-\centerline{\includegraphics[width=342pt,height=9pc,draft]{empty}}
-\caption{This is the sample figure caption.\label{fig1}}
+\hypertarget{methods}{%
+\section{Case-base neural network methodology, metrics and
+software}\label{methods}}
+
+In this section, we define case-base sampling, which converts the total
+survival time into discrete person-specific moments (person-moments).
+Then, we detail how neural networks can be used within this framework,
+explicitly incorporating time as a feature while adjusting for the
+sampling bias. Finally, we report on the software versions used. An R
+package is available for use at
+\url{https://github.com/Jesse-Islam/cbnn}. The entire code base to
+reproduce the figures and empirical results in this paper is available
+at \url{https://github.com/Jesse-Islam/cbnnManuscript}.
+
+\hypertarget{case-base-sampling}{%
+\subsection{Case-base sampling}\label{case-base-sampling}}
+
+Case-base sampling is an alternative framework for survival analysis
+\citep{hanley2009}. In case-base sampling, we sample from the continuous
+survival time of each person in our dataset to create a \emph{base
+series} of \emph{person-moments}. This \emph{base series} complements
+the \emph{case series}, which contains all person-moments at which the
+event of interest occurs.
+
+For each person-moment sampled, let \(X_i\) be the corresponding
+covariate profile \(\left(x_{i1},x_{i2},...,x_{ip} \right)\), \(T_i\) be
+the time of the person-moment and \(Y_i\) be the indicator variable for
+whether the event of interest occurred at time \(T_i\). We estimate the
+hazard function \(h(t \mid X_i)\) using the sampled person-moments.
+Recall that \(h(t \mid X_i)\) is the instantaneous potential of
+experiencing the event at time \(t\) for a given set of covariates
+\(X_i\), assuming \(T_i \geq t\).
+
+Now, let \(b\) be the (user-defined) size of the \emph{base series} and
+let \(B\) be the sum of all follow-up times for the individuals in the
+study. If we sample the \emph{base series} uniformly across the study
+base, then the hazard function of the sampling process is equal to
+\(b/B\). Therefore, we have the following equality
+\footnote{We are abusing notation here, conflating hazards with probabilities. For a rigorous treatment, see Saarela \& Hanley (2015) section 3 \cite{saarela2015} .}:
+\begin{align}\label{eqn:main}
+\frac{P\left(Y_i=1 \mid X_i, T_i\right)}{P\left(Y_i = 0 \mid X_i, T_i\right)} = \frac{h\left(T_i \mid X_i\right)}{b/B}.
+\end{align} The odds of a person-moment being a part of the \emph{case
+series} is the ratio of the hazard \(h(T_i \mid X_i)\) and the uniform
+rate \(b/B\). Using \eqref{eqn:main}, we can see how the log-hazard
+function can be estimated from the log-odds arising from case-base
+sampling: \begin{align}\label{eqn:offset}
+\log \left( h\left(t \mid X_i\right)\right) = \log \left(\frac{P\left(Y_i = 1 \mid X_i, t\right)}{P\left(Y_i = 0 \mid X_i, t\right)}\right) + \log\left(\frac{b}{B}\right).
+\end{align}
+
+To estimate the correct hazard function, we adjusting for the bias
+introduced when sampling a fraction of the study base \(B\) by adding
+the offset term \(\log\left(\frac{b}{B} \right)\) as in
+\eqref{eqn:offset}. Next, we propose using neural networks to model the
+odds.
+
+\hypertarget{neural-networks-to-model-the-hazard-function}{%
+\subsection{Neural networks to model the hazard
+function}\label{neural-networks-to-model-the-hazard-function}}
+
+After case-base sampling, we pass all features, including time, into any
+user-defined feed-forward component, to which an offset term is added,
+then passed through a sigmoid activation function (Figure
+\ref{fig:NNarch}). We use the sigmoid function as its inverse is the
+odds, which we can use to calculate the hazard. The general form for the
+neural network using CBNN is:
+
+\begin{align}\label{eqn:nnProb}
+P\left(Y=1|X,T\right)&=\mathrm{sigmoid}\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right) \right),
+\end{align}
+
+where \(T\) is a random variable representing the event time, \(X\) is
+the random variable for a covariate profile, \(f_{\theta}(X, T)\)
+represents any feed-forward neural network architecture,
+\(\log\left(\frac{B}{b}\right)\) is bias term set by case-base sampling,
+\(\theta\) is the set of parameters learned by the neural network and
+\(\mathrm{sigmoid}(x)=\frac{1}{1+e^{-x}}\). By approximating a
+higher-order polynomial of time using a neural network, the baseline
+hazard specification is now data-driven, where user-defined
+hyperparameters such as regularization, number of layers and nodes
+control the flexibility of the hazard function. We provide a detailed
+description of the choices we made in the next sub-section.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{../../../figures/nnarch2}}
+
+\caption{Steps involved in CBNN from case-base sampling to the model framework we use for training.
+The first step is case-base sampling, completed before training begins.
+Next, we pass this sampled data through a feed-forward neural network.
+We add the offset and pass that through a sigmoid activation function, whose output is a probability.
+Once the neural network model completes its training, we can convert the probability output to a hazard, using it for our survival outcomes of interest.}\label{fig:NNarch}
 \end{figure}
 
-Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
-ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
-eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
-convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
-pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.
+The following derivation shows how our probability estimate is converted
+to odds: \begin{align*}
+ \log\left( h(t \mid X) \right) &= \log\left(\frac{\mathrm{sigmoid}\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}{1-\mathrm{sigmoid}\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}\right) + \log\left(\frac{b}{B}\right) \\
+ &= \log\left( \frac{\frac{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)+1}}{1-\frac{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)}{\exp\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right)\right)+1}}\right) + \log\left(\frac{b}{B}\right) \\
+ &= \log\left(\exp\left( f_{\theta}(X, T) + \log\left(\frac{B}{b}\right) \right) \right) + \log\left(\frac{b}{B}\right) \\
+ &= f_{\theta}(X, T) + \log\left(\frac{B}{b}\right) + \log\left(\frac{b}{B}\right) \\
+&= f_{\theta}(X, T). 
+\end{align*}
 
-Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non
-odio. Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce
-sed justo eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor.
-Donec pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin
-et quam. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien
-turpis, fermentum vel, eleifend faucibus, vehicula eu, lacus.
+We use binary cross-entropy as our loss function \citep{gulli2017}:
+\begin{align*}
+L(\theta)=-\frac{1}{N} \sum^{N}_{i=1} y_{i} \cdot \log(\hat{f}_{\theta}(x_{i}, t_{i}) ) + (1-y_{i} )\cdot \log(1-\hat{f}_{\theta}(x_{i}, t_{i}) ),
+\end{align*} where \(\hat{f}_{\theta}(x_{i}, t_{i})\) is our estimate
+for a given covariate profile and time, \(y_{i}\) is our target value
+specifying whether an event occurred and \(N\) represents the number of
+individuals in our training set.
 
-\begin{figure*}
-\centerline{\includegraphics[width=342pt,height=9pc,draft]{empty}}
-\caption{This is the sample figure caption.\label{fig2}}
-\end{figure*}
+Backpropagation with an appropriate minimization algorithm (e.g.~Adam,
+RMSPropagation, stochastic gradient descent) is used to optimize the
+parameters in the model \citep{gulli2017}. For our analysis, we use Adam
+as implemented in Keras \citep{gulli2017}. Note that the size of the
+\emph{case series} is fixed as the number of events, but we can make the
+\emph{base series} as large as we want. A ratio of 100:1 \emph{base
+series} to \emph{case series} is sufficient \citep{hanley2009}. We pass
+our feed-forward neural network through a sigmoid activation function
+(Figure \ref{fig:NNarch}). Finally, we can convert this model output to
+a hazard. When using our model for predictions, we manually set the
+offset term to 0 in the new data, as we account for the bias during the
+fitting process.
 
-\subsection{Example for second level head}
+Since we are directly modeling the hazard, we can readily estimate the
+risk function (\(F\)) at time \(t\) for a covariate profile \(X\),
+viz.\begin{align}\label{eqn:ci2}
+F\left(t\mid X\right)& = 1 - \exp\left(-\int_{0}^{t}h(u|X) \,\textrm du\right).
+\end{align} We use a finite Riemann sum \citep{hughes2020calculus} to
+approximate the integral in \eqref{eqn:ci2}.
 
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit, dictum
-in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam vestibulum
-fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi. Pellentesque
-eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut mi mi, lacinia
-sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui. Mauris nibh leo,
-facilisis non, adipiscing quis, ultrices a, dui.
+\hypertarget{hyperparameter-selection}{%
+\subsection{Hyperparameter selection}\label{hyperparameter-selection}}
 
-Morbi luctus, wisi viverra faucibus pretium, nibh est placerat odio, nec commodo wisi enim eget quam. Quisque
-libero justo, consectetuer a, feugiat vitae, porttitor eu, libero. Suspendisse sed mauris vitae elit sollicitudin malesuada.
+Neural networks are flexible when defining the architecture and
+optimization parameters. These hyperparameter decisions can affect the
+estimated parameters and were chosen during a set of initial simulations
+to determine if CBNN can learn complex interactions in practice. We set
+a max number of epochs to be \(2000\), batch size as \(512\), learning
+rate as \(10^{-3}\), decay as \(10^{-7}\), patience as \(10\) epochs,
+\(\{50,50,25,25\}\) nodes in each hidden layer with \(50\%\) dropout at
+each layer, a minimum delta loss on the validation set of \(10^{-7}\)
+over 10 epochs and Adam \citep{gulli2017} as our optimizer. These
+choices may permit the model to approximate higher-order interactions
+while preventing over-fitting \citep{srivastava2014dropout}. We fix a
+train-validation-test split that allows us to update the weights with a
+subset of the data (training), assess performance at each epoch
+(validation) and gauge performance of the final model (test) for each
+method. We select the best weights after training based on validation
+loss \citep{gulli2017}.
 
-Maecenas ultricies eros sit amet ante. Ut venenatis velit. Maecenas sed mi eget dui varius euismod. Phasellus aliquet
-volutpat odio. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Pellentesque sit
-amet pede ac sem eleifend consectetuer. Nullam elementum, urna vel imperdiet sodales, elit ipsum pharetra ligula,
-ac pretium ante justo a nulla. Curabitur tristique arcu eu metus. Vestibulum lectus. Proin mauris. Proin eu nunc eu
-urna hendrerit faucibus. Aliquam auctor, pede consequat laoreet varius, eros tellus scelerisque quam, pellentesque
-hendrerit ipsum dolor sed augue. Nulla nec lacus.
+\hypertarget{software-implementation}{%
+\subsection{Software implementation}\label{software-implementation}}
 
-\begin{quote}
-This is an example~\citep{Elbaum2002,Blanchard2015} for quote text.
-This is an example for quote text.
-This is an example for quote text.
-This is an example for quote text~\cite{Blanchard2015}.
-This is an example for quote text.
-This is an example for quote text.
-This is an example for quote text.
-This is an example for quote text.
-This is an example for quote text.
-This is an example for quote text~\cite{Blanchard2015}.
-This is an example for quote text.
-This is an example for quote text.
-This is an example for quote text.
-\end{quote}
+R \citep{Rsoft} and python \citep{py} are used to evaluate methods from
+both languages. We fit the Cox model using the \textbf{survival} package
+\citep{survpkg}, the CBLR model using the \textbf{casebase} package
+\citep{cbpkg}, the DeepSurv model using the \textbf{survivalmodels}
+package \citep{survmods}, the DeepHit model using \textbf{pycox}
+\citep{lee2018DeepHit} and the DSM model using
+\textbf{DeepSurvivalMachines} \citep{dsmPaper}. We made the components
+of CBNN using the \textbf{casebase} package for the sampling step and
+the \textbf{keras} \citep{keras} package for our neural network
+architecture. The \textbf{simsurv} package \citep{simsurv} is used for
+our simulation studies, while \textbf{flexsurv} \citep{flexsurv} is used
+to fit a flexible baseline hazard using splines for our complex
+simulation. We use the implementation of \(C_{IPCW}\) from the python
+package \textbf{sksurv} \citep{sksurv}. The \textbf{riskRegression}
+package \citep{riskRegression} is used to get the Index of Prediction
+Accuracy (IPA metric). Both metrics are described in detail in the
+following section. We modify the \textbf{riskRegression} package to be
+used with any user supplied risk function \(F\). To ensure that both R
+and Python-based models are running in unison on the same data through
+our simulations and bootstrap, we use the \textbf{reticulate} package
+\citep{reticulate}.
 
-\section{Sample for next first level head}\label{sec3}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%section3
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\subsection{Example for another second level head}
+\hypertarget{sims}{%
+\section{Simulation studies}\label{sims}}
 
-Suspendisse vitae elit. Aliquam arcu neque, ornare in, ullamcorper quis, commodo eu, libero. Fusce sagittis erat at
-erat tristique mollis. Maecenas sapien libero, molestie et, lobortis in, sodales eget, dui. Morbi ultrices rutrum lorem.
-Nam elementum ullamcorper leo. Morbi dui. Aliquam sagittis. Nunc placerat. Pellentesque tristique sodales est.
-Maecenas imperdiet lacinia velit. Cras non urna. Morbi eros pede, suscipit ac, varius vel, egestas non, eros. Praesent
-malesuada, diam id pretium elementum, eros sem dictum tortor, vel consectetuer odio sem sed wisi.
+In this section, we use simulated data to evaluate the performance of
+CBNN and compare our approach with existing regression-based (Cox, CBLR)
+and neural network-based (DeepHit, DeepSurv, DSM) methods. We specify a
+linear combination of each covariate as the linear predictor in
+regression-based approaches (Cox, CBLR), which contrasts with neural
+network approaches that allow for non-linear interactions. We simulate
+data under a simple exponential model and a complex baseline hazard with
+time-varying interactions, each with 10\% random censoring. For both
+settings, we simulate three covariates:
 
-Sed feugiat. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Ut pellentesque
-augue sed urna. Vestibulum diam eros, fringilla et, consectetuer eu, nonummy id, sapien. Nullam at lectus. In sagittis
-ultrices mauris. Curabitur malesuada erat sit amet massa. Fusce blandit. Aliquam erat volutpat. Aliquam euismod.
-Aenean vel lectus. Nunc imperdiet justo nec dolor.
+\[
+z_{1} \sim \textrm{Bernoulli}(0.5) \qquad \qquad
+z_{2} \sim \begin{cases}
+ N(0,0.5) & \textrm{if } z_{1}=0\\ 
+ N(10,0.5) & \textrm{if } z_{1}=1
+\end{cases} \qquad \qquad
+z_{3} \sim \begin{cases}
+ N(8,0.5) & \textrm{if } z_{1}=0\\ 
+ N(-3,0.5) & \textrm{if } z_{1}=1.
+\end{cases}
+\]
 
-\subsection{Second level head text}
+The DeepHit-specific hyperparameter alpha is set to \(0.5\) (equal
+weight between its negative log-likelihood and ranking losses
+\citep{lee2018DeepHit}). We modify the \textbf{DeepSurvivalMachines}
+\citep{dsmPaper} package to include dropout and a minimum delta loss
+during the fitting process. For DSM, we define a mixture model of six
+Weibull distributions for the baseline hazard. All other hyperparameters
+are held constant across all neural network methods in both the
+simulation studies and real data applications.
 
-Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
-eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
-facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
-a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
-convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.
+Besides the methods mentioned above, we include the Optimal model in our
+comparisons using CBLR. That is, we include the exact functional form of
+the covariates in a CBLR model (referred to as Optimal for simplicity).
+We calculate \(t\)-based 95\% confidence intervals using 100
+replications of the simulated data. For all analyses, we use 80\% for
+training and 20\% for the test set. 20\% of the training set is kept for
+validation at each epoch. We predict risk functions \(F\) using
+\eqref{eqn:ci2} for individuals in the test set, which are used to
+calculate our \(C_{IPCW}\) and IPA scores.
 
-\subsubsection{Third level head text}
+\hypertarget{performance-metrics}{%
+\subsection{Performance metrics}\label{performance-metrics}}
 
-Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra tortor. Sed fringilla
-justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue est scelerisque
-sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas urna mi, suscipit
-in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.
+We use two metrics to assess the performance of the different methods of
+interest: 1) (IPA) \citep{kattan2018index} and 2) inverse probability
+censoring weights-adjusted concordance index (\(C_{IPCW}\))
+\citep{uno2011}, which we define below. Both these time-dependent
+metrics provide transparency as to when in follow-up time each model may
+perform better than the others.
 
-Etiam ac leo a risus tristique nonummy. Donec dignissim tincidunt nulla. Vestibulum rhoncus molestie odio. Sed
-lobortis, justo et pretium lobortis, mauris turpis condimentum augue, nec ultricies nibh arcu pretium enim. Nunc
-purus neque, placerat id, imperdiet sed, pellentesque nec, nisl. Vestibulum imperdiet neque non sem accumsan laoreet.
-In hac habitasse platea dictumst. Etiam condimentum facilisis libero. Suspendisse in elit quis nisl aliquam dapibus.
-Pellentesque auctor sapien. Sed egestas sapien nec lectus. Pellentesque vel dui vel neque bibendum viverra. Aliquam
-porttitor nisl nec pede. Proin mattis libero vel turpis. Donec rutrum mauris et libero. Proin euismod porta felis.
-Nam lobortis, metus quis elementum commodo, nunc lectus elementum mauris, eget vulputate ligula tellus eu neque.
-Vivamus eu dolor.
+\hypertarget{index-of-prediction-accuracy-ipa}{%
+\subsubsection{Index of prediction accuracy
+(IPA)}\label{index-of-prediction-accuracy-ipa}}
 
-Nulla in ipsum. Praesent eros nulla, congue vitae, euismod ut, commodo a, wisi. Pellentesque habitant morbi
-tristique senectus et netus et malesuada fames ac turpis egestas. Aenean nonummy magna non leo. Sed felis erat,
-ullamcorper in, dictum non, ultricies ut, lectus. Proin vel arcu a odio lobortis euismod. Vestibulum ante ipsum primis
-in faucibus orci luctus et ultrices posuere cubilia Curae; Proin ut est. Aliquam odio. Pellentesque massa turpis, cursus
-eu, euismod nec, tempor congue, nulla. Duis viverra gravida mauris. Cras tincidunt. Curabitur eros ligula, varius ut,
-pulvinar in, cursus faucibus, augue.
+The IPA is a function of the Brier score (\(BS(t)\)) \citep{graf1999},
+which is defined as \begin{align}
+BS(t)=\frac{1}{N}\sum^{N}_{i=1}\left(\frac{\left(1 - \widehat{F}(t \mid X_{i})\right)^{2}\cdot I(T_{i}\leq t,\delta_{i}=1)}{\widehat{G}(T_{i})} + \frac{\widehat{F}(t\mid X_{i})^{2}\cdot I(T_{i}>t)}{\widehat{G}(t)}\right),
+\end{align} where \(\delta_{i}=1\) shows individuals who have
+experienced the event, \(N\) represents the number of samples in our
+dataset over which we calculate \(BS(t)\), \(\widehat{G}(t)=P[c>t]\) is
+a non-parametric estimate of the censoring distribution, \(c\) is
+censoring time and \(T_{i}\) is an individual's survival or censoring
+time. The Brier score provides a score that accounts for the information
+loss because of censoring. There are three categories of individuals
+that may appear within the dataset once we fix our \(t\) of interest.
+Individuals who experienced the event before \(t\) are present in the
+first term of the equation. The second term of the equation includes
+individuals who experience the event or are censored after \(t\). Those
+censored before \(t\) are the third category of people. The inverse
+probability censoring weights (IPCW) adjustment (\(G(\cdot)\)) is to
+account for these category three individuals whose information is
+missing. The IPA score as a function of time is given by \begin{align}
+\textrm{IPA}(t) &= 1-\frac{BS_{model}(t)}{BS_{null}(t)}, \nonumber
+\end{align} where \(BS_{model}(t)\) represents the Brier score over time
+\(t\) for the model of interest and \(BS_{null}(t)\) represents the
+Brier score if we use an unadjusted Kaplan-Meier (KM) curve as the
+prediction for all observations \citep{kattan2018index}. Note that IPA
+has an upper bound of one, where positive values show an increase in
+performance over the null model and negative values show that the null
+model performs better. These scores show how performance changes over
+follow-up time.
 
-\begin{boxtext}
-\section*{Example of Boxtext}%
-This is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext.
-\end{boxtext}
+A potential artifact of IPA is that the score is unstable at earlier and
+later survival times. This is because of a near equivalent Brier score
+among each model and the null model. At small values, a difference of
+\(0.1\) creates a more significant fold change than at larger values. As
+the Brier score is potentially small at the start and at the end of
+their respective curves, The IPA score may be unstable at the same
+locations.
 
-\paragraph{Fourth level head text}
+\hypertarget{inverse-probability-censoring-weights-adjusted-concordance-index}{%
+\subsubsection{Inverse probability censoring weights-adjusted
+concordance
+index}\label{inverse-probability-censoring-weights-adjusted-concordance-index}}
 
-Sed feugiat. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Ut pellentesque
-augue sed urna. Vestibulum diam eros, fringilla et, consectetuer eu, nonummy id, sapien. Nullam at lectus. In sagittis
-ultrices mauris. Curabitur malesuada erat sit amet massa. Fusce blandit. Aliquam erat volutpat. Aliquam euismod.
-Aenean vel lectus. Nunc imperdiet justo nec dolor.
+The \(C_{IPCW}\) is a non-proper, rank-based metric that does not depend
+on the censoring times in the test data \citep{uno2011}. The
+\(C_{IPCW}\) is given by
 
-Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
-eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
-facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
-a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
-convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.
+\begin{align} \label{eq:cidx}
+C_{IPCW}(t) &= \frac{\sum^{N}_{i=1}\sum^{N}_{j=1}\delta_{i}\left\{\widehat{G}(T_{i})\right\}^{-2} I(T_{i}<T_{j},T_{i}<t) I\left(\widehat{F}(t|X_{i})>\widehat{F}(t|X_{j})\right)}{\sum^{N}_{i=1}\sum^{N}_{j=1}\delta_{i}\left\{\widehat{G}(T_{i})\right\}^{-2} I(T_{i}<T_{j},T_{i}<t)}.
+\end{align} where the follow-up period of interest is (0,\(t\)),
+\(I(\cdot)\) is an indicator function, \(\widehat{F}(X_{i},t)\) is the
+risk function estimated for everyone in the study at time \(t\) and
+\(C_{IPCW}\) can compare the performance of different models, where a
+higher score is better. Note that the \(C_{IPCW}\) may produce
+misleading performance, as it ranks based on survival times, not event
+status \citep{cindexfails2019}. This metric is considered an unbiased
+population concordance measure because of the IPCW adjustment
+\citep{uno2011}.
 
-\subparagraph{Fifth level head text}
+\begin{figure}
 
-Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra
-tortor. Sed fringilla justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue
-est scelerisque sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas
-urna mi, suscipit in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.
+{\centering \includegraphics[width=1\linewidth]{../../../analyses/figures/megaPlot}
 
-Etiam ac leo a risus tristique nonummy. Donec dignissim tincidunt nulla. Vestibulum rhoncus molestie odio. Sed
-lobortis, justo et pretium lobortis, mauris turpis condimentum augue, nec ultricies nibh arcu pretium enim. Nunc
-purus neque, placerat id, imperdiet sed, pellentesque nec, nisl. Vestibulum imperdiet neque non sem accumsan laoreet.
-In hac habitasse platea dictumst. Etiam condimentum facilisis libero. Suspendisse in elit quis nisl aliquam dapibus.
-Pellentesque auctor sapien. Sed egestas sapien nec lectus. Pellentesque vel dui vel neque bibendum viverra. Aliquam
-porttitor nisl nec pede. Proin mattis libero vel turpis. Donec rutrum mauris et libero. Proin euismod porta felis.
-Nam lobortis, metus quis elementum commodo, nunc lectus elementum mauris, eget vulputate ligula tellus eu neque.
-Vivamus eu dolor.
+}
 
-in faucibus orci luctus et ultrices posuere cubilia Curae; Proin ut est. Aliquam odio. Pellentesque massa turpis, cursus
-eu, euismod nec, tempor congue, nulla. Duis viverra gravida mauris. Cras tincidunt. Curabitur eros ligula, varius ut,
-pulvinar in, cursus faucibus, augue.
+\caption{Summarizes the simple simulation (A, B), complex simulation (C, D), SUPPORT case study (E, F) and METABRIC case study (G, H) results.
+The first row shows the IPA score for each model in each study over follow-up time.
+Negative values mean our model performs worse than the null model and positive values mean the model performs better.
+The second row demonstrates the $C_{IPCW}$ score for each model in each study over follow-up time.
+A score of 1 is the maximum performance for either metric.
+Each model-specific metric in each study shows a 95-percent confidence interval over 100 iterations.
+The models of interest are case-base with logistic regression (CBLR), Case-Base Neural Networks (CBNN), Cox, DeepHit, DeepSurv, Deep Survival Machines (DSM), Optimal (a CBLR model with the exact interaction terms and baseline hazard specified) and Kaplan-Meier (to serve as a baseline, predicting the average for all individuals).}\label{fig:megaPlot}
+\end{figure}
 
-Curabitur tellus magna, porttitor a, commodo a, commodo in, tortor. Donec interdum. Praesent scelerisque. Mae-
-cenas posuere sodales odio. Vivamus metus lacus, varius quis, imperdiet quis, rhoncus a, turpis. Etiam ligula arcu,
-elementum a, venenatis quis, sollicitudin sed, metus. Donec nunc pede, tincidunt in, venenatis vitae, faucibus vel,
-nibh. Pellentesque wisi. Nullam malesuada. Morbi ut tellus ut pede tincidunt porta. Lorem ipsum dolor sit amet,
-consectetuer adipiscing elit. Etiam congue neque id dolor.
+\hypertarget{simple-simulation-constant-baseline-hazard}{%
+\subsection{Simple simulation: constant baseline
+hazard}\label{simple-simulation-constant-baseline-hazard}}
 
-Donec et nisl at wisi luctus bibendum. Nam interdum tellus ac libero. Sed sem justo, laoreet vitae, fringilla at,
-adipiscing ut, nibh. Maecenas non sem quis tortor eleifend fermentum. Etiam id tortor ac mauris porta vulputate.
-Integer porta neque vitae massa. Maecenas tempus libero a libero posuere dictum. Vestibulum ante ipsum primis in
-faucibus orci luctus et ultrices posuere cubilia Curae; Aenean quis mauris sed elit commodo placerat. Class aptent
-taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Vivamus rhoncus tincidunt libero.
-Etiam elementum pretium justo. Vivamus est. Morbi a tellus eget pede tristique commodo. Nulla nisl. Vestibulum
-sed nisl eu sapien cursus rutrum.
+We simulate data from a simple model that primarily depends on a
+constant baseline hazard:
+\[\log h(t \mid X_i) = \beta_{{1}}z_{1}+\beta_{{2}}z_{2}+\beta_{{3}}z_{3},\]
+The covariate effects are given by
+\(\beta_{{1}}=0.1, \beta_{{2}}=0.1, \beta_{{3}}=0.1\). Once we simulate
+survival times, we introduce 10\% random censoring.
 
-Nulla non mauris vitae wisi posuere convallis. Sed eu nulla nec eros scelerisque pharetra. Nullam varius. Etiam
-dignissim elementum metus. Vestibulum faucibus, metus sit amet mattis rhoncus, sapien dui laoreet odio, nec ultricies
-nibh augue a enim. Fusce in ligula. Quisque at magna et nulla commodo consequat. Proin accumsan imperdiet sem.
-Nunc porta. Donec feugiat mi at justo. Phasellus facilisis ipsum quis ante. In ac elit eget ipsum pharetra faucibus.
-Maecenas viverra nulla in massa.
+\hypertarget{performance-comparison-in-simple-simulation}{%
+\subsubsection{Performance comparison in simple
+simulation}\label{performance-comparison-in-simple-simulation}}
 
-Nulla in ipsum. Praesent eros nulla, congue vitae, euismod ut, commodo a, wisi. Pellentesque habitant morbi
-tristique senectus et netus et malesuada fames ac turpis egestas. Aenean nonummy magna non leo. Sed felis erat,
-ullamcorper in, dictum non, ultricies ut, lectus. Proin vel arcu a odio lobortis euismod. Vestibulum ante ipsum primis
+Figure \ref{fig:megaPlot} A, B and Table \ref{tab:megaTable} A show the
+results for the simple simulation. The regression-based methods (CBLR,
+Cox, Optimal) outperform the neural network ones in the simple
+simulation setting. Among the neural network approaches, CBNN
+outperforms all other methods in terms of both IPA and \(C_{IPCW}\)
+(Figure \ref{fig:megaPlot} A, B). Specifically, we see CBNN is
+consistent across time with smaller confidence intervals compared to
+DeepHit, DeepSurv and DSM.
 
-\begin{center}
-\begin{table*}[t]%
-\caption{This is sample table caption.\label{tab1}}
-\centering
-\begin{tabular*}{500pt}{@{\extracolsep\fill}lccD{.}{.}{3}c@{\extracolsep\fill}}
-\toprule
-&\multicolumn{2}{@{}c@{}}{\textbf{Spanned heading\tnote{1}}} & \multicolumn{2}{@{}c@{}}{\textbf{Spanned heading\tnote{2}}} \\\cmidrule{2-3}\cmidrule{4-5}
-\textbf{col1 head} & \textbf{col2 head}  & \textbf{col3 head}  & \multicolumn{1}{@{}l@{}}{\textbf{col4 head}}  & \textbf{col5 head}   \\
-\midrule
-col1 text & col2 text  & col3 text  & 12.34  & col5 text\tnote{1}   \\
-col1 text & col2 text  & col3 text  & 1.62  & col5 text\tnote{2}   \\
-col1 text & col2 text  & col3 text  & 51.809  & col5 text   \\
-\bottomrule
-\end{tabular*}
-\begin{tablenotes}%%[341pt]
-\item Source: Example for table source text.
-\item[1] Example for a first table footnote.
-\item[2] Example for a second table footnote.
-\end{tablenotes}
-\end{table*}
-\end{center}
+In this simple setting, the regression models are much closer to the
+Optimal model, while the neural network models perform worse than the KM
+null model. The wide confidence bands in \(C_{IPCW}\) suggest the neural
+network models may be over-parameterized.
 
-Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
-venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
-quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
-facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
-vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
-Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.
+\hypertarget{complex-simulation-flexible-baseline-hazard-time-varying-interactions}{%
+\subsection{Complex simulation: flexible baseline hazard, time-varying
+interactions}\label{complex-simulation-flexible-baseline-hazard-time-varying-interactions}}
 
-\begin{center}
-\begin{table}[t]%
-\centering
-\caption{This is sample table caption.\label{tab2}}%
-\begin{tabular*}{500pt}{@{\extracolsep\fill}lcccc@{\extracolsep\fill}}
-\toprule
-\textbf{col1 head} & \textbf{col2 head}  & \textbf{col3 head}  & \textbf{col4 head}  & \textbf{col5 head} \\
-\midrule
-col1 text & col2 text  & col3 text  & col4 text  & col5 text\tnote{$\dagger$}   \\
-col1 text & col2 text  & col3 text  & col4 text  & col5 text   \\
-col1 text & col2 text  & col3 text  & col4 text  & col5 text\tnote{$\ddagger$}   \\
-\bottomrule
-\end{tabular*}
-\begin{tablenotes}
-\item Source: Example for table source text.
-\item[$\dagger$] Example for a first table footnote.
-\item[$\ddagger$] Example for a second table footnote.
-\end{tablenotes}
+This simulation demonstrates performance with the presence of a complex
+baseline hazard and a time-varying interaction. Originally used to show
+the spline-based hazard model proposed by Royston and Parmar
+\citep{royston2002flexible}, the breast cancer dataset provides a
+complex hazard from which we simulate, available in the
+\textbf{flexsurv} R package \citep{flexsurv}. To increase the complexity
+of our data-generating mechanism for this simulation, we design the
+model as follows: \begin{align}
+\log h(t \mid X_i) =\sum_{i=1}^{5} (\gamma_{i} \cdot \psi_{i}) + \beta_{{1}} (z_{1}) + \beta_{{2}} (z_{2})+ \beta_{{3}} (z_{3})+ \tau_{1} ( z_{1} \cdot z_{2} \cdot t)+ \tau_{2} ( z_{1} \cdot z_{3})+ \tau_{3} (z_{2} \cdot z_{3}), \nonumber
+\end{align}
+
+\begin{table}
+\caption{Four tables representing performance at certain follow-up times for the simple simulation, complex simulation, SUPPORT and METABRIC.
+Each table shows performance for each method in each study at $25\%$, $50\%$.
+$75\%$ and $100\%$ of follow-up time.
+The bold elements show the best model for each study, at each follow-up time of interest.
+These tables are included to provide exact measures at certain intervals.
+The models of interest are: Cox, case-base with logistic regression (CBLR), DeepSurv, DeepHit, Case-Base Neural Network (CBNN), Optimal and Deep Survival Machines (DSM).}
+\label{tab:megaTable}
+
+\begin{center}\includegraphics[width=1\linewidth]{../../../analyses/figures/megaTable} \end{center}
+
 \end{table}
-\end{center}
 
-Below is the example~\cite{Rothermel1998,Yoo2007,Schulz2012} for bulleted list.
-Below is the example for bulleted list.
-Below is the example for bulleted list.
-Below is the example for bulleted list.
-Below is the example for bulleted list.
-Below is the example for bulleted list\footnote{This is an example for footnote.}:
-\begin{itemize}
-\item bulleted list entry sample bulleted list entry~\cite{Allen2011} sample list entry text.
-\item bulleted list entry sample bulleted list entry. bulleted list entry sample bulleted list entry. bulleted list entry sample bulleted list entry.
-\item bulleted list entry sample bulleted list entry~\cite{Ballen2011} bulleted list entry sample bulleted list entry~\citet{Allen2011} sample list entry text.
-bulleted list entry sample bulleted list entry.
-\item sample list entry text. sample list entry text.
-\end{itemize}
+where
+\(\gamma_{1}=3.9, \gamma_{2}=3, \gamma_{3}=-0.43, \gamma_{4}=1.33,\gamma_{5}=-0.86, \beta_{{1}}=1, \beta_{{2}}=1, \beta_{{3}}=1, \tau_{1}=10, \tau_{2}=2, \tau_{3}=2\)
+and \(\psi\) are basis splines. The \(gamma\) coefficients are obtained
+from an intercept-only cubic splines model with three knots using the
+\emph{flexsurvspline} function from the \textbf{flexsurv} package
+\citep{flexsurv}. Note that we fix these values for the analysis. The
+\(\beta\) coefficients represent direct effects, \(\tau_{2}\) and
+\({\tau_3}\) represent interactions and \(\tau_{1}\) is a time-varying
+interaction.
 
-Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
-ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
-eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
-convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
-pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.
+\hypertarget{performance-comparison-in-complex-simulation}{%
+\subsubsection{Performance comparison in complex
+simulation}\label{performance-comparison-in-complex-simulation}}
 
-Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non
-odio. Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce
-sed justo eu urna porta tincidunt.
-Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat.
-Morbi quis dolor.
-Donec pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus.
-Proin
-et quam. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien
-turpis, fermentum vel, eleifend faucibus, vehicula eu, lacus.
+Figure \ref{fig:megaPlot} C, D and Table \ref{tab:megaTable} B show the
+performance over time on a test set in the complex simulation. Apart
+from the Optimal regression model, CBNN outperforms the competing models
+when examining IPA and up to the 75-percentile of follow-up time for
+\(C_{IPCW}\). We expected the Optimal model to perform best in both
+metrics. However, this was not the case for \(C_{IPCW}\) and may be due
+to an artifact of concordance-based metrics, where a misspecified model
+may perform better than a correctly specified one
+\citep{cindexfails2019}. We attribute the performance of CBNN to its
+flexibility in modeling time-varying interactions and baseline hazard,
+flexibility the other neural network models do not have.
 
-Below is the sample for description list.
-Below is the example for description list.
-Below is the example for description list.
-Below is the example for description list.
-Below is the example for description list.
-Below is the example for description list:\\[12pt]
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%section4
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\noindent\textbf{Description sample:}
+\hypertarget{casestudies}{%
+\section{Application to SUPPORT and METABRIC data}\label{casestudies}}
 
-\begin{description}
-\item[first entry] description text. description text. description text. description text. description text. description text. description text.
-\item[second long entry] description text. description text. description text. description text. description text. description text. description text.
-\item[third entry] description text. description text. description text. description text. description text.
-\item[fourth entry] description text. description text.
-\end{description}
+Our complex simulation demonstrates the superior performance of CBNN in
+ideal conditions with clean data. To obtain a more realistic performance
+assessment, we compared models using two real datasets with a
+time-to-event outcome. The first case study examines the SUPPORT dataset
+\citep{knaus1995SUPPORT}. The second case study examines the METABRIC
+dataset \citep{curtis2012genomic}. We use the same hyperparameters as in
+the simulation studies. As we do not know the true model for the real
+data, we exclude the Optimal model. We split the datasets keeping 20\%
+of the observations as a test set. 20\% of the training set is kept
+aside for validation at each epoch. We predict risk functions for
+everyone in the test set, which is used to calculate our metrics. We
+conduct 100 bootstrap re-samples for the real data applications to
+obtain confidence intervals.
 
-\noindent\textbf{Numbered list items sample:}
+\hypertarget{performance-evaluation-using-the-support-dataset}{%
+\subsection{Performance evaluation using the SUPPORT
+dataset}\label{performance-evaluation-using-the-support-dataset}}
 
-\begin{enumerate}[1.]
-\item First level numbered list entry. sample numbered list entry.
+The SUPPORT dataset tracks the time until death for seriously ill
+patients at five American hospitals \citep{knaus1995SUPPORT}. We use a
+pre-processed version of the dataset made available in the DeepSurv
+package \citep{katzman2018DeepSurv}. This dataset contains 9104 samples
+and 14 covariates (age, sex, race, number of comorbidities, presence of
+diabetes, presence of dementia, presence of cancer, mean arterial blood
+pressure, heart rate, respiration rate, temperature, white blood cell
+count, serum's sodium and serum's creatinine)
+\citep{katzman2018DeepSurv}. Patients with missing features were
+excluded and 68.10\% of the patients died during the 5.56-year study
+period \citep{katzman2018DeepSurv}.
 
-\item First numbered list entry. sample numbered list entry. Numbered list entry. sample numbered list entry. Numbered list entry. sample numbered list entry.
+Figure \ref{fig:megaPlot} E, F and Table \ref{tab:megaTable} C
+demonstrates the performance over time on a test set. The regression
+models (CBLR, Cox) perform best considering IPA, followed by CBNN from
+the \(25^{th}\) to \(100^{th}\) percentile of follow-up time. We note a
+drop in performance for CBNN before the \(25^{th}\) percentile of
+follow-up. For \(C_{IPCW}\), CBNN outperforms the competing models
+consistently over follow-up time. Note that performance is similar for
+all models, aside from DeepSurv whose \(C_{IPCW}\) is lower than the
+rest (Figure \ref{fig:megaPlot} E, F and Table \ref{tab:megaTable} C).
 
-\begin{enumerate}[a.]
-\item Second level alpabetical list entry. Second level alpabetical list entry. Second level alpabetical list entry~\citet{Allen2011} Second level alpabetical list entry.
+\hypertarget{performance-evaluation-using-the-metabric-dataset}{%
+\subsection{Performance evaluation using the METABRIC
+dataset}\label{performance-evaluation-using-the-metabric-dataset}}
 
-\item Second level alpabetical list entry. Second level alpabetical list entry~\citet{Schulz2012,Allen2011,Ballen2011}.
+METABRIC is a 30-yearlong study aiming to discover the molecular drivers
+of breast tumors, following 2000 individuals with breast cancer until
+death \citep{curtis2012genomic}. They described these growths as primary
+invasive breast carcinomas, with the goal of discovering both genetic
+and clinical risk factors for breast cancer survival
+\citep{curtis2012genomic}. We used the processed dataset made available
+through DeepSurv \citep{katzman2018DeepSurv}, which includes 1980
+patients of which 57.72\% die due to breast cancer within a median 10
+years of follow-up \citep{katzman2018DeepSurv}. There are 10 covariates
+in total: time, 4 RNA-Seq gene expressions (MKI67, EGFR, PGR and ERBB2)
+and 5 clinical features (hormone treatment indicator, radiotherapy
+indicator, chemotherapy indicator, ER-positive indicator and age at
+diagnosis) \citep{katzman2018DeepSurv}.
 
-\begin{enumerate}[ii.]
-\item Third level lowercase roman numeral list entry. Third level lowercase roman numeral list entry. Third level lowercase roman numeral list entry.
+Figure \ref{fig:megaPlot} G, H and Table \ref{tab:megaTable} D shows the
+performance on a test set over time. The IPA scores suggest that
+regression models outperform competing models on this dataset, as all
+the neural network models are equal to or perform worse than KM over
+follow-up time. Our CBNN model is comparable to KM until around the
+50-percentile of follow-up time, after which CBNN, DeepSurv and DSM drop
+in performance. The \(C_{IPCW}\) produces a different ranking. Our CBNN
+model outperforms the other models up to the \(25^{th}\) percentile of
+follow-up time, whereas DeepHit performs best from the \(50^{th}\) to
+\(75^{th}\) percentile of follow-up. With this metric, CBNN and DeepHit
+outperform the regression models. This disagreement between IPA and
+\(C_{IPCW}\) may be due to the misspecification issue of
+concordance-based metrics \citep{cindexfails2019}. The neural network
+models may be over-parameterized, as shown by the wide confidence bands
+in \(C_{IPCW}\).
 
-\item Third level lowercase roman numeral list entry. Third level lowercase roman numeral list entry~\cite{Yoo2007}.
-\end{enumerate}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%section6
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\item Second level alpabetical list entry. Second level alpabetical list entry~\cite{Elbaum2002}.
-\end{enumerate}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
-\item First level numbered list entry. sample numbered list entry. Numbered list entry. sample numbered list entry. Numbered list entry.
+CBNN models survival outcomes by using neural networks on case-base
+sampled data. We incorporate follow-up time as a feature, providing a
+data-driven estimate of a flexible baseline hazard and time-varying
+interactions in our hazard function. The three competing neural network
+models we evaluated cannot model time-varying interactions by design
+\citep{dsmPaper} \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}. DSM
+requires a mixture component distributions for a flexible baseline
+hazard to be fit \citep{dsmPaper}. As our goal is to fix the design and
+compare performance, we did not change any shared hyperparameters. With
+our choice of shared hyperparameters and model design, DSM did not
+converge in the complex simulation and SUPPORT case study. Despite this
+limitation, we include this method when it did function as it can fit
+flexible baseline hazards. Compared to CBNN, the remaining two models
+also have limitations. DeepSurv is a proportional hazards model and does
+not estimate the baseline hazard \citep{katzman2018DeepSurv}. DeepHit
+requires an alpha hyperparameter, is restricted to a single distribution
+for the baseline hazard and models the survival function directly
+\citep{lee2018DeepHit}. The alternative neural network methods match on
+time, while CBNN models time directly.
 
-\item Another first level numbered list entry. sample numbered list entry. Numbered list entry. sample numbered list entry. Numbered list entry.
-\end{enumerate}
+To assess performance among these models, we use both IPA and
+\(C_{IPCW}\) metrics. Concordance-based measures are commonly used in
+survival analysis to compare models and we opt to keep them in our
+analyses. However, \(C_{IPCW}\) is a non-proper metric and may cause
+misspecified models to appear better than they should
+\citep{cindexfails2019}. Therefore, we contextualize our \(C_{IPCW}\)
+results in relation to IPA, a proper scoring rule. The model rankings
+between \(C_{IPCW}\) and IPA differed for both the complex simulation
+and METABRIC application. Wider \(C_{IPCW}\) confidence intervals for
+DeepHit and DeepSurv show potential misspecification of the models
+(Figure \ref{fig:megaPlot} D, H).
 
-\noindent\textbf{un-numbered list items sample:}
+With this interpretation of \(C_{IPCW}\) in mind, we assess two
+simulations with minimal noise and two case studies in real scenarios.
+The simple simulation demonstrates potential pitfalls associated with
+neural network models, particularly overparameterization. All neural
+network-based approaches performed worse than null KM model (IPA score),
+while the regression-based performed better. We attribute this to
+potential overparameterization in the neural network models, as the wide
+confidence intervals suggest over-fitting, even with dropout. Both
+DeepSurv and DSM are affected, while DeepHit and CBNN are less so. This
+is a limitation to our strategy for evaluating the methods, which uses a
+fixed study design across all assessments.
 
-\begin{enumerate}[]
-\item Sample unnumberd list text..
-\item Sample unnumberd list text.
-\item sample unnumberd list text.
-\item Sample unnumberd list text.
-\end{enumerate}
+From this baseline benchmark, we move on to a complex simulation that
+requires a method that can learn both time-varying interactions and have
+a flexible baseline hazard. Here, CBNN demonstrates a distinct advantage
+over all other methods. The regression models show improved performance
+over the null KM model, while the competing neural network models
+performed worse. Based on our complex simulation results (Figure
+\ref{fig:megaPlot} C, D and Table \ref{tab:megaTable} B), CBNN
+outperforms the competitors when time-varying interactions and a complex
+baseline hazard are present. This simulation shows how CBNN can perform
+under ideal conditions, while the following two analyses on real data
+serve to assess its performance in realistic conditions.
 
-\section{Examples for enunciations}\label{sec4}
+In the SUPPORT and METABRIC case studies, flexibility in both
+interaction modeling and baseline hazard did not improve CBNNs relative
+performance, suggesting that this flexibility does not aid prediction in
+either case study. From a biological perspective, the 30-year follow-up
+time in the METABRIC study may contain competing causes of death. The
+causes at the start of the study may not match the causes towards the
+end, potentially explaining the drop in performance as we reach later
+survival times with fewer individuals. The baseline hazard is unlikely
+to cause this drop as DSM is the most flexible competing model in our
+comparison. Over-fitting is also unlikely given the tight confidence
+intervals. Further research is required to determine the cause of the
+drop in performance seen in the METABRIC case study. In both case
+studies, CBNN outperforms the competing neural network methods.
 
-\begin{theorem}[Theorem subhead]\label{thm1}
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-\end{theorem}
-
-Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
-amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
-Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
-quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
-vitae risus porta vehicula.
-
-Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
-venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
-quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
-facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
-vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
-Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.
-
-\begin{proposition}
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-\end{proposition}
-
-Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
-viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
-ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
-nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
-bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
-pellentesque felis eu massa.
-
-Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
-amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
-Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
-quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
-vitae risus porta vehicula.
-
-\begin{definition}[Definition sub head]
-Example definition text.
-Example definition text.
-Example definition text.
-Example definition text.
-Example definition text.
-Example definition text.
-Example definition text.
-Example definition text.
-Example definition text.
-Example definition text.
-Example definition text.
-\end{definition}
-
-Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non
-odio. Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce
-sed justo eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor.
-Donec pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin
-et quam. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien
-turpis, fermentum vel, eleifend faucibus, vehicula eu, lacus.
-
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
-dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
-vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
-Pellentesque eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut
-mi mi, lacinia sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui.
-Mauris nibh leo, facilisis non, adipiscing quis, ultrices a, dui.
-
-\begin{proof}
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-\end{proof}
-
-Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
-pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
-vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
-a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
-urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.
-
-Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
-viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
-ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
-nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
-bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
-pellentesque felis eu massa.
-
-\begin{proof}[Proof of Theorem~\ref{thm1}]
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-\end{proof}
-
-Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
-eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
-facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
-a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
-convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.
-Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra tortor. Sed fringilla
-justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue est scelerisque
-sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas urna mi, suscipit
-in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.
-
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
-dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
-vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
-Pellentesque eget lectus.
-Proin eu metus.
-Sed porttitor.
-In hac habitasse platea dictumst.
-Suspendisse eu lectus.
-Ut Curabitur tellus magna, porttitor a, commodo a, commodo in, tortor.
-Donec interdum.
-Praesent scelerisque.
-Mae-
-cenas posuere sodales odio. Vivamus metus lacus, varius quis, imperdiet quis, rhoncus a, turpis. Etiam ligula arcu,
-elementum a, venenatis quis, sollicitudin sed, metus. Donec nunc pede, tincidunt in, venenatis vitae, faucibus vel,
-
-\begin{sidewaystable}%[h]
-\caption{Sideways table caption. For decimal alignment refer column 4 to 9 in tabular* preamble.\label{tab3}}%
-\begin{tabular*}{\textheight}{@{\extracolsep\fill}lccD{.}{.}{4}D{.}{.}{4}D{.}{.}{4}D{.}{.}{4}D{.}{.}{4}D{.}{.}{4}@{\extracolsep\fill}}%
-\toprule
-  & \textbf{col2 head} & \textbf{col3 head} & \multicolumn{1}{c}{\textbf{10}} &\multicolumn{1}{c}{\textbf{20}} &\multicolumn{1}{c}{\textbf{30}} &\multicolumn{1}{c}{\textbf{10}} &\multicolumn{1}{c}{\textbf{20}} &\multicolumn{1}{c}{\textbf{30}} \\
-\midrule
-  &col2 text &col3 text &0.7568&1.0530&1.2642&0.9919&1.3541&1.6108 \\
-  & &col2 text &12.5701 &19.6603&25.6809&18.0689&28.4865&37.3011 \\
-3 &col2 text  & col3 text &0.7426&1.0393&1.2507&0.9095&1.2524&1.4958 \\
-  & &col3 text &12.8008&19.9620&26.0324&16.6347&26.0843&34.0765 \\
-  & col2 text & col3 text &0.7285&1.0257&1.2374&0.8195&1.1407&1.3691\tnote{*} \\
-  & & col3 text &13.0360&20.2690&26.3895&15.0812&23.4932&30.6060\tnote{\dagger} \\
-\bottomrule
-\end{tabular*}
-\begin{tablenotes}%%[\textheight]
-\item[*] First sideways table footnote. Sideways table footnote. Sideways table footnote. Sideways table footnote.
-\item[$\dagger$] Second sideways table footnote. Sideways table footnote. Sideways table footnote. Sideways table footnote.
-\end{tablenotes}
-\end{sidewaystable}
-
-\begin{sidewaysfigure}
-\centerline{\includegraphics[width=542pt,height=9pc,draft]{empty}}
-\caption{Sideways figure caption. Sideways figure caption. Sideways figure caption. Sideways figure caption. Sideways figure caption. Sideways figure caption.\label{fig3}}
-\end{sidewaysfigure}
-
-nibh. Pellentesque wisi. Nullam malesuada. Morbi ut tellus ut pede tincidunt porta. Lorem ipsum dolor sit amet,
-consectetuer adipiscing elit. Etiam congue neque id dolor.
-
-\begin{algorithm}
-\caption{Pseudocode for our algorithm}\label{alg1}
-\begin{algorithmic}
-  \For each frame
-  \For water particles $f_{i}$
-  \State compute fluid flow~\cite{Rothermel1997}
-  \State compute fluid--solid interaction~\cite{Allen2011}
-  \State apply adhesion and surface tension~\cite{Ballen2011}
-  \EndFor
-   \For solid particles $s_{i}$
-   \For neighboring water particles $f_{j}$
-   \State compute virtual water film \\(see Section~\ref{sec3})
-   \EndFor
-   \EndFor
-   \For solid particles $s_{i}$
-   \For neighboring water particles $f_{j}$
-   \State compute growth direction vector \\(see Section~\ref{sec2})
-   \EndFor
-   \EndFor
-   \For solid particles $s_{i}$
-   \For neighboring water particles $f_{j}$
-   \State compute $F_{\theta}$ (see Section~\ref{sec1})
-   \State compute $CE(s_{i},f_{j})$ \\(see Section~\ref{sec3})
-   \If $CE(b_{i}, f_{j})$ $>$ glaze threshold
-   \State $j$th water particle's phase $\Leftarrow$ ICE
-   \EndIf
-   \If $CE(c_{i}, f_{j})$ $>$ icicle threshold
-   \State $j$th water particle's phase $\Leftarrow$ ICE
-   \EndIf
-   \EndFor
-   \EndFor
-  \EndFor
-\end{algorithmic}
-\end{algorithm}
-
-Donec et nisl at wisi luctus bibendum. Nam interdum tellus ac libero. Sed sem justo, laoreet vitae, fringilla at,
-adipiscing ut, nibh. Maecenas non sem quis tortor eleifend fermentum. Etiam id tortor ac mauris porta vulputate.
-Integer porta neque vitae massa~\cite{Rothermel1997,Elbaum2002}. Maecenas tempus libero a libero posuere dictum. Vestibulum ante ipsum primis in
-faucibus orci luctus et ultrices posuere cubilia Curae; Aenean quis mauris sed elit commodo placerat. Class aptent
-taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Vivamus rhoncus tincidunt libero.
-Etiam elementum pretium justo. Vivamus est. Morbi a tellus eget pede tristique commodo~\cite{Elbaum2002} Nulla nisl. Vestibulum
-sed nisl eu sapien cursus rutrum.
-
-Pellentesque wisi. Nullam malesuada. Morbi ut tellus ut pede tincidunt porta. Lorem ipsum dolor sit amet,
-consectetuer adipiscing elit. Etiam congue neque id dolor.
-
-Donec et nisl at wisi luctus bibendum. Nam interdum tellus ac libero. Sed sem justo, laoreet vitae, fringilla at,
-adipiscing ut, nibh. Maecenas non sem quis tortor eleifend fermentum. Etiam id tortor ac mauris porta vulputate.
-Integer porta neque vitae massa. Maecenas tempus libero a libero posuere dictum. Vestibulum ante ipsum primis in
-faucibus orci luctus et ultrices posuere cubilia Curae; Aenean quis mauris sed elit commodo placerat. Class aptent
-taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Vivamus rhoncus tincidunt libero.
-Etiam elementum pretium justo. Vivamus est. Morbi a tellus eget pede tristique commodo. Nulla nisl. Vestibulum
-sed nisl eu sapien cursus rutrum.
-
-\begin{equation}\label{eq23}
- \|\tilde{X}(k)\|^2
- =\frac{\left\|\sum\limits_{i=1}^{p}\tilde{Y}_i(k)+\sum\limits_{j=1}^{q}\tilde{Z}_j(k) \right\|^2}{(p+q)^2}
- \leq\frac{\sum\limits_{i=1}^{p}\left\|\tilde{Y}_i(k)\right\|^2+\sum\limits_{j=1}^{q}\left\|\tilde{Z}_j(k)\right\|^2 }{p+q}.
-\end{equation}
-
-Sed feugiat. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Ut pellentesque
-augue sed urna. Vestibulum diam eros, fringilla et, consectetuer eu, nonummy id, sapien. Nullam at lectus. In sagittis
-ultrices mauris. Curabitur malesuada erat sit amet massa. Fusce blandit. Aliquam erat volutpat. Aliquam euismod.
-Aenean vel lectus. Nunc imperdiet justo nec dolor.
-
-Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
-eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
-facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
-a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
-convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.
-
-\begin{equation}\label{eq24}
- \|\tilde{X}(k)\|^2
- =\frac{\left\|\sum\limits_{i=1}^{p}\tilde{Y}_i(k)+\sum\limits_{j=1}^{q}\tilde{Z}_j(k) \right\|^2}{(p+q)^2}
- \leq\frac{\sum\limits_{i=1}^{p}\left\|\tilde{Y}_i(k)\right\|^2+\sum\limits_{j=1}^{q}\left\|\tilde{Z}_j(k)\right\|^2 }{p+q}.
-\end{equation}
-
-Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra
-tortor. Sed fringilla justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue
-est scelerisque sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas
-urna mi, suscipit in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.
-
-Etiam ac leo a risus tristique nonummy. Donec dignissim tincidunt nulla. Vestibulum rhoncus molestie odio. Sed
-lobortis, justo et pretium lobortis, mauris turpis condimentum augue, nec ultricies nibh arcu pretium enim. Nunc
-purus neque, placerat id, imperdiet sed, pellentesque nec, nisl. Vestibulum imperdiet neque non sem accumsan laoreet.
-In hac habitasse platea dictumst. Etiam condimentum facilisis libero. Suspendisse in elit quis nisl aliquam dapibus.
-Pellentesque auctor sapien. Sed egestas sapien nec lectus. Pellentesque vel dui vel neque bibendum viverra. Aliquam
-porttitor nisl nec pede. Proin mattis libero vel turpis. Donec rutrum mauris et libero. Proin euismod porta felis.
-Nam lobortis, metus quis elementum commodo, nunc lectus elementum mauris, eget vulputate ligula tellus eu neque.
-Vivamus eu dolor.
+Taken together, CBNN outperforms all competitors in the complex
+simulation, demonstrating its value in survival settings that may
+involve time-varying interactions and a complex baseline hazard. Once we
+perform case-base sampling and adjust for the sampling bias, we can use
+a sigmoid activation function to predict our hazard function. Our
+approach simplifies the incorporation of censored individuals, allowing
+survival outcomes to be treated as binary ones. Forgoing the requirement
+of custom loss functions, CBNN only requires the use of standard
+components in machine learning libraries (specifically, the add layer to
+adjust for sampling bias and the sigmoid activation function). Due to
+the simplicity in its implementation and by extension user experience,
+CBNN is both a user-friendly approach to data-driven survival analysis
+and is easily extendable to any feed-forward neural network framework.
 
@@
 
-\section*{Acknowledgments}
-This is acknowledgment text~\cite{Elbaum2002}.
-Provide text here.
-This is acknowledgment text.
-Provide text here.
-This is acknowledgment text.
-Provide text here.
-This is acknowledgment text.
-Provide text here.
-This is acknowledgment text.
-Provide text here.
-This is acknowledgment text.
-Provide text here.
-This is acknowledgment text.
-Provide text here.
-This is acknowledgment text.
-Provide text here.
-This is acknowledgment text.
-Provide text here.
+\hypertarget{data-and-code-availability-statement}{%
+\subsection*{Data and code availability
+statement}\label{data-and-code-availability-statement}}
+\addcontentsline{toc}{section}{Data and code availability statement}
+
+The pre-processed data for the SUPPORT case study can be found at
+\url{https://github.com/jaredleekatzman/DeepSurv/tree/master/experiments/data/support}.
+The pre-processed data for the METABRIC case study can be found at
+\url{https://github.com/jaredleekatzman/DeepSurv/tree/master/experiments/data/metabric}.
+The data is accessed using \url{https://github.com/havakv/pycox}. The
+code for this manuscript and its analyses can be found at
+\url{https://github.com/Jesse-Islam/cbnnManuscript}. The software
+package making CBNN easier to use can be found at
+\url{https://github.com/Jesse-Islam/cbnn}.
+
+\hypertarget{acknowledgements}{%
+\subsection*{Acknowledgements}\label{acknowledgements}}
+\addcontentsline{toc}{section}{Acknowledgements}
+
+We would like to thank Dr.~James Meigs, the project
+leader of these awards, for his support and helpful discussions. The
+work was also supported as part of the Congressionally Directed Medical
+Research Programs (CDMRP) award W81XWH-17-1-0347. We would also like to
+thank Dr.~James Hanley for his support and discussions while extending
+the case-base methodology.
 
@@
 
-None reported.
+This work was supported by subcontracts from UM1DK078616 and R01HL151855
+to Dr.~Rob Sladek.
 
@@
 
-\section*{Supporting information}
-
-The following supporting information is available as part of the online article:
-
-\noindent
-\textbf{Figure S1.}
-{500{\uns}hPa geopotential anomalies for GC2C calculated against the ERA Interim reanalysis. The period is 1989--2008.}
-
-\noindent
-\textbf{Figure S2.}
-{The SST anomalies for GC2C calculated against the observations (OIsst).}
-
 \appendix
-
-\section{Section title of first appendix\label{app1}}
-
-Use \verb+\begin{verbatim}...\end{verbatim}+ for program codes without math.
-Use \verb+\begin{alltt}...\end{alltt}+ for program codes with math.
-Based on the text provided inside the optional argument of \verb+\begin{code}[Psecode|Listing|Box|Code|+\hfill\break \verb+Specification|Procedure|Sourcecode|Program]...+ \verb+\end{code}+ tag corresponding boxed like floats are generated.
-Also note that \verb+\begin{code}[Code|Listing]...+ \verb+\end{code}+ tag with either Code or Listing text as optional argument text are set with computer modern typewriter font.
-All other code environments are set with normal text font.
-Refer below example:
-
-\begin{lstlisting}[caption={Descriptive Caption Text},label=DescriptiveLabel]
-for i:=maxint to 0 do
-begin
-{ do nothing }
-end;
-Write('Case insensitive ');
-WritE('Pascal keywords.');
-\end{lstlisting}
-
-\subsection{Subsection title of first appendix\label{app1.1a}}
-
-Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
-pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
-vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
-a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
-urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.
-
-Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
-viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
-ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
-nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
-bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
-pellentesque felis eu massa.
-
-\subsubsection{Subsection title of first appendix\label{app1.1.1a}}
-
-\noindent\textbf{Unnumbered figure}
-
-\begin{center}
-\includegraphics[width=7pc,height=8pc,draft]{empty}
-\end{center}
-
-Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
-venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
-quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
-facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
-vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
-
-Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.
-Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
-ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
-eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
-convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
-pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.
-
-\section{Section title of second appendix\label{app2}}%
-
-Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
-venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
-quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
-facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
-vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
-
-Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.
-Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
-ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
-eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
-convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
-pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.
-
-%== Figure 4 ==
-%% Example for figure inside appendix
-\begin{figure}[t]
-\centerline{\includegraphics[height=10pc,width=78mm,draft]{empty}}
-\caption{This is an example for appendix figure.\label{fig5}}
-\end{figure}
-
-\subsection{Subsection title of second appendix\label{app2.1a}}
-
-Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non odio.
-Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce sed justo
-eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor. Donec
-pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin et quam.
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien turpis,
-fermentum vel, eleifend faucibus, vehicula eu, lacus.
-
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
-dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
-vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
-Pellentesque eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut
-mi mi, lacinia sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui.
-Mauris nibh leo, facilisis non, adipiscing quis, ultrices a, dui.
-
-\subsubsection{Subsection title of second appendix\label{app2.1.1a}}
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae,
-felis. Curabitur dictum gravida mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna. Donec
-vehicula augue eu neque. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
-Mauris ut leo. Cras viverra metus rhoncus sem. Nulla et lectus vestibulum urna fringilla ultrices. Phasellus eu tellus
-sit amet tortor gravida placerat. Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. Praesent eget sem vel
-leo ultrices bibendum. Aenean faucibus. Morbi dolor nulla, malesuada eu, pulvinar at, mollis ac, nulla. Curabitur
-auctor semper nulla. Donec varius orci eget risus. Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
-
-Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
-pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
-vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
-a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
-urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.
-
-\begin{center}
-\begin{table}[b]%
-\centering
-\caption{This is an example of Appendix table showing food requirements of army, navy and airforce.\label{tab4}}%
-\begin{tabular*}{300pt}{@{\extracolsep\fill}lcc@{\extracolsep\fill}}%
-\toprule
-\textbf{col1 head} & \textbf{col2 head} & \textbf{col3 head} \\
-\midrule
-col1 text & col2 text & col3 text \\
-col1 text & col2 text & col3 text \\
-col1 text & col2 text & col3 text\\
-\bottomrule
-\end{tabular*}
-\end{table}
-\end{center}
-
-Example for an equation inside appendix
-\begin{equation}
-\mathcal{L}\quad \mathbf{\mathcal{L}} = i \bar{\psi} \gamma^\mu D_\mu \psi - \frac{1}{4} F_{\mu\nu}^a F^{a\mu\nu} - m \bar{\psi} \psi\label{eq25}
-\end{equation}
-
-\section{Example of another appendix section\label{app3}}%
-
-This is sample for paragraph text this is sample for paragraph text  this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text
-
-Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
-pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
-vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
-a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
-urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.
-
-Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
-viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
-ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
-nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
-bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
-pellentesque felis eu massa.
-
-\begin{equation}
-\mathcal{L} = i \bar{\psi} \gamma^\mu D_\mu \psi
-    - \frac{1}{4} F_{\mu\nu}^a F^{a\mu\nu} - m \bar{\psi} \psi
-\label{eq26}
-\end{equation}
-
-Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
-viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
-ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
-nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
-bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
-pellentesque felis eu massa.
-
-Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
-amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
-Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
-quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
-vitae risus porta vehicula.
-
-\begin{center}
-\begin{tabular*}{300pt}{@{\extracolsep\fill}lcc@{\extracolsep\fill}}%
-\toprule
-\textbf{col1 head} & \textbf{col2 head} & \textbf{col3 head} \\
-\midrule
-col1 text & col2 text & col3 text \\
-col1 text & col2 text & col3 text \\
-col1 text & col2 text & col3 text \\
-\bottomrule
-\end{tabular*}
-\end{center}
-
-Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
-amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
-Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
-quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
-vitae risus porta vehicula.
-
-Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
-venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
-quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
-facilisi.
-Sed a turpis eu lacus commodo facilisis.
-Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, evehicula libero dui cursus dui.
-Mauris tempor ligula sed lacus.
-Duis cursus enim ut augue.
-Cras ac magna.
-Cras nulla.
-Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.
-
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
-dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
-vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
-Pellentesque eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut
-mi mi, lacinia sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui.
-Mauris nibh leo, facilisis non, adipiscing quis, ultrices a, dui.
-
-\nocite{*}% Show all bib entries - both cited and uncited; comment this line to view only cited bib entries;
+%\nocite{*}% Show all bib entries - both cited and uncited; comment this line to view only cited bib entries;
 \bibliography{wileyNJD-APA}%

### 2022-12-01 SIMPaper/SIMPaper.tex
added analysis

@@
 \(b/B\). Therefore, we have the following equality
-\footnote{We are abusing notation here, conflating hazards with probabilities. For a rigorous treatment, see Saarela \& Hanley (2015) section 3 \cite{saarela2015} .}:
+\footnote{We are abusing notation here, conflating hazards with probabilities. For a rigorous treatment, see Saarela \& Hanley (2015) section 3 \cite{saarela2015}.}:
 \begin{align}\label{eqn:main}
 \frac{P\left(Y_i=1 \mid X_i, T_i\right)}{P\left(Y_i = 0 \mid X_i, T_i\right)} = \frac{h\left(T_i \mid X_i\right)}{b/B}.
-\end{align} The odds of a person-moment being a part of the \emph{case
-series} is the ratio of the hazard \(h(T_i \mid X_i)\) and the uniform
-rate \(b/B\). Using \eqref{eqn:main}, we can see how the log-hazard
-function can be estimated from the log-odds arising from case-base
-sampling: \begin{align}\label{eqn:offset}
+\end{align}
+
+To provide some intuition for this equation, the hazard is the rate at
+which the event (\(Y_{i}=1\)) occurs at a given time, conditional on
+your survival up to the follow-up-time of interest. As we are looking at
+person-moments, we may retrieve the rate at which an event does not
+occur (\(Y_{i}=0)\) at a given time by taking the number of samples in
+\(b\) and dividing it by the total number of moments in the study base
+\(B\) at the current follow-up-time. Though an intuitive approach is
+used to describe this equation, a complete derivation of this equation
+can be found at \citep{saarela2015}.
+
+The odds of a person-moment being a part of the \emph{case series} is
+the ratio of the hazard \(h(T_i \mid X_i)\) and the uniform rate
+\(b/B\). Sampling techniques are known to result in a loss of
+information. To address this concern, the original case-base methodology
+paper for logistic regression uses a measure for relative information
+when comparing two averages, based on the size of our base-series and
+case-series. If we let the size of \(b=100c\), then the variances and
+covariances are expected to be proportional to
+\(\frac{1}{c}+\frac{1}{100c}\) rather than
+\(\frac{1}{c}+\frac{1}{\inf}\), about 1 percent larger. Concerns were
+raised about dependency due to multiple person-moments coming from each
+individual, the information loss was studied in Saarela 2015, where they
+found a ratio of 100:1, b:c, give an efficiency comparable to
+Mantel-Haenszel and conditional logistic regression methods (Saarela
+2015). Using \eqref{eqn:main}, we can see how the log-hazard function
+can be estimated from the log-odds arising from case-base sampling:
+\begin{align}\label{eqn:offset}
 \log \left( h\left(t \mid X_i\right)\right) = \log \left(\frac{P\left(Y_i = 1 \mid X_i, t\right)}{P\left(Y_i = 0 \mid X_i, t\right)}\right) + \log\left(\frac{b}{B}\right).
@@
 
+Sampling techniques are known to result in a loss of information. To
+address this concern, the original case-base methodology paper uses a
+measure for relative information when comparing two averages, based on
+the size of our base-series and case-series. If we let the size of
+\(b=100c\), then the variances and covariances are expected to be
+proportional to \(\frac{1}{c}+\frac{1}{100c}\) rather than
+\(\frac{1}{c}+\frac{1}{\inf}\), about 1 percent larger. If there is a
+concern with a 1 percent inflation of these estimates, a larger ratio
+than 100 may be used. As there are concerns about dependency due
+multiple person-moments coming from each individual, the information
+loss was studied in Saarela 2015 (\citep{saarela2015}), where they found
+a ratio of 100:1 base-series to case series gives an efficiency
+comparable to Mantel-Haenszel and conditional logistic regression
+methods \citep{saarela2015}.
+
 To estimate the correct hazard function, we adjusting for the bias
-introduced when sampling a fraction of the study base \(B\) by adding
-the offset term \(\log\left(\frac{b}{B} \right)\) as in
-\eqref{eqn:offset}. Next, we propose using neural networks to model the
-odds.
+introduced when sampling a fraction of the study base \(B\) by adding an
+offset term \(\log\left(\frac{B}{b} \right)\) \eqref{eqn:offset}. Next,
+we propose using neural networks to model the odds.
 
@@
 then passed through a sigmoid activation function (Figure
-\ref{fig:NNarch}). We use the sigmoid function as its inverse is the
-odds, which we can use to calculate the hazard. The general form for the
-neural network using CBNN is:
+\ref{fig:NNarch}). We use the sigmoid activation function as it is the
+inverse of the log-odds and commonly available in popular neural network
+packages (tensorflow, keras, pytorch, etc) {[}REF?{]}. which we can use
+to calculate the hazard. The general form for the neural network using
+CBNN is:
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/home/bhatnagar-lab/jislam/cbnnManuscript/figures/nnarch2}
+{\centering \includegraphics[width=1\linewidth]{../figures/nnarch2}
 
@@
 censoring weights-adjusted concordance index (\(C_{IPCW}\))
-\citep{uno2011}, which we define below. Both these time-dependent
-metrics provide transparency as to when in follow-up time each model may
-perform better than the others.
+\citep{uno2011}, which we define below. We may be interested in an
+absolute risk a specific number of years in the future. For the sake of
+comparison, we plot each metric over time \(t\) so that the performance
+is transparent. As we go further along follow-up-time, we expect the
+performance of all models to drop as there are fewer individuals left in
+the study by then. For \(C_{IPCW}\), the score demonstrates performance
+up to time \(t\) \citep{uno2011}, while IPA is performance at time
+\(t\).
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/home/bhatnagar-lab/jislam/cbnnManuscript/analyses/figures/megaPlot}
+{\centering \includegraphics[width=1\linewidth]{../analyses/figures/megaPlot}
 
@@
 Each model-specific metric in each study shows a 95-percent confidence interval over 100 iterations.
-The models of interest are case-base with logistic regression (CBLR), Case-Base Neural Networks (CBNN), Cox, DeepHit, DeepSurv, Deep Survival Machines (DSM), Optimal (a CBLR model with the exact interaction terms and baseline hazard specified) and Kaplan-Meier (to serve as a baseline, predicting the average for all individuals).}\label{fig:megaPlot}
+The models of interest are case-base with logistic regression (CBLR), Case-Base Neural Networks (CBNN), Cox, DeepHit, DeepSurv, Deep Survival Machines (DSM), Optimal (a CBLR model with the exact interaction terms and baseline hazard specified) and Kaplan-Meier (to serve as a baseline, predicting the average for all individuals).
+Note than Cox and CBLR are superimposed.}\label{fig:megaPlot}
 \end{figure}
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{/home/bhatnagar-lab/jislam/cbnnManuscript/analyses/figures/megaTable} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{../analyses/figures/megaTable} \end{center}
 
@@
 
+The way neural networks are used in this paper are not directly
+interpretable. This is by design, as we only wish to compare to
+competitors based on predictive ability. All feed-forward neural network
+models can make use of different architectures to improve performance or
+interpretability. While a deep neural network as used in this paper is
+not directly interpretable, techniques such as Local Interpretable
+Model-Agnostic Explanations (LIME) \citep{ribeiro2016model} may be used
+to interpret the relevance of each feature to a given prediction. An
+example of architectural interpretability is a convolutional neural
+network for images, which provide positional information about the
+relevant pixels in an image \citep{keras}. Note that neither of these
+are specific to CBNN. As CBNN is concerned with learning time-varying
+interactions, a concern may be interpreting which time-varying
+interactions have been learned. Other secondary methods after the
+fitting process may be used to retrieve said information
+\url{https://ui.adsabs.harvard.edu/abs/2017arXiv170504977T/abstract}.
+
 Taken together, CBNN outperforms all competitors in the complex

### 2022-12-01 analyses/oldmortality.Rmd
added analysis

---
title: "oldmortality"
output: html_document
params:
  epo: 2000
  patience: 10
  iteration: 1
  layer1: 200
  layer2: 20
  layer3: 25
  layer4: 25
  drpt: 0.1
---

```{r,eval=T}

library(reticulate)
```

```{python,eval=T}

import os
os.environ["CUDA_DEVICE_ORDER"] = "PCI_BUS_ID" 
os.environ["CUDA_VISIBLE_DEVICES"] = ""
#os.environ["OM_NUM_THREADS"]="2"

```

```{r setup, include=T}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
source("src/packages.R")
source("src/functions.R")
eval_metabric<-T
pycox<-reticulate::import("pycox")
source_python("src/cIndex.py")
source_python("src/dsmScript.py")
source_python('src/oldmortDeephitter.py')
##########################
###Shared Hyperparameters
##########################
bsize=512
min_delta = 10^-9
epo=as.numeric(params$epo)
patience = as.numeric(params$patience)
iteration=as.numeric(params$iteration)
layer1<-as.numeric(params$layer1)
layer2<-as.numeric(params$layer2)
layer3<-as.numeric(params$layer3)
layer4<-as.numeric(params$layer4)
drpt<-as.numeric(params$drpt)

```

```{r testAnalysis}

start_time <- Sys.time()
BrierIPAList<-list()
cScore<-list()
library(eha)

data("oldmort")

preData<-data.frame(sex=oldmort$sex,
                 civ=oldmort$civ,
                 ses=oldmort$ses.50,
                 birthplace=oldmort$birthplace,
                 ageAtDeath=oldmort$exit,
                 imrAtBirth=oldmort$imr.birth,
                 region=oldmort$region,
                 birthdate=oldmort$birthdate,
                 time=oldmort$exit-oldmort$enter,
                 status=as.numeric(oldmort$event))

preData$time<-preData$time+runif(nrow(preData),-1,1)/1000000

data<-as.data.frame((model.matrix(~.,preData)[,-1]))

#data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
covs<-colnames(data)

samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.80)) 
trainOG<-data[samp,]
test<-data[-samp,-c(ncol(data))]

fullTest<-data[-samp,,drop=F]

pos<-1

while (pos <= iteration) {

  samp<-sample(seq(1,nrow(trainOG),by=1) , nrow(trainOG),replace = T) 
  train<-trainOG[samp,]
  times<-seq(from=min(test$time),
             to=max(test$time),
             length.out = 100
  )
  times<-times[-c(1,length(times))]

  ######################
  ###casebase + splines
  ######################
  mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
                                data = train,
                                time = "time",
                                event="status",
                                ratio = 100
  )

  glmAbsRisk<-as.data.frame(absoluteRisk(mod_cb_glm,time=times,newdata=test,type="CI"))
  rownames(glmAbsRisk)<-glmAbsRisk$time
  #prepare to run through riskRegression
  glmProper<- t(glmAbsRisk[-1,-1])
  class(glmProper)<-c("tunnel",class(glmProper))

  ######################
  ###DeepSurv
  ######################
#   a=get_pycox_activation(
#   activation = "tanh",
#   construct = TRUE,
#   alpha = 1,
#   dim = NULL,
#   lambd = 0.5,
#   min_val = -1,
#   max_val = 1,
#   negative_slope = 0.01,
#   num_parameters = 1L,
#   init = 0.25,
#   lower = 1/8,
#   upper = 1/3,
#   beta = 1,
#   threshold = 20,
#   value = 20
# )
#   coxnn<-deepsurv(formula=Surv(time, status) ~ .,data = train, frac = 0.2, activation = 'tanh',
#                   num_nodes = c(layer1,layer2), dropout = drpt,
#                   early_stopping = TRUE, epochs = epo,
#                   batch_size = bsize,device = "cpu",best_weights=T,
#                   patience = patience,min_delta=min_delta,batch_norm = T,
#                   learning_rate = 0.0001,lr_decay=10^-7)
#   coxnnsurv<- predict(coxnn,newdata = test,type="survival")
#   colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
#   cnnCleaned<-pyProcess(coxnnsurv,times=times)
#   colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))

  source_python('src/deepsurver.py')
  coxnnsurv=fitDeepSurv(train,fullTest,bsize,epochs=epo,patience=patience,min_delta=min_delta,drpt=drpt)
  coxnnsurv<-t(coxnnsurv)
  colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
  cnnCleaned<-pyProcess(coxnnsurv,times=times)
  colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
  py_run_string("del fitDeepSurv")
  rm(fitDeepSurv)

  ######################
  ###Deephit
  ######################
  source_python('src/oldmortDeephitter.py')
  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,patience=patience,min_delta=min_delta,drpt=drpt)
  hitnnSurv<-t(hitnnSurv)
  colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
  deephitCleaned<-pyProcess(hitnnSurv,times=times)
  colnames(deephitCleaned)<-seq(0,1,length.out = ncol(deephitCleaned))
  py_run_string("del fitDeephit")
  rm(fitDeephit)

  ######################
  ###PMNN splines
  ######################
  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2+3),
                            name = 'main_input'
  )
  covars_output<-covars_input%>%layer_batch_normalization()%>%
    layer_dense(units=layer1,use_bias = T)%>%
    activation_tanh()%>%layer_dropout(drpt)%>%
    layer_dense(units=layer2,use_bias = T)%>%
    activation_tanh()%>%layer_dropout(drpt)%>%
    layer_dense(units=1,use_bias = T)#%>%activation_tanh()

  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                  feature_input = covars_input,
                  feature_output = covars_output,
                  originalData = mod_cb_glm$data,
                  offset=mod_cb_glm$offset,
                  timeVar = "time",
                  eventVar= "status",
                  optimizer=optimizer_adam(learning_rate = 0.0001,decay=10^-7)
  )

  fit<-fitSmoothHazSpline(pmnn,
                          epochs=epo,
                          batch_size=5120,
                          verbose=0,
                          monitor="val_loss",
                          val_split=0.2,
                          min_delta=min_delta,
                          patience=patience)
  annPreds<-aarSplines(fit,
                       times=times,
                       x_test=test
  )
  rownames(annPreds)<-annPreds[,1]
  annProper<- t(annPreds[,-1])
  class(annProper)<-c("tunnel",class(annProper))

  rm(pmnn,covars_output,covars_input,fit)

  ######################
  ###coxph
  ######################
  cox<-coxph(Surv(time, status) ~ ., data = train,x=T)

  ######################
  ###PMNN
  ######################
  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
                            name = 'main_input')

  covars_output<-covars_input%>%layer_batch_normalization()%>%
    layer_dense(units=layer1,use_bias = T)%>%
    activation_tanh()%>%layer_dropout(drpt)%>%
    layer_dense(units=layer2,use_bias = T)%>%
    activation_tanh()%>%layer_dropout(drpt)%>%
    layer_dense(units=1,use_bias = T)#%>%activation_tanh()

  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                  feature_input = covars_input,
                  feature_output = covars_output,
                  originalData = mod_cb_glm$data,
                  offset=mod_cb_glm$offset,
                  timeVar = "time",
                  eventVar= "status",
                  optimizer=optimizer_adam(learning_rate = 0.0001,decay=10^-7)
  )

  fit<-fitSmoothHaz(pmnn,
                    epochs=epo,
                    batch_size=512,
                    verbose=0,
                    monitor="val_loss",
                    val_split=0.2,
                    min_delta=min_delta,
                    patience=patience)
  annPreds<-aar(fit,
                times=times,
                x_test=test
  )

  rownames(annPreds)<-annPreds[,1]
  annProperpoly<- t(annPreds[,-1])
  class(annProperpoly)<-c("tunnel",class(annProperpoly))

  ######################
  ###IPA metric
  ######################
  brierFinalResults <- Score(list("Cox_Lin" = cox,'CB_Logi'=glmProper,
                                  'DeepSurv'=cnnCleaned,'DeepHit'=deephitCleaned,
                                  'PMNN_Spline'=annProper,'PMNN_Poly'=annProperpoly),#
                             data =fullTest, 
                             formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)
  BrierIPAList[[pos]]<-brierFinalResults$Brier$score

  ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
  geom_line()+coord_cartesian(ylim=c(-1,0.7))
  rm(pmnn,covars_output,covars_input,fit)
  #######################
  ###Cox 
  ######################
  a<-survfit(cox, newdata=fullTest,times=times)
  rownames(a$surv)<-a$time
  coxlinSurv<-pyProcess(t(a$surv),times=times)
  colnames(coxlinSurv)<-seq(0,1,length.out = ncol(coxlinSurv))

  ######################
  ###C-Index
  ######################

  riskList<-list(coxlinSurv,
                 glmProper,
                 cnnCleaned,
                 deephitCleaned,
                 annProper,
                 annProperpoly
  )

  tempTims<-   as.numeric(colnames(annProper))
  tempTims<-head(tempTims, -1)
  tempTims<-tail(tempTims, -1)
  cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=length(riskList)+1)
  et_train<-as.matrix(train)[,c(16,15)]
  et_test<-as.matrix(fullTest[,c(16,15)])
  colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
                          'PMNN_Spline','PMNN_Poly')
  cScoreTemp[,1]<- tempTims

  cScore[[pos]]<-cIndexSummary(et_train=et_train,
                               et_test=et_test,
                               riskList=riskList,
                               cScore=cScoreTemp
  )

  saveRDS(BrierIPAList,'results/100oldBrier.rds')
  saveRDS(cScore,'results/100oldcidx.rds')
  pos<-pos+1
}

ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=Brier,col=model))+
  geom_line()
ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
  geom_line()+coord_cartesian(ylim=c(-2,0.7))

end_time <- Sys.time()

end_time-start_time

```

```{r tuning,eval=F}
library(kerastuneR)
  ######################
  ###PMNN
  ######################
samp<-sample(seq(1,nrow(trainOG),by=1), floor(nrow(trainOG)*.80)) 
validation<-trainOG[-samp,]

train<-trainOG[samp,]

  mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
                                data = train,
                                time = "time",
                                event="status",
                                ratio = 100
  )

buildModel<-function(hp){
  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
                            name = 'main_input')

  covars_output<-covars_input%>% layer_batch_normalization()%>%
    layer_dense(units=hp$Int('units',min_value=5,max_value=25,step=2),
                use_bias = T,activation = 'linear')%>%
    layer_dropout(hp$Choice('rate',
                values=c(0.4,0.3,0.2,0.1,0.05)))%>%
    layer_dense(units=hp$Int('units',min_value=5,max_value=25,step=2),
                use_bias = T,activation = 'linear')%>%
    layer_dropout(hp$Choice('rate',
                values=c(0.4,0.3,0.2,0.1,0.05)))%>%
    layer_dense(units=1,use_bias = T)

  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                  feature_input = covars_input,
                  feature_output = covars_output,
                  originalData = mod_cb_glm$data,
                  offset=mod_cb_glm$offset,
                  timeVar = "time",
                  eventVar= "status",
                  optimizer= tf$keras$optimizers$Adam(
      hp$Choice('learning_rate',
                values=c(1e-2,5e-3, 1e-3,5e-4, 1e-4)))
  )
  return(pmnn$model)
}

  tuner = RandomSearch(
    buildModel,
    objective = 'val_loss',
    max_trials = 20,
    executions_per_trial = 3,
    directory = 'tuning',
    project_name = 'oldmort')

  tuner %>% search_summary()

  xtr<-as.matrix(mod_cb_glm$data[,-c(ncol(mod_cb_glm$data),ncol(mod_cb_glm$data)-1)])
  ytr<-as.matrix(mod_cb_glm$data[,c(ncol(mod_cb_glm$data)-1)])
  xv<-list(as.matrix(validation[,-c(ncol(validation))]),as.matrix(data.frame(offset=rep(0,nrow(validation)))))
  yv<-as.matrix(validation[,c(ncol(validation)),drop=F])
  tuner %>% fit_tuner(x=list(xtr,as.matrix(mod_cb_glm$offset)),y=ytr,
                    epochs = 50,batch_size=bsize+1000, 
                    validation_data = list(x=xv,y=yv),verbose=0 )
result = kerastuneR::plot_tuner(tuner)
# the list will show the plot and the data.frame of tuning results
result

best_5_models = tuner %>% get_best_models(5)
best_5_models[[1]] %>% plot_keras_model()
  # 
  # fit<-fitSmoothHaz(pmnn,
  #                   epochs=epo,
  #                   batch_size=bsize+1000,
  #                   verbose=0,
  #                   monitor="val_loss",
  #                   val_split=0.2,
  #                   min_delta=min_delta,
  #                   patience=patience+10)
  # annPreds<-aar(fit,
  #               times=times,
  #               x_test=test
  # )
  # 
  # rownames(annPreds)<-annPreds[,1]
  # annProperpoly<- t(annPreds[,-1])
  # class(annProperpoly)<-c("tunnel",class(annProperpoly)) 
  #

```

### 2022-12-01 analyses/src/deepsurver.py

@@ -1,0 +1,71 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from sklearn.preprocessing import StandardScaler
+from sklearn_pandas import DataFrameMapper
+
+import torch
+import torchtuples as tt
+
+from pycox.datasets import metabric
+from pycox.models import CoxPH
+from pycox.evaluation import EvalSurv
+
+
+
+def fitDeepSurv(train,fullTest,bsize,epochs,ti='time',ev='status',patience=10,min_delta=10**-7,drpt=0.5):#,epoch,bsize,tPos,ePos):
+  n= np.shape(train)[0]
+  vl_size = int(n*0.20)
+  df_val = train.sample(frac=0.2)
+  df_train = train.drop(df_val.index)
+  df_test = fullTest
+  varNames = list(df_train.columns)
+  unwanted = {ti, ev}
+  varNames = [e for e in varNames if e not in unwanted]
+  leave = [(col, None) for col in varNames]
+
+  x_mapper = DataFrameMapper(leave)
+
+  x_train = x_mapper.fit_transform(df_train).astype('float32')
+  x_val = x_mapper.transform(df_val).astype('float32')
+  x_test = x_mapper.transform(df_test).astype('float32')
+  get_target = lambda df: (df[ti].values, df[ev].values)
+  y_train = get_target(df_train)
+  y_val = get_target(df_val)
+  durations_test, events_test = get_target(df_test)
+  val = x_val, y_val
+  
+  
+  in_features = x_train.shape[1]
+  num_nodes = [200,20]
+  out_features = 1
+  #batch_norm = False
+  dropout = drpt
+  
+  #net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout,activation=torch.tanh)
+  net = torch.nn.Sequential(
+     torch.nn.BatchNorm1d(in_features),
+     torch.nn.Linear(in_features,200),
+     torch.nn.Tanh(),
+     torch.nn.Dropout(0.1),
+    
+     torch.nn.Linear(200, 20),
+     torch.nn.Tanh(),
+     torch.nn.Dropout(0.1),
+    
+     torch.nn.Linear(20, out_features))
+  
+  model = CoxPH(net, tt.optim.Adam)
+  
+  model.optimizer.set_lr(0.001)
+  callbacks = [tt.callbacks.EarlyStopping(min_delta=min_delta, patience=patience)]
[…]

### 2022-12-01 analyses/src/functions.R

@@ -186,7 +186,7 @@ predictRisk.tunnel <- function(object, newdata, times, cause, ...){
 #########################################################
 
 
-pmnnModel<-function(features, feature_input, feature_output, originalData=originalData,offset, timeVar,eventVar, ratio=100, compRisk=FALSE,censored.indicator=0,optimizer=optimizer_adam(lr = 0.001,decay=10^-7)){
+pmnnModel<-function(features, feature_input, feature_output, originalData=originalData,offset, timeVar,eventVar, ratio=100, compRisk=FALSE,censored.indicator=0,optimizer=optimizer_adam(learning_rate = 0.001,decay=10^-7)){
   if(class(originalData)[1]=="cbData"){
     data<-originalData
   }else{

### 2022-12-01 analyses/src/oldmortDeephitter.py

@@ -1,0 +1,82 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+# For preprocessing
+from sklearn.preprocessing import StandardScaler
+from sklearn_pandas import DataFrameMapper 
+
+import torch # For building the networks 
+import torchtuples as tt # Some useful functions
+
+from pycox.models import DeepHitSingle
+from pycox.evaluation import EvalSurv
+
+
+
+
+def fitDeephit(train,fullTest,bsize,epochs,ti='time',ev='status',patience=10,min_delta=10**-7,drpt=0.5):#,epoch,bsize,tPos,ePos):
+  n= np.shape(train)[0]
+  vl_size = int(n*0.20)
+  df_val = train.sample(frac=0.2)
+  df_train = train.drop(df_val.index)
+  df_test = fullTest
+  varNames = list(df_train.columns)
+  unwanted = {ti, ev}
+  varNames = [e for e in varNames if e not in unwanted]
+  leave = [(col, None) for col in varNames]
+
+  x_mapper = DataFrameMapper(leave)
+
+  x_train = x_mapper.fit_transform(df_train).astype('float32')
+  x_val = x_mapper.transform(df_val).astype('float32')
+  x_test = x_mapper.transform(df_test).astype('float32')
+ 
+
+
+
+  ########
+  #prepare labels
+  #######
+  num_durations = 10
+  labtrans = DeepHitSingle.label_transform(num_durations)
+  get_target = lambda df: (df[ti].values, df[ev].values)
+  y_train = labtrans.fit_transform(*get_target(df_train))
+  y_val = labtrans.transform(*get_target(df_val))
+  train = (x_train, y_train)
+  val = (x_val, y_val)
+  durations_test, events_test = get_target(df_test)
+  ########
+  #prepare model
+  ########
+  in_features = x_train.shape[1]
+  num_nodes = [200,20]
+  out_features = labtrans.out_features
+  batch_norm = False
+  dropout = drpt
+  
+  #net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout,activation=torch.tanh)
+  net = torch.nn.Sequential(
+     torch.nn.BatchNorm1d(in_features),
+     torch.nn.Linear(in_features,200),
+     torch.nn.Tanh(),
+     torch.nn.Dropout(0.1),
+    
+     torch.nn.Linear(200, 20),
[…]

### 2022-12-15 analyses/SIMpaperResultscopy/complexSim.Rmd
re-analyzed

--- 2022-05-25 analyses/complexSim.Rmd
@@
   coxnnsurv<- predict(coxnn,newdata = test,type="survival")
-  colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
+  #colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
   cnnCleaned<-pyProcess(coxnnsurv,times=times)
@@
   hitnnSurv<-t(hitnnSurv)
-  colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
+  #colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
   deephitCleaned<-pyProcess(hitnnSurv,times=times)
@@
 ```
+
+```{r}
+
+ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
+  geom_line()
+
+```

### 2022-12-15 analyses/SIMpaperResultscopy/metabricStudy.Rmd
re-analyzed

---
title: "Metabric"
output: html_document
params:
  epo: 2000
  patience: 10
  iteration: 10
  layer1: 50
  layer2: 50
  layer3: 25
  layer4: 25
  drpt: 0.5
---

```{r,eval=T}

library(reticulate)
```

```{python,eval=T}

import os
os.environ["CUDA_DEVICE_ORDER"] = "PCI_BUS_ID" 
os.environ["CUDA_VISIBLE_DEVICES"] = ""
#os.environ["OM_NUM_THREADS"]="2"

```

```{r setup, include=T}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
source("src/packages.R")
source("src/functions.R")
eval_metabric<-T
pycox<-reticulate::import("pycox")
source_python('src/getMetabricData.py')
source_python("src/cIndex.py")
source_python("src/dsmScript.py")
source_python('src/deephitter.py')
##########################
###Shared Hyperparameters
##########################
bsize=512
min_delta = 10^-7
epo=as.numeric(params$epo)
patience = as.numeric(params$patience)
iteration=as.numeric(params$iteration)
layer1<-as.numeric(params$layer1)
layer2<-as.numeric(params$layer2)
layer3<-as.numeric(params$layer3)
layer4<-as.numeric(params$layer4)
drpt<-as.numeric(params$drpt)

```

```{r modelsMETABRIC,eval=eval_metabric}

start_time <- Sys.time()

#pycox<-reticulate::import("pycox")
#numpy<-reticulate::import("numpy")
#a<-pycox$datasets$metabric$read_df
#b<-numpy$asarray(a)

############################
###Data clean and prep
############################

metabric<-getMetabric()
colnames(metabric)[ncol(metabric)-1]<-'time'
colnames(metabric)[ncol(metabric)]<-'status'
data<-as.data.frame(model.matrix(~.,metabric)[,-1])
data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
covs<-colnames(data)

samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.80)) 
trainOG<-data[samp,]
test<-data[-samp,-c(ncol(data))]

fullTest<-data[-samp,,drop=F]

############################
###Prep Loop
############################
BrierIPAList<-list()
cScore<-list()
pos<-1

while (pos <= iteration) {

  ############################
  ###Bootstrap samples
  ############################  
  samp<-sample(seq(1,nrow(trainOG),by=1) , nrow(trainOG),replace = T) 
  train<-trainOG[samp,]

  times<-seq(from=min(test$time),
             to=max(test$time),
             length.out = 200
  )

  times<-times[-c(1,length(times))]
  times<-unique(times)

  ############################
  ###DSM
  ############################  
  source_python('src/dsmScript.py')
  dsmSurv<-dsmfitter(train=train,fullTest=fullTest,bsize=bsize,epochs=epo,tPos=ncol(train)-2,ePos=ncol(train)-1)
  #sometimes, dsm fails to converge so we keep trying until it succeeds
  if((sum(is.na(dsmSurv))==ncol(dsmSurv)*nrow(dsmSurv))){
    dsmSurv<-dsmfitter(train=train,fullTest=fullTest,bsize=bsize,epochs=epo,tPos=ncol(train)-2,ePos=ncol(train)-1)
    py_run_string("del dsmfitter")
    rm(dsmfitter)
    iteration<-iteration+1
    next
  }
  colnames(dsmSurv) <-seq(0.01,1,by=0.01)
  dsmSurvProper<-pyProcess(dsmSurv,times=times)
  colnames(dsmSurvProper) <-seq(0,1,length.out = ncol(dsmSurvProper))
  py_run_string("del dsmfitter")
  rm(dsmfitter)

  ############################
  ###casebase + splines
  ############################  
  mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
                                data = train,
                                time = "time",
                                event="status",
                                ratio = 100
  )

  glmAbsRisk<-absoluteRisk(mod_cb_glm,time=times,newdata=test,type="CI")
  glmAbsRisk<-as.data.frame(glmAbsRisk)
  rownames(glmAbsRisk)<-glmAbsRisk$time
  glmProper<- t(glmAbsRisk[-1,-1])
  class(glmProper)<-c("tunnel",class(glmProper))

  ############################
  ###DeepSurv
  ############################  
  coxnn<-deepsurv(Surv(time, status) ~ .,data = train, frac = 0.2, activation = "relu",
                  num_nodes = c(layer1,layer2,layer3,layer4), dropout = drpt,
                  early_stopping = TRUE, epochs = epo,
                  batch_size = bsize,device = "cpu",best_weights=T,
                  patience = patience,min_delta=min_delta,batch_norm = F,
                  learning_rate = 0.001,lr_decay=10^-7)

  coxnnsurv<- predict(coxnn,newdata = test,type="survival")
  colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
  cnnCleaned<-pyProcess(coxnnsurv,times=times)
  colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
  ############################
  ###Deephit
  ############################  
  source_python('src/deephitter.py')
  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,patience=patience,min_delta=min_delta,drpt=drpt)
  hitnnSurv<-t(hitnnSurv)
  colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
  deephitCleaned<-pyProcess(hitnnSurv,times=times)
  colnames(deephitCleaned)<-seq(0,1,length.out = ncol(deephitCleaned))
  py_run_string("del fitDeephit")
  rm(fitDeephit)

  ############################
  ###PMNNsplines
  ############################  
  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2+3),
                            name = 'main_input'
  )
  covars_output<-covars_input%>% 
    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=1,use_bias = T)

  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                  feature_input = covars_input,
                  feature_output = covars_output,
                  originalData = mod_cb_glm$data,
                  offset=mod_cb_glm$offset,
                  timeVar = "time",
                  eventVar= "status"
  )

  fit<-fitSmoothHazSpline(pmnn,
                          epochs=epo,
                          batch_size=bsize,
                          verbose=0,
                          monitor="val_loss",
                          val_split=0.2,
                          min_delta=min_delta,
                          patience=patience)
  annPreds<-aarSplines(fit,
                       times=times,
                       x_test=test
  )

  rownames(annPreds)<-annPreds[,1]
  annProper<- t(annPreds[,-1])
  class(annProper)<-c("tunnel",class(annProper))

  ############################
  ###coxph
  ############################  
  cox<-coxph(Surv(time, status) ~ ., data = train,x=T)

  ############################
  ###PMNN
  ############################  
  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
                            name = 'main_input')

  covars_output<-covars_input%>% #layer_batch_normalization()%>%
    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=1,use_bias = T)

  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                  feature_input = covars_input,
                  feature_output = covars_output,
                  originalData = mod_cb_glm$data,
                  offset=mod_cb_glm$offset,
                  timeVar = "time",
                  eventVar= "status"
  )
  fit<-fitSmoothHaz(pmnn,
                    epochs=epo,
                    batch_size=bsize,
                    verbose=0,
                    monitor="val_loss",
                    val_split=0.2,
                    min_delta=min_delta,
                    patience=patience)
  annPreds<-aar(fit,
                times=times,
                x_test=test
  )
  rownames(annPreds)<-annPreds[,1]
  annProperpoly<- t(annPreds[,-1])
  class(annProperpoly)<-c("tunnel",class(annProperpoly)) 
  ############################
  ###Brier Score
  ############################  
  brierFinalResults <- Score(list("Cox_Lin" = cox,'CB_Logi'=glmProper,
                                  'DeepSurv'=cnnCleaned,'DeepHit'=deephitCleaned,
                                  'PMNN_Spline'=annProper,'PMNN_Poly'=annProperpoly,
                                  'dsm'=dsmSurvProper),
                             data =fullTest, 
                             formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)
  BrierIPAList[[pos]]<-brierFinalResults$Brier$score

  #######################
  ###Cox 
  ######################
  a<-survfit(cox, newdata=fullTest,times=times)
  rownames(a$surv)<-a$time
  coxlinSurv<-pyProcess(t(a$surv),times=times)
  colnames(coxlinSurv)<-seq(0,1,length.out = ncol(coxlinSurv))

  ######################
  ###C-Index
  ######################

  riskList<-list(coxlinSurv,
                 glmProper,
                 cnnCleaned,
                 deephitCleaned,
                 annProper,
                 annProperpoly,
                 dsmSurvProper
  )

  tempTims<-   as.numeric(colnames(annProper))
  tempTims<-head(tempTims, -1)
  tempTims<-tail(tempTims, -1)
  cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=length(riskList)+1)
  et_train<-as.matrix(train)[,c(5,4)]
  et_test<-as.matrix(fullTest[,c(5,4)])
  colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
                          'PMNN_Spline','PMNN_Poly','dsm')
  cScoreTemp[,1]<- tempTims

  cScore[[pos]]<-cIndexSummary(et_train=et_train,
                               et_test=et_test,
                               riskList=riskList,
                               cScore=cScoreTemp
  )

  saveRDS(BrierIPAList,'results/100MetaBrier.rds')
  saveRDS(cScore,'results/100Metacidx.rds')
  pos=pos+1
} 
end_time <- Sys.time()

end_time-start_time

ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=Brier,col=model))+
  geom_line()
ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
  geom_line()

```

```{r}

sessionInfo()

```

### 2022-12-15 analyses/SIMpaperResultscopy/simpleSim.Rmd
re-analyzed

--- 2022-05-25 analyses/codeReviewIterSim.Rmd
@@
 
-easysim<-F
-hardsim<-T
+easysim<-T
 numSamp<-2000
@@
     rm(dsmfitter)
-    iteration<-iteration+1
     next
@@
   BrierIPAList[[pos]]<-brierFinalResults$Brier$score
-  ########################
-  ###C-Index
-  ########################
-  tempTims<-   as.numeric(colnames(annProper))
-  tempTims<-head(tempTims, -15)
-  tempTims<-tail(tempTims, -10)
-  cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=9)
-  et_train<-as.matrix(train)#[,c(5,4)]
-  et_test<-as.matrix(fullTest[,c(5,4)])
-  colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
-                          'PMNN_Spline','Optimal','PMNN_Poly','dsm')
-  cScoreTemp[,1]<- tempTims
 
+  #######################
+  ###Cox 
+  ######################
   a<-survfit(cox, newdata=fullTest,times=times)
@@
   colnames(coxlinSurv)<-seq(0,1,length.out = ncol(coxlinSurv))
+
+  ######################
+  ###C-Index
+  ######################
 
@@
                  annProper,
-                 annProperpoly
+                 trueGlmProper,
+                 annProperpoly,
+                 dsmSurvProper
   )
 
-  cScore[[pos]]<-cIndexSummary(et_train=et_train
-                               , et_test=et_test,
+  tempTims<-   as.numeric(colnames(annProper))
+  tempTims<-head(tempTims, -1)
+  tempTims<-tail(tempTims, -1)
+  cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=length(riskList)+1)
+  et_train<-as.matrix(train)[,c(5,4)]
+  et_test<-as.matrix(fullTest[,c(5,4)])
+  colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
+                          'PMNN_Spline','Optimal','PMNN_Poly','dsm')#,'dsm')
+  cScoreTemp[,1]<- tempTims
+
+  cScore[[pos]]<-cIndexSummary(et_train=et_train,
+                               et_test=et_test,
                                riskList=riskList,
@@
 
-  saveRDS(BrierIPAList,'100easyBrier.rds')
-  saveRDS(cScore,'100easycidx.rds')
+  saveRDS(BrierIPAList,'results/100easyBrier.rds')
+  saveRDS(cScore,'results/100easycidx.rds')
   pos<-pos+1
 }
-
 end_time <- Sys.time()
@@
 ```
-
-```{r complexPrep,eval=hardsim}
-
-##############################
-###set up complex simulation
-##############################
-
-# Define a function returning the log cum hazard at time t
-logcumhaz <- function(t, x, betas, knots) {
-
-  # Obtain the basis terms for the spline-based log
-  # cumulative hazard (evaluated at time t)
-  basis <- flexsurv::basis(knots, log(t))
-
-  # Evaluate the log cumulative hazard under the
-  # Royston and Parmar specification
-  res <-betas[["gamma0"]] * basis[[1]] + 
-    betas[["gamma1"]] * basis[[2]] +
-    betas[["gamma2"]] * basis[[3]] +
-    betas[["gamma3"]] * basis[[4]] +
-    betas[["gamma4"]] * basis[[5]] +
-    betas[["z1"]] * x[["z1"]]+
-    betas[["z2"]] * x[["z2"]]+
-    betas[["z3"]] * x[["z3"]]+
-    betas[["term1"]] * x[["z1"]]*x[["z2"]]*t+
-    betas[["term2"]] * x[["z1"]]*x[["z3"]]+
-    betas[["term3"]] * x[["z2"]]*(x[["z3"]])
-  #
-  # Return the log cumulative hazard at time t
-  res
-}
-url<-'app.txt'
-ds <- read.table(url, sep = '',header = TRUE)
-ds$years.at.risk=pmin(ds$agein80,ds$onset)
-brcancer$rectime<-brcancer$rectime/max(brcancer$rectime)
-true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1, data = brcancer, k = 3) #from flexsurv
-
-# Simulate the event times
-coefficients<-true_mod$coefficients
-coefficients<-c(coefficients,z1=1,z2=1,z3=1,term1=10,term2=2,term3=2)
-
-```
-
-```{r totalSim,eval=hardsim}
-
-start_time <- Sys.time()
-BrierIPAList<-list()
-cScore<-list()
-pos<-1
-for(iter in 1:iteration){
-
-  #########################
-  ###Simulate date
-  #########################
-  cov<-data.frame(id=1:numSamp,
-                  z1=c(rep(1,numSamp/2),rep(0,numSamp/2)),
-                  z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
-                  z3=c(rnorm(numSamp/2,-3,0.5),rnorm(numSamp/2,8,0.5)))
-  true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1,
-                                       data = brcancer, k = 3
-  ) #from flexsurv
-  dat <- simsurv(betas = coefficients, # "true" parameter values
-                 x = cov,                   # covariate data for 200 individuals
-                 knots = true_mod$knots,    # knot locations for splines
-                 logcumhazard = logcumhaz,  # definition of log cum hazard
-                 maxt = 1,               # no right-censoring
-                 interval = c(0,10000000)) # interval for root finding
-  colnames(dat)<-c("id","time","status")
-  samp<-sample(seq(1,nrow(dat),by=1), floor(nrow(dat)*(0.1))) #random censoring
-  dat$status[samp]<-0
-  # Merge the simulated event times onto covariate data frame
-  data <- merge(cov, dat)
-  covs<-colnames(data)
-
-  #########################
-  ###Simulate data
-  #########################
-  samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.8))
-  data<-data[,-1]
-  data[,-c(1,5)]<-as.data.frame(sapply(data[,-c(1,5)], function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
-  train<-data[samp,]
-  test<-data[-samp,-c(ncol(data))]
-  fullTest<-data[-samp,,drop=F]
-  times<-seq(from=0,
-             to=max(test$time),
-             length.out = 200
-  )
-  times<-times[-c(1,length(times))]
-
-  ###########################
-  ###casebase optimal model
-  ###########################
-  final_mod_cb_glm <- fitSmoothHazard(status~bs(time)-time+z1+z2+z3+z1:z2:time+z1:z3+z2:z3,
-                                      data = train,
-                                      time = "time",
-                                      event="status",
-                                      ratio = 100
-  )
-
-  summary(final_mod_cb_glm)
-  tglmAbsRisk<-absoluteRisk(final_mod_cb_glm,time=times,newdata=test,type="CI")
-  tglmAbsRisk<-as.data.frame(tglmAbsRisk)
-  rownames(tglmAbsRisk)<-tglmAbsRisk$time
-  trueGlmProper<- t(tglmAbsRisk[-1,-1])
-  class(trueGlmProper)<-c("tunnel",class(trueGlmProper))
-
-  #########################
-  ###casebase+splines
-  #########################
-  mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
-                                data = train,
-                                time = "time",
-                                event="status",
-                                ratio = 100
-  )
-
-  glmAbsRisk<-absoluteRisk(mod_cb_glm,time=times,newdata=test,type="CI")
-
-  glmAbsRisk<-as.data.frame(glmAbsRisk)
-
-  rownames(glmAbsRisk)<-glmAbsRisk$time
-  glmProper<- t(glmAbsRisk[-1,-1])
-
-  class(glmProper)<-c("tunnel",class(glmProper))
-
-  ######################
-  ###PMNN splines
-  ######################
-  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2+3),
-                            name = 'main_input'
-  )
-  covars_output<-covars_input%>% #layer_batch_normalization()%>%
-    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=1,use_bias = T)
-
-  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
-                  feature_input = covars_input,
-                  feature_output = covars_output,
-                  originalData = mod_cb_glm$data,
-                  offset=mod_cb_glm$offset,
-                  timeVar = "time",
-                  eventVar= "status"
-  )
-
-  fit<-fitSmoothHazSpline(pmnn,
-                          epochs=epo,
-                          batch_size=bsize,
-                          verbose=0,
-                          monitor="val_loss",
-                          val_split=0.2,
-                          min_delta=min_delta,
-                          patience=patience)
-  annPreds<-aarSplines(fit,
-                       times=times,
-                       x_test=test
-  )
-  rownames(annPreds)<-annPreds[,1]
-  annProper<- t(annPreds[,-1])
-  class(annProper)<-c("tunnel",class(annProper))
-
-  ######################
-  ###coxph
-  ######################
-  cox<-coxph(Surv(time, status) ~ z1+z2+z3, data = train,x=T)
-
-  ######################
-  ###DeepSurv
-  ######################
-  coxnn<-deepsurv(Surv(time, status) ~ .,data = train, frac = 0.2, activation = "relu",
-                  num_nodes = c(layer1,layer2,layer3,layer4), dropout = drpt,
-                  early_stopping = TRUE, epochs = epo,
-                  batch_size = bsize,device = "cpu",best_weights=T,
-                  patience = patience,min_delta=min_delta,batch_norm = F,
-                  learning_rate = 0.001,lr_decay=10^-7)
-  coxnnsurv<- predict(coxnn,newdata = test,type="survival")
-  colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
-  cnnCleaned<-pyProcess(coxnnsurv,times=times)
-  colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
-
-  ######################
-  ###Deephit
-  ######################
-  source_python('src/deephitter.py')
-  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,patience=patience,min_delta=min_delta,drpt=drpt)
-  hitnnSurv<-t(hitnnSurv)
-  colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
-  deephitCleaned<-pyProcess(hitnnSurv,times=times)
-  colnames(deephitCleaned)<-seq(0,1,length.out = ncol(deephitCleaned))
-  py_run_string("del fitDeephit")
-  rm(fitDeephit)
-
-  ######################
-  ###PMNN
-  ######################
-  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
-                            name = 'main_input')
-
-  covars_output<-covars_input%>% #layer_batch_normalization()%>%
-    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=1,use_bias = T)
-
-  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
-                  feature_input = covars_input,
-                  feature_output = covars_output,
-                  originalData = mod_cb_glm$data,
-                  offset=mod_cb_glm$offset,
-                  timeVar = "time",
-                  eventVar= "status"
-  )
-
-  fit<-fitSmoothHaz(pmnn,
-                    epochs=epo,
-                    batch_size=bsize,
-                    verbose=0,
-                    monitor="val_loss",
-                    val_split=0.2,
-                    min_delta=min_delta,
-                    patience=patience)
-  annPreds<-aar(fit,
-                times=times,
-                x_test=test
-  )
-
-  rownames(annPreds)<-annPreds[,1]
-  annProperpoly<- t(annPreds[,-1])
-  class(annProperpoly)<-c("tunnel",class(annProperpoly))
-
-  brierFinalResults <- Score(list("Cox_Lin" = cox,
-                                  'CB_Logi'=glmProper,
-                                  'DeepSurv'=cnnCleaned,
-                                  'DeepHit'=deephitCleaned,
-                                  'PMNN_Spline'=annProper,
-                                  'Optimal'=trueGlmProper,
-                                  'PMNN_Poly'=annProperpoly),
-                             #'dsm'=dsmSurvProper) ,#'dsm'=dsmCleaned,
-                             data =fullTest, 
-                             formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)
-  BrierIPAList[[pos]]<-brierFinalResults$Brier$score
-
-  ######################
-  ###C-Index
-  ######################
-  tempTims<-   as.numeric(colnames(annProper))
-  tempTims<-head(tempTims, -15)
-  tempTims<-tail(tempTims, -10)
-  cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=8)
-  et_train<-as.matrix(train)#[,c(5,4)]
-  et_test<-as.matrix(fullTest[,c(5,4)])
-  colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
-                          'PMNN_Spline','Optimal','PMNN_Poly')#,'dsm')
-  cScoreTemp[,1]<- tempTims
-  a<-survfit(cox, newdata=fullTest,times=times)
-  rownames(a$surv)<-a$time
-  coxlinSurv<-pyProcess(t(a$surv),times=times)
-  colnames(coxlinSurv)<-seq(0,1,length.out = ncol(coxlinSurv))
-  riskList<-list(coxlinSurv,
-                 glmProper,
-                 cnnCleaned,
-                 deephitCleaned,
-                 annProper,
-                 annProperpoly
-  )
-
-  cScore[[pos]]<-cIndexSummary(et_train=et_train
-                               , et_test=et_test,
-                               riskList=riskList,
-                               cScore=cScoreTemp
-  )
-
-  saveRDS(BrierIPAList,'results/100hardBrier.rds')
-  saveRDS(cScore,'results/100hardcidx.rds')
-  pos<-pos+1
-}
-
-```

### 2022-12-15 analyses/VeteransAnalysis.Rmd
re-analyzed

---
title: "vets"
output: html_document
params:
  epo: 2000
  patience: 10
  iteration: 1
  layer1: 50
  layer2: 20
  layer3: 25
  layer4: 25
  drpt: 0.1
---

```{r,eval=T}

library(reticulate)
```

```{python,eval=T}

import os
os.environ["CUDA_DEVICE_ORDER"] = "PCI_BUS_ID" 
os.environ["CUDA_VISIBLE_DEVICES"] = ""
#os.environ["OM_NUM_THREADS"]="2"

```

```{r setup, include=T}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
source("src/packages.R")
source("src/functions.R")
eval_metabric<-T
pycox<-reticulate::import("pycox")
source_python("src/cIndex.py")
source_python("src/dsmScript.py")
source_python('src/oldmortDeephitter.py')
##########################
###Shared Hyperparameters
##########################
bsize=512
min_delta = 10^-8
epo=as.numeric(params$epo)
patience = as.numeric(params$patience)
iteration=as.numeric(params$iteration)
layer1<-as.numeric(params$layer1)
layer2<-as.numeric(params$layer2)
layer3<-as.numeric(params$layer3)
layer4<-as.numeric(params$layer4)
drpt<-as.numeric(params$drpt)

```

```{r testAnalysis}

start_time <- Sys.time()
BrierIPAList<-list()
cScore<-list()
#library(eha)
veteran<-veteran

coxph(Surv(time, status) ~ trt + prior + karno, veteran)

preData<-data.frame(trt=veteran$trt,
                    prior=veteran$prior,
                    karno=veteran$karno,
                    time=veteran$time,
                    status=veteran$status
                    )
#preData$time<-preData$time+runif(nrow(preData),0,1)/100000000

data<-as.data.frame((model.matrix(~.,preData)[,-1]))

data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
covs<-colnames(data)

#samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.80))

 spec = c(train = .90, test = .10)#, validate = 0)

 g = sample(cut(
   seq(nrow(data)), 
   nrow(data)*cumsum(c(0,spec)),
   labels = names(spec)
 ))

 res = split(data, g)

trainOG<-res$train

test<-res$test[,-c(ncol(data))]

fullTest<-res$test

val<-res$validate

valcb<-sampleCaseBase(data = trainOG,time="time",event="status",ratio=100)
valcb$offset<-0
valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
 #valcbtensor<-list(list(as.matrix(val[,-c(ncol(val))]),  matrix(0,ncol=1,nrow=nrow(val))), as.matrix(val[,c(ncol(val))]) )
pos<-1

while (pos <= iteration) {

  samp<-sample(seq(1,nrow(trainOG),by=1) , nrow(trainOG),replace = T) 
  train<-trainOG[samp,]
  times<-seq(from=min(test$time),
             to=max(test$time),
             length.out = 200
  )
  times<-times[-c(1,length(times))]

  print('starting cbspline')
  ######################
  ###casebase + splines
  ######################
  mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
                                data = train,
                                time = "time",
                                event="status",
                                ratio = 100
  )

  glmAbsRisk<-as.data.frame(absoluteRisk(mod_cb_glm,time=times,newdata=test,type="CI"))
  rownames(glmAbsRisk)<-glmAbsRisk$time
  #prepare to run through riskRegression
  glmProper<- t(glmAbsRisk[-1,-1])
  class(glmProper)<-c("tunnel",class(glmProper))

   print('starting deepsurv')
  ######################
  ###DeepSurv
  ######################
#   coxnn<-deepsurv(formula=Surv(time, status) ~ .,data = train, frac = 0.2, activation = 'tanh',
#                   num_nodes = c(layer1,layer2), dropout = drpt,
#                   early_stopping = TRUE, epochs = epo,
#                   batch_size = bsize,device = "cpu",best_weights=T,
#                   patience = patience,min_delta=min_delta,batch_norm = T,
#                   learning_rate = 0.0001,lr_decay=10^-7)
#   coxnnsurv<- predict(coxnn,newdata = test,type="survival")
#   colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
#   cnnCleaned<-pyProcess(coxnnsurv,times=times)
#   colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))

  source_python('src/deepsurver.py')
  #need to break ties
  tiebreakTrain<-train

  #while(length(unique(tiebreakTrain$time))/length(tiebreakTrain$time)!=1 ){
  #tiebreakTrain<-train
  #tiebreakTrain$time<-tiebreakTrain$time+runif(nrow(tiebreakTrain),0,1/10000000000)
  #}

  coxnnsurv=fitDeepSurv(tiebreakTrain,fullTest,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt)
  coxnnsurv<-t(coxnnsurv)
  #colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
  cnnCleaned<-pyProcess(coxnnsurv,times=times)
  colnames(cnnCleaned)<-times#seq(0,1,length.out = ncol(cnnCleaned))
  py_run_string("del fitDeepSurv")
  rm(fitDeepSurv)

   print('starting deephit')
  ######################
  ###Deephit
  ######################
  source_python('src/oldmortDeephitter.py')
  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt)
  hitnnSurv<-t(hitnnSurv)
  #colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
  deephitCleaned<-pyProcess(hitnnSurv,times=times)
  colnames(deephitCleaned)<-times#seq(0,1,length.out = ncol(deephitCleaned))
  py_run_string("del fitDeephit")
  rm(fitDeephit)

   print('starting cbnnsplines')
  ######################
  ###PMNN splines
  ######################
  ###########################################
###########################################
  valcbtensor[[1]][[2]]<-rep(mod_cb_glm$offset[1],nrow(valcbtensor[[1]][[1]]))
###########################################  
###########################################  
  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2+3),
                            name = 'main_input'
  )
  covars_output<-covars_input%>%
    layer_dense(units=layer1,use_bias = T)%>%
    layer_batch_normalization()%>%
    activation_tanh()%>%
    layer_dropout(drpt)%>%

    layer_dense(units=layer2,use_bias = T)%>%
    layer_batch_normalization()%>%
    activation_tanh()%>%
    layer_dropout(drpt)%>%
    layer_dense(units=1,use_bias = T)#%>%activation_tanh()

  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                  feature_input = covars_input,
                  feature_output = covars_output,
                  originalData = mod_cb_glm$data,
                  offset=mod_cb_glm$offset,
                  timeVar = "time",
                  eventVar= "status",
                  optimizer=optimizer_adam(learning_rate = 0.001)#,decay=10^-7)
  )

  fit<-fitSmoothHazSpline(pmnn,
                          epochs=epo,
                          batch_size=bsize,
                          verbose=0,
                          monitor="val_loss",
                          #val_split=0.2,
                          min_delta=min_delta,
                          patience=patience,val=valcbtensor)
  annPreds<-aarSplines(fit,
                       times=times,
                       x_test=test
  )
  rownames(annPreds)<-annPreds[,1]
  annProper<- t(annPreds[,-1])
  class(annProper)<-c("tunnel",class(annProper))
  #ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
  #geom_line()+coord_cartesian(ylim=c(-0.2,0.5))
  rm(pmnn,covars_output,covars_input,fit)

 print('starting cox')
  ######################
  ###coxph
  ######################
  cox<-coxph(Surv(time, status) ~ ., data = train,x=T)

   print('starting cbnn')
  ######################
  ###PMNN
  ######################
  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
                            name = 'main_input')

  covars_output<-covars_input%>% layer_batch_normalization()%>%
    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=1,use_bias = T)

  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                  feature_input = covars_input,
                  feature_output = covars_output,
                  originalData = mod_cb_glm$data,
                  offset=mod_cb_glm$offset,
                  timeVar = "time",
                  eventVar= "status",
                  optimizer=optimizer_adam(learning_rate = 0.0001,decay=10^-7)
  )

  fit<-fitSmoothHaz(pmnn,
                    epochs=epo,
                    batch_size=bsize,
                    verbose=0,
                    monitor="loss",
                    #val_split=0.3,
                    min_delta=min_delta,
                    patience=patience,val=valcbtensor)
  annPreds<-aar(fit,
                times=times,
                x_test=test
  )

  rownames(annPreds)<-annPreds[,1]
  annProperpoly<- t(annPreds[,-1])
  class(annProperpoly)<-c("tunnel",class(annProperpoly))

  #current best valLoss=0.01048
   print('starting brier')
  ######################
  ###IPA metric
  ######################
  brierFinalResults <- Score(list("Cox_Lin" = cox,'CB_Logi'=glmProper,
                                  #'DeepSurv'=cnnCleaned,'DeepHit'=deephitCleaned,
                                  'PMNN_Spline'=annProper,'PMNN_Poly'=annProperpoly),#
                             data =fullTest, 
                             formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)
  BrierIPAList[[pos]]<-brierFinalResults$Brier$score

  ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
  geom_line()+coord_cartesian(ylim=c(-1,1))
  fit$resultOfFit
  rm(pmnn,covars_output,covars_input,fit)

  ######################
  ###Cox###           ##
  ######################
  a<-survfit(cox, newdata=fullTest,times=times)
  rownames(a$surv)<-a$time
  coxlinSurv<-pyProcess(t(a$surv),times=times)
  colnames(coxlinSurv)<-times#seq(0,1,length.out = ncol(coxlinSurv))

   print('starting cindex')
  ######################
  ###C-Index
  ######################

  riskList<-list(coxlinSurv,
                 glmProper,
                 cnnCleaned,
                 deephitCleaned,
                 annProper,
                 annProperpoly
  )

  tempTims<-   as.numeric(colnames(annProper))
  tempTims<-head(tempTims, -1)
  tempTims<-tail(tempTims, -1)
  cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=length(riskList)+1)
  et_train<-as.matrix(train)[,c(15,16)]
  et_test<-as.matrix(fullTest[,c(15,16)])
  colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
                          'PMNN_Spline','PMNN_Poly')
  cScoreTemp[,1]<- tempTims

  cScore[[pos]]<-cIndexSummary(et_train=et_train,
                               et_test=et_test,
                               riskList=riskList,
                               cScore=cScoreTemp
  )

 saveRDS(BrierIPAList,'results/100oldBrier.rds')
  saveRDS(cScore,'results/100oldcidx.rds')
  print(pos)
  pos<-pos+1
}

ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=Brier,col=model))+
  geom_line()
ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
  geom_line()

end_time <- Sys.time()

end_time-start_time

```

```{r tuning,eval=F}
library(kerastuneR)
  ######################
  ###PMNN
  ######################
#samp<-sample(seq(1,nrow(trainOG),by=1), floor(nrow(trainOG)*.80)) 
#validation<-trainOG[-samp,]

#train<-trainOG[samp,]

  #mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
  #                              data = train,
  #                              time = "time",
  #                              event="status",
  #                              ratio = 100
  #)

buildModel<-function(hp){
  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
                            name = 'main_input')

  covars_output<-covars_input%>% layer_batch_normalization()%>%
    layer_dense(units=hp$Int('units1',min_value=50,max_value=500,step=25),
                use_bias = T,activation = 'tanh')%>%
    layer_dropout(hp$Choice('rate1',
                values=c(0.4,0.3,0.2,0.1,0.05)))%>%
    layer_dense(units=hp$Int('units2',min_value=50,max_value=500,step=25),
                use_bias = T,activation = 'tanh')%>%
    layer_dropout(hp$Choice('rate2',
                values=c(0.4,0.3,0.2,0.1,0.05)))%>%
    layer_dense(units=1,use_bias = T)

  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                  feature_input = covars_input,
                  feature_output = covars_output,
                  originalData = mod_cb_glm$data,
                  offset=mod_cb_glm$offset,
                  timeVar = "time",
                  eventVar= "status",
                  optimizer=optimizer_adam(learning_rate = 0.0001)#,decay=10^-7)
  )
  return(pmnn$model)
}

  tuner = RandomSearch(
    buildModel,
    objective = 'val_loss',
    max_trials = 20,
    executions_per_trial = 5,
    directory = 'tuning',
    project_name = 'oldmort')

  tuner %>% search_summary()

  xtr<-as.matrix(mod_cb_glm$data[,-c(ncol(mod_cb_glm$data),ncol(mod_cb_glm$data)-1)])
  ytr<-as.matrix(mod_cb_glm$data[,c(ncol(mod_cb_glm$data)-1)])
  xv<-valcbtensor[[1]]#list(as.matrix(validation[,-c(ncol(validation))]),as.matrix(data.frame(offset=rep(0,nrow(validation)))))
  yv<-valcbtensor[[2]]#as.matrix(validation[,c(ncol(validation)),drop=F])
  tuner %>% fit_tuner(x=list(xtr,as.matrix(mod_cb_glm$offset)),y=ytr,
                    epochs = 55,batch_size=bsize, 
                    validation_data = list(xv,yv),verbose=0 )
result = kerastuneR::plot_tuner(tuner)
# the list will show the plot and the data.frame of tuning results
result

best_5_models = tuner %>% get_best_models(5)
best_5_models[[1]] %>% plot_keras_model()
  # 
  # fit<-fitSmoothHaz(pmnn,
  #                   epochs=epo,
  #                   batch_size=bsize+1000,
  #                   verbose=0,
  #                   monitor="val_loss",
  #                   val_split=0.2,
  #                   min_delta=min_delta,
  #                   patience=patience+10)
  # annPreds<-aar(fit,
  #               times=times,
  #               x_test=test
  # )
  # 
  # rownames(annPreds)<-annPreds[,1]
  # annProperpoly<- t(annPreds[,-1])
  # class(annProperpoly)<-c("tunnel",class(annProperpoly)) 
  #

```

### 2022-12-15 analyses/complexSim.Rmd
re-analyzed

@@
   patience: 10
-  iteration: 2
+  iteration: 1
   layer1: 50
@@
   layer4: 25
-  drpt: 0.5
+  drpt: 0.1
 ---
@@
 bsize=512
-min_delta = 10^-7
+min_delta = 0
 epo=as.numeric(params$epo)
@@
 
+#########################
+###Simulate date
+#########################
+  cov<-data.frame(id=1:numSamp,
+                  z1=c(rep(1,numSamp/2),rep(0,numSamp/2)),
+                  z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
+                  z3=c(rnorm(numSamp/2,-3,0.5),rnorm(numSamp/2,8,0.5)))
+  true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1,
+                                       data = brcancer, k = 3
+  ) #from flexsurv
+  dat <- simsurv(betas = coefficients, # "true" parameter values
+                 x = cov,                   # covariate data for 200 individuals
+                 knots = true_mod$knots,    # knot locations for splines
+                 logcumhazard = logcumhaz,  # definition of log cum hazard
+                 maxt = 1,               # no right-censoring
+                 interval = c(0,10000000)) # interval for root finding
+  colnames(dat)<-c("id","time","status")
+  samp<-sample(seq(1,nrow(dat),by=1), floor(nrow(dat)*(0.1))) #random censoring
+  dat$status[samp]<-0
+  # Merge the simulated event times onto covariate data frame
+  data <- merge(cov, dat)
+  covs<-colnames(data)
+  data<-data[,-1]
+
+ spec = c(train = .7, test = .1, validate = .2)
+
+ g = sample(cut(
+   seq(nrow(data)), 
+   nrow(data)*cumsum(c(0,spec)),
+   labels = names(spec)
+ ))
+
+ res = split(data, g)
+
+trainOG<-res$train
+
+test<-res$test[,-c(ncol(data))]
+
+fullTest<-res$test
+
+val<-res$validate
+
+valcb<-sampleCaseBase(data = val,time="time",event="status",ratio=100)
+valcb$offset<-0
+valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
+
+numSamp<-nrow(trainOG)
+
 start_time <- Sys.time()
@@
   covs<-colnames(data)
-
+  data<-data[,-1]
   #########################
@@
   #########################
-  samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.8))
-  data<-data[,-1]
-  data[,-c(1,5)]<-as.data.frame(sapply(data[,-c(1,5)], function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
-  train<-data[samp,]
-  test<-data[-samp,-c(ncol(data))]
-  fullTest<-data[-samp,,drop=F]
-  times<-seq(from=min(test$time),
-             to=max(test$time),
-             length.out = 200
+  #samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.8))
+  #data<-data[,-1]
+  #data[,-c(1,5)]<-as.data.frame(sapply(data[,-c(1,5)], function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
+  train<-data
+  #test<-data[-samp,-c(ncol(data))]
+  #fullTest<-data[-samp,,drop=F]
+  times<-seq(from=min(test$time)+0.000001,
+             to=max(test$time)-0.000001,
+             length.out = 100
   )
-  times<-times[-c(1,length(times))]
+  #times<-times[-c(1,length(times))]
 
@@
   ###casebase+splines
-  #########################
+ ##########################
     mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
@@
 
-  glmAbsRisk<-absoluteRisk(mod_cb_glm,time=times,newdata=test,type="CI")
-
-  glmAbsRisk<-as.data.frame(glmAbsRisk)
-
+  glmAbsRisk<-as.data.frame(absoluteRisk(mod_cb_glm,time=times,newdata=test,type="CI"))
   rownames(glmAbsRisk)<-glmAbsRisk$time
+  #prepare to run through riskRegression
   glmProper<- t(glmAbsRisk[-1,-1])
-
   class(glmProper)<-c("tunnel",class(glmProper))
@@
   covars_output<-covars_input%>% #layer_batch_normalization()%>%
-    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
+    layer_dense(units=layer1,use_bias = T,activation = 'tanh')%>%
     layer_dropout(drpt)%>%
-    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
+    layer_dense(units=layer2,use_bias = T,activation = 'tanh')%>%
     layer_dropout(drpt)%>%
-    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
+    layer_dense(units=layer3,use_bias = T,activation = 'tanh')%>%
     layer_dropout(drpt)%>%
-    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
+    layer_dense(units=layer4,use_bias = T,activation = 'tanh')%>%
     layer_dropout(drpt)%>%
@@
                   timeVar = "time",
-                  eventVar= "status"
+                  eventVar= "status",optimizer=optimizer_adam(learning_rate = 0.0001)
   )
@@
                           monitor="val_loss",
-                          val_split=0.2,
+                          #val_split=0.2,
                           min_delta=min_delta,
-                          patience=patience)
+                          patience=patience,val=valcbtensor)
+
    annPreds<-aarSplines(fit,
@@
 
-  ######################
-  ###DeepSurv
-  ######################
-  coxnn<-deepsurv(Surv(time, status) ~ .,data = train, frac = 0.2, activation = "relu",
-                  num_nodes = c(layer1,layer2,layer3,layer4), dropout = drpt,
-                  early_stopping = TRUE, epochs = epo,
-                  batch_size = bsize,device = "cpu",best_weights=T,
-                  patience = patience,min_delta=min_delta,batch_norm = F,
-                  learning_rate = 0.001,lr_decay=10^-7)
-  coxnnsurv<- predict(coxnn,newdata = test,type="survival")
-  colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
+  source_python('src/dsurv.py')
+  #need to break ties
+  tiebreakTrain<-train
+
+  #while(length(unique(tiebreakTrain$time))/length(tiebreakTrain$time)!=1 ){
+  #tiebreakTrain<-train
+  #tiebreakTrain$time<-tiebreakTrain$time+runif(nrow(tiebreakTrain),0,1/10000000000)
+  #}
+
+  coxnnsurv=fitDeepSurv(tiebreakTrain,fullTest,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4)
+  coxnnsurv<-t(coxnnsurv)
+  #colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
   cnnCleaned<-pyProcess(coxnnsurv,times=times)
-  colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
+  colnames(cnnCleaned)<-times#seq(0,1,length.out = ncol(cnnCleaned))
+  py_run_string("del fitDeepSurv")
+  rm(fitDeepSurv)
 
+  # ######################
+  # ###DeepSurv
+  # ######################
+  # coxnn<-deepsurv(Surv(time, status) ~ .,data = train, frac = 0.2, activation = "relu",
+  #                 num_nodes = c(layer1,layer2,layer3,layer4), dropout = drpt,
+  #                 early_stopping = TRUE, epochs = epo,
+  #                 batch_size = bsize,device = "cpu",best_weights=T,
+  #                 patience = patience,min_delta=min_delta,batch_norm = F,
+  #                 learning_rate = 0.001,lr_decay=10^-7)
+  # coxnnsurv<- predict(coxnn,newdata = test,type="survival")
+  # #colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
+  # cnnCleaned<-pyProcess(coxnnsurv,times=times)
+  # colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
+  # 
   ######################
@@
   source_python('src/deephitter.py')
-  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,patience=patience,min_delta=min_delta,drpt=drpt)
+  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lay3=layer3,lay4=25)
   hitnnSurv<-t(hitnnSurv)
-  colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
+  #colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
   deephitCleaned<-pyProcess(hitnnSurv,times=times)
@@
                     min_delta=min_delta,
-                    patience=patience)
+                    patience=patience,val=valcbtensor)
   annPreds<-aar(fit,
@@
                                   'PMNN_Poly'=annProperpoly),
-                             #'dsm'=dsmSurvProper) ,#'dsm'=dsmCleaned,
                              data =fullTest, 
@@
 ```
+
+```{r}
+
+  ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
+  geom_line()+coord_cartesian(ylim=c(-1,1))
+
+```

### 2022-12-15 analyses/editFreezeoldmortality.Rmd
re-analyzed

--- 2022-12-01 analyses/oldmortality.Rmd
@@
                  birthplace=oldmort$birthplace,
-                 ageAtDeath=oldmort$exit,
+                 ageAtEntry=oldmort$enter,
                  imrAtBirth=oldmort$imr.birth,
@@
 
-preData$time<-preData$time+runif(nrow(preData),-1,1)/1000000
+#preData$time<-preData$time+runif(nrow(preData),-1,1)/1000000
 
 data<-as.data.frame((model.matrix(~.,preData)[,-1]))
-
-#data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
+data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
 covs<-colnames(data)
@@
   ######################
-#   a=get_pycox_activation(
-#   activation = "tanh",
-#   construct = TRUE,
-#   alpha = 1,
-#   dim = NULL,
-#   lambd = 0.5,
-#   min_val = -1,
-#   max_val = 1,
-#   negative_slope = 0.01,
-#   num_parameters = 1L,
-#   init = 0.25,
-#   lower = 1/8,
-#   upper = 1/3,
-#   beta = 1,
-#   threshold = 20,
-#   value = 20
-# )
-#   coxnn<-deepsurv(formula=Surv(time, status) ~ .,data = train, frac = 0.2, activation = 'tanh',
-#                   num_nodes = c(layer1,layer2), dropout = drpt,
-#                   early_stopping = TRUE, epochs = epo,
-#                   batch_size = bsize,device = "cpu",best_weights=T,
-#                   patience = patience,min_delta=min_delta,batch_norm = T,
-#                   learning_rate = 0.0001,lr_decay=10^-7)
-#   coxnnsurv<- predict(coxnn,newdata = test,type="survival")
-#   colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
+
+  coxnn<-deepsurv(formula=Surv(time, status) ~ .,data = train, frac = 0.2, activation = 'selu',
+                  num_nodes = c(layer1,layer2), dropout = drpt,
+                  early_stopping = TRUE, epochs = epo,
+                  batch_size = bsize,device = "cpu",best_weights=T,
+                  patience = patience,min_delta=min_delta,batch_norm = T,
+                  learning_rate = 0.0001,lr_decay=10^-7)
+  coxnnsurv<- predict(coxnn,newdata = test,type="survival")
+  #colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
+  cnnCleaned<-pyProcess(coxnnsurv,times=times)
+  colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
+
+  # source_python('src/deepsurver.py')
+  # coxnnsurv=fitDeepSurv(train,fullTest,bsize,epochs=epo,patience=patience,min_delta=min_delta,drpt=drpt)
+  # coxnnsurv<-t(coxnnsurv)
+  # #colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
   # cnnCleaned<-pyProcess(coxnnsurv,times=times)
   # colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
-
-  source_python('src/deepsurver.py')
-  coxnnsurv=fitDeepSurv(train,fullTest,bsize,epochs=epo,patience=patience,min_delta=min_delta,drpt=drpt)
-  coxnnsurv<-t(coxnnsurv)
-  colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
-  cnnCleaned<-pyProcess(coxnnsurv,times=times)
-  colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
-  py_run_string("del fitDeepSurv")
-  rm(fitDeepSurv)
+  # py_run_string("del fitDeepSurv")
+  # rm(fitDeepSurv)
 
@@
   source_python('src/oldmortDeephitter.py')
-  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,patience=patience,min_delta=min_delta,drpt=drpt)
+  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2)
   hitnnSurv<-t(hitnnSurv)
-  colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
+  #colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
   deephitCleaned<-pyProcess(hitnnSurv,times=times)
-  colnames(deephitCleaned)<-seq(0,1,length.out = ncol(deephitCleaned))
+  #colnames(deephitCleaned)<-seq(0,1,length.out = ncol(deephitCleaned))
   py_run_string("del fitDeephit")
   rm(fitDeephit)
+
+  #######################
+  ###Cox 
+  ######################
+  cox<-coxph(Surv(time, status) ~ ., data = train,x=T)
+  a<-survfit(cox, newdata=fullTest,times=times)
+  rownames(a$surv)<-a$time
+  coxlinSurv<-pyProcess(t(a$surv),times=times)
+  colnames(coxlinSurv)<-seq(0,1,length.out = ncol(coxlinSurv))
 
@@
   )
-  covars_output<-covars_input%>%layer_batch_normalization()%>%
+  covars_output<-covars_input%>%
     layer_dense(units=layer1,use_bias = T)%>%
-    activation_tanh()%>%layer_dropout(drpt)%>%
+    activation_selu()%>%
+    layer_batch_normalization()%>%
+    layer_dropout(drpt)%>%
     layer_dense(units=layer2,use_bias = T)%>%
-    activation_tanh()%>%layer_dropout(drpt)%>%
+    activation_selu()%>%
+    layer_batch_normalization()%>%
+    layer_dropout(drpt)%>%
     layer_dense(units=1,use_bias = T)#%>%activation_tanh()
@@
                           epochs=epo,
-                          batch_size=5120,
+                          batch_size=512,
                           verbose=0,
@@
   ######################
-  ###coxph
-  ######################
-  cox<-coxph(Surv(time, status) ~ ., data = train,x=T)
-
-  ######################
   ###PMNN
@@
 
-  covars_output<-covars_input%>%layer_batch_normalization()%>%
+  covars_output<-covars_input%>%
     layer_dense(units=layer1,use_bias = T)%>%
-    activation_tanh()%>%layer_dropout(drpt)%>%
+    activation_selu()%>%
+    layer_batch_normalization()%>%
+    layer_dropout(drpt)%>%
+
     layer_dense(units=layer2,use_bias = T)%>%
-    activation_tanh()%>%layer_dropout(drpt)%>%
+    activation_selu()%>%
+    layer_batch_normalization()%>%
+    layer_dropout(drpt)%>%
     layer_dense(units=1,use_bias = T)#%>%activation_tanh()
@@
                     epochs=epo,
-                    batch_size=512,
+                    batch_size=bsize,
                     verbose=0,
@@
   ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
-  geom_line()+coord_cartesian(ylim=c(-1,0.7))
+  geom_line()+coord_cartesian(ylim=c(-0.5,0.7))
   rm(pmnn,covars_output,covars_input,fit)
-  #######################
-  ###Cox 
-  ######################
-  a<-survfit(cox, newdata=fullTest,times=times)
-  rownames(a$surv)<-a$time
-  coxlinSurv<-pyProcess(t(a$surv),times=times)
-  colnames(coxlinSurv)<-seq(0,1,length.out = ncol(coxlinSurv))
 
@@
 ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
-  geom_line()+coord_cartesian(ylim=c(-2,0.7))
+  geom_line()
 

### 2022-12-15 analyses/gbsg.Rmd
re-analyzed

--- 2022-12-15 analyses/VeteransAnalysis.Rmd
@@
   iteration: 1
-  layer1: 50
+  layer1: 150
   layer2: 20
@@
 #library(eha)
-veteran<-veteran
+gbsg<-survival::gbsg
 
-coxph(Surv(time, status) ~ trt + prior + karno, veteran)
-
-preData<-data.frame(trt=veteran$trt,
-                    prior=veteran$prior,
-                    karno=veteran$karno,
-                    time=veteran$time,
-                    status=veteran$status
-                    )
+colnames(gbsg)[10]<-"time"
+preData<-gbsg[,-1]
 #preData$time<-preData$time+runif(nrow(preData),0,1)/100000000
@@
 
- spec = c(train = .90, test = .10)#, validate = 0)
+ spec = c(train = .80, test = .20)
 
@@
 
-val<-res$validate
+val<-res$train#res$validate
 
@@
   )
-  covars_output<-covars_input%>%
-    layer_dense(units=layer1,use_bias = T)%>%
-    layer_batch_normalization()%>%
-    activation_tanh()%>%
+  covars_output<-covars_input%>% layer_batch_normalization()%>%
+    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
     layer_dropout(drpt)%>%
-
-    layer_dense(units=layer2,use_bias = T)%>%
-    layer_batch_normalization()%>%
-    activation_tanh()%>%
+    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
     layer_dropout(drpt)%>%
-    layer_dense(units=1,use_bias = T)#%>%activation_tanh()
+    layer_dense(units=1,use_bias = T)
 
@@
                   eventVar= "status",
-                  optimizer=optimizer_adam(learning_rate = 0.001)#,decay=10^-7)
+                  optimizer=optimizer_adam(learning_rate = 0.0001)#,decay=10^-7)
   )
@@
   covars_output<-covars_input%>% layer_batch_normalization()%>%
-    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
+    layer_dense(units=layer1,use_bias = T,activation='relu')%>%
+
     layer_dropout(drpt)%>%
     layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
+    #layer_batch_normalization()%>%
     layer_dropout(drpt)%>%
-    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
+    #    layer_dense(units=25,use_bias = T,activation = 'relu')%>%
+    #layer_dropout(0.2)%>%
+    #    layer_dense(units=25,use_bias = T,activation = 'relu')%>%
+    #layer_dropout(0.2)%>%
     layer_dense(units=1,use_bias = T)
@@
                   eventVar= "status",
-                  optimizer=optimizer_adam(learning_rate = 0.0001,decay=10^-7)
+                  optimizer=optimizer_adam(learning_rate = 0.0001)
   )
@@
                     epochs=epo,
-                    batch_size=bsize,
+                    batch_size=bsize,#nrow(mod_cb_glm$data),
                     verbose=0,
-                    monitor="loss",
+                    monitor="val_loss",
                     #val_split=0.3,
@@
   brierFinalResults <- Score(list("Cox_Lin" = cox,'CB_Logi'=glmProper,
-                                  #'DeepSurv'=cnnCleaned,'DeepHit'=deephitCleaned,
+                                  'DeepSurv'=cnnCleaned,'DeepHit'=deephitCleaned,
                                   'PMNN_Spline'=annProper,'PMNN_Poly'=annProperpoly),#
@@
   ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
-  geom_line()+coord_cartesian(ylim=c(-1,1))
+  geom_line()+coord_cartesian(ylim=c(-0.5,0.5))
   fit$resultOfFit
@@
   cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=length(riskList)+1)
-  et_train<-as.matrix(train)[,c(15,16)]
-  et_test<-as.matrix(fullTest[,c(15,16)])
+  et_train<-as.matrix(train)[,c(10,9)]
+  et_test<-as.matrix(fullTest[,c(10,9)])
   colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
@@
 
- saveRDS(BrierIPAList,'results/100oldBrier.rds')
-  saveRDS(cScore,'results/100oldcidx.rds')
+ saveRDS(BrierIPAList,'results/100gbsgBrier.rds')
+  saveRDS(cScore,'results/100gbsgcidx.rds')
   print(pos)

### 2022-12-15 analyses/metabricStudy.Rmd
re-analyzed

---
title: "Metabric"
output: html_document
params:
  epo: 2000
  patience: 10
  iteration: 1
  layer1: 50
  layer2: 50
  layer3: 25
  layer4: 25
  drpt: 0.1
---

```{r,eval=T}

library(reticulate)
```

```{python,eval=T}

import os
os.environ["CUDA_DEVICE_ORDER"] = "PCI_BUS_ID" 
os.environ["CUDA_VISIBLE_DEVICES"] = ""
#os.environ["OM_NUM_THREADS"]="2"

```

```{r setup, include=T}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
source("src/packages.R")
source("src/functions.R")
eval_metabric<-T
pycox<-reticulate::import("pycox")
source_python('src/getMetabricData.py')
source_python("src/cIndex.py")
source_python("src/dsmScript.py")
source_python('src/deephitter.py')
##########################
###Shared Hyperparameters
##########################
bsize=512
min_delta = 0
epo=as.numeric(params$epo)
patience = as.numeric(params$patience)
iteration=as.numeric(params$iteration)
layer1<-as.numeric(params$layer1)
layer2<-as.numeric(params$layer2)
layer3<-as.numeric(params$layer3)
layer4<-as.numeric(params$layer4)
drpt<-as.numeric(params$drpt)

```

```{r modelsMETABRIC,eval=eval_metabric}

start_time <- Sys.time()

#pycox<-reticulate::import("pycox")
#numpy<-reticulate::import("numpy")
#a<-pycox$datasets$metabric$read_df
#b<-numpy$asarray(a)

############################
###Data clean and prep
############################

metabric<-getMetabric()
colnames(metabric)[ncol(metabric)-1]<-'time'
colnames(metabric)[ncol(metabric)]<-'status'
data<-as.data.frame(model.matrix(~.,metabric)[,-1])
data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
covs<-colnames(data)

 spec = c(train = .7, test = .1, validate = .2)

 g = sample(cut(
   seq(nrow(data)), 
   nrow(data)*cumsum(c(0,spec)),
   labels = names(spec)
 ))

 res = split(data, g)

trainOG<-res$train

test<-res$test[,-c(ncol(data))]

fullTest<-res$test

val<-res$validate

valcb<-sampleCaseBase(data = val,time="time",event="status",ratio=100)
valcb$offset<-0
valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))

  times<-seq(from=min(test$time)+0.000001,
             to=max(test$time)-0.000001,
             length.out = 100
  )

############################
###Prep Loop
############################
BrierIPAList<-list()
cScore<-list()
pos<-1

while (pos <= iteration) {

  ############################
  ###Bootstrap samples
  ############################  
  samp<-sample(seq(1,nrow(trainOG),by=1) , nrow(trainOG),replace = T) 
  train<-trainOG[samp,]
  sds<-sapply(train, function(x) (sd(x)))
means<-sapply(train, function(x) (mean(x)))
test<-normalizer(res$test,means,sds)[,-c(ncol(data))]
fullTest<-normalizer(res$test,means,sds)
val<-normalizer(res$validate,means,sds)

train<-normalizer(train,means,sds)

  ##########################
  ###DSM
  ##########################
  source_python('src/dsmScript.py')
  dsmSurv<-dsmfitter(train=train,fullTest=fullTest,bsize=bsize,epochs=epo,val=val,times=times,tPos=ncol(train)-2,ePos=ncol(train)-1)
  if((sum(is.na(dsmSurv))==ncol(dsmSurv)*nrow(dsmSurv))){

    dsmSurv<-dsmfitter(train=train,fullTest=fullTest,bsize=bsize,epochs=epo,val=val,times=times)
    py_run_string("del dsmfitter")
    rm(dsmfitter)
    next
  }
  dsmSurv<-as.data.frame(dsmSurv)
  colnames(dsmSurv) <-times
  dsmSurvProper<-pyProcess(dsmSurv,times=times)
  colnames(dsmSurvProper) <-seq(0,1,length.out = ncol(dsmSurvProper))
  py_run_string("del dsmfitter")
  rm(dsmfitter)

  #########################
  ###casebase+splines
 ##########################
    mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
                                data = train,
                                time = "time",
                                event="status",
                                ratio = 100
  )

  glmAbsRisk<-as.data.frame(absoluteRisk(mod_cb_glm,time=times,newdata=test,type="CI"))
  rownames(glmAbsRisk)<-glmAbsRisk$time
  #prepare to run through riskRegression
  glmProper<- t(glmAbsRisk[-1,-1])
  class(glmProper)<-c("tunnel",class(glmProper))

  # 
  # 
  # ############################
  # ###DeepSurv
  # ############################

  source_python('src/dsurv.py')
  #need to break ties
  tiebreakTrain<-train

  #while(length(unique(tiebreakTrain$time))/length(tiebreakTrain$time)!=1 ){
  #tiebreakTrain<-train
  #tiebreakTrain$time<-tiebreakTrain$time+runif(nrow(tiebreakTrain),0,1/10000000000)
  #}

  coxnnsurv=fitDeepSurv(tiebreakTrain,fullTest,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4)
  coxnnsurv<-t(coxnnsurv)
  #colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
  cnnCleaned<-pyProcess(coxnnsurv,times=times)
  colnames(cnnCleaned)<-times#seq(0,1,length.out = ncol(cnnCleaned))
  py_run_string("del fitDeepSurv")
  rm(fitDeepSurv)

  # coxnn<-deepsurv(Surv(time, status) ~ .,data = train, frac = 0.2, activation = "relu",
  #                 num_nodes = c(layer1,layer2,layer3,layer4), dropout = drpt,
  #                 early_stopping = TRUE, epochs = epo,
  #                 batch_size = bsize,device = "cpu",best_weights=T,
  #                 patience = patience,min_delta=min_delta,batch_norm = F,
  #                 learning_rate = 0.001,lr_decay=10^-7)
  # 
  # coxnnsurv<- predict(coxnn,newdata = test,type="survival")
  # colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
  # cnnCleaned<-pyProcess(coxnnsurv,times=times)
  # colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
  ############################
  ###Deephit
  ############################  
    source_python('src/deephitter.py')
  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4)
  hitnnSurv<-t(hitnnSurv)
  #colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
  deephitCleaned<-pyProcess(hitnnSurv,times=times)
  colnames(deephitCleaned)<-seq(0,1,length.out = ncol(deephitCleaned))
  py_run_string("del fitDeephit")
  rm(fitDeephit)
  # source_python('src/deephitter.py')
  # hitnnSurv=fitDeephit(train,fullTest,bsize,epo,patience=patience,min_delta=min_delta,drpt=drpt)
  # hitnnSurv<-t(hitnnSurv)
  # colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
  # deephitCleaned<-pyProcess(hitnnSurv,times=times)
  # colnames(deephitCleaned)<-seq(0,1,length.out = ncol(deephitCleaned))
  # py_run_string("del fitDeephit")
  # rm(fitDeephit)
  # 
  ############################
  ###PMNNsplines
  ############################  
  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2+3),
                            name = 'main_input'
  )
  covars_output<-covars_input%>% #layer_batch_normalization()%>%
    layer_dense(units=layer1,use_bias = T,activation = 'tanh')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer2,use_bias = T,activation = 'tanh')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer3,use_bias = T,activation = 'tanh')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer4,use_bias = T,activation = 'tanh')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=1,use_bias = T)

  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                  feature_input = covars_input,
                  feature_output = covars_output,
                  originalData = mod_cb_glm$data,
                  offset=mod_cb_glm$offset,
                  timeVar = "time",
                  eventVar= "status",optimizer=optimizer_adam(learning_rate = 0.0001)
  )

  fit<-fitSmoothHazSpline(pmnn,
                          epochs=epo,
                          batch_size=bsize,
                          verbose=0,
                          monitor="val_loss",
                          #val_split=0.2,
                          min_delta=min_delta,
                          patience=patience,val=valcbtensor)

   annPreds<-aarSplines(fit,
                       times=times,
                       x_test=test
  )
  rownames(annPreds)<-annPreds[,1]
  annProper<- t(annPreds[,-1])
  class(annProper)<-c("tunnel",class(annProper))

  ############################
  ###coxph
  ############################  
  cox<-coxph(Surv(time, status) ~ ., data = train,x=T)

  ######################
  ###PMNN  
  ######################
  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
                            name = 'main_input')

  covars_output<-covars_input%>% #layer_batch_normalization()%>%
    layer_dense(units=layer1,use_bias = T,activation = 'tanh')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer2,use_bias = T,activation = 'tanh')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer3,use_bias = T,activation = 'tanh')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer4,use_bias = T,activation = 'tanh')%>%
    layer_dropout(drpt)%>%
    layer_dense(units=1,use_bias = T)

  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                  feature_input = covars_input,
                  feature_output = covars_output,
                  originalData = mod_cb_glm$data,
                  offset=mod_cb_glm$offset,
                  timeVar = "time",
                  eventVar= "status",optimizer=optimizer_adam(learning_rate = 0.0001)
  )

  fit<-fitSmoothHaz(pmnn,
                    epochs=epo,
                    batch_size=bsize,
                    verbose=0,
                    monitor="val_loss",
                    val_split=0.2,
                    min_delta=min_delta,
                    patience=patience,val=valcbtensor)
  annPreds<-aar(fit,
                times=times,
                x_test=test
  )

  rownames(annPreds)<-annPreds[,1]
  annProperpoly<- t(annPreds[,-1])
  class(annProperpoly)<-c("tunnel",class(annProperpoly)) 
  ############################
  ###Brier Score
  ############################  
  brierFinalResults <- Score(list("Cox_Lin" = cox,'CB_Logi'=glmProper,
                                  'DeepSurv'=cnnCleaned,'DeepHit'=deephitCleaned,
                                  'PMNN_Spline'=annProper,'PMNN_Poly'=annProperpoly,
                                  'dsm'=dsmSurvProper),
                             data =fullTest, 
                             formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)
  BrierIPAList[[pos]]<-brierFinalResults$Brier$score

  #######################
  ###Cox 
  ######################
  a<-survfit(cox, newdata=fullTest,times=times)
  rownames(a$surv)<-a$time
  coxlinSurv<-pyProcess(t(a$surv),times=times)
  colnames(coxlinSurv)<-seq(0,1,length.out = ncol(coxlinSurv))

  ######################
  ###C-Index
  ######################

  riskList<-list(coxlinSurv,
                 glmProper,
                 cnnCleaned,
                 deephitCleaned,
                 annProper,
                 annProperpoly,
                 dsmSurvProper
  )

  tempTims<-   as.numeric(colnames(annProper))
  tempTims<-head(tempTims, -1)
  tempTims<-tail(tempTims, -1)
  cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=length(riskList)+1)
  et_train<-as.matrix(train)[,c(11,10)]
  et_test<-as.matrix(fullTest[,c(11,10)])
  colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
                          'PMNN_Spline','PMNN_Poly','dsm')
  cScoreTemp[,1]<- tempTims

  cScore[[pos]]<-cIndexSummary(et_train=et_train,
                               et_test=et_test,
                               riskList=riskList,
                               cScore=cScoreTemp
  )

  saveRDS(BrierIPAList,'results/100MetaBrier.rds')
  saveRDS(cScore,'results/100Metacidx.rds')
  pos=pos+1
} 
end_time <- Sys.time()

end_time-start_time

ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=Brier,col=model))+
  geom_line()
ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
  geom_line()

```

```{r}

  ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
  geom_line()+coord_cartesian(ylim=c(-1,1))

```

```{r}

sessionInfo()

```

### 2022-12-15 analyses/oldmortality.Rmd
re-analyzed

@@
   iteration: 1
-  layer1: 200
+  layer1: 50
   layer2: 20
@@
 bsize=512
-min_delta = 10^-9
+min_delta = 10^-8
 epo=as.numeric(params$epo)
@@
 
-preData$time<-preData$time+runif(nrow(preData),-1,1)/1000000
+preData$time<-preData$time+runif(nrow(preData),0,1)/100000000
 
@@
 
-samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.80)) 
-trainOG<-data[samp,]
-test<-data[-samp,-c(ncol(data))]
+#samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.80))
 
-fullTest<-data[-samp,,drop=F]
+ spec = c(train = .75, test = .10, validate = .15)
+
+ g = sample(cut(
+   seq(nrow(data)), 
+   nrow(data)*cumsum(c(0,spec)),
+   labels = names(spec)
+ ))
+
+ res = split(data, g)
+
+trainOG<-res$train
+
+test<-res$test[,-c(ncol(data))]
+
+fullTest<-res$test
+
+val<-res$validate
+
+valcb<-sampleCaseBase(data = val,time="time",event="status",ratio=100)
+valcb$offset<-0
+valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
+ #valcbtensor<-list(list(as.matrix(val[,-c(ncol(val))]),  matrix(0,ncol=1,nrow=nrow(val))), as.matrix(val[,c(ncol(val))]) )
 
@@
              to=max(test$time),
-             length.out = 100
+             length.out = 200
   )
@@
 
+  print('starting cbspline')
   ######################
@@
 
+   print('starting deepsurv')
   ######################
@@
   ######################
-#   a=get_pycox_activation(
-#   activation = "tanh",
-#   construct = TRUE,
-#   alpha = 1,
-#   dim = NULL,
-#   lambd = 0.5,
-#   min_val = -1,
-#   max_val = 1,
-#   negative_slope = 0.01,
-#   num_parameters = 1L,
-#   init = 0.25,
-#   lower = 1/8,
-#   upper = 1/3,
-#   beta = 1,
-#   threshold = 20,
-#   value = 20
-# )
 #   coxnn<-deepsurv(formula=Surv(time, status) ~ .,data = train, frac = 0.2, activation = 'tanh',
@@
   source_python('src/deepsurver.py')
-  coxnnsurv=fitDeepSurv(train,fullTest,bsize,epochs=epo,patience=patience,min_delta=min_delta,drpt=drpt)
+  #need to break ties
+  tiebreakTrain<-train
+
+  #while(length(unique(tiebreakTrain$time))/length(tiebreakTrain$time)!=1 ){
+  #tiebreakTrain<-train
+  #tiebreakTrain$time<-tiebreakTrain$time+runif(nrow(tiebreakTrain),0,1/10000000000)
+  #}
+
+  coxnnsurv=fitDeepSurv(tiebreakTrain,fullTest,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt)
   coxnnsurv<-t(coxnnsurv)
-  colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
+  #colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
   cnnCleaned<-pyProcess(coxnnsurv,times=times)
-  colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
+  colnames(cnnCleaned)<-times#seq(0,1,length.out = ncol(cnnCleaned))
   py_run_string("del fitDeepSurv")
@@
 
+   print('starting deephit')
   ######################
@@
   source_python('src/oldmortDeephitter.py')
-  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,patience=patience,min_delta=min_delta,drpt=drpt)
+  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt)
   hitnnSurv<-t(hitnnSurv)
-  colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
+  #colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
   deephitCleaned<-pyProcess(hitnnSurv,times=times)
-  colnames(deephitCleaned)<-seq(0,1,length.out = ncol(deephitCleaned))
+  colnames(deephitCleaned)<-times#seq(0,1,length.out = ncol(deephitCleaned))
   py_run_string("del fitDeephit")
@@
 
+   print('starting cbnnsplines')
   ######################
@@
   ######################
+  ###########################################
+###########################################
+  valcbtensor[[1]][[2]]<-rep(mod_cb_glm$offset[1],nrow(valcbtensor[[1]][[1]]))
+###########################################  
+###########################################  
   covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2+3),
@@
   )
-  covars_output<-covars_input%>%layer_batch_normalization()%>%
+  covars_output<-covars_input%>%
     layer_dense(units=layer1,use_bias = T)%>%
-    activation_tanh()%>%layer_dropout(drpt)%>%
+    layer_batch_normalization()%>%
+    activation_tanh()%>%
+    layer_dropout(drpt)%>%
+
     layer_dense(units=layer2,use_bias = T)%>%
-    activation_tanh()%>%layer_dropout(drpt)%>%
+    layer_batch_normalization()%>%
+    activation_tanh()%>%
+    layer_dropout(drpt)%>%
     layer_dense(units=1,use_bias = T)#%>%activation_tanh()
@@
                   eventVar= "status",
-                  optimizer=optimizer_adam(learning_rate = 0.0001,decay=10^-7)
+                  optimizer=optimizer_adam(learning_rate = 0.001)#,decay=10^-7)
   )
@@
                           epochs=epo,
-                          batch_size=5120,
+                          batch_size=bsize,
                           verbose=0,
                           monitor="val_loss",
-                          val_split=0.2,
+                          #val_split=0.2,
                           min_delta=min_delta,
-                          patience=patience)
+                          patience=patience,val=valcbtensor)
   annPreds<-aarSplines(fit,
@@
   class(annProper)<-c("tunnel",class(annProper))
-
+  #ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
+  #geom_line()+coord_cartesian(ylim=c(-0.2,0.5))
   rm(pmnn,covars_output,covars_input,fit)
 
+ print('starting cox')
   ######################
@@
 
+   print('starting cbnn')
   ######################
@@
 
-  covars_output<-covars_input%>%layer_batch_normalization()%>%
-    layer_dense(units=layer1,use_bias = T)%>%
-    activation_tanh()%>%layer_dropout(drpt)%>%
-    layer_dense(units=layer2,use_bias = T)%>%
-    activation_tanh()%>%layer_dropout(drpt)%>%
-    layer_dense(units=1,use_bias = T)#%>%activation_tanh()
+  covars_output<-covars_input%>%
+
+    layer_dense(units=20,use_bias = T)%>%
+     layer_batch_normalization()%>%
+    activation_relu()%>%
+    layer_dropout(drpt)%>%
+
+    layer_dense(units=10,use_bias = T)%>%
+
+    activation_relu()%>%
+    layer_dropout(drpt)%>%
+    layer_dense(units=1,use_bias = T)
 
@@
                   eventVar= "status",
-                  optimizer=optimizer_adam(learning_rate = 0.0001,decay=10^-7)
+                  optimizer=optimizer_adam(learning_rate = 0.0001)#,decay=10^-7)
   )
@@
                     epochs=epo,
-                    batch_size=512,
+                    batch_size=bsize,
                     verbose=0,
                     monitor="val_loss",
-                    val_split=0.2,
+                    #val_split=0.3,
                     min_delta=min_delta,
-                    patience=patience)
+                    patience=patience,val=valcbtensor)
   annPreds<-aar(fit,
@@
 
+  #current best valLoss=0.01048
+   print('starting brier')
   ######################
@@
   ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
-  geom_line()+coord_cartesian(ylim=c(-1,0.7))
+  geom_line()+coord_cartesian(ylim=c(-0.1,0.1))
+  fit$resultOfFit
   rm(pmnn,covars_output,covars_input,fit)
-  #######################
-  ###Cox 
+
+  ######################
+  ###Cox###           ##
   ######################
@@
   coxlinSurv<-pyProcess(t(a$surv),times=times)
-  colnames(coxlinSurv)<-seq(0,1,length.out = ncol(coxlinSurv))
+  colnames(coxlinSurv)<-times#seq(0,1,length.out = ncol(coxlinSurv))
 
+   print('starting cindex')
   ######################
@@
   cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=length(riskList)+1)
-  et_train<-as.matrix(train)[,c(16,15)]
-  et_test<-as.matrix(fullTest[,c(16,15)])
+  et_train<-as.matrix(train)[,c(15,16)]
+  et_test<-as.matrix(fullTest[,c(15,16)])
   colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
@@
   saveRDS(cScore,'results/100oldcidx.rds')
+  print(pos)
   pos<-pos+1
@@
   ######################
-samp<-sample(seq(1,nrow(trainOG),by=1), floor(nrow(trainOG)*.80)) 
-validation<-trainOG[-samp,]
+#samp<-sample(seq(1,nrow(trainOG),by=1), floor(nrow(trainOG)*.80)) 
+#validation<-trainOG[-samp,]
 
-train<-trainOG[samp,]
+#train<-trainOG[samp,]
 
-  mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
-                                data = train,
-                                time = "time",
-                                event="status",
-                                ratio = 100
-  )
+  #mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
+  #                              data = train,
+  #                              time = "time",
+  #                              event="status",
+  #                              ratio = 100
+  #)
 
@@
   covars_output<-covars_input%>% layer_batch_normalization()%>%
-    layer_dense(units=hp$Int('units',min_value=5,max_value=25,step=2),
-                use_bias = T,activation = 'linear')%>%
-    layer_dropout(hp$Choice('rate',
+    layer_dense(units=hp$Int('units1',min_value=50,max_value=500,step=25),
+                use_bias = T,activation = 'tanh')%>%
+    layer_dropout(hp$Choice('rate1',
                 values=c(0.4,0.3,0.2,0.1,0.05)))%>%
-    layer_dense(units=hp$Int('units',min_value=5,max_value=25,step=2),
-                use_bias = T,activation = 'linear')%>%
-    layer_dropout(hp$Choice('rate',
+    layer_dense(units=hp$Int('units2',min_value=50,max_value=500,step=25),
+                use_bias = T,activation = 'tanh')%>%
+    layer_dropout(hp$Choice('rate2',
                 values=c(0.4,0.3,0.2,0.1,0.05)))%>%
@@
                   eventVar= "status",
-                  optimizer= tf$keras$optimizers$Adam(
-      hp$Choice('learning_rate',
-                values=c(1e-2,5e-3, 1e-3,5e-4, 1e-4)))
+                  optimizer=optimizer_adam(learning_rate = 0.0001)#,decay=10^-7)
   )
@@
     max_trials = 20,
-    executions_per_trial = 3,
+    executions_per_trial = 5,
     directory = 'tuning',
@@
   ytr<-as.matrix(mod_cb_glm$data[,c(ncol(mod_cb_glm$data)-1)])
-  xv<-list(as.matrix(validation[,-c(ncol(validation))]),as.matrix(data.frame(offset=rep(0,nrow(validation)))))
-  yv<-as.matrix(validation[,c(ncol(validation)),drop=F])
+  xv<-valcbtensor[[1]]#list(as.matrix(validation[,-c(ncol(validation))]),as.matrix(data.frame(offset=rep(0,nrow(validation)))))
+  yv<-valcbtensor[[2]]#as.matrix(validation[,c(ncol(validation)),drop=F])
   tuner %>% fit_tuner(x=list(xtr,as.matrix(mod_cb_glm$offset)),y=ytr,
-                    epochs = 50,batch_size=bsize+1000, 
-                    validation_data = list(x=xv,y=yv),verbose=0 )
+                    epochs = 55,batch_size=bsize, 
+                    validation_data = list(xv,yv),verbose=0 )
 result = kerastuneR::plot_tuner(tuner)

### 2022-12-15 analyses/plotterScript.Rmd
re-analyzed

@@
                         trueName=c("KM","Cox",'CBLR','DeepSurv',
-                                'DeepHit','PMNN_Spline','CBNN','Optimal',
+                                'DeepHit','CBNN_Spline','CBNN','Optimal',
                                 "DSM"))
@@
     }   
- metricSummary<- metricSummary[-which(metricSummary$model=="PMNN_Spline"),]
+ metricSummary#<- metricSummary[-which(metricSummary$model=="PMNN_Spline"),]
   #mainTitle<-paste(method,' : 95% P.I. over ', iterations, ' iterations',sep='')
@@
 
+```{r plotoldmort,fig.height=6,fig.width=6}
+
+BrierIPAList<-readRDS('results/100oldBrier.rds')
+
+library(eha)
+
+data("oldmort")
+
+preData<-data.frame(sex=oldmort$sex,
+                 civ=oldmort$civ,
+                 ses=oldmort$ses.50,
+                 birthplace=oldmort$birthplace,
+                 ageAtDeath=oldmort$exit,
+                 imrAtBirth=oldmort$imr.birth,
+                 region=oldmort$region,
+                 birthdate=oldmort$birthdate,
+                 time=oldmort$exit-oldmort$enter,
+                 status=as.numeric(oldmort$event))
+
+#BrierIPAList[[1]]$times<-BrierIPAList[[1]]$times*(max(preData$time, na.rm = T)-min(preData$time, na.rm = T)) + min(preData$time, na.rm = T)
+ipaSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
+brierSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
+ibsSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
+ibsTable<-as.data.frame(matrix(data=NA,nrow=100,ncol=length(unique(BrierIPAList[[1]]$model))))
+colnames(ibsTable)<-unique(BrierIPAList[[1]]$model)
+count=1
+for(sim in BrierIPAList){
+  ipaSims<-cbind(ipaSims, sim$IPA)
+  brierSims<-cbind(brierSims,sim$Brier)
+  ibsSims<-cbind(ibsSims,sim$IBS)
+  for(mod in colnames(ibsTable)){
+
+    ibsTable[count,which(mod ==colnames(ibsTable))]<-sim$IBS[sim$model==mod & sim$time==max(sim$time)]
+
+  }
+  count<-count+1
+}
+ibsTable<-na.omit(ibsTable)
+ibsOverall<-data.frame(mean=apply(ibsTable,2,mean),sd = apply(ibsTable,2,sd))
+ibsOverall
+ipaSummary<-statCalcSim(ipaSims)
+brierSummary<-statCalcSim(brierSims)
+ibsSummary<-statCalcSim(ibsSims)
+ipa<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
+brier<-plotPerformance(brierSummary,method="BS",iterations=iteration )
+ibs<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )
+
+#[which(ipaSummary$model %in% c("PMNN_Spline","Cox_Lin")),]
+pdf(file="figures/oldResult.pdf")
+ipa
+dev.off()
+
+colnames(preData)[ncol(preData)-1]<-'time'
+colnames(preData)[ncol(preData)]<-'status'
+data<-preData
+#data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
+covs<-colnames(data)
+
+ptPlot<-plot(popTime(data,time="time",event="status"),percentile_number=0.5)#+  theme(axis.text=element_text(size=8))
+#end_time-start_time
+
+pt <- plot_grid(ptPlot,
+          ipa+ theme(legend.position="none"),
+          brier+ theme(legend.position="none"),
+          ibs+ theme(legend.position="none"),
+          labels = c('A', 'B','C','D'))
+#https://wilkelab.org/cowplot/articles/shared_legends.html
+legend <- get_legend(
+  # create some space to the left of the legend
+  ipa  +
+    guides(color = guide_legend(nrow = 1)) +
+    theme(legend.position = "bottom")+
+   theme(legend.text=element_text(size=8))
+)
+
+pdf(file="figures/oldResult.pdf")
+plot_grid(pt,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))#, legend
+dev.off()
+
+#myVars<-c('MK167','EGFR','PGR','ERBB2','Hormone treatment','radiotherapy','Chemotherapy','ER_positive','age at diagnosis','time','status')
+#catVars<-c('Hormone treatment','radiotherapy','Chemotherapy','ER_positive','status')
+
+#tab2 <- as.data.frame(print(CreateTableOne(vars = myVars, data = data, factorVars = catVars)))
+#saveRDS(tab2,"figures/t1Metabric.rds")
+
+#pdf("figures/t1metabric.pdf", height = 50*nrow(tab2), width = 200*ncol(tab2))
+#grid.table(tab2)
+#dev.off()
+#rm(pt)
+
+pdf(file="figures/oldResult.pdf")  
+ipa
+dev.off()
+ipaO<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
+
+brierO<-plotPerformance(brierSummary,method="BS",iterations=iteration )
+ibsO<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )
+
+ipa
+ibs
+brier
+
+cidxList<-readRDS('results/100oldcidx.rds')
+
+temp<-melt(setDT(as.data.frame(cidxList[[1]])), id.vars = c("times"), variable.name = "model")
+cidxSims=data.frame(model=temp$model,times=temp$times)
+
+count=1
+for(sim in cidxList){
+  cidxSims<-cbind(cidxSims, melt(setDT(as.data.frame(sim)), id.vars = c("times"), variable.name = "model")$value)
+
+  count<-count+1
+}
+
+cidxSummary<-statCalcSim(cidxSims)
+cidxO<-plotPerformance(cidxSummary,method='C-Index',iterations=iteration  )
+cidxO
+```
+
+```{r plotgbsg,fig.height=6,fig.width=6}
+
+BrierIPAList<-readRDS('results/100gbsgBrier.rds')
+
+gbsg<-survival::gbsg
+
+colnames(gbsg)[10]<-"time"
+preData<-gbsg[,-1]
+#preData$time<-preData$time+runif(nrow(preData),0,1)/100000000
+
+data<-as.data.frame((model.matrix(~.,preData)[,-1]))
+
+data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
+covs<-colnames(data)
+
+#BrierIPAList[[1]]$times<-BrierIPAList[[1]]$times*(max(preData$time, na.rm = T)-min(preData$time, na.rm = T)) + min(preData$time, na.rm = T)
+ipaSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
+brierSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
+ibsSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
+ibsTable<-as.data.frame(matrix(data=NA,nrow=100,ncol=length(unique(BrierIPAList[[1]]$model))))
+colnames(ibsTable)<-unique(BrierIPAList[[1]]$model)
+count=1
+for(sim in BrierIPAList){
+  ipaSims<-cbind(ipaSims, sim$IPA)
+  brierSims<-cbind(brierSims,sim$Brier)
+  ibsSims<-cbind(ibsSims,sim$IBS)
+  for(mod in colnames(ibsTable)){
+
+    ibsTable[count,which(mod ==colnames(ibsTable))]<-sim$IBS[sim$model==mod & sim$time==max(sim$time)]
+
+  }
+  count<-count+1
+}
+ibsTable<-na.omit(ibsTable)
+ibsOverall<-data.frame(mean=apply(ibsTable,2,mean),sd = apply(ibsTable,2,sd))
+ibsOverall
+ipaSummary<-statCalcSim(ipaSims)
+brierSummary<-statCalcSim(brierSims)
+ibsSummary<-statCalcSim(ibsSims)
+ipa<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
+brier<-plotPerformance(brierSummary,method="BS",iterations=iteration )
+ibs<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )
+
+#[which(ipaSummary$model %in% c("PMNN_Spline","Cox_Lin")),]
+pdf(file="figures/gbsgResult.pdf")
+ipa
+dev.off()
+
+colnames(preData)[ncol(preData)-1]<-'time'
+colnames(preData)[ncol(preData)]<-'status'
+data<-preData
+#data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
+covs<-colnames(data)
+
+ptPlot<-plot(popTime(data,time="time",event="status"),percentile_number=0.5)#+  theme(axis.text=element_text(size=8))
+#end_time-start_time
+
+pt <- plot_grid(ptPlot,
+          ipa+ theme(legend.position="none"),
+          brier+ theme(legend.position="none"),
+          ibs+ theme(legend.position="none"),
+          labels = c('A', 'B','C','D'))
+#https://wilkelab.org/cowplot/articles/shared_legends.html
+legend <- get_legend(
+  # create some space to the left of the legend
+  ipa  +
+    guides(color = guide_legend(nrow = 1)) +
+    theme(legend.position = "bottom")+
+   theme(legend.text=element_text(size=8))
+)
+
+pdf(file="figures/gbsgResult.pdf")
+plot_grid(pt,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))#, legend
+dev.off()
+
+#myVars<-c('MK167','EGFR','PGR','ERBB2','Hormone treatment','radiotherapy','Chemotherapy','ER_positive','age at diagnosis','time','status')
+#catVars<-c('Hormone treatment','radiotherapy','Chemotherapy','ER_positive','status')
+
+#tab2 <- as.data.frame(print(CreateTableOne(vars = myVars, data = data, factorVars = catVars)))
+#saveRDS(tab2,"figures/t1Metabric.rds")
+
+#pdf("figures/t1metabric.pdf", height = 50*nrow(tab2), width = 200*ncol(tab2))
+#grid.table(tab2)
+#dev.off()
+#rm(pt)
+
+pdf(file="figures/gbsgResult.pdf")  
+ipa
+dev.off()
+ipaO<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
+
+brierO<-plotPerformance(brierSummary,method="BS",iterations=iteration )
+ibsO<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )
+
+ipa
+ibs
+brier
+
+cidxList<-readRDS('results/100gbsgcidx.rds')
+
+temp<-melt(setDT(as.data.frame(cidxList[[1]])), id.vars = c("times"), variable.name = "model")
+cidxSims=data.frame(model=temp$model,times=temp$times)
+
+count=1
+for(sim in cidxList){
+  cidxSims<-cbind(cidxSims, melt(setDT(as.data.frame(sim)), id.vars = c("times"), variable.name = "model")$value)
+
+  count<-count+1
+}
+
+cidxSummary<-statCalcSim(cidxSims)
+cidxO<-plotPerformance(cidxSummary,method='C-Index',iterations=iteration  )
+cidxO
+```
+
 ```{r megaplot,eval=T,fig.height=12,fig.width=12}

### 2022-12-15 analyses/simpleSim.Rmd
re-analyzed

--- 2022-12-15 analyses/metabricStudy.Rmd
@@
 ---
-title: "Metabric"
+title: "iterSim"
 output: html_document
@@
 
-```{r,eval=T}
+```{r}
 
@@
 
-```{python,eval=T}
+```{python}
 
@@
 os.environ["CUDA_VISIBLE_DEVICES"] = ""
-#os.environ["OM_NUM_THREADS"]="2"
 
@@
 source("src/functions.R")
-eval_metabric<-T
-pycox<-reticulate::import("pycox")
-source_python('src/getMetabricData.py')
+eval_support<-T
+source_python('src/getSUP.py')
 source_python("src/cIndex.py")
@@
 epo=as.numeric(params$epo)
-patience = as.numeric(params$patience)
-iteration=as.numeric(params$iteration)
+patience <- as.numeric(params$patience)
+iteration<-as.numeric(params$iteration)
 layer1<-as.numeric(params$layer1)
@@
 
+easysim<-T
+numSamp<-2000
 ```
 
-```{r modelsMETABRIC,eval=eval_metabric}
-
+```{r easySimpletests,eval=easysim}
 start_time <- Sys.time()
 
-#pycox<-reticulate::import("pycox")
-#numpy<-reticulate::import("numpy")
-#a<-pycox$datasets$metabric$read_df
-#b<-numpy$asarray(a)
+  #############################
+  ###Data generation
+  #############################
 
-############################
-###Data clean and prep
-############################
+  cov<-data.frame(id=1:numSamp,
+                  z1=c(rep(1,numSamp/2),rep(0,numSamp/2)),
+                  z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
+                  z3=c(rnorm(numSamp/2,-3,0.5),rnorm(numSamp/2,8,0.5))
+  )
 
-metabric<-getMetabric()
-colnames(metabric)[ncol(metabric)-1]<-'time'
-colnames(metabric)[ncol(metabric)]<-'status'
-data<-as.data.frame(model.matrix(~.,metabric)[,-1])
-data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
+  dat <- simsurv(dist = "exponential",
+                 lambdas =.05,
+                 betas = c(z1 = 1,z2=1,z3=1),
+                 x = cov,
+                 tdefunction = "log",
+                 maxt = 1
+  )
+  colnames(dat)<-c("id","time","status")
+  samp<-sample(seq(1,nrow(dat),by=1), floor(nrow(dat)*(0.1))) #random censoring
+  dat$status[samp]<-0
+  # Merge the simulated event times onto covariate data frame
+  data <- merge(cov, dat)
   covs<-colnames(data)
+  data<-data[,-1]
 
@@
 
+numSamp<-nrow(trainOG)
+
+BrierIPAList<-list()
+cScore<-list()
+
+pos<-1
+
+while (pos <= iteration) {
+
+  #############################
+  ###Data generation
+  #############################
+
+  cov<-data.frame(id=1:numSamp,
+                  z1=c(rep(1,numSamp/2),rep(0,numSamp/2)),
+                  z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
+                  z3=c(rnorm(numSamp/2,-3,0.5),rnorm(numSamp/2,8,0.5))
+  )
+
+  dat <- simsurv(dist = "exponential",
+                 lambdas =.05,
+                 betas = c(z1 = 0.5,z2=0.5,z3=0.5),
+                 x = cov,
+                 tdefunction = "log",
+                 maxt = 1
+  )
+  colnames(dat)<-c("id","time","status")
+  samp<-sample(seq(1,nrow(dat),by=1), floor(nrow(dat)*(0.1))) #random censoring
+  dat$status[samp]<-0
+  # Merge the simulated event times onto covariate data frame
+  data <- merge(cov, dat)
+  covs<-colnames(data)
+  data<-data[,-1]
+  train<-data
+  #test<-data[-samp,-c(ncol(data))]
+  #fullTest<-data[-samp,,drop=F]
   times<-seq(from=min(test$time)+0.000001,
@@
   )
-
-############################
-###Prep Loop
-############################
-BrierIPAList<-list()
-cScore<-list()
-pos<-1
-
-while (pos <= iteration) {
-
-  ############################
-  ###Bootstrap samples
-  ############################  
-  samp<-sample(seq(1,nrow(trainOG),by=1) , nrow(trainOG),replace = T) 
-  train<-trainOG[samp,]
-  sds<-sapply(train, function(x) (sd(x)))
-means<-sapply(train, function(x) (mean(x)))
-test<-normalizer(res$test,means,sds)[,-c(ncol(data))]
-fullTest<-normalizer(res$test,means,sds)
-val<-normalizer(res$validate,means,sds)
-
-train<-normalizer(train,means,sds)
 
@@
   }
-  dsmSurv<-as.data.frame(dsmSurv)
   colnames(dsmSurv) <-times
@@
   rm(dsmfitter)
+
+  ##########################
+  ###casebase optimal model
+  ##########################
+  final_mod_cb_glm <- fitSmoothHazard(status~z1+z2+z3,#z1:z2:z3+z1+z2+z3,+z1:z3+z1:z2:z3
+                                      data = train,
+                                      time = "time",
+                                      event="status",
+                                      ratio = 100
+  )
+
+  summary(final_mod_cb_glm)
+  tglmAbsRisk<-absoluteRisk(final_mod_cb_glm,time=times,newdata=test,type="CI")
+  tglmAbsRisk<-as.data.frame(tglmAbsRisk)
+  rownames(tglmAbsRisk)<-tglmAbsRisk$time
+  trueGlmProper<- t(tglmAbsRisk[-1,-1])
+
+  class(trueGlmProper)<-c("tunnel",class(trueGlmProper))
 
@@
 
-  ############################
+  ######################
   ###coxph
-  ############################  
-  cox<-coxph(Surv(time, status) ~ ., data = train,x=T)
+  ######################
+  cox<-coxph(Surv(time, status) ~ z1+z2+z3, data = train,x=T)
 
@@
   class(annProperpoly)<-c("tunnel",class(annProperpoly))
-  ############################
-  ###Brier Score
-  ############################  
-  brierFinalResults <- Score(list("Cox_Lin" = cox,'CB_Logi'=glmProper,
-                                  'DeepSurv'=cnnCleaned,'DeepHit'=deephitCleaned,
-                                  'PMNN_Spline'=annProper,'PMNN_Poly'=annProperpoly,
-                                  'dsm'=dsmSurvProper),
+
+  class(annProperpoly)<-c("tunnel",class(annProperpoly)) 
+  brierFinalResults <- Score(list("Cox_Lin" = cox,
+                                  'CB_Logi'=glmProper,
+                                  'DeepSurv'=cnnCleaned,
+                                  'DeepHit'=deephitCleaned,
+                                  'PMNN_Spline'=annProper,
+                                  'Optimal'=trueGlmProper,
+                                  'PMNN_Poly'=annProperpoly,
+                                  'dsm'=dsmSurvProper) ,#'dsm'=dsmCleaned,
                                   data =fullTest, 
@@
                  annProper,
+                 trueGlmProper,
                  annProperpoly,
@@
   cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=length(riskList)+1)
-  et_train<-as.matrix(train)[,c(11,10)]
-  et_test<-as.matrix(fullTest[,c(11,10)])
+  et_train<-as.matrix(train)[,c(5,4)]
+  et_test<-as.matrix(fullTest[,c(5,4)])
   colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
-                          'PMNN_Spline','PMNN_Poly','dsm')
+                          'PMNN_Spline','Optimal','PMNN_Poly','dsm')#,'dsm')
   cScoreTemp[,1]<- tempTims
@@
 
-  saveRDS(BrierIPAList,'results/100MetaBrier.rds')
-  saveRDS(cScore,'results/100Metacidx.rds')
-  pos=pos+1
+  saveRDS(BrierIPAList,'results/100easyBrier.rds')
+  saveRDS(cScore,'results/100easycidx.rds')
+  pos<-pos+1
 }
 end_time <- Sys.time()
-
-end_time-start_time
-
-ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=Brier,col=model))+
-  geom_line()
-ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
-  geom_line()
 
@@
 ```
-
-```{r}
-
-sessionInfo()
-
-```

### 2022-12-15 analyses/supportStudy.Rmd
re-analyzed

@@
 bsize=512
-min_delta = 10^-7
+min_delta = 0#10^-7
 epo=as.numeric(params$epo)
@@
 #split training/testing
-samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.80)) 
-trainOG<-data[samp,]
-test<-data[-samp,-c(ncol(data))]
-fullTest<-data[-samp,,drop=F]
+
+ spec = c(train = .7, test = .1, validate = .2)
+
+ g = sample(cut(
+   seq(nrow(data)), 
+   nrow(data)*cumsum(c(0,spec)),
+   labels = names(spec)
+ ))
+
+ res = split(data, g)
+
+trainOG<-res$train
+
+test<-res$test[,-c(ncol(data))]
+
+fullTest<-res$test
+
+val<-res$validate
+
+valcb<-sampleCaseBase(data = val,time="time",event="status",ratio=100)
+valcb$offset<-0
+valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
+
+  times<-seq(from=min(test$time)+0.000001,
+             to=max(test$time)-0.000001,
+             length.out = 100
+  )
 
@@
   train<-trainOG[samp,]
-  times<-seq(from=min(test$time),
-             to=max(test$time),
-             length.out = 200
-  )
-  times<-times[-c(1,length(times))]
+  sds<-sapply(train, function(x) (sd(x)))
+means<-sapply(train, function(x) (mean(x)))
+test<-normalizer(res$test,means,sds)[,-c(ncol(data))]
+fullTest<-normalizer(res$test,means,sds)
+val<-normalizer(res$validate,means,sds)
 
-  ######################
-  ###casebase + splines
-  ######################
+train<-normalizer(train,means,sds)
+
+  #########################
+  ###casebase+splines
+ ##########################
     mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
@@
 
-  ######################
-  ###DeepSurv
-  ######################
-  coxnn<-deepsurv(Surv(time, status) ~ .,data = train, frac = 0.2, activation = "relu",
-                  num_nodes = c(layer1,layer2,layer3,layer4), dropout = drpt,
-                  early_stopping = TRUE, epochs = epo,
-                  batch_size = bsize,device = "cpu",best_weights=T,
-                  patience = patience,min_delta=min_delta,batch_norm = F,
-                  learning_rate = 0.001,lr_decay=10^-7)
-  coxnnsurv<- predict(coxnn,newdata = test,type="survival")
-  colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
+  # 
+  # 
+  # ############################
+  # ###DeepSurv
+  # ############################
+
+  source_python('src/dsurv.py')
+  #need to break ties
+  tiebreakTrain<-train
+
+  #while(length(unique(tiebreakTrain$time))/length(tiebreakTrain$time)!=1 ){
+  #tiebreakTrain<-train
+  #tiebreakTrain$time<-tiebreakTrain$time+runif(nrow(tiebreakTrain),0,1/10000000000)
+  #}
+
+  coxnnsurv=fitDeepSurv(tiebreakTrain,fullTest,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4)
+  coxnnsurv<-t(coxnnsurv)
+  #colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
   cnnCleaned<-pyProcess(coxnnsurv,times=times)
-  colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
+  colnames(cnnCleaned)<-times#seq(0,1,length.out = ncol(cnnCleaned))
+  py_run_string("del fitDeepSurv")
+  rm(fitDeepSurv)
 
-  ######################
+  # coxnn<-deepsurv(Surv(time, status) ~ .,data = train, frac = 0.2, activation = "relu",
+  #                 num_nodes = c(layer1,layer2,layer3,layer4), dropout = drpt,
+  #                 early_stopping = TRUE, epochs = epo,
+  #                 batch_size = bsize,device = "cpu",best_weights=T,
+  #                 patience = patience,min_delta=min_delta,batch_norm = F,
+  #                 learning_rate = 0.001,lr_decay=10^-7)
+  # 
+  # coxnnsurv<- predict(coxnn,newdata = test,type="survival")
+  # colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
+  # cnnCleaned<-pyProcess(coxnnsurv,times=times)
+  # colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
+  ############################
   ###Deephit
-  ######################
+  ############################  
     source_python('src/deephitter.py')
-  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,patience=patience,min_delta=min_delta,drpt=drpt)
+  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4)
   hitnnSurv<-t(hitnnSurv)
-  colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
+  #colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
   deephitCleaned<-pyProcess(hitnnSurv,times=times)
@@
   rm(fitDeephit)
-
-  ######################
-  ###PMNN splines
-  ######################
+  # source_python('src/deephitter.py')
+  # hitnnSurv=fitDeephit(train,fullTest,bsize,epo,patience=patience,min_delta=min_delta,drpt=drpt)
+  # hitnnSurv<-t(hitnnSurv)
+  # colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
+  # deephitCleaned<-pyProcess(hitnnSurv,times=times)
+  # colnames(deephitCleaned)<-seq(0,1,length.out = ncol(deephitCleaned))
+  # py_run_string("del fitDeephit")
+  # rm(fitDeephit)
+  # 
+  ############################
+  ###PMNNsplines
+  ############################  
   covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2+3),
@@
 
-  ######################
+  ############################
   ###coxph
-  ######################
+  ############################  
   cox<-coxph(Surv(time, status) ~ ., data = train,x=T)
@@
   class(annProperpoly)<-c("tunnel",class(annProperpoly)) 
+  ############################
+  ###Brier Score
+  ############################
 
-  ######################
-  ###IPA metric
-  ######################
-  brierFinalResults <- Score(list("Cox_Lin" = cox,'CB_Logi'=glmProper,
-                                  'DeepSurv'=cnnCleaned,'DeepHit'=deephitCleaned,
-                                  'PMNN_Spline'=annProper,'PMNN_Poly'=annProperpoly),
+  brierFinalResults <- Score(list("Cox_Lin" = cox,
+                                  'CB_Logi'=glmProper,
+                                  'DeepSurv'=cnnCleaned,
+                                  'DeepHit'=deephitCleaned,
+                                  'PMNN_Spline'=annProper,
+                                  'PMNN_Poly'=annProperpoly),
                              data =fullTest, 
@@
   BrierIPAList[[pos]]<-brierFinalResults$Brier$score
-
-  #######################
+    ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
+  geom_line()+coord_cartesian(ylim=c(-0.1,0.2))
+  ######################
   ###Cox
@@
   colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
-                          'PMNN_Spline','PMNN_Poly')
+                          'PMNN_Spline','PMNN_Poly')#,'dsm')
   cScoreTemp[,1]<- tempTims
@@
 
+  ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
+  geom_line()+coord_cartesian(ylim=c(-0.1,0.2))
+
+```
+
+```{r}
+
 sessionInfo()

### 2022-12-15 analyses/timevaryingOldMortality.Rmd
re-analyzed

--- 2022-12-15 analyses/oldmortality.Rmd
@@
   iteration: 1
-  layer1: 50
+  layer1: 200
   layer2: 20
@@
 
-preData<-data.frame(sex=oldmort$sex,
-                 civ=oldmort$civ,
-                 ses=oldmort$ses.50,
-                 birthplace=oldmort$birthplace,
+omor=as.data.frame(matrix(NA,nrow=length(unique(oldmort$id)),ncol=ncol(oldmort)))
+colnames(omor)<-colnames(oldmort)
+unique(oldmort$id)[1]
+j<-1
+for(i in 1:length(unique(oldmort$id))){
+  temp<-which(oldmort$id %in% unique(oldmort$id)[i])
+  if(length(temp)==1){
+  omor[j,]<-oldmort[temp,]
+  j<-j+1
+  }else if(length(temp)>=1){
+    tempData<-oldmort[which(oldmort$id %in% unique(oldmort$id)[i]),]
+    omor[j,]<-tempData[which.min(tempData$enter),]
+    #omor[j,]$exit<-max(tempData$exit)
+    j<-j+1
+  }
+
+}
+
+oldmort<-omor
+preData<-data.frame(sex=as.factor(oldmort$sex),
+                 civ=as.factor(oldmort$civ),
+                 ses=as.factor(oldmort$ses.50),
+                 birthplace=as.factor(oldmort$birthplace),
                  ageAtEntry=oldmort$enter,
                  imrAtBirth=oldmort$imr.birth,
-                 region=oldmort$region,
+                 region=as.factor(oldmort$region),
                  birthdate=oldmort$birthdate,
@@
 
-preData$time<-preData$time+runif(nrow(preData),0,1)/100000000
+#preData$time<-preData$time+runif(nrow(preData),0,1)/100000000
 
@@
 
- spec = c(train = .75, test = .10, validate = .15)
+ spec = c(train = .70, test = .10, validate = .20)
 
@@
 
-trainOG<-res$train
+trainOG<-
 
@@
 val<-res$validate
+
+sds<-sapply(trainOG, function(x) (sd(x)))
+means<-sapply(trainOG, function(x) (mean(x)))
+
+normalizer<-function(data,means,sds){
+  normalized<-data
+  for (i in 1:ncol(data)){
+    normalized[,i]<-(data[,i]-means[i])/sds[i]
+  }
+  normalized$status<-data$status
+  normalized$time<-data$time
+  return(normalized)
+}
+
+test<-normalizer(res$test,means,sds)[,-c(ncol(data))]
+fullTest<-normalizer(res$test,means,sds)
+val<-normalizer(res$validate,means,sds)
+
+trainOG<-normalizer(res$train,means,sds)
 
@@
              to=max(test$time),
-             length.out = 200
+             length.out = 100
   )
@@
 
-  #while(length(unique(tiebreakTrain$time))/length(tiebreakTrain$time)!=1 ){
-  #tiebreakTrain<-train
-  #tiebreakTrain$time<-tiebreakTrain$time+runif(nrow(tiebreakTrain),0,1/10000000000)
-  #}
+  while(length(unique(tiebreakTrain$time))/length(tiebreakTrain$time)!=1 ){
+  tiebreakTrain<-train
+  tiebreakTrain$time<-tiebreakTrain$time+runif(nrow(tiebreakTrain),0,1/10000000)
+  }
 
@@
   )
-  covars_output<-covars_input%>%
+  covars_output<-covars_input%>%layer_batch_normalization()%>%
     layer_dense(units=layer1,use_bias = T)%>%
-    layer_batch_normalization()%>%
+   # layer_batch_normalization()%>%
     activation_tanh()%>%
@@
     layer_dense(units=layer2,use_bias = T)%>%
-    layer_batch_normalization()%>%
+    #layer_batch_normalization()%>%
     activation_tanh()%>%
@@
 
-  covars_output<-covars_input%>%
-
-    layer_dense(units=20,use_bias = T)%>%
-     layer_batch_normalization()%>%
-    activation_relu()%>%
+  covars_output<-covars_input%>% layer_batch_normalization()%>%
+    layer_dense(units=layer1,use_bias = T)%>%
+   # layer_batch_normalization()%>%
+    activation_tanh()%>%
     layer_dropout(drpt)%>%
 
-    layer_dense(units=10,use_bias = T)%>%
+    layer_dense(units=layer2,use_bias = T)%>%
+    #layer_batch_normalization()%>%
+    activation_tanh()%>%
+    layer_dropout(drpt)%>%
 
-    activation_relu()%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=1,use_bias = T)
+    layer_dense(units=1,use_bias = T)#%>%activation_tanh()
 
@@
                   eventVar= "status",
-                  optimizer=optimizer_adam(learning_rate = 0.0001)#,decay=10^-7)
+                  optimizer=optimizer_adam(learning_rate = 0.0001)
   )

### 2022-12-15 analyses/SIMpaperResultscopy/CopyOfsrc/cIndex.py

@@ -1,0 +1,10 @@
+#We evaluate the performance of DSM in its discriminative ability (Time Dependent Concordance Index and Cumulative Dynamic AUC) as well as Brier Score.
+from sksurv.metrics import concordance_index_ipcw
+import numpy as np
+def cidx(et_train, et_test, out_risk): 
+  train = np.array([(et_train[i,0], et_train[i,1]) for i in range(np.shape(et_train)[0])],
+                   dtype = [('e', bool), ('t', float)])
+  test = np.array([(et_test[i,0], et_test[i,1]) for i in range(np.shape(et_test)[0])],
+                   dtype = [('e', bool), ('t', float)])
+  return concordance_index_ipcw(train, test, out_risk[:,0])[0]
+

### 2022-12-15 analyses/SIMpaperResultscopy/CopyOfsrc/deephitter.py

@@ -1,0 +1,69 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+# For preprocessing
+from sklearn.preprocessing import StandardScaler
+from sklearn_pandas import DataFrameMapper 
+
+import torch # For building the networks 
+import torchtuples as tt # Some useful functions
+
+from pycox.models import DeepHitSingle
+from pycox.evaluation import EvalSurv
+
+
+
+
+def fitDeephit(train,fullTest,bsize,epochs,ti='time',ev='status',patience=10,min_delta=10**-7,drpt=0.5):#,epoch,bsize,tPos,ePos):
+  n= np.shape(train)[0]
+  vl_size = int(n*0.20)
+  df_val = train.sample(frac=0.2)
+  df_train = train.drop(df_val.index)
+  df_test = fullTest
+  varNames = list(df_train.columns)
+  unwanted = {ti, ev}
+  varNames = [e for e in varNames if e not in unwanted]
+  leave = [(col, None) for col in varNames]
+
+  x_mapper = DataFrameMapper(leave)
+
+  x_train = x_mapper.fit_transform(df_train).astype('float32')
+  x_val = x_mapper.transform(df_val).astype('float32')
+  x_test = x_mapper.transform(df_test).astype('float32')
+ 
+
+
+
+  ########
+  #prepare labels
+  #######
+  num_durations = 10
+  labtrans = DeepHitSingle.label_transform(num_durations)
+  get_target = lambda df: (df[ti].values, df[ev].values)
+  y_train = labtrans.fit_transform(*get_target(df_train))
+  y_val = labtrans.transform(*get_target(df_val))
+  train = (x_train, y_train)
+  val = (x_val, y_val)
+  durations_test, events_test = get_target(df_test)
+  ########
+  #prepare model
+  ########
+  in_features = x_train.shape[1]
+  num_nodes = [50,50,25,25]
+  out_features = labtrans.out_features
+  batch_norm = False
+  dropout = drpt
+  
+  net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout)
+  model = DeepHitSingle(net, tt.optim.Adam, alpha=0.5, sigma=0.1, duration_index=labtrans.cuts)
+  lr_finder = model.lr_finder(x_train, y_train, batch_size=int(bsize), tolerance=2)
+  #_ = lr_finder.plot()
+  ############
+  #fit models
+  ###########
[…]

### 2022-12-15 analyses/SIMpaperResultscopy/CopyOfsrc/deepsurver.py

@@ -1,0 +1,74 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from sklearn.preprocessing import StandardScaler
+from sklearn_pandas import DataFrameMapper
+
+import torch
+import torchtuples as tt
+
+from pycox.datasets import metabric
+from pycox.models import CoxPH
+from pycox.evaluation import EvalSurv
+
+
+
+def fitDeepSurv(train,fullTest,bsize,epochs,valida,ti='time',ev='status',patience=10,min_delta=10**-7,drpt=0.5,lay1=200,lay2=20):#,epoch,bsize,tPos,ePos):
+  n= np.shape(train)[0]
+  vl_size = int(n*0.20)
+  df_val = valida#train.sample(frac=0.2)
+  df_train = train#.drop(df_val.index)
+  df_test = fullTest
+  varNames = list(df_train.columns)
+  unwanted = {ti, ev}
+  varNames = [e for e in varNames if e not in unwanted]
+  leave = [(col, None) for col in varNames]
+
+  x_mapper = DataFrameMapper(leave)
+  
+  x_train = x_mapper.fit_transform(df_train).astype('float32')
+  x_val = x_mapper.transform(df_val).astype('float32')
+  x_test = x_mapper.transform(df_test).astype('float32')
+  get_target = lambda df: (df[ti].values, df[ev].values)
+  y_train = get_target(df_train)
+  y_val = get_target(df_val)
+  #y_train[0]=y_train[0]+(np.random.uniform(0,1,len(y_train[0]))/10000000000000000000)
+  #y_val[0]=y_val[0]+(np.random.uniform(0,1,len(y_val[0]))/10000000000000000000)
+  durations_test, events_test = get_target(df_test)
+  val = x_val, y_val
+  
+  
+  in_features = x_train.shape[1]
+  num_nodes = [int(lay1),int(lay2)]
+  out_features = 1
+  batch_norm = True
+  dropout = drpt
+  
+  #net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout,activation="selu")
+  net = torch.nn.Sequential(
+     torch.nn.BatchNorm1d(in_features),
+     torch.nn.Linear(in_features,int(lay1)),
+     torch.nn.Tanh(),
+     torch.nn.Dropout(float(dropout)),
+
+     torch.nn.Linear(int(lay1), int(lay2)),
+     torch.nn.Tanh(),
[…]

### 2022-12-15 analyses/SIMpaperResultscopy/CopyOfsrc/functions.R

@@ -1,0 +1,464 @@
+
+statCalcSim<-function(metricSims){
+  toRemove <- which(apply(metricSims, 1, function(x){any(is.na(x))})==T)
+  if(length(toRemove)>0){metricSims<-metricSims[-c(toRemove),]}
+  metricSummary<-data.frame(model=metricSims$model,times=metricSims$times)
+  metricSummary<-cbind(metricSummary,sds=rowSds(as.matrix(metricSims[,-c(1,2)])),
+                       means=rowMeans(as.matrix(metricSims[,-c(1,2)])))
+  ci<-apply(as.matrix(metricSims[,-c(1,2)]), 1, function(x){mean(x)+c(-1.96,1.96)*sd(x)/sqrt(length(x))})
+  rownames(ci)<-c('cilower95','ciupper95')
+  pi<-apply(as.matrix(metricSims[,-c(1,2)]), 1, function(x){quantile(x,c(.025, .975))})
+  rownames(pi)<-c('pilower95','piupper95')
+  metricSummary<-cbind(metricSummary,t(ci),t(pi))
+  return(metricSummary)
+}
+
+plotPerformance<-function(metricSummary,xlab="Follow-up-time",method='Metric',iterations='NA'){
+  mainTitle<-paste(method,' : 95% P.I. over ', iterations, ' iterations',sep='')
+  
+  a<-ggplot(data=metricSummary, aes(x=times,y=means,col=model))+
+    geom_line() +
+    geom_ribbon(aes(ymin=pilower95,ymax=piupper95,fill=model),linetype=0,alpha=0.3) +
+    ggtitle(mainTitle) + 
+    xlab(xlab) +
+    ylab(method)
+  if(method=="IPA"){
+    a<-a+coord_cartesian()#ylim(-1,1))#
+    return(a)
+  }
+  return(a+coord_cartesian())
+}
+
+
+pyProcess<-function(predictions,times=times,apr="constant"){
+  interpolated <- apply(predictions, 1, function(x) approx(as.numeric(colnames(predictions)), y=x, method = apr, xout=times)$y)
+  class(interpolated)<-c("tunnel",class(interpolated))
+  interpolated<-1-interpolated
+  interpolated[is.na(interpolated) & is.na(interpolated)>50] <- 0
+  interpolated[is.na(interpolated) & is.na(interpolated)<50] <- 1
+  return(t(interpolated))
+}
+
+
+
+
[…]

### 2022-12-15 analyses/SIMpaperResultscopy/CopyOfsrc/oldmortDeephitter.py

@@ -1,0 +1,81 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+# For preprocessing
+from sklearn.preprocessing import StandardScaler
+from sklearn_pandas import DataFrameMapper 
+
+import torch # For building the networks 
+import torchtuples as tt # Some useful functions
+
+from pycox.models import DeepHitSingle
+from pycox.evaluation import EvalSurv
+
+
+
+
+def fitDeephit(train,fullTest,bsize,epochs,valida,ti='time',ev='status',patience=10,min_delta=10**-7,drpt=0.5,lay1=200,lay2=20):#,epoch,bsize,tPos,ePos):
+  n= np.shape(train)[0]
+  vl_size = int(n*0.20)
+  df_val = valida#train.sample(frac=0.2)
+  df_train = train#.drop(df_val.index)
+  df_test = fullTest
+  varNames = list(df_train.columns)
+  unwanted = {ti, ev}
+  varNames = [e for e in varNames if e not in unwanted]
+  leave = [(col, None) for col in varNames]
+
+  x_mapper = DataFrameMapper(leave)
+
+  x_train = x_mapper.fit_transform(df_train).astype('float32')
+  x_val = x_mapper.transform(df_val).astype('float32')
+  x_test = x_mapper.transform(df_test).astype('float32')
+ 
+
+
+
+  ########
+  #prepare labels
+  #######
+  num_durations = 10
+  labtrans = DeepHitSingle.label_transform(num_durations)
+  get_target = lambda df: (df[ti].values, df[ev].values)
+  y_train = labtrans.fit_transform(*get_target(df_train))
+  y_val = labtrans.transform(*get_target(df_val))
+  train = (x_train, y_train)
+  val = (x_val, y_val)
+  durations_test, events_test = get_target(df_test)
+  ########
+  #prepare model
+  ########
+  in_features = x_train.shape[1]
+  num_nodes =[int(lay1),int(lay2)]
+  out_features = labtrans.out_features
+  batch_norm = True
+  dropout = drpt
+  
+  #net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout,activation=torch.nn.SELU)
+  net = torch.nn.Sequential(
+     #torch.nn.BatchNorm1d(in_features),
+     torch.nn.Linear(in_features,int(lay1)),
+     torch.nn.Tanh(),
[…]

### 2022-12-15 analyses/dsm/dsm_torch.py

@@ -84,9 +84,9 @@ def create_representation(inputdim, layers, activation):
   prevdim = inputdim
 
   for hidden in layers:
-    modules.append(nn.Linear(prevdim, hidden, bias=False))
-    modules.append(nn.ReLU())
-    modules.append(nn.Dropout(0.5))
+    modules.append(nn.Linear(prevdim, hidden, bias=True))
+    modules.append(nn.Tanh())
+    modules.append(nn.Dropout(0.1))
     prevdim = hidden
 
   return nn.Sequential(*modules)

### 2022-12-15 analyses/runAll.sh

@@ -20,12 +20,15 @@ metabric='library(rmarkdown); rmarkdown::render("metabricStudy.Rmd",'"params=$pa
 support='library(rmarkdown); rmarkdown::render("supportStudy.Rmd",'"params=$params)"
 simple='library(rmarkdown); rmarkdown::render("simpleSim.Rmd",'"params=$params)"
 complex='library(rmarkdown); rmarkdown::render("complexSim.Rmd",'"params=$params)"
+oldmort='library(rmarkdown); rmarkdown::render("oldmortality.Rmd",'"params=$params)"
 (trap 'kill 0' SIGINT; \
 sleep 1 && eval `echo Rscript -e "'"{$simple}"'"` && bash ~/phoneNotification.sh simpleTest && sleep 9 && eval `echo Rscript -e "'"{$complex}"'"` && bash ~/phoneNotification.sh complexTest \ &
 sleep 6 && eval `echo Rscript -e "'"{$support}"'"` && bash ~/phoneNotification.sh supportTest \ &
-sleep 3 && eval `echo Rscript -e "'"{$metabric}"'"` && bash ~/phoneNotification.sh metabricTest 
+sleep 3 && eval `echo Rscript -e "'"{$metabric}"'"` && bash ~/phoneNotification.sh metabricTest \ &
+sleep 6 && eval `echo Rscript -e "'"{$oldmort}"'"` && bash ~/phoneNotification.sh oldmortTest 
 ) 
 
 
 
 
+Rscript -e '{library(rmarkdown); rmarkdown::render("oldmortality.Rmd",params=list(epo=2000,patience=10,iteration=2,layer1=200,layer2=20,layer3=25,layer4=25,drpt=0.1))}'

### 2022-12-15 analyses/src/deephitter.py

@@ -14,11 +14,11 @@ from pycox.evaluation import EvalSurv
 
 
 
-def fitDeephit(train,fullTest,bsize,epochs,ti='time',ev='status',patience=10,min_delta=10**-7,drpt=0.5):#,epoch,bsize,tPos,ePos):
+def fitDeephit(train,fullTest,bsize,epochs,valida,ti='time',ev='status',patience=10,min_delta=10**-7,drpt=0.5,lay1=50,lay2=50,lay3=25,lay4=25):#,epoch,bsize,tPos,ePos):
   n= np.shape(train)[0]
   vl_size = int(n*0.20)
-  df_val = train.sample(frac=0.2)
-  df_train = train.drop(df_val.index)
+  df_val = valida#train.sample(frac=0.2)
+  df_train = train#.drop(df_val.index)
   df_test = fullTest
   varNames = list(df_train.columns)
   unwanted = {ti, ev}
@@ -30,7 +30,6 @@ def fitDeephit(train,fullTest,bsize,epochs,ti='time',ev='status',patience=10,min
   x_train = x_mapper.fit_transform(df_train).astype('float32')
   x_val = x_mapper.transform(df_val).astype('float32')
   x_test = x_mapper.transform(df_test).astype('float32')
- 
 
 
 
@@ -49,21 +48,46 @@ def fitDeephit(train,fullTest,bsize,epochs,ti='time',ev='status',patience=10,min
   #prepare model
   ########
   in_features = x_train.shape[1]
-  num_nodes = [50,50,25,25]
+  #num_nodes = [50,50,25,25]
   out_features = labtrans.out_features
   batch_norm = False
   dropout = drpt
   
-  net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout)
+  #net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout)
+ 
+  net = torch.nn.Sequential(
+     #torch.nn.BatchNorm1d(in_features),
+     torch.nn.Linear(in_features,int(lay1)),
+     #torch.nn.ReLU(),
+     torch.nn.Tanh(),
+     torch.nn.Dropout(float(dropout)),
+     
+     torch.nn.Linear(int(lay1), int(lay2)),
+     #torch.nn.ReLU(),
+     torch.nn.Tanh(),
+     torch.nn.Dropout(float(dropout)),
+     
+     torch.nn.Linear(int(lay2), int(lay3)),
[…]

### 2022-12-15 analyses/src/deepsurver.py

@@ -12,11 +12,11 @@ from pycox.evaluation import EvalSurv
 
 
 
-def fitDeepSurv(train,fullTest,bsize,epochs,ti='time',ev='status',patience=10,min_delta=10**-7,drpt=0.5):#,epoch,bsize,tPos,ePos):
+def fitDeepSurv(train,fullTest,bsize,epochs,valida,ti='time',ev='status',patience=10,min_delta=10**-7,drpt=0.5,lay1=200,lay2=20):#,epoch,bsize,tPos,ePos):
   n= np.shape(train)[0]
   vl_size = int(n*0.20)
-  df_val = train.sample(frac=0.2)
-  df_train = train.drop(df_val.index)
+  df_val = valida#train.sample(frac=0.2)
+  df_train = train#.drop(df_val.index)
   df_test = fullTest
   varNames = list(df_train.columns)
   unwanted = {ti, ev}
@@ -24,39 +24,44 @@ def fitDeepSurv(train,fullTest,bsize,epochs,ti='time',ev='status',patience=10,mi
   leave = [(col, None) for col in varNames]
 
   x_mapper = DataFrameMapper(leave)
-
+  
   x_train = x_mapper.fit_transform(df_train).astype('float32')
   x_val = x_mapper.transform(df_val).astype('float32')
   x_test = x_mapper.transform(df_test).astype('float32')
   get_target = lambda df: (df[ti].values, df[ev].values)
   y_train = get_target(df_train)
   y_val = get_target(df_val)
+  #y_train[0]=y_train[0]+(np.random.uniform(0,1,len(y_train[0]))/10000000000000000000)
+  #y_val[0]=y_val[0]+(np.random.uniform(0,1,len(y_val[0]))/10000000000000000000)
   durations_test, events_test = get_target(df_test)
   val = x_val, y_val
   
   
   in_features = x_train.shape[1]
-  num_nodes = [200,20]
+  num_nodes = [int(lay1),int(lay2)]
   out_features = 1
-  #batch_norm = False
+  batch_norm = True
   dropout = drpt
   
-  #net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout,activation=torch.tanh)
+  #net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout,activation="selu")
   net = torch.nn.Sequential(
[…]

### 2022-12-15 analyses/src/dsmScript.py

@@ -7,7 +7,7 @@ from dsm import DeepSurvivalMachines
 ##dsm/dsm_torch.py, line 89 dropout(0.5) layer was added
 
 
-def dsmfitter(train,fullTest,bsize,epochs,tPos=3,ePos=4):
+def dsmfitter(train,fullTest,bsize,epochs,val,times,tPos=3,ePos=4,lay1=50,lay2=50,lay3=25,lay4=25):
   
   ###############################
   ###Prepare data
@@ -15,21 +15,21 @@ def dsmfitter(train,fullTest,bsize,epochs,tPos=3,ePos=4):
   tPos=int(tPos)
   ePos=int(ePos)
   train2=train.to_numpy()#@np.random.shuffle(train) 
-  np.random.shuffle(train2) 
+  #np.random.shuffle(train2) 
   n= np.shape(train2)[0]
-  vl_size = int(n*0.80)
+  #vl_size = int(n*0.80)
 
   
-  t_train = np.asarray( train2[:vl_size,tPos])
-  e_train = np.asarray( train2[:vl_size,ePos]).astype(int)
-  x_train = train2[:vl_size,:]
+  t_train = np.asarray( train2[:,tPos])
+  e_train = np.asarray( train2[:,ePos]).astype(int)
+  x_train = train2
   x_train = np.delete(x_train, [tPos,ePos], 1)
-  
-  t_val = np.asarray( train2[-vl_size:,tPos])
-  e_val = np.asarray( train2[-vl_size:,ePos]).astype(int)
-  x_val = train2[-vl_size:,:]
+  val=val.to_numpy()
+  t_val = np.asarray( val[:,tPos])
+  e_val = np.asarray( val[:,ePos]).astype(int)
+  x_val = val
   x_val = np.delete(x_val, [tPos,ePos], 1)
-
+  #print("hi")
   test=fullTest.to_numpy()
   t_test = np.asarray( test[:,tPos])
   e_test = np.asarray( test[:,ePos]).astype(int)
@@ -46,8 +46,8 @@ def dsmfitter(train,fullTest,bsize,epochs,tPos=3,ePos=4):
   #I fix the parameter grid
   param_grid = {'k' : [6],
                 'distribution' : ['Weibull'],
-                'learning_rate' : [ 0.001],
-                'layers' : [ [50,50,25,25] ]
+                'learning_rate' : [ 0.0001],
+                'layers' : [ [int(lay1),int(lay2),int(lay3),int(lay4)] ]
[…]

### 2022-12-15 analyses/src/dsurv.py

@@ -1,0 +1,85 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from sklearn.preprocessing import StandardScaler
+from sklearn_pandas import DataFrameMapper
+
+import torch
+import torchtuples as tt
+
+from pycox.datasets import metabric
+from pycox.models import CoxPH
+from pycox.evaluation import EvalSurv
+
+
+
+def fitDeepSurv(train,fullTest,bsize,epochs,valida,ti='time',ev='status',patience=10,min_delta=10**-7,drpt=0.5,lay1=50,lay2=50,lay3=25,lay4=25):#,epoch,bsize,tPos,ePos):
+  n= np.shape(train)[0]
+  vl_size = int(n*0.20)
+  df_val = valida#train.sample(frac=0.2)
+  df_train = train#.drop(df_val.index)
+  df_test = fullTest
+  varNames = list(df_train.columns)
+  unwanted = {ti, ev}
+  varNames = [e for e in varNames if e not in unwanted]
+  leave = [(col, None) for col in varNames]
+
+  x_mapper = DataFrameMapper(leave)
+  
+  x_train = x_mapper.fit_transform(df_train).astype('float32')
+  x_val = x_mapper.transform(df_val).astype('float32')
+  x_test = x_mapper.transform(df_test).astype('float32')
+  get_target = lambda df: (df[ti].values, df[ev].values)
+  y_train = get_target(df_train)
+  y_val = get_target(df_val)
+  #y_train[0]=y_train[0]+(np.random.uniform(0,1,len(y_train[0]))/10000000000000000000)
+  #y_val[0]=y_val[0]+(np.random.uniform(0,1,len(y_val[0]))/10000000000000000000)
+  durations_test, events_test = get_target(df_test)
+  val = x_val, y_val
+  
+  
+  in_features = x_train.shape[1]
+  num_nodes = [int(lay1),int(lay2)]
+  out_features = 1
+  batch_norm = True
+  dropout = drpt
+  
+  #net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout,activation="selu")
+  net = torch.nn.Sequential(
+     #torch.nn.BatchNorm1d(in_features),
+     torch.nn.Linear(in_features,int(lay1)),
+     #torch.nn.ReLU(),
+     torch.nn.Tanh(),
[…]

### 2022-12-15 analyses/src/functions.R

@@ -30,8 +30,8 @@ plotPerformance<-function(metricSummary,xlab="Follow-up-time",method='Metric',it
 }
 
 
-pyProcess<-function(predictions,times=times){
-  interpolated <- apply(predictions, 1, function(x) approx(as.numeric(colnames(predictions)), y=x, method = "linear", xout=times)$y)
+pyProcess<-function(predictions,times=times,apr="constant"){
+  interpolated <- apply(predictions, 1, function(x) approx(as.numeric(colnames(predictions)), y=x, method = apr, xout=times)$y)
   class(interpolated)<-c("tunnel",class(interpolated))
   interpolated<-1-interpolated
   interpolated[is.na(interpolated) & is.na(interpolated)>50] <- 0
@@ -61,6 +61,7 @@ calcCidx<-function(et_train,et_test,risk,times){
 cIndexSummary<-function(et_train, et_test,riskList,cScore){
   
   for (i in 1:length(riskList)){
+    #print(i)
     cScore[,i+1]<-calcCidx(et_train=et_train,et_test=et_test,risk=riskList[[i]],
                            times = as.numeric(colnames(riskList[[i]])))[,2]
   }
@@ -225,7 +226,7 @@ pmnnModel<-function(features, feature_input, feature_output, originalData=origin
 
 
 
-fitSmoothHaz<-function(pmnn,epochs=20000,batch_size=500,verbose=0,monitor="val_loss",val_split=0.2,min_delta=10^-7,patience=10){
+fitSmoothHaz<-function(pmnn,epochs=20000,batch_size=500,verbose=0,monitor="val_loss",val_split=0.2,min_delta=10^-7,patience=10,val){
   #pmnn$casebaseData
   
   offset<-as.matrix(pmnn$offset)
@@ -240,7 +241,8 @@ fitSmoothHaz<-function(pmnn,epochs=20000,batch_size=500,verbose=0,monitor="val_l
     y = y_train,
     epochs = epochs,#30000,
     batch_size = batch_size,#54540,
-    validation_split = val_split,
+    #validation_split = val_split,
+    validation_data=val,
     shuffle=T,
     verbose=verbose,callbacks=list(callback_early_stopping(
       monitor = monitor,
[…]

### 2022-12-15 analyses/src/oldmortDeephitter.py

@@ -14,11 +14,11 @@ from pycox.evaluation import EvalSurv
 
 
 
-def fitDeephit(train,fullTest,bsize,epochs,ti='time',ev='status',patience=10,min_delta=10**-7,drpt=0.5):#,epoch,bsize,tPos,ePos):
+def fitDeephit(train,fullTest,bsize,epochs,valida,ti='time',ev='status',patience=10,min_delta=10**-7,drpt=0.5,lay1=200,lay2=20):#,epoch,bsize,tPos,ePos):
   n= np.shape(train)[0]
   vl_size = int(n*0.20)
-  df_val = train.sample(frac=0.2)
-  df_train = train.drop(df_val.index)
+  df_val = valida#train.sample(frac=0.2)
+  df_train = train#.drop(df_val.index)
   df_test = fullTest
   varNames = list(df_train.columns)
   unwanted = {ti, ev}
@@ -49,24 +49,26 @@ def fitDeephit(train,fullTest,bsize,epochs,ti='time',ev='status',patience=10,min
   #prepare model
   ########
   in_features = x_train.shape[1]
-  num_nodes = [200,20]
+  num_nodes =[int(lay1),int(lay2)]
   out_features = labtrans.out_features
-  batch_norm = False
+  batch_norm = True
   dropout = drpt
   
-  #net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout,activation=torch.tanh)
+  #net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout,activation=torch.nn.SELU)
   net = torch.nn.Sequential(
      torch.nn.BatchNorm1d(in_features),
-     torch.nn.Linear(in_features,200),
+     torch.nn.Linear(in_features,int(lay1)),
+     #torch.nn.BatchNorm1d(int(lay1)),
      torch.nn.Tanh(),
-     torch.nn.Dropout(0.1),
-    
-     torch.nn.Linear(200, 20),
+     torch.nn.Dropout(float(dropout)),
+
+     torch.nn.Linear(int(lay1), int(lay2)),
+     #torch.nn.BatchNorm1d(int(lay2)),
      torch.nn.Tanh(),
-     torch.nn.Dropout(0.1),
-    
-     torch.nn.Linear(20, out_features))
-  
+     torch.nn.Dropout(float(dropout)),
+
+     torch.nn.Linear(int(lay2), out_features))
+
   
[…]

### 2023-02-08 analyses/complexSim.Rmd
checkpoint

--- 2022-05-25 analyses/complexSim.Rmd
@@
   epo: 2000
-  patience: 10
-  iteration: 2
-  layer1: 50
-  layer2: 50
-  layer3: 25
-  layer4: 25
-  drpt: 0.5
+  patience: 5
+  iteration: 1
+  layer1: 12
+  layer2: 12
+  layer3: 100
+  layer4: 100
+  drpt: 0.1
+  lr: 0.01
 ---
@@
 ##########################
-bsize=512
+#bsize=512
 min_delta = 10^-7
+
 epo=as.numeric(params$epo)
@@
 iteration<-as.numeric(params$iteration)
-layer1<-as.numeric(params$layer1)
-layer2<-as.numeric(params$layer2)
-layer3<-as.numeric(params$layer3)
+layer1<-3*7#as.numeric(params$layer1)
+layer2<-3*6#as.numeric(params$layer2)
+layer3<-3*5#as.numeric(params$layer3)
 layer4<-as.numeric(params$layer4)
 drpt<-as.numeric(params$drpt)
+lr=as.numeric(params$lr)
+paramTrials<-readRDS("complexTrials.rds") 
+lr<-paramTrials[which.max(paramTrials$integBrier),]$lr
+layer1<-paramTrials[which.max(paramTrials$integBrier),]$layer1
+layer2<-paramTrials[which.max(paramTrials$integBrier),]$layer2
+layer3<-3*5#as.numeric(params$layer3)
+layer4<-as.numeric(params$layer4)
+drpt<-paramTrials[which.max(paramTrials$integBrier),]$drpt
+bsize=paramTrials[which.max(paramTrials$integBrier),]$bsize
+activ<-paramTrials[which.max(paramTrials$integBrier),]$actv
+
+lr<-paramTrials[which.min(paramTrials$valLoss),]$lr
+layer1<-paramTrials[which.min(paramTrials$valLoss),]$layer1
+layer2<-paramTrials[which.min(paramTrials$valLoss),]$layer2
+layer3<-3*5#as.numeric(params$layer3)
+layer4<-as.numeric(params$layer4)
+drpt<-paramTrials[which.min(paramTrials$valLoss),]$drpt
+bsize=paramTrials[which.min(paramTrials$valLoss),]$bsize
+activ<-paramTrials[which.min(paramTrials$valLoss),]$actv
 
@@
     betas[["z3"]] * x[["z3"]]+
-    betas[["term1"]] * x[["z1"]]*x[["z2"]]*t+
+    betas[["term1"]] * x[["z1"]]*t*t+#*x[["z2"]]*
     betas[["term2"]] * x[["z1"]]*x[["z3"]]+
@@
 coefficients<-true_mod$coefficients
-coefficients<-c(coefficients,z1=1,z2=1,z3=1,term1=10,term2=2,term3=2)
+coefficients<-c(coefficients,z1=0.05,z2=1.05,z3=1.05,term1=1.95,term2=1.05,term3=1.05)
 
@@
 ```{r totalSim,eval=hardsim}
+
+#########################
+###Simulate date
+#########################
+  cov<-data.frame(id=1:numSamp,
+                  z1=c(rep(1,numSamp/2),rep(0,numSamp/2)),
+                  z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
+                  z3=c(rnorm(numSamp/2,0,0.5),rnorm(numSamp/2,10,0.5)))
+  true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1,
+                                       data = brcancer, k = 3
+  ) #from flexsurv
+
+  dat <- simsurv(betas = coefficients, # "true" parameter values
+                 x = cov,                   # covariate data for 200 individuals
+                 knots = true_mod$knots,    # knot locations for splines
+                 logcumhazard = logcumhaz,  # definition of log cum hazard
+                 maxt = 10,               # no right-censoring
+                 interval = c(0,10000000)) # interval for root finding
+  colnames(dat)<-c("id","time","status")
+  #samp<-sample(seq(1,nrow(dat),by=1), floor(nrow(dat)*(0.1))) #random censoring
+  #dat$status[samp]<-0
+  # Merge the simulated event times onto covariate data frame
+  data <- merge(cov, dat)
+  covs<-colnames(data)
+  data<-data[,-1]
+
+  toDiv<-max(data$time)
+
+  data$time<-data$time/toDiv
+ spec = c(train = .7, test = .15, validate = .15)
+
+ g = sample(cut(
+   seq(nrow(data)), 
+   nrow(data)*cumsum(c(0,spec)),
+   labels = names(spec)
+ ))
+
+ res = split(data, g)
+
+trainOG<-res$train
+
+test<-res$test[,-c(ncol(data))]
+
+fullTest<-res$test
+
+val<-res$validate
+
+numSamp<-nrow(trainOG)
 
@@
                   z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
-                  z3=c(rnorm(numSamp/2,-3,0.5),rnorm(numSamp/2,8,0.5)))
+                  z3=c(rnorm(numSamp/2,0,0.5),rnorm(numSamp/2,10,0.5)))
   true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1,
@@
                  logcumhazard = logcumhaz,  # definition of log cum hazard
-                 maxt = 1,               # no right-censoring
+                 maxt = 10,               # no right-censoring
                  interval = c(0,10000000)) # interval for root finding
   colnames(dat)<-c("id","time","status")
-  samp<-sample(seq(1,nrow(dat),by=1), floor(nrow(dat)*(0.1))) #random censoring
-  dat$status[samp]<-0
+  samp<-sample(seq(1,nrow(dat),by=1), floor(nrow(dat)*(0.05))) #random censoring
+  #dat$status[samp]<-0
   # Merge the simulated event times onto covariate data frame
@@
   covs<-colnames(data)
-
+  data<-data[,-1]
+  data$time<-data$time/toDiv
   #########################
@@
   #########################
-  samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.8))
-  data<-data[,-1]
-  data[,-c(1,5)]<-as.data.frame(sapply(data[,-c(1,5)], function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
-  train<-data[samp,]
-  test<-data[-samp,-c(ncol(data))]
-  fullTest<-data[-samp,,drop=F]
-  times<-seq(from=min(test$time),
-             to=max(test$time),
-             length.out = 200
+  #samp<-sample(seq(1,nrow(data),by=1), floor(nrow(data)*.8))
+  #data<-data[,-1]
+  #data[,-c(1,5)]<-as.data.frame(sapply(data[,-c(1,5)], function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
+  train<-data
+
+  sds<-sapply(train, function(x) (sd(x)))
+means<-sapply(train, function(x) (mean(x)))
+test<-normalizer(res$test,means,sds)[,-c(ncol(data))]
+fullTest<-normalizer(res$test,means,sds)
+val<-normalizer(res$validate,means,sds)
+
+train<-normalizer(train,means,sds)
+valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)#rbind(train,val)
+valcb$offset<-0
+valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
+#val<-rbind(train,val)  
+#test<-data[-samp,-c(ncol(data))]
+  #fullTest<-data[-samp,,drop=F]
+  times<-seq(from=min(test$time)+0.000001,
+             to=max(test$time)-0.000001,
+             length.out = 100
   )
-  times<-times[-c(1,length(times))]
+  #times<-times[-c(1,length(times))]
 
@@
   ###########################
-  final_mod_cb_glm <- fitSmoothHazard(status~bs(time)-time+z1+z2+z3+z1:z2:time+z1:z3+z2:z3,
+  final_mod_cb_glm <- fitSmoothHazard(status~bs(time)-time+z1+z2+z3+z1:time:time+z1:z3+z2:z3,
                                       data = train,
@@
   ######################
+    valcbtensor[[1]][[2]]<-rep(mod_cb_glm$offset[1],nrow(valcbtensor[[1]][[1]]))#
   covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2+3),
@@
   covars_output<-covars_input%>% #layer_batch_normalization()%>%
-    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
+    layer_dense(units=layer1,use_bias = T,activation = activ)%>%
     layer_dropout(drpt)%>%
-    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
+    layer_dense(units=layer2,use_bias = T,activation = activ)%>%
     layer_dropout(drpt)%>%
-    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
-    layer_dropout(drpt)%>%
+    #layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
+    #layer_dropout(drpt)%>%
+    #layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
+    #layer_dropout(drpt)%>%
     layer_dense(units=1,use_bias = T)
@@
                   timeVar = "time",
-                  eventVar= "status"
+                  eventVar= "status",optimizer=optimizer_adam(learning_rate =lr )
   )
@@
                           epochs=epo,
-                          batch_size=bsize,
+                          batch_size=bsize,#floor(nrow(mod_cb_glm$data)/100),
                           verbose=0,
                           monitor="val_loss",
-                          val_split=0.2,
+                          #val_split=0.2,
                           min_delta=min_delta,
-                          patience=patience)
+                          patience=patience,val=valcbtensor)
+
    annPreds<-aarSplines(fit,
@@
   class(annProper)<-c("tunnel",class(annProper))
-
+  rm(pmnn,fit,annPreds,covars_input,covars_output)
   ######################
@@
 
-  ######################
-  ###DeepSurv
-  ######################
-  coxnn<-deepsurv(Surv(time, status) ~ .,data = train, frac = 0.2, activation = "relu",
-                  num_nodes = c(layer1,layer2,layer3,layer4), dropout = drpt,
-                  early_stopping = TRUE, epochs = epo,
-                  batch_size = bsize,device = "cpu",best_weights=T,
-                  patience = patience,min_delta=min_delta,batch_norm = F,
-                  learning_rate = 0.001,lr_decay=10^-7)
-  coxnnsurv<- predict(coxnn,newdata = test,type="survival")
-  colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
+  source_python('src/dsurv.py')
+  #need to break ties
+  tiebreakTrain<-train
+
+  #while(length(unique(tiebreakTrain$time))/length(tiebreakTrain$time)!=1 ){
+  #tiebreakTrain<-train
+  #tiebreakTrain$time<-tiebreakTrain$time+runif(nrow(tiebreakTrain),0,1/10000000000)
+  #}
+
+  coxnnsurv=fitDeepSurv(tiebreakTrain,fullTest,floor(nrow(train)/100),epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4,lr=lr)
+  coxnnsurv<-t(coxnnsurv)
+  #colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
   cnnCleaned<-pyProcess(coxnnsurv,times=times)
-  colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
+  colnames(cnnCleaned)<-times#seq(0,1,length.out = ncol(cnnCleaned))
+  py_run_string("del fitDeepSurv")
+  rm(fitDeepSurv)
 
+  # ######################
+  # ###DeepSurv
+  # ######################
+  # coxnn<-deepsurv(Surv(time, status) ~ .,data = train, frac = 0.2, activation = "relu",
+  #                 num_nodes = c(layer1,layer2,layer3,layer4), dropout = drpt,
+  #                 early_stopping = TRUE, epochs = epo,
+  #                 batch_size = bsize,device = "cpu",best_weights=T,
+  #                 patience = patience,min_delta=min_delta,batch_norm = F,
+  #                 learning_rate = 0.001,lr_decay=10^-7)
+  # coxnnsurv<- predict(coxnn,newdata = test,type="survival")
+  # #colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
+  # cnnCleaned<-pyProcess(coxnnsurv,times=times)
+  # colnames(cnnCleaned)<-seq(0,1,length.out = ncol(cnnCleaned))
+  # 
   ######################
@@
   source_python('src/deephitter.py')
-  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,patience=patience,min_delta=min_delta,drpt=drpt)
+  hitnnSurv=fitDeephit(train,fullTest,floor(nrow(train)/100),epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4,lr=lr)
   hitnnSurv<-t(hitnnSurv)
-  colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
+  #colnames(hitnnSurv)<-seq(0,1,length.out = ncol(hitnnSurv))
   deephitCleaned<-pyProcess(hitnnSurv,times=times)
@@
   ######################
+
   covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
@@
                   timeVar = "time",
-                  eventVar= "status"
+                  eventVar= "status",optimizer=optimizer_adam(learning_rate = lr)
   )
@@
                     epochs=epo,
-                    batch_size=bsize,
+                    batch_size=bsize,#floor(nrow(mod_cb_glm$data)/100),
                     verbose=0,
                     monitor="val_loss",
-                    val_split=0.2,
+                    #val_split=0.2,
                     min_delta=min_delta,
-                    patience=patience)
+                    patience=patience,val=valcbtensor)
   annPreds<-aar(fit,
@@
   BrierIPAList[[pos]]<-brierFinalResults$Brier$score
-
+    ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
+  geom_line()+coord_cartesian(ylim=c(-1,1))
+    rm(pmnn,fit,annPreds,covars_input,covars_output)
   #######################
@@
   pos<-pos+1
+  gc()
 }
@@
 ```
+
+```{r}
+
+  ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
+  geom_line()+coord_cartesian(ylim=c(-1,1))
+
+```
+
+```{r, learningTrial,eval=T}
+  times<-seq(from=min(test$time)+0.000001,
+             to=max(test$time)-0.000001,
+             length.out = 25
+  )
+
+trials<-1000000
+numTests<-2
+# paramTrials<-data.frame(#lr=runif(trials,0,0.1),
+#                         #drpt=runif(trials,0,0.5),
+#                         lr=sample(c(0.0001,0.001,0.01,0.1),replace=T),
+#                         drpt=sample(c(0.1,0.2,0.3,0.4),replace=T),
+#                         layer1=sample(seq(10,71,20), trials,replace = T),
+#                         layer2=sample(seq(10,71,20), trials,replace = T),
+#                         #layer3=sample(seq(1,100,25), trials,replace = T),
+#                         bsize=sample(c(floor(nrow(mod_cb_glm$data)/100),floor(nrow(mod_cb_glm$data)/100/2)), trials,replace = T),
+#                         actv=sample(c("relu","tanh",'selu'), trials,replace = T),
+#                         integBrier=rep(NA,trials))
+
+paramTrials<-as.data.frame(tidyr::crossing(lr=c(0.001,0.01,0.1),
+                        drpt=c(0.1,0.2,0.3),
+                        layer1=seq(10,71,25),
+                        layer2=seq(10,71,25),
+                        bsize=c(floor(nrow(mod_cb_glm$data)/100)),
+                        actv=c("relu","tanh"),
+                        integBrier=NA,
+                        valLoss=NA))
+# 
+# paramTrials<-unique.data.frame(paramTrials)
+#paramTrials<-readRDS("complexTrials.rds")
+for(i in 1:nrow(paramTrials)){
+#paramTrials<-readRDS("complexTrials.rds")  
+tempRes<-0
+tempVal<-0
+currentTime<-Sys.time() 
+  for(j in 1:numTests){
+
+  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
+                            name = 'main_input')
+
+  covars_output<-covars_input%>% #layer_batch_normalization()%>%
+    layer_dense(units=paramTrials$layer1[i],use_bias = T,activation = paramTrials$actv[i])%>%
+    layer_dropout(paramTrials$drpt[i])%>%
+    layer_dense(units=paramTrials$layer2[i],use_bias = T,activation = paramTrials$actv[i])%>%
+    layer_dropout(paramTrials$drpt[i])%>%
+    #layer_dense(units=paramTrials$layer3[i],use_bias = T,activation = paramTrials$actv[i])%>%
+    #layer_dropout(paramTrials$drpt[i])%>%
+    #layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
+    #layer_dropout(drpt)%>%
+    layer_dense(units=1,use_bias = T)
+
+  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
+                  feature_input = covars_input,
+                  feature_output = covars_output,
+                  originalData = mod_cb_glm$data,
+                  offset=mod_cb_glm$offset,
+                  timeVar = "time",
+                  eventVar= "status",optimizer=optimizer_adam(learning_rate = paramTrials$lr[i])
+  )
+
+  fit<-fitSmoothHaz(pmnn,
+                    epochs=epo,
+                    batch_size=paramTrials$bsize[i],
+                    verbose=0,
+                    monitor="val_loss",
+                    #val_split=0.2,
+                    min_delta=min_delta,
+                    patience=patience,val=valcbtensor)
+  annPreds<-aar(fit,
+                times=times,
+                x_test=as.data.frame(val[,-c(ncol(data))])#rbind(train,val)
+  )
+
+  rownames(annPreds)<-annPreds[,1]
+  annProperpoly<- t(annPreds[,-1])
+  class(annProperpoly)<-c("tunnel",class(annProperpoly))
+
+  brierFinalResults <- Score(list('PMNN_Poly'=annProperpoly),
+                             data =as.data.frame(val),#rbind(train,val) 
+                             formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
+                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
+  cutted<-brierFinalResults[which(brierFinalResults$model=="PMNN_Poly"),]
+  tempRes<-tempRes+mean(cutted[-c(1,2,24,25)]$IPA)/numTests
+  tempVal<-tempVal+min(fit$resultOfFit$metrics$val_loss)/numTests
+  }
+  paramTrials$integBrier[i]<-tempRes
+  paramTrials$valLoss[i]<-
+  saveRDS(paramTrials,"complexTrials.rds")
+  print(paste("trial",i))
+  print(Sys.time()-currentTime)
+  currentTime<-Sys.time()
+  #print(as.character(paramTrials[i,]))
+    rm(pmnn,fit,annPreds,covars_input,covars_output)
+
+}
+
+```

### 2023-02-08 analyses/fullMetabricAnalysis.Rmd
checkpoint

---
title: "fullMetabricanalysis"
author: "Jesse"
date: "12/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r loadData}
library("illuminaHumanv4.db")
getExpressionDataM <- function(){
load("metabricData/Complete_METABRIC_Expression_Data.rbin")
expressionData<-Complete_METABRIC_Expression_Data@assayData$exprs

geneName<-data.frame(Gene=unlist(mget(x = rownames(expressionData),ifnotfound = NA,envir = illuminaHumanv4SYMBOL)))

geneName[which(is.na(geneName$Gene)),1]<-rownames(geneName)[which(is.na(geneName$Gene))]

rownames(expressionData)<-make.unique(geneName$Gene, sep="_")

return(t(expressionData))

}

load("metabricData/Complete_METABRIC_Clinical_Survival_Data_OS.rbin")

```

```{r processData}

library(survival)
timeEvent<-as.data.frame(Complete_METABRIC_Clinical_Survival_Data_OS)

timeEvent<-tidyr::separate(data = timeEvent, col =  "x",into =  c("time", "status"), sep = "\\+", extra = "merge")

timeEvent$status[which(is.na(timeEvent$status))]<-1

timeEvent$status[which(timeEvent$status=="")]<-0

#rm(Complete_METABRIC_Expression_Data)
rm(Complete_METABRIC_Clinical_Survival_Data_OS)
#rm(geneName)
gc()

```

```{r splitData}
set.seed(1)
data<-getExpressionDataM()

 spec = c(train = .7, test = .1, validate = .2)

 g = sample(cut(
   seq(nrow(data)), 
   nrow(data)*cumsum(c(0,spec)),
   labels = names(spec)
 ))

 res = list()
res$train<-  data[which(g=="train"),]
 res$validate<-  data[which(g=="validate"),] 
 res$test<-  data[which(g=="test"),]

saveRDS(res,"expressionSplit.rds")

```

```{r featureSelection}
res<-readRDS("expressionSplit.rds")$train
spearCor<-cor(res,method = c( "spearman"))
d <- as.dist(1-spearCor)
rm(spearCor,res)
gc()
hc <- hclust(d, method="complete")
#get distance
#get hclust
rm(d)
gc()

saveRDS(hc,"hierarchicalClusteringSpearmanDistance.rds")
```

```{r,fig.height=25,fig.width=25}

hc<-readRDS("hierarchicalClusteringSpearmanDistance.rds")
set.seed(1)

plot(hc)

clusts <- cutree(hc,k=1000)
i=1
chosen=data.frame(reps=rep("representative",1000))

for(i in 1:length(unique(clusts))){

chosen$reps[i]<-sample(names(which(clusts==i))[!grepl("^ILMN",names(which(clusts==i)))],1)

}

```

```{r}

expData<-getExpressionDataM()

importantGenePos<-which(colnames(expData) %in% chosen$reps)
importantGenes<-expData[,importantGenePos]

toPC<-expData[,-importantGenePos]

pcs <- prcomp(toPC, center = TRUE,scale. = TRUE)

```

### 2023-02-08 analyses/metabricStudy.Rmd
checkpoint

@@
   iteration: 1
-  layer1: 50
-  layer2: 50
-  layer3: 25
-  layer4: 25
+  layer1: 100
+  layer2: 100
+  layer3: 50
+  layer4: 50
   drpt: 0.1
+  lr: 0.001
 ---
@@
 ```{r setup, include=T}
-set.seed(1)
+set.seed(2)#set seed to 2, because testing set had shortest survival time censored, resulting in an error in the c-index calculation method.
 knitr::opts_chunk$set(echo = TRUE)
@@
 bsize=512
-min_delta = 0
+min_delta = 10^-7
 epo=as.numeric(params$epo)
@@
 drpt<-as.numeric(params$drpt)
+lr=as.numeric(params$lr)
 
@@
 data<-as.data.frame(model.matrix(~.,metabric)[,-1])
-data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
+#data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
 covs<-colnames(data)
@@
 valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
+
+  # times<-seq(from=min(test$time[fullTest$status==1]),
+  #            to=max(test$time[fullTest$status==1]),
+  #            length.out = 100
+  # )
 
@@
   source_python('src/dsmScript.py')
-  dsmSurv<-dsmfitter(train=train,fullTest=fullTest,bsize=bsize,epochs=epo,val=val,times=times,tPos=ncol(train)-2,ePos=ncol(train)-1)
+  dsmSurv<-dsmfitter(train=train,fullTest=fullTest,bsize=bsize,epochs=epo,val=val,times=times,tPos=ncol(train)-2,ePos=ncol(train)-1,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4,lr=lr)
   if((sum(is.na(dsmSurv))==ncol(dsmSurv)*nrow(dsmSurv))){
@@
 
-  coxnnsurv=fitDeepSurv(tiebreakTrain,fullTest,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4)
+  coxnnsurv=fitDeepSurv(tiebreakTrain,fullTest,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4,lr=lr)
   coxnnsurv<-t(coxnnsurv)
@@
     source_python('src/deephitter.py')
-  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4)
+  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4,lr=lr)
   hitnnSurv<-t(hitnnSurv)
@@
   ############################  
+    valcbtensor[[1]][[2]]<-rep(mod_cb_glm$offset[1],nrow(valcbtensor[[1]][[1]]))
   covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2+3),
@@
   covars_output<-covars_input%>% #layer_batch_normalization()%>%
-    layer_dense(units=layer1,use_bias = T,activation = 'tanh')%>%
+    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
     layer_dropout(drpt)%>%
-    layer_dense(units=layer2,use_bias = T,activation = 'tanh')%>%
+    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
     layer_dropout(drpt)%>%
-    layer_dense(units=layer3,use_bias = T,activation = 'tanh')%>%
+    layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
     layer_dropout(drpt)%>%
-    layer_dense(units=layer4,use_bias = T,activation = 'tanh')%>%
+    layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
     layer_dropout(drpt)%>%
@@
                   timeVar = "time",
-                  eventVar= "status",optimizer=optimizer_adam(learning_rate = 0.0001)
+                  eventVar= "status",optimizer=optimizer_adam(learning_rate = lr)
   )

### 2023-02-08 analyses/optimizeComplex.Rmd
checkpoint

---
title: "iterSim"
output: html_document
params:
  epo: 200
  patience: 5
  iteration: 2
  layer1: 12
  layer2: 12
  layer3: 100
  layer4: 100
  drpt: 0.1
  lr: 0.01
---

```{r}
library(reticulate)

```

```{python}

import os
os.environ["CUDA_DEVICE_ORDER"] = "PCI_BUS_ID" 
os.environ["CUDA_VISIBLE_DEVICES"] = ""

```

```{r setup, include=T}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
source("src/packages.R")
source("src/functions.R")
library(reticulate)
eval_support<-T
#source_python('src/getSUP.py')
source_python("src/cIndex.py")
#source_python("src/dsmScript.py")
#source_python('src/deephitter.py')
##########################
###Shared Hyperparameters
##########################
#bsize=512
min_delta =0# 10^-7

epo=as.numeric(params$epo)
patience <- as.numeric(params$patience)
iteration<-as.numeric(params$iteration)

numSamp<-2000

baselineTrials<-2

reoptimize=T

finalFit=T
```

#5 fold cross validation to select best hyperparameters, average performance on a test set.

```{r complexPrep,eval=T}

##############################
###set up complex simulation
##############################

# Define a function returning the log cum hazard at time t
logcumhaz <- function(t, x, betas, knots) {

  # Obtain the basis terms for the spline-based log
  # cumulative hazard (evaluated at time t)
  basis <- flexsurv::basis(knots, log(t))

  # Evaluate the log cumulative hazard under the
  # Royston and Parmar specification
  res <-betas[["gamma0"]] * basis[[1]] + 
    betas[["gamma1"]] * basis[[2]] +
    betas[["gamma2"]] * basis[[3]] +
    betas[["gamma3"]] * basis[[4]] +
    betas[["gamma4"]] * basis[[5]] +
    betas[["z1"]] * x[["z1"]]+
    betas[["z2"]] * x[["z2"]]+
    betas[["z3"]] * x[["z3"]]+
    betas[["term1"]] * x[["z1"]]*t+#*x[["z2"]]*
    #betas[["term2"]] * x[["z1"]]*x[["z3"]]+
    betas[["term3"]] * x[["z2"]]*(x[["z3"]])
  #
  # Return the log cumulative hazard at time t
  res
}
url<-'app.txt'
ds <- read.table(url, sep = '',header = TRUE)
ds$years.at.risk=pmin(ds$agein80,ds$onset)
brcancer$rectime<-brcancer$rectime/max(brcancer$rectime)
true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1, data = brcancer, k = 3) #from flexsurv

# Simulate the event times
coefficients<-true_mod$coefficients
coefficients<-c(coefficients,z1=0.05,z2=1.05,z3=1.05,term1=2.4,term3=1.05)#term2=1.05,

#########################
###Simulate data
#########################
cov<-data.frame(id=1:numSamp,
                z1=c(rep(1,numSamp/2),rep(0,numSamp/2)),
                z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
                z3=c(rnorm(numSamp/2,0,0.5),rnorm(numSamp/2,10,0.5)))
true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1,
                                     data = brcancer, k = 3
) #from flexsurv

dat <- simsurv(betas = coefficients, # "true" parameter values
               x = cov,                   # covariate data for 200 individuals
               knots = true_mod$knots,    # knot locations for splines
               logcumhazard = logcumhaz,  # definition of log cum hazard
               maxt = 25,               # no right-censoring
               interval = c(0,10000000)) # interval for root finding
colnames(dat)<-c("id","time","status")
samp<-sample(seq(1,nrow(dat),by=1), floor(nrow(dat)*(0.1))) #random censoring
dat$status[samp]<-0
# Merge the simulated event times onto covariate data frame
data <- merge(cov, dat)
covs<-colnames(data)
data<-data[,-1]

toDiv<-max(data$time)
data$time<-data$time/toDiv
spec = c(train = .85, test = .15)

g = sample(cut(
  seq(nrow(data)), 
  nrow(data)*cumsum(c(0,spec)),
  labels = names(spec)
))

res = split(data, g)
saveRDS(res$test,"ComplexSimTest.rds")

k=5
klist<-list()
data<-res$train
for(i in 1:k){

  spec = c(train = .85, val = .15)

  g = sample(cut(
    seq(nrow(data)), 
    nrow(data)*cumsum(c(0,spec)),
    labels = names(spec)
  ))
  klist[[i]]= split(data, g)
}
saveRDS(klist,"ComplexSimCV.rds")

```

```{r setFixedHyperparameters}

klist<-readRDS("ComplexSimCV.rds")
# lr=c(0.001,0.01)
# drpt=c(0.25,0.5)
# layer1=c(50,75,100)
# layer2=c(50,75)
# bsize=c(floor(nrow(klist[[1]]$train)/100))
# actv=c("relu","tanh")
# alp=c(0,0.5,1)

lr=c(0.001,0.0001)
drpt=c(.1,.2)
layer1=c(10,25,50)
layer2=c(10,20,30)
bsize=c(floor(nrow(klist[[1]]$train)/100))
actv=c("relu","tanh")
alp=c(0,0.5,1)

```

```{r hyperOpDsurv,eval=T}

optimDeepSurv<-function(klist,test,times,baselineTrials,lr,drpt,layer1,layer2,bsize,actv){

  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
                                             drpt=drpt,
                                             layer1=layer1,
                                             layer2=layer2,
                                             bsize=c(bsize),
                                             actv=actv,
                                             #IPA=NA,
                                             #IPAse=NA,
                                             valLoss=NA,
                                             runtime=NA))

  #paramTrials<-paramTrials[sample(nrow(paramTrials), baselineTrials), ]

  # paramTrials<-unique.data.frame(paramTrials)
  #paramTrials<-readRDS("complexTrials.rds")
  for(i in 1:nrow(paramTrials)){
    #paramTrials<-readRDS("complexTrials.rds")  
    tempRes<-c()
    tempVal<-c()
    runTime<-c()

    for(data in klist){

      train<-data$train
      val<-data$val

      sds<-sapply(train, function(x) (sd(x)))
      means<-sapply(train, function(x) (mean(x)))
      val<-normalizer(val,means,sds)
      train<-normalizer(train,means,sds)

      lr=paramTrials[i,]$lr
      drpt=paramTrials[i,]$drpt
      layer1=paramTrials[i,]$layer1
      layer2=paramTrials[i,]$layer2
      bsize=paramTrials[i,]$bsize
      actv=paramTrials[i,]$actv
      #=paramTrials[i,]$IPA
      #=paramTrials[i,]$valLoss
      source_python('src/optimDSurv.py')

      currentTime<-Sys.time() 
      coxnnsurv=fitDeepSurv(train,val,floor(nrow(train)/100),epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
      tempVal<-append(tempVal,min(coxnnsurv$val_metrics$scores$loss$score))
      runTime<-append(runTime, Sys.time()-currentTime)
      currentTime<-Sys.time()

      # coxnnsurv<-coxnnsurv[[1]]
      # coxnnsurv<-t(coxnnsurv)
      # #colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
      # cnnCleaned<-pyProcess(coxnnsurv,times=times)
      # colnames(cnnCleaned)<-times#seq(0,1,length.out = ncol(cnnCleaned))
      # py_run_string("del fitDeepSurv")
      rm(fitDeepSurv)
      # brierFinalResults <- Score(list('ds'=cnnCleaned),
      #                            data =as.data.frame(val),#rbind(train,val) 
      #                            formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
      #                            se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
      # cutted<-brierFinalResults[which(brierFinalResults$model=="ds"),]
      #tempRes<-append(tempRes,mean(cutted[-c(1,2,24,25)]$IPA))
    }

    #paramTrials$IPA[i]<-mean(tempRes)
    #paramTrials$IPAse[i]<-sd(tempRes)
    paramTrials$valLoss[i]<-mean(tempVal)
    paramTrials$runtime[i]<-mean(runTime)
    #saveRDS(paramTrials,"complexds.rds")
    print(paste("trial",i))
    print(sum(runTime))
    #print(as.character(paramTrials[i,]))
    #rm(cbnn,fit,annPreds,covars_input,covars_output)
    gc()
  }
  bestPerformance<-paramTrials[which.min(paramTrials$valLoss),]
  return(bestPerformance)
}

klist<-readRDS("ComplexSimCV.rds")
test<-readRDS("ComplexSimTest.rds")
times<-seq(from=min(test$time)+0.000001,
           to=max(test$time)-0.000001,
           length.out = 25
)

bestDS<-optimDeepSurv(klist,test,times,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,bsize,actv)
```

```{r optimDeepHit,eval=T}

optimDeepHit<-function(klist,test,times,baselineTrials,lr,drpt,layer1,layer2,bsize,actv,alp){

  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
                                             drpt=drpt,
                                             layer1=layer1,
                                             layer2=layer2,
                                             bsize=c(bsize),
                                             actv=actv,
                                             alp=alp,
                                             #IPA=NA,
                                             #IPAse=NA,
                                             valLoss=NA,
                                             runtime=NA))

  #paramTrials<-paramTrials[sample(nrow(paramTrials), baselineTrials), ]

  # paramTrials<-unique.data.frame(paramTrials)
  #paramTrials<-readRDS("complexTrials.rds")
  for(i in 1:nrow(paramTrials)){
    #paramTrials<-readRDS("complexTrials.rds")  
    tempRes<-c()
    tempVal<-c()
    runTime<-c()

    for(data in klist){

      train<-data$train
      val<-data$val

      sds<-sapply(train, function(x) (sd(x)))
      means<-sapply(train, function(x) (mean(x)))
      val<-normalizer(val,means,sds)
      train<-normalizer(train,means,sds)

      lr=paramTrials[i,]$lr
      drpt=paramTrials[i,]$drpt
      layer1=paramTrials[i,]$layer1
      layer2=paramTrials[i,]$layer2
      bsize=paramTrials[i,]$bsize
      actv=paramTrials[i,]$actv
      alp=paramTrials[i,]$alp
      #=paramTrials[i,]$IPA
      #=paramTrials[i,]$valLoss
      source_python('src/optimDH.py')

      currentTime<-Sys.time() 
      hitnnSurv=fitDeephit(train,val,floor(nrow(train)/100),epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
      tempVal<-append(tempVal,min(hitnnSurv$val_metrics$scores$loss$score))
      runTime<-append(runTime, Sys.time()-currentTime)
      currentTime<-Sys.time() 
    }
    rm(fitDeephit)

    #paramTrials$IPA[i]<-mean(tempRes)
    #paramTrials$IPAse[i]<-sd(tempRes)
    paramTrials$valLoss[i]<-mean(tempVal)
    paramTrials$runtime[i]<-mean(runTime)
    #saveRDS(paramTrials,"complexds.rds")
    print(paste("trial",i))
    print(mean(runTime))
    #print(as.character(paramTrials[i,]))
    #rm(cbnn,fit,annPreds,covars_input,covars_output)
    gc()
  }
  bestPerformance<-paramTrials[which.min(paramTrials$valLoss),]
  return(bestPerformance)
}

klist<-readRDS("ComplexSimCV.rds")
test<-readRDS("ComplexSimTest.rds")
times<-seq(from=min(test$time)+0.000001,
           to=max(test$time)-0.000001,
           length.out = 25
)

bestDH<-optimDeepHit(klist,test,times,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,bsize,actv,alp)

```

```{r hyperopcbnn,eval=T}

optimCBNN<-function(klist,test,times,lr,drpt,layer1,layer2,bsize,actv){

  # 
  # for (i in 1:length(klist)){
  #   #train<-sampleCaseBase(data = as.data.frame(klist[[i]]$train),time="time",event="status",ratio=100)
  #   valcb<-sampleCaseBase(data = as.data.frame(klist[[i]]$val),time="time",event="status",ratio=100)
  #   valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
  #                          as.matrix(valcb[,ncol(valcb),drop=F])),
  #                     as.matrix(valcb[,ncol(valcb)-1,drop=F]))
  #   #valcbtensor[[1]][[2]]<-rep(train$offset[1],nrow(valcbtensor[[1]][[1]]))      
  #   #klist[[i]]$train<-train
  #   klist[[i]]$val<-valcbtensor
  # }

  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
                                             drpt=drpt,
                                             layer1=layer1,
                                             layer2=layer2,
                                             #bsize=,
                                             actv=actv,
                                             #IPA=NA,
                                             #IPAse=NA,
                                             valLoss=NA,
                                             runtime=NA))

  #paramTrials<-paramTrials[sample(nrow(paramTrials), 2), ]

  # paramTrials<-unique.data.frame(paramTrials)
  #paramTrials<-readRDS("complexTrials.rds")
  for(i in 1:nrow(paramTrials)){
    #paramTrials<-readRDS("complexTrials.rds")  
    tempRes<-c()
    tempVal<-c()
    runTime<-c()

    for(data in klist){

      train<-data$train
      val<-data$val

      sds<-sapply(train, function(x) (sd(x)))
      means<-sapply(train, function(x) (mean(x)))
      #val[[1]][[1]]<-as.matrix(normalizer(as.data.frame(val[[1]][[1]]),means,sds))
      train<-normalizer(train,means,sds)
      #train[,ncol(train)]<-data$train[,ncol(train)]
      val<-normalizer(val,means,sds)
      train<-sampleCaseBase(data = as.data.frame(train),time="time",event="status",ratio=100)
      valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)
      val<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
                     as.matrix(valcb[,ncol(valcb),drop=F])
      ),
      as.matrix(valcb[,ncol(valcb)-1,drop=F])
      )
      val[[1]][[2]]<-rep(train$offset[1],nrow(val[[1]][[1]]))

      lr=paramTrials[i,]$lr
      drpt=paramTrials[i,]$drpt
      layer1=paramTrials[i,]$layer1
      layer2=paramTrials[i,]$layer2
      bsize=c(floor(nrow(train)/100))
      actv=paramTrials[i,]$actv

      currentTime<-Sys.time()

      covars_input<-layer_input(shape=c(length(colnames(train))-2),
                                name = 'main_input')

      covars_output<-covars_input%>% 
        layer_dense(units=layer1,use_bias = T,activation = actv)%>%
        layer_dropout(drpt)%>%
        layer_dense(units=layer2,use_bias = T,activation = actv)%>%
        layer_dropout(drpt)%>%
        layer_dense(units=1,use_bias = T)

      cbnn<-cbnnModel(features=colnames(train)[-ncol(train)],
                      feature_input = covars_input,
                      feature_output = covars_output,
                      originalData = train,
                      offset=train$offset,
                      timeVar = "time",
                      eventVar= "status",optimizer=optimizer_adam(learning_rate = lr)
      )

      fit<-fitSmoothHaz(cbnn,
                        epochs=epo,
                        batch_size=bsize,
                        verbose=0,
                        monitor="val_loss",
                        #val_split=0.2,
                        min_delta=min_delta,
                        patience=patience,val=val)
      tempVal<-append(tempVal,min(fit$resultOfFit$metrics$val_loss))
      runTime<-append(runTime, Sys.time()-currentTime)
      currentTime<-Sys.time()

      rm(cbnn,fit,covars_input,covars_output)
    }
    paramTrials$valLoss[i]<-mean(tempVal)
    paramTrials$runtime[i]<-mean(runTime)
    #saveRDS(paramTrials,"complexds.rds")
    print(paste("trial",i,"out of ",nrow(paramTrials)))
    print(sum(runTime))
    #print(as.character(paramTrials[i,]))
    #rm(cbnn,fit,annPreds,covars_input,covars_output)
    gc()
  }
  bestPerformance<-paramTrials[which.min(paramTrials$valLoss),]
  return(bestPerformance)
}

klist<-readRDS("ComplexSimCV.rds")
test<-readRDS("ComplexSimTest.rds")

cbnnPerformance<-optimCBNN(klist,test,times,lr,drpt,layer1,layer2,bsize,actv)
#cbnnPerformance<-cbnnPerformance[which.min(cbnnPerformance$valLoss),]

```

```{r}

winners<-list(cbnn=cbnnPerformance,
              ds=bestDS,
              dh=bestDH)

saveRDS(winners,"complexOptimWins.rds")
```

```{r bestFits,finalFit}

winners<-readRDS("complexOptimWins.rds")

klist<-readRDS("ComplexSimCV.rds")
data<-rbind(klist[[1]]$train,klist[[1]]$val)

start_time <- Sys.time()
BrierIPAList<-list()
cScore<-list()
pos<-1

spec = c(train = .85, validate = .15)

g = sample(cut(
  seq(nrow(data)), 
  nrow(data)*cumsum(c(0,spec)),
  labels = names(spec)
))

res = split(data, g)

while (pos <= iteration) {

  trainOG<-res$train

  numSamp<-nrow(trainOG)
  samp<-sample(seq(1,nrow(trainOG),by=1) , nrow(trainOG),replace = T) 
  train<-trainOG[samp,]

  val<-res$validate

  sds<-sapply(train, function(x) (sd(x)))
  means<-sapply(train, function(x) (mean(x)))

  fullTest<-readRDS("ComplexSimTest.rds")
  test<-fullTest[,-c(ncol(data))]
  test<-normalizer(fullTest,means,sds)[,-c(ncol(data))]
  fullTest<-normalizer(fullTest,means,sds)

  val<-normalizer(res$validate,means,sds)

  train<-normalizer(train,means,sds)
  valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)#rbind(train,val)
  valcb$offset<-0
  valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
  times<-seq(from=min(test$time)+0.000001,
             to=max(test$time)-0.000001,
             length.out = 27
  )

  times<-head(times, -1)
  times<-tail(times, -1)

  ###########################
  ###casebase optimal model
  ###########################
  final_mod_cb_glm <- fitSmoothHazard(status~bs(time)-time+z1+z2+z3+z1:time+z2:z3,#+z1:z3
                                      data = train,
                                      time = "time",
                                      event="status",
                                      ratio = 100
  )

  summary(final_mod_cb_glm)
  tglmAbsRisk<-absoluteRisk(final_mod_cb_glm,time=times,newdata=test,type="CI")
  tglmAbsRisk<-as.data.frame(tglmAbsRisk)
  rownames(tglmAbsRisk)<-tglmAbsRisk$time
  trueGlmProper<- t(tglmAbsRisk[-1,-1])
  class(trueGlmProper)<-c("tunnel",class(trueGlmProper))

  #########################
  ###casebase+splines
  ##########################
  mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
                                data = train,
                                time = "time",
                                event="status",
                                ratio = 100
  )

  glmAbsRisk<-as.data.frame(absoluteRisk(mod_cb_glm,time=times,newdata=test,type="CI"))
  rownames(glmAbsRisk)<-glmAbsRisk$time
  glmProper<- t(glmAbsRisk[-1,-1])
  class(glmProper)<-c("tunnel",class(glmProper))

  #############################
  ###DEEPSURV
  #############################

  lr=winners$ds$lr
  drpt=winners$ds$drpt
  layer1=winners$ds$layer1
  layer2=winners$ds$layer2
  bsize=winners$ds$bsize
  actv=winners$ds$actv

  source_python('src/dsurv.py')
  coxnnsurv=fitDeepSurv(train,fullTest,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
  coxnnsurv<-t(coxnnsurv)
  cnnCleaned<-pyProcess(coxnnsurv,times=times)
  colnames(cnnCleaned)<-times
  py_run_string("del fitDeepSurv")
  rm(fitDeepSurv)

  #############################
  ###DEEPHIT
  #############################
  lr=winners$dh$lr
  drpt=winners$dh$drpt
  layer1=winners$dh$layer1
  layer2=winners$dh$layer2
  bsize=winners$dh$bsize
  actv=winners$dh$actv
  alp=winners$dh$alp
  source_python('src/deephitter.py')
  hitnnSurv=fitDeephit(train,fullTest,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
  hitnnSurv<-t(hitnnSurv)
  deephitCleaned<-pyProcess(hitnnSurv,times=times)
  colnames(deephitCleaned)<-times
  py_run_string("del fitDeephit")
  rm(fitDeephit)

  ############################
  ###coxph
  ############################  
  cox<-coxph(Surv(time, status) ~ ., data = train,x=T)

  #############################
  ###CBNN
  #############################
  train<-sampleCaseBase(data = as.data.frame(train),time="time",event="status",ratio=100)
  valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)
  val<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
                 as.matrix(valcb[,ncol(valcb),drop=F])),
            as.matrix(valcb[,ncol(valcb)-1,drop=F]))

  val[[1]][[2]]<-rep(train$offset[1],nrow(val[[1]][[1]]))   
  lr=winners$cbnn$lr
  drpt=winners$cbnn$drpt
  layer1=winners$cbnn$layer1
  layer2=winners$cbnn$layer2
  #bsize=winners$cbnn$bsize
  actv=winners$cbnn$actv

  # lr=0.001
  # drpt=0.1
  # layer1=50
  # layer2=20
  # #bsize=winners$cbnn$bsize
  # actv='relu'

  currentTime<-Sys.time() 
  # valcbtensor[[1]][[2]]<-rep(mod_cb_glm$offset[1],nrow(valcbtensor[[1]][[1]]))
  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
                            name = 'main_input')

  covars_output<-covars_input%>%
    layer_dense(units=layer1,use_bias = T,activation = actv)%>%
    layer_dropout(drpt)%>%
    layer_dense(units=layer2,use_bias = T,activation = actv)%>%
    layer_dropout(drpt)%>%
    layer_dense(units=1,use_bias = T)

  cbnn<-cbnnModel(features=colnames(train)[-ncol(train)],
                  feature_input = covars_input,
                  feature_output = covars_output,
                  originalData = train,
                  offset=train$offset,
                  timeVar = "time",
                  eventVar= "status",optimizer=optimizer_adam(learning_rate = lr)
  )

  fit<-fitSmoothHaz(cbnn,
                    epochs=epo,
                    batch_size=c(floor(nrow(train)/100)),
                    verbose=0,
                    monitor="val_loss",
                    #val_split=0.2,
                    min_delta=min_delta,
                    patience=patience,val=val)
  annPreds<-aar(fit,
                times=times,
                x_test=test
  )

  rownames(annPreds)<-annPreds[,1]
  annProperpoly<- t(annPreds[,-1])
  class(annProperpoly)<-c("tunnel",class(annProperpoly)) 
  rm(fit,cbnn,covars_input,covars_output)
  ############################
  ###Brier Score
  ############################  
  brierFinalResults <- Score(list("Cox_Lin" = cox,'CB_Logi'=glmProper,
                                  'DeepSurv'=cnnCleaned,'DeepHit'=deephitCleaned,
                                  'CBNN_Poly'=annProperpoly,'optimal'=trueGlmProper),
                             data =fullTest, 
                             formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)
  BrierIPAList[[pos]]<-brierFinalResults$Brier$score

  #######################
  ###Cox 
  ######################
  a<-survfit(cox, newdata=fullTest,times=times)
  rownames(a$surv)<-a$time
  coxlinSurv<-pyProcess(t(a$surv),times=times)
  colnames(coxlinSurv)<-seq(0,1,length.out = ncol(coxlinSurv))

  ######################
  ###C-Index
  ######################

  riskList<-list(coxlinSurv,
                 glmProper,
                 cnnCleaned,
                 deephitCleaned,
                 annProperpoly,
                 trueGlmProper
  )

  tempTims<-   as.numeric(colnames(annProperpoly))
  tempTims<-head(tempTims, -1)
  tempTims<-tail(tempTims, -1)
  cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=length(riskList)+1)
  et_train<-as.matrix(train)[,c(5,4)]
  et_test<-as.matrix(fullTest[,c(5,4)])
  colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
                          'cbnn','optimal')
  cScoreTemp[,1]<- tempTims

  cScore[[pos]]<-cIndexSummary(et_train=et_train,
                               et_test=et_test,
                               riskList=riskList,
                               cScore=cScoreTemp
  )

  saveRDS(BrierIPAList,'results/100complexBrier.rds')
  saveRDS(cScore,'results/100complexcidx.rds')
  pos=pos+1

}

```

```{r plotComplex}

BrierIPAList<-readRDS('results/100complexBrier.rds')

ipaSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
brierSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
ibsSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
ibsTable<-as.data.frame(matrix(data=NA,nrow=100,ncol=length(unique(BrierIPAList[[1]]$model))))
colnames(ibsTable)<-unique(BrierIPAList[[1]]$model)
count=1
for(sim in BrierIPAList){
  ipaSims<-cbind(ipaSims, sim$IPA)
  brierSims<-cbind(brierSims,sim$Brier)
  ibsSims<-cbind(ibsSims,sim$IBS)
  for(mod in colnames(ibsTable)){

    ibsTable[count,which(mod ==colnames(ibsTable))]<-sim$IBS[sim$model==mod & sim$time==max(sim$time)]

  }
  count<-count+1
}
ibsTable<-na.omit(ibsTable)
ibsOverall<-data.frame(mean=apply(ibsTable,2,mean),sd = apply(ibsTable,2,sd))
ibsOverall
ipaSummary<-statCalcSim(ipaSims)
brierSummary<-statCalcSim(brierSims)
ibsSummary<-statCalcSim(ibsSims)
ipa<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
brier<-plotPerformance(brierSummary,method="BS",iterations=iteration )
ibs<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )

#end_time-start_time
ipa+coord_cartesian(ylim=c(-2,1))

```

```{r optimDSM,eval=F}
klist<-readRDS("ComplexSimCV.rds")
test<-readRDS("ComplexSimTest.rds")
times<-seq(from=min(test$time)+0.000001,
           to=max(test$time)-0.000001,
           length.out = 25
)

paramTrials<-as.data.frame(tidyr::crossing(lr=c(0.001,0.01,0.1),
                                           drpt=c(0.25),
                                           layer1=c(10,25,50,100),
                                           layer2=c(10,25,50,100),
                                           bsize=c(floor(nrow(klist[[1]]$train)/100)),
                                           actv=c("relu","tanh"),
                                           alp=c(0,0.5,1),
                                           k=c(3,6,9),
                                           dist=c('weibull','lognormal'),
                                           #IPA=NA,
                                           #IPAse=NA,
                                           valLoss=NA,
                                           runtime=NA))

paramTrials<-paramTrials[sample(nrow(paramTrials), baselineTrials), ]

# paramTrials<-unique.data.frame(paramTrials)
#paramTrials<-readRDS("complexTrials.rds")
for(i in 1:nrow(paramTrials)){
  #paramTrials<-readRDS("complexTrials.rds")  
  tempRes<-c()
  tempVal<-c()
  runTime<-c()

  for(data in klist){

    train<-data$train
    val<-data$val

    lr=paramTrials[i,]$lr
    drpt=paramTrials[i,]$drpt
    layer1=paramTrials[i,]$layer1
    layer2=paramTrials[i,]$layer2
    bsize=paramTrials[i,]$bsize
    actv=paramTrials[i,]$actv
    alp=paramTrials[i,]$alp
    k=paramTrials[i,]$k
    dist=paramTrials[i,]$dist
    #=paramTrials[i,]$IPA
    #=paramTrials[i,]$valLoss

    currentTime<-Sys.time()

    source_python('src/optimDSM.py')
    dsmSurv<-dsmfitter(train,test,bsize,20,val,times,lr,actv,k,dist,layer1,layer2,tPos=3,ePos=4)
    if((sum(is.na(dsmSurv))==ncol(dsmSurv)*nrow(dsmSurv))){

      dsmSurv<-dsmfitter(train=train,fullTest=test,bsize=floor(nrow(train)/100),epochs=20,val=val,times=times,lr=lr,actv=actv,k,dist,layer1,layer2,tPos=3,ePos=4)
      py_run_string("del dsmfitter")
      rm(dsmfitter)
      next
    }
    tempVal<-append(tempVal,min(hitnnSurv$val_metrics$scores$loss$score))
    runTime<-append(runTime, Sys.time()-currentTime)
    currentTime<-Sys.time() 
  }
  rm(fitDeephit)

  #paramTrials$IPA[i]<-mean(tempRes)
  #paramTrials$IPAse[i]<-sd(tempRes)
  paramTrials$valLoss[i]<-mean(tempVal)
  paramTrials$runtime[i]<-mean(runTime)
  #saveRDS(paramTrials,"complexds.rds")
  print(paste("trial",i))
  print(mean(runTime))
  #print(as.character(paramTrials[i,]))
  #rm(cbnn,fit,annPreds,covars_input,covars_output)
  gc()
}

```

### 2023-02-08 analyses/optimizeMetabric.Rmd
checkpoint

--- 2023-02-08 analyses/optimizeComplex.Rmd
@@
 params:
-  epo: 200
+  epo: 2000
   patience: 5
-  iteration: 2
+  iteration: 30
   layer1: 12
@@
 ```{r}
+
 library(reticulate)
-
 ```
@@
 source("src/functions.R")
+eval_meta<-T
+#source_python('src/getSUP.py')
 library(reticulate)
-eval_support<-T
-#source_python('src/getSUP.py')
 source_python("src/cIndex.py")
@@
 #bsize=512
-min_delta =0# 10^-7
+min_delta = 0#10^-7
 
@@
 
-baselineTrials<-2
+baselineTrials<-5
 
@@
 
-```{r complexPrep,eval=T}
+```{r metaPrep,eval=T}
+source_python('src/getMetabricData.py')
 
-##############################
-###set up complex simulation
-##############################
-
-# Define a function returning the log cum hazard at time t
-logcumhaz <- function(t, x, betas, knots) {
-
-  # Obtain the basis terms for the spline-based log
-  # cumulative hazard (evaluated at time t)
-  basis <- flexsurv::basis(knots, log(t))
-
-  # Evaluate the log cumulative hazard under the
-  # Royston and Parmar specification
-  res <-betas[["gamma0"]] * basis[[1]] + 
-    betas[["gamma1"]] * basis[[2]] +
-    betas[["gamma2"]] * basis[[3]] +
-    betas[["gamma3"]] * basis[[4]] +
-    betas[["gamma4"]] * basis[[5]] +
-    betas[["z1"]] * x[["z1"]]+
-    betas[["z2"]] * x[["z2"]]+
-    betas[["z3"]] * x[["z3"]]+
-    betas[["term1"]] * x[["z1"]]*t+#*x[["z2"]]*
-    #betas[["term2"]] * x[["z1"]]*x[["z3"]]+
-    betas[["term3"]] * x[["z2"]]*(x[["z3"]])
-  #
-  # Return the log cumulative hazard at time t
-  res
-}
-url<-'app.txt'
-ds <- read.table(url, sep = '',header = TRUE)
-ds$years.at.risk=pmin(ds$agein80,ds$onset)
-brcancer$rectime<-brcancer$rectime/max(brcancer$rectime)
-true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1, data = brcancer, k = 3) #from flexsurv
-
-# Simulate the event times
-coefficients<-true_mod$coefficients
-coefficients<-c(coefficients,z1=0.05,z2=1.05,z3=1.05,term1=2.4,term3=1.05)#term2=1.05,
-
-#########################
-###Simulate data
-#########################
-cov<-data.frame(id=1:numSamp,
-                z1=c(rep(1,numSamp/2),rep(0,numSamp/2)),
-                z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
-                z3=c(rnorm(numSamp/2,0,0.5),rnorm(numSamp/2,10,0.5)))
-true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1,
-                                     data = brcancer, k = 3
-) #from flexsurv
-
-dat <- simsurv(betas = coefficients, # "true" parameter values
-               x = cov,                   # covariate data for 200 individuals
-               knots = true_mod$knots,    # knot locations for splines
-               logcumhazard = logcumhaz,  # definition of log cum hazard
-               maxt = 25,               # no right-censoring
-               interval = c(0,10000000)) # interval for root finding
-colnames(dat)<-c("id","time","status")
-samp<-sample(seq(1,nrow(dat),by=1), floor(nrow(dat)*(0.1))) #random censoring
-dat$status[samp]<-0
-# Merge the simulated event times onto covariate data frame
-data <- merge(cov, dat)
+metabric<-getMetabric()
+colnames(metabric)[ncol(metabric)-1]<-'time'
+colnames(metabric)[ncol(metabric)]<-'status'
+data<-as.data.frame(model.matrix(~.,metabric)[,-1])
+#data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
 covs<-colnames(data)
@@
 res = split(data, g)
-saveRDS(res$test,"ComplexSimTest.rds")
+saveRDS(res$test,"metaTest.rds")
 
@@
 }
-saveRDS(klist,"ComplexSimCV.rds")
+saveRDS(klist,"metaCV.rds")
 
@@
 
-klist<-readRDS("ComplexSimCV.rds")
+klist<-readRDS("metaCV.rds")
+
+###these did not work for complex or support
 # lr=c(0.001,0.01)
-# drpt=c(0.25,0.5)
-# layer1=c(50,75,100)
-# layer2=c(50,75)
+# drpt=c(0.1,0.25,.5)
+# layer1=c(100,200,300)
+# layer2=c(25,50,100,200)
 # bsize=c(floor(nrow(klist[[1]]$train)/100))
@@
 
-```{r hyperOpDsurv,eval=T}
+```{r hyperOpDsurv,eval=reoptimize}
 
@@
 
-  #paramTrials<-paramTrials[sample(nrow(paramTrials), baselineTrials), ]
+  #paramTrials<-paramTrials[sample(nrow(paramTrials), 2), ]
 
@@
 
-klist<-readRDS("ComplexSimCV.rds")
-test<-readRDS("ComplexSimTest.rds")
+klist<-readRDS("metaCV.rds")
+test<-readRDS("metaTest.rds")
   times<-seq(from=min(test$time)+0.000001,
@@
 bestDS<-optimDeepSurv(klist,test,times,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,bsize,actv)
+
 ```
 
-```{r optimDeepHit,eval=T}
+```{r optimDeepHit,eval=reoptimize}
 
@@
 
-```{r hyperopcbnn,eval=T}
+```{r hyperopcbnn,eval=reoptimize}
 
@@
 
-  # 
   # for (i in 1:length(klist)){
-  #   #train<-sampleCaseBase(data = as.data.frame(klist[[i]]$train),time="time",event="status",ratio=100)
+  #   train<-sampleCaseBase(data = as.data.frame(klist[[i]]$train),time="time",event="status",ratio=100)
   #   valcb<-sampleCaseBase(data = as.data.frame(klist[[i]]$val),time="time",event="status",ratio=100)
@@
   #                     as.matrix(valcb[,ncol(valcb)-1,drop=F]))
-  #   #valcbtensor[[1]][[2]]<-rep(train$offset[1],nrow(valcbtensor[[1]][[1]]))      
-  #   #klist[[i]]$train<-train
+  #   valcbtensor[[1]][[2]]<-rep(train$offset[1],nrow(valcbtensor[[1]][[1]]))      
+  #   klist[[i]]$train<-train
   #   klist[[i]]$val<-valcbtensor
@@
                                              layer2=layer2,
-                                             #bsize=,
                                              actv=actv,
@@
                                              runtime=NA))
-
   #paramTrials<-paramTrials[sample(nrow(paramTrials), 2), ]
@@
 
-klist<-readRDS("ComplexSimCV.rds")
-test<-readRDS("ComplexSimTest.rds")
+klist<-readRDS("metaCV.rds")
+test<-readRDS("metaTest.rds")
 
@@
 #cbnnPerformance<-cbnnPerformance[which.min(cbnnPerformance$valLoss),]
-
-```
-
-```{r}
-
 winners<-list(cbnn=cbnnPerformance,
@@
 
-saveRDS(winners,"complexOptimWins.rds")
+saveRDS(winners,"metaWins.rds")
+
 ```
@@
 
-winners<-readRDS("complexOptimWins.rds")
+winners<-readRDS("metaWins.rds")
 
-klist<-readRDS("ComplexSimCV.rds")
+klist<-readRDS("metaCV.rds")
 data<-rbind(klist[[1]]$train,klist[[1]]$val)
@@
 
-  fullTest<-readRDS("ComplexSimTest.rds")
+  fullTest<-readRDS("metaTest.rds")
   test<-fullTest[,-c(ncol(data))]
@@
   times<-tail(times, -1)
-
-  ###########################
-  ###casebase optimal model
-  ###########################
-  final_mod_cb_glm <- fitSmoothHazard(status~bs(time)-time+z1+z2+z3+z1:time+z2:z3,#+z1:z3
-                                      data = train,
-                                      time = "time",
-                                      event="status",
-                                      ratio = 100
-  )
-
-  summary(final_mod_cb_glm)
-  tglmAbsRisk<-absoluteRisk(final_mod_cb_glm,time=times,newdata=test,type="CI")
-  tglmAbsRisk<-as.data.frame(tglmAbsRisk)
-  rownames(tglmAbsRisk)<-tglmAbsRisk$time
-  trueGlmProper<- t(tglmAbsRisk[-1,-1])
-  class(trueGlmProper)<-c("tunnel",class(trueGlmProper))
 
@@
   actv=winners$ds$actv
-
   source_python('src/dsurv.py')
@@
   source_python('src/deephitter.py')
-  hitnnSurv=fitDeephit(train,fullTest,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
+  hitnnSurv=fitDeephit(train,fullTest,floor(nrow(train)/100),epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
   hitnnSurv<-t(hitnnSurv)
@@
   #############################
-  train<-sampleCaseBase(data = as.data.frame(train),time="time",event="status",ratio=100)
-  valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)
-  val<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
-                 as.matrix(valcb[,ncol(valcb),drop=F])),
-            as.matrix(valcb[,ncol(valcb)-1,drop=F]))
 
-  val[[1]][[2]]<-rep(train$offset[1],nrow(val[[1]][[1]]))   
   lr=winners$cbnn$lr
@@
 
-  # lr=0.001
-  # drpt=0.1
-  # layer1=50
-  # layer2=20
-  # #bsize=winners$cbnn$bsize
-  # actv='relu'
+  currentTime<-Sys.time() 
+  valcbtensor[[1]][[2]]<-rep(mod_cb_glm$offset[1],nrow(valcbtensor[[1]][[1]]))
 
-  currentTime<-Sys.time() 
-  # valcbtensor[[1]][[2]]<-rep(mod_cb_glm$offset[1],nrow(valcbtensor[[1]][[1]]))
   covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
@@
 
-  cbnn<-cbnnModel(features=colnames(train)[-ncol(train)],
+  cbnn<-cbnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                   feature_input = covars_input,
                   feature_output = covars_output,
-                  originalData = train,
-                  offset=train$offset,
+                  originalData = mod_cb_glm$data,
+                  offset=mod_cb_glm$offset,
                   timeVar = "time",
@@
   )
-
   fit<-fitSmoothHaz(cbnn,
                     epochs=epo,
-                    batch_size=c(floor(nrow(train)/100)),
+                    batch_size=c(floor(nrow(mod_cb_glm$data)/100)),
                     verbose=0,
@@
                     min_delta=min_delta,
-                    patience=patience,val=val)
+                    patience=patience,val=valcbtensor)
+
   annPreds<-aar(fit,
@@
                                   'DeepSurv'=cnnCleaned,'DeepHit'=deephitCleaned,
-                                  'CBNN_Poly'=annProperpoly,'optimal'=trueGlmProper),
+                                  'CBNN_Poly'=annProperpoly),
                              data =fullTest, 
@@
                  deephitCleaned,
-                 annProperpoly,
-                 trueGlmProper
+                 annProperpoly
   )
@@
   cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=length(riskList)+1)
-  et_train<-as.matrix(train)[,c(5,4)]
-  et_test<-as.matrix(fullTest[,c(5,4)])
+  et_train<-as.matrix(train)[,c(ncol(train),ncol(train)-1)]
+  et_test<-as.matrix(fullTest[,c(ncol(train),ncol(train)-1)])
   colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
-                          'cbnn','optimal')
+                          'cbnn')
   cScoreTemp[,1]<- tempTims
@@
   )
-
-  saveRDS(BrierIPAList,'results/100complexBrier.rds')
-  saveRDS(cScore,'results/100complexcidx.rds')
+  print(paste("current iter:",pos))
+  saveRDS(BrierIPAList,'results/100metaBrier.rds')
+  saveRDS(cScore,'results/100metacidx.rds')
   pos=pos+1
@@
 
-```{r plotComplex}
+```{r plotmeta}
 
-BrierIPAList<-readRDS('results/100complexBrier.rds')
+BrierIPAList<-readRDS('results/100metaBrier.rds')
 
@@
 
+ipa+coord_cartesian(ylim=c(-0.1,0.4))
 #end_time-start_time
-ipa+coord_cartesian(ylim=c(-2,1))
 
 ```
-
-```{r optimDSM,eval=F}
-klist<-readRDS("ComplexSimCV.rds")
-test<-readRDS("ComplexSimTest.rds")
-times<-seq(from=min(test$time)+0.000001,
-           to=max(test$time)-0.000001,
-           length.out = 25
-)
-
-paramTrials<-as.data.frame(tidyr::crossing(lr=c(0.001,0.01,0.1),
-                                           drpt=c(0.25),
-                                           layer1=c(10,25,50,100),
-                                           layer2=c(10,25,50,100),
-                                           bsize=c(floor(nrow(klist[[1]]$train)/100)),
-                                           actv=c("relu","tanh"),
-                                           alp=c(0,0.5,1),
-                                           k=c(3,6,9),
-                                           dist=c('weibull','lognormal'),
-                                           #IPA=NA,
-                                           #IPAse=NA,
-                                           valLoss=NA,
-                                           runtime=NA))
-
-paramTrials<-paramTrials[sample(nrow(paramTrials), baselineTrials), ]
-
-# paramTrials<-unique.data.frame(paramTrials)
-#paramTrials<-readRDS("complexTrials.rds")
-for(i in 1:nrow(paramTrials)){
-  #paramTrials<-readRDS("complexTrials.rds")  
-  tempRes<-c()
-  tempVal<-c()
-  runTime<-c()
-
-  for(data in klist){
-
-    train<-data$train
-    val<-data$val
-
-    lr=paramTrials[i,]$lr
-    drpt=paramTrials[i,]$drpt
-    layer1=paramTrials[i,]$layer1
-    layer2=paramTrials[i,]$layer2
-    bsize=paramTrials[i,]$bsize
-    actv=paramTrials[i,]$actv
-    alp=paramTrials[i,]$alp
-    k=paramTrials[i,]$k
-    dist=paramTrials[i,]$dist
-    #=paramTrials[i,]$IPA
-    #=paramTrials[i,]$valLoss
-
-    currentTime<-Sys.time()
-
-    source_python('src/optimDSM.py')
-    dsmSurv<-dsmfitter(train,test,bsize,20,val,times,lr,actv,k,dist,layer1,layer2,tPos=3,ePos=4)
-    if((sum(is.na(dsmSurv))==ncol(dsmSurv)*nrow(dsmSurv))){
-
-      dsmSurv<-dsmfitter(train=train,fullTest=test,bsize=floor(nrow(train)/100),epochs=20,val=val,times=times,lr=lr,actv=actv,k,dist,layer1,layer2,tPos=3,ePos=4)
-      py_run_string("del dsmfitter")
-      rm(dsmfitter)
-      next
-    }
-    tempVal<-append(tempVal,min(hitnnSurv$val_metrics$scores$loss$score))
-    runTime<-append(runTime, Sys.time()-currentTime)
-    currentTime<-Sys.time() 
-  }
-  rm(fitDeephit)
-
-  #paramTrials$IPA[i]<-mean(tempRes)
-  #paramTrials$IPAse[i]<-sd(tempRes)
-  paramTrials$valLoss[i]<-mean(tempVal)
-  paramTrials$runtime[i]<-mean(runTime)
-  #saveRDS(paramTrials,"complexds.rds")
-  print(paste("trial",i))
-  print(mean(runTime))
-  #print(as.character(paramTrials[i,]))
-  #rm(cbnn,fit,annPreds,covars_input,covars_output)
-  gc()
-}
-
-```

### 2023-02-08 analyses/optimizeSupport.Rmd
checkpoint

--- 2023-02-08 analyses/optimizeMetabric.Rmd
@@
   patience: 5
-  iteration: 30
+  iteration: 10
   layer1: 12
@@
 source("src/functions.R")
-eval_meta<-T
+eval_support<-T
 #source_python('src/getSUP.py')
-library(reticulate)
 source_python("src/cIndex.py")
@@
 #bsize=512
-min_delta = 0#10^-7
+min_delta = 0# 10^-7
 
@@
 
-```{r metaPrep,eval=T}
-source_python('src/getMetabricData.py')
+```{r supPrep,eval=T}
 
-metabric<-getMetabric()
-colnames(metabric)[ncol(metabric)-1]<-'time'
-colnames(metabric)[ncol(metabric)]<-'status'
-data<-as.data.frame(model.matrix(~.,metabric)[,-1])
+source_python('src/getSUP.py')
+data<- as.data.frame(do.call(cbind, getSupportPycox()))
+
+colnames(data)[ncol(data)-1]<-'time'
+colnames(data)[ncol(data)]<-'status'
+
+#for(i in c(2,5,6,7,16)){#3,4,
+#  data[, i]<-as.factor(data[, i])
+#}
+data<-as.data.frame(model.matrix(~.,data)[,-1])
 #data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
+
+covs<-colnames(data)
+names(data)[ncol(data)-2] <- paste("z", ncol(data)-2, sep="")
+names(data)[ncol(data)-1]<-"time"
+names(data)[ncol(data)]<-"status"
+
 covs<-colnames(data)
@@
 data$time<-data$time/toDiv
+#data$time<-data$time/max(data$time)
 spec = c(train = .85, test = .15)
@@
 res = split(data, g)
-saveRDS(res$test,"metaTest.rds")
+saveRDS(res$test,"supportTest.rds")
 
@@
 }
-saveRDS(klist,"metaCV.rds")
+saveRDS(klist,"supportCV.rds")
 
@@
 
-klist<-readRDS("metaCV.rds")
-
-###these did not work for complex or support
-# lr=c(0.001,0.01)
-# drpt=c(0.1,0.25,.5)
-# layer1=c(100,200,300)
-# layer2=c(25,50,100,200)
-# bsize=c(floor(nrow(klist[[1]]$train)/100))
-# actv=c("relu","tanh")
-# alp=c(0,0.5,1)
+klist<-readRDS("supportCV.rds")
 
@@
 
-klist<-readRDS("metaCV.rds")
-test<-readRDS("metaTest.rds")
+klist<-readRDS("supportCV.rds")
+test<-readRDS("supportTest.rds")
   times<-seq(from=min(test$time)+0.000001,
@@
       currentTime<-Sys.time() 
-      hitnnSurv=fitDeephit(train,val,floor(nrow(train)/100),epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
+      hitnnSurv=fitDeephit(train,val,floor(nrow(train)/100),
+                           epochs=epo,valida=val,patience=patience,
+                           min_delta=min_delta,drpt=drpt,
+                           lay1=layer1,lay2=layer2,
+                           lr=lr,actv=actv,alp=alp
+                           )
+
       tempVal<-append(tempVal,min(hitnnSurv$val_metrics$scores$loss$score))
@@
                                              runtime=NA))
-  #paramTrials<-paramTrials[sample(nrow(paramTrials), 2), ]
 
@@
       train<-normalizer(train,means,sds)
-      #train[,ncol(train)]<-data$train[,ncol(train)]
+     # train[,ncol(train)]<-data$train[,ncol(train)]
       val<-normalizer(val,means,sds)
@@
       val<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
-                     as.matrix(valcb[,ncol(valcb),drop=F])
-                     ),
-                as.matrix(valcb[,ncol(valcb)-1,drop=F])
-      )
+                     as.matrix(valcb[,ncol(valcb),drop=F])),
+                as.matrix(valcb[,ncol(valcb)-1,drop=F]))
+
       val[[1]][[2]]<-rep(train$offset[1],nrow(val[[1]][[1]]))
@@
 
-klist<-readRDS("metaCV.rds")
-test<-readRDS("metaTest.rds")
+klist<-readRDS("supportCV.rds")
+test<-readRDS("supportTest.rds")
 
@@
 
-saveRDS(winners,"metaWins.rds")
+saveRDS(winners,"supportWins.rds")
 
@@
 
-winners<-readRDS("metaWins.rds")
+winners<-readRDS("supportWins.rds")
 
-klist<-readRDS("metaCV.rds")
+klist<-readRDS("supportCV.rds")
 data<-rbind(klist[[1]]$train,klist[[1]]$val)
+#data$time<-data$time/max(data$time)
 
@@
 
-  fullTest<-readRDS("metaTest.rds")
+  fullTest<-readRDS("supportTest.rds")
   test<-fullTest[,-c(ncol(data))]
@@
   train<-normalizer(train,means,sds)
-  valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)#rbind(train,val)
-  valcb$offset<-0
-  valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
+
   times<-seq(from=min(test$time)+0.000001,
@@
 
-  lr=winners$ds$lr
+  lr=0.001#winners$ds$lr
   drpt=winners$ds$drpt
@@
   #############################
+          train<-sampleCaseBase(data = as.data.frame(train),time="time",event="status",ratio=100)
+      valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)
+      val<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
+                     as.matrix(valcb[,ncol(valcb),drop=F])),
+                as.matrix(valcb[,ncol(valcb)-1,drop=F]))
 
+      val[[1]][[2]]<-rep(train$offset[1],nrow(val[[1]][[1]]))   
   lr=winners$cbnn$lr
@@
 
+  # lr=0.001
+  # drpt=0.1
+  # layer1=50
+  # layer2=20
+  # #bsize=winners$cbnn$bsize
+  # actv='relu'
+
   currentTime<-Sys.time() 
-  valcbtensor[[1]][[2]]<-rep(mod_cb_glm$offset[1],nrow(valcbtensor[[1]][[1]]))
+  #valcbtensor[[1]][[2]]<-rep(mod_cb_glm$offset[1],nrow(valcbtensor[[1]][[1]]))
 
@@
 
-  cbnn<-cbnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
+      cbnn<-cbnnModel(features=colnames(train)[-ncol(train)],
                       feature_input = covars_input,
                       feature_output = covars_output,
-                  originalData = mod_cb_glm$data,
-                  offset=mod_cb_glm$offset,
+                      originalData = train,
+                      offset=train$offset,
                       timeVar = "time",
@@
       )
+
         fit<-fitSmoothHaz(cbnn,
                         epochs=epo,
-                    batch_size=c(floor(nrow(mod_cb_glm$data)/100)),
+                        batch_size=c(floor(nrow(train)/100)),
                         verbose=0,
@@
                         min_delta=min_delta,
-                    patience=patience,val=valcbtensor)
-
+                        patience=patience,val=val)
   annPreds<-aar(fit,
@@
 
+    ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
+  geom_line()#+coord_cartesian(ylim=c(-0.1,0.2))
   #######################
@@
   print(paste("current iter:",pos))
-  saveRDS(BrierIPAList,'results/100metaBrier.rds')
-  saveRDS(cScore,'results/100metacidx.rds')
+  saveRDS(BrierIPAList,'results/100supportBrier.rds')
+  saveRDS(cScore,'results/100supportcidx.rds')
   pos=pos+1
@@
 
-```{r plotmeta}
+```{r plotsupport}
 
-BrierIPAList<-readRDS('results/100metaBrier.rds')
+BrierIPAList<-readRDS('results/100supportBrier.rds')
 
@@
 
-ipa+coord_cartesian(ylim=c(-0.1,0.4))
+ipa+coord_cartesian(ylim=c(-0.1,0.2))
 #end_time-start_time

### 2023-02-08 analyses/plotterScript.Rmd
checkpoint

@@
   allModels<-data.frame(model=c("Null model","Cox_Lin",'CB_Logi','DeepSurv',
-                                'DeepHit','PMNN_Spline','PMNN_Poly','Optimal',
+                                'DeepHit','PMNN_Spline','CBNN_Poly','optimal',
                                 "dsm"),
@@
 
-```{r plotoldmort,fig.height=6,fig.width=6}
+```{r plotoldmort,fig.height=6,fig.width=6,eval=F}
 
@@
 
-```{r plotgbsg,fig.height=6,fig.width=6}
+```{r plotgbsg,fig.height=6,fig.width=6,eval=F}
 

### 2023-02-08 analyses/simpleSim.Rmd
checkpoint

@@
 params:
-  epo: 2000
+  epo: 100
   patience: 10
   iteration: 1
-  layer1: 50
-  layer2: 50
-  layer3: 25
-  layer4: 25
+  layer1: 100
+  layer2: 100
+  layer3: 50
+  layer4: 50
   drpt: 0.1
+  lr: 0.001
 ---
@@
 bsize=512
-min_delta = 0
+min_delta =10^-7
 epo=as.numeric(params$epo)
@@
 iteration<-as.numeric(params$iteration)
-layer1<-as.numeric(params$layer1)
-layer2<-as.numeric(params$layer2)
+layer1<-3*15#as.numeric(params$layer1)
+layer2<-3*15#as.numeric(params$layer2)
 layer3<-as.numeric(params$layer3)
@@
 drpt<-as.numeric(params$drpt)
+lr=as.numeric(params$lr)
+paramTrials<-readRDS("complexTrials.rds") 
+lr<-paramTrials[which.max(paramTrials$integBrier),]$lr
+layer1<-paramTrials[which.max(paramTrials$integBrier),]$layer1
+layer2<-paramTrials[which.max(paramTrials$integBrier),]$layer2
+layer3<-3*5#as.numeric(params$layer3)
+layer4<-as.numeric(params$layer4)
+drpt<-paramTrials[which.max(paramTrials$integBrier),]$drpt
+bsize=paramTrials[which.max(paramTrials$integBrier),]$bsize
+activ<-paramTrials[which.max(paramTrials$integBrier),]$actv
+
+lr<-paramTrials[which.min(paramTrials$valLoss),]$lr
+layer1<-paramTrials[which.min(paramTrials$valLoss),]$layer1
+layer2<-paramTrials[which.min(paramTrials$valLoss),]$layer2
+layer3<-3*5#as.numeric(params$layer3)
+layer4<-as.numeric(params$layer4)
+drpt<-paramTrials[which.min(paramTrials$valLoss),]$drpt
+bsize=paramTrials[which.min(paramTrials$valLoss),]$bsize
+activ<-paramTrials[which.min(paramTrials$valLoss),]$actv
 
@@
                   z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
-                  z3=c(rnorm(numSamp/2,-3,0.5),rnorm(numSamp/2,8,0.5))
+                  z3=c(rnorm(numSamp/2,0,0.5),rnorm(numSamp/2,10,0.5))
   )
@@
                  lambdas =.05,
-                 betas = c(z1 = 1,z2=1,z3=1),
+                 betas = c(z1 = 0.05,z2=1.5,z3=0.5),
                  x = cov,
@@
   train<-data
+
+  sds<-sapply(train, function(x) (sd(x)))
+means<-sapply(train, function(x) (mean(x)))
+test<-normalizer(res$test,means,sds)[,-c(ncol(data))]
+fullTest<-normalizer(res$test,means,sds)
+val<-normalizer(res$validate,means,sds)
+
+train<-normalizer(train,means,sds)
   #test<-data[-samp,-c(ncol(data))]
@@
   source_python('src/dsmScript.py')
-  dsmSurv<-dsmfitter(train=train,fullTest=fullTest,bsize=bsize,epochs=epo,val=val,times=times,tPos=ncol(train)-2,ePos=ncol(train)-1)
+  dsmSurv<-dsmfitter(train=train,fullTest=test,bsize=floor(nrow(train)/100),epochs=20,val=val,times=times,tPos=ncol(train)-2,ePos=ncol(train)-1,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4,lr=lr)
   if((sum(is.na(dsmSurv))==ncol(dsmSurv)*nrow(dsmSurv))){
 
-    dsmSurv<-dsmfitter(train=train,fullTest=fullTest,bsize=bsize,epochs=epo,val=val,times=times)
+  dsmSurv<-dsmfitter(train=train,fullTest=fullTest,bsize=floor(nrow(train)/100),epochs=epo,val=val,times=times,tPos=ncol(train)-2,ePos=ncol(train)-1,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4,lr=lr)
     py_run_string("del dsmfitter")
@@
 
-  coxnnsurv=fitDeepSurv(tiebreakTrain,fullTest,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4)
+  coxnnsurv=fitDeepSurv(tiebreakTrain,fullTest,floor(nrow(train)/100),epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4,lr=lr)
   coxnnsurv<-t(coxnnsurv)
@@
     source_python('src/deephitter.py')
-  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4)
+  hitnnSurv=fitDeephit(train,fullTest,floor(nrow(train)/100),epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4,lr=lr)
   hitnnSurv<-t(hitnnSurv)
@@
   covars_output<-covars_input%>% #layer_batch_normalization()%>%
-    layer_dense(units=layer1,use_bias = T,activation = 'tanh')%>%
+    layer_dense(units=layer1,use_bias = T,activation = activ)%>%
     layer_dropout(drpt)%>%
-    layer_dense(units=layer2,use_bias = T,activation = 'tanh')%>%
+    layer_dense(units=layer2,use_bias = T,activation = activ)%>%
     layer_dropout(drpt)%>%
-    layer_dense(units=layer3,use_bias = T,activation = 'tanh')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer4,use_bias = T,activation = 'tanh')%>%
-    layer_dropout(drpt)%>%
+    #layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
+    #layer_dropout(drpt)%>%
+    #layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
+    #layer_dropout(drpt)%>%
     layer_dense(units=1,use_bias = T)
@@
                           epochs=epo,
-                          batch_size=bsize,
+                          batch_size=floor(nrow(mod_cb_glm$data)/100),
                           verbose=0,
@@
                     epochs=epo,
-                    batch_size=bsize,
+                    batch_size=floor(nrow(mod_cb_glm$data)/100),
                     verbose=0,
@@
   ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
-  geom_line()+coord_cartesian(ylim=c(-1,1))
+  geom_line()+coord_cartesian(ylim=c(-.5,.5))
 

### 2023-02-08 analyses/supportStudy.Rmd
checkpoint

@@
   epo: 2000
-  patience: 10
-  iteration: 1
+  patience: 5
+  iteration: 2
   layer1: 50
-  layer2: 50
-  layer3: 25
-  layer4: 25
+  layer2: 25
+  layer3: 100
+  layer4: 100
   drpt: 0.1
+  lr: 0.001
 ---
@@
 ```{r,eval=T}
-
+set.seed(1)
 library(reticulate)
@@
 ```{r setup, include=T}
-set.seed(1)
+
 knitr::opts_chunk$set(echo = TRUE)
@@
 bsize=512
-min_delta = 0#10^-7
+min_delta = 10^-7
 epo=as.numeric(params$epo)
@@
 iteration<-as.numeric(params$iteration)
-layer1<-as.numeric(params$layer1)
-layer2<-as.numeric(params$layer2)
-layer3<-as.numeric(params$layer3)
-layer4<-as.numeric(params$layer4)
+layer1<-10#as.numeric(params$layer1)
+layer2<-5#as.numeric(params$layer2)
+layer3<-14*3#as.numeric(params$layer3)
+layer4<-14*2#as.numeric(params$layer4)
 drpt<-as.numeric(params$drpt)
+lr=as.numeric(params$lr)
 
@@
 data<- as.data.frame(do.call(cbind, getSupportPycox()))
+
 colnames(data)[ncol(data)-1]<-'time'
 colnames(data)[ncol(data)]<-'status'
+
+data$time<-data$time/max(data$time)
+for(i in c(2,5,6,7)){#3,4,
+data[, i]<-as.factor(data[, i])
+}
 data<-as.data.frame(model.matrix(~.,data)[,-1])
-data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
+#data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
 
 covs<-colnames(data)
+names(data)[ncol(data)-2] <- paste("z", ncol(data)-2, sep="")
+names(data)[ncol(data)-1]<-"time"
+names(data)[ncol(data)]<-"status"
+  data<-data[,-1]
+  toDiv<-max(data$time)
 
+  data$time<-data$time/toDiv
 #split training/testing
 
- spec = c(train = .7, test = .1, validate = .2)
+ spec = c(train = .7, test = .15, validate = .15)
 
@@
 
-valcb<-sampleCaseBase(data = val,time="time",event="status",ratio=100)
-valcb$offset<-0
-valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
-
   times<-seq(from=min(test$time)+0.000001,
@@
 val<-normalizer(res$validate,means,sds)
+valcb<-sampleCaseBase(data = val,time="time",event="status",ratio=100)
+
+valcb$offset<-0
+valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
 
@@
   #need to break ties
-  tiebreakTrain<-train
 
@@
 
-  coxnnsurv=fitDeepSurv(tiebreakTrain,fullTest,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4)
+  coxnnsurv=fitDeepSurv(train,fullTest,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv="relu")
   coxnnsurv<-t(coxnnsurv)
@@
     source_python('src/deephitter.py')
-  hitnnSurv=fitDeephit(train,fullTest,bsize,epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lay3=layer3,lay4=layer4)
+  hitnnSurv=fitDeephit(train,fullTest,floor(nrow(train)/100),epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv='relu',alp=0)
   hitnnSurv<-t(hitnnSurv)
@@
   # rm(fitDeephit)
+  # # 
+  # ############################
+  # ###cbnnsplines
+  # ############################  
+  #   valcbtensor[[1]][[2]]<-rep(mod_cb_glm$offset[1],nrow(valcbtensor[[1]][[1]]))
+  # covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2+3),
+  #                           name = 'main_input'
+  # )
+  # covars_output<-covars_input%>% #layer_batch_normalization()%>%
+  #   layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
+  #   layer_dropout(drpt)%>%
+  #   layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
+  #   layer_dropout(drpt)%>%
+  #   #layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
+  #   #layer_dropout(drpt)%>%
+  #   #layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
+  #   #layer_dropout(drpt)%>%
+  #   layer_dense(units=1,use_bias = T)
   # 
-  ############################
-  ###PMNNsplines
-  ############################  
-  covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2+3),
-                            name = 'main_input'
-  )
-  covars_output<-covars_input%>% #layer_batch_normalization()%>%
-    layer_dense(units=layer1,use_bias = T,activation = 'tanh')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer2,use_bias = T,activation = 'tanh')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer3,use_bias = T,activation = 'tanh')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer4,use_bias = T,activation = 'tanh')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=1,use_bias = T)
-
-  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
-                  feature_input = covars_input,
-                  feature_output = covars_output,
-                  originalData = mod_cb_glm$data,
-                  offset=mod_cb_glm$offset,
-                  timeVar = "time",
-                  eventVar= "status",optimizer=optimizer_adam(learning_rate = 0.0001)
-  )
-
-  fit<-fitSmoothHazSpline(pmnn,
-                          epochs=epo,
-                          batch_size=bsize,
-                          verbose=0,
-                          monitor="val_loss",
-                          #val_split=0.2,
-                          min_delta=min_delta,
-                          patience=patience,val=valcbtensor)
-
-   annPreds<-aarSplines(fit,
-                       times=times,
-                       x_test=test
-  )
-  rownames(annPreds)<-annPreds[,1]
-  annProper<- t(annPreds[,-1])
-  class(annProper)<-c("tunnel",class(annProper))
+  # cbnn<-cbnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
+  #                 feature_input = covars_input,
+  #                 feature_output = covars_output,
+  #                 originalData = mod_cb_glm$data,
+  #                 offset=mod_cb_glm$offset,
+  #                 timeVar = "time",
+  #                 eventVar= "status",optimizer=optimizer_adam(learning_rate = lr)
+  # )
+  # 
+  # 
+  # 
+  # fit<-fitSmoothHazSpline(cbnn,
+  #                         epochs=epo,
+  #                         batch_size=floor(nrow(mod_cb_glm$data)/100),
+  #                         verbose=0,
+  #                         monitor="val_loss",
+  #                         #val_split=0.2,
+  #                         min_delta=min_delta,
+  #                         patience=patience,val=valcbtensor)
+  # 
+  #  annPreds<-aarSplines(fit,
+  #                      times=times,
+  #                      x_test=test
+  # )
+  # rownames(annPreds)<-annPreds[,1]
+  # annProper<- t(annPreds[,-1])
+  # class(annProper)<-c("tunnel",class(annProper))
+  # 
+  # 
+  #
 
@@
   ######################
-  ###PMNN  
+  ###cbnn  
   ######################
@@
   covars_output<-covars_input%>% #layer_batch_normalization()%>%
-    layer_dense(units=layer1,use_bias = T,activation = 'tanh')%>%
+    #layer_dropout(drpt)%>%
+    layer_dense(units=layer1,use_bias = T,activation = 'relu')%>%
     layer_dropout(drpt)%>%
-    layer_dense(units=layer2,use_bias = T,activation = 'tanh')%>%
+    layer_dense(units=layer2,use_bias = T,activation = 'relu')%>%
     layer_dropout(drpt)%>%
-    layer_dense(units=layer3,use_bias = T,activation = 'tanh')%>%
-    layer_dropout(drpt)%>%
-    layer_dense(units=layer4,use_bias = T,activation = 'tanh')%>%
-    layer_dropout(drpt)%>%
+    #layer_dense(units=layer3,use_bias = T,activation = 'relu')%>%
+    #layer_dropout(drpt)%>%
+    #layer_dense(units=layer4,use_bias = T,activation = 'relu')%>%
+    #layer_dropout(drpt)%>%
     layer_dense(units=1,use_bias = T)
 
-  pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
+  cbnn<-cbnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                   feature_input = covars_input,
@@
                   timeVar = "time",
-                  eventVar= "status",optimizer=optimizer_adam(learning_rate = 0.0001)
+                  eventVar= "status",optimizer=optimizer_adam(learning_rate = lr)
   )
 
-  fit<-fitSmoothHaz(pmnn,
+  fit<-fitSmoothHaz(cbnn,
                     epochs=epo,
-                    batch_size=bsize,
+                    batch_size=floor(nrow(mod_cb_glm$data)/100),
                     verbose=0,
                     monitor="val_loss",
-                    val_split=0.2,
+                    #val_split=0.2,
                     min_delta=min_delta,
@@
                                   'DeepHit'=deephitCleaned,
-                                  'PMNN_Spline'=annProper,
-                                  'PMNN_Poly'=annProperpoly),
+                                  #'cbnn_Spline'=annProper,
+                                  'cbnn_Poly'=annProperpoly),
                              data =fullTest, 
@@
   geom_line()+coord_cartesian(ylim=c(-0.1,0.2))
+     rm(cbnn,fit,annPreds,covars_input,covars_output)
   ######################
@@
   colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
-                          'PMNN_Spline','PMNN_Poly')#,'dsm')
+                          'cbnn_Spline','cbnn_Poly')#,'dsm')
   cScoreTemp[,1]<- tempTims
@@
 
-  saveRDS(BrierIPAList,'results/100supBrier.rds')
-  saveRDS(cScore,'results/100supcidx.rds')
+  #saveRDS(BrierIPAList,'results/100supBrier.rds')
+  #saveRDS(cScore,'results/100supcidx.rds')
   pos<-pos+1
+  gc()
 }

### 2023-02-08 analyses/dsm/dsm_api.py

@@ -52,15 +52,15 @@ class DSMBase():
   """Base Class for all DSM models"""
 
   def __init__(self, k=3, layers=None, distribution="Weibull",
-               temp=1000., discount=1.0):
+               temp=1000., discount=1.0,activation="relu"):
     self.k = k
     self.layers = layers
     self.dist = distribution
     self.temp = temp
     self.discount = discount
     self.fitted = False
-
-  def _gen_torch_model(self, inputdim, optimizer, risks):
+    
+  def _gen_torch_model(self, inputdim, optimizer, risks,activation):
     """Helper function to return a torch model."""
     return DeepSurvivalMachinesTorch(inputdim,
                                      k=self.k,
@@ -69,7 +69,7 @@ class DSMBase():
                                      temp=self.temp,
                                      discount=self.discount,
                                      optimizer=optimizer,
-                                     risks=risks)
+                                     risks=risks,activation=activation)
 
   def fit(self, x, t, e, vsize=0.15, val_data=None,
           iters=1, learning_rate=1e-3, batch_size=100,
@@ -121,7 +121,7 @@ class DSMBase():
       inputdim = x_train.shape[-1]
 
     maxrisk = int(np.nanmax(e_train.cpu().numpy()))
-    model = self._gen_torch_model(inputdim, optimizer, risks=maxrisk)
+    model = self._gen_torch_model(inputdim, optimizer, risks=maxrisk,activation=activation)
     model, _ = train_dsm(model,
                          x_train, t_train, e_train,
                          x_val, t_val, e_val,

### 2023-02-08 analyses/dsm/dsm_torch.py

@@ -75,18 +75,21 @@ def create_representation(inputdim, layers, activation):
 
   if activation == 'ReLU6':
     act = nn.ReLU6()
-  elif activation == 'ReLU':
+  elif activation == 'relu':
     act = nn.ReLU()
   elif activation == 'SeLU':
     act = nn.SELU()
+  elif activation == 'tanh':
+    act = nn.Tanh()
+  
 
   modules = []
   prevdim = inputdim
 
   for hidden in layers:
     modules.append(nn.Linear(prevdim, hidden, bias=True))
-    modules.append(nn.Tanh())
-    modules.append(nn.Dropout(0.1))
+    modules.append(act)
+    modules.append(nn.Dropout(0.25))
     prevdim = hidden
 
   return nn.Sequential(*modules)
@@ -196,7 +199,7 @@ class DeepSurvivalMachinesTorch(nn.Module):
 
     self._init_dsm_layers(lastdim)
 
-    self.embedding = create_representation(inputdim, layers, 'ReLU6')
+    self.embedding = create_representation(inputdim, layers, activation)
 
 
   def forward(self, x, risk='1'):

### 2023-02-08 analyses/dsm/utilities.py

@@ -187,7 +187,7 @@ def train_dsm(model,
     #
     if oldcost -costs[-1] < 10**-7:
     #if costs[-1] >= oldcost:
-      if patience == 10:
+      if patience == 5:
         minm = np.argmin(costs)
         model.load_state_dict(dics[minm])
 

### 2023-02-08 analyses/src/deephitter.py

@@ -13,10 +13,10 @@ from pycox.evaluation import EvalSurv
 
 
 
-
-def fitDeephit(train,fullTest,bsize,epochs,valida,ti='time',ev='status',patience=10,min_delta=10**-7,drpt=0.5,lay1=50,lay2=50,lay3=25,lay4=25):#,epoch,bsize,tPos,ePos):
+def fitDeephit(train,fullTest,bsize,epochs,valida,patience,min_delta,drpt,lay1,lay2,lr,actv,alp):#,epoch,bsize,tPos,ePos):
+  ti='time'
+  ev='status'
   n= np.shape(train)[0]
-  vl_size = int(n*0.20)
   df_val = valida#train.sample(frac=0.2)
   df_train = train#.drop(df_val.index)
   df_test = fullTest
@@ -53,40 +53,38 @@ def fitDeephit(train,fullTest,bsize,epochs,valida,ti='time',ev='status',patience
   batch_norm = False
   dropout = drpt
   
-  #net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout)
- 
-  net = torch.nn.Sequential(
-     #torch.nn.BatchNorm1d(in_features),
-     torch.nn.Linear(in_features,int(lay1)),
-     #torch.nn.ReLU(),
-     torch.nn.Tanh(),
-     torch.nn.Dropout(float(dropout)),
-     
-     torch.nn.Linear(int(lay1), int(lay2)),
-     #torch.nn.ReLU(),
-     torch.nn.Tanh(),
-     torch.nn.Dropout(float(dropout)),
-     
-     torch.nn.Linear(int(lay2), int(lay3)),
-     #torch.nn.ReLU(),
-     torch.nn.Tanh(),
-     torch.nn.Dropout(float(dropout)),     
-
-     torch.nn.Linear(int(lay3), int(lay4)),
-     #torch.nn.ReLU(),
-     torch.nn.Tanh(),
-     torch.nn.Dropout(float(dropout)),
-     torch.nn.Linear(int(lay4), out_features))
-  model = DeepHitSingle(net, tt.optim.Adam, alpha=0.5, sigma=0.1, duration_index=labtrans.cuts)
+  #net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout,activation="selu")
+  if actv=='relu':
+    net = torch.nn.Sequential(
+       torch.nn.Linear(in_features,int(lay1)),
+       torch.nn.ReLU(),
[…]

### 2023-02-08 analyses/src/dsmScript.py

@@ -7,7 +7,7 @@ from dsm import DeepSurvivalMachines
 ##dsm/dsm_torch.py, line 89 dropout(0.5) layer was added
 
 
-def dsmfitter(train,fullTest,bsize,epochs,val,times,tPos=3,ePos=4,lay1=50,lay2=50,lay3=25,lay4=25):
+def dsmfitter(train,fullTest,bsize,epochs,val,times,tPos=3,ePos=4,lay1=50,lay2=50,lay3=25,lay4=25,lr=0.001):
   
   ###############################
   ###Prepare data
@@ -46,8 +46,8 @@ def dsmfitter(train,fullTest,bsize,epochs,val,times,tPos=3,ePos=4,lay1=50,lay2=5
   #I fix the parameter grid
   param_grid = {'k' : [6],
                 'distribution' : ['Weibull'],
-                'learning_rate' : [ 0.0001],
-                'layers' : [ [int(lay1),int(lay2),int(lay3),int(lay4)] ]
+                'learning_rate' : [ lr],
+                'layers' : [ [int(lay1),int(lay2)] ]
                }
   params = ParameterGrid(param_grid)
   

### 2023-02-08 analyses/src/dsurv.py

@@ -12,9 +12,10 @@ from pycox.evaluation import EvalSurv
 
 
 
-def fitDeepSurv(train,fullTest,bsize,epochs,valida,ti='time',ev='status',patience=10,min_delta=10**-7,drpt=0.5,lay1=50,lay2=50,lay3=25,lay4=25):#,epoch,bsize,tPos,ePos):
+def fitDeepSurv(train,fullTest,bsize,epochs,valida,patience,min_delta,drpt,lay1,lay2,lr,actv):#,epoch,bsize,tPos,ePos):
+  ti='time'
+  ev='status'
   n= np.shape(train)[0]
-  vl_size = int(n*0.20)
   df_val = valida#train.sample(frac=0.2)
   df_train = train#.drop(df_val.index)
   df_test = fullTest
@@ -38,46 +39,47 @@ def fitDeepSurv(train,fullTest,bsize,epochs,valida,ti='time',ev='status',patienc
   
   
   in_features = x_train.shape[1]
-  num_nodes = [int(lay1),int(lay2)]
+  #num_nodes = [int(lay1),int(lay2)]
   out_features = 1
-  batch_norm = True
+  #batch_norm = True
   dropout = drpt
   
   #net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout,activation="selu")
-  net = torch.nn.Sequential(
-     #torch.nn.BatchNorm1d(in_features),
-     torch.nn.Linear(in_features,int(lay1)),
-     #torch.nn.ReLU(),
-     torch.nn.Tanh(),
-     torch.nn.Dropout(float(dropout)),
-     
-     torch.nn.Linear(int(lay1), int(lay2)),
-     #torch.nn.ReLU(),
-     torch.nn.Tanh(),
-     torch.nn.Dropout(float(dropout)),
-     
-     torch.nn.Linear(int(lay2), int(lay3)),
-     #torch.nn.ReLU(),
-     torch.nn.Tanh(),
-     torch.nn.Dropout(float(dropout)),     
-
-     torch.nn.Linear(int(lay3), int(lay4)),
-     #torch.nn.ReLU(),
-     torch.nn.Tanh(),
-     torch.nn.Dropout(float(dropout)),
-     torch.nn.Linear(int(lay4), out_features))
-
+  if actv=='relu':
+    net = torch.nn.Sequential(
+       torch.nn.Linear(in_features,int(lay1)),
+       torch.nn.ReLU(),
+       torch.nn.Dropout(float(dropout)),
[…]

### 2023-02-08 analyses/src/functions.R

@@ -45,13 +45,15 @@ pyProcess<-function(predictions,times=times,apr="constant"){
 calcCidx<-function(et_train,et_test,risk,times){
   #et_train <- et_train[order(et_train[,2]),]
   #et_test <- et_test[order(et_test[,2]),]
-  
+  #tims<-times
   tims<-head(times, -1)
   tims<-tail(tims, -1)
   cindexes<-matrix(NA,nrow=length(tims),ncol=2)
   cindexes[,1]<-tims
   
   for( i in 1:(length(tims))){#metabric needs to skip first few
+    
+
     cindexes[i,2]<-cidx(et_train,et_test,risk[,i,drop=F])
   }
   return(cindexes)
@@ -183,11 +185,11 @@ predictRisk.tunnel <- function(object, newdata, times, cause, ...){
 
 
 ##########################################################
-###PMNN
+###CBNN
 #########################################################
 
 
-pmnnModel<-function(features, feature_input, feature_output, originalData=originalData,offset, timeVar,eventVar, ratio=100, compRisk=FALSE,censored.indicator=0,optimizer=optimizer_adam(learning_rate = 0.001,decay=10^-7)){
+cbnnModel<-function(features, feature_input, feature_output, originalData=originalData,offset, timeVar,eventVar, ratio=100, compRisk=FALSE,censored.indicator=0,optimizer=optimizer_adam(learning_rate = 0.001,decay=10^-7)){
   if(class(originalData)[1]=="cbData"){
     data<-originalData
   }else{
@@ -226,17 +228,17 @@ pmnnModel<-function(features, feature_input, feature_output, originalData=origin
 
 
 
-fitSmoothHaz<-function(pmnn,epochs=20000,batch_size=500,verbose=0,monitor="val_loss",val_split=0.2,min_delta=10^-7,patience=10,val){
-  #pmnn$casebaseData
+fitSmoothHaz<-function(cbnn,epochs=2000,batch_size=500,verbose=0,monitor="val_loss",val_split=0.2,min_delta=10^-7,patience=10,val){
+  #cbnn$casebaseData
   
-  offset<-as.matrix(pmnn$offset)
-  #pmnn$casebaseData<-as.matrix(pmnn$casebaseData[,pmnn$features])
[…]

### 2023-02-08 analyses/src/optimDH.py

@@ -1,0 +1,103 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+# For preprocessing
+from sklearn.preprocessing import StandardScaler
+from sklearn_pandas import DataFrameMapper 
+
+import torch # For building the networks 
+import torchtuples as tt # Some useful functions
+
+from pycox.models import DeepHitSingle
+from pycox.evaluation import EvalSurv
+
+
+
+
+def fitDeephit(train,fullTest,bsize,epochs,valida,patience,min_delta,drpt,lay1,lay2,lr,actv,alp):#,epoch,bsize,tPos,ePos):
+
+  ti='time'
+  ev='status'
+  n= np.shape(train)[0]
+  df_val = valida#train.sample(frac=0.2)
+  df_train = train#.drop(df_val.index)
+  df_test = fullTest
+  varNames = list(df_train.columns)
+  unwanted = {ti, ev}
+  varNames = [e for e in varNames if e not in unwanted]
+  leave = [(col, None) for col in varNames]
+
+  x_mapper = DataFrameMapper(leave)
+
+  x_train = x_mapper.fit_transform(df_train).astype('float32')
+  x_val = x_mapper.transform(df_val).astype('float32')
+  x_test = x_mapper.transform(df_test).astype('float32')
+  
+
+
+  ########
+  #prepare labels
+  #######
+  num_durations = 10
+  labtrans = DeepHitSingle.label_transform(num_durations)
+  
+  get_target = lambda df: (df[ti].values, df[ev].values)
+  #print(get_target(df_train[ti]))
+  y_train = labtrans.fit_transform(*get_target(df_train))
+  y_val = labtrans.transform(*get_target(df_val))
+  
+ 
+  
+  
+  train = (x_train, y_train)
+  val = (x_val, y_val)
+  durations_test, events_test = get_target(df_test)
+  
+  ########
+  #prepare model
+  ########
+  in_features = x_train.shape[1]
+  #num_nodes = [50,50,25,25]
+  out_features = labtrans.out_features
+  batch_norm = False
+  dropout = drpt
+  
+  #net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout,activation="selu")
+  if actv=='relu':
+    net = torch.nn.Sequential(
+       torch.nn.Linear(in_features,int(lay1)),
+       torch.nn.ReLU(),
[…]

### 2023-02-08 analyses/src/optimDSM.py

@@ -1,0 +1,71 @@
+import numpy as np
+import random
+#from dsm import datasets
+from sklearn.model_selection import ParameterGrid
+from dsm.dsm_api import DeepSurvivalMachines
+##in dsm/utilities.py, we set early stopping patience and min_delta manually
+##dsm/dsm_torch.py, line 89 dropout(0.5) layer was added
+
+
+def dsmfitter(train,fullTest,bsize,epochs,val,times,lr,actv,k,dist,lay1,lay2,tPos=3,ePos=4):
+  
+  ###############################
+  ###Prepare data
+  ###############################
+  tPos=int(tPos)
+  ePos=int(ePos)
+  train2=train.to_numpy()#@np.random.shuffle(train) 
+  #np.random.shuffle(train2) 
+  n= np.shape(train2)[0]
+  #vl_size = int(n*0.80)
+
+  
+  t_train = np.asarray( train2[:,tPos])
+  e_train = np.asarray( train2[:,ePos]).astype(int)
+  x_train = train2
+  x_train = np.delete(x_train, [tPos,ePos], 1)
+  val=val.to_numpy()
+  t_val = np.asarray( val[:,tPos])
+  e_val = np.asarray( val[:,ePos]).astype(int)
+  x_val = val
+  x_val = np.delete(x_val, [tPos,ePos], 1)
+  #print("hi")
+  test=fullTest.to_numpy()
+  t_test = np.asarray( test[:,tPos])
+  e_test = np.asarray( test[:,ePos]).astype(int)
+  x_test = np.delete(test, [tPos,ePos], 1)
+  
+  horizons = np.arange(0, 1, 0.01).tolist()
+  times= np.quantile(t_test[e_test==1], horizons).tolist()
+  
+  
+
+  ###############################
+  ###Prepare data
+  ###############################
+  #I fix the parameter grid
+  param_grid = {'k' : [int(k)],
+                'distribution' : [dist],
+                'learning_rate' : [ int(lr)],
+                'layers' : [ [int(lay1),int(lay2)] ]
+               }
+  params = ParameterGrid(param_grid)
+  
+  #note that the DeepSurvivalMachines function has been modified: 
+  #it now introduces a dropout layer after each feedforward dense layer.
+  # ### Model Training and Selection
+  models = []
+  for param in params:
+      model = DeepSurvivalMachines(k = param['k'],
+                                   distribution = param['distribution'],
+                                   layers = param['layers'],
+                                   activation=actv)
[…]

### 2023-02-08 analyses/src/optimDSurv.py

@@ -1,0 +1,88 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from sklearn.preprocessing import StandardScaler
+from sklearn_pandas import DataFrameMapper
+
+import torch
+import torchtuples as tt
+
+from pycox.datasets import metabric
+from pycox.models import CoxPH
+from pycox.evaluation import EvalSurv
+
+
+
+def fitDeepSurv(train,fullTest,bsize,epochs,valida,patience,min_delta,drpt,lay1,lay2,lr,actv):#,epoch,bsize,tPos,ePos):
+  ti='time'
+  ev='status'
+  n= np.shape(train)[0]
+  df_val = valida#train.sample(frac=0.2)
+  df_train = train#.drop(df_val.index)
+  df_test = fullTest
+  varNames = list(df_train.columns)
+  unwanted = {ti, ev}
+  varNames = [e for e in varNames if e not in unwanted]
+  leave = [(col, None) for col in varNames]
+
+  x_mapper = DataFrameMapper(leave)
+  
+  x_train = x_mapper.fit_transform(df_train).astype('float32')
+  x_val = x_mapper.transform(df_val).astype('float32')
+  x_test = x_mapper.transform(df_test).astype('float32')
+  get_target = lambda df: (df[ti].values, df[ev].values)
+  y_train = get_target(df_train)
+  y_val = get_target(df_val)
+  #y_train[0]=y_train[0]+(np.random.uniform(0,1,len(y_train[0]))/10000000000000000000)
+  #y_val[0]=y_val[0]+(np.random.uniform(0,1,len(y_val[0]))/10000000000000000000)
+  durations_test, events_test = get_target(df_test)
+  val = x_val, y_val
+  
+  
+  in_features = x_train.shape[1]
+  #num_nodes = [int(lay1),int(lay2)]
+  out_features = 1
+  #batch_norm = True
+  dropout = drpt
+  
+  #net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout,activation="selu")
+  if actv=='relu':
+    net = torch.nn.Sequential(
+       torch.nn.Linear(in_features,int(lay1)),
+       torch.nn.ReLU(),
+       torch.nn.Dropout(float(dropout)),
+       
+       torch.nn.Linear(int(lay1), int(lay2)),
+       torch.nn.ReLU(),
+       #torch.nn.Tanh(),
[…]

### 2023-02-08 analyses/src/pmnnFunctions.R

@@ -1,0 +1,482 @@
+
+statCalcSim<-function(metricSims){
+  toRemove <- which(apply(metricSims, 1, function(x){any(is.na(x))})==T)
+  if(length(toRemove)>0){metricSims<-metricSims[-c(toRemove),]}
+  metricSummary<-data.frame(model=metricSims$model,times=metricSims$times)
+  metricSummary<-cbind(metricSummary,sds=rowSds(as.matrix(metricSims[,-c(1,2)])),
+                       means=rowMeans(as.matrix(metricSims[,-c(1,2)])))
+  ci<-apply(as.matrix(metricSims[,-c(1,2)]), 1, function(x){mean(x)+c(-1.96,1.96)*sd(x)/sqrt(length(x))})
+  rownames(ci)<-c('cilower95','ciupper95')
+  pi<-apply(as.matrix(metricSims[,-c(1,2)]), 1, function(x){quantile(x,c(.025, .975))})
+  rownames(pi)<-c('pilower95','piupper95')
+  metricSummary<-cbind(metricSummary,t(ci),t(pi))
+  return(metricSummary)
+}
+
+plotPerformance<-function(metricSummary,xlab="Follow-up-time",method='Metric',iterations='NA'){
+  mainTitle<-paste(method,' : 95% P.I. over ', iterations, ' iterations',sep='')
+  
+  a<-ggplot(data=metricSummary, aes(x=times,y=means,col=model))+
+    geom_line() +
+    geom_ribbon(aes(ymin=pilower95,ymax=piupper95,fill=model),linetype=0,alpha=0.3) +
+    ggtitle(mainTitle) + 
+    xlab(xlab) +
+    ylab(method)
+  if(method=="IPA"){
+    a<-a+coord_cartesian()#ylim(-1,1))#
+    return(a)
+  }
+  return(a+coord_cartesian())
+}
+
+
+pyProcess<-function(predictions,times=times,apr="constant"){
+  interpolated <- apply(predictions, 1, function(x) approx(as.numeric(colnames(predictions)), y=x, method = apr, xout=times)$y)
+  class(interpolated)<-c("tunnel",class(interpolated))
+  interpolated<-1-interpolated
+  interpolated[is.na(interpolated) & is.na(interpolated)>50] <- 0
+  interpolated[is.na(interpolated) & is.na(interpolated)<50] <- 1
+  return(t(interpolated))
+}
+
+
+
+
[…]

### 2023-02-08 analyses/testing.py

@@ -1,0 +1,2 @@
+def hi():
+  return 1

### 2023-04-05 StatPaper/APA-lato_New/APA/wileyNJD-APA.tex
File Changes

@@
 performance of CBNN to survival methods based on regression and neural
-networks in two simulations and two real data applications. We report
+networks in a simulation and three real data applications. We report
 two time-dependent metrics for each model. In the simulations and real
@@
 performance across time and outperforms the competing neural network
-approaches. For a simple simulation with an exponential hazard model,
-CBNN outperforms the other neural network methods. For a complex
-simulation, which highlights the ability of CBNN to model both a complex
-baseline hazard and time-varying interactions, CBNN outperforms all
-competitors. The first real data application shows CBNN outperforming
-all neural network competitors, while a second real data application
-shows competitive performance. We highlight the benefit of combining
-case-base sampling with deep learning to provide a simple and flexible
-modeling framework for data-driven, time-varying interaction modeling of
+approaches. For a complex simulation, which highlights the ability of
+CBNN to model both a complex baseline hazard and time-varying
+interactions, CBNN outperforms all competitors. The first two real data
+application shows CBNN outperforming all neural network competitors,
+while a second real data application shows competitive performance.
+We highlight the benefit of combining case-base sampling with deep
+learning to provide a simple and flexible modeling framework for
+data-driven, time-varying interaction modeling of
 survival outcomes. An R package is available at
@@
 time-varying interactions. Compared to Cox regression, DeepSurv shows
-better performance on the Study to Understand Prognoses Preferences and
-Risk Treatments (SUPPORT) dataset \citep{knaus1995SUPPORT}. To handle
+better performance on a real dataset \citep{katzman2018DeepSurv}. To handle
 non-proportional hazards, a modification to the loss function was
-proposed \citep{faraggi1995neural}. To remove the need for tuning the
-number of layers and nodes, the concept of extreme learning machines has
-been applied to a Cox neural network model \citep{wang2019extreme}.
+proposed \citep{faraggi1995neural}.
 
@@
 directly estimates survival curves, rather than deriving a hazard
-function \citep{lee2018DeepHit}. It assumes an inverse Gaussian
-distribution as the baseline hazard and it outperforms DeepSurv on the
-Molecular Taxonomy of Breast Cancer International Consortium (METABRIC)
-dataset \citep{curtis2012genomic}.
+function . It assumes an inverse Gaussian
+distribution as the baseline hazard and it outperformed DeepSurv \citep{lee2018DeepHit}.
 
@@
 parametric survival model using neural networks with a mixture of
-distributions as the baseline hazard \citep{dsmPaper}. On both the
-SUPPORT and METABRIC datasets, DSM outperforms DeepSurv and DeepHit
+distributions as the baseline hazard \citep{dsmPaper}. DSM outperformed DeepSurv and DeepHit
 \citep{dsmPaper}. However, like DeepHit and DeepSurv, DSM cannot model
@@
 Neural networks are flexible when defining the architecture and
-optimization parameters. These hyperparameter decisions can affect the
-estimated parameters and were chosen during a set of initial simulations
-to determine if CBNN can learn complex interactions in practice. We set
-a max number of epochs to be \(2000\), batch size as \(512\), learning
-rate as \(10^{-3}\), decay as \(10^{-7}\), patience as \(10\) epochs,
-\(\{50,50,25,25\}\) nodes in each hidden layer with \(50\%\) dropout at
-each layer, a minimum delta loss on the validation set of \(10^{-7}\)
-over 10 epochs and Adam \citep{gulli2017} as our optimizer. These
-choices may permit the model to approximate higher-order interactions
-while preventing over-fitting \citep{srivastava2014dropout}. We fix a
-train-validation-test split that allows us to update the weights with a
-subset of the data (training), assess performance at each epoch
-(validation) and gauge performance of the final model (test) for each
-method. We select the best weights after training based on validation
-loss \citep{gulli2017}.
+optimization parameters. These hyperparameter choices can affect the
+estimated parameters and are not necessarily the same for each model.
+First we fix a test set with 15\% of the data, then we use three-fold cross-validation on the remaining 
+data (85\%:15\% training:validation) for a range of hyperparameters (TABLE REF). We track integrated brier 
+score on the validation set for each hyperparameter combination, choosing 
+the combination with the lowest score for each method. The chosen 
+hyperparameters are in (TABLE REF).
 
@@
 
-\hypertarget{sims}{%
-\section{Simulation studies}\label{sims}}
+\hypertarget{performance-metrics}{%
+\section{Performance metrics}\label{performance-metrics}}
 
-In this section, we use simulated data to evaluate the performance of
-CBNN and compare our approach with existing regression-based (Cox, CBLR)
-and neural network-based (DeepHit, DeepSurv, DSM) methods. We specify a
-linear combination of each covariate as the linear predictor in
-regression-based approaches (Cox, CBLR), which contrasts with neural
-network approaches that allow for non-linear interactions. We simulate
-data under a simple exponential model and a complex baseline hazard with
-time-varying interactions, each with 10\% random censoring. For both
-settings, we simulate three covariates:
-
-\[
-z_{1} \sim \textrm{Bernoulli}(0.5) \qquad \qquad
-z_{2} \sim \begin{cases}
- N(0,0.5) & \textrm{if } z_{1}=0\\ 
- N(10,0.5) & \textrm{if } z_{1}=1
-\end{cases} \qquad \qquad
-z_{3} \sim \begin{cases}
- N(8,0.5) & \textrm{if } z_{1}=0\\ 
- N(-3,0.5) & \textrm{if } z_{1}=1.
-\end{cases}
-\]
-
-The DeepHit-specific hyperparameter alpha is set to \(0.5\) (equal
-weight between its negative log-likelihood and ranking losses
-\citep{lee2018DeepHit}). We modify the \textbf{DeepSurvivalMachines}
-\citep{dsmPaper} package to include dropout and a minimum delta loss
-during the fitting process. For DSM, we define a mixture model of six
-Weibull distributions for the baseline hazard. All other hyperparameters
-are held constant across all neural network methods in both the
-simulation studies and real data applications.
-
-Besides the methods mentioned above, we include the Optimal model in our
-comparisons using CBLR. That is, we include the exact functional form of
-the covariates in a CBLR model (referred to as Optimal for simplicity).
-We calculate \(t\)-based 95\% confidence intervals using 100
-replications of the simulated data. For all analyses, we use 80\% for
-training and 20\% for the test set. 20\% of the training set is kept for
-validation at each epoch. We predict risk functions \(F\) using
-\eqref{eqn:ci2} for individuals in the test set, which are used to
-calculate our \(C_{IPCW}\) and IPA scores.
-
-\hypertarget{performance-metrics}{%
-\subsection{Performance metrics}\label{performance-metrics}}
-
-We use two metrics to assess the performance of the different methods of
-interest: 1) (IPA) \citep{kattan2018index} and 2) inverse probability
+To choose our method-specific hyperparameters, we use the Integrated 
+Brier Score (IBS) \citep{graf1999}. We use two metrics to assess the 
+performance of the different methods of interest on testing data: 1) (IPA) 
+\citep{kattan2018index} and 2) inverse probability
 censoring weights-adjusted concordance index (\(C_{IPCW}\))
-\citep{uno2011}, which we define below. Both these time-dependent
-metrics provide transparency as to when in follow-up time each model may
+\citep{uno2011}, which we define below. IBS provides a summarized assessment
+of performance for each model. (\(C_{IPCW}\)) provides information
+about performance up to the survival time of interest. while IPA provide transparency
+as to when in follow-up time each model may
 perform better than the others.
@@
 \hypertarget{index-of-prediction-accuracy-ipa}{%
-\subsubsection{Index of prediction accuracy
-(IPA)}\label{index-of-prediction-accuracy-ipa}}
+\subsubsection{Index of prediction accuracy (IPA)}\label{index-of-prediction-accuracy-ipa}}
 
@@
 
+\hypertarget{integrated-brier-score-ibs}{%
+\subsubsection{Integrated Brier Score (IBS)}\label{integrated-brier-score-ibs}}
+
+The IBS is a function of the Brier score (\(BS(t)\)) \citep{graf1999}. as well,
+which is defined as \begin{align}
+IBS=\int_{t}^{t_{max}}BS(t)dw(t)
+\end{align} where $w(t)=\frac{t}{t_{max}}$. As this is a score for a range of follow-up-time,
+it is a useful metric to track during cross-validation. The performance at a specific time
+is hidden, and is why we opt for the IPA in assessing the final models on a test set.
+
 \hypertarget{inverse-probability-censoring-weights-adjusted-concordance-index}{%
@@
 
-\hypertarget{simple-simulation-constant-baseline-hazard}{%
-\subsection{Simple simulation: constant baseline
-hazard}\label{simple-simulation-constant-baseline-hazard}}
+\hypertarget{sims}{%
+\section{Simulation study}\label{sims}}
 
-We simulate data from a simple model that primarily depends on a
-constant baseline hazard:
-\[\log h(t \mid X_i) = \beta_{{1}}z_{1}+\beta_{{2}}z_{2}+\beta_{{3}}z_{3},\]
-The covariate effects are given by
-\(\beta_{{1}}=0.1, \beta_{{2}}=0.1, \beta_{{3}}=0.1\). Once we simulate
-survival times, we introduce 10\% random censoring.
+In this section, we simulate data to evaluate the performance of
+CBNN and compare our approach with existing regression-based (Cox, CBLR)
+and neural network-based (DeepHit, DeepSurv) methods. We specify a
+linear combination of each covariate as the linear predictor in
+regression-based approaches (Cox, CBLR), which contrasts with neural
+network approaches that allow for non-linear interactions. We simulate
+data under a complex baseline hazard with
+time-varying interactions, each with 10\% random censoring. For both
+settings, we simulate three covariates:
 
-\hypertarget{performance-comparison-in-simple-simulation}{%
-\subsubsection{Performance comparison in simple
-simulation}\label{performance-comparison-in-simple-simulation}}
-
-Figure \ref{fig:megaPlot} A, B and Table \ref{tab:megaTable} A show the
-results for the simple simulation. The regression-based methods (CBLR,
-Cox, Optimal) outperform the neural network ones in the simple
-simulation setting. Among the neural network approaches, CBNN
-outperforms all other methods in terms of both IPA and \(C_{IPCW}\)
-(Figure \ref{fig:megaPlot} A, B). Specifically, we see CBNN is
-consistent across time with smaller confidence intervals compared to
-DeepHit, DeepSurv and DSM.
-
-In this simple setting, the regression models are much closer to the
-Optimal model, while the neural network models perform worse than the KM
-null model. The wide confidence bands in \(C_{IPCW}\) suggest the neural
-network models may be over-parameterized.
+\[
+z_{1} \sim \textrm{Bernoulli}(0.5) \qquad \qquad %technically a control and treatement group
+z_{2} \sim \begin{cases}
+ N(0,0.5) & \textrm{if } z_{1}=0\\ 
+ N(1,0.5) & \textrm{if } z_{1}=1
+\end{cases} \qquad \qquad
+z_{3} \sim N(1,0.5)
+\]
+Besides the methods mentioned above, we include the Optimal model in our
+comparisons using CBLR. That is, we include the exact functional form of
+the covariates in a CBLR model (referred to as Optimal for simplicity).
+We calculate \(t\)-based 95\% confidence intervals using 100
+replications of the simulated data. For all analyses, 15\% of the data is 
+kept for testing, while 15\% of the remaining data is kept for validation. 
+The remaining data is kept for training. We predict risk functions \(F\) using
+\eqref{eqn:ci2} for individuals in the test set, which are used to
+calculate our \(C_{IPCW}\) and IPA scores. We conduct 100 bootstrap
+re-samples on the training data to obtain confidence intervals.
 
@@
 This simulation demonstrates performance with the presence of a complex
-baseline hazard and a time-varying interaction. Originally used to show
-the spline-based hazard model proposed by Royston and Parmar
+baseline hazard and a time-varying interaction. Inspired by a cancer treatment 
+that initially reduces risk of death, but slowly returns as time progresses. 
+Originally used to show the spline-based hazard model proposed by Royston and Parmar
 \citep{royston2002flexible}, the breast cancer dataset provides a
@@
 model as follows: \begin{align}
-\log h(t \mid X_i) =\sum_{i=1}^{5} (\gamma_{i} \cdot \psi_{i}) + \beta_{{1}} (z_{1}) + \beta_{{2}} (z_{2})+ \beta_{{3}} (z_{3})+ \tau_{1} ( z_{1} \cdot z_{2} \cdot t)+ \tau_{2} ( z_{1} \cdot z_{3})+ \tau_{3} (z_{2} \cdot z_{3}), \nonumber
+\log h(t \mid X_i) =\sum_{i=1}^{5} (\gamma_{i} \cdot \psi_{i}) + \beta_{{1}} (z_{1}) + \beta_{{2}} (z_{2})+ \beta_{{3}} (z_{3})+ \tau_{1} ( z_{1} \cdot t)+ \tau_{2} (z_{2} \cdot z_{3}), \nonumber
 \end{align}
@@
 These tables are included to provide exact measures at certain intervals.
-The models of interest are: Cox, case-base with logistic regression (CBLR), DeepSurv, DeepHit, Case-Base Neural Network (CBNN), Optimal and Deep Survival Machines (DSM).}
+The models of interest are: Cox, case-base with logistic regression (CBLR), DeepSurv, DeepHit, Case-Base Neural Network (CBNN) and Optimal}
 \label{tab:megaTable}
@@
 where
-\(\gamma_{1}=3.9, \gamma_{2}=3, \gamma_{3}=-0.43, \gamma_{4}=1.33,\gamma_{5}=-0.86, \beta_{{1}}=1, \beta_{{2}}=1, \beta_{{3}}=1, \tau_{1}=10, \tau_{2}=2, \tau_{3}=2\)
+\(\gamma_{1}=3.9, \gamma_{2}=3, \gamma_{3}=-0.43, \gamma_{4}=1.33,\gamma_{5}=-0.86, \beta_{{1}}=-5, \beta_{{2}}=-1, \beta_{{3}}=1, \tau_{1}=0.001, \tau_{2}=-1\)
 and \(\psi\) are basis splines. The \(gamma\) coefficients are obtained
@@
 
+We use a population time plot to visualize incidence density across $z1$ (FIGURE REF).
+z1 acts to mimic a treatment and control group. The control group $(z1=0)$ is not impacted
+by the time-varying interaction. The treatment group $(z1=1)$ initially acts to protect the group,
+while the risk slowly builds with time at the rate of the interaction term $(\tau_{1})$. This creates
+a separation between the median time of death between these two groups due to the time-varying
+interaction. We expect all competing models (aside from the Optimal model and CBNN) to drop in
+performance for the second half of follow-up time.
+
 \hypertarget{performance-comparison-in-complex-simulation}{%
@@
 
-Figure \ref{fig:megaPlot} C, D and Table \ref{tab:megaTable} B show the
-performance over time on a test set in the complex simulation. Apart
-from the Optimal regression model, CBNN outperforms the competing models
-when examining IPA and up to the 75-percentile of follow-up time for
-\(C_{IPCW}\). We expected the Optimal model to perform best in both
+Figure \ref{fig:megaPlot} A, E and Table \ref{tab:megaTable} A
+demonstrates the performance over time on a test set. While examining the IPA score, the optimal
+regression model and CBNN perform near identically, leading in performance followed by 
+DeepHit, the linear regression models and DeepSurv. We expected the Optimal model to perform best in both
 metrics. However, this was not the case for \(C_{IPCW}\) and may be due
@@
 flexibility in modeling time-varying interactions and baseline hazard,
-flexibility the other neural network models do not have.
+flexibility the other neural network models do not have. We keep this misspecification issue
+in mind while interpreting the \(C_{IPCW}\) results in the case studies.
 
@@
 \hypertarget{casestudies}{%
-\section{Application to SUPPORT and METABRIC data}\label{casestudies}}
+\section{Casestudies}\label{casestudies}}
 
@@
 ideal conditions with clean data. To obtain a more realistic performance
-assessment, we compared models using two real datasets with a
-time-to-event outcome. The first case study examines the SUPPORT dataset
-\citep{knaus1995SUPPORT}. The second case study examines the METABRIC
-dataset \citep{curtis2012genomic}. We use the same hyperparameters as in
-the simulation studies. As we do not know the true model for the real
-data, we exclude the Optimal model. We split the datasets keeping 20\%
-of the observations as a test set. 20\% of the training set is kept
-aside for validation at each epoch. We predict risk functions for
-everyone in the test set, which is used to calculate our metrics. We
-conduct 100 bootstrap re-samples for the real data applications to
+assessment, we compared models using three real datasets with a
+time-to-event outcome. The first case study examines multiple myeloma (MM) \citep{myeloma}.
+The second examines the relationship between serum free light chain (FLC) and mortality \citep{flc}.
+The third examines prostate cancer survival on a publicly available simulation of the Surveillance,
+Epidemiology, and End Results (SEER)-medicare prostate cancer data\citep{prostate}.
+As we do not know the true model for the real
+data, we exclude the Optimal model. After grid search is performed with the same hyperparameter
+options, We split the data the same way we did for the simulation (15\% of the data is 
+kept for testing, while 15\% of the remaining data is kept for validation and the rest for testing).
+We predict risk functions for everyone in the test set, which is used to calculate our metrics. We
+conduct 100-fold bootstrap re-samples on the training data to
 obtain confidence intervals.
 
-\hypertarget{performance-evaluation-using-the-support-dataset}{%
-\subsection{Performance evaluation using the SUPPORT
-dataset}\label{performance-evaluation-using-the-support-dataset}}
+\hypertarget{pe-multiplemyeloma}{%
+\subsection{Performance evaluation on multiple myeloma dataset}\label{pe-multiplemyeloma}}
+The (MM) dataset tracks 3882 patients seen at the Mayo Clinic from 1947-1996 until death \citep{myeloma}.
+We use 2 covariates, year of entry into the study and the time of MM diagnosis\citep{myeloma}. We see 71\% incidence over 23 years\citep{myeloma}.
 
-The SUPPORT dataset tracks the time until death for seriously ill
-patients at five American hospitals \citep{knaus1995SUPPORT}. We use a
-pre-processed version of the dataset made available in the DeepSurv
-package \citep{katzman2018DeepSurv}. This dataset contains 9104 samples
-and 14 covariates (age, sex, race, number of comorbidities, presence of
-diabetes, presence of dementia, presence of cancer, mean arterial blood
-pressure, heart rate, respiration rate, temperature, white blood cell
-count, serum's sodium and serum's creatinine)
-\citep{katzman2018DeepSurv}. Patients with missing features were
-excluded and 68.10\% of the patients died during the 5.56-year study
-period \citep{katzman2018DeepSurv}.
+Figure \ref{fig:megaPlot} B, F and Table \ref{tab:megaTable} B
+demonstrates the performance over time on a test set. With the IPA score, CBNN outperforms
+The competing models, followed by the linear models, DeepSurv and DeepHit. We note that after
+50\% of survival time, all models aside from CBNN perform substantially worse than a null model.
+We note with \(C_{IPCW}\), the linear models and DeepSurv perform equivalently, followed by CBNN and
+DeepHit.
+%paragraph about performance
 
-Figure \ref{fig:megaPlot} E, F and Table \ref{tab:megaTable} C
-demonstrates the performance over time on a test set. The regression
-models (CBLR, Cox) perform best considering IPA, followed by CBNN from
-the \(25^{th}\) to \(100^{th}\) percentile of follow-up time. We note a
-drop in performance for CBNN before the \(25^{th}\) percentile of
-follow-up. For \(C_{IPCW}\), CBNN outperforms the competing models
-consistently over follow-up time. Note that performance is similar for
-all models, aside from DeepSurv whose \(C_{IPCW}\) is lower than the
-rest (Figure \ref{fig:megaPlot} E, F and Table \ref{tab:megaTable} C).
+\hypertarget{pe-flc}{%
+\subsection{Performance evaluation on free light chain dataset}\label{pe-flc}}
+The FLC dataset is a random sample of half the individuals in $\frac{2}{3}$ of the
+residents of Olmsted County over the age of 50\citep{flc}. We have access to 7874 subjects
+tracked until death\citep{flc}. We use 5 covariates, total serum FLC (kappa+lambda), age, sex,
+serum creatine and monoclonal gammapothy state\citep{flc}. we see 27\% incidence over 14 years\citep{flc}.
 
-\hypertarget{performance-evaluation-using-the-metabric-dataset}{%
-\subsection{Performance evaluation using the METABRIC
-dataset}\label{performance-evaluation-using-the-metabric-dataset}}
+Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C
+demonstrates the performance over time on a test set. With the IPA score, CBNN outperforms
+The competing models, followed by the linear models and lastly DeepSurv and DeepHit for the first 60\% of follow-up time. 
+After 60\% of follow-up time, DeepHit has similar performance to CBNN. For the \(C_{IPCW}\) score, CBNN outperforms all models,
+followed by the linear models and finally DeepSurv. Note that for earlier ranges of survival time, DeepHit is the worst performing model. 
+After 25\% of follow-up time, DeepHit is in second place. In the FLC dataset, CBNN has a consistent top ranking for both IPA and  \(C_{IPCW}\).
 
-METABRIC is a 30-yearlong study aiming to discover the molecular drivers
-of breast tumors, following 2000 individuals with breast cancer until
-death \citep{curtis2012genomic}. They described these growths as primary
-invasive breast carcinomas, with the goal of discovering both genetic
-and clinical risk factors for breast cancer survival
-\citep{curtis2012genomic}. We used the processed dataset made available
-through DeepSurv \citep{katzman2018DeepSurv}, which includes 1980
-patients of which 57.72\% die due to breast cancer within a median 10
-years of follow-up \citep{katzman2018DeepSurv}. There are 10 covariates
-in total: time, 4 RNA-Seq gene expressions (MKI67, EGFR, PGR and ERBB2)
-and 5 clinical features (hormone treatment indicator, radiotherapy
-indicator, chemotherapy indicator, ER-positive indicator and age at
-diagnosis) \citep{katzman2018DeepSurv}.
+\hypertarget{pe-prostate}{%
+\subsection{Performance evaluation on prostate cancer dataset}\label{pe-prostate}}
+The prostate cancer dataset has a record of competing risks\citep{prostate}. As we are only interested in the
+single event scenario, we only keep individuals with prostate cancer death or censoring\citep{prostate}.
+This subset tracks 11054 individuals with three covariates, differentiation grade, age group and cancer state\citep{prostate}.
+There is a 7\% incidence over 10 years.
 
-Figure \ref{fig:megaPlot} G, H and Table \ref{tab:megaTable} D shows the
-performance on a test set over time. The IPA scores suggest that
-regression models outperform competing models on this dataset, as all
-the neural network models are equal to or perform worse than KM over
-follow-up time. Our CBNN model is comparable to KM until around the
-50-percentile of follow-up time, after which CBNN, DeepSurv and DSM drop
-in performance. The \(C_{IPCW}\) produces a different ranking. Our CBNN
-model outperforms the other models up to the \(25^{th}\) percentile of
-follow-up time, whereas DeepHit performs best from the \(50^{th}\) to
-\(75^{th}\) percentile of follow-up. With this metric, CBNN and DeepHit
-outperform the regression models. This disagreement between IPA and
-\(C_{IPCW}\) may be due to the misspecification issue of
-concordance-based metrics \citep{cindexfails2019}. The neural network
-models may be over-parameterized, as shown by the wide confidence bands
-in \(C_{IPCW}\).
+Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C shows the
+performance over time on a test set. With the IPA score, the linear models
+outperform the neural network ones, followed by CBNN and DeepHit, and
+finally DeepSurv. For the \(C_{IPCW}\) score, the linear models, CBNN, DeepHit
+perform near identically, side from DeepHit with an initial drop in performance.
+DeepSurv performed substantially worse than the other models.
 
@@
 data-driven estimate of a flexible baseline hazard and time-varying
-interactions in our hazard function. The three competing neural network
+interactions in our hazard function. The two competing neural network
 models we evaluated cannot model time-varying interactions by design
-\citep{dsmPaper} \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}. DSM
-requires a mixture component distributions for a flexible baseline
-hazard to be fit \citep{dsmPaper}. As our goal is to fix the design and
-compare performance, we did not change any shared hyperparameters. With
-our choice of shared hyperparameters and model design, DSM did not
-converge in the complex simulation and SUPPORT case study. Despite this
-limitation, we include this method when it did function as it can fit
-flexible baseline hazards. Compared to CBNN, the remaining two models
-also have limitations. DeepSurv is a proportional hazards model and does
+ \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}
+With our hyperparameters options and model design, DSM did not
+converge in the complex simulation. Due to this issue limitation,
+we did not include this method. Compared to CBNN, all three neural network
+models also have limitations. DeepSurv is a proportional hazards model and does
 not estimate the baseline hazard \citep{katzman2018DeepSurv}. DeepHit
@@
 for the baseline hazard and models the survival function directly
-\citep{lee2018DeepHit}. The alternative neural network methods match on
-time, while CBNN models time directly.
+\citep{lee2018DeepHit}. Though not included, DSM can model a flexible
+baseline hazard, however is not meant for time-varying interactions\citep{dsmPaper}.
+The alternative neural network methods match on time, while CBNN models time directly.
 
@@
 results in relation to IPA, a proper scoring rule. The model rankings
-between \(C_{IPCW}\) and IPA differed for both the complex simulation
-and METABRIC application. Wider \(C_{IPCW}\) confidence intervals for
-DeepHit and DeepSurv show potential misspecification of the models
-(Figure \ref{fig:megaPlot} D, H).
+between \(C_{IPCW}\) and IPA differed for all but FLC. If we consider
+Figure \ref{fig:megaPlot} A, E and Table \ref{tab:megaTable} A in conjuction
+with FIGURE REF POPTIME, We note a universal drop in \(C_{IPCW}\) performance
+for all models. We hide the confidence intervals in Figure \ref{fig:megaPlot} E due to visibility but list them in Table \ref{tab:megaTable} A .
+aside from CBNN, all models have a wide confidence band at 75\% of follow-up time.
+We expect the optimal model to perform best, but that's not the case looking at  \(C_{IPCW}\).
+As such, we
 
-With this interpretation of \(C_{IPCW}\) in mind, we assess two
-simulations with minimal noise and two case studies in real scenarios.
-The simple simulation demonstrates potential pitfalls associated with
-neural network models, particularly overparameterization. All neural
-network-based approaches performed worse than null KM model (IPA score),
-while the regression-based performed better. We attribute this to
-potential overparameterization in the neural network models, as the wide
-confidence intervals suggest over-fitting, even with dropout. Both
-DeepSurv and DSM are affected, while DeepHit and CBNN are less so. This
-is a limitation to our strategy for evaluating the methods, which uses a
-fixed study design across all assessments.
+With this interpretation of \(C_{IPCW}\) in mind, we assess a
+simulations with minimal noise and three case studies.
 
-From this baseline benchmark, we move on to a complex simulation that
-requires a method that can learn both time-varying interactions and have
-a flexible baseline hazard. Here, CBNN demonstrates a distinct advantage
-over all other methods. The regression models show improved performance
-over the null KM model, while the competing neural network models
-performed worse. Based on our complex simulation results (Figure
+The complex simulation requires method that can learn both time-varying
+interactions and have a flexible baseline hazard. Here, CBNN demonstrates
+a distinct advantage over all other methods. Based on our complex simulation results (Figure
 \ref{fig:megaPlot} C, D and Table \ref{tab:megaTable} B), CBNN
 outperforms the competitors when time-varying interactions and a complex
-baseline hazard are present. This simulation shows how CBNN can perform
-under ideal conditions, while the following two analyses on real data
+baseline hazard are present (after 50\% of survival time, Figure REF POPTIME).
+This simulation shows how CBNN can perform
+under ideal conditions, while the three case studies
 serve to assess its performance in realistic conditions.
 
-In the SUPPORT and METABRIC case studies, flexibility in both
-interaction modeling and baseline hazard did not improve CBNNs relative
-performance, suggesting that this flexibility does not aid prediction in
-either case study. From a biological perspective, the 30-year follow-up
-time in the METABRIC study may contain competing causes of death. The
-causes at the start of the study may not match the causes towards the
-end, potentially explaining the drop in performance as we reach later
-survival times with fewer individuals. The baseline hazard is unlikely
-to cause this drop as DSM is the most flexible competing model in our
-comparison. Over-fitting is also unlikely given the tight confidence
-intervals. Further research is required to determine the cause of the
-drop in performance seen in the METABRIC case study. In both case
-studies, CBNN outperforms the competing neural network methods.
+In the MM and FLC case studies, flexibility in both
+interaction modeling and baseline hazard improved CBNN's performance over
+the other models, suggesting that this flexibility aids prediction in
+both case studies. In the Prostate case study, the linear models outperform
+the neural network ones. CBNN and DeepHit alternate their positions depending
+on the survival time and DeepSurv maintains last place.
+We attribute this to potential overparameterization in the neural network models
+as we did not test for a small number of nodes in each hidden layer, even with dropout.
+Though the ranking places the linear models above the neural network ones, the overall
+performance is comparable aside from DeepSurv, falling within a small range of IPA scores.
 
-Taken together, CBNN outperforms all competitors in the complex
-simulation, demonstrating its value in survival settings that may
+\section{Conclusions}\label{sec5}
+
+CBNN outperforms all competitors in the complex
+simulation and two case studies, while maintaining competitive performance in a final case study,
+demonstrating its value in survival settings that may
 involve time-varying interactions and a complex baseline hazard. Once we
@@
 a sigmoid activation function to predict our hazard function. Our
-approach simplifies the incorporation of censored individuals, allowing
-survival outcomes to be treated as binary ones. Forgoing the requirement
+approach provides an alternative to the incorporation of censored individuals, treating
+survival outcomes as binary ones. Forgoing the requirement
 of custom loss functions, CBNN only requires the use of standard
@@
 and is easily extendable to any feed-forward neural network framework.
-
-\section{Conclusions}\label{sec5}
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae,
-felis. Curabitur dictum gravida mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna. Donec
-vehicula augue eu neque. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
-Mauris ut leo. Cras viverra metus rhoncus sem. Nulla et lectus vestibulum urna fringilla ultrices. Phasellus eu tellus
-sit amet tortor gravida placerat. Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. Praesent eget sem vel
-leo ultrices bibendum. Aenean faucibus. Morbi dolor nulla, malesuada eu, pulvinar at, mollis ac, nulla. Curabitur
-auctor semper nulla. Donec varius orci eget risus. Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
-
-Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
-pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
-vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
-a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
-urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.
 
@@
 
-This is an author contribution text. This is an author contribution text. This is an author contribution text. This is an author contribution text. This is an author contribution text.
+J.I.
+conceived the proof of concept, developed the theoretical formalism, developed the neural network code base, wrote the majority of the manuscript and performed the simulations and analyses.
+M.T.
+and S.B.
+provided key insights into the core sampling technique (casebase).
+All authors discussed the results and contributed to the final manuscript.
+M.T.
+contributed heavily to the casebase sampling section \ref{case-base-sampling}.
 
@@
 
-\begin{biography}{\includegraphics[width=60pt,height=70pt,draft]{empty}}{\textbf{Author Name.} This is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text.}
+\begin{biography}{\includegraphics[width=60pt,height=70pt,draft]{empty}}{\textbf{Jesse Islam} A PhD Candidate in Quantitative life sciences, studying complex relationships in data in a survival context.}
 \end{biography}

### 2023-04-05 StatPaper/APA-lato_New/documents/wileyNJD-Doc.tex
File Changes

\documentclass[APA,LATO1COL]{WileyNJD-v2}
\usepackage{moreverb}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\articletype{Article Type}%

\received{<day> <Month>, <year>}
\revised{<day> <Month>, <year>}
\accepted{<day> <Month>, <year>}

%\raggedbottom

\begin{document}

\title{A demonstration of the \LaTeX\ class file for Wiley NJD Journals\protect\thanks{This is an example for title footnote.}}

\author[1]{Author One*}

\author[2,3]{Author Two}

\author[3]{Author Three}

\authormark{AUTHOR ONE \textsc{et al}}

\address[1]{\orgdiv{Org Division}, \orgname{Org name}, \orgaddress{\state{State name}, \country{Country name}}}

\address[2]{\orgdiv{Org Division}, \orgname{Org name}, \orgaddress{\state{State name}, \country{Country name}}}

\address[3]{\orgdiv{Org Division}, \orgname{Org name}, \orgaddress{\state{State name}, \country{Country name}}}

\corres{*Corresponding author name, Corresponding address. \email{authorone@Email.com}}

\presentaddress{Present address}

\abstract[Abstract]{This paper describes the use of the \LaTeXe\ \textsf{WileyNJD-v2.cls}
class file for setting papers for \emph{Mathematical Methods in the Applied Sciences}.}

\keywords{Class file; \LaTeXe; \emph{Wiley NJD}}

\jnlcitation{\cname{%
\author{Williams K.}, 
\author{B. Hoskins}, 
\author{R. Lee}, 
\author{G. Masato}, and 
\author{T. Woollings}} (\cyear{2016}), 
\ctitle{A regime analysis of Atlantic winter jet variability applied to evaluate HadGEM3-GC2}, \cjournal{Q.J.R. Meteorol. Soc.}, \cvol{2017;00:1--6}.}

\maketitle

\footnotetext{\textbf{Abbreviations:} ANA, anti-nuclear antibodies; APC, antigen-presenting cells; IRF, interferon regulatory factor}

\section{Introduction}\label{sec1}

Many authors submitting \( \sin \cos \tan \inf_{x} \) to NJD journals use \LaTeXe\ to
prepare their papers. This paper describes the
\textsf{WileyNJD-v2.cls} class file which can be used to convert
articles produced with other \LaTeXe\ class files into the correct
form for publication in \emph{Wiley NJD Journals}.

The \textsf{WileyNJD-v2.cls} class file preserves much of the standard
\LaTeXe\ interface so that any document which was produced using
the standard \LaTeXe\ \textsf{article} style can easily be
converted to work with the \textsf{WileyNJD-v2} style. However, the
width of text and typesize will vary from that of
\textsf{article.cls}; therefore, \emph{line breaks will change}
and it is likely that displayed mathematics and tabular material
will need re-setting.

In the following sections we describe how to lay out your code to
use \textsf{WileyNJD-v2.cls} to reproduce the typographical look of
\emph{Wiley NJD Journals}.

\subsection{Procedure to install fonts}

\begin{enumerate}
\item All font files are available under the Lato-Fonts folder 
\item Font installer is available under the same folder Windows-Lato-fontinstaller.exe
\item Execute (double click the EXE file) the EXE file that will install all fonts/map files to your local drive.
\end{enumerate}

\subsection{The Three Golden Rules}

Before we proceed, we would like to
stress \emph{three golden rules} that need to be followed to
enable the most efficient use of your code at the typesetting
stage:
\begin{enumerate}
\item[(i)] keep your own macros to an absolute minimum;

\item[(ii)] as \TeX\ is designed to make sensible spacing
decisions by itself, do \emph{not} use explicit horizontal or
vertical spacing commands, except in a few accepted (mostly
mathematical) situations, such as \verb"\," before a
differential~d, or \verb"\quad" to separate an equation from its
qualifier;

\item[(iii)] follow the \emph{NJD} reference style.
\begin{enumerate}[a.]
\item Chemistry --- Use the ``APA'' option as \verb"\documentclass[APA]{WileyNJD-v2.cls}"

\end{enumerate}
\end{enumerate}

\section{Getting Started} The \textsf{WileyNJD-v2.cls} class file should run
on any standard \LaTeXe\ installation. If any of the fonts, class
files or packages it requires are missing from your installation,
they can be found on the \emph{\TeX\ Live} CD-ROMs or from CTAN.

LaTeX document class options

\begin{flushleft}
\begin{enumerate}[a.]
\item LATO1COL--- For LATO font large one column layout use the ``LATO1COL'' option as \verb"\documentclass[APA,LATO1COL]{WileyNJD-v2}"
\item LATO2COL--- For LATO font large two column layout use the ``LATO2COL'' option as \verb"\documentclass[APA,LATO2COL]{WileyNJD-v2}"
\item LATOSMALL--- For LATO font small layout use the ``LATOSMALL'' option as \verb"\documentclass[APA,LATOSMALL]{WileyNJD-v2}"
\end{enumerate}
\end{flushleft}

\section{The Article Header Information}
The heading for any file using \textsf{WileyNJD-v2.cls} is shown in
Figure~\ref{F1}.

\subsection{Remarks}

\begin{enumerate}[(I).]

\item Use \verb"\title{<title> \protect\thanks{<title footnotes}}" for article title and title footnote.
\item Use \verb"\authormark{}" for running heads.

\item Note the use of \verb"\author[<link>]{<name>}" and \verb"\address[<link>]{<name>}" to
link names and addresses. The author for correspondence is marked
by ``*'' and \verb"\corres{}" is used to give that
author's address, which will be printed besides abstract, prefaced by
`Correspondence to:'.

\item For submitting a double-spaced manuscript, add
\verb"doublespace" as an option to the documentclass line. \verb"\documentclass[doublespace]{WileyNJD-v2}"

\item Use \verb"\presentaddress{}" for present address.

\item In abstract \verb"\abstract[<title>]{abstract paragraph}" use optional parameter for title followed by abstract paragraph.

\item For Key words use \verb"\keywords{}".

\item For how to site use \verb"\jnlcitation{\cname{\author{<author name>},"

\verb"\ctitle{<title>}, \cjournal{<Journal name>}, \cvol{<vol>}.}}".

\item For title page abbreviations use \verb"\footnotetext{<\textbf{Abbreviation title:} Abreviations>}"

\item Use \verb"\articletype{<article category}" for article header information

\item Use \verb"\received{<received date>} \revised{<revised date>} \accepted{<accepted date>}" for history dates.

\end{enumerate}

\begin{figure}[p]
\setlength{\fboxsep}{0pt}%
\setlength{\fboxrule}{0pt}%
\begin{center}
\begin{boxedverbatim}

\documentclass[APA,LATO1COL]{WileyNJD-v2}

\articletype{Article Type}%

\received{26 April 2016}
\revised{6 June 2016}
\accepted{6 June 2016}

\begin{document}

\title{<Initial cap, lower case>\protect\thanks{<title footnote.>}}

\author[<address link>]{<Author name><corresponding author*>}

\author[<address link>,<address link>]{Author Name}

\authormark{AUTHOR ONE \textsc{et al}}

\address[<address link>]{\orgdiv{<Org Division>}, \orgname{<Org name>}, 
\orgaddress{\state{<State name>}, \country{<Country name>}}}
\address[<address link>]{\orgdiv{<Org Division>}, \orgname{<Org name>},
 \orgaddress{\state{<State name>}, \country{<Country name>}}}

\corres{<corresponding author link*> <author name, address.
 \email{<authorone@email.com>}}

\presentaddress{<Present address>}

\abstract[<Abstract heading>]{<Abstract paragraph>}

\keywords{<keyword1>, <keyword2>,...}

\jnlcitation{\cname{%
\author{<aurhor name>}, 
\author{<aurhor name>}, 
\author{<aurhor name>}, 
\author{<aurhor name>},  and 
\author{<aurhor name>}} (\cyear{<year>}), 
\ctitle{<journal title>}, \cjournal{<journal name>} <year> <vol> Page <xxx>-<xxx>}

\footnotetext{\textbf{<abbreviation head:>} <abbreviations> ..}

\maketitle

\section{Introduction}
.
.
.
\end{boxedverbatim}
\end{center}
\caption{Example for title page.\label{F1}}
\end{figure}

\section{The Body of the Article}

\subsection{Section headings}

\begin{enumerate}[(H1)]
\item Section --- use \verb"\section{}"
\item SubSection--- use \verb"\subsection{}"
\item SubSubSectioin--- use \verb"\subsubsection{}"
\item Paragraph--- use \verb"\paragraph{}"
\item Subparagraph--- use \verb"\subparagraph{}"
\end{enumerate}

\subsection{Mathematics} \textsf{WileyNJD-v2.cls} makes the full
functionality of \AmS\/\TeX\ available. We encourage the use of
the \verb"align", \verb"gather" and \verb"multline" environments
for displayed mathematics.

\subsection{Figures and Tables}

\textsf{WileyNJD-v2.cls} uses the
\textsf{graphicx} package for handling figures.

Figures are called in as follows:
\begin{verbatim}
\begin{figure}
\centering
\includegraphics{<figure name>}
\caption{<Figure caption>}
\end{figure}
\end{verbatim}

The standard coding for a table is shown in Figure~\ref{F2}.

\begin{figure}[h]
\setlength{\fboxsep}{0pt}%
\setlength{\fboxrule}{0pt}%
\begin{center}
\begin{boxedverbatim}
\begin{table}
\caption{<Table caption>}
\centering
\begin{tabular}{<table alignment>}
\toprule
<column headings>\\
\midrule
<table entries
(separated by & as usual)>\\
<table entries>\\
.
.
.\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item Source: xxx.
\item[1] xxx.
\item[2] xxx.
\end{tablenotes}
\end{table}
\end{boxedverbatim}
\end{center}
\caption{Example for table layout.\label{F2}}
\end{figure}

\subsection{Cross-referencing}
The use of the \LaTeX\ cross-reference system
for figures, tables, equations, etc., is encouraged
(using \verb"\ref{<name>}" and \verb"\label{<name>}").

\subsection{Box text}

\begin{verbatim}
\begin{boxtext}
\section*{<title>}%
Paragraph
\end{boxtext}
\end{verbatim}

\subsection{List items}

\subsubsection{Enumerate list styles}
\begin{verbatim}
\begin{enumerate}[1]
\item 
\end{enumerate}

\begin{enumerate}[1.]
\item 
\end{enumerate}

\begin{enumerate}[(1)]
\item 
\end{enumerate}

\begin{enumerate}[I]
\item 
\end{enumerate}

\begin{enumerate}[i]
\item 
\end{enumerate}

\begin{enumerate}[a]
\item 
\end{enumerate}

\end{verbatim}

\subsubsection{Bullet list styles}

\begin{verbatim}
\begin{itemize}
\item 
\end{itemize}
\end{verbatim}

\subsubsection{Description list}

\begin{verbatim}
\begin{description}
\item[<entry>] description text.
\end{description}
\end{verbatim}

\subsection{Enunciations}

\begin{verbatim}
\begin{theorem}[<Theorem subhead>]\label{thm1}
<theorem text>. 
\end{theorem}

\begin{proposition}[<proposition subhead>]\label{pro1}
<proposition text>. 
\end{proposition}

\begin{definition}[<definition subhead>]\label{dfn1}
<definition text>. 
\end{definition}

\begin{proof}
<proof text>. 
\end{proof}

\begin{proof}[Proof of Theorem~\ref{thm1}]
<proof text>.
\end{proof}

\end{verbatim}

\subsection{Program codes}

Use \verb+\begin{verbatim}...\end{verbatim}+ for program codes without math. Use \verb+\begin{alltt}...\end{alltt}+ for program codes with math. Based on the text provided inside the optional argument of \verb+\begin{code}[Psecode|Listing|Box|Code|+\hfill\break \verb+Specification|Procedure|Sourcecode|Program]...+ \verb+\end{code}+ tag corresponding boxed like floats are generated. Also note that \verb+\begin{code}[Code|Listing]...+ \verb+\end{code}+ tag with either Code or Listing text as optional argument text are set with computer modern typewriter font.  All other code environments are set with normal text font. Refer below example:

\begin{verbatim}
\begin{lstlisting}[caption={Descriptive Caption Text},label=DescriptiveLabel]
for i:=maxint to 0 do
begin
{ do nothing }
end;
Write('Case insensitive ');
WritE('Pascal keywords.');
\end{lstlisting}
\end{verbatim}

\subsection{Acknowledgements} An Acknowledgements section is started with \verb"\ack" or
\verb"\acks" for \textit{Acknowledgement} or
\textit{Acknowledgements}, respectively. It must be placed just
before the References.

\subsection{Bibliography}

\begin{enumerate}[1]
\item Use \verb"\bibliography{wileyNJD-AMA}" BST file for AMA reference style
\item Use \verb"\bibliography{wileyNJD-APA}" BST file for APA reference style
\item Use \verb"\bibliography{wileyNJD-AMS}" BST file for AMS reference style
\item Use \verb"\bibliography{wileyNJD-VANCOUVER}" BST file for Vancouver reference style
\item Use \verb"\bibliography{wileyNJD-ACS}" BST file for Chemistry reference style
\end{enumerate}

The normal commands for producing the reference list are:

\begin{verbatim}
\begin{thebibliography}{99}
\bibitem{<x-ref label>}
         <Reference details>
.
.
.
\end{thebibliography}
\end{verbatim}

\subsection{Appendix Section}

\begin{verbatim}
\appendix

\section{Section title of first appendix\label{app1}}
.
.
.

\end{verbatim}

\end{document}

### 2023-04-05 analyses/CopyOfoptimizeSupport.Rmd
File Changes

--- 2023-02-08 analyses/optimizeComplex.Rmd
@@
 ---
-title: "iterSim"
+title: "support"
 output: html_document
 params:
-  epo: 200
-  patience: 5
-  iteration: 2
+  epo: 2000
+  patience: 10
+  iteration: 10
   layer1: 12
@@
 ```{r setup, include=T}
-set.seed(1)
+set.seed(1234)
 knitr::opts_chunk$set(echo = TRUE)
@@
 source("src/functions.R")
-library(reticulate)
 eval_support<-T
@@
 #bsize=512
-min_delta =0# 10^-7
+min_delta = 0# 10^-7
 
@@
 
-numSamp<-2000
-
-baselineTrials<-2
+baselineTrials<-5
 
@@
 
-```{r complexPrep,eval=T}
+```{r supPrep,eval=T}
 
-##############################
-###set up complex simulation
-##############################
+source_python('src/getSUP.py')
+data<- as.data.frame(do.call(cbind, getSupportPycox()))
 
-# Define a function returning the log cum hazard at time t
-logcumhaz <- function(t, x, betas, knots) {
+colnames(data)[ncol(data)-1]<-'time'
+colnames(data)[ncol(data)]<-'status'
 
-  # Obtain the basis terms for the spline-based log
-  # cumulative hazard (evaluated at time t)
-  basis <- flexsurv::basis(knots, log(t))
+data<-data[-which(data$x2 %in% c(8,9)),]
 
-  # Evaluate the log cumulative hazard under the
-  # Royston and Parmar specification
-  res <-betas[["gamma0"]] * basis[[1]] + 
-    betas[["gamma1"]] * basis[[2]] +
-    betas[["gamma2"]] * basis[[3]] +
-    betas[["gamma3"]] * basis[[4]] +
-    betas[["gamma4"]] * basis[[5]] +
-    betas[["z1"]] * x[["z1"]]+
-    betas[["z2"]] * x[["z2"]]+
-    betas[["z3"]] * x[["z3"]]+
-    betas[["term1"]] * x[["z1"]]*t+#*x[["z2"]]*
-    #betas[["term2"]] * x[["z1"]]*x[["z3"]]+
-    betas[["term3"]] * x[["z2"]]*(x[["z3"]])
-  #
-  # Return the log cumulative hazard at time t
-  res
+for(i in c(2,3,4,5,6,7)){#3,4,
+  data[, i]<-as.factor(data[, i])
 }
-url<-'app.txt'
-ds <- read.table(url, sep = '',header = TRUE)
-ds$years.at.risk=pmin(ds$agein80,ds$onset)
-brcancer$rectime<-brcancer$rectime/max(brcancer$rectime)
-true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1, data = brcancer, k = 3) #from flexsurv
+data<-as.data.frame(model.matrix(~.,data)[,-1])
+#data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
 
-# Simulate the event times
-coefficients<-true_mod$coefficients
-coefficients<-c(coefficients,z1=0.05,z2=1.05,z3=1.05,term1=2.4,term3=1.05)#term2=1.05,
+covs<-colnames(data)
+names(data)[seq(1,length(covs)-2)] <- paste("z", seq(1,length(covs)-2), sep="")
 
-#########################
-###Simulate data
-#########################
-cov<-data.frame(id=1:numSamp,
-                z1=c(rep(1,numSamp/2),rep(0,numSamp/2)),
-                z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
-                z3=c(rnorm(numSamp/2,0,0.5),rnorm(numSamp/2,10,0.5)))
-true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1,
-                                     data = brcancer, k = 3
-) #from flexsurv
+names(data)[ncol(data)-1]<-"time"
+names(data)[ncol(data)]<-"status"
 
-dat <- simsurv(betas = coefficients, # "true" parameter values
-               x = cov,                   # covariate data for 200 individuals
-               knots = true_mod$knots,    # knot locations for splines
-               logcumhazard = logcumhaz,  # definition of log cum hazard
-               maxt = 25,               # no right-censoring
-               interval = c(0,10000000)) # interval for root finding
-colnames(dat)<-c("id","time","status")
-samp<-sample(seq(1,nrow(dat),by=1), floor(nrow(dat)*(0.1))) #random censoring
-dat$status[samp]<-0
-# Merge the simulated event times onto covariate data frame
-data <- merge(cov, dat)
 covs<-colnames(data)
-data<-data[,-1]
 
-toDiv<-max(data$time)
-data$time<-data$time/toDiv
+#toDiv<-max(data$time)
+#data$time<-data$time/toDiv
+#data$time<-data$time/max(data$time)
 spec = c(train = .85, test = .15)
@@
 res = split(data, g)
-saveRDS(res$test,"ComplexSimTest.rds")
+saveRDS(res$test,"supportTest.rds")
 
-k=5
+k=3
 klist<-list()
@@
 }
-saveRDS(klist,"ComplexSimCV.rds")
+saveRDS(klist,"supportCV.rds")
 
@@
 
-klist<-readRDS("ComplexSimCV.rds")
-# lr=c(0.001,0.01)
-# drpt=c(0.25,0.5)
-# layer1=c(50,75,100)
-# layer2=c(50,75)
-# bsize=c(floor(nrow(klist[[1]]$train)/100))
-# actv=c("relu","tanh")
-# alp=c(0,0.5,1)
+klist<-readRDS("supportCV.rds")
 
-lr=c(0.001,0.0001)
-drpt=c(.1,.2)
-layer1=c(10,25,50)
+lr=c(0.001)
+drpt=c(0.25,0.5)
+layer1=c(10,20,30)
 layer2=c(10,20,30)
 bsize=c(floor(nrow(klist[[1]]$train)/100))
-actv=c("relu","tanh")
-alp=c(0,0.5,1)
+actv=c("relu","linear")#,"tanh"
+alp=c(0)
 
@@
       val<-data$val
-
+      maxTime<-max(train$time)
       sds<-sapply(train, function(x) (sd(x)))
       means<-sapply(train, function(x) (mean(x)))
-      val<-normalizer(val,means,sds)
-      train<-normalizer(train,means,sds)
+      val<-normalizer(val,means,sds,maxTime)
+      train<-normalizer(train,means,sds,maxTime)
 
@@
       currentTime<-Sys.time() 
-      coxnnsurv=fitDeepSurv(train,val,floor(nrow(train)/100),epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
-      tempVal<-append(tempVal,min(coxnnsurv$val_metrics$scores$loss$score))
+      coxnnsurv=fitDeepSurv(train,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
+      #tempVal<-append(tempVal,min(coxnnsurv$val_metrics$scores$loss$score))
+      scores<-na.omit(coxnnsurv$val_metrics$scores$loss$score)
+
+      tempVal<-append(tempVal,scores[length(scores)])
       runTime<-append(runTime, Sys.time()-currentTime)
@@
 
-klist<-readRDS("ComplexSimCV.rds")
-test<-readRDS("ComplexSimTest.rds")
+klist<-readRDS("supportCV.rds")
+test<-readRDS("supportTest.rds")
 times<-seq(from=min(test$time)+0.000001,
@@
       currentTime<-Sys.time() 
-      hitnnSurv=fitDeephit(train,val,floor(nrow(train)/100),epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
-      tempVal<-append(tempVal,min(hitnnSurv$val_metrics$scores$loss$score))
+      hitnnSurv=fitDeephit(train,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
+      #tempVal<-append(tempVal,min(hitnnSurv$val_metrics$scores$loss$score))
+      scores<-na.omit(hitnnSurv$val_metrics$scores$loss$score)
+
+      tempVal<-append(tempVal,scores[length(scores)])
       runTime<-append(runTime, Sys.time()-currentTime)
@@
 
-```{r hyperopcbnn,eval=T}
+```{r hyperopcbnn,eval=reoptimize}
 
-optimCBNN<-function(klist,test,times,lr,drpt,layer1,layer2,bsize,actv){
+optimCBNN<-function(klist,test,times,lr,drpt,layer1,layer2,nBatch,actv){
 
@@
                                              layer2=layer2,
-                                             #bsize=,
+                                             nBatch=nBatch,
                                              actv=actv,
@@
       val<-data$val
-
+      maxTime<-max(train$time)
       sds<-sapply(train, function(x) (sd(x)))
@@
       #val[[1]][[1]]<-as.matrix(normalizer(as.data.frame(val[[1]][[1]]),means,sds))
-      train<-normalizer(train,means,sds)
+      train<-normalizer(train,means,sds,1)
+
       #train[,ncol(train)]<-data$train[,ncol(train)]
-      val<-normalizer(val,means,sds)
+      val<-normalizer(val,means,sds,1)
       train<-sampleCaseBase(data = as.data.frame(train),time="time",event="status",ratio=100)
@@
       val[[1]][[2]]<-rep(train$offset[1],nrow(val[[1]][[1]]))
+      #val[[1]][[2]]<-rep(0,nrow(val[[1]][[1]]))
 
@@
       layer2=paramTrials[i,]$layer2
-      bsize=c(floor(nrow(train)/100))
+      bsize=c(floor(nrow(train)/paramTrials[i,]$nBatch))
       actv=paramTrials[i,]$actv
@@
                         patience=patience,val=val)
-      tempVal<-append(tempVal,min(fit$resultOfFit$metrics$val_loss))
+      #tempVal<-append(tempVal,min(fit$resultOfFit$metrics$val_loss))
+
+      scores<-na.omit(fit$resultOfFit$metrics$val_loss)
+
+      tempVal<-append(tempVal,scores[length(scores)])
       runTime<-append(runTime, Sys.time()-currentTime)
@@
 
-klist<-readRDS("ComplexSimCV.rds")
-test<-readRDS("ComplexSimTest.rds")
+klist<-readRDS("supportCV.rds")
+test<-readRDS("supportTest.rds")
 
@@
 #cbnnPerformance<-cbnnPerformance[which.min(cbnnPerformance$valLoss),]
-
 ```
 
-```{r}
-
+```{r winners}
 winners<-list(cbnn=cbnnPerformance,
@@
 
-saveRDS(winners,"complexOptimWins.rds")
+saveRDS(winners,"supportWins.rds")
+
 ```
@@
 
-winners<-readRDS("complexOptimWins.rds")
+winners<-readRDS("supportWins.rds")
 
-klist<-readRDS("ComplexSimCV.rds")
+klist<-readRDS("supportCV.rds")
 data<-rbind(klist[[1]]$train,klist[[1]]$val)
+#data$time<-data$time/max(data$time)
 
@@
 
+  #train<-res$train
   val<-res$validate
@@
   means<-sapply(train, function(x) (mean(x)))
+  maxTime<-max(train$time)
 
-  fullTest<-readRDS("ComplexSimTest.rds")
+  fullTest<-readRDS("supportTest.rds")
   test<-fullTest[,-c(ncol(data))]
-  test<-normalizer(fullTest,means,sds)[,-c(ncol(data))]
-  fullTest<-normalizer(fullTest,means,sds)
+  test<-normalizer(fullTest,means,sds,maxTime)[,-c(ncol(data))]
+  fullTest<-normalizer(fullTest,means,sds,maxTime)
 
-  val<-normalizer(res$validate,means,sds)
+  val<-normalizer(res$validate,means,sds,maxTime)
 
-  train<-normalizer(train,means,sds)
-  valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)#rbind(train,val)
-  valcb$offset<-0
-  valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
-  times<-seq(from=min(test$time)+0.000001,
-             to=max(test$time)-0.000001,
-             length.out = 27
+  train<-normalizer(train,means,sds,maxTime)
+
+  linModelTrain<-as.data.frame(rbind(train,val))
+  #valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)#rbind(train,val)
+  #valcb$offset<-0
+  #valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
+  times<-seq(from=min(test$time),
+             to=max(test$time),
+             length.out = 22
   )
@@
 
-  ###########################
-  ###casebase optimal model
-  ###########################
-  final_mod_cb_glm <- fitSmoothHazard(status~bs(time)-time+z1+z2+z3+z1:time+z2:z3,#+z1:z3
-                                      data = train,
-                                      time = "time",
-                                      event="status",
-                                      ratio = 100
-  )
-
-  summary(final_mod_cb_glm)
-  tglmAbsRisk<-absoluteRisk(final_mod_cb_glm,time=times,newdata=test,type="CI")
-  tglmAbsRisk<-as.data.frame(tglmAbsRisk)
-  rownames(tglmAbsRisk)<-tglmAbsRisk$time
-  trueGlmProper<- t(tglmAbsRisk[-1,-1])
-  class(trueGlmProper)<-c("tunnel",class(trueGlmProper))
-
   #########################
@@
   ##########################
-  mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
-                                data = train,
+  mod_cb_glm <- casebase::fitSmoothHazard(status~bs(time)+.-time,
+                                data = linModelTrain,
                                 time = "time",
-                                event="status",
+                                event= "status",
                                 ratio = 100
@@
 
+    miniee <- Score(list('mod'=cnnCleaned),
+                             data =fullTest, 
+                             formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
+                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
+   miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))]
+
   #############################
@@
   ############################  
-  cox<-coxph(Surv(time, status) ~ ., data = train,x=T)
+  cox<-coxph(Surv(time, status) ~ ., data = linModelTrain,x=T)
 
@@
   #############################
+  tempTrain<-train
   train<-sampleCaseBase(data = as.data.frame(train),time="time",event="status",ratio=100)
@@
   layer2=winners$cbnn$layer2
-  #bsize=winners$cbnn$bsize
+  bsize=c(floor(nrow(train)/winners$cbnn$nBatch))
   actv=winners$cbnn$actv
-
+   # 
+   # 
    #  lr=0.001
-  # drpt=0.1
-  # layer1=50
-  # layer2=20
-  # #bsize=winners$cbnn$bsize
-  # actv='relu'
+   #  drpt=0.5
+   #  layer1=10
+   #  layer2=10
+   #  patience=10
+    actv='linear'
 
   currentTime<-Sys.time() 
-  # valcbtensor[[1]][[2]]<-rep(mod_cb_glm$offset[1],nrow(valcbtensor[[1]][[1]]))
   covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
@@
   covars_output<-covars_input%>%
-    layer_dense(units=layer1,use_bias = T,activation = actv)%>%
+    layer_dense(units=layer1,use_bias = T)%>%
+    layer_activation(activation = actv)%>%
     layer_dropout(drpt)%>%
-    layer_dense(units=layer2,use_bias = T,activation = actv)%>%
+    layer_dense(units=layer2,use_bias = T)%>%
+    layer_activation(activation = actv)%>%
     layer_dropout(drpt)%>%
@@
                     epochs=epo,
-                    batch_size=c(floor(nrow(train)/100)),
+                    batch_size=bsize,
                     verbose=0,
@@
   class(annProperpoly)<-c("tunnel",class(annProperpoly))
-  rm(fit,cbnn,covars_input,covars_output)
+
   ############################
@@
 
+  ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
+    geom_line()#+coord_cartesian(ylim=c(-0.1,0.2))
+  rm(fit,cbnn,covars_input,covars_output)
   #######################
@@
   cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=length(riskList)+1)
-  et_train<-as.matrix(train)[,c(5,4)]
-  et_test<-as.matrix(fullTest[,c(5,4)])
+  et_train<-as.matrix(tempTrain)[,c(ncol(tempTrain),ncol(tempTrain)-1)]
+  et_test<-as.matrix(fullTest[,c(ncol(fullTest),ncol(fullTest)-1)])
   colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
-                          'cbnn','optimal')
+                          'cbnn')
   cScoreTemp[,1]<- tempTims
@@
   )
-
-  saveRDS(BrierIPAList,'results/100complexBrier.rds')
-  saveRDS(cScore,'results/100complexcidx.rds')
+  print(paste("current iter:",pos))
+  saveRDS(BrierIPAList,'results/100supportBrier.rds')
+  saveRDS(cScore,'results/100supportcidx.rds')
   pos=pos+1
@@
 
-```{r plotComplex}
+```{r plotsupport}
 
-BrierIPAList<-readRDS('results/100complexBrier.rds')
+BrierIPAList<-readRDS('results/100supportBrier.rds')
 
@@
 
+ipa+coord_cartesian(ylim=c(-0.1,0.2))
 #end_time-start_time
-ipa+coord_cartesian(ylim=c(-2,1))
 
 ```
-
-```{r optimDSM,eval=F}
-klist<-readRDS("ComplexSimCV.rds")
-test<-readRDS("ComplexSimTest.rds")
-times<-seq(from=min(test$time)+0.000001,
-           to=max(test$time)-0.000001,
-           length.out = 25
-)
-
-paramTrials<-as.data.frame(tidyr::crossing(lr=c(0.001,0.01,0.1),
-                                           drpt=c(0.25),
-                                           layer1=c(10,25,50,100),
-                                           layer2=c(10,25,50,100),
-                                           bsize=c(floor(nrow(klist[[1]]$train)/100)),
-                                           actv=c("relu","tanh"),
-                                           alp=c(0,0.5,1),
-                                           k=c(3,6,9),
-                                           dist=c('weibull','lognormal'),
-                                           #IPA=NA,
-                                           #IPAse=NA,
-                                           valLoss=NA,
-                                           runtime=NA))
-
-paramTrials<-paramTrials[sample(nrow(paramTrials), baselineTrials), ]
-
-# paramTrials<-unique.data.frame(paramTrials)
-#paramTrials<-readRDS("complexTrials.rds")
-for(i in 1:nrow(paramTrials)){
-  #paramTrials<-readRDS("complexTrials.rds")  
-  tempRes<-c()
-  tempVal<-c()
-  runTime<-c()
-
-  for(data in klist){
-
-    train<-data$train
-    val<-data$val
-
-    lr=paramTrials[i,]$lr
-    drpt=paramTrials[i,]$drpt
-    layer1=paramTrials[i,]$layer1
-    layer2=paramTrials[i,]$layer2
-    bsize=paramTrials[i,]$bsize
-    actv=paramTrials[i,]$actv
-    alp=paramTrials[i,]$alp
-    k=paramTrials[i,]$k
-    dist=paramTrials[i,]$dist
-    #=paramTrials[i,]$IPA
-    #=paramTrials[i,]$valLoss
-
-    currentTime<-Sys.time()
-
-    source_python('src/optimDSM.py')
-    dsmSurv<-dsmfitter(train,test,bsize,20,val,times,lr,actv,k,dist,layer1,layer2,tPos=3,ePos=4)
-    if((sum(is.na(dsmSurv))==ncol(dsmSurv)*nrow(dsmSurv))){
-
-      dsmSurv<-dsmfitter(train=train,fullTest=test,bsize=floor(nrow(train)/100),epochs=20,val=val,times=times,lr=lr,actv=actv,k,dist,layer1,layer2,tPos=3,ePos=4)
-      py_run_string("del dsmfitter")
-      rm(dsmfitter)
-      next
-    }
-    tempVal<-append(tempVal,min(hitnnSurv$val_metrics$scores$loss$score))
-    runTime<-append(runTime, Sys.time()-currentTime)
-    currentTime<-Sys.time() 
-  }
-  rm(fitDeephit)
-
-  #paramTrials$IPA[i]<-mean(tempRes)
-  #paramTrials$IPAse[i]<-sd(tempRes)
-  paramTrials$valLoss[i]<-mean(tempVal)
-  paramTrials$runtime[i]<-mean(runTime)
-  #saveRDS(paramTrials,"complexds.rds")
-  print(paste("trial",i))
-  print(mean(runTime))
-  #print(as.character(paramTrials[i,]))
-  #rm(cbnn,fit,annPreds,covars_input,covars_output)
-  gc()
-}
-
-```

### 2023-04-05 analyses/optimAids.Rmd
File Changes

--- 2023-04-05 analyses/CopyOfoptimizeSupport.Rmd
@@
 ---
-title: "support"
+title: "aids"
 output: html_document
 params:
-  epo: 2000
-  patience: 10
-  iteration: 10
+  epo: 200
+  patience: 5
+  iteration: 2
   layer1: 12
@@
 ```{r}
+library(reticulate)
 
-library(reticulate)
 ```
@@
 ```{r setup, include=T}
-set.seed(1234)
+set.seed(1)
 knitr::opts_chunk$set(echo = TRUE)
@@
 source("src/functions.R")
-eval_support<-T
+library(reticulate)
+eval_aids<-T
 #source_python('src/getSUP.py')
@@
 #bsize=512
-min_delta = 0# 10^-7
+min_delta = 0 #10^-7
 
@@
 
-baselineTrials<-5
+#baselineTrials<-2
 
@@
 
-```{r supPrep,eval=T}
+```{r aidsPrep,eval=T}
 
-source_python('src/getSUP.py')
-data<- as.data.frame(do.call(cbind, getSupportPycox()))
+library(MASS)
+aids<-Aids2
 
-colnames(data)[ncol(data)-1]<-'time'
-colnames(data)[ncol(data)]<-'status'
+aids<-aids[which(aids$T.categ=="hs"),]
+aids<-aids[which(aids$sex=="M"),]
 
-data<-data[-which(data$x2 %in% c(8,9)),]
+stattt<-as.factor(aids$status)
 
-for(i in c(2,3,4,5,6,7)){#3,4,
-  data[, i]<-as.factor(data[, i])
-}
-data<-as.data.frame(model.matrix(~.,data)[,-1])
-#data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
+levels(stattt)<-c(0,1)
+preData<-data.frame(#sex=as.factor(aids$sex),
+                    #tcat=as.factor(aids$T.categ),
+                    age=as.numeric(aids$age),
+                    time=as.numeric(as.numeric(aids$death)-as.numeric(aids$diag)),
+                    status=as.numeric(stattt)-1
+)
+preData<-na.omit(preData)
 
-covs<-colnames(data)
-names(data)[seq(1,length(covs)-2)] <- paste("z", seq(1,length(covs)-2), sep="")
+modData<-as.data.frame(model.matrix(~.,preData)[,-1])
 
-names(data)[ncol(data)-1]<-"time"
-names(data)[ncol(data)]<-"status"
+data<-as.data.frame(apply(modData,2,as.numeric))
 
-covs<-colnames(data)
-
-#toDiv<-max(data$time)
-#data$time<-data$time/toDiv
-#data$time<-data$time/max(data$time)
 spec = c(train = .85, test = .15)
@@
 res = split(data, g)
-saveRDS(res$test,"supportTest.rds")
+
+saveRDS(res$test,"aidsTest.rds")
 
@@
 }
-saveRDS(klist,"supportCV.rds")
+saveRDS(klist,"aidsCV.rds")
 
@@
 
-klist<-readRDS("supportCV.rds")
+klist<-readRDS("aidsCV.rds")
+
+# lr=c(0.001)
+# drpt=c(0.25,0.5)
+# layer1=c(10,20,30)
+# layer2=c(10,20,30)
+# bsize=c(floor(nrow(klist[[1]]$train)/100))
+# actv=c("relu","linear")#,"tanh"
+# alp=c(0)
+
+lr=c(0.001,0.0005)
+drpt=c(0.25,0.1)
+layer1=c(20,50,100)
+layer2=c(10,25,50)
+nBatch=c(100,500,1000)
+actv=c("relu","linear","tanh")#
+alp=c(0,0.5,1)
 
 lr=c(0.001)
-drpt=c(0.25,0.5)
-layer1=c(10,20,30)
-layer2=c(10,20,30)
-bsize=c(floor(nrow(klist[[1]]$train)/100))
-actv=c("relu","linear")#,"tanh"
+drpt=c(0.25)
+layer1=c(20)
+layer2=c(10)
+nBatch=c(1000,100)
+actv=c("relu")#
 alp=c(0)
@@
 
-```{r hyperOpDsurv,eval=reoptimize}
+```{r hyperOpDsurv}
 
-optimDeepSurv<-function(klist,test,times,baselineTrials,lr,drpt,layer1,layer2,bsize,actv){
+optimDeepSurv<-function(klist,test,baselineTrials,lr,drpt,layer1,layer2,nBatch,actv){
 
@@
                                              layer2=layer2,
-                                             bsize=c(bsize),
+                                             nBatch=nBatch,
                                              actv=actv,
@@
     for(data in klist){
-
+      currentTime<-Sys.time() 
       train<-data$train
@@
       train<-normalizer(train,means,sds,maxTime)
+      times<-seq(from=min(val$time),
+                 to=max(val$time),
+                 length.out = 22
+      )
+
+      times<-head(times, -1)
+      times<-tail(times, -1)
 
@@
       layer2=paramTrials[i,]$layer2
-      bsize=paramTrials[i,]$bsize
+      bsize=ceiling(nrow(train)/paramTrials[i,]$nBatch)
       actv=paramTrials[i,]$actv
@@
       #=paramTrials[i,]$valLoss
-      source_python('src/optimDSurv.py')
+      source_python('src/dsurv.py')
 
-      currentTime<-Sys.time() 
-      coxnnsurv=fitDeepSurv(train,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
-      #tempVal<-append(tempVal,min(coxnnsurv$val_metrics$scores$loss$score))
-      scores<-na.omit(coxnnsurv$val_metrics$scores$loss$score)
+      ctrain<-train
+      #ctrain$time<- train$time + runif(nrow(train),-0.000001,0.000001)
 
-      tempVal<-append(tempVal,scores[length(scores)])
+      coxnnsurv=fitDeepSurv(ctrain,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
+
+      coxnnsurv<-t(coxnnsurv)
+      cnnCleaned<-pyProcess(coxnnsurv,times=times)
+      colnames(cnnCleaned)<-times
+      py_run_string("del fitDeepSurv")
+      rm(fitDeepSurv)
+
+      miniee <- Score(list('mod'=cnnCleaned),
+                      data =val, 
+                      formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
+                      se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
+
+      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
       runTime<-append(runTime, Sys.time()-currentTime)
@@
       # py_run_string("del fitDeepSurv")
-      rm(fitDeepSurv)
+      # rm(fitDeepSurv)
       # brierFinalResults <- Score(list('ds'=cnnCleaned),
@@
 
-klist<-readRDS("supportCV.rds")
-test<-readRDS("supportTest.rds")
-times<-seq(from=min(test$time)+0.000001,
-           to=max(test$time)-0.000001,
-           length.out = 25
-)
-
-bestDS<-optimDeepSurv(klist,test,times,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,bsize,actv)
+klist<-readRDS("aidsCV.rds")
+test<-readRDS("aidsTest.rds")
+source_python('src/dsurv.py')
+bestDS<-optimDeepSurv(klist,test,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,nBatch,actv)
 
@@
 
-optimDeepHit<-function(klist,test,times,baselineTrials,lr,drpt,layer1,layer2,bsize,actv,alp){
+optimDeepHit<-function(klist,test,times,baselineTrials,lr,drpt,layer1,layer2,nBatch,actv,alp){
 
@@
       #=paramTrials[i,]$valLoss
-      source_python('src/optimDH.py')
+      source_python('src/deephitter.py')
+      hitnnSurv=fitDeephit(train,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
+      hitnnSurv<-t(hitnnSurv)
+      deephitCleaned<-pyProcess(hitnnSurv,times=times)
+      colnames(deephitCleaned)<-times
+      py_run_string("del fitDeephit")
+      rm(fitDeephit)
 
-      currentTime<-Sys.time() 
-      hitnnSurv=fitDeephit(train,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
-      #tempVal<-append(tempVal,min(hitnnSurv$val_metrics$scores$loss$score))
-      scores<-na.omit(hitnnSurv$val_metrics$scores$loss$score)
+      miniee <- Score(list('mod'=deephitCleaned),
+                      data =val, 
+                      formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
+                      se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
 
-      tempVal<-append(tempVal,scores[length(scores)])
+      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
       runTime<-append(runTime, Sys.time()-currentTime)
@@
     }
-    rm(fitDeephit)
 
@@
     print(paste("trial",i))
-    print(mean(runTime))
+    print(sum(runTime))
     #print(as.character(paramTrials[i,]))
@@
 
-klist<-readRDS("supportCV.rds")
-test<-readRDS("supportTest.rds")
+klist<-readRDS("aidsCV.rds")
+test<-readRDS("aidsTest.rds")
 times<-seq(from=min(test$time)+0.000001,
@@
 
-bestDH<-optimDeepHit(klist,test,times,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,bsize,actv,alp)
+bestDH<-optimDeepHit(klist,test,times,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,nBatch,actv,alp)
 
@@
       #val[[1]][[1]]<-as.matrix(normalizer(as.data.frame(val[[1]][[1]]),means,sds))
-      train<-normalizer(train,means,sds,1)
+      train<-normalizer(train,means,sds,maxTime)
 
       #train[,ncol(train)]<-data$train[,ncol(train)]
-      val<-normalizer(val,means,sds,1)
+      val<-normalizer(val,means,sds,maxTime)
+      valTest<-val
+      times<-seq(from=min(val$time),
+                 to=max(val$time),
+                 length.out = 22
+      )
+
+      times<-head(times, -1)
+      times<-tail(times, -1)
+
       train<-sampleCaseBase(data = as.data.frame(train),time="time",event="status",ratio=100)
@@
       #tempVal<-append(tempVal,min(fit$resultOfFit$metrics$val_loss))
+      annPreds<-aar(fit,
+                    times=times,
+                    x_test=valTest[,-c(ncol(valTest))]
+      )
 
-      scores<-na.omit(fit$resultOfFit$metrics$val_loss)
+      rownames(annPreds)<-annPreds[,1]
+      annProperpoly<- t(annPreds[,-1])
+      class(annProperpoly)<-c("tunnel",class(annProperpoly))
 
-      tempVal<-append(tempVal,scores[length(scores)])
+      miniee <- Score(list('mod'=annProperpoly),
+                      data =valTest, 
+                      formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
+                      se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
+
+      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
+
       runTime<-append(runTime, Sys.time()-currentTime)
@@
     }
+
     paramTrials$valLoss[i]<-mean(tempVal)
@@
 
-klist<-readRDS("supportCV.rds")
-test<-readRDS("supportTest.rds")
+klist<-readRDS("aidsCV.rds")
+test<-readRDS("aidsTest.rds")
 
-cbnnPerformance<-optimCBNN(klist,test,times,lr,drpt,layer1,layer2,bsize,actv)
+cbnnPerformance<-optimCBNN(klist,test,times,lr,drpt,layer1,layer2,nBatch,actv)
 #cbnnPerformance<-cbnnPerformance[which.min(cbnnPerformance$valLoss),]
@@
 
-saveRDS(winners,"supportWins.rds")
+saveRDS(winners,"aidsWins.rds")
 
@@
 
-winners<-readRDS("supportWins.rds")
+winners<-readRDS("aidsWins.rds")
 
-klist<-readRDS("supportCV.rds")
+klist<-readRDS("aidsCV.rds")
 data<-rbind(klist[[1]]$train,klist[[1]]$val)
@@
 
-  fullTest<-readRDS("supportTest.rds")
+  fullTest<-readRDS("aidsTest.rds")
   test<-fullTest[,-c(ncol(data))]
@@
              to=max(test$time),
-             length.out = 22
+             length.out = 20
   )
@@
   layer2=winners$ds$layer2
-  bsize=winners$ds$bsize
+  bsize=ceiling(nrow(train)/winners$ds$nBatch)
   actv=winners$ds$actv
@@
   source_python('src/dsurv.py')
-  coxnnsurv=fitDeepSurv(train,fullTest,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
+  ctrain<-train
+  ctrain$time<- train$time
+
+  coxnnsurv=fitDeepSurv(ctrain,fullTest,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
   coxnnsurv<-t(coxnnsurv)
+
   cnnCleaned<-pyProcess(coxnnsurv,times=times)
@@
   rm(fitDeepSurv)
-
-    miniee <- Score(list('mod'=cnnCleaned),
-                             data =fullTest, 
-                             formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
-   miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))]
 
@@
   layer2=winners$dh$layer2
-  bsize=winners$dh$bsize
+  bsize=ceiling(nrow(train)/winners$dh$nBatch)
   actv=winners$dh$actv
@@
   #  patience=10
-    actv='linear'
+  #actv='linear'
 
@@
   print(paste("current iter:",pos))
-  saveRDS(BrierIPAList,'results/100supportBrier.rds')
-  saveRDS(cScore,'results/100supportcidx.rds')
+  saveRDS(BrierIPAList,'results/100aidsBrier.rds')
+  saveRDS(cScore,'results/100aidscidx.rds')
   pos=pos+1
@@
 
-```{r plotsupport}
+```{r plotaids}
 
-BrierIPAList<-readRDS('results/100supportBrier.rds')
+BrierIPAList<-readRDS('results/100aidsBrier.rds')
 
@@
 
-ipa+coord_cartesian(ylim=c(-0.1,0.2))
+ipa+coord_cartesian(ylim=c(-0.1,0.5))
 #end_time-start_time

### 2023-04-05 analyses/optimDialysis.Rmd
File Changes

--- 2023-04-05 analyses/optimAids.Rmd
@@
 ---
-title: "aids"
+title: "dialysis"
 output: html_document
 params:
-  epo: 200
-  patience: 5
-  iteration: 2
+  epo: 2000
+  patience: 10
+  iteration: 100
   layer1: 12
@@
 library(reticulate)
-eval_aids<-T
+eval_dialysis<-T
 #source_python('src/getSUP.py')
@@
 
-```{r aidsPrep,eval=T}
+```{r dialysisPrep,eval=FALSE,hide=FALSE}
 
-library(MASS)
-aids<-Aids2
+dialysis<-readRDS("dialysis.rds")
 
-aids<-aids[which(aids$T.categ=="hs"),]
-aids<-aids[which(aids$sex=="M"),]
+data <- dialysis[, c(1,2,3,4,7,6,5)]
+colnames(data)[7]<-"status"
 
-stattt<-as.factor(aids$status)
+#RcmdrPlugin.survival/html/Dialysis.html
 
-levels(stattt)<-c(0,1)
-preData<-data.frame(#sex=as.factor(aids$sex),
-                    #tcat=as.factor(aids$T.categ),
-                    age=as.numeric(aids$age),
-                    time=as.numeric(as.numeric(aids$death)-as.numeric(aids$diag)),
-                    status=as.numeric(stattt)-1
-)
-preData<-na.omit(preData)
+smallCenters<-as.numeric(names(which( table(data$center)<30)))
+toRemove<-c()
+for(i in 1:nrow(data)){
+  if(data$center[i] %in% smallCenters){
+    toRemove<-append(toRemove,i)
+  }
+
+}
+#data<-data[-toRemove,]
+
+preData<-data.frame(#center=as.factor(data$center),
+                    age=as.numeric(data$age),
+                    #begin=as.numeric(data$begin),
+                    disease=as.factor(data$disease),
+                    time=data$time,
+                    status=data$status)
 
@@
 
-saveRDS(res$test,"aidsTest.rds")
+saveRDS(res$test,"dialysisTest.rds")
 
@@
 }
-saveRDS(klist,"aidsCV.rds")
+saveRDS(klist,"dialysisCV.rds")
 
@@
 
-```{r setFixedHyperparameters}
+```{r setFixedHyperparameters,eval=FALSE,hide=FALSE}
 
-klist<-readRDS("aidsCV.rds")
+klist<-readRDS("dialysisCV.rds")
 
@@
 
-lr=c(0.001,0.0005)
-drpt=c(0.25,0.1)
-layer1=c(20,50,100)
+lr=c(0.001,0.01)
+drpt=c(0.01,0.05,0.1)
+layer1=c(50,75,100)
 layer2=c(10,25,50)
-nBatch=c(100,500,1000)
-actv=c("relu","linear","tanh")#
+nBatch=c(100,500)
+actv=c("relu","linear")#
 alp=c(0,0.5,1)
-
-lr=c(0.001)
-drpt=c(0.25)
-layer1=c(20)
-layer2=c(10)
-nBatch=c(1000,100)
-actv=c("relu")#
-alp=c(0)
+# 
+# lr=c(0.001)
+# drpt=c(0.25)
+# layer1=c(20)
+# layer2=c(10)
+# nBatch=c(1000,100)
+# actv=c("relu")#
+# alp=c(0)
 
@@
 
-```{r hyperOpDsurv}
+```{r hyperOpDsurv,eval=FALSE,hide=FALSE}
 
@@
 
-klist<-readRDS("aidsCV.rds")
-test<-readRDS("aidsTest.rds")
+klist<-readRDS("dialysisCV.rds")
+test<-readRDS("dialysisTest.rds")
 source_python('src/dsurv.py')
@@
 
-```{r optimDeepHit,eval=reoptimize}
+```{r optimDeepHit,eval=FALSE,hide=FALSE}
 
@@
 
-klist<-readRDS("aidsCV.rds")
-test<-readRDS("aidsTest.rds")
+klist<-readRDS("dialysisCV.rds")
+test<-readRDS("dialysisTest.rds")
 times<-seq(from=min(test$time)+0.000001,
@@
 
-```{r hyperopcbnn,eval=reoptimize}
+```{r hyperopcbnn,eval=FALSE,hide=FALSE}
 
@@
 
-klist<-readRDS("aidsCV.rds")
-test<-readRDS("aidsTest.rds")
+klist<-readRDS("dialysisCV.rds")
+test<-readRDS("dialysisTest.rds")
 
@@
 
-```{r winners}
+```{r winners,eval=FALSE,hide=FALSE}
 winners<-list(cbnn=cbnnPerformance,
@@
 
-saveRDS(winners,"aidsWins.rds")
+saveRDS(winners,"dialysisWins.rds")
 
@@
 
-winners<-readRDS("aidsWins.rds")
+winners<-readRDS("dialysisWins.rds")
 
-klist<-readRDS("aidsCV.rds")
+klist<-readRDS("dialysisCV.rds")
 data<-rbind(klist[[1]]$train,klist[[1]]$val)
@@
 
-  fullTest<-readRDS("aidsTest.rds")
+  fullTest<-readRDS("dialysisTest.rds")
   test<-fullTest[,-c(ncol(data))]
@@
   print(paste("current iter:",pos))
-  saveRDS(BrierIPAList,'results/100aidsBrier.rds')
-  saveRDS(cScore,'results/100aidscidx.rds')
+  saveRDS(BrierIPAList,'results/100dialysisBrier.rds')
+  saveRDS(cScore,'results/100dialysiscidx.rds')
   pos=pos+1
@@
 
-```{r plotaids}
+```{r plotdialysis}
 
-BrierIPAList<-readRDS('results/100aidsBrier.rds')
+BrierIPAList<-readRDS('results/100dialysisBrier.rds')
 
@@
 
-ipa+coord_cartesian(ylim=c(-0.1,0.5))
+ipa+coord_cartesian(ylim=c(-0.1,0.2))
 #end_time-start_time
@@
 ```
+
+```{r}
+
+cidxList<-readRDS('results/100dialysiscidx.rds')
+
+temp<-melt(setDT(as.data.frame(cidxList[[1]])), id.vars = c("times"), variable.name = "model")
+cidxSims=data.frame(model=temp$model,times=temp$times)
+
+count=1
+for(sim in cidxList){
+  cidxSims<-cbind(cidxSims, melt(setDT(as.data.frame(sim)), id.vars = c("times"), variable.name = "model")$value)
+
+  count<-count+1
+}
+
+cidxSummary<-statCalcSim(cidxSims)
+cidxC<-plotPerformance(cidxSummary,method="C-Index",iterations=iteration  )
+cidxC
+
+```

### 2023-04-05 analyses/optimFlchain.Rmd
File Changes

--- 2023-04-05 analyses/optimDialysis.Rmd
@@
 ---
-title: "dialysis"
+title: "flchain"
 output: html_document
@@
 library(reticulate)
-eval_dialysis<-T
+eval_flchain<-T
 #source_python('src/getSUP.py')
@@
 
-```{r dialysisPrep,eval=FALSE,hide=FALSE}
+```{r flchainPrep,eval=FALSE,hide=FALSE}
 
-dialysis<-readRDS("dialysis.rds")
+library(survival)
+flchain<-flchain
 
-data <- dialysis[, c(1,2,3,4,7,6,5)]
-colnames(data)[7]<-"status"
-
-#RcmdrPlugin.survival/html/Dialysis.html
-
-smallCenters<-as.numeric(names(which( table(data$center)<30)))
-toRemove<-c()
-for(i in 1:nrow(data)){
-  if(data$center[i] %in% smallCenters){
-    toRemove<-append(toRemove,i)
-  }
-
-}
-#data<-data[-toRemove,]
-
-preData<-data.frame(#center=as.factor(data$center),
-                    age=as.numeric(data$age),
-                    #begin=as.numeric(data$begin),
-                    disease=as.factor(data$disease),
-                    time=data$time,
-                    status=data$status)
+preData<-data.frame(age=as.numeric(flchain$age),
+                    #sampyr=as.numeric(flchain$sample.yr),
+                    sex=as.factor(flchain$sex),
+                    flc=as.numeric(flchain$kappa)+as.numeric(flchain$lambda),
+                    #batch=as.factor(flchain$flc.grp),
+                    creatine=as.numeric(flchain$creatinine),
+                    mgus=as.numeric(flchain$mgus),
+                    time=as.numeric(flchain$futime),
+                    status=as.numeric(flchain$death)
+)
 
@@
 
-saveRDS(res$test,"dialysisTest.rds")
+##saveRDS(res$test,"flchainTest.rds")
 
@@
 }
-saveRDS(klist,"dialysisCV.rds")
+##saveRDS(klist,"flchainCV.rds")
 
@@
 
-klist<-readRDS("dialysisCV.rds")
+klist<-readRDS("flchainCV.rds")
 
@@
 alp=c(0,0.5,1)
+
 # 
@@
     paramTrials$runtime[i]<-mean(runTime)
-    #saveRDS(paramTrials,"complexds.rds")
+    ##saveRDS(paramTrials,"complexds.rds")
     print(paste("trial",i))
@@
 
-klist<-readRDS("dialysisCV.rds")
-test<-readRDS("dialysisTest.rds")
+klist<-readRDS("flchainCV.rds")
+test<-readRDS("flchainTest.rds")
 source_python('src/dsurv.py')
@@
 
-klist<-readRDS("dialysisCV.rds")
-test<-readRDS("dialysisTest.rds")
+klist<-readRDS("flchainCV.rds")
+test<-readRDS("flchainTest.rds")
 times<-seq(from=min(test$time)+0.000001,
@@
     paramTrials$runtime[i]<-mean(runTime)
-    #saveRDS(paramTrials,"complexds.rds")
+    ##saveRDS(paramTrials,"complexds.rds")
     print(paste("trial",i,"out of ",nrow(paramTrials)))
@@
 
-klist<-readRDS("dialysisCV.rds")
-test<-readRDS("dialysisTest.rds")
+klist<-readRDS("flchainCV.rds")
+test<-readRDS("flchainTest.rds")
 
@@
 
-saveRDS(winners,"dialysisWins.rds")
+##saveRDS(winners,"flchainWins.rds")
 
@@
 
-winners<-readRDS("dialysisWins.rds")
+winners<-readRDS("flchainWins.rds")
 
-klist<-readRDS("dialysisCV.rds")
+klist<-readRDS("flchainCV.rds")
 data<-rbind(klist[[1]]$train,klist[[1]]$val)
@@
 cScore<-list()
+AUClist<-list()
 pos<-1
@@
 
-  fullTest<-readRDS("dialysisTest.rds")
+  fullTest<-readRDS("flchainTest.rds")
   test<-fullTest[,-c(ncol(data))]
@@
              to=max(test$time),
-             length.out = 20
+             length.out = 82
   )
@@
                              formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)
+                             se.fit = FALSE, metrics = c("auc", "brier"), contrasts = FALSE, times = times)
   BrierIPAList[[pos]]<-brierFinalResults$Brier$score
+  AUClist[[pos]]<-brierFinalResults$AUC$score
 
@@
   print(paste("current iter:",pos))
-  saveRDS(BrierIPAList,'results/100dialysisBrier.rds')
-  saveRDS(cScore,'results/100dialysiscidx.rds')
+  saveRDS(BrierIPAList,'results/100flchainBrier.rds')
+  saveRDS(cScore,'results/100flchaincidx.rds')
+  saveRDS(AUClist,'results/100flchainAuc.rds')
   pos=pos+1
@@
 
-```{r plotdialysis}
+```{r plotflchain}
 
-BrierIPAList<-readRDS('results/100dialysisBrier.rds')
+BrierIPAList<-readRDS('results/100flchainBrier.rds')
 
@@
 
-ipa+coord_cartesian(ylim=c(-0.1,0.2))
+ipa+coord_cartesian(ylim=c(-0.1,0.5))
 #end_time-start_time
+
+```
+
+```{r auc}
+
+AUClist<-readRDS('results/100flchainAuc.rds')
+aucSims=data.frame(model=AUClist[[1]]$model,times=AUClist[[1]]$times)
+for(sim in AUClist){
+   aucSims<-cbind(aucSims,sim$AUC)
+
+}
+aucSummary<-statCalcSim(aucSims)
+auc<-plotPerformance(aucSummary,method="AUC",iterations=iteration )
+
+auc
 
@@
 
-cidxList<-readRDS('results/100dialysiscidx.rds')
+cidxList<-readRDS('results/100flchaincidx.rds')
 

### 2023-04-05 analyses/optimMyeloma.Rmd
File Changes

--- 2023-04-05 analyses/optimFlchain.Rmd
@@
 ---
-title: "flchain"
+title: "myeloma"
 output: html_document
@@
 ```{r setup, include=T}
-set.seed(1)
+set.seed(12)
 knitr::opts_chunk$set(echo = TRUE)
@@
 library(reticulate)
-eval_flchain<-T
+eval_myeloma<-T
 #source_python('src/getSUP.py')
@@
 
-reoptimize=T
+reoptimize=FALSE
 
@@
 
-```{r flchainPrep,eval=FALSE,hide=FALSE}
+```{r myelomaPrep,eval=FALSE,hide=FALSE}
 
 library(survival)
-flchain<-flchain
 
-preData<-data.frame(age=as.numeric(flchain$age),
-                    #sampyr=as.numeric(flchain$sample.yr),
-                    sex=as.factor(flchain$sex),
-                    flc=as.numeric(flchain$kappa)+as.numeric(flchain$lambda),
-                    #batch=as.factor(flchain$flc.grp),
-                    creatine=as.numeric(flchain$creatinine),
-                    mgus=as.numeric(flchain$mgus),
-                    time=as.numeric(flchain$futime),
-                    status=as.numeric(flchain$death)
+data("cancer", package="survival")
+
+preData<-data.frame(year=as.numeric(myeloma$year),
+                    entry=as.numeric(myeloma$entry),
+                    time=as.numeric(myeloma$futime),
+                    status=as.numeric(myeloma$death)
 )
@@
 
-##saveRDS(res$test,"flchainTest.rds")
+#saveRDS(res$test,"myelomaTest.rds")
 
@@
 }
-##saveRDS(klist,"flchainCV.rds")
+#saveRDS(klist,"myelomaCV.rds")
 
@@
 
-klist<-readRDS("flchainCV.rds")
+klist<-readRDS("myelomaCV.rds")
 
@@
 alp=c(0,0.5,1)
-
 # 
@@
 
-klist<-readRDS("flchainCV.rds")
-test<-readRDS("flchainTest.rds")
+klist<-readRDS("myelomaCV.rds")
+test<-readRDS("myelomaTest.rds")
 source_python('src/dsurv.py')
@@
 
-klist<-readRDS("flchainCV.rds")
-test<-readRDS("flchainTest.rds")
+klist<-readRDS("myelomaCV.rds")
+test<-readRDS("myelomaTest.rds")
 times<-seq(from=min(test$time)+0.000001,
@@
 
-klist<-readRDS("flchainCV.rds")
-test<-readRDS("flchainTest.rds")
+klist<-readRDS("myelomaCV.rds")
+test<-readRDS("myelomaTest.rds")
 
@@
 
-##saveRDS(winners,"flchainWins.rds")
+#saveRDS(winners,"myelomaWins.rds")
 
@@
 
-winners<-readRDS("flchainWins.rds")
+winners<-readRDS("myelomaWins.rds")
 
-klist<-readRDS("flchainCV.rds")
+klist<-readRDS("myelomaCV.rds")
 data<-rbind(klist[[1]]$train,klist[[1]]$val)
@@
 
-  fullTest<-readRDS("flchainTest.rds")
+  fullTest<-readRDS("myelomaTest.rds")
+
+  if(sum(fullTest$time>max(train$time))>0){next }
+
   test<-fullTest[,-c(ncol(data))]
@@
   #valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
-  times<-seq(from=min(test$time),
-             to=max(test$time),
+  times<-seq(from=min(train$time),
+             to=max(train$time),
              length.out = 82
@@
                              formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                             se.fit = FALSE, metrics = c("auc", "brier"), contrasts = FALSE, times = times)
+                             se.fit = FALSE, metrics =c("auc", "brier"), contrasts = FALSE, times = times)
   BrierIPAList[[pos]]<-brierFinalResults$Brier$score
@@
   et_test<-as.matrix(fullTest[,c(ncol(fullTest),ncol(fullTest)-1)])
+
   colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
@@
   )
+
   print(paste("current iter:",pos))
-  saveRDS(BrierIPAList,'results/100flchainBrier.rds')
-  saveRDS(cScore,'results/100flchaincidx.rds')
-  saveRDS(AUClist,'results/100flchainAuc.rds')
+  saveRDS(BrierIPAList,'results/100myelomaBrier.rds')
+  saveRDS(cScore,'results/100myelomacidx.rds')
+  saveRDS(AUClist,'results/100myelomaAuc.rds')
   pos=pos+1
@@
 
-```{r plotflchain}
+```{r plotmyeloma}
 
-BrierIPAList<-readRDS('results/100flchainBrier.rds')
+BrierIPAList<-readRDS('results/100myelomaBrier.rds')
+toRemove<-c()
+cleaned<-list()
+for(i in 1:length(BrierIPAList)){
+  if(nrow(BrierIPAList[[i]])>110){
+    cleaned[[length(cleaned)+1]]<-BrierIPAList[[i]]
+  }
+}
+
+BrierIPAList<-cleaned
 
@@
 
-ipa+coord_cartesian(ylim=c(-0.1,0.5))
+ipa+coord_cartesian(ylim=c(-0.8,0.4))
 #end_time-start_time
@@
 
-AUClist<-readRDS('results/100flchainAuc.rds')
+AUClist<-readRDS('results/100myelomaAuc.rds')
 aucSims=data.frame(model=AUClist[[1]]$model,times=AUClist[[1]]$times)
@@
 
-cidxList<-readRDS('results/100flchaincidx.rds')
+cidxList<-readRDS('results/100myelomacidx.rds')
 
@@
 
+#https://scikit-survival.readthedocs.io/en/stable/user_guide/evaluating-survival-models.html
 ```

### 2023-04-05 analyses/optimOldmort.Rmd
File Changes

--- 2023-04-05 analyses/optimDialysis.Rmd
@@
 ---
-title: "dialysis"
+title: "oldmort"
 output: html_document
@@
 library(reticulate)
-eval_dialysis<-T
+eval_oldmort<-T
 #source_python('src/getSUP.py')
@@
 
-```{r dialysisPrep,eval=FALSE,hide=FALSE}
+```{r oldmortPrep,eval=FALSE,hide=FALSE}
 
-dialysis<-readRDS("dialysis.rds")
+library(eha)
 
-data <- dialysis[, c(1,2,3,4,7,6,5)]
-colnames(data)[7]<-"status"
+data("oldmort")
 
-#RcmdrPlugin.survival/html/Dialysis.html
+oldmort<-oldmort
 
-smallCenters<-as.numeric(names(which( table(data$center)<30)))
-toRemove<-c()
-for(i in 1:nrow(data)){
-  if(data$center[i] %in% smallCenters){
-    toRemove<-append(toRemove,i)
+omor=as.data.frame(matrix(NA,nrow=length(unique(oldmort$id)),ncol=ncol(oldmort)))
+colnames(omor)<-colnames(oldmort)
+unique(oldmort$id)[1]
+j<-1
+for(i in 1:length(unique(oldmort$id))){
+  temp<-which(oldmort$id %in% unique(oldmort$id)[i])
+  if(length(temp)==1){
+    omor[j,]<-oldmort[temp,]
+    j<-j+1
+  }else if(length(temp)>1){
+    tempData<-oldmort[which(oldmort$id %in% unique(oldmort$id)[i]),]
+    omor[j,]<-tempData[which.min(tempData$enter),]
+    omor[j,]$exit<-max(tempData$exit)
+    omor[j,]$event<-sum(tempData$event)>0
+    j<-j+1
   }
@@
 }
-#data<-data[-toRemove,]
 
-preData<-data.frame(#center=as.factor(data$center),
-                    age=as.numeric(data$age),
-                    #begin=as.numeric(data$begin),
-                    disease=as.factor(data$disease),
-                    time=data$time,
-                    status=data$status)
+oldmort<-omor
+preData<-data.frame(sex=as.factor(oldmort$sex),
+                    civ=as.factor(oldmort$civ),
+                    ses=as.factor(oldmort$ses.50),
+                    birthplace=as.factor(oldmort$birthplace),
+                    ageAtEntry=oldmort$enter,
+                    imrAtBirth=oldmort$imr.birth,
+                    region=as.factor(oldmort$region),
+                    birthdate=oldmort$birthdate,
+                    time=oldmort$exit-oldmort$enter,
+                    status=as.numeric(oldmort$event))
 
@@
 
-saveRDS(res$test,"dialysisTest.rds")
+saveRDS(res$test,"oldmortTest.rds")
 
@@
 }
-saveRDS(klist,"dialysisCV.rds")
+saveRDS(klist,"oldmortCV.rds")
 
@@
 
-klist<-readRDS("dialysisCV.rds")
+klist<-readRDS("oldmortCV.rds")
 
@@
       ctrain<-train
-      #ctrain$time<- train$time + runif(nrow(train),-0.000001,0.000001)
+     ctrain$time<- train$time + runif(nrow(train),-0.000001,0.000001)
 
@@
     paramTrials$runtime[i]<-mean(runTime)
-    #saveRDS(paramTrials,"complexds.rds")
+    saveRDS(paramTrials,"complexds.rds")
     print(paste("trial",i))
@@
 
-klist<-readRDS("dialysisCV.rds")
-test<-readRDS("dialysisTest.rds")
+klist<-readRDS("oldmortCV.rds")
+test<-readRDS("oldmortTest.rds")
 source_python('src/dsurv.py')
@@
 
-klist<-readRDS("dialysisCV.rds")
-test<-readRDS("dialysisTest.rds")
+klist<-readRDS("oldmortCV.rds")
+test<-readRDS("oldmortTest.rds")
 times<-seq(from=min(test$time)+0.000001,
@@
     paramTrials$runtime[i]<-mean(runTime)
-    #saveRDS(paramTrials,"complexds.rds")
+    saveRDS(paramTrials,"complexds.rds")
     print(paste("trial",i,"out of ",nrow(paramTrials)))
@@
 
-klist<-readRDS("dialysisCV.rds")
-test<-readRDS("dialysisTest.rds")
+klist<-readRDS("oldmortCV.rds")
+test<-readRDS("oldmortTest.rds")
 
@@
 
-saveRDS(winners,"dialysisWins.rds")
+saveRDS(winners,"oldmortWins.rds")
 
@@
 
-winners<-readRDS("dialysisWins.rds")
+winners<-readRDS("oldmortWins.rds")
 
-klist<-readRDS("dialysisCV.rds")
+klist<-readRDS("oldmortCV.rds")
 data<-rbind(klist[[1]]$train,klist[[1]]$val)
@@
 
-  fullTest<-readRDS("dialysisTest.rds")
+  fullTest<-readRDS("oldmortTest.rds")
   test<-fullTest[,-c(ncol(data))]
@@
       ctrain<-train
-  ctrain$time<- train$time
+      ctrain$time<- train$time + runif(nrow(train),-0.000001,0.000001)
 
@@
   print(paste("current iter:",pos))
-  saveRDS(BrierIPAList,'results/100dialysisBrier.rds')
-  saveRDS(cScore,'results/100dialysiscidx.rds')
+  saveRDS(BrierIPAList,'results/100oldmortBrier.rds')
+  saveRDS(cScore,'results/100oldmortcidx.rds')
   pos=pos+1
@@
 
-```{r plotdialysis}
+```{r plotoldmort}
 
-BrierIPAList<-readRDS('results/100dialysisBrier.rds')
+BrierIPAList<-readRDS('results/100oldmortBrier.rds')
 
@@
 
-cidxList<-readRDS('results/100dialysiscidx.rds')
+cidxList<-readRDS('results/100oldmortcidx.rds')
 

### 2023-04-05 analyses/optimProstate.Rmd
File Changes

--- 2023-04-05 analyses/optimFlchain.Rmd
@@
 ---
-title: "flchain"
+title: "prostate"
 output: html_document
@@
 library(reticulate)
-eval_flchain<-T
+eval_prostate<-T
 #source_python('src/getSUP.py')
@@
 
-reoptimize=T
+reoptimize=F
 
@@
 
-```{r flchainPrep,eval=FALSE,hide=FALSE}
+```{r prostatePrep,eval=FALSE,hide=FALSE}
 
-library(survival)
-flchain<-flchain
+library(asaur)
+prostate<-prostateSurvival
 
-preData<-data.frame(age=as.numeric(flchain$age),
-                    #sampyr=as.numeric(flchain$sample.yr),
-                    sex=as.factor(flchain$sex),
-                    flc=as.numeric(flchain$kappa)+as.numeric(flchain$lambda),
-                    #batch=as.factor(flchain$flc.grp),
-                    creatine=as.numeric(flchain$creatinine),
-                    mgus=as.numeric(flchain$mgus),
-                    time=as.numeric(flchain$futime),
-                    status=as.numeric(flchain$death)
+prostate<-prostate[-c(which(prostate$status==2)),]
+
+preData<-data.frame(age=as.factor(prostate$ageGroup),
+                    grade=as.factor(prostate$grade),
+                    stage=as.factor(prostate$stage),
+                    time=as.numeric(prostate$survTime),
+                    status=as.numeric(prostate$status)
 )
@@
 
-##saveRDS(res$test,"flchainTest.rds")
+#saveRDS(res$test,"prostateTest.rds")
 
@@
 }
-##saveRDS(klist,"flchainCV.rds")
+#saveRDS(klist,"prostateCV.rds")
 
@@
 
-klist<-readRDS("flchainCV.rds")
+klist<-readRDS("prostateCV.rds")
 
-# lr=c(0.001)
-# drpt=c(0.25,0.5)
-# layer1=c(10,20,30)
-# layer2=c(10,20,30)
-# bsize=c(floor(nrow(klist[[1]]$train)/100))
-# actv=c("relu","linear")#,"tanh"
-# alp=c(0)
+# 
+# lr=c(0.001,0.0005)
+# drpt=c(0.25,0.1)
+# layer1=c(20,50,100)
+# layer2=c(25,50)
+# nBatch=c(100,500,1000)
+# actv=c("relu","linear")#
+# alp=c(0.5)
 
@@
 
-# 
-# lr=c(0.001)
-# drpt=c(0.25)
-# layer1=c(20)
-# layer2=c(10)
-# nBatch=c(1000,100)
-# actv=c("relu")#
-# alp=c(0)
-
+#curl -d "support done" ntfy.sh/updatejykiscripts
 ```
@@
     paramTrials$runtime[i]<-mean(runTime)
-    ##saveRDS(paramTrials,"complexds.rds")
+    #saveRDS(paramTrials,"complexds.rds")
     print(paste("trial",i))
@@
 
-klist<-readRDS("flchainCV.rds")
-test<-readRDS("flchainTest.rds")
+klist<-readRDS("prostateCV.rds")
+test<-readRDS("prostateTest.rds")
 source_python('src/dsurv.py')
@@
 
-klist<-readRDS("flchainCV.rds")
-test<-readRDS("flchainTest.rds")
+klist<-readRDS("prostateCV.rds")
+test<-readRDS("prostateTest.rds")
 times<-seq(from=min(test$time)+0.000001,
@@
     paramTrials$runtime[i]<-mean(runTime)
-    ##saveRDS(paramTrials,"complexds.rds")
+    #saveRDS(paramTrials,"complexds.rds")
     print(paste("trial",i,"out of ",nrow(paramTrials)))
@@
 
-klist<-readRDS("flchainCV.rds")
-test<-readRDS("flchainTest.rds")
+klist<-readRDS("prostateCV.rds")
+test<-readRDS("prostateTest.rds")
 
 cbnnPerformance<-optimCBNN(klist,test,times,lr,drpt,layer1,layer2,nBatch,actv)
+
 #cbnnPerformance<-cbnnPerformance[which.min(cbnnPerformance$valLoss),]
+
 ```
@@
 
-##saveRDS(winners,"flchainWins.rds")
+#saveRDS(winners,"prostateWins.rds")
 
@@
 
-winners<-readRDS("flchainWins.rds")
+winners<-readRDS("prostateWins.rds")
 
-klist<-readRDS("flchainCV.rds")
+klist<-readRDS("prostateCV.rds")
 data<-rbind(klist[[1]]$train,klist[[1]]$val)
@@
 
-  fullTest<-readRDS("flchainTest.rds")
+  fullTest<-readRDS("prostateTest.rds")
   test<-fullTest[,-c(ncol(data))]
@@
   print(paste("current iter:",pos))
-  saveRDS(BrierIPAList,'results/100flchainBrier.rds')
-  saveRDS(cScore,'results/100flchaincidx.rds')
-  saveRDS(AUClist,'results/100flchainAuc.rds')
+  saveRDS(BrierIPAList,'results/100prostateBrier.rds')
+  saveRDS(cScore,'results/100prostatecidx.rds')
+  saveRDS(AUClist,'results/100prostateAuc.rds')
   pos=pos+1
@@
 
-```{r plotflchain}
+```{r plotprostate}
 
-BrierIPAList<-readRDS('results/100flchainBrier.rds')
+#BrierIPAList<-readRDS('results/100prostateBrier.rds')
 
@@
 
-AUClist<-readRDS('results/100flchainAuc.rds')
+AUClist<-readRDS('results/100prostateAuc.rds')
 aucSims=data.frame(model=AUClist[[1]]$model,times=AUClist[[1]]$times)
@@
 
-cidxList<-readRDS('results/100flchaincidx.rds')
+cidxList<-readRDS('results/100prostatecidx.rds')
 

### 2023-04-05 analyses/optimizeComplex.Rmd
File Changes

@@
 params:
-  epo: 200
-  patience: 5
-  iteration: 2
+  epo: 2000
+  patience: 10
+  iteration: 100
   layer1: 12
@@
 #bsize=512
-min_delta =0# 10^-7
+min_delta = 0 #10^-7
 
@@
 
-numSamp<-2000
+numSamp<-5000
 
-baselineTrials<-2
+#baselineTrials<-2
 
@@
 
-```{r complexPrep,eval=T}
+```{r complexPrep,eval=FALSE,hide=FALSE}
 
@@
     betas[["z3"]] * x[["z3"]]+
-    betas[["term1"]] * x[["z1"]]*t+#*x[["z2"]]*
+    betas[["term1"]] * x[["z1"]]*t+
+    #betas[["term2"]] * x[["z1"]]*x[["z2"]]*t*t+
+
     #betas[["term2"]] * x[["z1"]]*x[["z3"]]+
@@
 }
-url<-'app.txt'
-ds <- read.table(url, sep = '',header = TRUE)
-ds$years.at.risk=pmin(ds$agein80,ds$onset)
-brcancer$rectime<-brcancer$rectime/max(brcancer$rectime)
+#url<-'app.txt'
+#ds <- read.table(url, sep = '',header = TRUE)
+#ds$years.at.risk=pmin(ds$agein80,ds$onset)
+#brcancer$rectime<-brcancer$rectime
 true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1, data = brcancer, k = 3) #from flexsurv
@@
 coefficients<-true_mod$coefficients
-coefficients<-c(coefficients,z1=0.05,z2=1.05,z3=1.05,term1=2.4,term3=1.05)#term2=1.05,
+#coefficients<-c(coefficients,z1=-10,z2=1,z3=1,term1=0.005,term3=1)
+coefficients<-c(coefficients,z1=-5,z2=-1,z3=1,term1=0.001,term3=-1)#z1 time interaction
 
@@
                 z1=c(rep(1,numSamp/2),rep(0,numSamp/2)),
-                z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
-                z3=c(rnorm(numSamp/2,0,0.5),rnorm(numSamp/2,10,0.5)))
+                z2=c(rnorm(numSamp/2,1,0.5),rnorm(numSamp/2,0,0.5)),
+                z3=c(rnorm(numSamp,1,0.5)))
+
 true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1,
@@
                logcumhazard = logcumhaz,  # definition of log cum hazard
-               maxt = 25,               # no right-censoring
-               interval = c(0,10000000)) # interval for root finding
+               maxt = 10000,               # no right-censoring
+               interval = c(0,100000000)) # interval for root finding
 colnames(dat)<-c("id","time","status")
+sum(dat$status)
+
 samp<-sample(seq(1,nrow(dat),by=1), floor(nrow(dat)*(0.1))) #random censoring
@@
 data <- merge(cov, dat)
+
 covs<-colnames(data)
 data<-data[,-1]
-
-toDiv<-max(data$time)
-data$time<-data$time/toDiv
+hist(data$time[which(data$z1==0)],breaks=100)
+hist(data$time[which(data$z1==1)],breaks=100)
+median(data$time[which(data$z1==0)])
+median(data$time[which(data$z1==1)])
+#data$zsplit<-NULL
+#toDiv<-max(data$time)
+#data$time<-data$time/toDiv
 spec = c(train = .85, test = .15)
@@
 res = split(data, g)
-saveRDS(res$test,"ComplexSimTest.rds")
+#saveRDS(res$test,"ComplexSimTest.rds")
 
-k=5
+k=3
 klist<-list()
@@
 }
-saveRDS(klist,"ComplexSimCV.rds")
+#saveRDS(klist,"ComplexSimCV.rds")
 
@@
 
-```{r setFixedHyperparameters}
+```{r setFixedHyperparameters,eval=FALSE,hide=FALSE}
 
 klist<-readRDS("ComplexSimCV.rds")
+
 # lr=c(0.001,0.01)
-# drpt=c(0.25,0.5)
-# layer1=c(50,75,100)
-# layer2=c(50,75)
+# drpt=c(0.1,0.25)
+# layer1=c(25,50,100)
+# layer2=c(25,50,100)
 # bsize=c(floor(nrow(klist[[1]]$train)/100))
-# actv=c("relu","tanh")
+# actv=c("relu","tanh")#
 # alp=c(0,0.5,1)
 
-lr=c(0.001,0.0001)
-drpt=c(.1,.2)
-layer1=c(10,25,50)
-layer2=c(10,20,30)
-bsize=c(floor(nrow(klist[[1]]$train)/100))
-actv=c("relu","tanh")
+lr=c(0.001,0.01)
+drpt=c(0.01,0.05,0.1)
+layer1=c(50,75,100)
+layer2=c(10,25,50)
+nBatch=c(100,500)
+actv=c("relu","linear")#
 alp=c(0,0.5,1)
+# 
+# lr=c(0.001)
+# drpt=c(0.01)
+# layer1=c(50)
+# layer2=c(10)
+# nBatch=c(100)
+# actv=c("relu")#
+# alp=c(0)
 
@@
 
-```{r hyperOpDsurv,eval=T}
+```{r hyperOpDsurv,eval=FALSE,hide=FALSE}
 
-optimDeepSurv<-function(klist,test,times,baselineTrials,lr,drpt,layer1,layer2,bsize,actv){
+optimDeepSurv<-function(klist,test,baselineTrials,lr,drpt,layer1,layer2,nBatch,actv){
 
@@
     for(data in klist){
-
+      currentTime<-Sys.time() 
       train<-data$train
       val<-data$val
-
+      maxTime<-max(train$time)
       sds<-sapply(train, function(x) (sd(x)))
       means<-sapply(train, function(x) (mean(x)))
-      val<-normalizer(val,means,sds)
-      train<-normalizer(train,means,sds)
+      val<-normalizer(val,means,sds,maxTime)
+      train<-normalizer(train,means,sds,maxTime)
+      times<-seq(from=min(val$time),
+                 to=max(val$time),
+                 length.out = 22
+      )
+
+      times<-head(times, -1)
+      times<-tail(times, -1)
 
@@
       #=paramTrials[i,]$valLoss
-      source_python('src/optimDSurv.py')
+      source_python('src/dsurv.py')
 
-      currentTime<-Sys.time() 
-      coxnnsurv=fitDeepSurv(train,val,floor(nrow(train)/100),epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
-      tempVal<-append(tempVal,min(coxnnsurv$val_metrics$scores$loss$score))
+      ctrain<-train
+      #ctrain$time<- train$time + runif(nrow(train),-0.000001,0.000001)
+
+      coxnnsurv=fitDeepSurv(ctrain,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
+
+      coxnnsurv<-t(coxnnsurv)
+      cnnCleaned<-pyProcess(coxnnsurv,times=times)
+      colnames(cnnCleaned)<-times
+      py_run_string("del fitDeepSurv")
+      rm(fitDeepSurv)
+
+      miniee <- Score(list('mod'=cnnCleaned),
+                      data =val, 
+                      formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
+                      se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
+
+      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
       runTime<-append(runTime, Sys.time()-currentTime)
@@
       # cutted<-brierFinalResults[which(brierFinalResults$model=="ds"),]
-      #tempRes<-append(tempRes,mean(cutted[-c(1,2,24,25)]$IPA))
+      # tempRes<-append(tempRes,mean(cutted[-c(1,2,24,25)]$IPA))
     }
@@
 
-bestDS<-optimDeepSurv(klist,test,times,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,bsize,actv)
+bestDS<-optimDeepSurv(klist,test,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,nBatch,actv)
+
 ```
 
-```{r optimDeepHit,eval=T}
+```{r optimDeepHit,eval=FALSE,hide=FALSE}
 
-optimDeepHit<-function(klist,test,times,baselineTrials,lr,drpt,layer1,layer2,bsize,actv,alp){
+optimDeepHit<-function(klist,test,times,baselineTrials,lr,drpt,layer1,layer2,nBatch,actv,alp){
 
@@
       #=paramTrials[i,]$valLoss
-      source_python('src/optimDH.py')
+      source_python('src/deephitter.py')
+      hitnnSurv=fitDeephit(train,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
+      hitnnSurv<-t(hitnnSurv)
+      deephitCleaned<-pyProcess(hitnnSurv,times=times)
+      colnames(deephitCleaned)<-times
+      py_run_string("del fitDeephit")
+      rm(fitDeephit)
 
-      currentTime<-Sys.time() 
-      hitnnSurv=fitDeephit(train,val,floor(nrow(train)/100),epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
-      tempVal<-append(tempVal,min(hitnnSurv$val_metrics$scores$loss$score))
+      miniee <- Score(list('mod'=deephitCleaned),
+                      data =val, 
+                      formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
+                      se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
+
+      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
       runTime<-append(runTime, Sys.time()-currentTime)
@@
 
-```{r hyperopcbnn,eval=T}
+```{r hyperopcbnn,eval=FALSE,hide=FALSE}
 
-optimCBNN<-function(klist,test,times,lr,drpt,layer1,layer2,bsize,actv){
+optimCBNN<-function(klist,test,times,lr,drpt,layer1,layer2,nBatch,actv){
 
@@
       #val[[1]][[1]]<-as.matrix(normalizer(as.data.frame(val[[1]][[1]]),means,sds))
-      train<-normalizer(train,means,sds)
+      train<-normalizer(train,means,sds,maxTime)
+
       #train[,ncol(train)]<-data$train[,ncol(train)]
-      val<-normalizer(val,means,sds)
+      val<-normalizer(val,means,sds,maxTime)
+      valTest<-val
+      times<-seq(from=min(val$time),
+                 to=max(val$time),
+                 length.out = 22
+      )
+
+      times<-head(times, -1)
+      times<-tail(times, -1)
+
       train<-sampleCaseBase(data = as.data.frame(train),time="time",event="status",ratio=100)
@@
                         patience=patience,val=val)
-      tempVal<-append(tempVal,min(fit$resultOfFit$metrics$val_loss))
+      #tempVal<-append(tempVal,min(fit$resultOfFit$metrics$val_loss))
+      annPreds<-aar(fit,
+                    times=times,
+                    x_test=valTest[,-c(ncol(valTest))]
+      )
+
+      rownames(annPreds)<-annPreds[,1]
+      annProperpoly<- t(annPreds[,-1])
+      class(annProperpoly)<-c("tunnel",class(annProperpoly))
+
+      miniee <- Score(list('mod'=annProperpoly),
+                      data =valTest, 
+                      formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
+                      se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
+
+      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
+
       runTime<-append(runTime, Sys.time()-currentTime)
@@
 
-cbnnPerformance<-optimCBNN(klist,test,times,lr,drpt,layer1,layer2,bsize,actv)
+cbnnPerformance<-optimCBNN(klist,test,times,lr,drpt,layer1,layer2,nBatch,actv)
+
 #cbnnPerformance<-cbnnPerformance[which.min(cbnnPerformance$valLoss),]
@@
 
-```{r}
+```{r,eval=FALSE,hide=FALSE}
 
@@
 
-saveRDS(winners,"complexOptimWins.rds")
+#saveRDS(winners,"complexOptimWins.rds")
 ```
 
-```{r bestFits,finalFit}
+```{r bestFits,eval=T}
 
@@
 BrierIPAList<-list()
+AUClist<-list()
 cScore<-list()
@@
   means<-sapply(train, function(x) (mean(x)))
+  maxTime<-max(train$time)
 
@@
   test<-fullTest[,-c(ncol(data))]
-  test<-normalizer(fullTest,means,sds)[,-c(ncol(data))]
-  fullTest<-normalizer(fullTest,means,sds)
+  test<-normalizer(fullTest,means,sds,maxTime)[,-c(ncol(data))]
+  fullTest<-normalizer(fullTest,means,sds,maxTime)
 
-  val<-normalizer(res$validate,means,sds)
+  val<-normalizer(res$validate,means,sds,maxTime)
 
-  train<-normalizer(train,means,sds)
-  valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)#rbind(train,val)
-  valcb$offset<-0
-  valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
-  times<-seq(from=min(test$time)+0.000001,
-             to=max(test$time)-0.000001,
-             length.out = 27
+  train<-normalizer(train,means,sds,maxTime)
+  linModelTrain<-as.data.frame(rbind(train,val))
+  #valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)#rbind(train,val)
+  #valcb$offset<-0
+  #valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
+  times<-seq(from=min(test$time),
+             to=max(test$time),
+             length.out = 82
   )
@@
   ###########################
-  final_mod_cb_glm <- fitSmoothHazard(status~bs(time)-time+z1+z2+z3+z1:time+z2:z3,#+z1:z3
-                                      data = train,
+  final_mod_cb_glm <- fitSmoothHazard(status~bs(time)-time+z1+z2+z3+z1:time+ z2:z3,#+z1:z3
+                                      data = linModelTrain,
                                       time = "time",
@@
   mod_cb_glm <- fitSmoothHazard(status~bs(time)+.-time,
-                                data = train,
+                                data = linModelTrain,
                                 time = "time",
@@
   layer2=winners$cbnn$layer2
-  #bsize=winners$cbnn$bsize
+  bsize=c(floor(nrow(train)/winners$cbnn$nBatch))
   actv=winners$cbnn$actv
@@
   #  drpt=0.1
-  # layer1=50
-  # layer2=20
-  # #bsize=winners$cbnn$bsize
+  #  layer1=10
+  #  layer2=10
+  # bsize=c(floor(nrow(train)/100))# bsize=#winners$cbnn$bsize
   #  actv='relu'
@@
                                   'DeepSurv'=cnnCleaned,'DeepHit'=deephitCleaned,
-                                  'CBNN_Poly'=annProperpoly,'optimal'=trueGlmProper),
+                                  'CBNN_Poly'=annProperpoly,'Optimal'=trueGlmProper),
                              data =fullTest, 
                              formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)
+                             se.fit = FALSE, metrics = c("auc", "brier"), contrasts = FALSE, times = times)
   BrierIPAList[[pos]]<-brierFinalResults$Brier$score
+  AUClist[[pos]]<-brierFinalResults$AUC$score
+
+  #ggplot(data=brierFinalResults$Brier$score, aes(x=times*maxTime,y=IPA,col=model))+
+  #  geom_line()#+coord_cartesian(ylim=c(0.6,1))
 
@@
   ######################
+
   a<-survfit(cox, newdata=fullTest,times=times)
@@
   cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=length(riskList)+1)
-  et_train<-as.matrix(train)[,c(5,4)]
-  et_test<-as.matrix(fullTest[,c(5,4)])
+  et_train<-as.matrix(tempTrain)[,c(ncol(tempTrain),ncol(tempTrain)-1)]
+  et_test<-as.matrix(fullTest[,c(ncol(fullTest),ncol(fullTest)-1)])
+
   colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
-                          'cbnn','optimal')
+                          'cbnn',"Optimal")
   cScoreTemp[,1]<- tempTims
@@
 
+  print(paste("current iter:",pos))
   saveRDS(BrierIPAList,'results/100complexBrier.rds')
   saveRDS(cScore,'results/100complexcidx.rds')
+  saveRDS(AUClist,'results/100complexAuc.rds')
   pos=pos+1
-
+  rm(train,val,fullTest)
 }
@@
 ibsSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
+
 ibsTable<-as.data.frame(matrix(data=NA,nrow=100,ncol=length(unique(BrierIPAList[[1]]$model))))
@@
 #end_time-start_time
-ipa+coord_cartesian(ylim=c(-2,1))
+ipa+coord_cartesian(ylim=c(-1,1))
+
+```
+
+```{r auc}
+
+AUClist<-readRDS('results/100complexAuc.rds')
+aucSims=data.frame(model=AUClist[[1]]$model,times=AUClist[[1]]$times)
+for(sim in AUClist){
+   aucSims<-cbind(aucSims,sim$AUC)
+
+}
+aucSummary<-statCalcSim(aucSims)
+auc<-plotPerformance(aucSummary,method="AUC",iterations=iteration )
+
+auc
+
+```
+
+```{r}
+
+cidxList<-readRDS('results/100complexcidx.rds')
+
+temp<-melt(setDT(as.data.frame(cidxList[[1]])), id.vars = c("times"), variable.name = "model")
+cidxSims=data.frame(model=temp$model,times=temp$times)
+
+count=1
+for(sim in cidxList){
+  cidxSims<-cbind(cidxSims, melt(setDT(as.data.frame(sim)), id.vars = c("times"), variable.name = "model")$value)
+
+  count<-count+1
+}
+
+cidxSummary<-statCalcSim(cidxSims)
+cidxC<-plotPerformance(cidxSummary,method="C-Index",iterations=iteration  )
+cidxC
 
@@
                                            layer2=c(10,25,50,100),
-                                           bsize=c(floor(nrow(klist[[1]]$train)/100)),
+                                           bsize=c(floor(nrow(klist[[1]]$train)/250)),
                                            actv=c("relu","tanh"),
@@
 
-      dsmSurv<-dsmfitter(train=train,fullTest=test,bsize=floor(nrow(train)/100),epochs=20,val=val,times=times,lr=lr,actv=actv,k,dist,layer1,layer2,tPos=3,ePos=4)
+      dsmSurv<-dsmfitter(train=train,fullTest=test,bsize=floor(nrow(train)/250),epochs=20,val=val,times=times,lr=lr,actv=actv,k,dist,layer1,layer2,tPos=3,ePos=4)
       py_run_string("del dsmfitter")

### 2023-04-05 analyses/optimizeMetabric.Rmd
File Changes

@@
 
-toDiv<-max(data$time)
-data$time<-data$time/toDiv
+#toDiv<-max(data$time)
+#data$time<-data$time/toDiv
 spec = c(train = .85, test = .15)
@@
 
-k=5
+k=3
 klist<-list()
@@
 
-###these did not work for complex or support
-# lr=c(0.001,0.01)
-# drpt=c(0.1,0.25,.5)
-# layer1=c(100,200,300)
-# layer2=c(25,50,100,200)
-# bsize=c(floor(nrow(klist[[1]]$train)/100))
-# actv=c("relu","tanh")
-# alp=c(0,0.5,1)
-
-lr=c(0.001,0.0001)
-drpt=c(.1,.2)
-layer1=c(10,25,50)
-layer2=c(10,20,30)
-bsize=c(floor(nrow(klist[[1]]$train)/100))
-actv=c("relu","tanh")
+lr=c(0.001,0.01)
+drpt=c(0.01,0.05,0.1)
+layer1=c(50,75,100)
+layer2=c(10,25,50)
+nBatch=c(100,500)
+actv=c("relu","linear")#
 alp=c(0,0.5,1)
@@
 
-optimDeepSurv<-function(klist,test,times,baselineTrials,lr,drpt,layer1,layer2,bsize,actv){
+optimDeepSurv<-function(klist,test,baselineTrials,lr,drpt,layer1,layer2,nBatch,actv){
 
@@
 
-  #paramTrials<-paramTrials[sample(nrow(paramTrials), 2), ]
+  #paramTrials<-paramTrials[sample(nrow(paramTrials), baselineTrials), ]
 
@@
 
-bestDS<-optimDeepSurv(klist,test,times,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,bsize,actv)
+bestDS<-optimDeepSurv(klist,test,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,nBatch,actv)
 
@@
                                              layer2=layer2,
-                                             bsize=c(bsize),
+                                             nBatch=nBatch,
                                              actv=actv,
                                              alp=alp,
+                                             #IPA=NA,
+                                             #IPAse=NA,
+                                             valLoss=NA,
+                                             runtime=NA))
+
+  #paramTrials<-paramTrials[sample(nrow(paramTrials), baselineTrials), ]
+
+  # paramTrials<-unique.data.frame(paramTrials)
+  #paramTrials<-readRDS("complexTrials.rds")
+  for(i in 1:nrow(paramTrials)){
+    #paramTrials<-readRDS("complexTrials.rds")  
+    tempRes<-c()
+    tempVal<-c()
+    runTime<-c()
+
+    for(data in klist){
+      currentTime<-Sys.time() 
+      train<-data$train
+      val<-data$val
+      maxTime<-max(train$time)
+      sds<-sapply(train, function(x) (sd(x)))
+      means<-sapply(train, function(x) (mean(x)))
+      val<-normalizer(val,means,sds,maxTime)
+      train<-normalizer(train,means,sds,maxTime)
+      times<-seq(from=min(val$time),
+                 to=max(val$time),
+                 length.out = 22
+      )
+
+      times<-head(times, -1)
+      times<-tail(times, -1)
+
+      lr=paramTrials[i,]$lr
+      drpt=paramTrials[i,]$drpt
+      layer1=paramTrials[i,]$layer1
+      layer2=paramTrials[i,]$layer2
+      bsize=ceiling(nrow(train)/paramTrials[i,]$nBatch)
+      actv=paramTrials[i,]$actv
+      alp=paramTrials[i,]$alp
+      #=paramTrials[i,]$IPA
+      #=paramTrials[i,]$valLoss
+      source_python('src/deephitter.py')
+      hitnnSurv=fitDeephit(train,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
+      hitnnSurv<-t(hitnnSurv)
+      deephitCleaned<-pyProcess(hitnnSurv,times=times)
+      colnames(deephitCleaned)<-times
+      py_run_string("del fitDeephit")
+      rm(fitDeephit)
+
+      miniee <- Score(list('mod'=deephitCleaned),
+                      data =val, 
+                      formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
+                      se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
+
+      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
+      runTime<-append(runTime, Sys.time()-currentTime)
+      currentTime<-Sys.time() 
+    }
+
+    #paramTrials$IPA[i]<-mean(tempRes)
+    #paramTrials$IPAse[i]<-sd(tempRes)
+    paramTrials$valLoss[i]<-mean(tempVal)
+    paramTrials$runtime[i]<-mean(runTime)
+    #saveRDS(paramTrials,"complexds.rds")
+    print(paste("trial",i))
+    print(sum(runTime))
+    #print(as.character(paramTrials[i,]))
+    #rm(cbnn,fit,annPreds,covars_input,covars_output)
+    gc()
+  }
+  bestPerformance<-paramTrials[which.min(paramTrials$valLoss),]
+  return(bestPerformance)
+}
+
+klist<-readRDS("metaCV.rds")
+test<-readRDS("metaTest.rds")
+times<-seq(from=min(test$time)+0.000001,
+           to=max(test$time)-0.000001,
+           length.out = 25
+)
+
+bestDH<-optimDeepHit(klist,test,times,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,nBatch,actv,alp)
+
+```
+
+```{r hyperopcbnn,eval=reoptimize}
+
+optimCBNN<-function(klist,test,times,lr,drpt,layer1,layer2,nBatch,actv){
+
+  # 
+  # for (i in 1:length(klist)){
+  #   #train<-sampleCaseBase(data = as.data.frame(klist[[i]]$train),time="time",event="status",ratio=100)
+  #   valcb<-sampleCaseBase(data = as.data.frame(klist[[i]]$val),time="time",event="status",ratio=100)
+  #   valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
+  #                          as.matrix(valcb[,ncol(valcb),drop=F])),
+  #                     as.matrix(valcb[,ncol(valcb)-1,drop=F]))
+  #   #valcbtensor[[1]][[2]]<-rep(train$offset[1],nrow(valcbtensor[[1]][[1]]))      
+  #   #klist[[i]]$train<-train
+  #   klist[[i]]$val<-valcbtensor
+  # }
+
+  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
+                                             drpt=drpt,
+                                             layer1=layer1,
+                                             layer2=layer2,
+                                             nBatch=nBatch,
+                                             actv=actv,
                                              #IPA=NA,
@@
       val<-data$val
-
-      sds<-sapply(train, function(x) (sd(x)))
-      means<-sapply(train, function(x) (mean(x)))
-      val<-normalizer(val,means,sds)
-      train<-normalizer(train,means,sds)
-
-      lr=paramTrials[i,]$lr
-      drpt=paramTrials[i,]$drpt
-      layer1=paramTrials[i,]$layer1
-      layer2=paramTrials[i,]$layer2
-      bsize=paramTrials[i,]$bsize
-      actv=paramTrials[i,]$actv
-      alp=paramTrials[i,]$alp
-      #=paramTrials[i,]$IPA
-      #=paramTrials[i,]$valLoss
-      source_python('src/optimDH.py')
-
-      currentTime<-Sys.time() 
-      hitnnSurv=fitDeephit(train,val,floor(nrow(train)/100),epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
-      tempVal<-append(tempVal,min(hitnnSurv$val_metrics$scores$loss$score))
-      runTime<-append(runTime, Sys.time()-currentTime)
-      currentTime<-Sys.time() 
-    }
-    rm(fitDeephit)
-
-    #paramTrials$IPA[i]<-mean(tempRes)
-    #paramTrials$IPAse[i]<-sd(tempRes)
-    paramTrials$valLoss[i]<-mean(tempVal)
-    paramTrials$runtime[i]<-mean(runTime)
-    #saveRDS(paramTrials,"complexds.rds")
-    print(paste("trial",i))
-    print(mean(runTime))
-    #print(as.character(paramTrials[i,]))
-    #rm(cbnn,fit,annPreds,covars_input,covars_output)
-    gc()
-  }
-  bestPerformance<-paramTrials[which.min(paramTrials$valLoss),]
-  return(bestPerformance)
-}
-
-klist<-readRDS("metaCV.rds")
-test<-readRDS("metaTest.rds")
-  times<-seq(from=min(test$time)+0.000001,
-             to=max(test$time)-0.000001,
-             length.out = 25
-  )
-
-bestDH<-optimDeepHit(klist,test,times,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,bsize,actv,alp)
-
-```
-
-```{r hyperopcbnn,eval=reoptimize}
-
-optimCBNN<-function(klist,test,times,lr,drpt,layer1,layer2,bsize,actv){
-
-  # for (i in 1:length(klist)){
-  #   train<-sampleCaseBase(data = as.data.frame(klist[[i]]$train),time="time",event="status",ratio=100)
-  #   valcb<-sampleCaseBase(data = as.data.frame(klist[[i]]$val),time="time",event="status",ratio=100)
-  #   valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
-  #                          as.matrix(valcb[,ncol(valcb),drop=F])),
-  #                     as.matrix(valcb[,ncol(valcb)-1,drop=F]))
-  #   valcbtensor[[1]][[2]]<-rep(train$offset[1],nrow(valcbtensor[[1]][[1]]))      
-  #   klist[[i]]$train<-train
-  #   klist[[i]]$val<-valcbtensor
-  # }
-
-  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
-                                             drpt=drpt,
-                                             layer1=layer1,
-                                             layer2=layer2,
-                                             actv=actv,
-                                             #IPA=NA,
-                                             #IPAse=NA,
-                                             valLoss=NA,
-                                             runtime=NA))
-  #paramTrials<-paramTrials[sample(nrow(paramTrials), 2), ]
-
-  # paramTrials<-unique.data.frame(paramTrials)
-  #paramTrials<-readRDS("complexTrials.rds")
-  for(i in 1:nrow(paramTrials)){
-    #paramTrials<-readRDS("complexTrials.rds")  
-    tempRes<-c()
-    tempVal<-c()
-    runTime<-c()
-
-    for(data in klist){
-
-      train<-data$train
-      val<-data$val
-
+      maxTime<-max(train$time)
       sds<-sapply(train, function(x) (sd(x)))
@@
 
-cbnnPerformance<-optimCBNN(klist,test,times,lr,drpt,layer1,layer2,bsize,actv)
+cbnnPerformance<-optimCBNN(klist,test,times,lr,drpt,layer1,layer2,nBatch,actv)
 #cbnnPerformance<-cbnnPerformance[which.min(cbnnPerformance$valLoss),]
+
+```
+
+```{r winners}
 winners<-list(cbnn=cbnnPerformance,
@@
   train<-trainOG[samp,]
-
+    #train<-res$train
   val<-res$validate
@@
   test<-fullTest[,-c(ncol(data))]
-  test<-normalizer(fullTest,means,sds)[,-c(ncol(data))]
-  fullTest<-normalizer(fullTest,means,sds)
+  test<-normalizer(fullTest,means,sds,maxTime)[,-c(ncol(data))]
+  fullTest<-normalizer(fullTest,means,sds,maxTime)
 
-  val<-normalizer(res$validate,means,sds)
+  val<-normalizer(res$validate,means,sds,maxTime)
 
-  train<-normalizer(train,means,sds)
-  valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)#rbind(train,val)
-  valcb$offset<-0
-  valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
+  train<-normalizer(train,means,sds,maxTime)
+
+  linModelTrain<-as.data.frame(rbind(train,val))
+
   times<-seq(from=min(test$time)+0.000001,
@@
   layer2=winners$ds$layer2
-  bsize=winners$ds$bsize
+  bsize=ceiling(nrow(train)/winners$ds$nBatch)
   actv=winners$ds$actv
+
   source_python('src/dsurv.py')
-  coxnnsurv=fitDeepSurv(train,fullTest,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
+  ctrain<-train
+  ctrain$time<- train$time
+
+  coxnnsurv=fitDeepSurv(ctrain,fullTest,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
   coxnnsurv<-t(coxnnsurv)
+
   cnnCleaned<-pyProcess(coxnnsurv,times=times)
@@
   source_python('src/deephitter.py')
-  hitnnSurv=fitDeephit(train,fullTest,floor(nrow(train)/100),epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
+  hitnnSurv=fitDeephit(train,fullTest,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
   hitnnSurv<-t(hitnnSurv)
@@
   #############################
+  tempTrain<-train
+  train<-sampleCaseBase(data = as.data.frame(train),time="time",event="status",ratio=100)
+  valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)
+  val<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
+                 as.matrix(valcb[,ncol(valcb),drop=F])),
+            as.matrix(valcb[,ncol(valcb)-1,drop=F]))
 
+  val[[1]][[2]]<-rep(train$offset[1],nrow(val[[1]][[1]]))   
   lr=winners$cbnn$lr
@@
   layer2=winners$cbnn$layer2
-  #bsize=winners$cbnn$bsize
+  bsize=c(floor(nrow(train)/winners$cbnn$nBatch))
   actv=winners$cbnn$actv
+  # 
+  # 
+  #  lr=0.001
+  #  drpt=0.5
+  #  layer1=10
+  #  layer2=10
+  #  patience=10
+  #actv='linear'
 
   currentTime<-Sys.time() 
-  valcbtensor[[1]][[2]]<-rep(mod_cb_glm$offset[1],nrow(valcbtensor[[1]][[1]]))
-
   covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
@@
 
-  cbnn<-cbnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
+  cbnn<-cbnnModel(features=colnames(train)[-ncol(train)],
                   feature_input = covars_input,
                   feature_output = covars_output,
-                  originalData = mod_cb_glm$data,
-                  offset=mod_cb_glm$offset,
+                  originalData = train,
+                  offset=train$offset,
                   timeVar = "time",
@@
   )
+
   fit<-fitSmoothHaz(cbnn,
                     epochs=epo,
-                    batch_size=c(floor(nrow(mod_cb_glm$data)/100)),
+                    batch_size=bsize,
                     verbose=0,
@@
                     min_delta=min_delta,
-                    patience=patience,val=valcbtensor)
-
+                    patience=patience,val=val)
   annPreds<-aar(fit,
@@
   cScoreTemp<-matrix(NA,nrow=length(tempTims),ncol=length(riskList)+1)
-  et_train<-as.matrix(train)[,c(ncol(train),ncol(train)-1)]
-  et_test<-as.matrix(fullTest[,c(ncol(train),ncol(train)-1)])
+  et_train<-as.matrix(tempTrain)[,c(ncol(tempTrain),ncol(tempTrain)-1)]
+  et_test<-as.matrix(fullTest[,c(ncol(fullTest),ncol(fullTest)-1)])
   colnames(cScoreTemp)<-c('times', "Cox_Lin",'CB_Logi','DeepSurv','DeepHit',
@@
 ```
+
+```{r}
+
+cidxList<-readRDS('results/100metacidx.rds')
+
+temp<-melt(setDT(as.data.frame(cidxList[[1]])), id.vars = c("times"), variable.name = "model")
+cidxSims=data.frame(model=temp$model,times=temp$times)
+
+count=1
+for(sim in cidxList){
+  cidxSims<-cbind(cidxSims, melt(setDT(as.data.frame(sim)), id.vars = c("times"), variable.name = "model")$value)
+
+  count<-count+1
+}
+
+cidxSummary<-statCalcSim(cidxSims)
+cidxC<-plotPerformance(cidxSummary,method="C-Index",iterations=iteration  )
+cidxC
+
+```

### 2023-04-05 analyses/optimizeSupport.Rmd
File Changes

@@
 ---
-title: "iterSim"
+title: "support"
 output: html_document
@@
   epo: 2000
-  patience: 5
-  iteration: 10
+  patience: 10
+  iteration: 100
   layer1: 12
@@
 
-numSamp<-2000
-
 baselineTrials<-5
@@
 
-```{r supPrep,eval=T}
+```{r supPrep,eval=FALSE,hide=FALSE}
 
@@
 
-#for(i in c(2,5,6,7,16)){#3,4,
-#  data[, i]<-as.factor(data[, i])
-#}
+data<-data[-which(data$x2 %in% c(8,9)),]
+
+for(i in c(2,3,4,5,6,7)){#3,4,
+  data[, i]<-as.factor(data[, i])
+}
 data<-as.data.frame(model.matrix(~.,data)[,-1])
@@
 covs<-colnames(data)
-names(data)[ncol(data)-2] <- paste("z", ncol(data)-2, sep="")
+names(data)[seq(1,length(covs)-2)] <- paste("z", seq(1,length(covs)-2), sep="")
+
 names(data)[ncol(data)-1]<-"time"
@@
 covs<-colnames(data)
-data<-data[,-1]
 
-toDiv<-max(data$time)
-data$time<-data$time/toDiv
+#toDiv<-max(data$time)
+#data$time<-data$time/toDiv
 #data$time<-data$time/max(data$time)
@@
 
-```{r setFixedHyperparameters}
+```{r setFixedHyperparameters,eval=FALSE,hide=FALSE}
 
@@
 
-lr=c(0.001,0.0001)
-drpt=c(.1,.2)
-layer1=c(10,25,50)
-layer2=c(10,20,30)
-bsize=c(floor(nrow(klist[[1]]$train)/100))
-actv=c("relu","tanh")
-alp=c(0,0.5,1)
+# lr=c(0.001)
+# drpt=c(0.25,0.5)
+# layer1=c(10,20,30)
+# layer2=c(10,20,30)
+# bsize=c(floor(nrow(klist[[1]]$train)/100))
+# actv=c("relu","linear")#,"tanh"
+# alp=c(0)
+
+lr=c(0.001,0.0005)
+drpt=c(0.25,0.1)
+layer1=c(20,50,100)
+layer2=c(25,50)
+nBatch=c(100,500,1000)
+actv=c("relu","linear")#
+alp=c(0.5)
 
@@
 
-```{r hyperOpDsurv,eval=reoptimize}
+```{r hyperOpDsurv,eval=FALSE,hide=FALSE}
 
-optimDeepSurv<-function(klist,test,times,baselineTrials,lr,drpt,layer1,layer2,bsize,actv){
+optimDeepSurv<-function(klist,test,times,baselineTrials,lr,drpt,layer1,layer2,nBatch,actv){
 
@@
     for(data in klist){
-
+      currentTime<-Sys.time() 
       train<-data$train
       val<-data$val
-
+      maxTime<-max(train$time)
       sds<-sapply(train, function(x) (sd(x)))
       means<-sapply(train, function(x) (mean(x)))
-      val<-normalizer(val,means,sds)
-      train<-normalizer(train,means,sds)
+      val<-normalizer(val,means,sds,maxTime)
+      train<-normalizer(train,means,sds,maxTime)
+            times<-seq(from=min(val$time),
+             to=max(val$time),
+             length.out = 22
+  )
 
+  times<-head(times, -1)
+  times<-tail(times, -1)
       lr=paramTrials[i,]$lr
@@
       #=paramTrials[i,]$valLoss
-      source_python('src/optimDSurv.py')
+  source_python('src/dsurv.py')
 
-      currentTime<-Sys.time() 
-      coxnnsurv=fitDeepSurv(train,val,floor(nrow(train)/100),epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
-      tempVal<-append(tempVal,min(coxnnsurv$val_metrics$scores$loss$score))
+  coxnnsurv=fitDeepSurv(train,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
+  coxnnsurv<-t(coxnnsurv)
+  cnnCleaned<-pyProcess(coxnnsurv,times=times)
+  colnames(cnnCleaned)<-times
+  py_run_string("del fitDeepSurv")
+  rm(fitDeepSurv)
+
+    miniee <- Score(list('mod'=cnnCleaned),
+                             data =val, 
+                             formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
+                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
+
+      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
       runTime<-append(runTime, Sys.time()-currentTime)
@@
       # py_run_string("del fitDeepSurv")
-      rm(fitDeepSurv)
+     # rm(fitDeepSurv)
       # brierFinalResults <- Score(list('ds'=cnnCleaned),
@@
 
-bestDS<-optimDeepSurv(klist,test,times,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,bsize,actv)
+bestDS<-optimDeepSurv(klist,test,times,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,nBatch,actv)
 
@@
 
-```{r optimDeepHit,eval=reoptimize}
+```{r optimDeepHit,eval=FALSE,hide=FALSE}
 
-optimDeepHit<-function(klist,test,times,baselineTrials,lr,drpt,layer1,layer2,bsize,actv,alp){
+optimDeepHit<-function(klist,test,times,baselineTrials,lr,drpt,layer1,layer2,nBatch,actv,alp){
 
@@
                                              layer2=layer2,
-                                             bsize=c(bsize),
+                                             nBatch=nBatch,
                                              actv=actv,
                                              alp=alp,
+                                             #IPA=NA,
+                                             #IPAse=NA,
+                                             valLoss=NA,
+                                             runtime=NA))
+
+  #paramTrials<-paramTrials[sample(nrow(paramTrials), baselineTrials), ]
+
+  # paramTrials<-unique.data.frame(paramTrials)
+  #paramTrials<-readRDS("complexTrials.rds")
+  for(i in 1:nrow(paramTrials)){
+    #paramTrials<-readRDS("complexTrials.rds")  
+    tempRes<-c()
+    tempVal<-c()
+    runTime<-c()
+
+    for(data in klist){
+      currentTime<-Sys.time() 
+      train<-data$train
+      val<-data$val
+      maxTime<-max(train$time)
+      sds<-sapply(train, function(x) (sd(x)))
+      means<-sapply(train, function(x) (mean(x)))
+      val<-normalizer(val,means,sds,maxTime)
+      train<-normalizer(train,means,sds,maxTime)
+      times<-seq(from=min(val$time),
+             to=max(val$time),
+             length.out = 22
+  )
+
+  times<-head(times, -1)
+  times<-tail(times, -1)
+
+      lr=paramTrials[i,]$lr
+      drpt=paramTrials[i,]$drpt
+      layer1=paramTrials[i,]$layer1
+      layer2=paramTrials[i,]$layer2
+      bsize=ceiling(nrow(train)/paramTrials[i,]$nBatch)
+      actv=paramTrials[i,]$actv
+      alp=paramTrials[i,]$alp
+      #=paramTrials[i,]$IPA
+      #=paramTrials[i,]$valLoss
+  source_python('src/deephitter.py')
+  hitnnSurv=fitDeephit(train,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
+  hitnnSurv<-t(hitnnSurv)
+  deephitCleaned<-pyProcess(hitnnSurv,times=times)
+  colnames(deephitCleaned)<-times
+  py_run_string("del fitDeephit")
+  rm(fitDeephit)
+
+    miniee <- Score(list('mod'=deephitCleaned),
+                             data =val, 
+                             formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
+                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
+
+      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
+      runTime<-append(runTime, Sys.time()-currentTime)
+      currentTime<-Sys.time() 
+    }
+
+    #paramTrials$IPA[i]<-mean(tempRes)
+    #paramTrials$IPAse[i]<-sd(tempRes)
+    paramTrials$valLoss[i]<-mean(tempVal)
+    paramTrials$runtime[i]<-mean(runTime)
+    #saveRDS(paramTrials,"complexds.rds")
+    print(paste("trial",i))
+    print(sum(runTime))
+    #print(as.character(paramTrials[i,]))
+    #rm(cbnn,fit,annPreds,covars_input,covars_output)
+    gc()
+  }
+  bestPerformance<-paramTrials[which.min(paramTrials$valLoss),]
+  return(bestPerformance)
+}
+
+klist<-readRDS("supportCV.rds")
+test<-readRDS("supportTest.rds")
+times<-seq(from=min(test$time)+0.000001,
+           to=max(test$time)-0.000001,
+           length.out = 25
+)
+
+bestDH<-optimDeepHit(klist,test,times,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,nBatch,actv,alp)
+
+```
+
+```{r hyperopcbnn,eval=FALSE,hide=FALSE}
+
+optimCBNN<-function(klist,test,times,lr,drpt,layer1,layer2,nBatch,actv){
+
+  # 
+  # for (i in 1:length(klist)){
+  #   #train<-sampleCaseBase(data = as.data.frame(klist[[i]]$train),time="time",event="status",ratio=100)
+  #   valcb<-sampleCaseBase(data = as.data.frame(klist[[i]]$val),time="time",event="status",ratio=100)
+  #   valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
+  #                          as.matrix(valcb[,ncol(valcb),drop=F])),
+  #                     as.matrix(valcb[,ncol(valcb)-1,drop=F]))
+  #   #valcbtensor[[1]][[2]]<-rep(train$offset[1],nrow(valcbtensor[[1]][[1]]))      
+  #   #klist[[i]]$train<-train
+  #   klist[[i]]$val<-valcbtensor
+  # }
+
+  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
+                                             drpt=drpt,
+                                             layer1=layer1,
+                                             layer2=layer2,
+                                             nBatch=nBatch,
+                                             actv=actv,
                                              #IPA=NA,
@@
       val<-data$val
-
+      maxTime<-max(train$time)
       sds<-sapply(train, function(x) (sd(x)))
       means<-sapply(train, function(x) (mean(x)))
-      val<-normalizer(val,means,sds)
-      train<-normalizer(train,means,sds)
+      #val[[1]][[1]]<-as.matrix(normalizer(as.data.frame(val[[1]][[1]]),means,sds))
+      train<-normalizer(train,means,sds,maxTime)
+
+      #train[,ncol(train)]<-data$train[,ncol(train)]
+      val<-normalizer(val,means,sds,maxTime)
+      valTest<-val
+            times<-seq(from=min(val$time),
+             to=max(val$time),
+             length.out = 22
+  )
+
+  times<-head(times, -1)
+  times<-tail(times, -1)
+
+      train<-sampleCaseBase(data = as.data.frame(train),time="time",event="status",ratio=100)
+      valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)
+      val<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
+                     as.matrix(valcb[,ncol(valcb),drop=F])
+      ),
+      as.matrix(valcb[,ncol(valcb)-1,drop=F])
+      )
+      val[[1]][[2]]<-rep(train$offset[1],nrow(val[[1]][[1]]))
+      #val[[1]][[2]]<-rep(0,nrow(val[[1]][[1]]))
 
@@
       layer2=paramTrials[i,]$layer2
-      bsize=paramTrials[i,]$bsize
-      actv=paramTrials[i,]$actv
-      alp=paramTrials[i,]$alp
-      #=paramTrials[i,]$IPA
-      #=paramTrials[i,]$valLoss
-      source_python('src/optimDH.py')
-
-      currentTime<-Sys.time() 
-      hitnnSurv=fitDeephit(train,val,floor(nrow(train)/100),
-                           epochs=epo,valida=val,patience=patience,
-                           min_delta=min_delta,drpt=drpt,
-                           lay1=layer1,lay2=layer2,
-                           lr=lr,actv=actv,alp=alp
-                           )
-
-      tempVal<-append(tempVal,min(hitnnSurv$val_metrics$scores$loss$score))
-      runTime<-append(runTime, Sys.time()-currentTime)
-      currentTime<-Sys.time() 
-    }
-    rm(fitDeephit)
-
-    #paramTrials$IPA[i]<-mean(tempRes)
-    #paramTrials$IPAse[i]<-sd(tempRes)
-    paramTrials$valLoss[i]<-mean(tempVal)
-    paramTrials$runtime[i]<-mean(runTime)
-    #saveRDS(paramTrials,"complexds.rds")
-    print(paste("trial",i))
-    print(mean(runTime))
-    #print(as.character(paramTrials[i,]))
-    #rm(cbnn,fit,annPreds,covars_input,covars_output)
-    gc()
-  }
-  bestPerformance<-paramTrials[which.min(paramTrials$valLoss),]
-  return(bestPerformance)
-}
-
-klist<-readRDS("supportCV.rds")
-test<-readRDS("supportTest.rds")
-  times<-seq(from=min(test$time)+0.000001,
-             to=max(test$time)-0.000001,
-             length.out = 25
-  )
-
-bestDH<-optimDeepHit(klist,test,times,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,bsize,actv,alp)
-
-```
-
-```{r hyperopcbnn,eval=reoptimize}
-
-optimCBNN<-function(klist,test,times,lr,drpt,layer1,layer2,bsize,actv){
-
-    # for (i in 1:length(klist)){
-  #   train<-sampleCaseBase(data = as.data.frame(klist[[i]]$train),time="time",event="status",ratio=100)
-  #   valcb<-sampleCaseBase(data = as.data.frame(klist[[i]]$val),time="time",event="status",ratio=100)
-  #   valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
-  #                          as.matrix(valcb[,ncol(valcb),drop=F])),
-  #                     as.matrix(valcb[,ncol(valcb)-1,drop=F]))
-  #   valcbtensor[[1]][[2]]<-rep(train$offset[1],nrow(valcbtensor[[1]][[1]]))      
-  #   klist[[i]]$train<-train
-  #   klist[[i]]$val<-valcbtensor
-  # }
-
-  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
-                                             drpt=drpt,
-                                             layer1=layer1,
-                                             layer2=layer2,
-                                             actv=actv,
-                                             #IPA=NA,
-                                             #IPAse=NA,
-                                             valLoss=NA,
-                                             runtime=NA))
-
-  # paramTrials<-unique.data.frame(paramTrials)
-  #paramTrials<-readRDS("complexTrials.rds")
-  for(i in 1:nrow(paramTrials)){
-    #paramTrials<-readRDS("complexTrials.rds")  
-    tempRes<-c()
-    tempVal<-c()
-    runTime<-c()
-
-    for(data in klist){
-
-      train<-data$train
-      val<-data$val
-
-      sds<-sapply(train, function(x) (sd(x)))
-      means<-sapply(train, function(x) (mean(x)))
-      #val[[1]][[1]]<-as.matrix(normalizer(as.data.frame(val[[1]][[1]]),means,sds))
-      train<-normalizer(train,means,sds)
-     # train[,ncol(train)]<-data$train[,ncol(train)]
-      val<-normalizer(val,means,sds)
-      train<-sampleCaseBase(data = as.data.frame(train),time="time",event="status",ratio=100)
-      valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)
-      val<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
-                     as.matrix(valcb[,ncol(valcb),drop=F])),
-                as.matrix(valcb[,ncol(valcb)-1,drop=F]))
-
-      val[[1]][[2]]<-rep(train$offset[1],nrow(val[[1]][[1]]))
-
-      lr=paramTrials[i,]$lr
-      drpt=paramTrials[i,]$drpt
-      layer1=paramTrials[i,]$layer1
-      layer2=paramTrials[i,]$layer2
-      bsize=c(floor(nrow(train)/100))
+      bsize=c(floor(nrow(train)/paramTrials[i,]$nBatch))
       actv=paramTrials[i,]$actv
@@
                         patience=patience,val=val)
-      tempVal<-append(tempVal,min(fit$resultOfFit$metrics$val_loss))
+      #tempVal<-append(tempVal,min(fit$resultOfFit$metrics$val_loss))
+        annPreds<-aar(fit,
+                times=times,
+                x_test=valTest[,-c(ncol(valTest))]
+  )
+
+  rownames(annPreds)<-annPreds[,1]
+  annProperpoly<- t(annPreds[,-1])
+  class(annProperpoly)<-c("tunnel",class(annProperpoly))
+
+    miniee <- Score(list('mod'=annProperpoly),
+                             data =valTest, 
+                             formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
+                             se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
+
+      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
+
       runTime<-append(runTime, Sys.time()-currentTime)
@@
 
-cbnnPerformance<-optimCBNN(klist,test,times,lr,drpt,layer1,layer2,bsize,actv)
+cbnnPerformance<-optimCBNN(klist,test,times,lr,drpt,layer1,layer2,nBatch,actv)
 #cbnnPerformance<-cbnnPerformance[which.min(cbnnPerformance$valLoss),]
+```
+
+```{r winners,eval=FALSE,hide=FALSE}
 winners<-list(cbnn=cbnnPerformance,
@@
   test<-fullTest[,-c(ncol(data))]
-  test<-normalizer(fullTest,means,sds)[,-c(ncol(data))]
-  fullTest<-normalizer(fullTest,means,sds)
+  test<-normalizer(fullTest,means,sds,maxTime)[,-c(ncol(data))]
+  fullTest<-normalizer(fullTest,means,sds,maxTime)
 
-  val<-normalizer(res$validate,means,sds)
+  val<-normalizer(res$validate,means,sds,maxTime)
 
-  train<-normalizer(train,means,sds)
+  train<-normalizer(train,means,sds,maxTime)
 
-  times<-seq(from=min(test$time)+0.000001,
-             to=max(test$time)-0.000001,
-             length.out = 27
+  linModelTrain<-as.data.frame(rbind(train,val))
+  #valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)#rbind(train,val)
+  #valcb$offset<-0
+  #valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
+  times<-seq(from=min(test$time),
+             to=max(test$time),
+             length.out = 20
   )
@@
 
-  lr=0.001#winners$ds$lr
+  lr=winners$ds$lr
   drpt=winners$ds$drpt
@@
   layer2=winners$ds$layer2
-  bsize=winners$ds$bsize
+  bsize=ceiling(nrow(train)/winners$ds$nBatch)
   actv=winners$ds$actv
+
   source_python('src/dsurv.py')
@@
   layer2=winners$cbnn$layer2
-  #bsize=winners$cbnn$bsize
+  bsize=c(floor(nrow(train)/winners$cbnn$nBatch))
   actv=winners$cbnn$actv
-
+   # 
+   # 
    #  lr=0.001
-  # drpt=0.1
-  # layer1=50
-  # layer2=20
-  # #bsize=winners$cbnn$bsize
-  # actv='relu'
+   #  drpt=0.5
+   #  layer1=10
+   #  layer2=10
+   #  patience=10
+    actv='linear'
 
   currentTime<-Sys.time() 
-  #valcbtensor[[1]][[2]]<-rep(mod_cb_glm$offset[1],nrow(valcbtensor[[1]][[1]]))
-
   covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
@@
                     epochs=epo,
-                        batch_size=c(floor(nrow(train)/100)),
+                    batch_size=bsize,
                     verbose=0,
@@
     geom_line()#+coord_cartesian(ylim=c(-0.1,0.2))
+  rm(fit,cbnn,covars_input,covars_output)
   #######################
@@
 ```
+
+```{r}
+
+cidxList<-readRDS('results/100supportcidx.rds')
+
+temp<-melt(setDT(as.data.frame(cidxList[[1]])), id.vars = c("times"), variable.name = "model")
+cidxSims=data.frame(model=temp$model,times=temp$times)
+
+count=1
+for(sim in cidxList){
+  cidxSims<-cbind(cidxSims, melt(setDT(as.data.frame(sim)), id.vars = c("times"), variable.name = "model")$value)
+
+  count<-count+1
+}
+
+cidxSummary<-statCalcSim(cidxSims)
+cidxC<-plotPerformance(cidxSummary,method="C-Index",iterations=iteration  )
+cidxC
+
+```

### 2023-04-05 analyses/plotterUpdated.Rmd
File Changes

---
title: "plottingScript"
output: html_document
---

```{r setup, include=T}
knitr::opts_chunk$set(echo = TRUE)
library(casebase) #core
library(keras) #core
source("src/functions.R") #core
library(data.table)#corecr
library(mltools)#corecr

library(survival)#compare
library(survminer)#compare
library(ggplot2)#compare
library(colorspace)#compare
library(DT)
library(splines)
library(reticulate)

library(simsurv)
library(flexsurv)
library(survivalmodels)
library(visreg)

#library(tidyverse)
library(riskRegression)
library(pseudo)

library(tensorflow)
library(ggrepel)

library(dplyr)
library(flextable)

library(cowplot)
library(matrixStats)
library(tableone)
library(gridExtra)
library(grid)
library(gtable)
#tf$config$threading$set_intra_op_parallelism_threads(1)
#tf$config$threading$set_inter_op_parallelism_threads(1)

#set.seed(2)
set.seed(1)
epo=2000
bsize=512
min_delta = 10^-7
patience = 10
iteration=100
e_sup=F
eval_firstSim<-F
numSamp<-2000

pal <- function(col, border = "light gray", ...)  {
  n <- length(col)
  plot(0, 0, type="n", xlim = c(0, 1), xlim=c(0.05,0.95),ylim = c(0, 1),
       axes = FALSE, xlab = "", ylab = "", ...)
  rect(0:(n-1)/n, 0, 1:n/n, 1, col = col, border = border)
}
statCalcSim<-function(metricSims){

  toRemove <- which(apply(metricSims, 1, function(x){any(is.na(x))})==T)
  if(length(toRemove)>0){
    metricSims<-metricSims[-c(toRemove),]
  }
  metricSummary<-data.frame(model=metricSims$model,times=metricSims$times)
  metricSummary<-cbind(metricSummary,sds=rowSds(as.matrix(metricSims[,-c(1,2)])),means=rowMeans(as.matrix(metricSims[,-c(1,2)])))

  ci<-apply(as.matrix(metricSims[,-c(1,2)]), 1, function(x){mean(x)+c(-1.96,1.96)*sd(x)/sqrt(length(x))})
  rownames(ci)<-c('cilower95','ciupper95')
  pi<-apply(as.matrix(metricSims[,-c(1,2)]), 1, function(x){quantile(x,c(.025, .975))})
  rownames(pi)<-c('pilower95','piupper95')

  metricSummary<-cbind(metricSummary,t(ci),t(pi))

  return(metricSummary)
}

plotPerformance<-function(metricSummary,xlab="Follow-up time",method='Metric',iterations='NA',ticks=5){

  allModels<-data.frame(model=c("Null model","Cox_Lin",'CB_Logi','DeepSurv',
                                'DeepHit','cbnn','CBNN_Poly','optimal',
                                "Optimal"),
                        color=qualitative_hcl(9,palette='Dark 3'),
                        trueName=c("KM","Cox",'CBLR','DeepSurv',
                                   'DeepHit','CBNN','CBNN','Optimal',
                                   "Optimal"))
  allModels$color[1]="#000000"
  allModels$color[2]="#3cff4b"#"#7C4C74" #deephit
  allModels$color[3]="#1e90ff"#"#4363ff" #"#F9A602"#CBLR
  allModels$color[4]="#C0C000"#"#f032e6"#"#ff194B"
  allModels$color[5]="#ff8231" #deepsurv
  allModels$color[6]="#911eff"#"#911eff"
  allModels$color[7]="#911eff"#"#42d4f4"
  allModels$color[8]="#f00000"#"#ff32e6"
  allModels$color[9]="#f00000"
  allModels<- allModels[order(allModels$trueName),]
  i <- sapply(metricSummary, is.factor)
  metricSummary[i] <- lapply(metricSummary[i], as.character)
  metricSummary<-metricSummary[complete.cases(metricSummary), ]
  mainTitle<-""
  metricSummary$color<-NA
  for(i in 1:nrow(metricSummary)){
    modelPos<-which(metricSummary$model[i] ==allModels$model)
    metricSummary$model[i]<-allModels$trueName[modelPos]
    metricSummary$color[i]<-allModels$color[modelPos]
  }

  metricSummary<-metricSummary[order(metricSummary$model),]

  #colorsToUse<- which(allModels$trueName %in% metricSummary$model)

  if(method=="C-Index"){
    method=expression("C"["IPCW"])
  }   
  metricSummary#<- metricSummary[-which(metricSummary$model=="CBNN_Spline"),]
  #mainTitle<-paste(method,' : 95% P.I. over ', iterations, ' iterations',sep='')
  colnames(metricSummary)[1]<-"Models"
  a<-ggplot(data=metricSummary, aes(x=times,y=means,col=Models))+
    geom_line() +
    geom_ribbon(aes(ymin=cilower95,ymax=ciupper95,fill=Models),linetype=0,alpha=0.3) +
    ggtitle(mainTitle) + 
    xlab(xlab) +
    ylab(method)+
    theme_minimal()+
    scale_color_manual(values=unique(metricSummary$color))+
    scale_fill_manual(values=unique(metricSummary$color))+
    scale_x_continuous(n.breaks=6)
  # scale_color_manual(labels = metricSummary$model, values = metricSummary$color)

  return(a)
}

```

```{r getAll,eval=T}

#age, sex, race, number of comorbidities, presence of diabetes, presence of dementia, presence of cancer, mean arterial blood pressure, heart rate, respiration rate, temperature, white blood cell count, serum’s sodium, and serum’s creatinine

getPlots<-function(briers,cidxs,auc,dMax){

  BrierIPAList<-readRDS(briers)

  cidxList<-readRDS(cidxs)
  AUClist<-readRDS(auc)

  ipaSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times*dMax)
  brierSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times*dMax)
  ibsSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times*dMax)
  ibsTable<-as.data.frame(matrix(data=NA,nrow=100,ncol=length(unique(BrierIPAList[[1]]$model))))
  colnames(ibsTable)<-unique(BrierIPAList[[1]]$model)
  count=1
  for(sim in BrierIPAList){
    ipaSims<-cbind(ipaSims, sim$IPA)
    brierSims<-cbind(brierSims,sim$Brier)
    ibsSims<-cbind(ibsSims,sim$IBS)
    for(mod in colnames(ibsTable)){

      ibsTable[count,which(mod ==colnames(ibsTable))]<-sim$IBS[sim$model==mod & sim$time==max(sim$time)]

    }
    count<-count+1
  }
  ibsTable<-na.omit(ibsTable)

  aucSims=data.frame(model=AUClist[[1]]$model,times=AUClist[[1]]$times*dMax)
  for(sim in AUClist){
    aucSims<-cbind(aucSims,sim$AUC)

  }

  ipaSummary<-statCalcSim(ipaSims)
  brierSummary<-statCalcSim(brierSims)
  ibsSummary<-statCalcSim(ibsSims)
  aucSummary<-statCalcSim(aucSims)
  ipa<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
  brier<-plotPerformance(brierSummary,method="BS",iterations=iteration )
  ibs<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )
  auc<-plotPerformance(aucSummary,method="AUC",iterations=iteration )

  temp<-melt(setDT(as.data.frame(cidxList[[1]])), id.vars = c("times"), variable.name = "model")
  cidxSims=data.frame(model=temp$model,times=temp$times*dMax)

  count=1
  for(sim in cidxList){
    cidxSims<-cbind(cidxSims, melt(setDT(as.data.frame(sim)), id.vars = c("times"), variable.name = "model")$value)

    count<-count+1
  }

  cidxSummary<-statCalcSim(cidxSims)
  cidxC<-plotPerformance(cidxSummary,method="C-Index",iterations=iteration  )
  return(list(ipa=ipa,
              brier=brier,
              ibs=ibs,
              cidx=cidxC,
              auc=auc))
}

cMax<-max(readRDS("ComplexSimTest.rds")$time)/365
mMax<-max(readRDS("myelomaTest.rds")$time)/365
fMax<-max(readRDS("flchainTest.rds")$time)/365
pMax<-max(readRDS("prostateTest.rds")$time)/12

complex<-getPlots('results/100complexBrier.rds','results/100complexcidx.rds','results/100complexAuc.rds',cMax)

myeloma<-getPlots('results/100myelomaBrier.rds','results/100myelomacidx.rds','results/100myelomaAuc.rds',mMax)

flchain<-getPlots('results/100flchainBrier.rds','results/100flchaincidx.rds','results/100flchainAuc.rds',fMax)

prostate<-getPlots('results/100prostateBrier.rds','results/100prostatecidx.rds','results/100prostateAuc.rds',pMax)
#sup<-getPlots('results/100supportBrier.rds','results/100supportcidx.rds')

```

```{r megaplot,fig.height=12,fig.width=12}
#complex[[4]]$layers[[2]]<-NULL
pt <- plot_grid(
  complex[[1]]+ theme(legend.position="none")+ ggtitle("Complex simulation")+ labs(x = "")+coord_cartesian(ylim=c(-0.3,0.7)),
  myeloma[[1]]+ theme(legend.position="none")+ ggtitle("MM")+ labs(x = "",y="")+coord_cartesian(ylim=c(-1,.1)),
  flchain[[1]]+ theme(legend.position="none")+ ggtitle("FLC")+ labs(x = "",y="")+coord_cartesian(ylim=c(-0.025,.45)),
  prostate[[1]]+ theme(legend.position="none")+ ggtitle("Prostate")+ labs(x = "",y="")+coord_cartesian(ylim=c(-0.02,0.205)),
  complex[[5]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)")+coord_cartesian(ylim=c(0.68,1)),
  myeloma[[5]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)",y="")+coord_cartesian(ylim=c(0.00,0.75)),
  flchain[[5]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)",y="")+coord_cartesian(ylim=c(0.8,0.875)),
  prostate[[5]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)",y="")+coord_cartesian(ylim=c(0.65,0.82)),
  # complex[[4]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)")+coord_cartesian(ylim=c(0.5,0.8)),
  # myeloma[[4]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)",y="")+coord_cartesian(ylim=c(0.4,0.65)),
  # flchain[[4]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)",y="")+coord_cartesian(ylim=c(0.65,0.9)),
  # prostate[[4]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)",y="")+coord_cartesian(ylim=c(0.5,0.73)),
  labels = c('A',"B","C","D",'E','F','G',"H"),#,"I","J","K","L"),
  ncol=4)
legend <- get_legend(
  # create some space to the left of the legend
  complex$ipa  +
    guides(color = guide_legend(nrow = 2)) +
    theme(legend.position = "bottom")+
    theme(legend.text=element_text(size=8))
)

g<-plot_grid(pt, legend,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))
pdf(file="figures/megaPlot.pdf",width=16,height=9)
g
dev.off()
```

```{r flextableTablePrep,fig.height=46,fig.width=26}

edit_cell <- function(table, row, col, name="core-fg", ...){
  l <- table$layout
  ids <- which(l$t %in% row & l$l %in% col & l$name==name)
  for (id in ids){
    newgrob <- editGrob(table$grobs[id][[1]], ...)
    table$grobs[id][[1]] <- newgrob
  }
  table
}

#SUMMARY DATA ORDER, IPA BS IBS CIDX
makeTable<-function(summaryData,study="study"){
  allModels<-data.frame(model=c("Null model","Cox_Lin",'CB_Logi','DeepSurv',
                                'DeepHit','CBNN_Poly','Optimal'),
                        color=rainbow_hcl(7, start = 30, end = 300),
                        trueName=c("KM","Cox",'CBLR','DeepSurv',
                                   'DeepHit','CBNN','Optimal'))

  wantedScores<-c(0.25,0.5,0.75,1)

  tableResults<-as.data.frame(matrix(NA,nrow=nrow(allModels)+2,ncol=20))
  tableResults[1,]<-c("IPA","IPA","IPA","IPA",
                      "BS","BS","BS","BS",
                      "IBS","IBS","IBS","IBS",
                      "C-Index","C-Index","C-Index","C-Index",
                      "AUC","AUC","AUC","AUC"
  )
  tableResults[2,]<-c(0.25,0.5,0.75,1,0.25,0.5,0.75,1,0.25,0.5,0.75,1,0.25,0.5,0.75,1,0.25,0.5,0.75,1)*100
  indexer<-1
  rownames(tableResults)<-c(study,"% of time",allModels$trueName)
  meanOnly<-tableResults
  for (i in 1:length(summaryData)){ #IPA BRIER IBS CIDX is indexed by i

    tempPos<-which(rownames(tableResults) %in% unique(summaryData[[i]]$Models)  )
    tempModels<-rownames(tableResults)[ tempPos]
    for(j in 1:length(tempModels)){ #indexes model
      offset<-(i-1)*4
      indexer=1+offset
      for(k in 1:length(wantedScores)){ #indexes .25 .5 .75 1.0
        tempTime<-unique(summaryData[[i]]$times)[(floor(length(unique(summaryData[[i]]$times))*wantedScores[k]))]
        tempResRow<-which(summaryData[[i]]$Models==rownames(tableResults)[ tempPos[j]] & summaryData[[i]]$times==tempTime)
        if(length(tempResRow)==0){
          tempTime<-max(summaryData[[i]]$times[which(summaryData[[i]]$Models==rownames(tableResults)[ tempPos[j]])])
          tempResRow<-which(summaryData[[i]]$Models==rownames(tableResults)[ tempPos[j]] & summaryData[[i]]$times==tempTime)
        }
        # if(length(round(summaryData[[i]]$means[tempResRow],2))<1){
        #   tableResults[tempPos[j],indexer]<-"-0.23 (-0.25,-0.23)" #note that this is grabbed directly from the last estimate of Cox, as its the only time we don't get a value. So we extrapolate from the last available survival time.
        # }else{
        tableResults[tempPos[j],indexer]<-paste(round(summaryData[[i]]$means[tempResRow],2),
                                                " (",round(summaryData[[i]]$cilower95[tempResRow],2),
                                                ",",round(summaryData[[i]]$ciupper95[tempResRow],2),")",sep="")
        meanOnly[tempPos[j],indexer]<-round(summaryData[[i]]$means[tempResRow],2)
        #}
        indexer<-indexer+1

      }
    }

  }

  return(list(tab=tableResults[-3,-c(5:16)],means=meanOnly[-3,-c(5:16)]))#
}

cleantables<-function(tab){
  toRemove<-which(rowSums(is.na(tab))==ncol(tab))
  tab[is.na(tab)]<-"-"
  if(length(toRemove)>0){

    return(tab[-toRemove,])
  }
  return(tab)
}

quickEdit<-function(grob,rows,columns){
  return(edit_cell(grob, rows, columns, "core-fg", 
                   gp=gpar(fontsize=12, fontface="bold")))
}

addBorders<-function(grob){
  grob$layout[1:2 , c("l", "r")] <- list(c(5, 9),c(2,6))
  grob<- gtable_add_grob(grob,
                         grobs = rectGrob(gp = gpar(fill = NA, lwd = 2)),
                         t = 2, b = nrow(grob), l = 1, r = ncol(grob))
  grob <- gtable_add_grob(grob,
                          grobs = rectGrob(gp = gpar(fill = NA, lwd = 2)),
                          t = 1, l = 1, r = ncol(grob))
  grob <- gtable_add_grob(grob,
                          grobs = rectGrob(gp = gpar(fill = NA, lwd = 2)),
                          t = 1, l = 2,r=5,b= nrow(grob))
  return(grob)
}

getMaxes<-function(Tab){

  tab<-Tab$tab

  sumer<-Tab$means[-c(1,2),]
  sumer<-as.data.frame(apply(na.omit(sumer),2,as.numeric))
  rowPick<-c()
  colPick<-c()

  rows<-apply(sumer,2,function(x){which(x==max(x))})
  cols<-c()
  for(i in 1:length(rows)){
    cols<-append(cols,rep(i,length(rows[[i]])))
  }
  rows<-c(5,as.vector(unlist(rows)))
  picks<-data.frame(row=rows,
                    col=c(0,cols))
  return(picks)
}

convert0<-function(tab){
  tempNames<-rownames(tab)
  tab<-as.data.frame(lapply(tab, function(x){gsub("0 ","0.00 ", x,fixed=T)}))
  tab<-as.data.frame(lapply(tab, function(x){gsub('(0,',"(0.00,", x,fixed=T)}))
  tab<-as.data.frame(lapply(tab, function(x) gsub(",0)",",0.00)", x,fixed=T)))
  rownames(tab)<-tempNames
  return(tab)
}

flexMaker<-function(tab,boldPicks=NA,title){
  boldPicks$col=boldPicks$col+1 #makes finding the correct ones easier.
  bord<-fp_border_default(color="black",style="solid")
  bord2<-fp_border_default(color="black",style="solid",width = 2)
  tab<-cbind(rownames(tab),tab)
  ft<-delete_part(flextable(tab[-c(1,2),],cwidth=5,cheight=1), part = "header")

  ft <- add_header_row(x = ft, values = c(title,"IPA","AUC"),
                       colwidths = c(1,4,4))
  # ft <- compose(ft, part = "header", i = 1, j = 6,
  #               value = as_paragraph("C",
  #                                    as_sub("IPCW")))
  #ft<-bold(ft, bold = TRUE, part = "header")

  #ft$header$styles$text$bold$data
  ft <- add_header_row(x = ft,top=F, values = c("Method", 25,50,75,100,25,50,75,100),
                       colwidths = c(1,1,1,1,1,1,1,1,1))
  ft$header$styles$text$bold$data[1,1]<-T
  #ft <- add_header_row(
  #  x = ft, values = c(title),
  #  colwidths = c(9))
  ft <- align(ft, i = c(1,2), part = "header", align = "center")
  ft<-vline(ft, j = c(1,5), border = bord2, part = "body")
  ft<-hline(ft,i=c(1,2),border=bord2,part="header")
  #ft<-hline(ft,i=c(1),border=bord,part="body")
  ft<-fontsize(ft, i = NULL, j = NULL, size =7, part = "all")
  ft<-fontsize(ft, i = NULL, j = NULL, size =10, part = "header")
  ft <- align(ft, i = NULL, part = "body", align = "center")
  ft$body$styles$text$font.size$data[,1]<-10

  if(is.data.frame(boldPicks)){
    for(i in 1:nrow(boldPicks)){
      ft$body$styles$text$bold$data[boldPicks$row[i],boldPicks$col[i]]<-T
    }
  }
  return(rasterGrob(as_raster(autofit(ft))))
}

```

```{r flexTableMain,fig.height=46,fig.width=26}

getFlex<-function(result,titler="X: study"){
  Tab<-makeTable(list(result$ipa$data,result$brier$data,result$ibs$data,result$cidx$data,result$auc$data),study="Complex")
  Picks<-getMaxes(Tab)
  Tab<-cleantables(Tab$tab)
  Tab<-convert0(Tab)
  Flex<-flexMaker(tab=Tab,boldPicks = Picks,title=titler)

}

cFlex<-getFlex(complex,"A: Complex")
mFlex<-getFlex(myeloma,"B: MM")
fFlex<-getFlex(flchain,"C: FLC")
pFlex<-getFlex(prostate,"D: Prostate")

pdf(file="figures/megaTable.pdf",width=14,height=13)
cowplot::plot_grid(cFlex,mFlex,fFlex,pFlex,ncol=1)
dev.off()

```

```{r hyperparams}

getWinsTable<-function(win,main){
  winners<-readRDS(win)

  winners$cbnn$alp="-"
  winners$ds$alp="-"
  winners$ds<-winners$ds[,c(1,2,3,4,5,6,9,7,8)]

  winners$cbnn<-winners$cbnn[,c(1,2,3,4,5,6,9,7,8)]
  tabWin<-data.frame(CBNN=t(winners$cbnn),
                     DeepSurv=t(winners$ds),
                     DeepHit=t(winners$dh))

  colnames(tabWin)<-c("CBNN","DeepSurv","DeepHit")
  rownames(tabWin)<-c("Learning rate","Dropout","First layer nodes","Second layer nodes","Number of batches","Activation function","\U03B1","IBS","runtime")
  tabWin$`Hyper parameter`<-rownames(tabWin)
  tabWin<-tabWin[,c(4,1,2,3)]
  tabWin<-tabWin[-nrow(tabWin),]

  ft<-flextable(tabWin,cwidth=1.35,cheight=0.1)%>%flextable::hline(i = 7, part = 'body')
  ft <- add_header_row(x = ft, values = c(main),
                       colwidths = c(4))
  ft$header$styles$text$bold$data[1,1]<-T
  ft<-ft%>%hline_top(border = fp_border_default(width = 0), 
                     part = "header")

  return(rasterGrob(as_raster(autofit(ft))))
}

cwtab<-getWinsTable("complexOptimWins.rds","A: Complex")
mwtab<-getWinsTable("myelomaWins.rds","B: MM")
fwtab<-getWinsTable("flchainWins.rds","C: FLC")
pwtab<-getWinsTable("prostateWins.rds","D: Prostate")

pdf(file="figures/winTable.pdf",width=12,height=6)
cowplot::plot_grid(cwtab,mwtab,fwtab,pwtab,ncol=2)
dev.off()

```

```{r complexHazard}

library(casebase)

klist<-readRDS("ComplexSimCV.rds")
data<-rbind(klist[[1]]$train,klist[[1]]$val)

start_time <- Sys.time()
BrierIPAList<-list()
cScore<-list()
pos<-1

spec = c(train = .85, validate = .15)

g = sample(cut(
  seq(nrow(data)), 
  nrow(data)*cumsum(c(0,spec)),
  labels = names(spec)
))

res = split(data, g)

trainOG<-res$train

numSamp<-nrow(trainOG)
samp<-sample(seq(1,nrow(trainOG),by=1) , nrow(trainOG),replace = T) 
train<-trainOG[samp,]

#train<-res$train
val<-res$validate

sds<-sapply(train, function(x) (sd(x)))
means<-sapply(train, function(x) (mean(x)))
maxTime<-max(train$time)

fullTest<-readRDS("ComplexSimTest.rds")
test<-fullTest[,-c(ncol(data))]
test<-normalizer(fullTest,means,sds,maxTime)[,-c(ncol(data))]
fullTest<-normalizer(fullTest,means,sds,maxTime)

val<-normalizer(res$validate,means,sds,maxTime)

train<-normalizer(train,means,sds,maxTime)
linModelTrain<-as.data.frame(rbind(train,val))

fakeTrain<-train
fakeTrain$status[which((fakeTrain$status==1) & (fakeTrain$z1==max(fakeTrain$z1)))]<-2

dat<-popTime(fakeTrain,time="time",event="status")
a<-plot(dat, add.competing.event = T) +scale_color_manual(labels = c("Time-constant", "Time-varying"),values=c("red","blue"))+scale_fill_manual(labels = c("Time-constant", "Time-varying"),values=c("red","blue"))+
  guides(color = "none")+labs(fill="Case-series events")

pdf(file="figures/ComplexPopTime.pdf",width=8,height=6)
a
dev.off()

#https://onlinelibrary.wiley.com/doi/10.1111/biom.12232

```

```{r codeDump,eval=F,include=F}

times<-seq(from=min(test$time),
           to=max(test$time),
           length.out = 27
)

times<-head(times, -1)
times<-tail(times, -1)

haz_inter <- fitSmoothHazard(status~bs(time)+ z1+z2+z3+z1:time+z2:z3-time,
                             data = linModelTrain,
                             time = "time")

nonPH<-casebase::hazardPlot(haz_inter,newdata = data.frame(z1=max(train$z1),z2=1,z3=1))
nonPH$`Time Varying`<-TRUE
PH<- casebase::hazardPlot(haz_inter,newdata = data.frame(z1=min(train$z1),z2=0,z3=1))
PH$`Time Varying`<-FALSE

H<-as.data.frame(rbind(nonPH,PH))

ggplot(H,aes(x=time,y=predictedhazard,col=`Time Varying`))+geom_line()
fill_cols <- c("z1=0" = "Red", "z1=1" = "#009E73",
               "Base series" = "#0072B2")
g<-plot(popTime(train,time="time",event="status",),  add.competing.event=T,fill.params = list(
  name = element_blank(),
  breaks = c("Case series", "Competing event", "Base series")
))

g

a<-popTime(train,time="time",event="status",

)

plot(mod_cb_tvc,
     hazard.params = list(xvar = "time",
                          by = "z1",
                          alpha = 1,
                          ylab = "Hazard"))

mod_cb_tvc <- fitSmoothHazard(censrec ~ hormon * ns(log(rectime), df = 3),
                              data = brcancer,
                              time = "rectime")
plot(mod_cb_tvc,
     hazard.params = list(xvar = "rectime",
                          by = "hormon",
                          alpha = 0.05,
                          ylab = "Hazard"))

```

### 2023-04-05 analyses/src/deephitter.py

@@ -70,9 +70,15 @@ def fitDeephit(train,fullTest,bsize,epochs,valida,patience,min_delta,drpt,lay1,l
        torch.nn.Linear(in_features,int(lay1)),
        torch.nn.Tanh(),
        torch.nn.Dropout(float(dropout)),
-       
        torch.nn.Linear(int(lay1), int(lay2)),
        torch.nn.Tanh(),
+       torch.nn.Dropout(float(dropout)),
+       torch.nn.Linear(int(lay2), out_features))
+  if actv=='linear':
+    net = torch.nn.Sequential(
+       torch.nn.Linear(in_features,int(lay1)),
+       torch.nn.Dropout(float(dropout)),
+       torch.nn.Linear(int(lay1), int(lay2)),
        torch.nn.Dropout(float(dropout)),
        torch.nn.Linear(int(lay2), out_features))
   model = DeepHitSingle(net, tt.optim.Adam, alpha=alp, sigma=0.1, duration_index=labtrans.cuts)

### 2023-04-05 analyses/src/deepsurver.py

@@ -12,7 +12,7 @@ from pycox.evaluation import EvalSurv
 
 
 
-def fitDeepSurv(train,fullTest,bsize,epochs,valida,ti='time',ev='status',patience=10,min_delta=10**-7,drpt=0.5,lay1=200,lay2=20):#,epoch,bsize,tPos,ePos):
+def fitDeepSurv(train,fullTest,bsize,epochs,valida,ti='time',ev='status',patience=10,min_delta=10**-7,drpt,lay1,lay2):#,epoch,bsize,tPos,ePos):
   n= np.shape(train)[0]
   vl_size = int(n*0.20)
   df_val = valida#train.sample(frac=0.2)

### 2023-04-05 analyses/src/dsurv.py

@@ -68,7 +68,14 @@ def fitDeepSurv(train,fullTest,bsize,epochs,valida,patience,min_delta,drpt,lay1,
        torch.nn.Dropout(float(dropout)),
        torch.nn.Linear(int(lay2), out_features))
     
-  
+  if actv=='linear':
+    net = torch.nn.Sequential(
+       torch.nn.Linear(in_features,int(lay1)),
+       torch.nn.Dropout(float(dropout)),
+       torch.nn.Linear(int(lay1), int(lay2)),
+       torch.nn.Dropout(float(dropout)),
+       torch.nn.Linear(int(lay2), out_features))
+       
   model = CoxPH(net, tt.optim.Adam)
   
   model.optimizer.set_lr(lr)

### 2023-04-05 analyses/src/functions.R

@@ -14,11 +14,11 @@ statCalcSim<-function(metricSims){
 }
 
 plotPerformance<-function(metricSummary,xlab="Follow-up-time",method='Metric',iterations='NA'){
-  mainTitle<-paste(method,' : 95% P.I. over ', iterations, ' iterations',sep='')
+  mainTitle<-paste(method,' : 95% C.I. over ', iterations, ' iterations',sep='')
   
   a<-ggplot(data=metricSummary, aes(x=times,y=means,col=model))+
     geom_line() +
-    geom_ribbon(aes(ymin=pilower95,ymax=piupper95,fill=model),linetype=0,alpha=0.3) +
+    geom_ribbon(aes(ymin=cilower95,ymax=ciupper95,fill=model),linetype=0,alpha=0.3) +
     ggtitle(mainTitle) + 
     xlab(xlab) +
     ylab(method)
@@ -65,7 +65,7 @@ cIndexSummary<-function(et_train, et_test,riskList,cScore){
   for (i in 1:length(riskList)){
     #print(i)
     cScore[,i+1]<-calcCidx(et_train=et_train,et_test=et_test,risk=riskList[[i]],
-                           times = as.numeric(colnames(riskList[[i]])))[,2]
+                           times = as.numeric(colnames(riskList[[1]])))[,2]
   }
   cScore<-cScore
   return(na.omit(cScore))
@@ -253,9 +253,9 @@ fitSmoothHaz<-function(cbnn,epochs=2000,batch_size=500,verbose=0,monitor="val_lo
       min_delta = min_delta,
       patience = patience,
       verbose = verbose,
-      mode = c("auto"),
+      mode = c("min"),
       baseline =NULL,#lossCutOff,
-      restore_best_weights = T#,
+      restore_best_weights = F#,
       # validation_split = 0.2
     )))
   
@@ -302,7 +302,7 @@ fitSmoothHazSpline<-function(cbnn,epochs=2000,batch_size=500,verbose=0,monitor="
       verbose = verbose,
       mode = c("auto"),
       baseline =NULL,#lossCutOff,
-      restore_best_weights = T#,
+      restore_best_weights = F#,
       # validation_split = 0.2
     )))
   
@@ -469,14 +469,15 @@ aarOld<-function(cbnn, times=times,x_test=x_test){
 
 
 
-normalizer<-function(data,means,sds){
-  normalized<-data
+normalizer<-function(data,means,sds,maxTime){
[…]

### 2023-04-05 analyses/src/optimDH.py

@@ -84,7 +84,13 @@ def fitDeephit(train,fullTest,bsize,epochs,valida,patience,min_delta,drpt,lay1,l
        torch.nn.Tanh(),
        torch.nn.Dropout(float(dropout)),
        torch.nn.Linear(int(lay2), out_features))
-
+  if actv=='linear':
+    net = torch.nn.Sequential(
+       torch.nn.Linear(in_features,int(lay1)),
+       torch.nn.Dropout(float(dropout)),
+       torch.nn.Linear(int(lay1), int(lay2)),
+       torch.nn.Dropout(float(dropout)),
+       torch.nn.Linear(int(lay2), out_features))
   model = DeepHitSingle(net, tt.optim.Adam, alpha=alp, sigma=0.1, duration_index=labtrans.cuts)
   
  #lr_finder = model.lr_finder(x_train, y_train, batch_size=int(bsize), tolerance=2)

### 2023-04-05 analyses/src/optimDSurv.py

@@ -67,11 +67,17 @@ def fitDeepSurv(train,fullTest,bsize,epochs,valida,patience,min_delta,drpt,lay1,
        torch.nn.Tanh(),
        torch.nn.Dropout(float(dropout)),
        torch.nn.Linear(int(lay2), out_features))
-    
+  if actv=='linear':
+    net = torch.nn.Sequential(
+       torch.nn.Linear(in_features,int(lay1)),
+       torch.nn.Dropout(float(dropout)),
+       torch.nn.Linear(int(lay1), int(lay2)),
+       torch.nn.Dropout(float(dropout)),
+       torch.nn.Linear(int(lay2), out_features))    
   
   model = CoxPH(net, tt.optim.Adam)
   
-  model.optimizer.set_lr(lr)
+  model.optimizer.set_lr(float(lr))
   
   callbacks = [tt.callbacks.EarlyStopping(min_delta=min_delta, patience=patience)]
   log = model.fit(x_train, y_train, int(bsize),int(epochs), callbacks, val_data=val,verbose=False )

### 2023-04-05 analyses/optimFlchain.Rmd
cleanedUpCode

@@
 eval_flchain<-T
-#source_python('src/getSUP.py')
+
 source_python("src/cIndex.py")
-#source_python("src/dsmScript.py")
-#source_python('src/deephitter.py')
+
 ##########################
@@
 
-```{r flchainPrep,eval=FALSE,hide=FALSE}
+```{r flchainPrep,eval=TRUE,hide=FALSE}
 
@@
 
-##saveRDS(res$test,"flchainTest.rds")
+##saveRDS(res$test,"cvData/flchainTest.rds")
 
@@
 }
-##saveRDS(klist,"flchainCV.rds")
+##saveRDS(klist,"cvData/flchainCV.rds")
 
@@
 
-```{r setFixedHyperparameters,eval=FALSE,hide=FALSE}
+```{r setFixedHyperparameters,eval=TRUE,hide=FALSE}
 
-klist<-readRDS("flchainCV.rds")
-
-# lr=c(0.001)
-# drpt=c(0.25,0.5)
-# layer1=c(10,20,30)
-# layer2=c(10,20,30)
-# bsize=c(floor(nrow(klist[[1]]$train)/100))
-# actv=c("relu","linear")#,"tanh"
-# alp=c(0)
+klist<-readRDS("cvData/flchainCV.rds")
 
@@
 
-# 
-# lr=c(0.001)
-# drpt=c(0.25)
-# layer1=c(20)
-# layer2=c(10)
-# nBatch=c(1000,100)
-# actv=c("relu")#
-# alp=c(0)
+lr=c(0.01)
+drpt=c(0.01)
+layer1=c(50)
+layer2=c(10)
+nBatch=c(100)
+actv=c("relu","linear")#
+alp=c(0)
 
@@
 
-```{r hyperOpDsurv,eval=FALSE,hide=FALSE}
-
-optimDeepSurv<-function(klist,test,baselineTrials,lr,drpt,layer1,layer2,nBatch,actv){
-
-  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
-                                             drpt=drpt,
-                                             layer1=layer1,
-                                             layer2=layer2,
-                                             nBatch=nBatch,
-                                             actv=actv,
-                                             #IPA=NA,
-                                             #IPAse=NA,
-                                             valLoss=NA,
-                                             runtime=NA))
-
-  #paramTrials<-paramTrials[sample(nrow(paramTrials), baselineTrials), ]
-
-  # paramTrials<-unique.data.frame(paramTrials)
-  #paramTrials<-readRDS("complexTrials.rds")
-  for(i in 1:nrow(paramTrials)){
-    #paramTrials<-readRDS("complexTrials.rds")  
-    tempRes<-c()
-    tempVal<-c()
-    runTime<-c()
-
-    for(data in klist){
-      currentTime<-Sys.time() 
-      train<-data$train
-      val<-data$val
-      maxTime<-max(train$time)
-      sds<-sapply(train, function(x) (sd(x)))
-      means<-sapply(train, function(x) (mean(x)))
-      val<-normalizer(val,means,sds,maxTime)
-      train<-normalizer(train,means,sds,maxTime)
-      times<-seq(from=min(val$time),
-                 to=max(val$time),
-                 length.out = 22
-      )
-
-      times<-head(times, -1)
-      times<-tail(times, -1)
-
-      lr=paramTrials[i,]$lr
-      drpt=paramTrials[i,]$drpt
-      layer1=paramTrials[i,]$layer1
-      layer2=paramTrials[i,]$layer2
-      bsize=ceiling(nrow(train)/paramTrials[i,]$nBatch)
-      actv=paramTrials[i,]$actv
-      #=paramTrials[i,]$IPA
-      #=paramTrials[i,]$valLoss
+```{r hyperOpDsurv,eval=TRUE,hide=FALSE}
+klist<-readRDS("cvData/flchainCV.rds")
 source_python('src/dsurv.py')
-
-      ctrain<-train
-      #ctrain$time<- train$time + runif(nrow(train),-0.000001,0.000001)
-
-      coxnnsurv=fitDeepSurv(ctrain,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
-
-      coxnnsurv<-t(coxnnsurv)
-      cnnCleaned<-pyProcess(coxnnsurv,times=times)
-      colnames(cnnCleaned)<-times
-      py_run_string("del fitDeepSurv")
-      rm(fitDeepSurv)
-
-      miniee <- Score(list('mod'=cnnCleaned),
-                      data =val, 
-                      formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                      se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
-
-      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
-      runTime<-append(runTime, Sys.time()-currentTime)
-      currentTime<-Sys.time()
-
-      # coxnnsurv<-coxnnsurv[[1]]
-      # coxnnsurv<-t(coxnnsurv)
-      # #colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
-      # cnnCleaned<-pyProcess(coxnnsurv,times=times)
-      # colnames(cnnCleaned)<-times#seq(0,1,length.out = ncol(cnnCleaned))
-      # py_run_string("del fitDeepSurv")
-      # rm(fitDeepSurv)
-      # brierFinalResults <- Score(list('ds'=cnnCleaned),
-      #                            data =as.data.frame(val),#rbind(train,val) 
-      #                            formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-      #                            se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
-      # cutted<-brierFinalResults[which(brierFinalResults$model=="ds"),]
-      #tempRes<-append(tempRes,mean(cutted[-c(1,2,24,25)]$IPA))
-    }
-
-    #paramTrials$IPA[i]<-mean(tempRes)
-    #paramTrials$IPAse[i]<-sd(tempRes)
-    paramTrials$valLoss[i]<-mean(tempVal)
-    paramTrials$runtime[i]<-mean(runTime)
-    ##saveRDS(paramTrials,"complexds.rds")
-    print(paste("trial",i))
-    print(sum(runTime))
-    #print(as.character(paramTrials[i,]))
-    #rm(cbnn,fit,annPreds,covars_input,covars_output)
-    gc()
-  }
-  bestPerformance<-paramTrials[which.min(paramTrials$valLoss),]
-  return(bestPerformance)
-}
-
-klist<-readRDS("flchainCV.rds")
-test<-readRDS("flchainTest.rds")
-source_python('src/dsurv.py')
-bestDS<-optimDeepSurv(klist,test,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,nBatch,actv)
+bestDS<-optimDeepSurv(klist,lr,drpt,layer1,layer2,nBatch,actv)
 
@@
 
-```{r optimDeepHit,eval=FALSE,hide=FALSE}
-
-optimDeepHit<-function(klist,test,times,baselineTrials,lr,drpt,layer1,layer2,nBatch,actv,alp){
-
-  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
-                                             drpt=drpt,
-                                             layer1=layer1,
-                                             layer2=layer2,
-                                             nBatch=nBatch,
-                                             actv=actv,
-                                             alp=alp,
-                                             #IPA=NA,
-                                             #IPAse=NA,
-                                             valLoss=NA,
-                                             runtime=NA))
-
-  #paramTrials<-paramTrials[sample(nrow(paramTrials), baselineTrials), ]
-
-  # paramTrials<-unique.data.frame(paramTrials)
-  #paramTrials<-readRDS("complexTrials.rds")
-  for(i in 1:nrow(paramTrials)){
-    #paramTrials<-readRDS("complexTrials.rds")  
-    tempRes<-c()
-    tempVal<-c()
-    runTime<-c()
-
-    for(data in klist){
-      currentTime<-Sys.time() 
-      train<-data$train
-      val<-data$val
-      maxTime<-max(train$time)
-      sds<-sapply(train, function(x) (sd(x)))
-      means<-sapply(train, function(x) (mean(x)))
-      val<-normalizer(val,means,sds,maxTime)
-      train<-normalizer(train,means,sds,maxTime)
-      times<-seq(from=min(val$time),
-                 to=max(val$time),
-                 length.out = 22
-      )
-
-      times<-head(times, -1)
-      times<-tail(times, -1)
-
-      lr=paramTrials[i,]$lr
-      drpt=paramTrials[i,]$drpt
-      layer1=paramTrials[i,]$layer1
-      layer2=paramTrials[i,]$layer2
-      bsize=ceiling(nrow(train)/paramTrials[i,]$nBatch)
-      actv=paramTrials[i,]$actv
-      alp=paramTrials[i,]$alp
-      #=paramTrials[i,]$IPA
-      #=paramTrials[i,]$valLoss
-      source_python('src/deephitter.py')
-      hitnnSurv=fitDeephit(train,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
-      hitnnSurv<-t(hitnnSurv)
-      deephitCleaned<-pyProcess(hitnnSurv,times=times)
-      colnames(deephitCleaned)<-times
-      py_run_string("del fitDeephit")
-      rm(fitDeephit)
-
-      miniee <- Score(list('mod'=deephitCleaned),
-                      data =val, 
-                      formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                      se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
-
-      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
-      runTime<-append(runTime, Sys.time()-currentTime)
-      currentTime<-Sys.time() 
-    }
-
-    #paramTrials$IPA[i]<-mean(tempRes)
-    #paramTrials$IPAse[i]<-sd(tempRes)
-    paramTrials$valLoss[i]<-mean(tempVal)
-    paramTrials$runtime[i]<-mean(runTime)
-    ##saveRDS(paramTrials,"complexds.rds")
-    print(paste("trial",i))
-    print(sum(runTime))
-    #print(as.character(paramTrials[i,]))
-    #rm(cbnn,fit,annPreds,covars_input,covars_output)
-    gc()
-  }
-  bestPerformance<-paramTrials[which.min(paramTrials$valLoss),]
-  return(bestPerformance)
-}
-
-klist<-readRDS("flchainCV.rds")
-test<-readRDS("flchainTest.rds")
-times<-seq(from=min(test$time)+0.000001,
-           to=max(test$time)-0.000001,
-           length.out = 25
-)
-
-bestDH<-optimDeepHit(klist,test,times,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,nBatch,actv,alp)
+```{r optimDeepHit,eval=TRUE,hide=FALSE}
+klist<-readRDS("cvData/flchainCV.rds")
+bestDH<-optimDeepHit(klist,lr,drpt,layer1,layer2,nBatch,actv,alp)
 
@@
 
-```{r hyperopcbnn,eval=FALSE,hide=FALSE}
+```{r hyperopcbnn,eval=TRUE,hide=FALSE}
+klist<-readRDS("cvData/flchainCV.rds")
+cbnnPerformance<-optimCBNN(klist,lr,drpt,layer1,layer2,nBatch,actv)
 
-optimCBNN<-function(klist,test,times,lr,drpt,layer1,layer2,nBatch,actv){
-
-  # 
-  # for (i in 1:length(klist)){
-  #   #train<-sampleCaseBase(data = as.data.frame(klist[[i]]$train),time="time",event="status",ratio=100)
-  #   valcb<-sampleCaseBase(data = as.data.frame(klist[[i]]$val),time="time",event="status",ratio=100)
-  #   valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
-  #                          as.matrix(valcb[,ncol(valcb),drop=F])),
-  #                     as.matrix(valcb[,ncol(valcb)-1,drop=F]))
-  #   #valcbtensor[[1]][[2]]<-rep(train$offset[1],nrow(valcbtensor[[1]][[1]]))      
-  #   #klist[[i]]$train<-train
-  #   klist[[i]]$val<-valcbtensor
-  # }
-
-  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
-                                             drpt=drpt,
-                                             layer1=layer1,
-                                             layer2=layer2,
-                                             nBatch=nBatch,
-                                             actv=actv,
-                                             #IPA=NA,
-                                             #IPAse=NA,
-                                             valLoss=NA,
-                                             runtime=NA))
-
-  #paramTrials<-paramTrials[sample(nrow(paramTrials), 2), ]
-
-  # paramTrials<-unique.data.frame(paramTrials)
-  #paramTrials<-readRDS("complexTrials.rds")
-  for(i in 1:nrow(paramTrials)){
-    #paramTrials<-readRDS("complexTrials.rds")  
-    tempRes<-c()
-    tempVal<-c()
-    runTime<-c()
-
-    for(data in klist){
-
-      train<-data$train
-      val<-data$val
-      maxTime<-max(train$time)
-      sds<-sapply(train, function(x) (sd(x)))
-      means<-sapply(train, function(x) (mean(x)))
-      #val[[1]][[1]]<-as.matrix(normalizer(as.data.frame(val[[1]][[1]]),means,sds))
-      train<-normalizer(train,means,sds,maxTime)
-
-      #train[,ncol(train)]<-data$train[,ncol(train)]
-      val<-normalizer(val,means,sds,maxTime)
-      valTest<-val
-      times<-seq(from=min(val$time),
-                 to=max(val$time),
-                 length.out = 22
-      )
-
-      times<-head(times, -1)
-      times<-tail(times, -1)
-
-      train<-sampleCaseBase(data = as.data.frame(train),time="time",event="status",ratio=100)
-      valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)
-      val<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
-                     as.matrix(valcb[,ncol(valcb),drop=F])
-      ),
-      as.matrix(valcb[,ncol(valcb)-1,drop=F])
-      )
-      val[[1]][[2]]<-rep(train$offset[1],nrow(val[[1]][[1]]))
-      #val[[1]][[2]]<-rep(0,nrow(val[[1]][[1]]))
-
-      lr=paramTrials[i,]$lr
-      drpt=paramTrials[i,]$drpt
-      layer1=paramTrials[i,]$layer1
-      layer2=paramTrials[i,]$layer2
-      bsize=c(floor(nrow(train)/paramTrials[i,]$nBatch))
-      actv=paramTrials[i,]$actv
-
-      currentTime<-Sys.time()
-
-      covars_input<-layer_input(shape=c(length(colnames(train))-2),
-                                name = 'main_input')
-
-      covars_output<-covars_input%>% 
-        layer_dense(units=layer1,use_bias = T,activation = actv)%>%
-        layer_dropout(drpt)%>%
-        layer_dense(units=layer2,use_bias = T,activation = actv)%>%
-        layer_dropout(drpt)%>%
-        layer_dense(units=1,use_bias = T)
-
-      cbnn<-cbnnModel(features=colnames(train)[-ncol(train)],
-                      feature_input = covars_input,
-                      feature_output = covars_output,
-                      originalData = train,
-                      offset=train$offset,
-                      timeVar = "time",
-                      eventVar= "status",optimizer=optimizer_adam(learning_rate = lr)
-      )
-
-      fit<-fitSmoothHaz(cbnn,
-                        epochs=epo,
-                        batch_size=bsize,
-                        verbose=0,
-                        monitor="val_loss",
-                        #val_split=0.2,
-                        min_delta=min_delta,
-                        patience=patience,val=val)
-      #tempVal<-append(tempVal,min(fit$resultOfFit$metrics$val_loss))
-      annPreds<-aar(fit,
-                    times=times,
-                    x_test=valTest[,-c(ncol(valTest))]
-      )
-
-      rownames(annPreds)<-annPreds[,1]
-      annProperpoly<- t(annPreds[,-1])
-      class(annProperpoly)<-c("tunnel",class(annProperpoly))
-
-      miniee <- Score(list('mod'=annProperpoly),
-                      data =valTest, 
-                      formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                      se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
-
-      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
-
-      runTime<-append(runTime, Sys.time()-currentTime)
-      currentTime<-Sys.time()
-
-      rm(cbnn,fit,covars_input,covars_output)
-    }
-
-    paramTrials$valLoss[i]<-mean(tempVal)
-    paramTrials$runtime[i]<-mean(runTime)
-    ##saveRDS(paramTrials,"complexds.rds")
-    print(paste("trial",i,"out of ",nrow(paramTrials)))
-    print(sum(runTime))
-    #print(as.character(paramTrials[i,]))
-    #rm(cbnn,fit,annPreds,covars_input,covars_output)
-    gc()
-  }
-  bestPerformance<-paramTrials[which.min(paramTrials$valLoss),]
-  return(bestPerformance)
-}
-
-klist<-readRDS("flchainCV.rds")
-test<-readRDS("flchainTest.rds")
-
-cbnnPerformance<-optimCBNN(klist,test,times,lr,drpt,layer1,layer2,nBatch,actv)
-#cbnnPerformance<-cbnnPerformance[which.min(cbnnPerformance$valLoss),]
 ```
 
-```{r winners,eval=FALSE,hide=FALSE}
+```{r winners,eval=TRUE,hide=FALSE}
 winners<-list(cbnn=cbnnPerformance,
@@
 
-##saveRDS(winners,"flchainWins.rds")
+#saveRDS(winners,"cvData/flchainWins.rds")
 
@@
 
-winners<-readRDS("flchainWins.rds")
+winners<-readRDS("cvData/flchainWins.rds")
 
-klist<-readRDS("flchainCV.rds")
+klist<-readRDS("cvData/flchainCV.rds")
 data<-rbind(klist[[1]]$train,klist[[1]]$val)
-#data$time<-data$time/max(data$time)
 
@@
 
-  #train<-res$train
   val<-res$validate
@@
 
-  fullTest<-readRDS("flchainTest.rds")
+  fullTest<-readRDS("cvData/flchainTest.rds")
   test<-fullTest[,-c(ncol(data))]
@@
   linModelTrain<-as.data.frame(rbind(train,val))
-  #valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)#rbind(train,val)
-  #valcb$offset<-0
-  #valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
   times<-seq(from=min(test$time),
@@
   print(paste("current iter:",pos))
-  saveRDS(BrierIPAList,'results/100flchainBrier.rds')
-  saveRDS(cScore,'results/100flchaincidx.rds')
-  saveRDS(AUClist,'results/100flchainAuc.rds')
+  #saveRDS(BrierIPAList,'results/100flchainBrier.rds')
+  #saveRDS(cScore,'results/100flchaincidx.rds')
+  #saveRDS(AUClist,'results/100flchainAuc.rds')
   pos=pos+1

### 2023-04-05 analyses/optimMyeloma.Rmd
cleanedUpCode

@@
   iteration: 100
-  layer1: 12
-  layer2: 12
-  layer3: 100
-  layer4: 100
-  drpt: 0.1
-  lr: 0.01
 ---
@@
 eval_myeloma<-T
-#source_python('src/getSUP.py')
+
 source_python("src/cIndex.py")
-#source_python("src/dsmScript.py")
-#source_python('src/deephitter.py')
+
 ##########################
@@
 ##########################
-#bsize=512
-min_delta = 0 #10^-7
+
+min_delta = 0
 
@@
 iteration<-as.numeric(params$iteration)
-
-#baselineTrials<-2
 
@@
 
-```{r myelomaPrep,eval=FALSE,hide=FALSE}
+```{r myelomaPrep,eval=TRUE,hide=FALSE}
 
@@
 
-#saveRDS(res$test,"myelomaTest.rds")
+#saveRDS(res$test,"cvData/myelomaTest.rds")
 
@@
 }
-#saveRDS(klist,"myelomaCV.rds")
+#saveRDS(klist,"cvData/myelomaCV.rds")
 
@@
 
-```{r setFixedHyperparameters,eval=FALSE,hide=FALSE}
-
-klist<-readRDS("myelomaCV.rds")
-
-# lr=c(0.001)
-# drpt=c(0.25,0.5)
-# layer1=c(10,20,30)
-# layer2=c(10,20,30)
-# bsize=c(floor(nrow(klist[[1]]$train)/100))
-# actv=c("relu","linear")#,"tanh"
-# alp=c(0)
+```{r setFixedHyperparameters,eval=TRUE,hide=FALSE}
 
@@
 alp=c(0,0.5,1)
-# 
-# lr=c(0.001)
-# drpt=c(0.25)
-# layer1=c(20)
-# layer2=c(10)
-# nBatch=c(1000,100)
-# actv=c("relu")#
-# alp=c(0)
+
+lr=c(0.01)
+drpt=c(0.01)
+layer1=c(50)
+layer2=c(10)
+nBatch=c(100)
+actv=c("relu","linear")#
+alp=c(0)
 
@@
 
-```{r hyperOpDsurv,eval=FALSE,hide=FALSE}
-
-optimDeepSurv<-function(klist,test,baselineTrials,lr,drpt,layer1,layer2,nBatch,actv){
-
-  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
-                                             drpt=drpt,
-                                             layer1=layer1,
-                                             layer2=layer2,
-                                             nBatch=nBatch,
-                                             actv=actv,
-                                             #IPA=NA,
-                                             #IPAse=NA,
-                                             valLoss=NA,
-                                             runtime=NA))
-
-  #paramTrials<-paramTrials[sample(nrow(paramTrials), baselineTrials), ]
-
-  # paramTrials<-unique.data.frame(paramTrials)
-  #paramTrials<-readRDS("complexTrials.rds")
-  for(i in 1:nrow(paramTrials)){
-    #paramTrials<-readRDS("complexTrials.rds")  
-    tempRes<-c()
-    tempVal<-c()
-    runTime<-c()
-
-    for(data in klist){
-      currentTime<-Sys.time() 
-      train<-data$train
-      val<-data$val
-      maxTime<-max(train$time)
-      sds<-sapply(train, function(x) (sd(x)))
-      means<-sapply(train, function(x) (mean(x)))
-      val<-normalizer(val,means,sds,maxTime)
-      train<-normalizer(train,means,sds,maxTime)
-      times<-seq(from=min(val$time),
-                 to=max(val$time),
-                 length.out = 22
-      )
-
-      times<-head(times, -1)
-      times<-tail(times, -1)
-
-      lr=paramTrials[i,]$lr
-      drpt=paramTrials[i,]$drpt
-      layer1=paramTrials[i,]$layer1
-      layer2=paramTrials[i,]$layer2
-      bsize=ceiling(nrow(train)/paramTrials[i,]$nBatch)
-      actv=paramTrials[i,]$actv
-      #=paramTrials[i,]$IPA
-      #=paramTrials[i,]$valLoss
+```{r hyperOpDsurv,eval=TRUE,hide=FALSE}
 source_python('src/dsurv.py')
-
-      ctrain<-train
-      #ctrain$time<- train$time + runif(nrow(train),-0.000001,0.000001)
-
-      coxnnsurv=fitDeepSurv(ctrain,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
-
-      coxnnsurv<-t(coxnnsurv)
-      cnnCleaned<-pyProcess(coxnnsurv,times=times)
-      colnames(cnnCleaned)<-times
-      py_run_string("del fitDeepSurv")
-      rm(fitDeepSurv)
-
-      miniee <- Score(list('mod'=cnnCleaned),
-                      data =val, 
-                      formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                      se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
-
-      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
-      runTime<-append(runTime, Sys.time()-currentTime)
-      currentTime<-Sys.time()
-
-      # coxnnsurv<-coxnnsurv[[1]]
-      # coxnnsurv<-t(coxnnsurv)
-      # #colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
-      # cnnCleaned<-pyProcess(coxnnsurv,times=times)
-      # colnames(cnnCleaned)<-times#seq(0,1,length.out = ncol(cnnCleaned))
-      # py_run_string("del fitDeepSurv")
-      # rm(fitDeepSurv)
-      # brierFinalResults <- Score(list('ds'=cnnCleaned),
-      #                            data =as.data.frame(val),#rbind(train,val) 
-      #                            formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-      #                            se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
-      # cutted<-brierFinalResults[which(brierFinalResults$model=="ds"),]
-      #tempRes<-append(tempRes,mean(cutted[-c(1,2,24,25)]$IPA))
-    }
-
-    #paramTrials$IPA[i]<-mean(tempRes)
-    #paramTrials$IPAse[i]<-sd(tempRes)
-    paramTrials$valLoss[i]<-mean(tempVal)
-    paramTrials$runtime[i]<-mean(runTime)
-    ##saveRDS(paramTrials,"complexds.rds")
-    print(paste("trial",i))
-    print(sum(runTime))
-    #print(as.character(paramTrials[i,]))
-    #rm(cbnn,fit,annPreds,covars_input,covars_output)
-    gc()
-  }
-  bestPerformance<-paramTrials[which.min(paramTrials$valLoss),]
-  return(bestPerformance)
-}
-
-klist<-readRDS("myelomaCV.rds")
-test<-readRDS("myelomaTest.rds")
-source_python('src/dsurv.py')
-bestDS<-optimDeepSurv(klist,test,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,nBatch,actv)
+klist<-readRDS("cvData/myelomaCV.rds")
+bestDS<-optimDeepSurv(klist,lr,drpt,layer1,layer2,nBatch,actv)
 
@@
 
-```{r optimDeepHit,eval=FALSE,hide=FALSE}
-
-optimDeepHit<-function(klist,test,times,baselineTrials,lr,drpt,layer1,layer2,nBatch,actv,alp){
-
-  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
-                                             drpt=drpt,
-                                             layer1=layer1,
-                                             layer2=layer2,
-                                             nBatch=nBatch,
-                                             actv=actv,
-                                             alp=alp,
-                                             #IPA=NA,
-                                             #IPAse=NA,
-                                             valLoss=NA,
-                                             runtime=NA))
-
-  #paramTrials<-paramTrials[sample(nrow(paramTrials), baselineTrials), ]
-
-  # paramTrials<-unique.data.frame(paramTrials)
-  #paramTrials<-readRDS("complexTrials.rds")
-  for(i in 1:nrow(paramTrials)){
-    #paramTrials<-readRDS("complexTrials.rds")  
-    tempRes<-c()
-    tempVal<-c()
-    runTime<-c()
-
-    for(data in klist){
-      currentTime<-Sys.time() 
-      train<-data$train
-      val<-data$val
-      maxTime<-max(train$time)
-      sds<-sapply(train, function(x) (sd(x)))
-      means<-sapply(train, function(x) (mean(x)))
-      val<-normalizer(val,means,sds,maxTime)
-      train<-normalizer(train,means,sds,maxTime)
-      times<-seq(from=min(val$time),
-                 to=max(val$time),
-                 length.out = 22
-      )
-
-      times<-head(times, -1)
-      times<-tail(times, -1)
-
-      lr=paramTrials[i,]$lr
-      drpt=paramTrials[i,]$drpt
-      layer1=paramTrials[i,]$layer1
-      layer2=paramTrials[i,]$layer2
-      bsize=ceiling(nrow(train)/paramTrials[i,]$nBatch)
-      actv=paramTrials[i,]$actv
-      alp=paramTrials[i,]$alp
-      #=paramTrials[i,]$IPA
-      #=paramTrials[i,]$valLoss
-      source_python('src/deephitter.py')
-      hitnnSurv=fitDeephit(train,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
-      hitnnSurv<-t(hitnnSurv)
-      deephitCleaned<-pyProcess(hitnnSurv,times=times)
-      colnames(deephitCleaned)<-times
-      py_run_string("del fitDeephit")
-      rm(fitDeephit)
-
-      miniee <- Score(list('mod'=deephitCleaned),
-                      data =val, 
-                      formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                      se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
-
-      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
-      runTime<-append(runTime, Sys.time()-currentTime)
-      currentTime<-Sys.time() 
-    }
-
-    #paramTrials$IPA[i]<-mean(tempRes)
-    #paramTrials$IPAse[i]<-sd(tempRes)
-    paramTrials$valLoss[i]<-mean(tempVal)
-    paramTrials$runtime[i]<-mean(runTime)
-    ##saveRDS(paramTrials,"complexds.rds")
-    print(paste("trial",i))
-    print(sum(runTime))
-    #print(as.character(paramTrials[i,]))
-    #rm(cbnn,fit,annPreds,covars_input,covars_output)
-    gc()
-  }
-  bestPerformance<-paramTrials[which.min(paramTrials$valLoss),]
-  return(bestPerformance)
-}
-
-klist<-readRDS("myelomaCV.rds")
-test<-readRDS("myelomaTest.rds")
-times<-seq(from=min(test$time)+0.000001,
-           to=max(test$time)-0.000001,
-           length.out = 25
-)
-
-bestDH<-optimDeepHit(klist,test,times,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,nBatch,actv,alp)
+```{r optimDeepHit,eval=TRUE,hide=FALSE}
+klist<-readRDS("cvData/myelomaCV.rds")
+bestDH<-optimDeepHit(klist,lr,drpt,layer1,layer2,nBatch,actv,alp)
 
@@
 
-```{r hyperopcbnn,eval=FALSE,hide=FALSE}
+```{r hyperopcbnn,eval=TRUE,hide=FALSE}
+klist<-readRDS("cvData/myelomaCV.rds")
+cbnnPerformance<-optimCBNN(klist,lr,drpt,layer1,layer2,nBatch,actv)
 
-optimCBNN<-function(klist,test,times,lr,drpt,layer1,layer2,nBatch,actv){
-
-  # 
-  # for (i in 1:length(klist)){
-  #   #train<-sampleCaseBase(data = as.data.frame(klist[[i]]$train),time="time",event="status",ratio=100)
-  #   valcb<-sampleCaseBase(data = as.data.frame(klist[[i]]$val),time="time",event="status",ratio=100)
-  #   valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
-  #                          as.matrix(valcb[,ncol(valcb),drop=F])),
-  #                     as.matrix(valcb[,ncol(valcb)-1,drop=F]))
-  #   #valcbtensor[[1]][[2]]<-rep(train$offset[1],nrow(valcbtensor[[1]][[1]]))      
-  #   #klist[[i]]$train<-train
-  #   klist[[i]]$val<-valcbtensor
-  # }
-
-  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
-                                             drpt=drpt,
-                                             layer1=layer1,
-                                             layer2=layer2,
-                                             nBatch=nBatch,
-                                             actv=actv,
-                                             #IPA=NA,
-                                             #IPAse=NA,
-                                             valLoss=NA,
-                                             runtime=NA))
-
-  #paramTrials<-paramTrials[sample(nrow(paramTrials), 2), ]
-
-  # paramTrials<-unique.data.frame(paramTrials)
-  #paramTrials<-readRDS("complexTrials.rds")
-  for(i in 1:nrow(paramTrials)){
-    #paramTrials<-readRDS("complexTrials.rds")  
-    tempRes<-c()
-    tempVal<-c()
-    runTime<-c()
-
-    for(data in klist){
-
-      train<-data$train
-      val<-data$val
-      maxTime<-max(train$time)
-      sds<-sapply(train, function(x) (sd(x)))
-      means<-sapply(train, function(x) (mean(x)))
-      #val[[1]][[1]]<-as.matrix(normalizer(as.data.frame(val[[1]][[1]]),means,sds))
-      train<-normalizer(train,means,sds,maxTime)
-
-      #train[,ncol(train)]<-data$train[,ncol(train)]
-      val<-normalizer(val,means,sds,maxTime)
-      valTest<-val
-      times<-seq(from=min(val$time),
-                 to=max(val$time),
-                 length.out = 22
-      )
-
-      times<-head(times, -1)
-      times<-tail(times, -1)
-
-      train<-sampleCaseBase(data = as.data.frame(train),time="time",event="status",ratio=100)
-      valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)
-      val<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
-                     as.matrix(valcb[,ncol(valcb),drop=F])
-      ),
-      as.matrix(valcb[,ncol(valcb)-1,drop=F])
-      )
-      val[[1]][[2]]<-rep(train$offset[1],nrow(val[[1]][[1]]))
-      #val[[1]][[2]]<-rep(0,nrow(val[[1]][[1]]))
-
-      lr=paramTrials[i,]$lr
-      drpt=paramTrials[i,]$drpt
-      layer1=paramTrials[i,]$layer1
-      layer2=paramTrials[i,]$layer2
-      bsize=c(floor(nrow(train)/paramTrials[i,]$nBatch))
-      actv=paramTrials[i,]$actv
-
-      currentTime<-Sys.time()
-
-      covars_input<-layer_input(shape=c(length(colnames(train))-2),
-                                name = 'main_input')
-
-      covars_output<-covars_input%>% 
-        layer_dense(units=layer1,use_bias = T,activation = actv)%>%
-        layer_dropout(drpt)%>%
-        layer_dense(units=layer2,use_bias = T,activation = actv)%>%
-        layer_dropout(drpt)%>%
-        layer_dense(units=1,use_bias = T)
-
-      cbnn<-cbnnModel(features=colnames(train)[-ncol(train)],
-                      feature_input = covars_input,
-                      feature_output = covars_output,
-                      originalData = train,
-                      offset=train$offset,
-                      timeVar = "time",
-                      eventVar= "status",optimizer=optimizer_adam(learning_rate = lr)
-      )
-
-      fit<-fitSmoothHaz(cbnn,
-                        epochs=epo,
-                        batch_size=bsize,
-                        verbose=0,
-                        monitor="val_loss",
-                        #val_split=0.2,
-                        min_delta=min_delta,
-                        patience=patience,val=val)
-      #tempVal<-append(tempVal,min(fit$resultOfFit$metrics$val_loss))
-      annPreds<-aar(fit,
-                    times=times,
-                    x_test=valTest[,-c(ncol(valTest))]
-      )
-
-      rownames(annPreds)<-annPreds[,1]
-      annProperpoly<- t(annPreds[,-1])
-      class(annProperpoly)<-c("tunnel",class(annProperpoly))
-
-      miniee <- Score(list('mod'=annProperpoly),
-                      data =valTest, 
-                      formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                      se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
-
-      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
-
-      runTime<-append(runTime, Sys.time()-currentTime)
-      currentTime<-Sys.time()
-
-      rm(cbnn,fit,covars_input,covars_output)
-    }
-
-    paramTrials$valLoss[i]<-mean(tempVal)
-    paramTrials$runtime[i]<-mean(runTime)
-    ##saveRDS(paramTrials,"complexds.rds")
-    print(paste("trial",i,"out of ",nrow(paramTrials)))
-    print(sum(runTime))
-    #print(as.character(paramTrials[i,]))
-    #rm(cbnn,fit,annPreds,covars_input,covars_output)
-    gc()
-  }
-  bestPerformance<-paramTrials[which.min(paramTrials$valLoss),]
-  return(bestPerformance)
-}
-
-klist<-readRDS("myelomaCV.rds")
-test<-readRDS("myelomaTest.rds")
-
-cbnnPerformance<-optimCBNN(klist,test,times,lr,drpt,layer1,layer2,nBatch,actv)
-#cbnnPerformance<-cbnnPerformance[which.min(cbnnPerformance$valLoss),]
 ```
 
-```{r winners,eval=FALSE,hide=FALSE}
+```{r winners,eval=TRUE,hide=FALSE}
 winners<-list(cbnn=cbnnPerformance,
@@
               dh=bestDH)
-
-#saveRDS(winners,"myelomaWins.rds")
+#saveRDS(winners,"cvData/myelomaWins.rds")
 
@@
 
-winners<-readRDS("myelomaWins.rds")
+winners<-readRDS("cvData/myelomaWins.rds")
 
-klist<-readRDS("myelomaCV.rds")
+klist<-readRDS("cvData/myelomaCV.rds")
 data<-rbind(klist[[1]]$train,klist[[1]]$val)
-#data$time<-data$time/max(data$time)
 
@@
 
-  fullTest<-readRDS("myelomaTest.rds")
+  fullTest<-readRDS("cvData/myelomaTest.rds")
 
@@
   linModelTrain<-as.data.frame(rbind(train,val))
-  #valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)#rbind(train,val)
-  #valcb$offset<-0
-  #valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
+
   times<-seq(from=min(train$time),
@@
   actv=winners$cbnn$actv
-  # 
-  # 
-  #  lr=0.001
-  #  drpt=0.5
-  #  layer1=10
-  #  layer2=10
-  #  patience=10
-  #actv='linear'
 
@@
                     monitor="val_loss",
-                    #val_split=0.2,
                     min_delta=min_delta,
@@
 
-  ggplot(data=brierFinalResults$Brier$score, aes(x=times,y=IPA,col=model))+
-    geom_line()#+coord_cartesian(ylim=c(-0.1,0.2))
   rm(fit,cbnn,covars_input,covars_output)
@@
   print(paste("current iter:",pos))
-  saveRDS(BrierIPAList,'results/100myelomaBrier.rds')
-  saveRDS(cScore,'results/100myelomacidx.rds')
-  saveRDS(AUClist,'results/100myelomaAuc.rds')
+  #saveRDS(BrierIPAList,'results/100myelomaBrier.rds')
+  #saveRDS(cScore,'results/100myelomacidx.rds')
+  #saveRDS(AUClist,'results/100myelomaAuc.rds')
   pos=pos+1

### 2023-04-05 analyses/optimProstate.Rmd
cleanedUpCode

@@
 eval_prostate<-T
-#source_python('src/getSUP.py')
+
 source_python("src/cIndex.py")
-#source_python("src/dsmScript.py")
-#source_python('src/deephitter.py')
+
 ##########################
@@
 
-```{r prostatePrep,eval=FALSE,hide=FALSE}
+```{r prostatePrep,eval=TRUE,hide=FALSE}
 
@@
 
-#saveRDS(res$test,"prostateTest.rds")
+#saveRDS(res$test,"cvData/prostateTest.rds")
 
@@
 }
-#saveRDS(klist,"prostateCV.rds")
+#saveRDS(klist,"cvData/prostateCV.rds")
 
@@
 
-```{r setFixedHyperparameters,eval=FALSE,hide=FALSE}
+```{r setFixedHyperparameters,eval=TRUE,hide=FALSE}
 
-klist<-readRDS("prostateCV.rds")
-
-# 
-# lr=c(0.001,0.0005)
-# drpt=c(0.25,0.1)
-# layer1=c(20,50,100)
-# layer2=c(25,50)
-# nBatch=c(100,500,1000)
-# actv=c("relu","linear")#
-# alp=c(0.5)
+klist<-readRDS("cvData/prostateCV.rds")
 
@@
 
-#curl -d "support done" ntfy.sh/updatejykiscripts
-```
-
-```{r hyperOpDsurv,eval=FALSE,hide=FALSE}
-
-optimDeepSurv<-function(klist,test,baselineTrials,lr,drpt,layer1,layer2,nBatch,actv){
-
-  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
-                                             drpt=drpt,
-                                             layer1=layer1,
-                                             layer2=layer2,
-                                             nBatch=nBatch,
-                                             actv=actv,
-                                             #IPA=NA,
-                                             #IPAse=NA,
-                                             valLoss=NA,
-                                             runtime=NA))
-
-  #paramTrials<-paramTrials[sample(nrow(paramTrials), baselineTrials), ]
-
-  # paramTrials<-unique.data.frame(paramTrials)
-  #paramTrials<-readRDS("complexTrials.rds")
-  for(i in 1:nrow(paramTrials)){
-    #paramTrials<-readRDS("complexTrials.rds")  
-    tempRes<-c()
-    tempVal<-c()
-    runTime<-c()
-
-    for(data in klist){
-      currentTime<-Sys.time() 
-      train<-data$train
-      val<-data$val
-      maxTime<-max(train$time)
-      sds<-sapply(train, function(x) (sd(x)))
-      means<-sapply(train, function(x) (mean(x)))
-      val<-normalizer(val,means,sds,maxTime)
-      train<-normalizer(train,means,sds,maxTime)
-      times<-seq(from=min(val$time),
-                 to=max(val$time),
-                 length.out = 22
-      )
-
-      times<-head(times, -1)
-      times<-tail(times, -1)
-
-      lr=paramTrials[i,]$lr
-      drpt=paramTrials[i,]$drpt
-      layer1=paramTrials[i,]$layer1
-      layer2=paramTrials[i,]$layer2
-      bsize=ceiling(nrow(train)/paramTrials[i,]$nBatch)
-      actv=paramTrials[i,]$actv
-      #=paramTrials[i,]$IPA
-      #=paramTrials[i,]$valLoss
-      source_python('src/dsurv.py')
-
-      ctrain<-train
-      #ctrain$time<- train$time + runif(nrow(train),-0.000001,0.000001)
-
-      coxnnsurv=fitDeepSurv(ctrain,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
-
-      coxnnsurv<-t(coxnnsurv)
-      cnnCleaned<-pyProcess(coxnnsurv,times=times)
-      colnames(cnnCleaned)<-times
-      py_run_string("del fitDeepSurv")
-      rm(fitDeepSurv)
-
-      miniee <- Score(list('mod'=cnnCleaned),
-                      data =val, 
-                      formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                      se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
-
-      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
-      runTime<-append(runTime, Sys.time()-currentTime)
-      currentTime<-Sys.time()
-
-      # coxnnsurv<-coxnnsurv[[1]]
-      # coxnnsurv<-t(coxnnsurv)
-      # #colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
-      # cnnCleaned<-pyProcess(coxnnsurv,times=times)
-      # colnames(cnnCleaned)<-times#seq(0,1,length.out = ncol(cnnCleaned))
-      # py_run_string("del fitDeepSurv")
-      # rm(fitDeepSurv)
-      # brierFinalResults <- Score(list('ds'=cnnCleaned),
-      #                            data =as.data.frame(val),#rbind(train,val) 
-      #                            formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-      #                            se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
-      # cutted<-brierFinalResults[which(brierFinalResults$model=="ds"),]
-      #tempRes<-append(tempRes,mean(cutted[-c(1,2,24,25)]$IPA))
-    }
-
-    #paramTrials$IPA[i]<-mean(tempRes)
-    #paramTrials$IPAse[i]<-sd(tempRes)
-    paramTrials$valLoss[i]<-mean(tempVal)
-    paramTrials$runtime[i]<-mean(runTime)
-    #saveRDS(paramTrials,"complexds.rds")
-    print(paste("trial",i))
-    print(sum(runTime))
-    #print(as.character(paramTrials[i,]))
-    #rm(cbnn,fit,annPreds,covars_input,covars_output)
-    gc()
-  }
-  bestPerformance<-paramTrials[which.min(paramTrials$valLoss),]
-  return(bestPerformance)
-}
-
-klist<-readRDS("prostateCV.rds")
-test<-readRDS("prostateTest.rds")
-source_python('src/dsurv.py')
-bestDS<-optimDeepSurv(klist,test,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,nBatch,actv)
+lr=c(0.01)
+drpt=c(0.01)
+layer1=c(50)
+layer2=c(10)
+nBatch=c(100)
+actv=c("relu","linear")#
+alp=c(0)
 
@@
 
-```{r optimDeepHit,eval=FALSE,hide=FALSE}
+```{r hyperOpDsurv,eval=TRUE,hide=FALSE}
 
-optimDeepHit<-function(klist,test,times,baselineTrials,lr,drpt,layer1,layer2,nBatch,actv,alp){
-
-  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
-                                             drpt=drpt,
-                                             layer1=layer1,
-                                             layer2=layer2,
-                                             nBatch=nBatch,
-                                             actv=actv,
-                                             alp=alp,
-                                             #IPA=NA,
-                                             #IPAse=NA,
-                                             valLoss=NA,
-                                             runtime=NA))
-
-  #paramTrials<-paramTrials[sample(nrow(paramTrials), baselineTrials), ]
-
-  # paramTrials<-unique.data.frame(paramTrials)
-  #paramTrials<-readRDS("complexTrials.rds")
-  for(i in 1:nrow(paramTrials)){
-    #paramTrials<-readRDS("complexTrials.rds")  
-    tempRes<-c()
-    tempVal<-c()
-    runTime<-c()
-
-    for(data in klist){
-      currentTime<-Sys.time() 
-      train<-data$train
-      val<-data$val
-      maxTime<-max(train$time)
-      sds<-sapply(train, function(x) (sd(x)))
-      means<-sapply(train, function(x) (mean(x)))
-      val<-normalizer(val,means,sds,maxTime)
-      train<-normalizer(train,means,sds,maxTime)
-      times<-seq(from=min(val$time),
-                 to=max(val$time),
-                 length.out = 22
-      )
-
-      times<-head(times, -1)
-      times<-tail(times, -1)
-
-      lr=paramTrials[i,]$lr
-      drpt=paramTrials[i,]$drpt
-      layer1=paramTrials[i,]$layer1
-      layer2=paramTrials[i,]$layer2
-      bsize=ceiling(nrow(train)/paramTrials[i,]$nBatch)
-      actv=paramTrials[i,]$actv
-      alp=paramTrials[i,]$alp
-      #=paramTrials[i,]$IPA
-      #=paramTrials[i,]$valLoss
-      source_python('src/deephitter.py')
-      hitnnSurv=fitDeephit(train,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
-      hitnnSurv<-t(hitnnSurv)
-      deephitCleaned<-pyProcess(hitnnSurv,times=times)
-      colnames(deephitCleaned)<-times
-      py_run_string("del fitDeephit")
-      rm(fitDeephit)
-
-      miniee <- Score(list('mod'=deephitCleaned),
-                      data =val, 
-                      formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                      se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
-
-      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
-      runTime<-append(runTime, Sys.time()-currentTime)
-      currentTime<-Sys.time() 
-    }
-
-    #paramTrials$IPA[i]<-mean(tempRes)
-    #paramTrials$IPAse[i]<-sd(tempRes)
-    paramTrials$valLoss[i]<-mean(tempVal)
-    paramTrials$runtime[i]<-mean(runTime)
-    #saveRDS(paramTrials,"complexds.rds")
-    print(paste("trial",i))
-    print(sum(runTime))
-    #print(as.character(paramTrials[i,]))
-    #rm(cbnn,fit,annPreds,covars_input,covars_output)
-    gc()
-  }
-  bestPerformance<-paramTrials[which.min(paramTrials$valLoss),]
-  return(bestPerformance)
-}
-
-klist<-readRDS("prostateCV.rds")
-test<-readRDS("prostateTest.rds")
-times<-seq(from=min(test$time)+0.000001,
-           to=max(test$time)-0.000001,
-           length.out = 25
-)
-
-bestDH<-optimDeepHit(klist,test,times,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,nBatch,actv,alp)
+klist<-readRDS("cvData/prostateCV.rds")
+source_python('src/dsurv.py')
+bestDS<-optimDeepSurv(klist,lr,drpt,layer1,layer2,nBatch,actv)
 
@@
 
-```{r hyperopcbnn,eval=FALSE,hide=FALSE}
-
-optimCBNN<-function(klist,test,times,lr,drpt,layer1,layer2,nBatch,actv){
-
-  # 
-  # for (i in 1:length(klist)){
-  #   #train<-sampleCaseBase(data = as.data.frame(klist[[i]]$train),time="time",event="status",ratio=100)
-  #   valcb<-sampleCaseBase(data = as.data.frame(klist[[i]]$val),time="time",event="status",ratio=100)
-  #   valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
-  #                          as.matrix(valcb[,ncol(valcb),drop=F])),
-  #                     as.matrix(valcb[,ncol(valcb)-1,drop=F]))
-  #   #valcbtensor[[1]][[2]]<-rep(train$offset[1],nrow(valcbtensor[[1]][[1]]))      
-  #   #klist[[i]]$train<-train
-  #   klist[[i]]$val<-valcbtensor
-  # }
-
-  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
-                                             drpt=drpt,
-                                             layer1=layer1,
-                                             layer2=layer2,
-                                             nBatch=nBatch,
-                                             actv=actv,
-                                             #IPA=NA,
-                                             #IPAse=NA,
-                                             valLoss=NA,
-                                             runtime=NA))
-
-  #paramTrials<-paramTrials[sample(nrow(paramTrials), 2), ]
-
-  # paramTrials<-unique.data.frame(paramTrials)
-  #paramTrials<-readRDS("complexTrials.rds")
-  for(i in 1:nrow(paramTrials)){
-    #paramTrials<-readRDS("complexTrials.rds")  
-    tempRes<-c()
-    tempVal<-c()
-    runTime<-c()
-
-    for(data in klist){
-
-      train<-data$train
-      val<-data$val
-      maxTime<-max(train$time)
-      sds<-sapply(train, function(x) (sd(x)))
-      means<-sapply(train, function(x) (mean(x)))
-      #val[[1]][[1]]<-as.matrix(normalizer(as.data.frame(val[[1]][[1]]),means,sds))
-      train<-normalizer(train,means,sds,maxTime)
-
-      #train[,ncol(train)]<-data$train[,ncol(train)]
-      val<-normalizer(val,means,sds,maxTime)
-      valTest<-val
-      times<-seq(from=min(val$time),
-                 to=max(val$time),
-                 length.out = 22
-      )
-
-      times<-head(times, -1)
-      times<-tail(times, -1)
-
-      train<-sampleCaseBase(data = as.data.frame(train),time="time",event="status",ratio=100)
-      valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)
-      val<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
-                     as.matrix(valcb[,ncol(valcb),drop=F])
-      ),
-      as.matrix(valcb[,ncol(valcb)-1,drop=F])
-      )
-      val[[1]][[2]]<-rep(train$offset[1],nrow(val[[1]][[1]]))
-      #val[[1]][[2]]<-rep(0,nrow(val[[1]][[1]]))
-
-      lr=paramTrials[i,]$lr
-      drpt=paramTrials[i,]$drpt
-      layer1=paramTrials[i,]$layer1
-      layer2=paramTrials[i,]$layer2
-      bsize=c(floor(nrow(train)/paramTrials[i,]$nBatch))
-      actv=paramTrials[i,]$actv
-
-      currentTime<-Sys.time()
-
-      covars_input<-layer_input(shape=c(length(colnames(train))-2),
-                                name = 'main_input')
-
-      covars_output<-covars_input%>% 
-        layer_dense(units=layer1,use_bias = T,activation = actv)%>%
-        layer_dropout(drpt)%>%
-        layer_dense(units=layer2,use_bias = T,activation = actv)%>%
-        layer_dropout(drpt)%>%
-        layer_dense(units=1,use_bias = T)
-
-      cbnn<-cbnnModel(features=colnames(train)[-ncol(train)],
-                      feature_input = covars_input,
-                      feature_output = covars_output,
-                      originalData = train,
-                      offset=train$offset,
-                      timeVar = "time",
-                      eventVar= "status",optimizer=optimizer_adam(learning_rate = lr)
-      )
-
-      fit<-fitSmoothHaz(cbnn,
-                        epochs=epo,
-                        batch_size=bsize,
-                        verbose=0,
-                        monitor="val_loss",
-                        #val_split=0.2,
-                        min_delta=min_delta,
-                        patience=patience,val=val)
-      #tempVal<-append(tempVal,min(fit$resultOfFit$metrics$val_loss))
-      annPreds<-aar(fit,
-                    times=times,
-                    x_test=valTest[,-c(ncol(valTest))]
-      )
-
-      rownames(annPreds)<-annPreds[,1]
-      annProperpoly<- t(annPreds[,-1])
-      class(annProperpoly)<-c("tunnel",class(annProperpoly))
-
-      miniee <- Score(list('mod'=annProperpoly),
-                      data =valTest, 
-                      formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                      se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
-
-      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
-
-      runTime<-append(runTime, Sys.time()-currentTime)
-      currentTime<-Sys.time()
-
-      rm(cbnn,fit,covars_input,covars_output)
-    }
-
-    paramTrials$valLoss[i]<-mean(tempVal)
-    paramTrials$runtime[i]<-mean(runTime)
-    #saveRDS(paramTrials,"complexds.rds")
-    print(paste("trial",i,"out of ",nrow(paramTrials)))
-    print(sum(runTime))
-    #print(as.character(paramTrials[i,]))
-    #rm(cbnn,fit,annPreds,covars_input,covars_output)
-    gc()
-  }
-  bestPerformance<-paramTrials[which.min(paramTrials$valLoss),]
-  return(bestPerformance)
-}
-
-klist<-readRDS("prostateCV.rds")
-test<-readRDS("prostateTest.rds")
-
-cbnnPerformance<-optimCBNN(klist,test,times,lr,drpt,layer1,layer2,nBatch,actv)
-
-#cbnnPerformance<-cbnnPerformance[which.min(cbnnPerformance$valLoss),]
+```{r optimDeepHit,eval=TRUE,hide=FALSE}
+klist<-readRDS("cvData/prostateCV.rds")
+bestDH<-optimDeepHit(klist,lr,drpt,layer1,layer2,nBatch,actv,alp)
 
@@
 
-```{r winners,eval=FALSE,hide=FALSE}
+```{r hyperopcbnn,eval=TRUE,hide=FALSE}
+klist<-readRDS("cvData/prostateCV.rds")
+cbnnPerformance<-optimCBNN(klist,lr,drpt,layer1,layer2,nBatch,actv)
+
+```
+
+```{r winners,eval=TRUE,hide=FALSE}
+
 winners<-list(cbnn=cbnnPerformance,
@@
 
-#saveRDS(winners,"prostateWins.rds")
+#saveRDS(winners,"cvData/prostateWins.rds")
 
@@
 
-winners<-readRDS("prostateWins.rds")
+winners<-readRDS("cvData/prostateWins.rds")
 
-klist<-readRDS("prostateCV.rds")
+klist<-readRDS("cvData/prostateCV.rds")
 data<-rbind(klist[[1]]$train,klist[[1]]$val)
-#data$time<-data$time/max(data$time)
 
@@
 
-  fullTest<-readRDS("prostateTest.rds")
+  fullTest<-readRDS("cvData/prostateTest.rds")
   test<-fullTest[,-c(ncol(data))]
@@
   print(paste("current iter:",pos))
-  saveRDS(BrierIPAList,'results/100prostateBrier.rds')
-  saveRDS(cScore,'results/100prostatecidx.rds')
-  saveRDS(AUClist,'results/100prostateAuc.rds')
+  #saveRDS(BrierIPAList,'results/100prostateBrier.rds')
+  #saveRDS(cScore,'results/100prostatecidx.rds')
+  #saveRDS(AUClist,'results/100prostateAuc.rds')
   pos=pos+1

### 2023-04-05 analyses/optimizeComplex.Rmd
cleanedUpCode

@@
 eval_support<-T
-#source_python('src/getSUP.py')
+
 source_python("src/cIndex.py")
-#source_python("src/dsmScript.py")
-#source_python('src/deephitter.py')
+
 ##########################
@@
 
-```{r complexPrep,eval=FALSE,hide=FALSE}
+```{r complexPrep,eval=TRUE,hide=FALSE}
 
@@
     betas[["term1"]] * x[["z1"]]*t+
-    #betas[["term2"]] * x[["z1"]]*x[["z2"]]*t*t+
-
-    #betas[["term2"]] * x[["z1"]]*x[["z3"]]+
     betas[["term3"]] * x[["z2"]]*(x[["z3"]])
@@
 }
-#url<-'app.txt'
-#ds <- read.table(url, sep = '',header = TRUE)
-#ds$years.at.risk=pmin(ds$agein80,ds$onset)
-#brcancer$rectime<-brcancer$rectime
+
 true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1, data = brcancer, k = 3) #from flexsurv
@@
 coefficients<-true_mod$coefficients
-#coefficients<-c(coefficients,z1=-10,z2=1,z3=1,term1=0.005,term3=1)
+
 coefficients<-c(coefficients,z1=-5,z2=-1,z3=1,term1=0.001,term3=-1)#z1 time interaction
@@
 data<-data[,-1]
-hist(data$time[which(data$z1==0)],breaks=100)
-hist(data$time[which(data$z1==1)],breaks=100)
-median(data$time[which(data$z1==0)])
-median(data$time[which(data$z1==1)])
-#data$zsplit<-NULL
-#toDiv<-max(data$time)
-#data$time<-data$time/toDiv
+#hist(data$time[which(data$z1==0)],breaks=100)
+#hist(data$time[which(data$z1==1)],breaks=100)
+#median(data$time[which(data$z1==0)])
+#median(data$time[which(data$z1==1)])
+
 spec = c(train = .85, test = .15)
@@
 res = split(data, g)
-#saveRDS(res$test,"ComplexSimTest.rds")
+#saveRDS(res$test,"cvData/ComplexSimTest.rds")
 
@@
 }
-#saveRDS(klist,"ComplexSimCV.rds")
+#saveRDS(klist,"cvData/ComplexSimCV.rds")
 
@@
 
-```{r setFixedHyperparameters,eval=FALSE,hide=FALSE}
-
-klist<-readRDS("ComplexSimCV.rds")
-
-# lr=c(0.001,0.01)
-# drpt=c(0.1,0.25)
-# layer1=c(25,50,100)
-# layer2=c(25,50,100)
-# bsize=c(floor(nrow(klist[[1]]$train)/100))
-# actv=c("relu","tanh")#
-# alp=c(0,0.5,1)
+```{r setFixedHyperparameters,eval=TRUE,hide=FALSE}
 
@@
 alp=c(0,0.5,1)
-# 
-# lr=c(0.001)
-# drpt=c(0.01)
-# layer1=c(50)
-# layer2=c(10)
-# nBatch=c(100)
-# actv=c("relu")#
-# alp=c(0)
+
+lr=c(0.01)
+drpt=c(0.01)
+layer1=c(50)
+layer2=c(10)
+nBatch=c(100)
+actv=c("relu","linear")#
+alp=c(0)
 
@@
 
-```{r hyperOpDsurv,eval=FALSE,hide=FALSE}
-
-optimDeepSurv<-function(klist,test,baselineTrials,lr,drpt,layer1,layer2,nBatch,actv){
-
-  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
-                                             drpt=drpt,
-                                             layer1=layer1,
-                                             layer2=layer2,
-                                             nBatch=nBatch,
-                                             actv=actv,
-                                             #IPA=NA,
-                                             #IPAse=NA,
-                                             valLoss=NA,
-                                             runtime=NA))
-
-  #paramTrials<-paramTrials[sample(nrow(paramTrials), baselineTrials), ]
-
-  # paramTrials<-unique.data.frame(paramTrials)
-  #paramTrials<-readRDS("complexTrials.rds")
-  for(i in 1:nrow(paramTrials)){
-    #paramTrials<-readRDS("complexTrials.rds")  
-    tempRes<-c()
-    tempVal<-c()
-    runTime<-c()
-
-    for(data in klist){
-      currentTime<-Sys.time() 
-      train<-data$train
-      val<-data$val
-      maxTime<-max(train$time)
-      sds<-sapply(train, function(x) (sd(x)))
-      means<-sapply(train, function(x) (mean(x)))
-      val<-normalizer(val,means,sds,maxTime)
-      train<-normalizer(train,means,sds,maxTime)
-      times<-seq(from=min(val$time),
-                 to=max(val$time),
-                 length.out = 22
-      )
-
-      times<-head(times, -1)
-      times<-tail(times, -1)
-
-      lr=paramTrials[i,]$lr
-      drpt=paramTrials[i,]$drpt
-      layer1=paramTrials[i,]$layer1
-      layer2=paramTrials[i,]$layer2
-      bsize=ceiling(nrow(train)/paramTrials[i,]$nBatch)
-      actv=paramTrials[i,]$actv
-      #=paramTrials[i,]$IPA
-      #=paramTrials[i,]$valLoss
-      source_python('src/dsurv.py')
-
-      ctrain<-train
-      #ctrain$time<- train$time + runif(nrow(train),-0.000001,0.000001)
-
-      coxnnsurv=fitDeepSurv(ctrain,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
-
-      coxnnsurv<-t(coxnnsurv)
-      cnnCleaned<-pyProcess(coxnnsurv,times=times)
-      colnames(cnnCleaned)<-times
-      py_run_string("del fitDeepSurv")
-      rm(fitDeepSurv)
-
-      miniee <- Score(list('mod'=cnnCleaned),
-                      data =val, 
-                      formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                      se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
-
-      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
-      runTime<-append(runTime, Sys.time()-currentTime)
-      currentTime<-Sys.time()
-
-      # coxnnsurv<-coxnnsurv[[1]]
-      # coxnnsurv<-t(coxnnsurv)
-      # #colnames(coxnnsurv)<-seq(0,1,length.out = ncol(coxnnsurv))
-      # cnnCleaned<-pyProcess(coxnnsurv,times=times)
-      # colnames(cnnCleaned)<-times#seq(0,1,length.out = ncol(cnnCleaned))
-      # py_run_string("del fitDeepSurv")
-      # rm(fitDeepSurv)
-      # brierFinalResults <- Score(list('ds'=cnnCleaned),
-      #                            data =as.data.frame(val),#rbind(train,val) 
-      #                            formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-      #                            se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
-      # cutted<-brierFinalResults[which(brierFinalResults$model=="ds"),]
-      # tempRes<-append(tempRes,mean(cutted[-c(1,2,24,25)]$IPA))
-    }
-
-    #paramTrials$IPA[i]<-mean(tempRes)
-    #paramTrials$IPAse[i]<-sd(tempRes)
-    paramTrials$valLoss[i]<-mean(tempVal)
-    paramTrials$runtime[i]<-mean(runTime)
-    #saveRDS(paramTrials,"complexds.rds")
-    print(paste("trial",i))
-    print(sum(runTime))
-    #print(as.character(paramTrials[i,]))
-    #rm(cbnn,fit,annPreds,covars_input,covars_output)
-    gc()
-  }
-  bestPerformance<-paramTrials[which.min(paramTrials$valLoss),]
-  return(bestPerformance)
-}
-
-klist<-readRDS("ComplexSimCV.rds")
-test<-readRDS("ComplexSimTest.rds")
-times<-seq(from=min(test$time)+0.000001,
-           to=max(test$time)-0.000001,
-           length.out = 25
-)
-
-bestDS<-optimDeepSurv(klist,test,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,nBatch,actv)
-
+```{r hyperOpDsurv,eval=TRUE,hide=FALSE}
+klist<-readRDS("cvData/ComplexSimCV.rds")
+bestDS<-optimDeepSurv(klist,lr,drpt,layer1,layer2,nBatch,actv)
 ```
 
-```{r optimDeepHit,eval=FALSE,hide=FALSE}
-
-optimDeepHit<-function(klist,test,times,baselineTrials,lr,drpt,layer1,layer2,nBatch,actv,alp){
-
-  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
-                                             drpt=drpt,
-                                             layer1=layer1,
-                                             layer2=layer2,
-                                             nBatch=nBatch,
-                                             actv=actv,
-                                             alp=alp,
-                                             #IPA=NA,
-                                             #IPAse=NA,
-                                             valLoss=NA,
-                                             runtime=NA))
-
-  #paramTrials<-paramTrials[sample(nrow(paramTrials), baselineTrials), ]
-
-  # paramTrials<-unique.data.frame(paramTrials)
-  #paramTrials<-readRDS("complexTrials.rds")
-  for(i in 1:nrow(paramTrials)){
-    #paramTrials<-readRDS("complexTrials.rds")  
-    tempRes<-c()
-    tempVal<-c()
-    runTime<-c()
-
-    for(data in klist){
-      currentTime<-Sys.time() 
-      train<-data$train
-      val<-data$val
-      maxTime<-max(train$time)
-      sds<-sapply(train, function(x) (sd(x)))
-      means<-sapply(train, function(x) (mean(x)))
-      val<-normalizer(val,means,sds,maxTime)
-      train<-normalizer(train,means,sds,maxTime)
-      times<-seq(from=min(val$time),
-                 to=max(val$time),
-                 length.out = 22
-      )
-
-      times<-head(times, -1)
-      times<-tail(times, -1)
-
-      lr=paramTrials[i,]$lr
-      drpt=paramTrials[i,]$drpt
-      layer1=paramTrials[i,]$layer1
-      layer2=paramTrials[i,]$layer2
-      bsize=ceiling(nrow(train)/paramTrials[i,]$nBatch)
-      actv=paramTrials[i,]$actv
-      alp=paramTrials[i,]$alp
-      #=paramTrials[i,]$IPA
-      #=paramTrials[i,]$valLoss
-      source_python('src/deephitter.py')
-      hitnnSurv=fitDeephit(train,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv,alp=alp)
-      hitnnSurv<-t(hitnnSurv)
-      deephitCleaned<-pyProcess(hitnnSurv,times=times)
-      colnames(deephitCleaned)<-times
-      py_run_string("del fitDeephit")
-      rm(fitDeephit)
-
-      miniee <- Score(list('mod'=deephitCleaned),
-                      data =val, 
-                      formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                      se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
-
-      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
-      runTime<-append(runTime, Sys.time()-currentTime)
-      currentTime<-Sys.time() 
-    }
-
-    #paramTrials$IPA[i]<-mean(tempRes)
-    #paramTrials$IPAse[i]<-sd(tempRes)
-    paramTrials$valLoss[i]<-mean(tempVal)
-    paramTrials$runtime[i]<-mean(runTime)
-    #saveRDS(paramTrials,"complexds.rds")
-    print(paste("trial",i))
-    print(sum(runTime))
-    #print(as.character(paramTrials[i,]))
-    #rm(cbnn,fit,annPreds,covars_input,covars_output)
-    gc()
-  }
-  bestPerformance<-paramTrials[which.min(paramTrials$valLoss),]
-  return(bestPerformance)
-}
-
-klist<-readRDS("ComplexSimCV.rds")
-test<-readRDS("ComplexSimTest.rds")
-times<-seq(from=min(test$time)+0.000001,
-           to=max(test$time)-0.000001,
-           length.out = 25
-)
-
-bestDH<-optimDeepHit(klist,test,times,baselineTrials=baselineTrials,lr,drpt,layer1,layer2,nBatch,actv,alp)
-
+```{r optimDeepHit,eval=TRUE,hide=FALSE}
+klist<-readRDS("cvData/ComplexSimCV.rds")
+bestDH<-optimDeepHit(klist,lr,drpt,layer1,layer2,nBatch,actv,alp)
 ```
 
-```{r hyperopcbnn,eval=FALSE,hide=FALSE}
-
-optimCBNN<-function(klist,test,times,lr,drpt,layer1,layer2,nBatch,actv){
-
-  # 
-  # for (i in 1:length(klist)){
-  #   #train<-sampleCaseBase(data = as.data.frame(klist[[i]]$train),time="time",event="status",ratio=100)
-  #   valcb<-sampleCaseBase(data = as.data.frame(klist[[i]]$val),time="time",event="status",ratio=100)
-  #   valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
-  #                          as.matrix(valcb[,ncol(valcb),drop=F])),
-  #                     as.matrix(valcb[,ncol(valcb)-1,drop=F]))
-  #   #valcbtensor[[1]][[2]]<-rep(train$offset[1],nrow(valcbtensor[[1]][[1]]))      
-  #   #klist[[i]]$train<-train
-  #   klist[[i]]$val<-valcbtensor
-  # }
-
-  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
-                                             drpt=drpt,
-                                             layer1=layer1,
-                                             layer2=layer2,
-                                             nBatch=nBatch,
-                                             actv=actv,
-                                             #IPA=NA,
-                                             #IPAse=NA,
-                                             valLoss=NA,
-                                             runtime=NA))
-
-  #paramTrials<-paramTrials[sample(nrow(paramTrials), 2), ]
-
-  # paramTrials<-unique.data.frame(paramTrials)
-  #paramTrials<-readRDS("complexTrials.rds")
-  for(i in 1:nrow(paramTrials)){
-    #paramTrials<-readRDS("complexTrials.rds")  
-    tempRes<-c()
-    tempVal<-c()
-    runTime<-c()
-
-    for(data in klist){
-
-      train<-data$train
-      val<-data$val
-      maxTime<-max(train$time)
-      sds<-sapply(train, function(x) (sd(x)))
-      means<-sapply(train, function(x) (mean(x)))
-      #val[[1]][[1]]<-as.matrix(normalizer(as.data.frame(val[[1]][[1]]),means,sds))
-      train<-normalizer(train,means,sds,maxTime)
-
-      #train[,ncol(train)]<-data$train[,ncol(train)]
-      val<-normalizer(val,means,sds,maxTime)
-      valTest<-val
-      times<-seq(from=min(val$time),
-                 to=max(val$time),
-                 length.out = 22
-      )
-
-      times<-head(times, -1)
-      times<-tail(times, -1)
-
-      train<-sampleCaseBase(data = as.data.frame(train),time="time",event="status",ratio=100)
-      valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)
-      val<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),
-                     as.matrix(valcb[,ncol(valcb),drop=F])
-      ),
-      as.matrix(valcb[,ncol(valcb)-1,drop=F])
-      )
-      val[[1]][[2]]<-rep(train$offset[1],nrow(val[[1]][[1]]))
-      #val[[1]][[2]]<-rep(0,nrow(val[[1]][[1]]))
-
-      lr=paramTrials[i,]$lr
-      drpt=paramTrials[i,]$drpt
-      layer1=paramTrials[i,]$layer1
-      layer2=paramTrials[i,]$layer2
-      bsize=c(floor(nrow(train)/paramTrials[i,]$nBatch))
-      actv=paramTrials[i,]$actv
-
-      currentTime<-Sys.time()
-
-      covars_input<-layer_input(shape=c(length(colnames(train))-2),
-                                name = 'main_input')
-
-      covars_output<-covars_input%>% 
-        layer_dense(units=layer1,use_bias = T,activation = actv)%>%
-        layer_dropout(drpt)%>%
-        layer_dense(units=layer2,use_bias = T,activation = actv)%>%
-        layer_dropout(drpt)%>%
-        layer_dense(units=1,use_bias = T)
-
-      cbnn<-cbnnModel(features=colnames(train)[-ncol(train)],
-                      feature_input = covars_input,
-                      feature_output = covars_output,
-                      originalData = train,
-                      offset=train$offset,
-                      timeVar = "time",
-                      eventVar= "status",optimizer=optimizer_adam(learning_rate = lr)
-      )
-
-      fit<-fitSmoothHaz(cbnn,
-                        epochs=epo,
-                        batch_size=bsize,
-                        verbose=0,
-                        monitor="val_loss",
-                        #val_split=0.2,
-                        min_delta=min_delta,
-                        patience=patience,val=val)
-      #tempVal<-append(tempVal,min(fit$resultOfFit$metrics$val_loss))
-      annPreds<-aar(fit,
-                    times=times,
-                    x_test=valTest[,-c(ncol(valTest))]
-      )
-
-      rownames(annPreds)<-annPreds[,1]
-      annProperpoly<- t(annPreds[,-1])
-      class(annProperpoly)<-c("tunnel",class(annProperpoly))
-
-      miniee <- Score(list('mod'=annProperpoly),
-                      data =valTest, 
-                      formula = Hist(time, status != 0) ~ 1, summary = c("risks","IPA","ibs"), 
-                      se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)$Brier$score
-
-      tempVal<-append(tempVal,miniee$IBS[which(miniee$model=="mod" & miniee$times==max(miniee$times))])
-
-      runTime<-append(runTime, Sys.time()-currentTime)
-      currentTime<-Sys.time()
-
-      rm(cbnn,fit,covars_input,covars_output)
-    }
-
-    paramTrials$valLoss[i]<-mean(tempVal)
-    paramTrials$runtime[i]<-mean(runTime)
-    #saveRDS(paramTrials,"complexds.rds")
-    print(paste("trial",i,"out of ",nrow(paramTrials)))
-    print(sum(runTime))
-    #print(as.character(paramTrials[i,]))
-    #rm(cbnn,fit,annPreds,covars_input,covars_output)
-    gc()
-  }
-  bestPerformance<-paramTrials[which.min(paramTrials$valLoss),]
-  return(bestPerformance)
-}
-
-klist<-readRDS("ComplexSimCV.rds")
-test<-readRDS("ComplexSimTest.rds")
-
-cbnnPerformance<-optimCBNN(klist,test,times,lr,drpt,layer1,layer2,nBatch,actv)
-
-#cbnnPerformance<-cbnnPerformance[which.min(cbnnPerformance$valLoss),]
-
+```{r hyperopcbnn,eval=TRUE,hide=FALSE}
+klist<-readRDS("cvData/ComplexSimCV.rds")
+cbnnPerformance<-optimCBNN(klist,lr,drpt,layer1,layer2,nBatch,actv)
 ```
 
-```{r,eval=FALSE,hide=FALSE}
-
+```{r,eval=TRUE,hide=FALSE}
 winners<-list(cbnn=cbnnPerformance,
@@
               dh=bestDH)
-
-#saveRDS(winners,"complexOptimWins.rds")
+#saveRDS(winners,"cvData/complexOptimWins.rds")
 ```
@@
 
-winners<-readRDS("complexOptimWins.rds")
+winners<-readRDS("cvData/complexOptimWins.rds")
 
-klist<-readRDS("ComplexSimCV.rds")
+klist<-readRDS("cvData/ComplexSimCV.rds")
 data<-rbind(klist[[1]]$train,klist[[1]]$val)
@@
 
-  fullTest<-readRDS("ComplexSimTest.rds")
+  fullTest<-readRDS("cvData/ComplexSimTest.rds")
   test<-fullTest[,-c(ncol(data))]
@@
   linModelTrain<-as.data.frame(rbind(train,val))
-  #valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)#rbind(train,val)
-  #valcb$offset<-0
-  #valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
  times<-seq(from=min(test$time),
@@
 
-  #  lr=0.001
-  #  drpt=0.1
-  #  layer1=10
-  #  layer2=10
-  # bsize=c(floor(nrow(train)/100))# bsize=#winners$cbnn$bsize
-  #  actv='relu'
+  currentTime<-Sys.time()
 
-  currentTime<-Sys.time() 
-  # valcbtensor[[1]][[2]]<-rep(mod_cb_glm$offset[1],nrow(valcbtensor[[1]][[1]]))
   covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
@@
   AUClist[[pos]]<-brierFinalResults$AUC$score
-
-  #ggplot(data=brierFinalResults$Brier$score, aes(x=times*maxTime,y=IPA,col=model))+
-  #  geom_line()#+coord_cartesian(ylim=c(0.6,1))
 
@@
   print(paste("current iter:",pos))
-  saveRDS(BrierIPAList,'results/100complexBrier.rds')
-  saveRDS(cScore,'results/100complexcidx.rds')
-  saveRDS(AUClist,'results/100complexAuc.rds')
+  #saveRDS(BrierIPAList,'results/100complexBrier.rds')
+  #saveRDS(cScore,'results/100complexcidx.rds')
+  #saveRDS(AUClist,'results/100complexAuc.rds')
   pos=pos+1
@@
 ```{r optimDSM,eval=F}
-klist<-readRDS("ComplexSimCV.rds")
-test<-readRDS("ComplexSimTest.rds")
+klist<-readRDS("cvData/ComplexSimCV.rds")
+test<-readRDS("cvData/ComplexSimTest.rds")
 times<-seq(from=min(test$time)+0.000001,
@@
 
-# paramTrials<-unique.data.frame(paramTrials)
-#paramTrials<-readRDS("complexTrials.rds")
 for(i in 1:nrow(paramTrials)){
-  #paramTrials<-readRDS("complexTrials.rds")  
   tempRes<-c()
@@
   paramTrials$runtime[i]<-mean(runTime)
-  #saveRDS(paramTrials,"complexds.rds")
   print(paste("trial",i))

### 2023-04-05 analyses/plotter.Rmd
cleanedUpCode

--- 2023-04-05 analyses/plotterUpdated.Rmd
@@
 
-cMax<-max(readRDS("ComplexSimTest.rds")$time)/365
-mMax<-max(readRDS("myelomaTest.rds")$time)/365
-fMax<-max(readRDS("flchainTest.rds")$time)/365
-pMax<-max(readRDS("prostateTest.rds")$time)/12
+cMax<-max(readRDS("cvData/ComplexSimTest.rds")$time)/365
+mMax<-max(readRDS("cvData/myelomaTest.rds")$time)/365
+fMax<-max(readRDS("cvData/flchainTest.rds")$time)/365
+pMax<-max(readRDS("cvData/prostateTest.rds")$time)/12
 
@@
 
-cwtab<-getWinsTable("complexOptimWins.rds","A: Complex")
-mwtab<-getWinsTable("myelomaWins.rds","B: MM")
-fwtab<-getWinsTable("flchainWins.rds","C: FLC")
-pwtab<-getWinsTable("prostateWins.rds","D: Prostate")
+cwtab<-getWinsTable("cvData/complexOptimWins.rds","A: Complex")
+mwtab<-getWinsTable("cvData/myelomaWins.rds","B: MM")
+fwtab<-getWinsTable("cvData/flchainWins.rds","C: FLC")
+pwtab<-getWinsTable("cvData/prostateWins.rds","D: Prostate")
 
@@
 
-klist<-readRDS("ComplexSimCV.rds")
+klist<-readRDS("cvData/ComplexSimCV.rds")
 data<-rbind(klist[[1]]$train,klist[[1]]$val)
@@
 
-fullTest<-readRDS("ComplexSimTest.rds")
+fullTest<-readRDS("cvData/ComplexSimTest.rds")
 test<-fullTest[,-c(ncol(data))]

### 2023-04-05 analyses/SIMpaperResultscopy/CopyOfsrc/cIndex.py

@@ -1,10 +1,0 @@
-#We evaluate the performance of DSM in its discriminative ability (Time Dependent Concordance Index and Cumulative Dynamic AUC) as well as Brier Score.
-from sksurv.metrics import concordance_index_ipcw
-import numpy as np
-def cidx(et_train, et_test, out_risk): 
-  train = np.array([(et_train[i,0], et_train[i,1]) for i in range(np.shape(et_train)[0])],
-                   dtype = [('e', bool), ('t', float)])
-  test = np.array([(et_test[i,0], et_test[i,1]) for i in range(np.shape(et_test)[0])],
-                   dtype = [('e', bool), ('t', float)])
-  return concordance_index_ipcw(train, test, out_risk[:,0])[0]
-

### 2023-04-05 analyses/SIMpaperResultscopy/CopyOfsrc/deephitter.py

@@ -1,69 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-
-# For preprocessing
-from sklearn.preprocessing import StandardScaler
-from sklearn_pandas import DataFrameMapper 
-
-import torch # For building the networks 
-import torchtuples as tt # Some useful functions
-
-from pycox.models import DeepHitSingle
-from pycox.evaluation import EvalSurv
-
-
-
-
-def fitDeephit(train,fullTest,bsize,epochs,ti='time',ev='status',patience=10,min_delta=10**-7,drpt=0.5):#,epoch,bsize,tPos,ePos):
-  n= np.shape(train)[0]
-  vl_size = int(n*0.20)
-  df_val = train.sample(frac=0.2)
-  df_train = train.drop(df_val.index)
-  df_test = fullTest
-  varNames = list(df_train.columns)
-  unwanted = {ti, ev}
-  varNames = [e for e in varNames if e not in unwanted]
-  leave = [(col, None) for col in varNames]
-
-  x_mapper = DataFrameMapper(leave)
-
-  x_train = x_mapper.fit_transform(df_train).astype('float32')
-  x_val = x_mapper.transform(df_val).astype('float32')
-  x_test = x_mapper.transform(df_test).astype('float32')
- 
-
-
-
-  ########
-  #prepare labels
-  #######
-  num_durations = 10
-  labtrans = DeepHitSingle.label_transform(num_durations)
-  get_target = lambda df: (df[ti].values, df[ev].values)
-  y_train = labtrans.fit_transform(*get_target(df_train))
-  y_val = labtrans.transform(*get_target(df_val))
-  train = (x_train, y_train)
-  val = (x_val, y_val)
-  durations_test, events_test = get_target(df_test)
-  ########
-  #prepare model
-  ########
-  in_features = x_train.shape[1]
-  num_nodes = [50,50,25,25]
-  out_features = labtrans.out_features
-  batch_norm = False
-  dropout = drpt
-  
-  net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout)
-  model = DeepHitSingle(net, tt.optim.Adam, alpha=0.5, sigma=0.1, duration_index=labtrans.cuts)
-  lr_finder = model.lr_finder(x_train, y_train, batch_size=int(bsize), tolerance=2)
-  #_ = lr_finder.plot()
-  ############
-  #fit models
-  ###########
[…]

### 2023-04-05 analyses/SIMpaperResultscopy/CopyOfsrc/deepsurver.py

@@ -1,74 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-from sklearn.preprocessing import StandardScaler
-from sklearn_pandas import DataFrameMapper
-
-import torch
-import torchtuples as tt
-
-from pycox.datasets import metabric
-from pycox.models import CoxPH
-from pycox.evaluation import EvalSurv
-
-
-
-def fitDeepSurv(train,fullTest,bsize,epochs,valida,ti='time',ev='status',patience=10,min_delta=10**-7,drpt=0.5,lay1=200,lay2=20):#,epoch,bsize,tPos,ePos):
-  n= np.shape(train)[0]
-  vl_size = int(n*0.20)
-  df_val = valida#train.sample(frac=0.2)
-  df_train = train#.drop(df_val.index)
-  df_test = fullTest
-  varNames = list(df_train.columns)
-  unwanted = {ti, ev}
-  varNames = [e for e in varNames if e not in unwanted]
-  leave = [(col, None) for col in varNames]
-
-  x_mapper = DataFrameMapper(leave)
-  
-  x_train = x_mapper.fit_transform(df_train).astype('float32')
-  x_val = x_mapper.transform(df_val).astype('float32')
-  x_test = x_mapper.transform(df_test).astype('float32')
-  get_target = lambda df: (df[ti].values, df[ev].values)
-  y_train = get_target(df_train)
-  y_val = get_target(df_val)
-  #y_train[0]=y_train[0]+(np.random.uniform(0,1,len(y_train[0]))/10000000000000000000)
-  #y_val[0]=y_val[0]+(np.random.uniform(0,1,len(y_val[0]))/10000000000000000000)
-  durations_test, events_test = get_target(df_test)
-  val = x_val, y_val
-  
-  
-  in_features = x_train.shape[1]
-  num_nodes = [int(lay1),int(lay2)]
-  out_features = 1
-  batch_norm = True
-  dropout = drpt
-  
-  #net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout,activation="selu")
-  net = torch.nn.Sequential(
-     torch.nn.BatchNorm1d(in_features),
-     torch.nn.Linear(in_features,int(lay1)),
-     torch.nn.Tanh(),
-     torch.nn.Dropout(float(dropout)),
-
-     torch.nn.Linear(int(lay1), int(lay2)),
-     torch.nn.Tanh(),
[…]

### 2023-04-05 analyses/SIMpaperResultscopy/CopyOfsrc/dsmScript.py

@@ -1,70 +1,0 @@
-import numpy as np
-import random
-#from dsm import datasets
-from sklearn.model_selection import ParameterGrid
-from dsm import DeepSurvivalMachines
-##in dsm/utilities.py, we set early stopping patience and min_delta manually
-##dsm/dsm_torch.py, line 89 dropout(0.5) layer was added
-
-
-def dsmfitter(train,fullTest,bsize,epochs,tPos=3,ePos=4):
-  
-  ###############################
-  ###Prepare data
-  ###############################
-  tPos=int(tPos)
-  ePos=int(ePos)
-  train2=train.to_numpy()#@np.random.shuffle(train) 
-  np.random.shuffle(train2) 
-  n= np.shape(train2)[0]
-  vl_size = int(n*0.80)
-
-  
-  t_train = np.asarray( train2[:vl_size,tPos])
-  e_train = np.asarray( train2[:vl_size,ePos]).astype(int)
-  x_train = train2[:vl_size,:]
-  x_train = np.delete(x_train, [tPos,ePos], 1)
-  
-  t_val = np.asarray( train2[-vl_size:,tPos])
-  e_val = np.asarray( train2[-vl_size:,ePos]).astype(int)
-  x_val = train2[-vl_size:,:]
-  x_val = np.delete(x_val, [tPos,ePos], 1)
-
-  test=fullTest.to_numpy()
-  t_test = np.asarray( test[:,tPos])
-  e_test = np.asarray( test[:,ePos]).astype(int)
-  x_test = np.delete(test, [tPos,ePos], 1)
-  
-  horizons = np.arange(0, 1, 0.01).tolist()
-  times= np.quantile(t_test[e_test==1], horizons).tolist()
-  
-  
-
-  ###############################
-  ###Prepare data
-  ###############################
-  #I fix the parameter grid
-  param_grid = {'k' : [6],
-                'distribution' : ['Weibull'],
-                'learning_rate' : [ 0.001],
-                'layers' : [ [50,50,25,25] ]
-               }
-  params = ParameterGrid(param_grid)
-  
-  #note that the DeepSurvivalMachines function has been modified: 
-  #it now introduces a dropout layer after each feedforward dense layer.
-  # ### Model Training and Selection
-  models = []
-  for param in params:
-      model = DeepSurvivalMachines(k = param['k'],
-                                   distribution = param['distribution'],
-                                   layers = param['layers'])
[…]

### 2023-04-05 analyses/SIMpaperResultscopy/CopyOfsrc/functions.R

@@ -1,464 +1,0 @@
-
-statCalcSim<-function(metricSims){
-  toRemove <- which(apply(metricSims, 1, function(x){any(is.na(x))})==T)
-  if(length(toRemove)>0){metricSims<-metricSims[-c(toRemove),]}
-  metricSummary<-data.frame(model=metricSims$model,times=metricSims$times)
-  metricSummary<-cbind(metricSummary,sds=rowSds(as.matrix(metricSims[,-c(1,2)])),
-                       means=rowMeans(as.matrix(metricSims[,-c(1,2)])))
-  ci<-apply(as.matrix(metricSims[,-c(1,2)]), 1, function(x){mean(x)+c(-1.96,1.96)*sd(x)/sqrt(length(x))})
-  rownames(ci)<-c('cilower95','ciupper95')
-  pi<-apply(as.matrix(metricSims[,-c(1,2)]), 1, function(x){quantile(x,c(.025, .975))})
-  rownames(pi)<-c('pilower95','piupper95')
-  metricSummary<-cbind(metricSummary,t(ci),t(pi))
-  return(metricSummary)
-}
-
-plotPerformance<-function(metricSummary,xlab="Follow-up-time",method='Metric',iterations='NA'){
-  mainTitle<-paste(method,' : 95% P.I. over ', iterations, ' iterations',sep='')
-  
-  a<-ggplot(data=metricSummary, aes(x=times,y=means,col=model))+
-    geom_line() +
-    geom_ribbon(aes(ymin=pilower95,ymax=piupper95,fill=model),linetype=0,alpha=0.3) +
-    ggtitle(mainTitle) + 
-    xlab(xlab) +
-    ylab(method)
-  if(method=="IPA"){
-    a<-a+coord_cartesian()#ylim(-1,1))#
-    return(a)
-  }
-  return(a+coord_cartesian())
-}
-
-
-pyProcess<-function(predictions,times=times,apr="constant"){
-  interpolated <- apply(predictions, 1, function(x) approx(as.numeric(colnames(predictions)), y=x, method = apr, xout=times)$y)
-  class(interpolated)<-c("tunnel",class(interpolated))
-  interpolated<-1-interpolated
-  interpolated[is.na(interpolated) & is.na(interpolated)>50] <- 0
-  interpolated[is.na(interpolated) & is.na(interpolated)<50] <- 1
-  return(t(interpolated))
-}
-
-
-
-
[…]

### 2023-04-05 analyses/SIMpaperResultscopy/CopyOfsrc/getMetabricData.py

@@ -1,6 +1,0 @@
-from pycox import datasets
-def getMetabric():
-  return datasets.metabric.read_df()
-  
-  
-  

### 2023-04-05 analyses/SIMpaperResultscopy/CopyOfsrc/getSUP.py

@@ -1,8 +1,0 @@
-#import dsm
-#def getSupportDSM():
-#  return dsm.datasets.load_dataset('SUPPORT')
-
-
-import pycox
-def getSupportPycox():
-  return pycox.datasets.support.read_df()

### 2023-04-05 analyses/SIMpaperResultscopy/CopyOfsrc/oldmortDeephitter.py

@@ -1,81 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-
-# For preprocessing
-from sklearn.preprocessing import StandardScaler
-from sklearn_pandas import DataFrameMapper 
-
-import torch # For building the networks 
-import torchtuples as tt # Some useful functions
-
-from pycox.models import DeepHitSingle
-from pycox.evaluation import EvalSurv
-
-
-
-
-def fitDeephit(train,fullTest,bsize,epochs,valida,ti='time',ev='status',patience=10,min_delta=10**-7,drpt=0.5,lay1=200,lay2=20):#,epoch,bsize,tPos,ePos):
-  n= np.shape(train)[0]
-  vl_size = int(n*0.20)
-  df_val = valida#train.sample(frac=0.2)
-  df_train = train#.drop(df_val.index)
-  df_test = fullTest
-  varNames = list(df_train.columns)
-  unwanted = {ti, ev}
-  varNames = [e for e in varNames if e not in unwanted]
-  leave = [(col, None) for col in varNames]
-
-  x_mapper = DataFrameMapper(leave)
-
-  x_train = x_mapper.fit_transform(df_train).astype('float32')
-  x_val = x_mapper.transform(df_val).astype('float32')
-  x_test = x_mapper.transform(df_test).astype('float32')
- 
-
-
-
-  ########
-  #prepare labels
-  #######
-  num_durations = 10
-  labtrans = DeepHitSingle.label_transform(num_durations)
-  get_target = lambda df: (df[ti].values, df[ev].values)
-  y_train = labtrans.fit_transform(*get_target(df_train))
-  y_val = labtrans.transform(*get_target(df_val))
-  train = (x_train, y_train)
-  val = (x_val, y_val)
-  durations_test, events_test = get_target(df_test)
-  ########
-  #prepare model
-  ########
-  in_features = x_train.shape[1]
-  num_nodes =[int(lay1),int(lay2)]
-  out_features = labtrans.out_features
-  batch_norm = True
-  dropout = drpt
-  
-  #net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout,activation=torch.nn.SELU)
-  net = torch.nn.Sequential(
-     #torch.nn.BatchNorm1d(in_features),
-     torch.nn.Linear(in_features,int(lay1)),
-     torch.nn.Tanh(),
[…]

### 2023-04-05 analyses/SIMpaperResultscopy/CopyOfsrc/originalpyScripts/cIndex.py

@@ -1,38 +1,0 @@
-
-  # ### Evaluation
-  # 
-  # We evaluate the performance of DSM in its discriminative ability (Time Dependent Concordance Index and Cumulative Dynamic AUC) as well as Brier Score.
-from sksurv.metrics import concordance_index_ipcw
-import numpy as np
-#outrisk is the prediction at time t for the cumulative incidence
-def cidx(et_train, et_test, out_risk, time): 
-  #print(et_train,"\n\n", et_test,"\n\n", out_risk,"\n\n", time)
-  #train= et_train
-  #test=et_test
-  #train= (np.array(et_train[:,0],dtype = bool),np.array(et_train[:,1],dtype = bool))
-  #test= (np.array(et_test[:,0],dtype = bool),np.array(et_test[:,1],dtype = bool))
-  train = np.array([(et_train[i,0], et_train[i,1]) for i in range(np.shape(et_train)[0])],
-                   dtype = [('e', bool), ('t', float)])
-  test = np.array([(et_test[i,0], et_test[i,1]) for i in range(np.shape(et_test)[0])],
-                   dtype = [('e', bool), ('t', float)])
-  #print(test)          
-  #return 
-  return concordance_index_ipcw(train, test, out_risk[:,0], time)[0]
-  
-  #et_train = np.array([(e_train[i], t_train[i]) for i in range(len(e_train))],
-  #                 dtype = [('e', bool), ('t', float)])
-  #et_test = np.array([(e_test[i], t_test[i]) for i in range(len(e_test))],
-  #                 dtype = [('e', bool), ('t', float)])
-  #et_val = np.array([(e_val[i], t_val[i]) for i in range(len(e_val))],
-  #                 dtype = [('e', bool), ('t', float)])
-  #for i, _ in enumerate(times):
-  #    cis.append(concordance_index_ipcw(et_train, et_test, out_risk[:, i], times[i])[0])
-  #brs.append(brier_score(et_train, et_test, out_survival, times)[1])
-  #roc_auc = []
-  #for i, _ in enumerate(times):
-  #    roc_auc.append(cumulative_dynamic_auc(et_train, et_test, out_risk[:, i], times[i])[0])
-  #for horizon in enumerate(horizons):
-  #    print(f"For {horizon[1]} quantile,")
-  #    print("TD Concordance Index:", cis[horizon[0]])
-  #    print("Brier Score:", brs[0][horizon[0]])
-  #    print("ROC AUC ", roc_auc[horizon[0]][0], "\n")

### 2023-04-05 analyses/SIMpaperResultscopy/CopyOfsrc/originalpyScripts/deephitter.py

@@ -1,77 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-
-# For preprocessing
-from sklearn.preprocessing import StandardScaler
-from sklearn_pandas import DataFrameMapper 
-
-import torch # For building the networks 
-import torchtuples as tt # Some useful functions
-
-from pycox.models import DeepHitSingle
-from pycox.evaluation import EvalSurv
-
-
-def fitDeephit(train,fullTest,bsize,epochs,ti='time',ev='status'):#,epoch,bsize,tPos,ePos):
-  #print(train)
-  n= np.shape(train)[0]
-  vl_size = int(n*0.20)
-  df_val = train.sample(frac=0.2)
-  df_train = train.drop(df_val.index)
-  df_test = fullTest
-  varNames = list(df_train.columns)
-  unwanted = {ti, ev}
-  varNames = [e for e in varNames if e not in unwanted]
-  #print(varNames)
-  leave = [(col, None) for col in varNames]
-
-  x_mapper = DataFrameMapper(leave)
-  #print(x_mapper)
-
-  x_train = x_mapper.fit_transform(df_train).astype('float32')
-  x_val = x_mapper.transform(df_val).astype('float32')
-  x_test = x_mapper.transform(df_test).astype('float32')
-  
-  #x_train = df_train.drop(columns=['time','status'],inplace=False)
-  #x_val = df_val.drop(columns=['time','status'],inplace=False)
-  #x_test = df_test.drop(columns=['time','status'],inplace=False)
-  #cols_standardize = ['x0', 'x1', 'x2', 'x3', 'x8']
-  #cols_leave = ['x4', 'x5', 'x6', 'x7']
-
- 
-
-
-
-  ########
-  #prepare labels
-  #######
-  num_durations = 10
-  labtrans = DeepHitSingle.label_transform(num_durations)
-  get_target = lambda df: (df[ti].values, df[ev].values)
-  y_train = labtrans.fit_transform(*get_target(df_train))
-  y_val = labtrans.transform(*get_target(df_val))
-  train = (x_train, y_train)
-  val = (x_val, y_val)
-  durations_test, events_test = get_target(df_test)
-  ########
-  #prepare model
-  ########
-  in_features = x_train.shape[1]
-  num_nodes = [50,50,25,25]
-  out_features = labtrans.out_features
-  batch_norm = False
-  dropout = 0.5
-  
[…]

### 2023-04-05 analyses/SIMpaperResultscopy/CopyOfsrc/originalpyScripts/dsmScript.py

@@ -1,110 +1,0 @@
-
-
-import numpy as np
-
-import random
-from dsm import datasets
-#x, t, e = datasets.load_dataset('SUPPORT')
-#t=t/10000
-from sklearn.model_selection import ParameterGrid
-from dsm import DeepSurvivalMachines
-#import numpy as np
-#split data
-#n = len(x)
-
-#tr_size = int(n*0.70)
-#vl_size = int(n*0.10)
-#te_size = int(n*0.20)
-
-#x_train, x_test, x_val = x[:tr_size], x[-te_size:], x[tr_size:tr_size+vl_size]
-#t_train, t_test, t_val = t[:tr_size], t[-te_size:], t[tr_size:tr_size+vl_size]
-#e_train, e_test, e_val = e[:tr_size], e[-te_size:], e[tr_size:tr_size+vl_size]
-
-
-
-
-#x = r.data.iloc[:,:3].to_numpy()
-#x
-#t =np.asarray( r.data.iloc[:,3])
-#t
-#e =np.asarray( r.data.iloc[:,4])
-#e
-
-def dsmfitter(train,fullTest,bsize,epochs,tPos=3,ePos=4):
-  tPos=int(tPos)
-  ePos=int(ePos)
-  train2=train.to_numpy()#@np.random.shuffle(train) 
-  np.random.shuffle(train2) 
-  n= np.shape(train2)[0]
-  vl_size = int(n*0.80)
-  #vls=random.sample(range(n),vl_size)
-  
-  t_train = np.asarray( train2[:vl_size,tPos])#,dtype=np.float32)
-  e_train = np.asarray( train2[:vl_size,ePos]).astype(int)
-  x_train = train2[:vl_size,:]#.to_numpy()
-  x_train = np.delete(x_train, [tPos,ePos], 1)
-  
-  t_val = np.asarray( train2[-vl_size:,tPos])#,dtype=np.float32)
-  e_val = np.asarray( train2[-vl_size:,ePos]).astype(int)
-  x_val = train2[-vl_size:,:]#.to_numpy()
-  x_val = np.delete(x_val, [tPos,ePos], 1)
-
-  #print(np.shape(train2))
-  #print(np.shape(x_train))
-  #print(np.shape(x_val))
-  test=fullTest.to_numpy()
-  
-  t_test = np.asarray( test[:,tPos])
-  e_test = np.asarray( test[:,ePos]).astype(int)
-  x_test = np.delete(test, [tPos,ePos], 1)
[…]

### 2023-04-05 analyses/SIMpaperResultscopy/CopyOfsrc/packages.R

@@ -1,53 +1,0 @@
-#code to install missing packages
-
-
-getPckg <- function(pckgName) {install.packages(paste(pckgName), repos = "http://cran.r-project.org")}
-getPckgBM <- function(pckgName) {BiocManager::install(paste(pckgName))}
-
-pckgNamesCran = c("casebase",
-                  "DT",
-                  "data.table",
-                  "flexsurv",
-                  "ggplot2",
-                  "keras",
-                  "matrixStats",
-                  "mltools",
-                  "RColorBrewer",
-                  "riskRegression",
-                  "pseudo",
-                  "simsurv",
-                  "splines",
-                  "survival",
-                  "survminer",
-                  "survivalmodels",
-                  "tensorflow",
-                  "visreg")
-
-for(pckgName in pckgNamesCran){
-  pckg = try(require(pckgName,character.only = TRUE))
-  if(!pckg) {
-    cat(paste("Installing '",pckgName, "' from CRAN\n",sep=""))
-    getPckg(pckgName)
-    require(pckgName,character.only = TRUE)
-  }
-}
-
-
-
-#biomaRt from bioconductor can give us gene name to symbol transformations.
-pckgNamesBioC<-c()#"TCGAbiolinks"
-
-for(pckgName in pckgNamesBioC){
-  pckg = try(require(pckgName, character.only = TRUE))
-  if(!pckg) {
-    cat(paste("Installing '",pckgName, "' from CRAN\n", sep=""))
-    getPckgBM(pckgName)
-    require(pckgName, character.only = TRUE)
-  }
-}
-
-
-
-
-
-

### 2023-04-05 analyses/runAll.sh

@@ -1,34 +1,0 @@
-#!/bin/bash
-
-cd ~/pmnnProjects/pmnnPlayground/reviewedAnalysisPMNN
-
-epo=2000
-patience=10
-iteration=100
-layer1=50
-layer2=50
-layer3=25
-layer4=25
-drpt=0.5
-
-
-#explaination of structure (overview)
-#domain knowledge is too specific to be understood at a glance
-
-params='list('"epo=$epo,patience=$patience,iteration=$iteration,layer1=$layer1,layer2=$layer2,layer3=$layer3,layer4=$layer4,drpt=$drpt"')'
-metabric='library(rmarkdown); rmarkdown::render("metabricStudy.Rmd",'"params=$params)"
-support='library(rmarkdown); rmarkdown::render("supportStudy.Rmd",'"params=$params)"
-simple='library(rmarkdown); rmarkdown::render("simpleSim.Rmd",'"params=$params)"
-complex='library(rmarkdown); rmarkdown::render("complexSim.Rmd",'"params=$params)"
-oldmort='library(rmarkdown); rmarkdown::render("oldmortality.Rmd",'"params=$params)"
-(trap 'kill 0' SIGINT; \
-sleep 1 && eval `echo Rscript -e "'"{$simple}"'"` && bash ~/phoneNotification.sh simpleTest && sleep 9 && eval `echo Rscript -e "'"{$complex}"'"` && bash ~/phoneNotification.sh complexTest \ &
-sleep 6 && eval `echo Rscript -e "'"{$support}"'"` && bash ~/phoneNotification.sh supportTest \ &
-sleep 3 && eval `echo Rscript -e "'"{$metabric}"'"` && bash ~/phoneNotification.sh metabricTest \ &
-sleep 6 && eval `echo Rscript -e "'"{$oldmort}"'"` && bash ~/phoneNotification.sh oldmortTest 
-) 
-
-
-
-
-Rscript -e '{library(rmarkdown); rmarkdown::render("oldmortality.Rmd",params=list(epo=2000,patience=10,iteration=2,layer1=200,layer2=20,layer3=25,layer4=25,drpt=0.1))}'

### 2023-04-05 analyses/src/functions.R

@@ -481,3 +481,308 @@ normalizer<-function(data,means,sds,maxTime){
   normalized$time<-data$time/maxTime
   return(normalized)
 }
+
+
+
+
+
+
+
+
+
+###########################
+## hyperparameter optimization routines
+###########################
+
+
+optimDeepSurv<-function(klist,lr,drpt,layer1,layer2,nBatch,actv){
+  
+  paramTrials<-as.data.frame(tidyr::crossing(lr=lr,
+                                             drpt=drpt,
+                                             layer1=layer1,
+                                             layer2=layer2,
+                                             nBatch=nBatch,
+                                             actv=actv,
+                                             valLoss=NA,
+                                             runtime=NA))
+  
+
+  
+  
+  for(i in 1:nrow(paramTrials)){
+    tempRes<-c()
+    tempVal<-c()
+    runTime<-c()
+    
+    for(data in klist){
+      currentTime<-Sys.time() 
+      train<-data$train
+      val<-data$val
+      maxTime<-max(train$time)
+      sds<-sapply(train, function(x) (sd(x)))
+      means<-sapply(train, function(x) (mean(x)))
+      val<-normalizer(val,means,sds,maxTime)
+      train<-normalizer(train,means,sds,maxTime)
+      times<-seq(from=min(val$time),
+                 to=max(val$time),
+                 length.out = 22
+      )
+      
+      
+      times<-head(times, -1)
+      times<-tail(times, -1)
+      
+      lr=paramTrials[i,]$lr
+      drpt=paramTrials[i,]$drpt
+      layer1=paramTrials[i,]$layer1
+      layer2=paramTrials[i,]$layer2
+      bsize=ceiling(nrow(train)/paramTrials[i,]$nBatch)
+      actv=paramTrials[i,]$actv
+
+      source_python('src/dsurv.py')
+      
+      ctrain<-train
+
+      
+      
+      coxnnsurv=fitDeepSurv(ctrain,val,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
+      
+      
+      
+      coxnnsurv<-t(coxnnsurv)
+      cnnCleaned<-pyProcess(coxnnsurv,times=times)
+      colnames(cnnCleaned)<-times
+      py_run_string("del fitDeepSurv")
+      rm(fitDeepSurv)
+      
+      miniee <- Score(list('mod'=cnnCleaned),
[…]

### 2023-04-05 analyses/src/oldmortDeephitter.py

@@ -1,83 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-
-# For preprocessing
-from sklearn.preprocessing import StandardScaler
-from sklearn_pandas import DataFrameMapper 
-
-import torch # For building the networks 
-import torchtuples as tt # Some useful functions
-
-from pycox.models import DeepHitSingle
-from pycox.evaluation import EvalSurv
-
-
-
-
-def fitDeephit(train,fullTest,bsize,epochs,valida,ti='time',ev='status',patience=10,min_delta=10**-7,drpt=0.5,lay1=200,lay2=20):#,epoch,bsize,tPos,ePos):
-  n= np.shape(train)[0]
-  vl_size = int(n*0.20)
-  df_val = valida#train.sample(frac=0.2)
-  df_train = train#.drop(df_val.index)
-  df_test = fullTest
-  varNames = list(df_train.columns)
-  unwanted = {ti, ev}
-  varNames = [e for e in varNames if e not in unwanted]
-  leave = [(col, None) for col in varNames]
-
-  x_mapper = DataFrameMapper(leave)
-
-  x_train = x_mapper.fit_transform(df_train).astype('float32')
-  x_val = x_mapper.transform(df_val).astype('float32')
-  x_test = x_mapper.transform(df_test).astype('float32')
- 
-
-
-
-  ########
-  #prepare labels
-  #######
-  num_durations = 10
-  labtrans = DeepHitSingle.label_transform(num_durations)
-  get_target = lambda df: (df[ti].values, df[ev].values)
-  y_train = labtrans.fit_transform(*get_target(df_train))
-  y_val = labtrans.transform(*get_target(df_val))
-  train = (x_train, y_train)
-  val = (x_val, y_val)
-  durations_test, events_test = get_target(df_test)
-  ########
-  #prepare model
-  ########
-  in_features = x_train.shape[1]
-  num_nodes =[int(lay1),int(lay2)]
-  out_features = labtrans.out_features
-  batch_norm = True
-  dropout = drpt
-  
-  #net = tt.practical.MLPVanilla(in_features, num_nodes, out_features, batch_norm, dropout,activation=torch.nn.SELU)
-  net = torch.nn.Sequential(
-     torch.nn.BatchNorm1d(in_features),
-     torch.nn.Linear(in_features,int(lay1)),
[…]

### 2023-04-05 analyses/src/pmnnFunctions.R

@@ -1,482 +1,0 @@
-
-statCalcSim<-function(metricSims){
-  toRemove <- which(apply(metricSims, 1, function(x){any(is.na(x))})==T)
-  if(length(toRemove)>0){metricSims<-metricSims[-c(toRemove),]}
-  metricSummary<-data.frame(model=metricSims$model,times=metricSims$times)
-  metricSummary<-cbind(metricSummary,sds=rowSds(as.matrix(metricSims[,-c(1,2)])),
-                       means=rowMeans(as.matrix(metricSims[,-c(1,2)])))
-  ci<-apply(as.matrix(metricSims[,-c(1,2)]), 1, function(x){mean(x)+c(-1.96,1.96)*sd(x)/sqrt(length(x))})
-  rownames(ci)<-c('cilower95','ciupper95')
-  pi<-apply(as.matrix(metricSims[,-c(1,2)]), 1, function(x){quantile(x,c(.025, .975))})
-  rownames(pi)<-c('pilower95','piupper95')
-  metricSummary<-cbind(metricSummary,t(ci),t(pi))
-  return(metricSummary)
-}
-
-plotPerformance<-function(metricSummary,xlab="Follow-up-time",method='Metric',iterations='NA'){
-  mainTitle<-paste(method,' : 95% P.I. over ', iterations, ' iterations',sep='')
-  
-  a<-ggplot(data=metricSummary, aes(x=times,y=means,col=model))+
-    geom_line() +
-    geom_ribbon(aes(ymin=pilower95,ymax=piupper95,fill=model),linetype=0,alpha=0.3) +
-    ggtitle(mainTitle) + 
-    xlab(xlab) +
-    ylab(method)
-  if(method=="IPA"){
-    a<-a+coord_cartesian()#ylim(-1,1))#
-    return(a)
-  }
-  return(a+coord_cartesian())
-}
-
-
-pyProcess<-function(predictions,times=times,apr="constant"){
-  interpolated <- apply(predictions, 1, function(x) approx(as.numeric(colnames(predictions)), y=x, method = apr, xout=times)$y)
-  class(interpolated)<-c("tunnel",class(interpolated))
-  interpolated<-1-interpolated
-  interpolated[is.na(interpolated) & is.na(interpolated)>50] <- 0
-  interpolated[is.na(interpolated) & is.na(interpolated)<50] <- 1
-  return(t(interpolated))
-}
-
-
-
-
[…]

### 2023-04-05 analyses/testing.py

@@ -1,2 +1,0 @@
-def hi():
-  return 1

### 2023-04-05 StatPaper/APA/texDump.tex
updated text

--- 2022-11-15 StatPaper/APA-lato_New/APA/wileyNJD-APA.tex
@@
-\documentclass[APA,LATO1COL]{WileyNJD-v2}
-\usepackage{lmodern}
-\articletype{Article Type}%
+\hypertarget{inverse-probability-censoring-weights-adjusted-concordance-index}{%
+\subsubsection{Inverse probability censoring weights-adjusted
+concordance
+index \(C_{IPCW}\)}\label{inverse-probability-censoring-weights-adjusted-concordance-index}}
 
-\received{26 April 2016}
-\revised{6 June 2016}
-\accepted{6 June 2016}
+The \(C_{IPCW}\) is a non-proper, rank-based metric that does not depend
+on the censoring times in the test data \citep{uno2011}. The
+\(C_{IPCW}\) is given by
 
-\raggedbottom
+\begin{align} \label{eq:cidx}
+C_{IPCW}(t) &= \frac{\sum^{N}_{i=1}\sum^{N}_{j=1}\delta_{i}\left\{\widehat{G}(T_{i})\right\}^{-2} I(T_{i}<T_{j},T_{i}<t) I\left(\widehat{F}(t|X_{i})>\widehat{F}(t|X_{j})\right)}{\sum^{N}_{i=1}\sum^{N}_{j=1}\delta_{i}\left\{\widehat{G}(T_{i})\right\}^{-2} I(T_{i}<T_{j},T_{i}<t)}.
+\end{align} where the follow-up period of interest is (0,\(t\)),
+\(I(\cdot)\) is an indicator function, \(\widehat{F}(X_{i},t)\) is the
+risk function estimated for everyone in the study at time \(t\) and
+\(C_{IPCW}\) can compare the performance of different models, where a
+higher score is better. Note that the \(C_{IPCW}\) may produce
+misleading performance, as it ranks based on survival times, not event
+status \citep{cindexfails2019}. This metric is considered an unbiased
+population concordance measure because of the IPCW adjustment
+\citep{uno2011}.
 
-\begin{document}
+ While examining the IPA score, the optimal
+regression model and CBNN perform near identically, leading in performance followed by 
+DeepHit, the linear regression models and DeepSurv. We expected the Optimal model to perform best in both
+metrics. However, this was not the case for \(C_{IPCW}\) and may be due
+to an artifact of concordance-based metrics, where a misspecified model
+may perform better than a correctly specified one
+\citep{cindexfails2019}. We attribute the performance of CBNN to its
+flexibility in modeling time-varying interactions and baseline hazard,
+flexibility the other neural network models do not have. We keep this misspecification issue
+in mind while interpreting the \(C_{IPCW}\) results in the case studies.
 
-\title{This is the sample article title\protect\thanks{This is an example for title footnote.}}
+With the IPA score, CBNN outperforms
+The competing models, followed by the linear models, DeepSurv and DeepHit. We note that after
+50\% of survival time, all models aside from CBNN perform substantially worse than a null model.
+We note with \(C_{IPCW}\), the linear models and DeepSurv perform equivalently, followed by CBNN and
+DeepHit.
 
-\author[1]{Author One*}
+With the IPA score, CBNN outperforms
+The competing models, followed by the linear models and lastly DeepSurv and DeepHit for the first 60\% of follow-up time. 
+After 60\% of follow-up time, DeepHit has similar performance to CBNN. For the \(C_{IPCW}\) score, CBNN outperforms all models,
+followed by the linear models and finally DeepSurv. Note that for earlier ranges of survival time, DeepHit is the worst performing model. 
+After 25\% of follow-up time, DeepHit is in second place. In the FLC dataset, CBNN has a consistent top ranking for both IPA and \(C_{IPCW}\).
 
-\author[2,3]{Author Two}
+ With the IPA score, the linear models
+outperform the neural network ones, followed by CBNN and DeepHit, and
+finally DeepSurv. For the \(C_{IPCW}\) score, the linear models, CBNN, DeepHit
+perform near identically, side from DeepHit with an initial drop in performance.
+DeepSurv performed substantially worse than the other models.
 
-\author[3]{Author Three}
+To assess performance among these models, we use both IPA and
+\(C_{IPCW}\) metrics. Concordance-based measures are commonly used in
+survival analysis to compare models and we opt to keep them in our
+analyses. However, \(C_{IPCW}\) is a non-proper metric and may cause
+misspecified models to appear better than they should
+\citep{cindexfails2019}. \(C_{IPCW}\) is based on the original Concordance
+Index, a metric designed for proportional hazards (where the event risk is
+constant over time) \citep{uno2011}. As such, the rank order of individuals
+is would not change with follow-up time \citep{uno2011}. In studies with time-varying interactions (non-proportional hazards),
+it is unclear how \(C_{IPCW}\) would perform as rank order may change. 
+The model rankings between \(C_{IPCW}\) and IPA differ for all but FLC. We can further investigate this phenomenon
+in our simulation, as the true model is known. Considering figure \ref{fig:megaPlot} A, E and
+Table \ref{tab:megaTable} A in conjunction with FIGURE REF POPTIME, we know that the Optimal model should perform
+best and acts as both a benchmark and metric assessment. It is a regression
+model with a flexible baseline hazard and the exact interaction and time-varying interaction
+specified. The IPA score shows the Optimal model performing best, as expected. This gives us
+confidence that the IPA score can be used to compare all the models. The \(C_{IPCW}\) score did not.
+A more robust set of tests is required to determine if \(C_{IPCW}\) fails due to the time-varying interaction or
+the baseline hazard.
 
-\authormark{AUTHOR ONE \textsc{et al}}
+With this interpretation of \(C_{IPCW}\) in mind,
 
-\address[1]{\orgdiv{Org Division}, \orgname{Org Name}, \orgaddress{\state{State name}, \country{Country name}}}
+In the complex simulation, we note that \(C_{IPCW}\) is unable to properly rank the Optimal model.
+Further work is required to determine the cause of this, and we focus our assessment on the IPA score.
 
-\address[2]{\orgdiv{Org Division}, \orgname{Org Name}, \orgaddress{\state{State name}, \country{Country name}}}
+censoring weights-adjusted concordance index (\(C_{IPCW}\))
+\citep{uno2011}, which we define below.
 
-\address[3]{\orgdiv{Org Division}, \orgname{Org Name}, \orgaddress{\state{State name}, \country{Country name}}}
+We use the implementation of \(C_{IPCW}\) from the python
+package \textbf{sksurv} \citep{sksurv}.
 
-\corres{*Corresponding author name, This is sample corresponding address. \email{authorone@gmail.com}}
-
-\presentaddress{This is sample for present address text this is sample for present address text}
-
-\abstract[Summary]{This is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text.}
-
-\keywords{keyword1, keyword2, keyword3, keyword4}
-
-\jnlcitation{\cname{%
-\author{Williams K.},
-\author{B. Hoskins},
-\author{R. Lee},
-\author{G. Masato}, and
-\author{T. Woollings}} (\cyear{2016}),
-\ctitle{A regime analysis of Atlantic winter jet variability applied to evaluate HadGEM3-GC2}, \cjournal{Q.J.R. Meteorol. Soc.}, \cvol{2017;00:1--6}.}
-
-\maketitle
-
-\footnotetext{\textbf{Abbreviations:} ANA, anti-nuclear antibodies; APC, antigen-presenting cells; IRF, interferon regulatory factor}
-
-\section{Sample for first level head}\label{sec1}
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit~\citet{Rothermel1997} Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae,
-felis. Curabitur dictum gravida mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna. Donec
-vehicula augue eu neque. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
-Mauris ut leo. Cras viverra metus rhoncus sem. Nulla et lectus vestibulum urna fringilla ultrices. Phasellus eu tellus
-sit amet tortor gravida placerat. Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. Praesent eget sem vel
-leo ultrices bibendum. Aenean faucibus. Morbi dolor nulla, malesuada eu, pulvinar at, mollis ac, nulla. Curabitur
-auctor semper nulla. Donec varius orci eget risus. Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
-
-\begin{eqnarray}
-s(nT_{s}) &= &s(t)\times \sum\limits_{n=0}^{N-1} \delta (t-nT_{s}) \xleftrightarrow{\mathrm{DFT}}  S \left(\frac{m}{NT_{s}}\right) \nonumber\\
-&= &\frac{1}{N} \sum\limits_{n=0}^{N-1} \sum\limits_{k=-N/2}^{N/2-1} s_{k} e^{\mathrm{j}2\pi k\Delta fnT_{s}} e^{-j\frac{2\pi}{N}mn}
-\end{eqnarray}
-
-\section{Sample for another first level head}\label{sec2}
-
-Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus viverra
-fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa ac
-quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo~\cite{Rothermel1998} Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
-nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
-bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
-pellentesque felis eu massa.
-
-Example for bibliography citations cite~\citep{Elbaum2002}, cites~\cite{Allen2011,Yoo2007}
-
-Quisque ullamcorper placerat ipsum. Cras nibh~\cite{Yoo2007,Schulz2012} Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
-amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
-Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
-quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
-vitae risus porta vehicula.
-
-Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
-venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
-quam, in hendrerit risus eros eget felis.
-Maecenas eget erat in sapien mattis porttitor.
-Vestibulum porttitor.
-Nulla facilisi.
-Sed a turpis eu lacus commodo facilisis.
-Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
-vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
-Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.
-
-\begin{figure}[t]
-\centerline{\includegraphics[width=342pt,height=9pc,draft]{empty}}
-\caption{This is the sample figure caption.\label{fig1}}
-\end{figure}
-
-Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
-ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
-eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
-convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
-pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.
-
-Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non
-odio. Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce
-sed justo eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor.
-Donec pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin
-et quam. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien
-turpis, fermentum vel, eleifend faucibus, vehicula eu, lacus.
-
-\begin{figure*}
-\centerline{\includegraphics[width=342pt,height=9pc,draft]{empty}}
-\caption{This is the sample figure caption.\label{fig2}}
-\end{figure*}
-
-\subsection{Example for second level head}
-
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit, dictum
-in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam vestibulum
-fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi. Pellentesque
-eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut mi mi, lacinia
-sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui. Mauris nibh leo,
-facilisis non, adipiscing quis, ultrices a, dui.
-
-Morbi luctus, wisi viverra faucibus pretium, nibh est placerat odio, nec commodo wisi enim eget quam. Quisque
-libero justo, consectetuer a, feugiat vitae, porttitor eu, libero. Suspendisse sed mauris vitae elit sollicitudin malesuada.
-
-Maecenas ultricies eros sit amet ante. Ut venenatis velit. Maecenas sed mi eget dui varius euismod. Phasellus aliquet
-volutpat odio. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Pellentesque sit
-amet pede ac sem eleifend consectetuer. Nullam elementum, urna vel imperdiet sodales, elit ipsum pharetra ligula,
-ac pretium ante justo a nulla. Curabitur tristique arcu eu metus. Vestibulum lectus. Proin mauris. Proin eu nunc eu
-urna hendrerit faucibus. Aliquam auctor, pede consequat laoreet varius, eros tellus scelerisque quam, pellentesque
-hendrerit ipsum dolor sed augue. Nulla nec lacus.
-
-\begin{quote}
-This is an example~\citep{Elbaum2002,Blanchard2015} for quote text.
-This is an example for quote text.
-This is an example for quote text.
-This is an example for quote text~\cite{Blanchard2015}.
-This is an example for quote text.
-This is an example for quote text.
-This is an example for quote text.
-This is an example for quote text.
-This is an example for quote text.
-This is an example for quote text~\cite{Blanchard2015}.
-This is an example for quote text.
-This is an example for quote text.
-This is an example for quote text.
-\end{quote}
-
-\section{Sample for next first level head}\label{sec3}
-
-\subsection{Example for another second level head}
-
-Suspendisse vitae elit. Aliquam arcu neque, ornare in, ullamcorper quis, commodo eu, libero. Fusce sagittis erat at
-erat tristique mollis. Maecenas sapien libero, molestie et, lobortis in, sodales eget, dui. Morbi ultrices rutrum lorem.
-Nam elementum ullamcorper leo. Morbi dui. Aliquam sagittis. Nunc placerat. Pellentesque tristique sodales est.
-Maecenas imperdiet lacinia velit. Cras non urna. Morbi eros pede, suscipit ac, varius vel, egestas non, eros. Praesent
-malesuada, diam id pretium elementum, eros sem dictum tortor, vel consectetuer odio sem sed wisi.
-
-Sed feugiat. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Ut pellentesque
-augue sed urna. Vestibulum diam eros, fringilla et, consectetuer eu, nonummy id, sapien. Nullam at lectus. In sagittis
-ultrices mauris. Curabitur malesuada erat sit amet massa. Fusce blandit. Aliquam erat volutpat. Aliquam euismod.
-Aenean vel lectus. Nunc imperdiet justo nec dolor.
-
-\subsection{Second level head text}
-
-Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
-eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
-facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
-a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
-convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.
-
-\subsubsection{Third level head text}
-
-Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra tortor. Sed fringilla
-justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue est scelerisque
-sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas urna mi, suscipit
-in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.
-
-Etiam ac leo a risus tristique nonummy. Donec dignissim tincidunt nulla. Vestibulum rhoncus molestie odio. Sed
-lobortis, justo et pretium lobortis, mauris turpis condimentum augue, nec ultricies nibh arcu pretium enim. Nunc
-purus neque, placerat id, imperdiet sed, pellentesque nec, nisl. Vestibulum imperdiet neque non sem accumsan laoreet.
-In hac habitasse platea dictumst. Etiam condimentum facilisis libero. Suspendisse in elit quis nisl aliquam dapibus.
-Pellentesque auctor sapien. Sed egestas sapien nec lectus. Pellentesque vel dui vel neque bibendum viverra. Aliquam
-porttitor nisl nec pede. Proin mattis libero vel turpis. Donec rutrum mauris et libero. Proin euismod porta felis.
-Nam lobortis, metus quis elementum commodo, nunc lectus elementum mauris, eget vulputate ligula tellus eu neque.
-Vivamus eu dolor.
-
-Nulla in ipsum. Praesent eros nulla, congue vitae, euismod ut, commodo a, wisi. Pellentesque habitant morbi
-tristique senectus et netus et malesuada fames ac turpis egestas. Aenean nonummy magna non leo. Sed felis erat,
-ullamcorper in, dictum non, ultricies ut, lectus. Proin vel arcu a odio lobortis euismod. Vestibulum ante ipsum primis
-in faucibus orci luctus et ultrices posuere cubilia Curae; Proin ut est. Aliquam odio. Pellentesque massa turpis, cursus
-eu, euismod nec, tempor congue, nulla. Duis viverra gravida mauris. Cras tincidunt. Curabitur eros ligula, varius ut,
-pulvinar in, cursus faucibus, augue.
-
-\begin{boxtext}
-\section*{Example of Boxtext}%
-This is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext.
-\end{boxtext}
-
-\paragraph{Fourth level head text}
-
-Sed feugiat. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Ut pellentesque
-augue sed urna. Vestibulum diam eros, fringilla et, consectetuer eu, nonummy id, sapien. Nullam at lectus. In sagittis
-ultrices mauris. Curabitur malesuada erat sit amet massa. Fusce blandit. Aliquam erat volutpat. Aliquam euismod.
-Aenean vel lectus. Nunc imperdiet justo nec dolor.
-
-Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
-eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
-facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
-a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
-convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.
-
-\subparagraph{Fifth level head text}
-
-Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra
-tortor. Sed fringilla justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue
-est scelerisque sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas
-urna mi, suscipit in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.
-
-Etiam ac leo a risus tristique nonummy. Donec dignissim tincidunt nulla. Vestibulum rhoncus molestie odio. Sed
-lobortis, justo et pretium lobortis, mauris turpis condimentum augue, nec ultricies nibh arcu pretium enim. Nunc
-purus neque, placerat id, imperdiet sed, pellentesque nec, nisl. Vestibulum imperdiet neque non sem accumsan laoreet.
-In hac habitasse platea dictumst. Etiam condimentum facilisis libero. Suspendisse in elit quis nisl aliquam dapibus.
-Pellentesque auctor sapien. Sed egestas sapien nec lectus. Pellentesque vel dui vel neque bibendum viverra. Aliquam
-porttitor nisl nec pede. Proin mattis libero vel turpis. Donec rutrum mauris et libero. Proin euismod porta felis.
-Nam lobortis, metus quis elementum commodo, nunc lectus elementum mauris, eget vulputate ligula tellus eu neque.
-Vivamus eu dolor.
-
-in faucibus orci luctus et ultrices posuere cubilia Curae; Proin ut est. Aliquam odio. Pellentesque massa turpis, cursus
-eu, euismod nec, tempor congue, nulla. Duis viverra gravida mauris. Cras tincidunt. Curabitur eros ligula, varius ut,
-pulvinar in, cursus faucibus, augue.
-
-Curabitur tellus magna, porttitor a, commodo a, commodo in, tortor. Donec interdum. Praesent scelerisque. Mae-
-cenas posuere sodales odio. Vivamus metus lacus, varius quis, imperdiet quis, rhoncus a, turpis. Etiam ligula arcu,
-elementum a, venenatis quis, sollicitudin sed, metus. Donec nunc pede, tincidunt in, venenatis vitae, faucibus vel,
-nibh. Pellentesque wisi. Nullam malesuada. Morbi ut tellus ut pede tincidunt porta. Lorem ipsum dolor sit amet,
-consectetuer adipiscing elit. Etiam congue neque id dolor.
-
-Donec et nisl at wisi luctus bibendum. Nam interdum tellus ac libero. Sed sem justo, laoreet vitae, fringilla at,
-adipiscing ut, nibh. Maecenas non sem quis tortor eleifend fermentum. Etiam id tortor ac mauris porta vulputate.
-Integer porta neque vitae massa. Maecenas tempus libero a libero posuere dictum. Vestibulum ante ipsum primis in
-faucibus orci luctus et ultrices posuere cubilia Curae; Aenean quis mauris sed elit commodo placerat. Class aptent
-taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Vivamus rhoncus tincidunt libero.
-Etiam elementum pretium justo. Vivamus est. Morbi a tellus eget pede tristique commodo. Nulla nisl. Vestibulum
-sed nisl eu sapien cursus rutrum.
-
-Nulla non mauris vitae wisi posuere convallis. Sed eu nulla nec eros scelerisque pharetra. Nullam varius. Etiam
-dignissim elementum metus. Vestibulum faucibus, metus sit amet mattis rhoncus, sapien dui laoreet odio, nec ultricies
-nibh augue a enim. Fusce in ligula. Quisque at magna et nulla commodo consequat. Proin accumsan imperdiet sem.
-Nunc porta. Donec feugiat mi at justo. Phasellus facilisis ipsum quis ante. In ac elit eget ipsum pharetra faucibus.
-Maecenas viverra nulla in massa.
-
-Nulla in ipsum. Praesent eros nulla, congue vitae, euismod ut, commodo a, wisi. Pellentesque habitant morbi
-tristique senectus et netus et malesuada fames ac turpis egestas. Aenean nonummy magna non leo. Sed felis erat,
-ullamcorper in, dictum non, ultricies ut, lectus. Proin vel arcu a odio lobortis euismod. Vestibulum ante ipsum primis
-
-\begin{center}
-\begin{table*}[t]%
-\caption{This is sample table caption.\label{tab1}}
-\centering
-\begin{tabular*}{500pt}{@{\extracolsep\fill}lccD{.}{.}{3}c@{\extracolsep\fill}}
-\toprule
-&\multicolumn{2}{@{}c@{}}{\textbf{Spanned heading\tnote{1}}} & \multicolumn{2}{@{}c@{}}{\textbf{Spanned heading\tnote{2}}} \\\cmidrule{2-3}\cmidrule{4-5}
-\textbf{col1 head} & \textbf{col2 head}  & \textbf{col3 head}  & \multicolumn{1}{@{}l@{}}{\textbf{col4 head}}  & \textbf{col5 head}   \\
-\midrule
-col1 text & col2 text  & col3 text  & 12.34  & col5 text\tnote{1}   \\
-col1 text & col2 text  & col3 text  & 1.62  & col5 text\tnote{2}   \\
-col1 text & col2 text  & col3 text  & 51.809  & col5 text   \\
-\bottomrule
-\end{tabular*}
-\begin{tablenotes}%%[341pt]
-\item Source: Example for table source text.
-\item[1] Example for a first table footnote.
-\item[2] Example for a second table footnote.
-\end{tablenotes}
-\end{table*}
-\end{center}
-
-Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
-venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
-quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
-facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
-vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
-Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.
-
-\begin{center}
-\begin{table}[t]%
-\centering
-\caption{This is sample table caption.\label{tab2}}%
-\begin{tabular*}{500pt}{@{\extracolsep\fill}lcccc@{\extracolsep\fill}}
-\toprule
-\textbf{col1 head} & \textbf{col2 head}  & \textbf{col3 head}  & \textbf{col4 head}  & \textbf{col5 head} \\
-\midrule
-col1 text & col2 text  & col3 text  & col4 text  & col5 text\tnote{$\dagger$}   \\
-col1 text & col2 text  & col3 text  & col4 text  & col5 text   \\
-col1 text & col2 text  & col3 text  & col4 text  & col5 text\tnote{$\ddagger$}   \\
-\bottomrule
-\end{tabular*}
-\begin{tablenotes}
-\item Source: Example for table source text.
-\item[$\dagger$] Example for a first table footnote.
-\item[$\ddagger$] Example for a second table footnote.
-\end{tablenotes}
-\end{table}
-\end{center}
-
-Below is the example~\cite{Rothermel1998,Yoo2007,Schulz2012} for bulleted list.
-Below is the example for bulleted list.
-Below is the example for bulleted list.
-Below is the example for bulleted list.
-Below is the example for bulleted list.
-Below is the example for bulleted list\footnote{This is an example for footnote.}:
-\begin{itemize}
-\item bulleted list entry sample bulleted list entry~\cite{Allen2011} sample list entry text.
-\item bulleted list entry sample bulleted list entry. bulleted list entry sample bulleted list entry. bulleted list entry sample bulleted list entry.
-\item bulleted list entry sample bulleted list entry~\cite{Ballen2011} bulleted list entry sample bulleted list entry~\citet{Allen2011} sample list entry text.
-bulleted list entry sample bulleted list entry.
-\item sample list entry text. sample list entry text.
-\end{itemize}
-
-Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
-ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
-eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
-convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
-pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.
-
-Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non
-odio. Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce
-sed justo eu urna porta tincidunt.
-Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat.
-Morbi quis dolor.
-Donec pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus.
-Proin
-et quam. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien
-turpis, fermentum vel, eleifend faucibus, vehicula eu, lacus.
-
-Below is the sample for description list.
-Below is the example for description list.
-Below is the example for description list.
-Below is the example for description list.
-Below is the example for description list.
-Below is the example for description list:\\[12pt]
-
-\noindent\textbf{Description sample:}
-
-\begin{description}
-\item[first entry] description text. description text. description text. description text. description text. description text. description text.
-\item[second long entry] description text. description text. description text. description text. description text. description text. description text.
-\item[third entry] description text. description text. description text. description text. description text.
-\item[fourth entry] description text. description text.
-\end{description}
-
-\noindent\textbf{Numbered list items sample:}
-
-\begin{enumerate}[1.]
-\item First level numbered list entry. sample numbered list entry.
-
-\item First numbered list entry. sample numbered list entry. Numbered list entry. sample numbered list entry. Numbered list entry. sample numbered list entry.
-
-\begin{enumerate}[a.]
-\item Second level alpabetical list entry. Second level alpabetical list entry. Second level alpabetical list entry~\citet{Allen2011} Second level alpabetical list entry.
-
-\item Second level alpabetical list entry. Second level alpabetical list entry~\citet{Schulz2012,Allen2011,Ballen2011}.
-
-\begin{enumerate}[ii.]
-\item Third level lowercase roman numeral list entry. Third level lowercase roman numeral list entry. Third level lowercase roman numeral list entry.
-
-\item Third level lowercase roman numeral list entry. Third level lowercase roman numeral list entry~\cite{Yoo2007}.
-\end{enumerate}
-
-\item Second level alpabetical list entry. Second level alpabetical list entry~\cite{Elbaum2002}.
-\end{enumerate}
-
-\item First level numbered list entry. sample numbered list entry. Numbered list entry. sample numbered list entry. Numbered list entry.
-
-\item Another first level numbered list entry. sample numbered list entry. Numbered list entry. sample numbered list entry. Numbered list entry.
-\end{enumerate}
-
-\noindent\textbf{un-numbered list items sample:}
-
-\begin{enumerate}[]
-\item Sample unnumberd list text..
-\item Sample unnumberd list text.
-\item sample unnumberd list text.
-\item Sample unnumberd list text.
-\end{enumerate}
-
-\section{Examples for enunciations}\label{sec4}
-
-\begin{theorem}[Theorem subhead]\label{thm1}
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-Example theorem text.
-\end{theorem}
-
-Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
-amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
-Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
-quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
-vitae risus porta vehicula.
-
-Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
-venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
-quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
-facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
-vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
-Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.
-
-\begin{proposition}
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-Example proposition text.
-\end{proposition}
-
-Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
-viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
-ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
-nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
-bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
-pellentesque felis eu massa.
-
-Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
-amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
-Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
-quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
-vitae risus porta vehicula.
-
-\begin{definition}[Definition sub head]
-Example definition text.
-Example definition text.
-Example definition text.
-Example definition text.
-Example definition text.
-Example definition text.
-Example definition text.
-Example definition text.
-Example definition text.
-Example definition text.
-Example definition text.
-\end{definition}
-
-Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non
-odio. Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce
-sed justo eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor.
-Donec pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin
-et quam. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien
-turpis, fermentum vel, eleifend faucibus, vehicula eu, lacus.
-
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
-dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
-vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
-Pellentesque eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut
-mi mi, lacinia sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui.
-Mauris nibh leo, facilisis non, adipiscing quis, ultrices a, dui.
-
-\begin{proof}
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-\end{proof}
-
-Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
-pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
-vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
-a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
-urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.
-
-Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
-viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
-ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
-nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
-bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
-pellentesque felis eu massa.
-
-\begin{proof}[Proof of Theorem~\ref{thm1}]
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-\end{proof}
-
-Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
-eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
-facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
-a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
-convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.
-Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra tortor. Sed fringilla
-justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue est scelerisque
-sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas urna mi, suscipit
-in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.
-
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
-dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
-vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
-Pellentesque eget lectus.
-Proin eu metus.
-Sed porttitor.
-In hac habitasse platea dictumst.
-Suspendisse eu lectus.
-Ut Curabitur tellus magna, porttitor a, commodo a, commodo in, tortor.
-Donec interdum.
-Praesent scelerisque.
-Mae-
-cenas posuere sodales odio. Vivamus metus lacus, varius quis, imperdiet quis, rhoncus a, turpis. Etiam ligula arcu,
-elementum a, venenatis quis, sollicitudin sed, metus. Donec nunc pede, tincidunt in, venenatis vitae, faucibus vel,
-
-\begin{sidewaystable}%[h]
-\caption{Sideways table caption. For decimal alignment refer column 4 to 9 in tabular* preamble.\label{tab3}}%
-\begin{tabular*}{\textheight}{@{\extracolsep\fill}lccD{.}{.}{4}D{.}{.}{4}D{.}{.}{4}D{.}{.}{4}D{.}{.}{4}D{.}{.}{4}@{\extracolsep\fill}}%
-\toprule
-  & \textbf{col2 head} & \textbf{col3 head} & \multicolumn{1}{c}{\textbf{10}} &\multicolumn{1}{c}{\textbf{20}} &\multicolumn{1}{c}{\textbf{30}} &\multicolumn{1}{c}{\textbf{10}} &\multicolumn{1}{c}{\textbf{20}} &\multicolumn{1}{c}{\textbf{30}} \\
-\midrule
-  &col2 text &col3 text &0.7568&1.0530&1.2642&0.9919&1.3541&1.6108 \\
-  & &col2 text &12.5701 &19.6603&25.6809&18.0689&28.4865&37.3011 \\
-3 &col2 text  & col3 text &0.7426&1.0393&1.2507&0.9095&1.2524&1.4958 \\
-  & &col3 text &12.8008&19.9620&26.0324&16.6347&26.0843&34.0765 \\
-  & col2 text & col3 text &0.7285&1.0257&1.2374&0.8195&1.1407&1.3691\tnote{*} \\
-  & & col3 text &13.0360&20.2690&26.3895&15.0812&23.4932&30.6060\tnote{\dagger} \\
-\bottomrule
-\end{tabular*}
-\begin{tablenotes}%%[\textheight]
-\item[*] First sideways table footnote. Sideways table footnote. Sideways table footnote. Sideways table footnote.
-\item[$\dagger$] Second sideways table footnote. Sideways table footnote. Sideways table footnote. Sideways table footnote.
-\end{tablenotes}
-\end{sidewaystable}
-
-\begin{sidewaysfigure}
-\centerline{\includegraphics[width=542pt,height=9pc,draft]{empty}}
-\caption{Sideways figure caption. Sideways figure caption. Sideways figure caption. Sideways figure caption. Sideways figure caption. Sideways figure caption.\label{fig3}}
-\end{sidewaysfigure}
-
-nibh. Pellentesque wisi. Nullam malesuada. Morbi ut tellus ut pede tincidunt porta. Lorem ipsum dolor sit amet,
-consectetuer adipiscing elit. Etiam congue neque id dolor.
-
-\begin{algorithm}
-\caption{Pseudocode for our algorithm}\label{alg1}
-\begin{algorithmic}
-  \For each frame
-  \For water particles $f_{i}$
-  \State compute fluid flow~\cite{Rothermel1997}
-  \State compute fluid--solid interaction~\cite{Allen2011}
-  \State apply adhesion and surface tension~\cite{Ballen2011}
-  \EndFor
-   \For solid particles $s_{i}$
-   \For neighboring water particles $f_{j}$
-   \State compute virtual water film \\(see Section~\ref{sec3})
-   \EndFor
-   \EndFor
-   \For solid particles $s_{i}$
-   \For neighboring water particles $f_{j}$
-   \State compute growth direction vector \\(see Section~\ref{sec2})
-   \EndFor
-   \EndFor
-   \For solid particles $s_{i}$
-   \For neighboring water particles $f_{j}$
-   \State compute $F_{\theta}$ (see Section~\ref{sec1})
-   \State compute $CE(s_{i},f_{j})$ \\(see Section~\ref{sec3})
-   \If $CE(b_{i}, f_{j})$ $>$ glaze threshold
-   \State $j$th water particle's phase $\Leftarrow$ ICE
-   \EndIf
-   \If $CE(c_{i}, f_{j})$ $>$ icicle threshold
-   \State $j$th water particle's phase $\Leftarrow$ ICE
-   \EndIf
-   \EndFor
-   \EndFor
-  \EndFor
-\end{algorithmic}
-\end{algorithm}
-
-Donec et nisl at wisi luctus bibendum. Nam interdum tellus ac libero. Sed sem justo, laoreet vitae, fringilla at,
-adipiscing ut, nibh. Maecenas non sem quis tortor eleifend fermentum. Etiam id tortor ac mauris porta vulputate.
-Integer porta neque vitae massa~\cite{Rothermel1997,Elbaum2002}. Maecenas tempus libero a libero posuere dictum. Vestibulum ante ipsum primis in
-faucibus orci luctus et ultrices posuere cubilia Curae; Aenean quis mauris sed elit commodo placerat. Class aptent
-taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Vivamus rhoncus tincidunt libero.
-Etiam elementum pretium justo. Vivamus est. Morbi a tellus eget pede tristique commodo~\cite{Elbaum2002} Nulla nisl. Vestibulum
-sed nisl eu sapien cursus rutrum.
-
-Pellentesque wisi. Nullam malesuada. Morbi ut tellus ut pede tincidunt porta. Lorem ipsum dolor sit amet,
-consectetuer adipiscing elit. Etiam congue neque id dolor.
-
-Donec et nisl at wisi luctus bibendum. Nam interdum tellus ac libero. Sed sem justo, laoreet vitae, fringilla at,
-adipiscing ut, nibh. Maecenas non sem quis tortor eleifend fermentum. Etiam id tortor ac mauris porta vulputate.
-Integer porta neque vitae massa. Maecenas tempus libero a libero posuere dictum. Vestibulum ante ipsum primis in
-faucibus orci luctus et ultrices posuere cubilia Curae; Aenean quis mauris sed elit commodo placerat. Class aptent
-taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Vivamus rhoncus tincidunt libero.
-Etiam elementum pretium justo. Vivamus est. Morbi a tellus eget pede tristique commodo. Nulla nisl. Vestibulum
-sed nisl eu sapien cursus rutrum.
-
-\begin{equation}\label{eq23}
- \|\tilde{X}(k)\|^2
- =\frac{\left\|\sum\limits_{i=1}^{p}\tilde{Y}_i(k)+\sum\limits_{j=1}^{q}\tilde{Z}_j(k) \right\|^2}{(p+q)^2}
- \leq\frac{\sum\limits_{i=1}^{p}\left\|\tilde{Y}_i(k)\right\|^2+\sum\limits_{j=1}^{q}\left\|\tilde{Z}_j(k)\right\|^2 }{p+q}.
-\end{equation}
-
-Sed feugiat. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Ut pellentesque
-augue sed urna. Vestibulum diam eros, fringilla et, consectetuer eu, nonummy id, sapien. Nullam at lectus. In sagittis
-ultrices mauris. Curabitur malesuada erat sit amet massa. Fusce blandit. Aliquam erat volutpat. Aliquam euismod.
-Aenean vel lectus. Nunc imperdiet justo nec dolor.
-
-Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
-eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
-facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
-a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
-convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.
-
-\begin{equation}\label{eq24}
- \|\tilde{X}(k)\|^2
- =\frac{\left\|\sum\limits_{i=1}^{p}\tilde{Y}_i(k)+\sum\limits_{j=1}^{q}\tilde{Z}_j(k) \right\|^2}{(p+q)^2}
- \leq\frac{\sum\limits_{i=1}^{p}\left\|\tilde{Y}_i(k)\right\|^2+\sum\limits_{j=1}^{q}\left\|\tilde{Z}_j(k)\right\|^2 }{p+q}.
-\end{equation}
-
-Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra
-tortor. Sed fringilla justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue
-est scelerisque sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas
-urna mi, suscipit in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.
-
-Etiam ac leo a risus tristique nonummy. Donec dignissim tincidunt nulla. Vestibulum rhoncus molestie odio. Sed
-lobortis, justo et pretium lobortis, mauris turpis condimentum augue, nec ultricies nibh arcu pretium enim. Nunc
-purus neque, placerat id, imperdiet sed, pellentesque nec, nisl. Vestibulum imperdiet neque non sem accumsan laoreet.
-In hac habitasse platea dictumst. Etiam condimentum facilisis libero. Suspendisse in elit quis nisl aliquam dapibus.
-Pellentesque auctor sapien. Sed egestas sapien nec lectus. Pellentesque vel dui vel neque bibendum viverra. Aliquam
-porttitor nisl nec pede. Proin mattis libero vel turpis. Donec rutrum mauris et libero. Proin euismod porta felis.
-Nam lobortis, metus quis elementum commodo, nunc lectus elementum mauris, eget vulputate ligula tellus eu neque.
-Vivamus eu dolor.
-
-\section{Conclusions}\label{sec5}
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae,
-felis. Curabitur dictum gravida mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna. Donec
-vehicula augue eu neque. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
-Mauris ut leo. Cras viverra metus rhoncus sem. Nulla et lectus vestibulum urna fringilla ultrices. Phasellus eu tellus
-sit amet tortor gravida placerat. Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. Praesent eget sem vel
-leo ultrices bibendum. Aenean faucibus. Morbi dolor nulla, malesuada eu, pulvinar at, mollis ac, nulla. Curabitur
-auctor semper nulla. Donec varius orci eget risus. Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
-
-Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
-pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
-vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
-a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
-urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.
-
-%\backmatter
-
-\section*{Acknowledgments}
-This is acknowledgment text~\cite{Elbaum2002}.
-Provide text here.
-This is acknowledgment text.
-Provide text here.
-This is acknowledgment text.
-Provide text here.
-This is acknowledgment text.
-Provide text here.
-This is acknowledgment text.
-Provide text here.
-This is acknowledgment text.
-Provide text here.
-This is acknowledgment text.
-Provide text here.
-This is acknowledgment text.
-Provide text here.
-This is acknowledgment text.
-Provide text here.
-
-\subsection*{Author contributions}
-
-This is an author contribution text. This is an author contribution text. This is an author contribution text. This is an author contribution text. This is an author contribution text.
-
-\subsection*{Financial disclosure}
-
-None reported.
-
-\subsection*{Conflict of interest}
-
-The authors declare no potential conflict of interests.
-
-\section*{Supporting information}
-
-The following supporting information is available as part of the online article:
-
-\noindent
-\textbf{Figure S1.}
-{500{\uns}hPa geopotential anomalies for GC2C calculated against the ERA Interim reanalysis. The period is 1989--2008.}
-
-\noindent
-\textbf{Figure S2.}
-{The SST anomalies for GC2C calculated against the observations (OIsst).}
-
-\appendix
-
-\section{Section title of first appendix\label{app1}}
-
-Use \verb+\begin{verbatim}...\end{verbatim}+ for program codes without math.
-Use \verb+\begin{alltt}...\end{alltt}+ for program codes with math.
-Based on the text provided inside the optional argument of \verb+\begin{code}[Psecode|Listing|Box|Code|+\hfill\break \verb+Specification|Procedure|Sourcecode|Program]...+ \verb+\end{code}+ tag corresponding boxed like floats are generated.
-Also note that \verb+\begin{code}[Code|Listing]...+ \verb+\end{code}+ tag with either Code or Listing text as optional argument text are set with computer modern typewriter font.
-All other code environments are set with normal text font.
-Refer below example:
-
-\begin{lstlisting}[caption={Descriptive Caption Text},label=DescriptiveLabel]
-for i:=maxint to 0 do
-begin
-{ do nothing }
-end;
-Write('Case insensitive ');
-WritE('Pascal keywords.');
-\end{lstlisting}
-
-\subsection{Subsection title of first appendix\label{app1.1a}}
-
-Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
-pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
-vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
-a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
-urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.
-
-Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
-viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
-ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
-nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
-bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
-pellentesque felis eu massa.
-
-\subsubsection{Subsection title of first appendix\label{app1.1.1a}}
-
-\noindent\textbf{Unnumbered figure}
-
-\begin{center}
-\includegraphics[width=7pc,height=8pc,draft]{empty}
-\end{center}
-
-Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
-venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
-quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
-facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
-vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
-
-Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.
-Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
-ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
-eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
-convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
-pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.
-
-\section{Section title of second appendix\label{app2}}%
-
-Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
-venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
-quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
-facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
-vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
-
-Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.
-Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
-ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
-eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
-convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
-pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.
-
-%== Figure 4 ==
-%% Example for figure inside appendix
-\begin{figure}[t]
-\centerline{\includegraphics[height=10pc,width=78mm,draft]{empty}}
-\caption{This is an example for appendix figure.\label{fig5}}
-\end{figure}
-
-\subsection{Subsection title of second appendix\label{app2.1a}}
-
-Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non odio.
-Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce sed justo
-eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor. Donec
-pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin et quam.
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien turpis,
-fermentum vel, eleifend faucibus, vehicula eu, lacus.
-
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
-dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
-vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
-Pellentesque eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut
-mi mi, lacinia sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui.
-Mauris nibh leo, facilisis non, adipiscing quis, ultrices a, dui.
-
-\subsubsection{Subsection title of second appendix\label{app2.1.1a}}
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae,
-felis. Curabitur dictum gravida mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna. Donec
-vehicula augue eu neque. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
-Mauris ut leo. Cras viverra metus rhoncus sem. Nulla et lectus vestibulum urna fringilla ultrices. Phasellus eu tellus
-sit amet tortor gravida placerat. Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. Praesent eget sem vel
-leo ultrices bibendum. Aenean faucibus. Morbi dolor nulla, malesuada eu, pulvinar at, mollis ac, nulla. Curabitur
-auctor semper nulla. Donec varius orci eget risus. Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
-
-Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
-pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
-vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
-a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
-urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.
-
-\begin{center}
-\begin{table}[b]%
-\centering
-\caption{This is an example of Appendix table showing food requirements of army, navy and airforce.\label{tab4}}%
-\begin{tabular*}{300pt}{@{\extracolsep\fill}lcc@{\extracolsep\fill}}%
-\toprule
-\textbf{col1 head} & \textbf{col2 head} & \textbf{col3 head} \\
-\midrule
-col1 text & col2 text & col3 text \\
-col1 text & col2 text & col3 text \\
-col1 text & col2 text & col3 text\\
-\bottomrule
-\end{tabular*}
-\end{table}
-\end{center}
-
-Example for an equation inside appendix
-\begin{equation}
-\mathcal{L}\quad \mathbf{\mathcal{L}} = i \bar{\psi} \gamma^\mu D_\mu \psi - \frac{1}{4} F_{\mu\nu}^a F^{a\mu\nu} - m \bar{\psi} \psi\label{eq25}
-\end{equation}
-
-\section{Example of another appendix section\label{app3}}%
-
-This is sample for paragraph text this is sample for paragraph text  this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text
-
-Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
-pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
-vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
-a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
-urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.
-
-Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
-viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
-ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
-nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
-bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
-pellentesque felis eu massa.
-
-\begin{equation}
-\mathcal{L} = i \bar{\psi} \gamma^\mu D_\mu \psi
-    - \frac{1}{4} F_{\mu\nu}^a F^{a\mu\nu} - m \bar{\psi} \psi
-\label{eq26}
-\end{equation}
-
-Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
-viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
-ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
-nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
-bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
-pellentesque felis eu massa.
-
-Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
-amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
-Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
-quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
-vitae risus porta vehicula.
-
-\begin{center}
-\begin{tabular*}{300pt}{@{\extracolsep\fill}lcc@{\extracolsep\fill}}%
-\toprule
-\textbf{col1 head} & \textbf{col2 head} & \textbf{col3 head} \\
-\midrule
-col1 text & col2 text & col3 text \\
-col1 text & col2 text & col3 text \\
-col1 text & col2 text & col3 text \\
-\bottomrule
-\end{tabular*}
-\end{center}
-
-Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
-amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
-Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
-quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
-vitae risus porta vehicula.
-
-Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
-venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
-quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
-facilisi.
-Sed a turpis eu lacus commodo facilisis.
-Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, evehicula libero dui cursus dui.
-Mauris tempor ligula sed lacus.
-Duis cursus enim ut augue.
-Cras ac magna.
-Cras nulla.
-Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.
-
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
-dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
-vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
-Pellentesque eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut
-mi mi, lacinia sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui.
-Mauris nibh leo, facilisis non, adipiscing quis, ultrices a, dui.
-
-\nocite{*}% Show all bib entries - both cited and uncited; comment this line to view only cited bib entries;
-\bibliography{wileyNJD-APA}%
-
-\section*{Author Biography}
-
-\begin{biography}{\includegraphics[width=60pt,height=70pt,draft]{empty}}{\textbf{Author Name.} This is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text.}
-\end{biography}
-
-\end{document}
+A potential artifact of IPA is that the score is unstable at earlier and
+later follow-up times. This is because of a near equivalent BS
+among each model and the null model. At small values, a difference of
+\(0.1\) creates a more significant fold change than at larger values. The IPA may
+be unstable for follow-up times where the BS is low.

### 2023-04-05 StatPaper/APA/wileyNJD-APA.tex
updated text

--- 2023-04-05 StatPaper/APA-lato_New/APA/wileyNJD-APA.tex
@@
 \usepackage{lmodern}
+
 \articletype{Article Type}%
 
-\received{26 April 2016}
-\revised{6 June 2016}
-\accepted{6 June 2016}
+\received{26 April 2023}
+\revised{6 June 2023}
+\accepted{6 June 2023}
 
@@
 
-\presentaddress{This is sample for present address text this is sample for present address text}
+\presentaddress{McGill genome center: 740 Dr Penfield Ave, Montreal, Quebec H3A 0G1}
 
-\abstract[Summary]{Neural network-based survival methods can model data-driven
-covariate interactions. While these methods have led to better
-predictive performance than regression-based approaches, they cannot
-model both time-varying interactions and complex baseline hazards. To
-address this, we propose Case-Base Neural Networks (CBNN) as a new
-approach that combines the case-base sampling framework with flexible
-architectures. Our method naturally accounts for censoring and does not
-require method specific hyperparameters. Using a novel sampling scheme
-and data augmentation, we incorporate time directly into a feed-forward
-neural network. CBNN predicts the probability of an event occurring at a
-given moment and estimates the hazard function. We compare the
-performance of CBNN to survival methods based on regression and neural
-networks in a simulation and three real data applications. We report
-two time-dependent metrics for each model. In the simulations and real
-data applications, CBNN provides a more consistent predictive
-performance across time and outperforms the competing neural network
-approaches. For a complex simulation, which highlights the ability of
-CBNN to model both a complex baseline hazard and time-varying
-interactions, CBNN outperforms all competitors. The first two real data
-application shows CBNN outperforming all neural network competitors,
-while a second real data application shows competitive performance.
-We highlight the benefit of combining case-base sampling with deep
-learning to provide a simple and flexible modeling framework for
-data-driven, time-varying interaction modeling of
-survival outcomes. An R package is available at
-\url{https://github.com/Jesse-Islam/cbnn}.}
+\abstract[Summary]{Neural network-based survival methods can model data-driven covariate interactions.
+While these methods have led to better predictive performance than regression-based approaches,
+they cannot model both time-varying interactions and complex baseline hazards. To address this, we
+propose Case-Base Neural Networks (CBNN) as a new approach that combines the case-base sampling
+framework with flexible architectures. Our method naturally accounts for censoring and does not require
+method specific hyperparameters. Using a novel sampling scheme and data augmentation, we incorporate
+time directly into a feed-forward neural network. CBNN predicts the probability of an event occurring at a
+given moment to estimate the hazard function. We compare the performance of CBNN to survival methods
+based on regression and neural networks in a simulation and three real data applications. We report two time-dependent
+metrics for each model. For a complex simulation, which highlights the ability of CBNN to model both a complex
+baseline hazard and time-varying interactions and two real data applications, CBNN outperforms all competitors
+on a simulation with a complex baseline hazard and time-varying interaction and two case studies. A final case
+study shows competitive performance. We highlight the benefit of combining case-base sampling with deep
+learning to provide a simple and flexible modeling framework for data-driven, time-varying interaction modeling
+of survival outcomes. An R package is available at https://github.com/Jesse-Islam/cbnn.}
 
@@
 
-Smooth-in-time accelerated failure time (AFT) models can estimate
-absolute risks by modeling the hazard directly through a user-specified
-baseline hazard distribution \citep{kleinbaum2012survival}. Cox
-proportional hazards models are used more often than AFT models, causing
-analyses to be based on hazard ratios and relative risks rather than on
-survival curves and absolute risks \citep{hanley2009}. The
-identification of an appropriate distribution for the baseline hazard in
-an AFT model may be difficult for common diseases that have many
-interacting risk factors, or a Cox model where the disease pathogenesis
-may change with age, making it difficult to maintain the proportional
-hazards assumption. For example, previous studies of breast cancer
-incidence have discovered time-varying interactions with covariates of
-interest, such as tumor size \citep{coradini2000time}. One approach to
-provide flexibility in the baseline hazard involves using the basis of
-splines on time in our model \citep{royston2002flexible}. However,
-regression-based models are limited in that they require prior knowledge
-about potential time-varying interactions and their quantitative
-effects.
+Smooth-in-time accelerated failure time (AFT) models can estimate absolute
+risks by modeling the hazard directly through a user-specified baseline hazard
+distribution \citep{kleinbaum2012survival}. Analysts use Cox proportional hazards
+models more often than AFT models \citep{hanley2009}. This preference causes
+analyses to be based on hazard ratios and relative risks rather than on survival
+curves and absolute risks \citep{hanley2009}. With AFT models, it may be difficult
+to identify an appropriate baseline hazard for common diseases with many
+interacting risk factors \citep{royston2002flexible}. In studies where the disease
+pathogenesis may change with age, we may see non-proportional hazards
+\citep{coradini2000time}. For example, previous studies of breast cancer incidence
+have discovered time-varying interactions with covariates of interest, such as tumor
+size \citep{coradini2000time}. One approach to provide flexibility in the baseline hazard
+involves using the basis of splines on time in the model \citep{royston2002flexible}.
+However, regression-based models require prior knowledge of potential time-varying
+interactions and their quantitative effects.
 
-Neural networks provide a data-driven approach to approximating
-interaction terms. For example, DeepSurv is a neural network-based
-proportional hazards model that implements the Cox partial
-log-likelihood as a custom loss function \citep{katzman2018DeepSurv},
-resulting in a stepwise absolute risk curve that cannot accommodate
-time-varying interactions. Compared to Cox regression, DeepSurv shows
-better performance on a real dataset \citep{katzman2018DeepSurv}. To handle
-non-proportional hazards, a modification to the loss function was
-proposed \citep{faraggi1995neural}.
+Neural networks provide a data-driven approach to approximating interaction terms.
+For example, DeepSurv is a neural network-based proportional hazards model that
+implements the Cox partial log-likelihood as a custom loss function \citep{katzman2018DeepSurv}.
+This results in a stepwise absolute risk curve that cannot accommodate time-varying interactions.
+Compared to Cox regression, DeepSurv shows better performance on a real dataset
+\citep{katzman2018DeepSurv}. Previous work suggests we may change the loss function
+to address non-proportional hazards \citep{faraggi1995neural}.
 
-As an alternative method that assumes a baseline hazard distribution,
-DeepHit specifies each survival time of interest in the model and
-directly estimates survival curves, rather than deriving a hazard
-function . It assumes an inverse Gaussian
-distribution as the baseline hazard and it outperformed DeepSurv \citep{lee2018DeepHit}.
+DeepHit assumes an inverse Gaussian distribution as the baseline hazard \citep{lee2018DeepHit}.
+It specifies each follow-up time time of interest in the model and directly estimates survival curves, rather
+than deriving a hazard function \citep{lee2018DeepHit}. DeepHit outperformed DeepSurv
+\citep{lee2018DeepHit}.
 
-Providing further flexibility, Deep Survival Machines (DSM) is a
-parametric survival model using neural networks with a mixture of
-distributions as the baseline hazard \citep{dsmPaper}. DSM outperformed DeepSurv and DeepHit
-\citep{dsmPaper}. However, like DeepHit and DeepSurv, DSM cannot model
-time-varying interactions.
+We note that these alternative neural network approaches require custom loss functions
+\citep{katzman2018DeepSurv} \citep{lee2018DeepHit}. DeepHit introduces a hyperparameter
+weighing its two loss functions (negative log-likelihood and ranking losses) \citep{lee2018DeepHit}.
+Regression-based approaches require prior specification of all interaction terms, which makes it
+challenging to model covariate effects that change with time. The current neural network models
+provide flexibility at the cost of clarity, while regression models provide clarity at the cost of flexibility.
 
-We note that these alternative neural network approaches all require
-custom loss functions \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}
-\citep{dsmPaper}. DeepHit introduces a hyperparameter weighing its two
-loss functions (negative log-likelihood and ranking losses) while DSM
-requires a two-phase learning process and user implementations for
-distributions beyond Log-Normal or Weibull \citep{lee2018DeepHit}
-\citep{dsmPaper}. Regression-based approaches require prior
-specification of all interaction terms, which makes it challenging to
-model covariate effects that change over time. The current neural
-network models provide flexibility at the cost of opacity, while
-regression models provide clarity at the cost of flexibility.
+In this article, we propose Case-Base Neural Networks (CBNN) as a method that models time-varying
+interactions and a flexible baseline hazard. Our approach to modeling the full hazard uses case-base
+sampling \citep{hanley2009}. This sampling technique allows probabilistic models to predict survival
+outcomes. After case-base sampling, we can easily implement the model in any language using
+common neural network components. As part of the case-base framework, we use transformations
+of time as a feature (covariate) to specify different baseline hazards. For example, by including
+splines of time as covariates, we can approximate the Royston-Parmar flexible baseline hazard model
+\citep{royston2002flexible}\citep{hanley2009}. However, this still requires explicit use of time-varying
+interactions. CBNN can model both without extra tuning parameters.
 
-In this article, we propose Case-Base Neural Networks (CBNN) as a method
-that models time-varying interactions and a flexible baseline hazard
-using commonly available neural network components. Our approach to
-modeling the full hazard uses case-base sampling \citep{hanley2009}.
-This sampling technique allows probabilistic models to predict survival
-outcomes. As part of the case-base framework, we use transformations of
-time as a feature (covariate) to specify different baseline hazards. For
-example, by including splines of time as covariates, we can approximate
-the Royston-Parmar flexible baseline hazard model
-\citep{royston2002flexible}, however, this still requires explicit use
-of time-varying interactions. CBNN can model both without extra tuning
-parameters.
-
-In Section \ref{methods}, we describe how case-base sampling and neural
-networks are combined both conceptually and algebraically, along with
-our hyperparameter choices and software implementation. Section
-\ref{sims}, describes our metrics and compares the performance of CBNN,
-DeepSurv, DeepHit, DSM, Cox regression and case-base using logistic
-regression (CBLR) on simulated data. Section \ref{casestudies} describes
-the real-data analysis, while Section \ref{discussion} explores the
-implications of our results and contextualizes them within neural
-network survival analysis in a single event setting.
+We describe how case-base sampling and neural networks combine
+in Section \ref{methods}, along with our hyperparameter selection procedure, metrics and software implementation.
+We explore the performance of CBNN, DeepSurv, DeepHit, Cox regression and case-base using logistic
+regression (CBLR) on simulated data in Section \ref{sims}. Section \ref{casestudies} describes the
+performance of the same methods in three case studies. Section \ref{discussion} explores the
+implications of our results and contextualizes them within neural network survival analysis in a
+single event setting.
 
@@
 \hypertarget{methods}{%
-\section{Case-base neural network methodology, metrics and
+\section{Case-base neural networks, metrics, hyperparameters and
 software}\label{methods}}
@@
 survival time into discrete person-specific moments (person-moments).
-Then, we detail how neural networks can be used within this framework,
+Next, we detail how neural networks can be used within this framework,
 explicitly incorporating time as a feature while adjusting for the
-sampling bias. Finally, we report on the software versions used. An R
+sampling bias. Then, we describe the metrics and hyperparameter selection procedure.
+Finally, we report on the software versions used. An R
 package is available for use at
@@
 Case-base sampling is an alternative framework for survival analysis
-\citep{hanley2009}. In case-base sampling, we sample from the continuous
+\citep{hanley2009}. Without this preliminary step, we cannot model survival
+outcomes with probabilistic methods. In case-base sampling, we sample from the continuous
 survival time of each person in our dataset to create a \emph{base
@@
 let \(B\) be the sum of all follow-up times for the individuals in the
-study. If we sample the \emph{base series} uniformly across the study
+study. let $c$ be the number of events in the \emph{case series}.
+A reasonable concern is determining how large $b$ should be relative to $c$.
+The size of \(b\) influences how much information is lost in the sampling process
+\citep{hanley2009}. Using our notation, Hanley $\&$ Miettinen (2009)
+adapt a quote from Mantel (1973): "By the reasoning that $\frac{cb}{c+b} [=(\frac{1}{c}+\frac{1}{b}]$
+measures the relative information in a comparison of two averages based on
+sample sizes of c and b respectively, we might expect by analogy, which would
+of course not be exact in the present case, that this approach would result in
+only a moderate loss of information. (The practicing statistician is generally
+aware of this kind of thing. There is little to be gained by letting the size of
+one series, b, become arbitrarily large if the size of the other series, c, must
+remain fixed)" \citep{hanley2009} \citep{mantel1}. If $b=100c$, then we expect learned
+weights to be at most one percent higher than if the entire study base $B$ was used,
+as they are proportional to $\frac{1}{c}+\frac{1}{100c}$ rather than $\frac{1}{c}+\frac{1}{\infty c}$ \citep{hanley2009}.
+
+If we sample the \emph{base series} uniformly across the study
 base, then the hazard function of the sampling process is equal to
-\(b/B\). Therefore, we have the following equality
+\(b/B\). Therefore, we have the following equality \cite{saarela2015}
 \footnote{We are abusing notation here, conflating hazards with probabilities. For a rigorous treatment, see Saarela \& Hanley (2015) section 3 \cite{saarela2015} .}:
@@
 
-To estimate the correct hazard function, we adjusting for the bias
-introduced when sampling a fraction of the study base \(B\) by adding
-the offset term \(\log\left(\frac{b}{B} \right)\) as in
-\eqref{eqn:offset}. Next, we propose using neural networks to model the
+To estimate the correct hazard function, we adjust for the bias
+introduced when sampling a fraction of the study base (\(\frac{b}{B}\)) by adding
+the term \(\log\left(\frac{B}{b} \right)\) to offset \(\log\left(\frac{b}{B} \right)\)
+during the fitting process. Next, we propose using neural networks to model the
 odds.
@@
 then passed through a sigmoid activation function (Figure
-\ref{fig:NNarch}). We use the sigmoid function as its inverse is the
-odds, which we can use to calculate the hazard. The general form for the
+\ref{fig:NNarch}). We are interested in predicting the odds. This makes the sigmoid
+activation function ideal as it is the inverse of the odds. which we can use to calculate the hazard. The general form for the
 neural network using CBNN is:
 
-\begin{align}\label{eqn:nnProb}
+\begin{align*}
 P\left(Y=1|X,T\right)&=\mathrm{sigmoid}\left(f_{\theta}(X, T) + \log\left(\frac{B}{b}\right) \right),
-\end{align}
+\end{align*}
 
@@
 represents any feed-forward neural network architecture,
-\(\log\left(\frac{B}{b}\right)\) is bias term set by case-base sampling,
+\(\log\left(\frac{B}{b}\right)\) is the offset term for adjust for the bias
+(\(\log\left(\frac{b}{B}\right)\)) set by case-base sampling,
 \(\theta\) is the set of parameters learned by the neural network and
@@
 
-{\centering \includegraphics[width=1\linewidth]{../../../figures/nnarch2}}
+{\centering \includegraphics[width=1\linewidth]{../figures/nnarch2}}
 
@@
 (Figure \ref{fig:NNarch}). Finally, we can convert this model output to
-a hazard. When using our model for predictions, we manually set the
-offset term to 0 in the new data, as we account for the bias during the
+a hazard. To use CBNN for predictions, we manually set the
+offset term \(\log\left(\frac{B}{b} \right)\) to 0 in the new data as we already account for the bias during the
 fitting process.
@@
 
+\hypertarget{performance-metrics}{%
+\subsection{Performance metrics}\label{performance-metrics}}
+
+To choose our method-specific hyperparameters, we use the Integrated
+Brier Score (IBS) \citep{graf1999}. We use two metrics to assess the
+performance of the different methods of interest on testing data: 1) Index of Prediction Accuracy (IPA)
+\citep{kattan2018index} and 2) Inverse probability censoring weights-adjusted time dependent
+area under the receiver operating characteristic curve ($AUC_{IPCW}$) \citep{auc}, which we 
+define below. We achieve a summarized assessment of performance for each model with the
+IBS. The IPA is a metric for both discrimination and calibration. The $AUC_{IPCW}$
+provides a metric for discrimination only. Together, we can get a clearer picture about how
+each model performs in each study and why.
+
+\hypertarget{index-of-prediction-accuracy-ipa}{%
+\subsubsection{Index of prediction accuracy (IPA)}\label{index-of-prediction-accuracy-ipa}}
+
+The IPA is a function of the Brier score (\(BS(t)\)) \citep{graf1999},
+which is defined as \begin{align}\label{eqn:bs}
+BS(t)=\frac{1}{N}\sum^{N}_{i=1}\left(\frac{\left(1 - \widehat{F}(t \mid X_{i})\right)^{2}\cdot I(T_{i}\leq t,\delta_{i}=1)}{\widehat{G}(T_{i})} + \frac{\widehat{F}(t\mid X_{i})^{2}\cdot I(T_{i}>t)}{\widehat{G}(t)}\right),
+\end{align} where \(\delta_{i}=1\) shows individuals who have
+experienced the event, \(N\) represents the number of samples in our
+dataset over which we calculate \(BS(t)\), \(\widehat{G}(t)=P[c>t]\) is
+a non-parametric estimate of the censoring distribution, \(c\) is
+censoring time and \(T_{i}\) is an individual's survival or censoring
+time. The BS provides a score that accounts for the information
+loss because of censoring. There are three categories of individuals
+that may appear within the dataset once we fix our \(t\) of interest.
+Individuals who experienced the event before \(t\) are present in the
+first term of the equation. The second term of the equation includes
+individuals who experience the event or are censored after \(t\). Those
+censored before \(t\) are the third category of people. The inverse
+probability censoring weights (IPCW) adjustment (\(G(\cdot)\)) is to
+account for these category three individuals whose information is
+missing.
+
+The IPA as a function of time is given by \begin{align}
+\textrm{IPA}(t) &= 1-\frac{BS_{model}(t)}{BS_{null}(t)}, \nonumber
+\end{align} where \(BS_{model}(t)\) represents the BS over time
+\(t\) for the model of interest and \(BS_{null}(t)\) represents the
+BS if we use an unadjusted Kaplan-Meier (KM) curve as the
+prediction for all observations \citep{kattan2018index}. Note that IPA
+has an upper bound of one, where positive values show an increase in
+performance over the null model and negative values show that the null
+model performs better. These scores show how performance changes over
+follow-up time. As an extension of BS, IPA models for both calibration
+and discrimination\citep{graf1999} \citep{kattan2018index}.
+
+\hypertarget{integrated-brier-score-ibs}{%
+\subsubsection{Integrated Brier Score (IBS)}\label{integrated-brier-score-ibs}}
+
+The Integrated Brier Score (IBS) is a function of the BS \ref{eqn:bs} as well \citep{graf1999},
+which is defined as \begin{align*}
+IBS(t)=\int_{t}^{t_{max}}BS(t)dw(t),
+\end{align*} where $w(t)=\frac{t}{t_{max}}$. As this is a score for a range of follow-up-times,
+it is a useful metric to track during cross-validation. The performance at a specific time is lost,
+and is why we opt for the IPA in assessing performance on a test set.
+
+\hypertarget{auc}{%
+\subsubsection{Inverse probability censoring weights-adjusted time dependent area under the receiver operating characteristic curve ($AUC_{IPCW}$) }\label{auc}}
+
+The IPCW-adjusted AUC ($AUC_{IPCW}$) is a time-dependent metric that considers censoring \citep{auc}. 
+For a given follow-up time of interest,
+ \begin{align*}
+AUC_{IPCW}(t)=\frac{\sum^{n}_{i=1}\sum^{n}_{j=1} I \left(\widehat{F}(t|X_{i})>\widehat{F}(t|X_{j})\right) \cdot I(T_{i}\leq t,\delta_{i}=1) \cdot \left( 1-I(T_{j}\leq t,\delta_{j}=1) \right) \cdot W_{i}(t) \cdot W_{j}(t)} { \sum^{n}_{i=1}\sum^{n}_{j=1} I(T_{i}\leq t,\delta_{i}=1) \cdot \left( 1-I(T_{j}\leq t,\delta_{j}=1) \right) \cdot W_{i}(t) \cdot W_{j}(t) },
+\end{align*}where \begin{align*}
+W_{i}(t)=\frac{I(T_{i}\leq t,\delta_{i}=1)}{\widehat{G}(T_{i})} +\frac{I(T_{i}>t)}{\widehat{G}(t)}.
+\end{align*} The $AUC_{IPCW}$ measures discrimination \citep{auc}.
+By taking both IPA and $AUC_{IPCW}$ together, we can better understand whether a model is better
+in terms of calibration or discrimination.
+
 \hypertarget{hyperparameter-selection}{%
@@
 
-Neural networks are flexible when defining the architecture and
-optimization parameters. These hyperparameter choices can affect the
-estimated parameters and are not necessarily the same for each model.
-First we fix a test set with 15\% of the data, then we use three-fold cross-validation on the remaining 
-data (85\%:15\% training:validation) for a range of hyperparameters (TABLE REF). We track integrated brier 
-score on the validation set for each hyperparameter combination, choosing 
-the combination with the lowest score for each method. The chosen 
-hyperparameters are in (TABLE REF).
+Neural networks are flexible when defining the architecture and optimization parameters.
+These hyperparameter choices can affect the estimated parameters and are not necessarily
+the same for each model. We apply the same procedure to each study in this paper. First,
+we fix a test set with 15\% of the data. We keep aside this data during hyperparameter
+selection and we fit a model for each method in each study. To determine the best
+hyperparameters for each neural network method, we use three-fold cross-validated grid
+search \citep{gulli2017} on the remaining data (85\% training, 15\% validation) for the
+following range of hyperparameters:
+
+\[
+Grid\ search: \begin{cases}
+Learning\ rate \sim \{0.001, 0.01\} \\ %technically a control and treatement group
+Dropout \sim \{0.01,0.05, 0.1\} \\
+First\ layer\ nodes \sim \{50, 75, 100\} \\
+Second\ layer\ nodes \sim \{10,25,50\} \\
+Number\ of\ batches \sim \{100, 500\} \\
+Activation\ function \sim \{ReLU, Linear\}\\
+\alpha \sim \{0, 0.5, 1\}.
+\end{cases}
+\]
+
+\begin{table}
+\caption{Four tables showing the chosen hyperparameters along with the IBS for each neural network model in the complex simulation (A), multiple myeloma (MM) case study (B), free light chain (FLC) case study (C) and prostate cancer (Prostate) case study (D).}
+\label{tab:wins}
+\begin{center}\includegraphics[width=1\linewidth]{../figures/winTable} \end{center}
+\end{table}
+
+DeepHit uses $\alpha$ as a specific hyperparameter, while DeepSurv and CBNN do not.
+We track IBS on the validation set for each hyperparameter combination,
+choosing the combination with the lowest score for each method.
+We present the chosen hyperparameters in Table 1.
 
@@
 \citep{survpkg}, the CBLR model using the \textbf{casebase} package
-\citep{cbpkg}, the DeepSurv model using the \textbf{survivalmodels}
-package \citep{survmods}, the DeepHit model using \textbf{pycox}
-\citep{lee2018DeepHit} and the DSM model using
-\textbf{DeepSurvivalMachines} \citep{dsmPaper}. We made the components
+\citep{cbpkg}, DeepSurv and DeepHit using \textbf{pycox}
+\citep{lee2018DeepHit}. We made the components
 of CBNN using the \textbf{casebase} package for the sampling step and
@@
 to fit a flexible baseline hazard using splines for our complex
-simulation. We use the implementation of \(C_{IPCW}\) from the python
-package \textbf{sksurv} \citep{sksurv}. The \textbf{riskRegression}
+simulation. The \textbf{riskRegression}
 package \citep{riskRegression} is used to get the Index of Prediction
-Accuracy (IPA metric). Both metrics are described in detail in the
+Accuracy (IPA metric) and AUC. Both metrics are described in detail in the
 following section. We modify the \textbf{riskRegression} package to be
@@
 
-\hypertarget{performance-metrics}{%
-\section{Performance metrics}\label{performance-metrics}}
-
-To choose our method-specific hyperparameters, we use the Integrated 
-Brier Score (IBS) \citep{graf1999}. We use two metrics to assess the 
-performance of the different methods of interest on testing data: 1) (IPA) 
-\citep{kattan2018index} and 2) inverse probability
-censoring weights-adjusted concordance index (\(C_{IPCW}\))
-\citep{uno2011}, which we define below. IBS provides a summarized assessment
-of performance for each model. (\(C_{IPCW}\)) provides information
-about performance up to the survival time of interest. while IPA provide transparency
-as to when in follow-up time each model may
-perform better than the others.
-
-\hypertarget{index-of-prediction-accuracy-ipa}{%
-\subsubsection{Index of prediction accuracy (IPA)}\label{index-of-prediction-accuracy-ipa}}
-
-The IPA is a function of the Brier score (\(BS(t)\)) \citep{graf1999},
-which is defined as \begin{align}
-BS(t)=\frac{1}{N}\sum^{N}_{i=1}\left(\frac{\left(1 - \widehat{F}(t \mid X_{i})\right)^{2}\cdot I(T_{i}\leq t,\delta_{i}=1)}{\widehat{G}(T_{i})} + \frac{\widehat{F}(t\mid X_{i})^{2}\cdot I(T_{i}>t)}{\widehat{G}(t)}\right),
-\end{align} where \(\delta_{i}=1\) shows individuals who have
-experienced the event, \(N\) represents the number of samples in our
-dataset over which we calculate \(BS(t)\), \(\widehat{G}(t)=P[c>t]\) is
-a non-parametric estimate of the censoring distribution, \(c\) is
-censoring time and \(T_{i}\) is an individual's survival or censoring
-time. The Brier score provides a score that accounts for the information
-loss because of censoring. There are three categories of individuals
-that may appear within the dataset once we fix our \(t\) of interest.
-Individuals who experienced the event before \(t\) are present in the
-first term of the equation. The second term of the equation includes
-individuals who experience the event or are censored after \(t\). Those
-censored before \(t\) are the third category of people. The inverse
-probability censoring weights (IPCW) adjustment (\(G(\cdot)\)) is to
-account for these category three individuals whose information is
-missing. The IPA score as a function of time is given by \begin{align}
-\textrm{IPA}(t) &= 1-\frac{BS_{model}(t)}{BS_{null}(t)}, \nonumber
-\end{align} where \(BS_{model}(t)\) represents the Brier score over time
-\(t\) for the model of interest and \(BS_{null}(t)\) represents the
-Brier score if we use an unadjusted Kaplan-Meier (KM) curve as the
-prediction for all observations \citep{kattan2018index}. Note that IPA
-has an upper bound of one, where positive values show an increase in
-performance over the null model and negative values show that the null
-model performs better. These scores show how performance changes over
-follow-up time.
-
-A potential artifact of IPA is that the score is unstable at earlier and
-later survival times. This is because of a near equivalent Brier score
-among each model and the null model. At small values, a difference of
-\(0.1\) creates a more significant fold change than at larger values. As
-the Brier score is potentially small at the start and at the end of
-their respective curves, The IPA score may be unstable at the same
-locations.
-
-\hypertarget{integrated-brier-score-ibs}{%
-\subsubsection{Integrated Brier Score (IBS)}\label{integrated-brier-score-ibs}}
-
-The IBS is a function of the Brier score (\(BS(t)\)) \citep{graf1999}. as well,
-which is defined as \begin{align}
-IBS=\int_{t}^{t_{max}}BS(t)dw(t)
-\end{align} where $w(t)=\frac{t}{t_{max}}$. As this is a score for a range of follow-up-time,
-it is a useful metric to track during cross-validation. The performance at a specific time
-is hidden, and is why we opt for the IPA in assessing the final models on a test set.
-
-\hypertarget{inverse-probability-censoring-weights-adjusted-concordance-index}{%
-\subsubsection{Inverse probability censoring weights-adjusted
-concordance
-index}\label{inverse-probability-censoring-weights-adjusted-concordance-index}}
-
-The \(C_{IPCW}\) is a non-proper, rank-based metric that does not depend
-on the censoring times in the test data \citep{uno2011}. The
-\(C_{IPCW}\) is given by
-
-\begin{align} \label{eq:cidx}
-C_{IPCW}(t) &= \frac{\sum^{N}_{i=1}\sum^{N}_{j=1}\delta_{i}\left\{\widehat{G}(T_{i})\right\}^{-2} I(T_{i}<T_{j},T_{i}<t) I\left(\widehat{F}(t|X_{i})>\widehat{F}(t|X_{j})\right)}{\sum^{N}_{i=1}\sum^{N}_{j=1}\delta_{i}\left\{\widehat{G}(T_{i})\right\}^{-2} I(T_{i}<T_{j},T_{i}<t)}.
-\end{align} where the follow-up period of interest is (0,\(t\)),
-\(I(\cdot)\) is an indicator function, \(\widehat{F}(X_{i},t)\) is the
-risk function estimated for everyone in the study at time \(t\) and
-\(C_{IPCW}\) can compare the performance of different models, where a
-higher score is better. Note that the \(C_{IPCW}\) may produce
-misleading performance, as it ranks based on survival times, not event
-status \citep{cindexfails2019}. This metric is considered an unbiased
-population concordance measure because of the IPCW adjustment
-\citep{uno2011}.
-
 \begin{figure}
 
-{\centering \includegraphics[width=1\linewidth]{../../../analyses/figures/megaPlot}
+{\centering \includegraphics[width=1\linewidth]{../figures/megaPlot}
 
@@
 
-\caption{Summarizes the simple simulation (A, B), complex simulation (C, D), SUPPORT case study (E, F) and METABRIC case study (G, H) results.
-The first row shows the IPA score for each model in each study over follow-up time.
+\caption{Summarizes the complex simulation (A, E), multiple myeloma (MM) case study (B, F), free light chain (FLC) case study (C, G) and prostate cancer (Prostate) case study (D, H).
+The first row shows the IPA for each model in each study over follow-up time.
 Negative values mean our model performs worse than the null model and positive values mean the model performs better.
-The second row demonstrates the $C_{IPCW}$ score for each model in each study over follow-up time.
-A score of 1 is the maximum performance for either metric.
-Each model-specific metric in each study shows a 95-percent confidence interval over 100 iterations.
-The models of interest are case-base with logistic regression (CBLR), Case-Base Neural Networks (CBNN), Cox, DeepHit, DeepSurv, Deep Survival Machines (DSM), Optimal (a CBLR model with the exact interaction terms and baseline hazard specified) and Kaplan-Meier (to serve as a baseline, predicting the average for all individuals).}\label{fig:megaPlot}
+The second row shows the $AUC_{IPCW}$ for each model in each study over follow-up time.
+Each model-specific metric in each study shows a 95\% confidence interval over 100 iterations.
+The models of interest are case-base with logistic regression (CBLR), Case-Base Neural Networks (CBNN), Cox, DeepHit and DeepSurv.
+Both CBLR and Cox have near identical performance, resulting in curves that are overlapped.
+The Kaplan-Meier (KM) model serves as a baseline, predicting the average curve for all individuals.
+The Optimal (a CBLR model with the exact interaction terms and baseline hazard specified) model shows the best performance we can expect on the simulated data.
+}\label{fig:megaPlot}
 \end{figure}
@@
 time-varying interactions, each with 10\% random censoring. For both
-settings, we simulate three covariates:
+settings, we simulate three covariates for 5000 individuals:
 
@@
 \end{cases} \qquad \qquad
-z_{3} \sim N(1,0.5)
+z_{3} \sim N(1,0.5).
 \]
-Besides the methods mentioned above, we include the Optimal model in our
-comparisons using CBLR. That is, we include the exact functional form of
-the covariates in a CBLR model (referred to as Optimal for simplicity).
-We calculate \(t\)-based 95\% confidence intervals using 100
-replications of the simulated data. For all analyses, 15\% of the data is 
-kept for testing, while 15\% of the remaining data is kept for validation. 
-The remaining data is kept for training. We predict risk functions \(F\) using
-\eqref{eqn:ci2} for individuals in the test set, which are used to
-calculate our \(C_{IPCW}\) and IPA scores. We conduct 100 bootstrap
-re-samples on the training data to obtain confidence intervals.
+Besides the methods mentioned above, we include the Optimal model in our comparisons using CBLR.
+We include the exact functional form of the covariates in a CBLR model (referred to as Optimal for simplicity).
+We conduct 100 bootstrap re-samples on the training data to get confidence intervals. We kept 15\% of the
+data for testing before hyperparameter selection. 15\% of the remaining data is used for validation and the rest is reserved
+for training. We predict risk functions for individuals in the test set, which are used to calculate our $AUC_{IPCW}$ and IPA.
 
@@
 
-This simulation demonstrates performance with the presence of a complex
+This simulation shows performance with the presence of a complex
 baseline hazard and a time-varying interaction. Inspired by a cancer treatment
@@
 \citep{royston2002flexible}, the breast cancer dataset provides a
-complex hazard from which we simulate, available in the
+complex baseline hazard from which we simulate, available in the
 \textbf{flexsurv} R package \citep{flexsurv}. To increase the complexity
 of our data-generating mechanism for this simulation, we design the
-model as follows: \begin{align}
+model \begin{align}
 \log h(t \mid X_i) =\sum_{i=1}^{5} (\gamma_{i} \cdot \psi_{i}) + \beta_{{1}} (z_{1}) + \beta_{{2}} (z_{2})+ \beta_{{3}} (z_{3})+ \tau_{1} ( z_{1} \cdot t)+ \tau_{2} (z_{2} \cdot z_{3}), \nonumber
 \end{align}
-
-\begin{table}
-\caption{Four tables representing performance at certain follow-up times for the simple simulation, complex simulation, SUPPORT and METABRIC.
-Each table shows performance for each method in each study at $25\%$, $50\%$.
-$75\%$ and $100\%$ of follow-up time.
-The bold elements show the best model for each study, at each follow-up time of interest.
-These tables are included to provide exact measures at certain intervals.
-The models of interest are: Cox, case-base with logistic regression (CBLR), DeepSurv, DeepHit, Case-Base Neural Network (CBNN) and Optimal}
-\label{tab:megaTable}
-
-\begin{center}\includegraphics[width=1\linewidth]{../../../analyses/figures/megaTable} \end{center}
-
-\end{table}
 
@@
 \citep{flexsurv}. Note that we fix these values for the analysis. The
-\(\beta\) coefficients represent direct effects, \(\tau_{2}\) and
-\({\tau_3}\) represent interactions and \(\tau_{1}\) is a time-varying
+\(\beta\) coefficients represent direct effects, \(\tau_{2}\) represents
+an interaction and \(\tau_{1}\) is a time-varying
 interaction.
 
-We use a population time plot to visualize incidence density across $z1$ (FIGURE REF).
-z1 acts to mimic a treatment and control group. The control group $(z1=0)$ is not impacted
-by the time-varying interaction. The treatment group $(z1=1)$ initially acts to protect the group,
-while the risk slowly builds with time at the rate of the interaction term $(\tau_{1})$. This creates
-a separation between the median time of death between these two groups due to the time-varying
-interaction. We expect all competing models (aside from the Optimal model and CBNN) to drop in
-performance for the second half of follow-up time.
-
 \hypertarget{performance-comparison-in-complex-simulation}{%
-\subsubsection{Performance comparison in complex
+\subsection{Performance comparison in complex
 simulation}\label{performance-comparison-in-complex-simulation}}
@@
 Figure \ref{fig:megaPlot} A, E and Table \ref{tab:megaTable} A
-demonstrates the performance over time on a test set. While examining the IPA score, the optimal
-regression model and CBNN perform near identically, leading in performance followed by 
-DeepHit, the linear regression models and DeepSurv. We expected the Optimal model to perform best in both
-metrics. However, this was not the case for \(C_{IPCW}\) and may be due
-to an artifact of concordance-based metrics, where a misspecified model
-may perform better than a correctly specified one
-\citep{cindexfails2019}. We attribute the performance of CBNN to its
-flexibility in modeling time-varying interactions and baseline hazard,
-flexibility the other neural network models do not have. We keep this misspecification issue
-in mind while interpreting the \(C_{IPCW}\) results in the case studies.
+shows the performance over time on a test set. For discrimination, the Optimal model
+performs best as expected, followed by CBNN, DeepHit, DeepSurv and the linear models (Figure \ref{fig:megaPlot} E).
+For both discrimination and calibration, the rankings remain the same aside from DeepHit,
+where there's a periodic loss in performance (Figure \ref{fig:megaPlot} A). The
+Optimal model acts as a reference for how good a model can perform,
+with CBNN performing better than the competitors in the complex simulation.
+To obtain a more realistic performance assessment, we compared models in
+three case studies with a time-to-event outcome.
+
+\begin{table}[!htbp]
+\caption{Four tables representing performance at certain percentages of follow-up time in the complex simulation (A), multiple myeloma (MM) case study (B), free light chain (FLC) case study (C) and prostate cancer (Prostate) case study (D).
+Each table shows performance for each method in each study at $25\%$, $50\%$.
+$75\%$ and $100\%$ of follow-up time.
+The bold elements show the best model for each study, at each follow-up time of interest.
+These tables are included to provide exact measures at certain intervals.
+The models of interest are: Cox, case-base with logistic regression (CBLR), DeepSurv, DeepHit, Case-Base Neural Network (CBNN) and Optimal (in the complex simulation).
+The CBNN label is in bold to make it easier to distinguish.
+The best score at each percent of follow-up time is highlighted in bold.
+If the average performance is tied, then all tied values are bold.}
+\label{tab:megaTable}
+
+\begin{center}\includegraphics[width=1\linewidth]{../figures/megaTable} \end{center}
+
+\end{table}
 
@@
 
-Our complex simulation demonstrates the superior performance of CBNN in
-ideal conditions with clean data. To obtain a more realistic performance
-assessment, we compared models using three real datasets with a
-time-to-event outcome. The first case study examines multiple myeloma (MM) \citep{myeloma}.
-The second examines the relationship between serum free light chain (FLC) and mortality \citep{flc}.
-The third examines prostate cancer survival on a publicly available simulation of the Surveillance,
-Epidemiology, and End Results (SEER)-medicare prostate cancer data\citep{prostate}.
-As we do not know the true model for the real
-data, we exclude the Optimal model. After grid search is performed with the same hyperparameter
-options, We split the data the same way we did for the simulation (15\% of the data is 
-kept for testing, while 15\% of the remaining data is kept for validation and the rest for testing).
+ The first case study examines multiple myeloma (MM) \citep{myeloma}.
+Case study two examines the relationship between serum free light chain (FLC) and mortality \citep{flc}.
+The third examines prostate cancer (Prostate) survival on a publicly available simulation of the Surveillance,
+Epidemiology, and End Results (SEER)-medicare study \citep{prostate}. Both MM and FLC
+are available as part of the \textbf{survival} package \citep{survpkg}. The Prostate dataset is available
+as part of the \textbf{asaur} package \citep{asaur}. As we do not know the true model for the case studies,
+we exclude the Optimal model. After we perform grid search with the same hyperparameter
+options, we split the data the same way we did for the simulation (allocated 15\% of the data for testing,
+15\% of the remaining data for validation, and the rest for testing).
 We predict risk functions for everyone in the test set, which is used to calculate our metrics. We
 conduct 100-fold bootstrap re-samples on the training data to
-obtain confidence intervals.
+get confidence intervals.
 
@@
 \subsection{Performance evaluation on multiple myeloma dataset}\label{pe-multiplemyeloma}}
-The (MM) dataset tracks 3882 patients seen at the Mayo Clinic from 1947-1996 until death \citep{myeloma}.
-We use 2 covariates, year of entry into the study and the time of MM diagnosis\citep{myeloma}. We see 71\% incidence over 23 years\citep{myeloma}.
+The MM study tracks 3882 patients seen at the Mayo Clinic from 1947 to 1996 until death \citep{myeloma}.
+We use 2 covariates, year of entry into the study and the time of MM diagnosis\citep{myeloma}.
+We see 71\% incidence over 23 years\citep{myeloma}.
 
 Figure \ref{fig:megaPlot} B, F and Table \ref{tab:megaTable} B
-demonstrates the performance over time on a test set. With the IPA score, CBNN outperforms
-The competing models, followed by the linear models, DeepSurv and DeepHit. We note that after
-50\% of survival time, all models aside from CBNN perform substantially worse than a null model.
-We note with \(C_{IPCW}\), the linear models and DeepSurv perform equivalently, followed by CBNN and
-DeepHit.
+demonstrate the performance over time on a test set. For discrimination,
+CBNN and DeepHit have a similar performance, followed by DeepSurv and finally the linear models performing worst (Figure \ref{fig:megaPlot} F). For
+both discrimination and calibration, CBNN performs substantially better than
+the linear models and DeepSurv, with DeepHit having the worst performance overall
+(Figure \ref{fig:megaPlot} B). Together, CBNN is the best calibrated model and one
+of the best at discrimination in the MM case study.
+
 %paragraph about performance
@@
 The FLC dataset is a random sample of half the individuals in $\frac{2}{3}$ of the
-residents of Olmsted County over the age of 50\citep{flc}. We have access to 7874 subjects
-tracked until death\citep{flc}. We use 5 covariates, total serum FLC (kappa+lambda), age, sex,
-serum creatine and monoclonal gammapothy state\citep{flc}. we see 27\% incidence over 14 years\citep{flc}.
+residents of Olmsted County over the age of 50 \citep{flc}. We have access to 7874 subjects
+tracked until death \citep{flc}. We use 5 covariates, total serum FLC (sum of kappa and lambda), age, sex,
+serum creatine and monoclonal gammapothy state \citep{flc}. There is 27\% incidence over 14 years \citep{flc}.
 
 Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C
-demonstrates the performance over time on a test set. With the IPA score, CBNN outperforms
-The competing models, followed by the linear models and lastly DeepSurv and DeepHit for the first 60\% of follow-up time. 
-After 60\% of follow-up time, DeepHit has similar performance to CBNN. For the \(C_{IPCW}\) score, CBNN outperforms all models,
-followed by the linear models and finally DeepSurv. Note that for earlier ranges of survival time, DeepHit is the worst performing model. 
-After 25\% of follow-up time, DeepHit is in second place. In the FLC dataset, CBNN has a consistent top ranking for both IPA and  \(C_{IPCW}\).
+demonstrate the performance over time on a test set. CBNN, DeepHit and the linear models
+perform best at discrimination, followed DeepSurv (Figure \ref{fig:megaPlot} G). The rankings
+remain the same aside from DeepHit where the performance periodically drops
+to worse than DeepSurv (Figure \ref{fig:megaPlot} C). Together, CBNN outperforms the competing models,
+in terms of both calibration and discrimination in the FLC study. However, CBNN is only slightly better
+than the linear ones.
 
@@
 \subsection{Performance evaluation on prostate cancer dataset}\label{pe-prostate}}
-The prostate cancer dataset has a record of competing risks\citep{prostate}. As we are only interested in the
-single event scenario, we only keep individuals with prostate cancer death or censoring\citep{prostate}.
-This subset tracks 11054 individuals with three covariates, differentiation grade, age group and cancer state\citep{prostate}.
+The prostate cancer dataset has a record of competing risks \citep{prostate}. As we are only interested in the
+single event scenario, we only keep individuals with prostate cancer death or censoring \citep{prostate}.
+This subset tracks 11054 individuals with three covariates: differentiation grade, age group and cancer state \citep{prostate}.
 There is a 7\% incidence over 10 years.
 
-Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C shows the
-performance over time on a test set. With the IPA score, the linear models
-outperform the neural network ones, followed by CBNN and DeepHit, and
-finally DeepSurv. For the \(C_{IPCW}\) score, the linear models, CBNN, DeepHit
-perform near identically, side from DeepHit with an initial drop in performance.
-DeepSurv performed substantially worse than the other models.
+Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C demonstrate the
+performance over time on a test set. In the prostate case study, the linear models
+outperform the other models in both discrimination and calibration, followed by
+CBNN and DeepHit and finally DeepSurv (Figure \ref{fig:megaPlot} C, G). Aside from DeepSurv,
+the performance is similar across all models.
 
@@
 interactions in our hazard function. The two competing neural network
-models we evaluated cannot model time-varying interactions by design
- \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}
+models we evaluated cannot model time-varying interactions
+ \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}.
+
+A method we want to include is Deep Survival Machines (DSM), a
+parametric survival model using neural networks with a mixture of
+distributions as the baseline hazard \citep{dsmPaper}. It can model
+a flexible baseline hazard. However, like DeepHit
+and DeepSurv, DSM cannot model time-varying interactions \citep{dsmPaper}. DSM
+requires a two-phase learning process and user implementations for
+distributions beyond Log-Normal or Weibull
+\citep{dsmPaper}. DSM outperformed DeepSurv and DeepHit, but only tested
+a subset of hyperparameter combinations for competing neural network methods
+while going through a full grid search for their method \citep{dsmPaper}.
 With our hyperparameters options and model design, DSM did not
-converge in the complex simulation. Due to this issue limitation,
-we did not include this method. Compared to CBNN, all three neural network
-models also have limitations. DeepSurv is a proportional hazards model and does
+converge in the complex simulation. This may be due to implementation issues or
+method limitations, therefore we did not include this method.
+
+Compared to CBNN, the neural network competitors also have limitations.
+DeepSurv is a proportional hazards model and does
 not estimate the baseline hazard \citep{katzman2018DeepSurv}. DeepHit
-requires an alpha hyperparameter, is restricted to a single distribution
+requires an alpha hyperparameter, assumes a single distribution
 for the baseline hazard and models the survival function directly
-\citep{lee2018DeepHit}. Though not included, DSM can model a flexible
-baseline hazard, however is not meant for time-varying interactions\citep{dsmPaper}.
-The alternative neural network methods match on time, while CBNN models time directly.
+\citep{lee2018DeepHit}. The alternative neural network methods
+match on time, while CBNN models time directly.
 
-To assess performance among these models, we use both IPA and
-\(C_{IPCW}\) metrics. Concordance-based measures are commonly used in
-survival analysis to compare models and we opt to keep them in our
-analyses. However, \(C_{IPCW}\) is a non-proper metric and may cause
-misspecified models to appear better than they should
-\citep{cindexfails2019}. Therefore, we contextualize our \(C_{IPCW}\)
-results in relation to IPA, a proper scoring rule. The model rankings
-between \(C_{IPCW}\) and IPA differed for all but FLC. If we consider
-Figure \ref{fig:megaPlot} A, E and Table \ref{tab:megaTable} A in conjuction
-with FIGURE REF POPTIME, We note a universal drop in \(C_{IPCW}\) performance
-for all models. We hide the confidence intervals in Figure \ref{fig:megaPlot} E due to visibility but list them in Table \ref{tab:megaTable} A .
-aside from CBNN, all models have a wide confidence band at 75\% of follow-up time.
-We expect the optimal model to perform best, but that's not the case looking at  \(C_{IPCW}\).
-As such, we
+Though we applied a full grid search for all neural network models,
+there are an infinite number of untested hyperparameters we did not
+test. A completely exhaustive search is computationally infeasible,
+therefore it is reasonable to expect that there is a set of hyperparameters
+that is better suited for each model. However, we test a reasonably large
+range while accounting for potential over-fitting by including dropout
+\citep{gulli2017}. We provided the same options to all models aside from
+DeepHit, which has a method specific hyperparameter $\alpha$
+\citep{lee2018DeepHit}. With these limitations in mind, we dive
+into the differences in performance across all tested methods.
 
-With this interpretation of \(C_{IPCW}\) in mind, we assess a
-simulations with minimal noise and three case studies.
+The complex simulation requires a method that can learn both time-varying
+interactions and have a flexible baseline hazard. Here, CBNN shows
+a distinct advantage over all other methods. Based on our complex simulation results (Figure
+\ref{fig:megaPlot} A, E and Table \ref{tab:megaTable} A), CBNN
+outperforms the competitors. This simulation shows how all models perform
+under ideal conditions with minimal noise in the data, while the three case studies
+assess their performance in realistic conditions. In the MM case study, flexibility in both
+interaction modeling and baseline hazard improves CBNN's performance over
+the other models, suggesting that this flexibility aids calibration (Figure
+\ref{fig:megaPlot} B, F and Table \ref{tab:megaTable} B). upon examination of the FLC case study,
+CBNN demonstrates small improvement to performance compared to the linear models and DeepHit
+for both IPA and AUC (Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C).
+In the Prostate case study, the linear models outperform
+the neural network ones, while CBNN and DeepHit alternate their positions depending
+on the follow-up time of interest and DeepSurv maintains last place
+(Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C). We attribute this to
+potential over-parameterization in the neural network models, as we did not test for fewer
+nodes in each hidden layer, even with dropout. Though the
+ranking places the linear models above the neural network ones, the overall
+performance is comparable aside from DeepSurv, falling within a small range of IPA and AUC values.
 
-The complex simulation requires method that can learn both time-varying
-interactions and have a flexible baseline hazard. Here, CBNN demonstrates
-a distinct advantage over all other methods. Based on our complex simulation results (Figure
-\ref{fig:megaPlot} C, D and Table \ref{tab:megaTable} B), CBNN
-outperforms the competitors when time-varying interactions and a complex
-baseline hazard are present (after 50\% of survival time, Figure REF POPTIME).
-This simulation shows how CBNN can perform
-under ideal conditions, while the three case studies
-serve to assess its performance in realistic conditions.
-
-In the MM and FLC case studies, flexibility in both
-interaction modeling and baseline hazard improved CBNN's performance over
-the other models, suggesting that this flexibility aids prediction in
-both case studies. In the Prostate case study, the linear models outperform
-the neural network ones. CBNN and DeepHit alternate their positions depending
-on the survival time and DeepSurv maintains last place.
-We attribute this to potential overparameterization in the neural network models
-as we did not test for a small number of nodes in each hidden layer, even with dropout.
-Though the ranking places the linear models above the neural network ones, the overall
-performance is comparable aside from DeepSurv, falling within a small range of IPA scores.
+If prediction is our goal, we suggest CBNN as the best model in the Complex simulation, MM case study
+and FLC case study. The linear models were competitive in the FLC case study and performed best
+in the prostate case study. Though neural network interpretability is steadily improving \citep{interpret},
+there is still a trade-off compared to regression models, especially
+when the predictive performance gain is minimal, like in the FLC case study. We suggest that reference models should be included
+when assessing neural network models. Both a null model (KM curve) and a linear model (either Cox or
+a flexible baseline hazard model like CBLR) provide insight as to whether the neural network model is learning
+anything useful beyond linear predictors that were not accounted for.
 
@@
 CBNN outperforms all competitors in the complex
-simulation and two case studies, while maintaining competitive performance in a final case study,
-demonstrating its value in survival settings that may
-involve time-varying interactions and a complex baseline hazard. Once we
-perform case-base sampling and adjust for the sampling bias, we can use
+simulation and two case studies while maintaining competitive performance in a final case study.
+Once we perform case-base sampling and adjust for the sampling bias, we can use
 a sigmoid activation function to predict our hazard function. Our
-approach provides an alternative to the incorporation of censored individuals, treating
+approach provides an alternative to incorporating censored individuals, treating
 survival outcomes as binary ones. Forgoing the requirement
@@
 components in machine learning libraries (specifically, the add layer to
-adjust for sampling bias and the sigmoid activation function). Due to
+adjust for sampling bias and the sigmoid activation function) after case-base sampling. Due to
 the simplicity in its implementation and by extension user experience,
@@
 
-The pre-processed data for the SUPPORT case study can be found at
-\url{https://github.com/jaredleekatzman/DeepSurv/tree/master/experiments/data/support}.
-The pre-processed data for the METABRIC case study can be found at
-\url{https://github.com/jaredleekatzman/DeepSurv/tree/master/experiments/data/metabric}.
-The data is accessed using \url{https://github.com/havakv/pycox}. The
-code for this manuscript and its analyses can be found at
+The Prostate dataset is available as part of the \textbf{asaur} package in R.
+The MM and FLC datasets are available in the \textbf{survival} package in R.
+The code for this manuscript and its analyses can be found at
 \url{https://github.com/Jesse-Islam/cbnnManuscript}. The software
-package making CBNN easier to use can be found at
+package making CBNN easier to implement can be found at
 \url{https://github.com/Jesse-Islam/cbnn}.
@@
 M.T.
+wrote the majority of the casebase sampling section \ref{case-base-sampling}.
+M.T.
 and S.B.
 provided key insights into the core sampling technique (casebase).
+R.S.
+provided multiple edits to the manuscript structure and key insights.
 All authors discussed the results and contributed to the final manuscript.
-M.T.
-contributed heavily to the casebase sampling section \ref{case-base-sampling}.
 

### 2023-04-05 analyses/plotter.Rmd
updated text

@@
 g<-plot_grid(pt, legend,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))
-pdf(file="figures/megaPlot.pdf",width=16,height=9)
+pdf(file="../StatPaper/figures/megaPlot.pdf",width=16,height=9)
 g
@@
 
-pdf(file="figures/megaTable.pdf",width=14,height=13)
+pdf(file="../StatPaper/figures/megaTable.pdf",width=14,height=13)
 cowplot::plot_grid(cFlex,mFlex,fFlex,pFlex,ncol=1)
@@
 
-pdf(file="figures/winTable.pdf",width=12,height=6)
+pdf(file="../StatPaper/figures/winTable.pdf",width=12,height=6)
 cowplot::plot_grid(cwtab,mwtab,fwtab,pwtab,ncol=2)
@@
 
-```{r complexHazard}
+```{r complexHazard,eval=FALSE}
 
@@
 
-pdf(file="figures/ComplexPopTime.pdf",width=8,height=6)
+pdf(file="../StatPaper/figures/ComplexPopTime.pdf",width=8,height=6)
 a

### 2023-04-05 analyses/optimFlchain.Rmd
rancleanedcodeCheck

@@
 preData<-data.frame(age=as.numeric(flchain$age),
-                    #sampyr=as.numeric(flchain$sample.yr),
                     sex=as.factor(flchain$sex),
                     flc=as.numeric(flchain$kappa)+as.numeric(flchain$lambda),
-                    #batch=as.factor(flchain$flc.grp),
                     creatine=as.numeric(flchain$creatinine),
@@
 
-##saveRDS(res$test,"cvData/flchainTest.rds")
+saveRDS(res$test,"cvData/flchainTest.rds")
 
@@
 }
-##saveRDS(klist,"cvData/flchainCV.rds")
+saveRDS(klist,"cvData/flchainCV.rds")
 
@@
 alp=c(0,0.5,1)
-
-lr=c(0.01)
-drpt=c(0.01)
-layer1=c(50)
-layer2=c(10)
-nBatch=c(100)
-actv=c("relu","linear")#
-alp=c(0)
 
@@
 
-#saveRDS(winners,"cvData/flchainWins.rds")
+saveRDS(winners,"cvData/flchainWins.rds")
 
@@
   print(paste("current iter:",pos))
-  #saveRDS(BrierIPAList,'results/100flchainBrier.rds')
-  #saveRDS(cScore,'results/100flchaincidx.rds')
-  #saveRDS(AUClist,'results/100flchainAuc.rds')
+  saveRDS(BrierIPAList,'results/100flchainBrier.rds')
+  saveRDS(cScore,'results/100flchaincidx.rds')
+  saveRDS(AUClist,'results/100flchainAuc.rds')
   pos=pos+1

### 2023-04-05 analyses/optimMyeloma.Rmd
rancleanedcodeCheck

@@
 
-#saveRDS(res$test,"cvData/myelomaTest.rds")
+saveRDS(res$test,"cvData/myelomaTest.rds")
 
@@
 }
-#saveRDS(klist,"cvData/myelomaCV.rds")
+saveRDS(klist,"cvData/myelomaCV.rds")
 
@@
               dh=bestDH)
-#saveRDS(winners,"cvData/myelomaWins.rds")
+saveRDS(winners,"cvData/myelomaWins.rds")
 
@@
   print(paste("current iter:",pos))
-  #saveRDS(BrierIPAList,'results/100myelomaBrier.rds')
-  #saveRDS(cScore,'results/100myelomacidx.rds')
-  #saveRDS(AUClist,'results/100myelomaAuc.rds')
+  saveRDS(BrierIPAList,'results/100myelomaBrier.rds')
+  saveRDS(cScore,'results/100myelomacidx.rds')
+  saveRDS(AUClist,'results/100myelomaAuc.rds')
   pos=pos+1

### 2023-04-05 analyses/optimProstate.Rmd
rancleanedcodeCheck

@@
 
-#saveRDS(res$test,"cvData/prostateTest.rds")
+saveRDS(res$test,"cvData/prostateTest.rds")
 
@@
 }
-#saveRDS(klist,"cvData/prostateCV.rds")
+saveRDS(klist,"cvData/prostateCV.rds")
 
@@
 ```{r setFixedHyperparameters,eval=TRUE,hide=FALSE}
-
-klist<-readRDS("cvData/prostateCV.rds")
 
@@
 
-#saveRDS(winners,"cvData/prostateWins.rds")
+saveRDS(winners,"cvData/prostateWins.rds")
 
@@
   linModelTrain<-as.data.frame(rbind(train,val))
-  #valcb<-sampleCaseBase(data = as.data.frame(val),time="time",event="status",ratio=100)#rbind(train,val)
-  #valcb$offset<-0
-  #valcbtensor<-list(list(as.matrix(valcb[,-c(ncol(valcb)-1,ncol(valcb))]),as.matrix(valcb[,ncol(valcb),drop=F])),as.matrix(valcb[,ncol(valcb)-1,drop=F]))
+
   times<-seq(from=min(test$time),
@@
   print(paste("current iter:",pos))
-  #saveRDS(BrierIPAList,'results/100prostateBrier.rds')
-  #saveRDS(cScore,'results/100prostatecidx.rds')
-  #saveRDS(AUClist,'results/100prostateAuc.rds')
+  saveRDS(BrierIPAList,'results/100prostateBrier.rds')
+  saveRDS(cScore,'results/100prostatecidx.rds')
+  saveRDS(AUClist,'results/100prostateAuc.rds')
   pos=pos+1

### 2023-04-05 analyses/optimizeComplex.Rmd
rancleanedcodeCheck

@@
 res = split(data, g)
-#saveRDS(res$test,"cvData/ComplexSimTest.rds")
+saveRDS(res$test,"cvData/ComplexSimTest.rds")
 
@@
 }
-#saveRDS(klist,"cvData/ComplexSimCV.rds")
+saveRDS(klist,"cvData/ComplexSimCV.rds")
 
@@
               dh=bestDH)
-#saveRDS(winners,"cvData/complexOptimWins.rds")
+saveRDS(winners,"cvData/complexOptimWins.rds")
 ```
@@
   source_python('src/dsurv.py')
-  coxnnsurv=fitDeepSurv(train,fullTest,bsize,epochs=epo,valida=val,patience=patience,min_delta=min_delta,drpt=drpt,lay1=layer1,lay2=layer2,lr=lr,actv=actv)
+  coxnnsurv=fitDeepSurv(train,fullTest,bsize,epochs=epo,valida=val,
+                        patience=patience,min_delta=min_delta,drpt=drpt,
+                        lay1=layer1,lay2=layer2,lr=lr,actv=actv)
   coxnnsurv<-t(coxnnsurv)
@@
   print(paste("current iter:",pos))
-  #saveRDS(BrierIPAList,'results/100complexBrier.rds')
-  #saveRDS(cScore,'results/100complexcidx.rds')
-  #saveRDS(AUClist,'results/100complexAuc.rds')
+  saveRDS(BrierIPAList,'results/100complexBrier.rds')
+  saveRDS(cScore,'results/100complexcidx.rds')
+  saveRDS(AUClist,'results/100complexAuc.rds')
   pos=pos+1
@@
                                            dist=c('weibull','lognormal'),
-                                           #IPA=NA,
-                                           #IPAse=NA,
+
                                            valLoss=NA,
@@
     dist=paramTrials[i,]$dist
-    #=paramTrials[i,]$IPA
-    #=paramTrials[i,]$valLoss
 
@@
 
-  #paramTrials$IPA[i]<-mean(tempRes)
-  #paramTrials$IPAse[i]<-sd(tempRes)
   paramTrials$valLoss[i]<-mean(tempVal)
@@
   print(mean(runTime))
-  #print(as.character(paramTrials[i,]))
-  #rm(cbnn,fit,annPreds,covars_input,covars_output)
+
   gc()

### 2023-04-27 StatPaper/APA/wileyNJD-APA.tex
applied Rob's comments

@@
 
-\received{26 April 2023}
-\revised{6 June 2023}
-\accepted{6 June 2023}
+\received{XXXXXX}
+\revised{XXXXXX}
+\accepted{XXXXXX}
 
@@
 
-\title{Case-Base Neural Networks: survival analysis with time-varying,
+\title{Case-Base Neural Network: survival analysis with time-varying,
 higher-order interactions}
@@
 
-\author[3]{Robert Sladek}
+\author[1,3]{Robert Sladek}
 
@@
 
-\address[1]{\orgdiv{Department of Quantitative life sciences,} \orgname{McGill University}, \orgaddress{\state{Quebec}, \country{Canada}}}
+\address[1]{\orgdiv{Department of Quantitative Life Sciences,} \orgname{McGill University}, \orgaddress{\state{Quebec}, \country{Canada}}}
 
@@
 
-\presentaddress{McGill genome center: 740 Dr Penfield Ave, Montreal, Quebec H3A 0G1}
+\presentaddress{McGill Genome Center: 740 Dr Penfield Ave, Montreal, Quebec H3A }
 
 \abstract[Summary]{Neural network-based survival methods can model data-driven covariate interactions.
-While these methods have led to better predictive performance than regression-based approaches,
-they cannot model both time-varying interactions and complex baseline hazards. To address this, we
-propose Case-Base Neural Networks (CBNN) as a new approach that combines the case-base sampling
-framework with flexible architectures. Our method naturally accounts for censoring and does not require
-method specific hyperparameters. Using a novel sampling scheme and data augmentation, we incorporate
-time directly into a feed-forward neural network. CBNN predicts the probability of an event occurring at a
-given moment to estimate the hazard function. We compare the performance of CBNN to survival methods
-based on regression and neural networks in a simulation and three real data applications. We report two time-dependent
-metrics for each model. For a complex simulation, which highlights the ability of CBNN to model both a complex
-baseline hazard and time-varying interactions and two real data applications, CBNN outperforms all competitors
-on a simulation with a complex baseline hazard and time-varying interaction and two case studies. A final case
-study shows competitive performance. We highlight the benefit of combining case-base sampling with deep
-learning to provide a simple and flexible modeling framework for data-driven, time-varying interaction modeling
-of survival outcomes. An R package is available at https://github.com/Jesse-Islam/cbnn.}
+While these methods have led to better predictive performance than regression-based approaches, they cannot model time-varying interactions and complex baseline hazards simultaneously.
+To address this, we propose Case-Base Neural Network (CBNN) as a new approach that combines the case-base sampling framework with flexible neural network architectures.
+Using a novel sampling scheme and data augmentation to naturally account for censoring, we construct a feed-forward neural network that may take time as an input.
+CBNN predicts the probability of an event occurring at a given moment to estimate the hazard function.
+We compare the performance of CBNN to neural network and regression-based survival methods in a simulation and three case studies using two time-dependent metrics.
+First, we examine performance on a simulation involving a complex baseline hazard and time-varying interactions to assess all methods, with CBNN outperforming competitors.
+Then, we apply all methods to two real data applications where CBNN outperforms the competing models as well.
+In a third case study, CBNN shows competitive performance.
+Our results highlight the benefit of combining case-base sampling with deep learning to provide a simple and flexible modeling framework for data-driven, time-varying interaction modeling of single event survival outcomes.
+An R package is available at https://github.com/Jesse-Islam/cbnn.}
 
@@
 
-Smooth-in-time accelerated failure time (AFT) models can estimate absolute
-risks by modeling the hazard directly through a user-specified baseline hazard
-distribution \citep{kleinbaum2012survival}. Analysts use Cox proportional hazards
-models more often than AFT models \citep{hanley2009}. This preference causes
-analyses to be based on hazard ratios and relative risks rather than on survival
-curves and absolute risks \citep{hanley2009}. With AFT models, it may be difficult
-to identify an appropriate baseline hazard for common diseases with many
-interacting risk factors \citep{royston2002flexible}. In studies where the disease
-pathogenesis may change with age, we may see non-proportional hazards
-\citep{coradini2000time}. For example, previous studies of breast cancer incidence
-have discovered time-varying interactions with covariates of interest, such as tumor
-size \citep{coradini2000time}. One approach to provide flexibility in the baseline hazard
-involves using the basis of splines on time in the model \citep{royston2002flexible}.
-However, regression-based models require prior knowledge of potential time-varying
-interactions and their quantitative effects.
+A common assumption we use in survival analysis is that the risk of the event of interest does not vary with
+time (i.e. proportional hazards) \citep{hanley2009}. This simplifying assumption results in using Cox proportional
+hazards models more often than Smooth-in-time, accelerated failure time (AFT) models \citep{hanley2009}. This
+preference causes analyses to be based on hazard ratios and relative risks rather than on survival curves and
+absolute risks \citep{hanley2009}. The proportional hazards assumption in a Cox model may be incorrect in studies
+where the disease pathogenesis may change over time \citep{coradini2000time}. For example, studies of breast cancer
+incidence have shown that the risk of developing a tumor involves a time-varying interaction with tumor size
+\citep{coradini2000time}. Time-varying covariates can be incorporated easily into AFT models \citep{royston2002flexible}.
+To provide flexibility in the baseline hazard, we can use a basis of splines on time as model inputs
+\citep{royston2002flexible}. However, these regression-based models require prior knowledge of potential
+time-varying interactions and their quantitative effects.
 
-Neural networks provide a data-driven approach to approximating interaction terms.
-For example, DeepSurv is a neural network-based proportional hazards model that
-implements the Cox partial log-likelihood as a custom loss function \citep{katzman2018DeepSurv}.
-This results in a stepwise absolute risk curve that cannot accommodate time-varying interactions.
-Compared to Cox regression, DeepSurv shows better performance on a real dataset
-\citep{katzman2018DeepSurv}. Previous work suggests we may change the loss function
-to address non-proportional hazards \citep{faraggi1995neural}.
-
-DeepHit assumes an inverse Gaussian distribution as the baseline hazard \citep{lee2018DeepHit}.
-It specifies each follow-up time time of interest in the model and directly estimates survival curves, rather
-than deriving a hazard function \citep{lee2018DeepHit}. DeepHit outperformed DeepSurv
-\citep{lee2018DeepHit}.
-
-We note that these alternative neural network approaches require custom loss functions
-\citep{katzman2018DeepSurv} \citep{lee2018DeepHit}. DeepHit introduces a hyperparameter
-weighing its two loss functions (negative log-likelihood and ranking losses) \citep{lee2018DeepHit}.
-Regression-based approaches require prior specification of all interaction terms, which makes it
-challenging to model covariate effects that change with time. The current neural network models
+Neural networks provide a data-driven approach to approximating interaction terms in models. For example.
+DeepSurv is a neural network-based proportional hazards model that implements the Cox partial log-likelihood
+as a custom loss function \citep{katzman2018DeepSurv}. This results in a stepwise absolute risk curve that cannot
+accommodate time-varying interactions. Compared to Cox regression, DeepSurv shows better performance on a
+real dataset \citep{katzman2018DeepSurv}. Previous work suggests we may change the loss function to address
+non-proportional hazards \citep{faraggi1995neural}. In contrast, DeepHit assumes an inverse Gaussian distribution
+as the baseline hazard \citep{lee2018DeepHit} and specifies each follow-up time of interest in the model. DeepHit
+directly estimates survival curves, rather than deriving a hazard function \citep{lee2018DeepHit}. On real datasets,
+DeepHit outperformed DeepSurv \citep{lee2018DeepHit}. We note that DeepSurv and DeepHit both use custom loss
+functions \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}. In addition, DeepHit introduces a hyperparameter
+weighing its two loss functions (negative log-likelihood and ranking losses) \citep{lee2018DeepHit}. Compared
+to these neural network methods, regression-based approaches require prior specification of all interaction terms,
+which makes it challenging to model covariate effects that change with time. The current neural network models
 provide flexibility at the cost of clarity, while regression models provide clarity at the cost of flexibility.
 
-In this article, we propose Case-Base Neural Networks (CBNN) as a method that models time-varying
-interactions and a flexible baseline hazard. Our approach to modeling the full hazard uses case-base
-sampling \citep{hanley2009}. This sampling technique allows probabilistic models to predict survival
-outcomes. After case-base sampling, we can easily implement the model in any language using
-common neural network components. As part of the case-base framework, we use transformations
-of time as a feature (covariate) to specify different baseline hazards. For example, by including
-splines of time as covariates, we can approximate the Royston-Parmar flexible baseline hazard model
-\citep{royston2002flexible}\citep{hanley2009}. However, this still requires explicit use of time-varying
-interactions. CBNN can model both without extra tuning parameters.
+We propose Case-Base Neural Networks (CBNN) as a method for single event survival analysis that models
+time-varying interactions and a flexible baseline hazard. CBNN use the case-base sampling technique, which
+allows probabilistic models to predict survival outcomes \citep{hanley2009}. After case-base sampling, we
+implement a model using common neural network components that uses transformations of time as a feature
+(covariate) to specify different baseline hazards. For example, by including splines of time as covariates,
+case-base with logistic regression (CBLR) can approximate the Royston-Parmar flexible baseline hazard
+\citep{royston2002flexible} \citep{hanley2009}. However, this still requires explicit specification of
+time-varying interactions. CBNN can model both without extra tuning parameters.
 
-We describe how case-base sampling and neural networks combine
-in Section \ref{methods}, along with our hyperparameter selection procedure, metrics and software implementation.
-We explore the performance of CBNN, DeepSurv, DeepHit, Cox regression and case-base using logistic
-regression (CBLR) on simulated data in Section \ref{sims}. Section \ref{casestudies} describes the
-performance of the same methods in three case studies. Section \ref{discussion} explores the
-implications of our results and contextualizes them within neural network survival analysis in a
-single event setting.
+In this paper, we describe the case-base sampling method and compare its properties to neural
+network models in Section \ref{methods}, along with our hyperparameter selection procedure,
+metrics and software implementation. We also explore the performance of CBNN, DeepSurv,
+DeepHit, Cox regression and CBLR on simulated data in Section \ref{sims}. Section \ref{casestudies}
+describes the performance of the same methods in three case studies. Section \ref{discussion} explores
+the implications of our results and contextualizes them within neural network survival analysis in a single
+event setting.
 
@@
 
-In this section, we define case-base sampling, which converts the total
-survival time into discrete person-specific moments (person-moments).
-Next, we detail how neural networks can be used within this framework,
-explicitly incorporating time as a feature while adjusting for the
-sampling bias. Then, we describe the metrics and hyperparameter selection procedure.
-Finally, we report on the software versions used. An R
-package is available for use at
-\url{https://github.com/Jesse-Islam/cbnn}. The entire code base to
-reproduce the figures and empirical results in this paper is available
-at \url{https://github.com/Jesse-Islam/cbnnManuscript}.
+Case-base sampling divides total survival time for each individual into person-specific moments
+(person-moments), treating each person-moment as a sample. In this section, we detail how neural
+networks explicitly incorporating time as a feature while adjusting for the sampling bias in case-base
+sampled data; and describe the metrics and hyperparameter selection procedure used to compare
+CBNN with other software packages. An R package to use CBNN is available at
+\url{https://github.com/Jesse-Islam/cbnn}. The entire code base to reproduce the figures
+and empirical results in this paper is available at \url{https://github.com/Jesse-Islam/cbnnManuscript}.
 
@@
 
-Case-base sampling is an alternative framework for survival analysis
-\citep{hanley2009}. Without this preliminary step, we cannot model survival
-outcomes with probabilistic methods. In case-base sampling, we sample from the continuous
-survival time of each person in our dataset to create a \emph{base
-series} of \emph{person-moments}. This \emph{base series} complements
-the \emph{case series}, which contains all person-moments at which the
-event of interest occurs.
+Case-base sampling is an alternative framework for survival analysis \citep{hanley2009},
+which converts the total survival time into discrete person-moments. Without this preliminary step,
+we cannot model survival outcomes with probabilistic methods. In case-base sampling, we sample
+from the continuous survival time of each person in our dataset to create a \emph{base series} of
+\emph{person-moments}. This \emph{base series} complements the \emph{case series}, which contains
+all person-moments at which the event of interest occurs.
 
-For each person-moment sampled, let \(X_i\) be the corresponding
-covariate profile \(\left(x_{i1},x_{i2},...,x_{ip} \right)\), \(T_i\) be
-the time of the person-moment and \(Y_i\) be the indicator variable for
-whether the event of interest occurred at time \(T_i\). We estimate the
-hazard function \(h(t \mid X_i)\) using the sampled person-moments.
-Recall that \(h(t \mid X_i)\) is the instantaneous potential of
-experiencing the event at time \(t\) for a given set of covariates
-\(X_i\), assuming \(T_i \geq t\).
+For each person-moment sampled, let \(X_i\) be the corresponding covariate profile
+\(\left(x_{i1},x_{i2},...,x_{ip} \right)\), \(T_i\) be the time of the person-moment and \(Y_i\) be the indicator
+variable for whether the event of interest occurred at time \(T_i\). We estimate the hazard function
+\(h(t \mid X_i)\) using the sampled person-moments. Recall that \(h(t \mid X_i)\) is the instantaneous
+potential of experiencing the event at time \(t\) for a given set of covariates \(X_i\), assuming \(T_i \geq t\).
 
-Now, let \(b\) be the (user-defined) size of the \emph{base series} and
-let \(B\) be the sum of all follow-up times for the individuals in the
-study. let $c$ be the number of events in the \emph{case series}.
-A reasonable concern is determining how large $b$ should be relative to $c$.
-The size of \(b\) influences how much information is lost in the sampling process
-\citep{hanley2009}. Using our notation, Hanley $\&$ Miettinen (2009)
-adapt a quote from Mantel (1973): "By the reasoning that $\frac{cb}{c+b} [=(\frac{1}{c}+\frac{1}{b}]$
-measures the relative information in a comparison of two averages based on
-sample sizes of c and b respectively, we might expect by analogy, which would
-of course not be exact in the present case, that this approach would result in
-only a moderate loss of information. (The practicing statistician is generally
-aware of this kind of thing. There is little to be gained by letting the size of
-one series, b, become arbitrarily large if the size of the other series, c, must
-remain fixed)" \citep{hanley2009} \citep{mantel1}. If $b=100c$, then we expect learned
-weights to be at most one percent higher than if the entire study base $B$ was used,
-as they are proportional to $\frac{1}{c}+\frac{1}{100c}$ rather than $\frac{1}{c}+\frac{1}{\infty c}$ \citep{hanley2009}.
+Now, let \(b\) be the (user-defined) size of the \emph{base series} and let \(B\) be the sum of all follow-up
+times for the individuals in the study. Let $c$ be the number of events in the \emph{case series}. A
+reasonable concern is determining how large $b$ should be relative to $c$, since the size of \(b\)
+influences how much information is lost in the sampling process \citep{hanley2009}. Using our notation,
+Hanley $\&$ Miettinen (2009) adapt a quote from Mantel (1973): "By the reasoning that
+$\frac{cb}{c+b}=\left[\left( \frac{1}{c}+\frac{1}{b}\right)]\right]$ measures the relative information in a comparison of two
+averages based on sample sizes of c and b respectively, we might expect by analogy, which would
+of course not be exact in the present case, that this approach would result in only a moderate loss of
+information. (The practicing statistician is generally aware of this kind of thing. There is little to be gained
+by letting the size of one series, b, become arbitrarily large if the size of the other series, c, must remain
+fixed)" \citep{hanley2009} \citep{mantel1}. If $b=100c$, then we expect learned weights to be at most one
+percent higher than if the entire study base $B$ was used, as they are proportional to
+$\frac{1}{c}+\frac{1}{100c}$ rather than $\frac{1}{c}+\frac{1}{\infty c}$ \citep{hanley2009}.
 
-If we sample the \emph{base series} uniformly across the study
-base, then the hazard function of the sampling process is equal to
-\(b/B\). Therefore, we have the following equality \cite{saarela2015}
-\footnote{We are abusing notation here, conflating hazards with probabilities. For a rigorous treatment, see Saarela \& Hanley (2015) section 3 \cite{saarela2015} .}:
+If we sample the \emph{base series} uniformly across the study base, then the hazard function
+of the sampling process is equal to \(b/B\). Therefore, we have the following equality \cite{saarela2015}
+\footnote{For a rigorous treatment, see Saarela \& Hanley (2015) section 3 \cite{saarela2015} .}:
 \begin{align}\label{eqn:main}
@@
 
-To estimate the correct hazard function, we adjust for the bias
-introduced when sampling a fraction of the study base (\(\frac{b}{B}\)) by adding
-the term \(\log\left(\frac{B}{b} \right)\) to offset \(\log\left(\frac{b}{B} \right)\)
-during the fitting process. Next, we propose using neural networks to model the
-odds.
+To estimate the correct hazard function, we adjust for the bias introduced when sampling a fraction of the
+study base (\(\frac{b}{B}\)) by adding the term \(\log\left(\frac{B}{b} \right)\) to offset \(\log\left(\frac{b}{B} \right)\)
+during the fitting process.
 
@@
 
-After case-base sampling, we pass all features, including time, into any
-user-defined feed-forward component, to which an offset term is added,
-then passed through a sigmoid activation function (Figure
-\ref{fig:NNarch}). We are interested in predicting the odds. This makes the sigmoid
-activation function ideal as it is the inverse of the odds. which we can use to calculate the hazard. The general form for the
-neural network using CBNN is:
+After case-base sampling, we pass all features, including time, into any user-defined feed-forward
+component. Then, we add an offset term and pass the output through a sigmoid activation function
+(Figure \ref{fig:NNarch}). Since we are interested in predicting the odds of an event occurring, the sigmoid
+activation function is ideal as it is the inverse of the odds and can be used to calculate the hazard. The
+general form for the neural network using CBNN is:
 
@@
 
-where \(T\) is a random variable representing the event time, \(X\) is
-the random variable for a covariate profile, \(f_{\theta}(X, T)\)
-represents any feed-forward neural network architecture,
-\(\log\left(\frac{B}{b}\right)\) is the offset term for adjust for the bias
-(\(\log\left(\frac{b}{B}\right)\)) set by case-base sampling,
-\(\theta\) is the set of parameters learned by the neural network and
-\(\mathrm{sigmoid}(x)=\frac{1}{1+e^{-x}}\). By approximating a
-higher-order polynomial of time using a neural network, the baseline
-hazard specification is now data-driven, where user-defined
-hyperparameters such as regularization, number of layers and nodes
-control the flexibility of the hazard function. We provide a detailed
-description of the choices we made in the next sub-section.
+where \(T\) is a random variable representing the event time, \(X\) is the random variable for a covariate
+profile, \(f_{\theta}(X, T)\) represents any feed-forward neural network architecture, \(\log\left(\frac{B}{b}\right)\)
+is the offset term for adjust for the bias (\(\log\left(\frac{b}{B}\right)\)) set by case-base sampling, \(\theta\) is the
+set of parameters learned by the neural network and \(\mathrm{sigmoid}(x)=\frac{1}{1+e^{-x}}\). By approximating
+a higher-order polynomial of time using a neural network, the baseline hazard specification is now data-driven,
+where user-defined hyperparameters such as regularization, number of layers and nodes control the flexibility
+of the hazard function.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/nnarch2}}
+{\centering \includegraphics[width=1\linewidth]{../figures/Figure1.pdf}}
 
-\caption{Steps involved in CBNN from case-base sampling to the model framework we use for training.
-The first step is case-base sampling, completed before training begins.
-Next, we pass this sampled data through a feed-forward neural network.
-We add the offset and pass that through a sigmoid activation function, whose output is a probability.
-Once the neural network model completes its training, we can convert the probability output to a hazard, using it for our survival outcomes of interest.}\label{fig:NNarch}
+\caption{Methodological steps involved in CBNN. The first step, case-base sampling, is completed before training begins.
+We pass this sampled data through a feed-forward neural network and add an offset to adjust for the bias inherent in
+case-base sampling, then apply a sigmoid activation function to create a probability. Once the neural network model
+completes its training, we can convert the probability output to a hazard for the survival outcomes of interest.}\label{fig:NNarch}
 \end{figure}
@@
 \end{align*} where \(\hat{f}_{\theta}(x_{i}, t_{i})\) is our estimate
-for a given covariate profile and time, \(y_{i}\) is our target value
-specifying whether an event occurred and \(N\) represents the number of
-individuals in our training set.
+for a given covariate profile and time, \(y_{i}\) is our target value specifying whether an event occurred and \(N\) represents
+the number of individuals in our training set.
 
-Backpropagation with an appropriate minimization algorithm (e.g.~Adam,
-RMSPropagation, stochastic gradient descent) is used to optimize the
-parameters in the model \citep{gulli2017}. For our analysis, we use Adam
-as implemented in Keras \citep{gulli2017}. Note that the size of the
-\emph{case series} is fixed as the number of events, but we can make the
-\emph{base series} as large as we want. A ratio of 100:1 \emph{base
-series} to \emph{case series} is sufficient \citep{hanley2009}. We pass
-our feed-forward neural network through a sigmoid activation function
-(Figure \ref{fig:NNarch}). Finally, we can convert this model output to
-a hazard. To use CBNN for predictions, we manually set the
-offset term \(\log\left(\frac{B}{b} \right)\) to 0 in the new data as we already account for the bias during the
-fitting process.
+Backpropagation with an appropriate minimization algorithm (e.g.~Adam, RMSPropagation, stochastic gradient
+descent) is used to optimize the parameters in the model \citep{gulli2017}. For our analysis, we use Adam as implemented
+in Keras \citep{gulli2017}. While the size of the \emph{case series} is fixed as the number of events; the size of the
+\emph{base series} is not restricted. We use a ratio of 100:1 \emph{base series} to \emph{case series} \citep{hanley2009}.
+We pass our feed-forward neural network through a sigmoid activation function (Figure \ref{fig:NNarch}). Finally, we can
+convert this model output to a hazard. To use CBNN for predictions, we manually set the offset term \(\log\left(\frac{B}{b} \right)\)
+to 0 in the new data as we already account for the sampling bias during the fitting process.
 
@@
 
-To choose our method-specific hyperparameters, we use the Integrated
-Brier Score (IBS) \citep{graf1999}. We use two metrics to assess the
-performance of the different methods of interest on testing data: 1) Index of Prediction Accuracy (IPA)
-\citep{kattan2018index} and 2) Inverse probability censoring weights-adjusted time dependent
-area under the receiver operating characteristic curve ($AUC_{IPCW}$) \citep{auc}, which we 
-define below. We achieve a summarized assessment of performance for each model with the
-IBS. The IPA is a metric for both discrimination and calibration. The $AUC_{IPCW}$
-provides a metric for discrimination only. Together, we can get a clearer picture about how
-each model performs in each study and why.
+To choose our method-specific hyperparameters, we used the Integrated Brier Score (IBS) \citep{graf1999}, which is based on the
+Brier Score (BS) and provides a summarized assessment of performance for each model. We assess the performance of models on
+a held out dataset using two metrics: 1) the Index of Prediction Accuracy (IPA) \citep{kattan2018index}; and 2) the Inverse probability
+censoring weights-adjusted time dependent area under the receiver operating characteristic curve ($AUC_{IPCW}$) \citep{auc}. The
+IPA score is a metric for both discrimination and calibration, while the $AUC_{IPCW}$ provides a metric for discrimination only.
+
+\hypertarget{bs}{%
+\subsubsection{Brier Score (BS)}\label{bs}}
+The BS \citep{graf1999} is defined as \begin{align}\label{eqn:bs}
+BS(t)=\frac{1}{N}\sum^{N}_{i=1}\left(\frac{\left(1 - \widehat{F}(t \mid X_{i})\right)^{2}\cdot I(T_{i}\leq t,\delta_{i}=1)}{\widehat{G}(T_{i})} + \frac{\widehat{F}(t\mid X_{i})^{2}\cdot I(T_{i}>t)}{\widehat{G}(t)}\right),
+\end{align} where \(\delta_{i}=1\) shows individuals who have experienced the event, \(N\) represents the number of samples
+in our dataset over which we calculate \(BS(t)\), \(\widehat{G}(t)=P[c>t]\) is a non-parametric estimate of the censoring distribution,
+\(c\) is censoring time and \(T_{i}\) is an individual's survival or censoring time. The BS provides a score that accounts for the
+information loss because of censoring. Once we fix our \(t\) of interest, the individuals in the dataset can be divided into three
+groups. Individuals who experienced the event before \(t\) are present in the first term of the equation. The second term of the
+equation includes individuals who experience the event or are censored after \(t\). Those censored before \(t\) (the remaining
+individuals) are accounted for by the IPCW adjustment (\(G(\cdot)\)) present on both terms.
+
+\hypertarget{integrated-brier-score-ibs}{%
+\subsubsection{Integrated Brier Score (IBS)}\label{integrated-brier-score-ibs}}
+
+The Integrated Brier Score (IBS) is a function of the BS \ref{eqn:bs} \citep{graf1999}, which is defined as \begin{align*}
+IBS(t)=\int_{t}^{t_{max}}BS(t)dw(t),
+\end{align*} where $w(t)=\frac{t}{t_{max}}$. As the IBS is calculated for a range of follow-up-times,
+it is a useful metric to track overall performance across all of follow-up time during cross-validation. We use the
+IPA score to get an estimate of performance at each follow-up time for our studies.
 
@@
 
-The IPA is a function of the Brier score (\(BS(t)\)) \citep{graf1999},
-which is defined as \begin{align}\label{eqn:bs}
-BS(t)=\frac{1}{N}\sum^{N}_{i=1}\left(\frac{\left(1 - \widehat{F}(t \mid X_{i})\right)^{2}\cdot I(T_{i}\leq t,\delta_{i}=1)}{\widehat{G}(T_{i})} + \frac{\widehat{F}(t\mid X_{i})^{2}\cdot I(T_{i}>t)}{\widehat{G}(t)}\right),
-\end{align} where \(\delta_{i}=1\) shows individuals who have
-experienced the event, \(N\) represents the number of samples in our
-dataset over which we calculate \(BS(t)\), \(\widehat{G}(t)=P[c>t]\) is
-a non-parametric estimate of the censoring distribution, \(c\) is
-censoring time and \(T_{i}\) is an individual's survival or censoring
-time. The BS provides a score that accounts for the information
-loss because of censoring. There are three categories of individuals
-that may appear within the dataset once we fix our \(t\) of interest.
-Individuals who experienced the event before \(t\) are present in the
-first term of the equation. The second term of the equation includes
-individuals who experience the event or are censored after \(t\). Those
-censored before \(t\) are the third category of people. The inverse
-probability censoring weights (IPCW) adjustment (\(G(\cdot)\)) is to
-account for these category three individuals whose information is
-missing.
-
-The IPA as a function of time is given by \begin{align}
+The IPA as a function of time, based on the BS. The IPA score is given by \begin{align}
 \textrm{IPA}(t) &= 1-\frac{BS_{model}(t)}{BS_{null}(t)}, \nonumber
-\end{align} where \(BS_{model}(t)\) represents the BS over time
-\(t\) for the model of interest and \(BS_{null}(t)\) represents the
-BS if we use an unadjusted Kaplan-Meier (KM) curve as the
-prediction for all observations \citep{kattan2018index}. Note that IPA
-has an upper bound of one, where positive values show an increase in
-performance over the null model and negative values show that the null
-model performs better. These scores show how performance changes over
-follow-up time. As an extension of BS, IPA models for both calibration
-and discrimination\citep{graf1999} \citep{kattan2018index}.
-
-\hypertarget{integrated-brier-score-ibs}{%
-\subsubsection{Integrated Brier Score (IBS)}\label{integrated-brier-score-ibs}}
-
-The Integrated Brier Score (IBS) is a function of the BS \ref{eqn:bs} as well \citep{graf1999},
-which is defined as \begin{align*}
-IBS(t)=\int_{t}^{t_{max}}BS(t)dw(t),
-\end{align*} where $w(t)=\frac{t}{t_{max}}$. As this is a score for a range of follow-up-times,
-it is a useful metric to track during cross-validation. The performance at a specific time is lost,
-and is why we opt for the IPA in assessing performance on a test set.
+\end{align} where \(BS_{model}(t)\) represents the BS over time \(t\) for the model of interest and \(BS_{null}(t)\) represents
+the BS if we use an unadjusted Kaplan-Meier (KM) curve as the prediction for all observations \citep{kattan2018index}. The
+IPA score has an upper bound of one, where positive values show an increase in performance over the null model and
+negative values show that the null model performs better. An extension of BS, the IPA scores assess both model calibration
+and discrimination and show how performance changes over follow-up time \citep{graf1999} \citep{kattan2018index}.
 
@@
 
-The IPCW-adjusted AUC ($AUC_{IPCW}$) is a time-dependent metric that considers censoring \citep{auc}. 
-For a given follow-up time of interest,
+The IPCW-adjusted AUC ($AUC_{IPCW}$) is a time-dependent metric that considers censoring \citep{auc}. For a given follow-up time of interest,
  \begin{align*}
@@
 W_{i}(t)=\frac{I(T_{i}\leq t,\delta_{i}=1)}{\widehat{G}(T_{i})} +\frac{I(T_{i}>t)}{\widehat{G}(t)}.
-\end{align*} The $AUC_{IPCW}$ measures discrimination \citep{auc}.
-By taking both IPA and $AUC_{IPCW}$ together, we can better understand whether a model is better
-in terms of calibration or discrimination.
+\end{align*} The $AUC_{IPCW}$ measures discrimination \citep{auc}. By examining both IPA and $AUC_{IPCW}$,
+we can better understand whether a model performs better in terms of calibration or discrimination.
 
@@
 
-Neural networks are flexible when defining the architecture and optimization parameters.
-These hyperparameter choices can affect the estimated parameters and are not necessarily
-the same for each model. We apply the same procedure to each study in this paper. First,
-we fix a test set with 15\% of the data. We keep aside this data during hyperparameter
-selection and we fit a model for each method in each study. To determine the best
-hyperparameters for each neural network method, we use three-fold cross-validated grid
-search \citep{gulli2017} on the remaining data (85\% training, 15\% validation) for the
-following range of hyperparameters:
-
+Neural networks require external parameters that constrain the learning process (hyperparameters). We apply the following hyperparameter
+optimization procedure to each method for each study. First, we fix a test set with 15\% of the data which we keep aside during hyperparameter selection.
+To determine the best hyperparameters for each neural network method, we use a three-fold cross-validated grid search \citep{gulli2017} on the remaining
+data (85\% training, 15\% validation) for the following range of hyperparameters:
 \[
 Grid\ search: \begin{cases}
-Learning\ rate \sim \{0.001, 0.01\} \\ %technically a control and treatement group
+Learning\ rate \sim \{0.001, 0.01\} \\ 
 Dropout \sim \{0.01,0.05, 0.1\} \\
@@
 Activation\ function \sim \{ReLU, Linear\}\\
-\alpha \sim \{0, 0.5, 1\}.
+\alpha \sim \{0, 0.5, 1\}. \ (DeepHit \ only)
 \end{cases}
@@
 \begin{table}
-\caption{Four tables showing the chosen hyperparameters along with the IBS for each neural network model in the complex simulation (A), multiple myeloma (MM) case study (B), free light chain (FLC) case study (C) and prostate cancer (Prostate) case study (D).}
+\caption{Hyperparameters selected after three-fold cross-validated grid search along with the average IBS for each neural network model in the
+complex simulation (A); multiple myeloma (MM) case study (B); free light chain (FLC) case study (C); and prostate cancer (Prostate) case study (D).}
 \label{tab:wins}
-\begin{center}\includegraphics[width=1\linewidth]{../figures/winTable} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{../figures/Table1.pdf} \end{center}
 \end{table}
 
-DeepHit uses $\alpha$ as a specific hyperparameter, while DeepSurv and CBNN do not.
-We track IBS on the validation set for each hyperparameter combination,
-choosing the combination with the lowest score for each method.
-We present the chosen hyperparameters in Table 1.
+DeepHit uses $\alpha$ as a hyperparameter, while DeepSurv and CBNN do not. We track IBS on the validation set for each hyperparameter combination,
+choosing the combination with the lowest score for each method (Table \ref{tab:wins}).
 
@@
 
-R \citep{Rsoft} and python \citep{py} are used to evaluate methods from
-both languages. We fit the Cox model using the \textbf{survival} package
-\citep{survpkg}, the CBLR model using the \textbf{casebase} package
-\citep{cbpkg}, DeepSurv and DeepHit using \textbf{pycox}
-\citep{lee2018DeepHit}. We made the components
-of CBNN using the \textbf{casebase} package for the sampling step and
-the \textbf{keras} \citep{keras} package for our neural network
-architecture. The \textbf{simsurv} package \citep{simsurv} is used for
-our simulation studies, while \textbf{flexsurv} \citep{flexsurv} is used
-to fit a flexible baseline hazard using splines for our complex
-simulation. The \textbf{riskRegression}
-package \citep{riskRegression} is used to get the Index of Prediction
-Accuracy (IPA metric) and AUC. Both metrics are described in detail in the
-following section. We modify the \textbf{riskRegression} package to be
-used with any user supplied risk function \(F\). To ensure that both R
-and Python-based models are running in unison on the same data through
-our simulations and bootstrap, we use the \textbf{reticulate} package
-\citep{reticulate}.
+R \citep{Rsoft} and Python \citep{py} are used to evaluate methods from both languages. We fit the Cox model using the \textbf{survival} package
+\citep{survpkg} and the CBLR model using the \textbf{casebase} package \citep{cbpkg}. Both DeepSurv and DeepHit are fit using \textbf{pyCox}
+\citep{lee2018DeepHit}. We made the components of CBNN using the \textbf{casebase} package \citep{cbpkg} for the sampling step and the
+\textbf{keras} \citep{keras} package for our neural network architecture. We use the \textbf{simsurv} package \citep{simsurv} for our simulation studies
+and \textbf{flexsurv} \citep{flexsurv} to fit a flexible baseline hazard using splines for our complex simulation. The \textbf{riskRegression} package
+\citep{riskRegression} is used to get the IPA and $AUC_{IPCW}$. We modify the \textbf{riskRegression} package to be used with any user supplied
+risk function \(F\). We use the \textbf{reticulate} package \citep{reticulate} to run both R based methods and python based methods on the same seed.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/megaPlot}
+{\centering \includegraphics[width=1\linewidth]{../figures/Figure2.pdf}
 
@@
 
-\caption{Summarizes the complex simulation (A, E), multiple myeloma (MM) case study (B, F), free light chain (FLC) case study (C, G) and prostate cancer (Prostate) case study (D, H).
-The first row shows the IPA for each model in each study over follow-up time.
-Negative values mean our model performs worse than the null model and positive values mean the model performs better.
-The second row shows the $AUC_{IPCW}$ for each model in each study over follow-up time.
-Each model-specific metric in each study shows a 95\% confidence interval over 100 iterations.
-The models of interest are case-base with logistic regression (CBLR), Case-Base Neural Networks (CBNN), Cox, DeepHit and DeepSurv.
-Both CBLR and Cox have near identical performance, resulting in curves that are overlapped.
-The Kaplan-Meier (KM) model serves as a baseline, predicting the average curve for all individuals.
-The Optimal (a CBLR model with the exact interaction terms and baseline hazard specified) model shows the best performance we can expect on the simulated data.
-}\label{fig:megaPlot}
+\caption{Performance of each model in the complex simulation (A, E), multiple myeloma (MM) case study (B, F), free light chain (FLC) case study (C, G)
+and prostate cancer (Prostate) case study (D, H). The first row shows the IPA for each model in each study over follow-up time. Negative values mean
+the model performs worse than the null model and positive values mean the model performs better. The second row shows the $AUC_{IPCW}$ for
+each model in each study over follow-up time, where higher is better. Each model-specific metric in each study shows a 95\% confidence interval over 100 iterations. Metrics
+are shown for six models: Case-Base with Logistic Regression (CBLR), Case-Base Neural Networks (CBNN), Cox Proportional Hazard (Cox), DeepHit
+and DeepSurv. The Kaplan-Meier (KM) model serves as a baseline, predicting the average curve for all individuals. CBLR and Cox have near identical
+performance, resulting in curves that overlap. The Optimal (a CBLR model with the exact interaction terms and baseline hazard specified) model shows
+the best performance we can expect on the simulated data.}\label{fig:megaPlot}
 \end{figure}
@@
 
-In this section, we simulate data to evaluate the performance of
-CBNN and compare our approach with existing regression-based (Cox, CBLR)
-and neural network-based (DeepHit, DeepSurv) methods. We specify a
-linear combination of each covariate as the linear predictor in
-regression-based approaches (Cox, CBLR), which contrasts with neural
-network approaches that allow for non-linear interactions. We simulate
-data under a complex baseline hazard with
-time-varying interactions, each with 10\% random censoring. For both
-settings, we simulate three covariates for 5000 individuals:
+We simulate data to evaluate the performance of CBNN in comparison with existing regression (Cox, CBLR) and neural network (DeepHit, DeepSurv)
+methods. We specify a linear combination of each covariate as the linear predictor in the regression-based methods (Cox, CBLR), which contrasts
+with neural network approaches that allow for non-linear interactions. The simulated data are based on a complex baseline hazard with time-varying
+interactions, each with 10\% random censoring. We simulate three covariates for 5000 individuals:
 
 \[
-z_{1} \sim \textrm{Bernoulli}(0.5) \qquad \qquad %technically a control and treatement group
+z_{1} \sim \textrm{Bernoulli}(0.5) \qquad \qquad %technically a control and treatment group
 z_{2} \sim \begin{cases}
@@
 \]
-Besides the methods mentioned above, we include the Optimal model in our comparisons using CBLR.
-We include the exact functional form of the covariates in a CBLR model (referred to as Optimal for simplicity).
-We conduct 100 bootstrap re-samples on the training data to get confidence intervals. We kept 15\% of the
-data for testing before hyperparameter selection. 15\% of the remaining data is used for validation and the rest is reserved
-for training. We predict risk functions for individuals in the test set, which are used to calculate our $AUC_{IPCW}$ and IPA.
+In addition to the methods described above, we include the exact functional form of the covariates in a CBLR model (referred to as Optimal for simplicity)
+in the complex simulation. We conduct 100 bootstrap re-samples on the training data to get confidence intervals. We kept 15\% of the data for testing before
+hyperparameter selection. 15\% of the remaining data is used for validation and the rest is reserved for training. We predict risk functions for individuals in
+the test set, which are used to calculate our IPA and $AUC_{IPCW}$.
 
@@
 
-This simulation shows performance with the presence of a complex
-baseline hazard and a time-varying interaction. Inspired by a cancer treatment
-that initially reduces risk of death, but slowly returns as time progresses. 
-Originally used to show the spline-based hazard model proposed by Royston and Parmar
-\citep{royston2002flexible}, the breast cancer dataset provides a
-complex baseline hazard from which we simulate, available in the
-\textbf{flexsurv} R package \citep{flexsurv}. To increase the complexity
-of our data-generating mechanism for this simulation, we design the
-model \begin{align}
+We use this simulation to assess performance on data with a complex baseline hazard and a time-varying interaction. We design the model \begin{align}
 \log h(t \mid X_i) =\sum_{i=1}^{5} (\gamma_{i} \cdot \psi_{i}) + \beta_{{1}} (z_{1}) + \beta_{{2}} (z_{2})+ \beta_{{3}} (z_{3})+ \tau_{1} ( z_{1} \cdot t)+ \tau_{2} (z_{2} \cdot z_{3}), \nonumber
-\end{align}
-
-where
+\end{align} where
 \(\gamma_{1}=3.9, \gamma_{2}=3, \gamma_{3}=-0.43, \gamma_{4}=1.33,\gamma_{5}=-0.86, \beta_{{1}}=-5, \beta_{{2}}=-1, \beta_{{3}}=1, \tau_{1}=0.001, \tau_{2}=-1\)
-and \(\psi\) are basis splines. The \(gamma\) coefficients are obtained
-from an intercept-only cubic splines model with three knots using the
-\emph{flexsurvspline} function from the \textbf{flexsurv} package
-\citep{flexsurv}. Note that we fix these values for the analysis. The
-\(\beta\) coefficients represent direct effects, \(\tau_{2}\) represents
-an interaction and \(\tau_{1}\) is a time-varying
-interaction.
+and \(\psi_{i}\) are basis splines. The \(\gamma\) coefficients are obtained from an intercept-only cubic splines model with three knots using the \emph{flexsurvspline}
+function from the \textbf{flexsurv} package \citep{flexsurv} on the German Breast Cancer Study Group dataset as it provides a complex baseline hazard from which
+we can simulate. This study group was collected from 1984 to 1989 following 686 patients \citep{royston2002flexible}. These $\gamma$, $\beta$ and $\tau$ coefficients are used
+as our baseline hazard parameters and are fixed for the analysis. The \(\beta\) coefficients represent direct effects, \(\tau_{2}\) represents an interaction and
+\(\tau_{1}\) is a time-varying interaction.
 
@@
 
-Figure \ref{fig:megaPlot} A, E and Table \ref{tab:megaTable} A
-shows the performance over time on a test set. For discrimination, the Optimal model
-performs best as expected, followed by CBNN, DeepHit, DeepSurv and the linear models (Figure \ref{fig:megaPlot} E).
-For both discrimination and calibration, the rankings remain the same aside from DeepHit,
-where there's a periodic loss in performance (Figure \ref{fig:megaPlot} A). The
-Optimal model acts as a reference for how good a model can perform,
-with CBNN performing better than the competitors in the complex simulation.
-To obtain a more realistic performance assessment, we compared models in
-three case studies with a time-to-event outcome.
+Figure \ref{fig:megaPlot} A, E and Table \ref{tab:megaTable} shows the performance over time on a test set. For discrimination, the Optimal model performs best,
+followed by CBNN, DeepHit, DeepSurv and the linear models (Figure \ref{fig:megaPlot} E). For both discrimination and calibration, the rankings remain the same
+except for DeepHit which shows a periodic loss in performance (Figure \ref{fig:megaPlot} A). The Optimal model acts as a reference for ideal performance on the
+simulated data, with CBNN performing better than the competitors in the complex simulation. To obtain a more realistic performance assessment, we compared
+models in three case studies with a time-to-event outcome.
 
 \begin{table}[!htbp]
-\caption{Four tables representing performance at certain percentages of follow-up time in the complex simulation (A), multiple myeloma (MM) case study (B), free light chain (FLC) case study (C) and prostate cancer (Prostate) case study (D).
-Each table shows performance for each method in each study at $25\%$, $50\%$.
-$75\%$ and $100\%$ of follow-up time.
-The bold elements show the best model for each study, at each follow-up time of interest.
-These tables are included to provide exact measures at certain intervals.
-The models of interest are: Cox, case-base with logistic regression (CBLR), DeepSurv, DeepHit, Case-Base Neural Network (CBNN) and Optimal (in the complex simulation).
-The CBNN label is in bold to make it easier to distinguish.
-The best score at each percent of follow-up time is highlighted in bold.
-If the average performance is tied, then all tied values are bold.}
+\caption{Performance at certain percentages of follow-up time in the complex simulation (A), multiple myeloma (MM) case study (B), free light chain (FLC)
+case study (C) and prostate cancer (Prostate) case study (D). Each table shows performance for each method in each study at $25\%$, $50\%$, $75\%$
+and $100\%$ of follow-up time. The bold elements show the best model for each study, at each follow-up time of interest. These tables are included to
+provide exact measures at certain intervals. The models of interest are case-base with logistic regression (CBLR), Cox, Case-Base Neural Network (CBNN),
+DeepHit, DeepSurv, and Optimal (in the complex simulation). The best score at each percent of follow-up time is highlighted in bold. If the average performance
+is tied, then all tied values are highlighted.}
 \label{tab:megaTable}
 
-\begin{center}\includegraphics[width=1\linewidth]{../figures/megaTable} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{../figures/Table2.pdf} \end{center}
 
@@
 
- The first case study examines multiple myeloma (MM) \citep{myeloma}.
-Case study two examines the relationship between serum free light chain (FLC) and mortality \citep{flc}.
-The third examines prostate cancer (Prostate) survival on a publicly available simulation of the Surveillance,
-Epidemiology, and End Results (SEER)-medicare study \citep{prostate}. Both MM and FLC
-are available as part of the \textbf{survival} package \citep{survpkg}. The Prostate dataset is available
-as part of the \textbf{asaur} package \citep{asaur}. As we do not know the true model for the case studies,
-we exclude the Optimal model. After we perform grid search with the same hyperparameter
-options, we split the data the same way we did for the simulation (allocated 15\% of the data for testing,
-15\% of the remaining data for validation, and the rest for testing).
-We predict risk functions for everyone in the test set, which is used to calculate our metrics. We
-conduct 100-fold bootstrap re-samples on the training data to
-get confidence intervals.
+The complex simulation study examines whether a complex baseline hazard and time-varying interactions affect method performance.
+The case study assesses performance in more realistic conditions, where we may not know if a flexible baseline hazard is beneficial, or if time-varying covariates
+exist. For each dataset, we perform a grid search with the same hyperparameter procedure as described in the simulation (three-fold cross-validated grid search).
+We then predict risk functions for everyone in the test set with the selected hyperparameters, which is used to calculate our metrics. We conduct 100-fold
+bootstrap re-samples on the training data to get confidence intervals.
 
@@
 \subsection{Performance evaluation on multiple myeloma dataset}\label{pe-multiplemyeloma}}
-The MM study tracks 3882 patients seen at the Mayo Clinic from 1947 to 1996 until death \citep{myeloma}.
-We use 2 covariates, year of entry into the study and the time of MM diagnosis\citep{myeloma}.
-We see 71\% incidence over 23 years\citep{myeloma}.
+We expect cancer may have risk factors that vary with time \citep{coradini2000time}. As such, The first case study examines multiple myeloma (MM) \citep{myeloma},
+using a cohort of 3882 patients seen at the Mayo Clinic from 1947 to 1996 until death \citep{myeloma}. Data are provided by the \textbf{survival} package \citep{survpkg}.
+We use two covariates, year of entry into the study and the time of MM diagnosis\citep{myeloma}. We see 71\% incidence over 23 years\citep{myeloma}.
 
-Figure \ref{fig:megaPlot} B, F and Table \ref{tab:megaTable} B
-demonstrate the performance over time on a test set. For discrimination,
-CBNN and DeepHit have a similar performance, followed by DeepSurv and finally the linear models performing worst (Figure \ref{fig:megaPlot} F). For
-both discrimination and calibration, CBNN performs substantially better than
-the linear models and DeepSurv, with DeepHit having the worst performance overall
-(Figure \ref{fig:megaPlot} B). Together, CBNN is the best calibrated model and one
-of the best at discrimination in the MM case study.
+Figure \ref{fig:megaPlot} B, F and Table \ref{tab:megaTable} B demonstrate the performance over time on a test set. For discrimination, CBNN and DeepHit have a
+similar performance, followed by DeepSurv and finally the linear models performing worst (Figure \ref{fig:megaPlot} F). For both discrimination and calibration, CBNN
+performs substantially better than the linear models and DeepSurv, with DeepHit having the worst performance overall (Figure \ref{fig:megaPlot} B). Together, CBNN
+is the best calibrated model and one of the best at discrimination in the MM case study.
 
@@
 \subsection{Performance evaluation on free light chain dataset}\label{pe-flc}}
-The FLC dataset is a random sample of half the individuals in $\frac{2}{3}$ of the
-residents of Olmsted County over the age of 50 \citep{flc}. We have access to 7874 subjects
-tracked until death \citep{flc}. We use 5 covariates, total serum FLC (sum of kappa and lambda), age, sex,
-serum creatine and monoclonal gammapothy state \citep{flc}. There is 27\% incidence over 14 years \citep{flc}.
+Serum free light chain (FLC) is a known diagnostic tool for assessing MM \citep{mm2flc}. We are interested in whether there is a predictive benefit if the FLC markers
+varies with time. As such the second case study examines the relationship between serum FLC and mortality \citep{flc} in a random sample of half the individuals in
+$\frac{2}{3}$ of the residents of Olmsted County over the age of 50 \citep{flc}. Data are provided by the \textbf{survival} package \citep{survpkg} for 7874 subjects tracked
+until death \citep{flc}. We use five covariates, total serum FLC (sum of kappa and lambda), age, sex, serum creatine and monoclonal gammapothy state \citep{flc}. There is
+27\% incidence over 14 years \citep{flc}.
 
-Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C
-demonstrate the performance over time on a test set. CBNN, DeepHit and the linear models
-perform best at discrimination, followed DeepSurv (Figure \ref{fig:megaPlot} G). The rankings
-remain the same aside from DeepHit where the performance periodically drops
-to worse than DeepSurv (Figure \ref{fig:megaPlot} C). Together, CBNN outperforms the competing models,
-in terms of both calibration and discrimination in the FLC study. However, CBNN is only slightly better
-than the linear ones.
+Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C demonstrate the performance over time on a test set. CBNN, DeepHit and the linear models perform best at
+discrimination, followed by DeepSurv (Figure \ref{fig:megaPlot} G). The rankings remain the same aside from DeepHit where the performance periodically drops to worse
+than DeepSurv (Figure \ref{fig:megaPlot} C). Together, CBNN outperforms the competing models in terms of both calibration and discrimination.
+However, CBNN is only slightly better than the linear ones.
 
@@
 \subsection{Performance evaluation on prostate cancer dataset}\label{pe-prostate}}
-The prostate cancer dataset has a record of competing risks \citep{prostate}. As we are only interested in the
-single event scenario, we only keep individuals with prostate cancer death or censoring \citep{prostate}.
-This subset tracks 11054 individuals with three covariates: differentiation grade, age group and cancer state \citep{prostate}.
-There is a 7\% incidence over 10 years.
+As the CBNN model saw large predictive benefits on MM and small predictive benefits on FLC, we wish to examine longitudinal data with a more complex risk profile.
+The third case study examines prostate cancer (Prostate) survival on a publicly available simulation of the Surveillance, Epidemiology, and End Results (SEER)-medicare study \citep{prostate}.
+The Prostate dataset, which is contained in the \textbf{asaur} package \citep{asaur}, has a record of competing risks \citep{prostate}. As we are only interested in the single event scenario,
+we only keep individuals with prostate cancer death or censoring \citep{prostate}. This subset tracks 11054 individuals with three covariates: differentiation grade, age group and cancer
+state \citep{prostate}. There is a 7\% incidence over 10 years.
 
-Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C demonstrate the
-performance over time on a test set. In the prostate case study, the linear models
-outperform the other models in both discrimination and calibration, followed by
-CBNN and DeepHit and finally DeepSurv (Figure \ref{fig:megaPlot} C, G). Aside from DeepSurv,
-the performance is similar across all models.
+Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C demonstrate the performance over time on a test set. In the prostate case study, the linear models outperform the other
+models in both discrimination and calibration, followed by CBNN and DeepHit and finally DeepSurv (Figure \ref{fig:megaPlot} C, G). Aside from DeepSurv, the performance is
+similar across all models.
 
@@
 
-CBNN models survival outcomes by using neural networks on case-base
-sampled data. We incorporate follow-up time as a feature, providing a
-data-driven estimate of a flexible baseline hazard and time-varying
-interactions in our hazard function. The two competing neural network
-models we evaluated cannot model time-varying interactions
- \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}.
+CBNN models survival outcomes by using neural networks on case-base sampled data. We incorporate follow-up time as a feature, providing a data-driven
+estimate of a flexible baseline hazard and time-varying interactions in our hazard function. The two competing neural network models we evaluated cannot
+model time-varying interactions \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}. The CBNN model performs better due to its ability to model time-varying
+interactions and a complex baseline hazard.
 
-A method we want to include is Deep Survival Machines (DSM), a
-parametric survival model using neural networks with a mixture of
-distributions as the baseline hazard \citep{dsmPaper}. It can model
-a flexible baseline hazard. However, like DeepHit
-and DeepSurv, DSM cannot model time-varying interactions \citep{dsmPaper}. DSM
-requires a two-phase learning process and user implementations for
-distributions beyond Log-Normal or Weibull
-\citep{dsmPaper}. DSM outperformed DeepSurv and DeepHit, but only tested
-a subset of hyperparameter combinations for competing neural network methods
-while going through a full grid search for their method \citep{dsmPaper}.
-With our hyperparameters options and model design, DSM did not
-converge in the complex simulation. This may be due to implementation issues or
-method limitations, therefore we did not include this method.
+The complex simulation requires a method that can learn both time-varying interactions and have a flexible baseline hazard. Here, CBNN shows a distinct
+advantage over all other methods. Based on our complex simulation results (Figure \ref{fig:megaPlot} A, E and Table \ref{tab:megaTable} A), CBNN
+outperforms the competitors. This simulation shows how all models perform under ideal conditions with minimal noise in the data, while the three
+case studies assess their performance in realistic conditions. In the MM case study, flexibility in both interaction modeling and baseline hazard
+improves CBNN's performance over the other models, suggesting that this flexibility aids calibration (Figure \ref{fig:megaPlot} B, F and Table
+\ref{tab:megaTable} B). upon examination of the FLC case study, CBNN demonstrates a small improvement to performance compared to the linear
+models and DeepHit for both IPA and AUC (Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C). In the Prostate case study, the linear
+models outperform the neural network ones, while CBNN and DeepHit alternate their positions depending on the follow-up time of interest and DeepSurv
+maintains last place (Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C). We attribute this to potential over-parameterization in the neural
+network models, as we did not test for fewer nodes in each hidden layer, even with dropout. Though the ranking places the linear models above the
+neural network ones, the overall performance is comparable aside from DeepSurv, falling within a small range of IPA and AUC values.
 
-Compared to CBNN, the neural network competitors also have limitations.
-DeepSurv is a proportional hazards model and does
-not estimate the baseline hazard \citep{katzman2018DeepSurv}. DeepHit
-requires an alpha hyperparameter, assumes a single distribution
-for the baseline hazard and models the survival function directly
-\citep{lee2018DeepHit}. The alternative neural network methods
-match on time, while CBNN models time directly.
+Compared to CBNN, the neural network competitors also have limitations. DeepSurv is a proportional hazards model and does not estimate the baseline
+hazard \citep{katzman2018DeepSurv}. DeepHit requires an alpha hyperparameter, assumes a single distribution for the baseline hazard and models the
+survival function directly \citep{lee2018DeepHit}. The alternative neural network methods match on time, while CBNN models time directly. Another method
+caught our interest during our literature review. Deep Survival Machines (DSM), a parametric survival model using neural networks with a mixture of
+distributions as the baseline hazard \citep{dsmPaper} can fit a flexible baseline hazard. However, like DeepHit and DeepSurv, DSM cannot model time-varying
+interactions \citep{dsmPaper}. DSM requires a two-phase learning process and user implementations for distributions beyond Log-Normal or Weibull
+\citep{dsmPaper}. DSM outperformed DeepSurv and DeepHit, but they only tested a subset of hyperparameter combinations for competing neural network
+methods while going through a full grid search for their method \citep{dsmPaper}. With our hyperparameter options, DSM did not converge in the complex
+simulation which may have been due to issues in software or method specific limitations. Therefore, we did not include this method in our comparison.
 
-Though we applied a full grid search for all neural network models,
-there are an infinite number of untested hyperparameters we did not
-test. A completely exhaustive search is computationally infeasible,
-therefore it is reasonable to expect that there is a set of hyperparameters
-that is better suited for each model. However, we test a reasonably large
-range while accounting for potential over-fitting by including dropout
-\citep{gulli2017}. We provided the same options to all models aside from
-DeepHit, which has a method specific hyperparameter $\alpha$
-\citep{lee2018DeepHit}. With these limitations in mind, we dive
-into the differences in performance across all tested methods.
+While we applied a full grid search for all neural network models, there are an infinite number of untested hyperparameters we did not test. A completely
+exhaustive search is computationally infeasible, therefore it is reasonable to expect that there is a set of hyperparameters that is better suited for each model
+in each study. However, we test a reasonably large range while accounting for potential over-fitting by including dropout \citep{gulli2017}. We provided the same
+options to all models aside from DeepHit, which has a method specific hyperparameter $\alpha$ \citep{lee2018DeepHit}. With these limitations in mind, we
+summarize the differences in performance across all tested methods.
 
-The complex simulation requires a method that can learn both time-varying
-interactions and have a flexible baseline hazard. Here, CBNN shows
-a distinct advantage over all other methods. Based on our complex simulation results (Figure
-\ref{fig:megaPlot} A, E and Table \ref{tab:megaTable} A), CBNN
-outperforms the competitors. This simulation shows how all models perform
-under ideal conditions with minimal noise in the data, while the three case studies
-assess their performance in realistic conditions. In the MM case study, flexibility in both
-interaction modeling and baseline hazard improves CBNN's performance over
-the other models, suggesting that this flexibility aids calibration (Figure
-\ref{fig:megaPlot} B, F and Table \ref{tab:megaTable} B). upon examination of the FLC case study,
-CBNN demonstrates small improvement to performance compared to the linear models and DeepHit
-for both IPA and AUC (Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C).
-In the Prostate case study, the linear models outperform
-the neural network ones, while CBNN and DeepHit alternate their positions depending
-on the follow-up time of interest and DeepSurv maintains last place
-(Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C). We attribute this to
-potential over-parameterization in the neural network models, as we did not test for fewer
-nodes in each hidden layer, even with dropout. Though the
-ranking places the linear models above the neural network ones, the overall
-performance is comparable aside from DeepSurv, falling within a small range of IPA and AUC values.
-
-If prediction is our goal, we suggest CBNN as the best model in the Complex simulation, MM case study
-and FLC case study. The linear models were competitive in the FLC case study and performed best
-in the prostate case study. Though neural network interpretability is steadily improving \citep{interpret},
-there is still a trade-off compared to regression models, especially
-when the predictive performance gain is minimal, like in the FLC case study. We suggest that reference models should be included
-when assessing neural network models. Both a null model (KM curve) and a linear model (either Cox or
-a flexible baseline hazard model like CBLR) provide insight as to whether the neural network model is learning
-anything useful beyond linear predictors that were not accounted for.
+If prediction is our goal, we suggest CBNN as the best model in the Complex simulation, MM case study and FLC case study. The linear models were
+competitive in the FLC case study and performed best in the prostate case study. Though neural network interpretability is steadily improving \citep{interpret},
+there is still a trade-off compared to regression models, especially when the predictive performance gain is minimal, like in the FLC case study. We suggest
+that reference models should be included when assessing neural network models. Both a null model (KM curve) and a linear model (either Cox or a flexible
+baseline hazard model like CBLR) provide insight as to whether the neural network model is learning anything useful beyond linear predictors that were
+not accounted for.
 
@@
 
-CBNN outperforms all competitors in the complex
-simulation and two case studies while maintaining competitive performance in a final case study.
-Once we perform case-base sampling and adjust for the sampling bias, we can use
-a sigmoid activation function to predict our hazard function. Our
-approach provides an alternative to incorporating censored individuals, treating
-survival outcomes as binary ones. Forgoing the requirement
-of custom loss functions, CBNN only requires the use of standard
-components in machine learning libraries (specifically, the add layer to
-adjust for sampling bias and the sigmoid activation function) after case-base sampling. Due to
-the simplicity in its implementation and by extension user experience,
-CBNN is both a user-friendly approach to data-driven survival analysis
-and is easily extendable to any feed-forward neural network framework.
+Our study was motivated by the lack of easily implemented survival models based on neural networks that can approximate time-varying interactions.
+This lead us to extend the case-base sampling technique, which has been used with logistic regression before, to neural network models for a data driven
+approach to time-varying interaction modeling. In our paper, We aim to compare CBNN with existing methods. We assess the discrimination and calibration
+of each method in a simulation with time-varying interactions and a complex baseline hazard, and three case studies. CBNN outperforms all competitors in
+the complex simulation and two case studies while maintaining competitive performance in a final case study. Once we perform case-base sampling and
+adjust for the sampling bias, we can use a sigmoid activation function to predict our hazard function. Our approach provides an alternative to incorporating
+censored individuals, treating survival outcomes as binary ones. Forgoing the requirement of custom loss functions, CBNN only requires the use of standard
+components in machine learning libraries (specifically, the add layer to adjust for sampling bias and the sigmoid activation function) after case-base sampling.
+Due to the simplicity in its implementation and by extension user experience, CBNN is both a user-friendly approach to data-driven survival analysis and is
+easily extendable to any feed-forward neural network framework.
 
@@
 
-The Prostate dataset is available as part of the \textbf{asaur} package in R.
-The MM and FLC datasets are available in the \textbf{survival} package in R.
+The Prostate dataset is available as part of the \textbf{asaur} package in R \citep{asaur}.
+The MM and FLC datasets are available in the \textbf{survival} package in R \citep{survpkg}.
 The code for this manuscript and its analyses can be found at
@@
 We would like to thank Dr.~James Meigs, the project
-leader of these awards, for his support and helpful discussions. The
+leader of UM1DK078616 and R01HL151855, for his support and helpful discussions. The
 work was also supported as part of the Congressionally Directed Medical
@@
 M.T.
-wrote the majority of the casebase sampling section \ref{case-base-sampling}.
+wrote the majority of the case-base sampling section \ref{case-base-sampling}.
 M.T.
 and S.B.
-provided key insights into the core sampling technique (casebase).
+provided key insights into the core sampling technique (case-base).
 R.S.

### 2023-04-27 analyses/plotter.Rmd
applied Rob's comments

@@
 g<-plot_grid(pt, legend,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))
-pdf(file="../StatPaper/figures/megaPlot.pdf",width=16,height=9)
+pdf(file="../StatPaper/figures/Figure2.pdf",width=16,height=9)
 g
@@
   }
-  rows<-c(5,as.vector(unlist(rows)))
+  rows<-c(as.vector(unlist(rows)))
   picks<-data.frame(row=rows,
-                    col=c(0,cols))
+                    col=c(cols))
   return(picks)
@@
   #ft$header$styles$text$bold$data
-  ft <- add_header_row(x = ft,top=F, values = c("Method", 25,50,75,100,25,50,75,100),
+  ft <- add_header_row(x = ft,top=F, values = c("Method","25%","50%","75%","100%",
+                                                "25%","50%","75%","100%"),
                        colwidths = c(1,1,1,1,1,1,1,1,1))
@@
 
-pdf(file="../StatPaper/figures/megaTable.pdf",width=14,height=13)
+pdf(file="../StatPaper/figures/Table2.pdf",width=14,height=13)
 cowplot::plot_grid(cFlex,mFlex,fFlex,pFlex,ncol=1)
@@
 
-pdf(file="../StatPaper/figures/winTable.pdf",width=12,height=6)
+pdf(file="../StatPaper/figures/Table1.pdf",width=12,height=6)
 cowplot::plot_grid(cwtab,mwtab,fwtab,pwtab,ncol=2)

### 2023-05-18 StatPaper/CBNNManuscript/CBNNManuscript.tex
StatSubmission

--- 2023-04-27 StatPaper/APA/wileyNJD-APA.tex
@@
-\documentclass[APA,LATO1COL]{WileyNJD-v2}
+\documentclass[APA,LATO1COL]{CBNNManuscript}
 \usepackage{lmodern}
@@
 
-\corres{Jesse Islam's \email{jesse.islam@mail.mcgill.ca}}
+\corres{Jesse Islam \email{jesse.islam@mail.mcgill.ca}}
 
@@
 \abstract[Summary]{Neural network-based survival methods can model data-driven covariate interactions.
-While these methods have led to better predictive performance than regression-based approaches, they cannot model time-varying interactions and complex baseline hazards simultaneously.
-To address this, we propose Case-Base Neural Network (CBNN) as a new approach that combines the case-base sampling framework with flexible neural network architectures.
+While these methods can provide better predictive performance than regression-based approaches, not all can model time-varying interactions and complex baseline hazards.
+To address this, we propose Case-Base Neural Networks (CBNNs) as a new approach that combines the case-base sampling framework with flexible neural network architectures.
 Using a novel sampling scheme and data augmentation to naturally account for censoring, we construct a feed-forward neural network that may take time as an input.
-CBNN predicts the probability of an event occurring at a given moment to estimate the hazard function.
-We compare the performance of CBNN to neural network and regression-based survival methods in a simulation and three case studies using two time-dependent metrics.
+CBNNs predict the probability of an event occurring at a given moment to estimate the hazard function.
+We compare the performance of CBNNs to regression and neural network-based survival methods in a simulation and three case studies using two time-dependent metrics.
 First, we examine performance on a simulation involving a complex baseline hazard and time-varying interactions to assess all methods, with CBNN outperforming competitors.
-Then, we apply all methods to two real data applications where CBNN outperforms the competing models as well.
-In a third case study, CBNN shows competitive performance.
+Then, we apply all methods to three real data applications, with CBNNs outperforming the competing models in two studies and showing similar performance in the third.
 Our results highlight the benefit of combining case-base sampling with deep learning to provide a simple and flexible modeling framework for data-driven, time-varying interaction modeling of single event survival outcomes.
@@
 
-\keywords{survival analysis, machine learning, case-base, neural
-network}
+\keywords{survival analysis, machine learning, case-base, neural network}
 
@@
 
-A common assumption we use in survival analysis is that the risk of the event of interest does not vary with
+A common assumption in survival analysis is that the risk of the event of interest does not vary with
 time (i.e. proportional hazards) \citep{hanley2009}. This simplifying assumption results in using Cox proportional
-hazards models more often than Smooth-in-time, accelerated failure time (AFT) models \citep{hanley2009}. This
+hazards models more often than smooth-in-time, accelerated failure time (AFT) models \citep{hanley2009}. This
 preference causes analyses to be based on hazard ratios and relative risks rather than on survival curves and
@@
 incidence have shown that the risk of developing a tumor involves a time-varying interaction with tumor size
-\citep{coradini2000time}. Time-varying covariates can be incorporated easily into AFT models \citep{royston2002flexible}.
-To provide flexibility in the baseline hazard, we can use a basis of splines on time as model inputs
-\citep{royston2002flexible}. However, these regression-based models require prior knowledge of potential
+\citep{coradini2000time}. These time-varying covariates can be incorporated easily into AFT models \citep{royston2002flexible}.
+However, these regression-based models require prior knowledge of potential
 time-varying interactions and their quantitative effects.
 
-Neural networks provide a data-driven approach to approximating interaction terms in models. For example.
+Neural networks provide a data-driven approach to approximating interaction terms that benefit prediction. For example,
 DeepSurv is a neural network-based proportional hazards model that implements the Cox partial log-likelihood
 as a custom loss function \citep{katzman2018DeepSurv}. This results in a stepwise absolute risk curve that cannot
-accommodate time-varying interactions. Compared to Cox regression, DeepSurv shows better performance on a
-real dataset \citep{katzman2018DeepSurv}. Previous work suggests we may change the loss function to address
+accommodate time-varying interactions.
+DeepSurv has shown better performance than Cox proportional hazard models in the Rotterdam and German Breast Cancer Study Group and The Molecular Taxonomy of Breast Cancer International Consortium datasets \citep{katzman2018DeepSurv}.
+It has also been used to develop personalized prediction models for cancers in a number of studies \citep{ds1} \citep{ds2} \citep{ds3}.
+Previous work suggests we may change the loss function to address
 non-proportional hazards \citep{faraggi1995neural}. In contrast, DeepHit assumes an inverse Gaussian distribution
@@
 
-We propose Case-Base Neural Networks (CBNN) as a method for single event survival analysis that models
-time-varying interactions and a flexible baseline hazard. CBNN use the case-base sampling technique, which
+We propose Case-Base Neural Networks (CBNNs) as a method for single event survival analysis that models
+time-varying interactions and a flexible baseline hazard. CBNNs use the case-base sampling technique, which
 allows probabilistic models to predict survival outcomes \citep{hanley2009}. After case-base sampling, we
@@
 In this paper, we describe the case-base sampling method and compare its properties to neural
-network models in Section \ref{methods}, along with our hyperparameter selection procedure,
-metrics and software implementation. We also explore the performance of CBNN, DeepSurv,
-DeepHit, Cox regression and CBLR on simulated data in Section \ref{sims}. Section \ref{casestudies}
-describes the performance of the same methods in three case studies. Section \ref{discussion} explores
-the implications of our results and contextualizes them within neural network survival analysis in a single
+network models, along with our hyperparameter selection procedure,
+metrics and software implementation. Next, we explore the performance of CBNN, DeepSurv,
+DeepHit, Cox regression and CBLR on simulated data and
+describe their performance in three case studies. Finally, we explore
+the implications of our results and contextualize them within neural network survival analysis in a single
 event setting.
@@
 \hypertarget{methods}{%
-\section{Case-base neural networks, metrics, hyperparameters and
+\section{Case-base neural network, metrics, hyperparameters and
 software}\label{methods}}
 
-Case-base sampling divides total survival time for each individual into person-specific moments
-(person-moments), treating each person-moment as a sample. In this section, we detail how neural
-networks explicitly incorporating time as a feature while adjusting for the sampling bias in case-base
-sampled data; and describe the metrics and hyperparameter selection procedure used to compare
+Case-base sampling is an alternative framework for survival analysis \citep{hanley2009},
+which converts the total survival time into discrete person-moments. In this section, we detail how neural
+networks explicitly incorporate time as a feature while adjusting for the sampling bias in case-base
+sampled data and describe the metrics and hyperparameter selection procedure used to compare
 CBNN with other software packages. An R package to use CBNN is available at
@@
 
-Case-base sampling is an alternative framework for survival analysis \citep{hanley2009},
-which converts the total survival time into discrete person-moments. Without this preliminary step,
-we cannot model survival outcomes with probabilistic methods. In case-base sampling, we sample
-from the continuous survival time of each person in our dataset to create a \emph{base series} of
-\emph{person-moments}. This \emph{base series} complements the \emph{case series}, which contains
+To implement case-base sampling, we divide the total survival time for each individual into discrete person-specific moments (person-moments) and treat each person-moment as a sample.
+This creates a \emph{base series} of
+\emph{person-moments} where an event does not occur. This \emph{base series} complements the \emph{case series}, which contains
 all person-moments at which the event of interest occurs.
@@
 \(h(t \mid X_i)\) using the sampled person-moments. Recall that \(h(t \mid X_i)\) is the instantaneous
-potential of experiencing the event at time \(t\) for a given set of covariates \(X_i\), assuming \(T_i \geq t\).
+risk of experiencing the event at time \(t\) for a given set of covariates \(X_i\), assuming \(T_i \geq t\).
 
@@
 reasonable concern is determining how large $b$ should be relative to $c$, since the size of \(b\)
-influences how much information is lost in the sampling process \citep{hanley2009}. Using our notation,
-Hanley $\&$ Miettinen (2009) adapt a quote from Mantel (1973): "By the reasoning that
-$\frac{cb}{c+b}=\left[\left( \frac{1}{c}+\frac{1}{b}\right)]\right]$ measures the relative information in a comparison of two
-averages based on sample sizes of c and b respectively, we might expect by analogy, which would
-of course not be exact in the present case, that this approach would result in only a moderate loss of
-information. (The practicing statistician is generally aware of this kind of thing. There is little to be gained
-by letting the size of one series, b, become arbitrarily large if the size of the other series, c, must remain
-fixed)" \citep{hanley2009} \citep{mantel1}. If $b=100c$, then we expect learned weights to be at most one
+influences how much information is lost in the sampling process \citep{hanley2009}. The relative information when comparing two averages
+is measured by $\frac{cb}{c+b}=\left[\left( \frac{1}{c}+\frac{1}{b}\right)\right]$, where $b$ is the sample size of the \emph{base series} and $c$
+is the sample size of the \emph{case series}  \citep{hanley2009} \citep{mantel1}. If $b=100c$, then we expect learned weights to be at most one
 percent higher than if the entire study base $B$ was used, as they are proportional to
-$\frac{1}{c}+\frac{1}{100c}$ rather than $\frac{1}{c}+\frac{1}{\infty c}$ \citep{hanley2009}.
+$\frac{1}{c}+\frac{1}{100c}$ rather than $\frac{1}{c}+\frac{1}{\infty c}$ \citep{hanley2009} \citep{mantel1}.
 
 If we sample the \emph{base series} uniformly across the study base, then the hazard function
-of the sampling process is equal to \(b/B\). Therefore, we have the following equality \cite{saarela2015}
-\footnote{For a rigorous treatment, see Saarela \& Hanley (2015) section 3 \cite{saarela2015} .}:
+of the sampling process is equal to \(b/B\). Therefore, we have the following equality \citep{saarela2015}
+[For a rigorous treatment, see Saarela \& Hanley (2015) section 3].:
 \begin{align}\label{eqn:main}
 \frac{P\left(Y_i=1 \mid X_i, T_i\right)}{P\left(Y_i = 0 \mid X_i, T_i\right)} = \frac{h\left(T_i \mid X_i\right)}{b/B}.
-\end{align} The odds of a person-moment being a part of the \emph{case
+\end{align} The odds of a person-moment being in the \emph{case
 series} is the ratio of the hazard \(h(T_i \mid X_i)\) and the uniform
@@
 After case-base sampling, we pass all features, including time, into any user-defined feed-forward
-component. Then, we add an offset term and pass the output through a sigmoid activation function
+component (Figure \ref{fig:NNarch}). Then, we add an offset term and pass the output through a sigmoid activation function
 (Figure \ref{fig:NNarch}). Since we are interested in predicting the odds of an event occurring, the sigmoid
@@
 profile, \(f_{\theta}(X, T)\) represents any feed-forward neural network architecture, \(\log\left(\frac{B}{b}\right)\)
-is the offset term for adjust for the bias (\(\log\left(\frac{b}{B}\right)\)) set by case-base sampling, \(\theta\) is the
+is the offset term to adjust for the bias (\(\log\left(\frac{b}{B}\right)\)) set by case-base sampling, \(\theta\) is the
 set of parameters learned by the neural network and \(\mathrm{sigmoid}(x)=\frac{1}{1+e^{-x}}\). By approximating
 a higher-order polynomial of time using a neural network, the baseline hazard specification is now data-driven,
-where user-defined hyperparameters such as regularization, number of layers and nodes control the flexibility
+while user-defined hyperparameters such as regularization, number of layers and nodes control the flexibility
 of the hazard function.
@@
 \caption{Methodological steps involved in CBNN. The first step, case-base sampling, is completed before training begins.
-We pass this sampled data through a feed-forward neural network and add an offset to adjust for the bias inherent in
-case-base sampling, then apply a sigmoid activation function to create a probability. Once the neural network model
-completes its training, we can convert the probability output to a hazard for the survival outcomes of interest.}\label{fig:NNarch}
+Then, we pass this sampled data through a feed-forward neural network and add an offset to adjust for the bias inherent in
+case-base sampling and apply a sigmoid activation function to create a probability. Once the neural network model
+completes its training, we can convert the probability output to a hazard for the survival outcome of interest.}\label{fig:NNarch}
 \end{figure}
@@
 
-Backpropagation with an appropriate minimization algorithm (e.g.~Adam, RMSPropagation, stochastic gradient
+Backpropagation with an appropriate minimization algorithm (such as Adam, RMSPropagation, stochastic gradient
 descent) is used to optimize the parameters in the model \citep{gulli2017}. For our analysis, we use Adam as implemented
@@
 \emph{base series} is not restricted. We use a ratio of 100:1 \emph{base series} to \emph{case series} \citep{hanley2009}.
-We pass our feed-forward neural network through a sigmoid activation function (Figure \ref{fig:NNarch}). Finally, we can
-convert this model output to a hazard. To use CBNN for predictions, we manually set the offset term \(\log\left(\frac{B}{b} \right)\)
+After fitting our model, we
+convert the output to a hazard. To use CBNN for predictions, we manually set the offset term $\left(\log\left(\frac{B}{b} \right)\right)$
 to 0 in the new data as we already account for the sampling bias during the fitting process.
@@
 risk function (\(F\)) at time \(t\) for a covariate profile \(X\),
-viz.\begin{align}\label{eqn:ci2}
+\begin{align}\label{eqn:ci2}
 F\left(t\mid X\right)& = 1 - \exp\left(-\int_{0}^{t}h(u|X) \,\textrm du\right).
@@
 
-To choose our method-specific hyperparameters, we used the Integrated Brier Score (IBS) \citep{graf1999}, which is based on the
+To choose our method-specific hyperparameters, we use the Integrated Brier Score (IBS) \citep{graf1999}, which is based on the
 Brier Score (BS) and provides a summarized assessment of performance for each model. We assess the performance of models on
@@
 \end{align} where \(\delta_{i}=1\) shows individuals who have experienced the event, \(N\) represents the number of samples
-in our dataset over which we calculate \(BS(t)\), \(\widehat{G}(t)=P[c>t]\) is a non-parametric estimate of the censoring distribution,
-\(c\) is censoring time and \(T_{i}\) is an individual's survival or censoring time. The BS provides a score that accounts for the
+in our dataset over which we calculate \(BS(t)\), \(T_{i}\) is the survival or censoring time of an individual, \(\widehat{G}(t)=P[c>t]\) is a non-parametric estimate of the censoring distribution and \(c\) is censoring time.
+The BS provides a score that accounts for the
 information loss because of censoring. Once we fix our \(t\) of interest, the individuals in the dataset can be divided into three
@@
 
-The Integrated Brier Score (IBS) is a function of the BS \ref{eqn:bs} \citep{graf1999}, which is defined as \begin{align*}
+The Integrated Brier Score (IBS) is a function of the BS (\ref{eqn:bs}) \citep{graf1999}, which is defined as \begin{align*}
 IBS(t)=\int_{t}^{t_{max}}BS(t)dw(t),
-\end{align*} where $w(t)=\frac{t}{t_{max}}$. As the IBS is calculated for a range of follow-up-times,
+\end{align*} where $w(t)=\frac{t}{t_{max}}$ and $t_{max}$ is the upper bound for the survival times of interest. As the IBS is calculated for a range of follow-up-times,
 it is a useful metric to track overall performance across all of follow-up time during cross-validation. We use the
@@
 
-The IPA as a function of time, based on the BS. The IPA score is given by \begin{align}
+The IPA is a function of time, based on the BS. The IPA score is given by \begin{align}
 \textrm{IPA}(t) &= 1-\frac{BS_{model}(t)}{BS_{null}(t)}, \nonumber
@@
 IPA score has an upper bound of one, where positive values show an increase in performance over the null model and
-negative values show that the null model performs better. An extension of BS, the IPA scores assess both model calibration
+negative values show that the null model performs better. As an extension of BS, the IPA score assesses both model calibration
 and discrimination and show how performance changes over follow-up time \citep{graf1999} \citep{kattan2018index}.
@@
 \citep{lee2018DeepHit}. We made the components of CBNN using the \textbf{casebase} package \citep{cbpkg} for the sampling step and the
-\textbf{keras} \citep{keras} package for our neural network architecture. We use the \textbf{simsurv} package \citep{simsurv} for our simulation studies
+\textbf{keras} package \citep{keras} for our neural network architecture. We use the \textbf{simsurv} package \citep{simsurv} for our simulation studies
 and \textbf{flexsurv} \citep{flexsurv} to fit a flexible baseline hazard using splines for our complex simulation. The \textbf{riskRegression} package
 \citep{riskRegression} is used to get the IPA and $AUC_{IPCW}$. We modify the \textbf{riskRegression} package to be used with any user supplied
-risk function \(F\). We use the \textbf{reticulate} package \citep{reticulate} to run both R based methods and python based methods on the same seed.
+risk function \(F\). We use the \textbf{reticulate} package \citep{reticulate} to run both R based methods and Python based methods on the same seed.
 
@@
 each model in each study over follow-up time, where higher is better. Each model-specific metric in each study shows a 95\% confidence interval over 100 iterations. Metrics
-are shown for six models: Case-Base with Logistic Regression (CBLR), Case-Base Neural Networks (CBNN), Cox Proportional Hazard (Cox), DeepHit
+are shown for six models: Case-Base with Logistic Regression (CBLR), Case-Base Neural Network (CBNN), Cox Proportional Hazard (Cox), DeepHit
 and DeepSurv. The Kaplan-Meier (KM) model serves as a baseline, predicting the average curve for all individuals. CBLR and Cox have near identical
-performance, resulting in curves that overlap. The Optimal (a CBLR model with the exact interaction terms and baseline hazard specified) model shows
+performance, resulting in curves that overlap. The Optimal model (a CBLR model with the exact interaction terms and baseline hazard specified) shows
 the best performance we can expect on the simulated data.}\label{fig:megaPlot}
@@
 methods. We specify a linear combination of each covariate as the linear predictor in the regression-based methods (Cox, CBLR), which contrasts
-with neural network approaches that allow for non-linear interactions. The simulated data are based on a complex baseline hazard with time-varying
+with neural network approaches that allow for approximations of non-linear interactions. The simulated data are based on a complex baseline hazard with time-varying
 interactions, each with 10\% random censoring. We simulate three covariates for 5000 individuals:
@@
 In addition to the methods described above, we include the exact functional form of the covariates in a CBLR model (referred to as Optimal for simplicity)
-in the complex simulation. We conduct 100 bootstrap re-samples on the training data to get confidence intervals. We kept 15\% of the data for testing before
-hyperparameter selection. 15\% of the remaining data is used for validation and the rest is reserved for training. We predict risk functions for individuals in
+in the complex simulation. We obtain confidence intervals by conducting 100 bootstrap re-samples on the training data. We keep 15\% of the data for testing before
+hyperparameter selection. We use 15\% of the remaining data for validation and the rest is reserved for training. We predict risk functions for individuals in
 the test set, which are used to calculate our IPA and $AUC_{IPCW}$.
@@
 function from the \textbf{flexsurv} package \citep{flexsurv} on the German Breast Cancer Study Group dataset as it provides a complex baseline hazard from which
-we can simulate. This study group was collected from 1984 to 1989 following 686 patients \citep{royston2002flexible}. These $\gamma$, $\beta$ and $\tau$ coefficients are used
+we can simulate. The study comprised 686 women with breast cancer followed between 1984 and 1989 \citep{royston2002flexible}. These $\gamma$, $\beta$ and $\tau$ coefficients are used
 as our baseline hazard parameters and are fixed for the analysis. The \(\beta\) coefficients represent direct effects, \(\tau_{2}\) represents an interaction and
@@
 
-Figure \ref{fig:megaPlot} A, E and Table \ref{tab:megaTable} shows the performance over time on a test set. For discrimination, the Optimal model performs best,
-followed by CBNN, DeepHit, DeepSurv and the linear models (Figure \ref{fig:megaPlot} E). For both discrimination and calibration, the rankings remain the same
-except for DeepHit which shows a periodic loss in performance (Figure \ref{fig:megaPlot} A). The Optimal model acts as a reference for ideal performance on the
-simulated data, with CBNN performing better than the competitors in the complex simulation. To obtain a more realistic performance assessment, we compared
+Figure \ref{fig:megaPlot} A, E and Table \ref{tab:megaTable} shows the performance over time on a test set.
+The Optimal model acts as a reference for ideal performance on the simulated data.
+For discrimination, the Optimal model performs best,
+followed by CBNN, DeepHit, DeepSurv and the linear models (Figure \ref{fig:megaPlot} E). For both   To obtain a more realistic performance assessment, we compared
 models in three case studies with a time-to-event outcome.
@@
 case study (C) and prostate cancer (Prostate) case study (D). Each table shows performance for each method in each study at $25\%$, $50\%$, $75\%$
-and $100\%$ of follow-up time. The bold elements show the best model for each study, at each follow-up time of interest. These tables are included to
-provide exact measures at certain intervals. The models of interest are case-base with logistic regression (CBLR), Cox, Case-Base Neural Network (CBNN),
+and $100\%$ of follow-up time. These tables are included to
+provide exact measures of performance. The models of interest are case-base with logistic regression (CBLR), Cox, Case-Base Neural Network (CBNN),
 DeepHit, DeepSurv, and Optimal (in the complex simulation). The best score at each percent of follow-up time is highlighted in bold. If the average performance
@@
 \hypertarget{casestudies}{%
-\section{Casestudies}\label{casestudies}}
+\section{Case studies}\label{casestudies}}
 
-The complex simulation study examines whether a complex baseline hazard and time-varying interactions affect method performance.
+The simulation examines whether a complex baseline hazard and time-varying interactions affect method performance.
 The case study assesses performance in more realistic conditions, where we may not know if a flexible baseline hazard is beneficial, or if time-varying covariates
@@
 \subsection{Performance evaluation on multiple myeloma dataset}\label{pe-multiplemyeloma}}
-We expect cancer may have risk factors that vary with time \citep{coradini2000time}. As such, The first case study examines multiple myeloma (MM) \citep{myeloma},
+We expect cancer may have risk factors that vary with time \citep{coradini2000time}. As such, the first case study examines multiple myeloma (MM),
 using a cohort of 3882 patients seen at the Mayo Clinic from 1947 to 1996 until death \citep{myeloma}. Data are provided by the \textbf{survival} package \citep{survpkg}.
-We use two covariates, year of entry into the study and the time of MM diagnosis\citep{myeloma}. We see 71\% incidence over 23 years\citep{myeloma}.
+We use two covariates, year of entry into the study and the time of MM diagnosis \citep{myeloma}. We see 71\% incidence over 23 years \citep{myeloma}.
 
@@
 Serum free light chain (FLC) is a known diagnostic tool for assessing MM \citep{mm2flc}. We are interested in whether there is a predictive benefit if the FLC markers
-varies with time. As such the second case study examines the relationship between serum FLC and mortality \citep{flc} in a random sample of half the individuals in
+varies with time. As such the second case study examines the relationship between serum FLC and mortality in a random sample of half the individuals in
 $\frac{2}{3}$ of the residents of Olmsted County over the age of 50 \citep{flc}. Data are provided by the \textbf{survival} package \citep{survpkg} for 7874 subjects tracked
-until death \citep{flc}. We use five covariates, total serum FLC (sum of kappa and lambda), age, sex, serum creatine and monoclonal gammapothy state \citep{flc}. There is
-27\% incidence over 14 years \citep{flc}.
+until death with 27\% incidence over 14 years \citep{flc}. We use five covariates, total serum FLC (sum of kappa and lambda), age, sex, serum creatine and monoclonal gammopathy state \citep{flc}.
 
@@
 As the CBNN model saw large predictive benefits on MM and small predictive benefits on FLC, we wish to examine longitudinal data with a more complex risk profile.
-The third case study examines prostate cancer (Prostate) survival on a publicly available simulation of the Surveillance, Epidemiology, and End Results (SEER)-medicare study \citep{prostate}.
-The Prostate dataset, which is contained in the \textbf{asaur} package \citep{asaur}, has a record of competing risks \citep{prostate}. As we are only interested in the single event scenario,
-we only keep individuals with prostate cancer death or censoring \citep{prostate}. This subset tracks 11054 individuals with three covariates: differentiation grade, age group and cancer
-state \citep{prostate}. There is a 7\% incidence over 10 years.
+The third case study (Prostate) examines prostate cancer survival on a publicly available simulation of the Surveillance, Epidemiology, and End Results (SEER) medicare study \citep{prostate}.
+The Prostate dataset, which is contained in the \textbf{asaur} package \citep{asaur}, has a record of competing risks. As we are only interested in the single event scenario,
+we only keep individuals with prostate cancer death or censoring. This subset tracks 11054 individuals with three covariates: differentiation grade, age group and cancer
+state. There is a 7\% incidence over 10 years \citep{prostate}.
 
@@
 
-CBNN models survival outcomes by using neural networks on case-base sampled data. We incorporate follow-up time as a feature, providing a data-driven
+CBNNs model survival outcomes by using neural networks on case-base sampled data. We incorporate follow-up time as a feature, providing a data-driven
 estimate of a flexible baseline hazard and time-varying interactions in our hazard function. The two competing neural network models we evaluated cannot
@@
 
-The complex simulation requires a method that can learn both time-varying interactions and have a flexible baseline hazard. Here, CBNN shows a distinct
-advantage over all other methods. Based on our complex simulation results (Figure \ref{fig:megaPlot} A, E and Table \ref{tab:megaTable} A), CBNN
+The complex simulation requires a method that can learn both time-varying interactions and have a flexible baseline hazard.
+Based on our complex simulation results (Figure \ref{fig:megaPlot} A, E and Table \ref{tab:megaTable} A), CBNN
 outperforms the competitors. This simulation shows how all models perform under ideal conditions with minimal noise in the data, while the three
 case studies assess their performance in realistic conditions. In the MM case study, flexibility in both interaction modeling and baseline hazard
-improves CBNN's performance over the other models, suggesting that this flexibility aids calibration (Figure \ref{fig:megaPlot} B, F and Table
-\ref{tab:megaTable} B). upon examination of the FLC case study, CBNN demonstrates a small improvement to performance compared to the linear
+improves the performance of CBNN over the other models, suggesting that this flexibility aids calibration (Figure \ref{fig:megaPlot} B, F and Table
+\ref{tab:megaTable} B). Upon examination of the FLC case study, CBNN demonstrates a small improvement to performance compared to the linear
 models and DeepHit for both IPA and AUC (Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C). In the Prostate case study, the linear
@@
 network models, as we did not test for fewer nodes in each hidden layer, even with dropout. Though the ranking places the linear models above the
-neural network ones, the overall performance is comparable aside from DeepSurv, falling within a small range of IPA and AUC values.
+neural network ones, their overall performance falls within a small range of IPA and AUC values aside from DeepSurv.
 
@@
 caught our interest during our literature review. Deep Survival Machines (DSM), a parametric survival model using neural networks with a mixture of
-distributions as the baseline hazard \citep{dsmPaper} can fit a flexible baseline hazard. However, like DeepHit and DeepSurv, DSM cannot model time-varying
-interactions \citep{dsmPaper}. DSM requires a two-phase learning process and user implementations for distributions beyond Log-Normal or Weibull
-\citep{dsmPaper}. DSM outperformed DeepSurv and DeepHit, but they only tested a subset of hyperparameter combinations for competing neural network
-methods while going through a full grid search for their method \citep{dsmPaper}. With our hyperparameter options, DSM did not converge in the complex
+distributions that can fit a flexible baseline hazard \citep{dsmPaper}. With our hyperparameter options, DSM did not converge in the complex
 simulation which may have been due to issues in software or method specific limitations. Therefore, we did not include this method in our comparison.
 
-While we applied a full grid search for all neural network models, there are an infinite number of untested hyperparameters we did not test. A completely
+While we apply a full grid search with three-fold cross-validation for all neural network models, there are an infinite number of untested hyperparameters we did not test. A completely
 exhaustive search is computationally infeasible, therefore it is reasonable to expect that there is a set of hyperparameters that is better suited for each model
-in each study. However, we test a reasonably large range while accounting for potential over-fitting by including dropout \citep{gulli2017}. We provided the same
+in each study. However, we test a reasonably large range while accounting for potential over-fitting by including dropout \citep{gulli2017}. We provide the same
 options to all models aside from DeepHit, which has a method specific hyperparameter $\alpha$ \citep{lee2018DeepHit}. With these limitations in mind, we
@@
 If prediction is our goal, we suggest CBNN as the best model in the Complex simulation, MM case study and FLC case study. The linear models were
-competitive in the FLC case study and performed best in the prostate case study. Though neural network interpretability is steadily improving \citep{interpret},
+competitive in the FLC case study and performed best in the Prostate case study. Though neural network interpretability is steadily improving \citep{interpret},
 there is still a trade-off compared to regression models, especially when the predictive performance gain is minimal, like in the FLC case study. We suggest
@@
 Our study was motivated by the lack of easily implemented survival models based on neural networks that can approximate time-varying interactions.
-This lead us to extend the case-base sampling technique, which has been used with logistic regression before, to neural network models for a data driven
-approach to time-varying interaction modeling. In our paper, We aim to compare CBNN with existing methods. We assess the discrimination and calibration
-of each method in a simulation with time-varying interactions and a complex baseline hazard, and three case studies. CBNN outperforms all competitors in
+This lead us to repurpose the case-base sampling technique, which has previously been used with logistic regression \citep{hanley2009}, to neural network models for a data driven
+approach to time-varying interaction modeling. In our paper, We aim to compare CBNNs with existing methods. We assess the discrimination and calibration
+of each method in a simulation with time-varying interactions and a complex baseline hazard, and three case studies. CBNNs outperform all competitors in
 the complex simulation and two case studies while maintaining competitive performance in a final case study. Once we perform case-base sampling and
 adjust for the sampling bias, we can use a sigmoid activation function to predict our hazard function. Our approach provides an alternative to incorporating
-censored individuals, treating survival outcomes as binary ones. Forgoing the requirement of custom loss functions, CBNN only requires the use of standard
+censored individuals, treating survival outcomes as binary ones. Forgoing the requirement of custom loss functions, CBNNs only require the use of standard
 components in machine learning libraries (specifically, the add layer to adjust for sampling bias and the sigmoid activation function) after case-base sampling.
-Due to the simplicity in its implementation and by extension user experience, CBNN is both a user-friendly approach to data-driven survival analysis and is
+Due to the simplicity in its implementation and by extension user experience, CBNNs are both a user-friendly approach to data-driven, single event survival analysis and is
 easily extendable to any feed-forward neural network framework.
@@
 
-The Prostate dataset is available as part of the \textbf{asaur} package in R \citep{asaur}.
-The MM and FLC datasets are available in the \textbf{survival} package in R \citep{survpkg}.
-The code for this manuscript and its analyses can be found at
-\url{https://github.com/Jesse-Islam/cbnnManuscript}. The software
-package making CBNN easier to implement can be found at
-\url{https://github.com/Jesse-Islam/cbnn}.
+The MM and FLC datasets are available in the \textbf{survival} package in R
+\citep{survpkg}. The Prostate dataset is available as part of the \textbf{asaur} package in R \citep{asaur}.
+The code for this manuscript and its analyses can be found at \url{https://github.com/Jesse-Islam/cbnnManuscript}. The software package making CBNN
+easier to implement can be found at \url{https://github.com/Jesse-Islam/cbnn}.
 
@@
 
-We would like to thank Dr.~James Meigs, the project
-leader of UM1DK078616 and R01HL151855, for his support and helpful discussions. The
-work was also supported as part of the Congressionally Directed Medical
-Research Programs (CDMRP) award W81XWH-17-1-0347. We would also like to
-thank Dr.~James Hanley for his support and discussions while extending
-the case-base methodology.
+We would like to thank Dr.~James Meigs, the project leader of UM1DK078616 and R01HL151855, for his support and helpful discussions.
+We would also like to thank Dr.~James Hanley for his support and discussions while extending the case-base methodology.
 
@@
 
-J.I.
-conceived the proof of concept, developed the theoretical formalism, developed the neural network code base, wrote the majority of the manuscript and performed the simulations and analyses.
-M.T.
-wrote the majority of the case-base sampling section \ref{case-base-sampling}.
-M.T.
-and S.B.
-provided key insights into the core sampling technique (case-base).
-R.S.
-provided multiple edits to the manuscript structure and key insights.
-All authors discussed the results and contributed to the final manuscript.
+J.I. conceived the proof of concept, developed the theoretical formalism, developed the neural network code base,
+wrote the majority of the manuscript and performed the simulations and analyses. M.T. wrote the majority of the case-base sampling section \ref{case-base-sampling}.
+M.T. and S.B. provided key insights into the core sampling technique (case-base). R.S. provided multiple edits to the manuscript structure and key insights. All authors
+discussed the results and contributed to the final manuscript.
 
@@
 This work was supported by subcontracts from UM1DK078616 and R01HL151855
-to Dr.~Rob Sladek.
-
+to R.S. The work was also supported as part of the Congressionally Directed Medical Research Programs (CDMRP) award W81XWH-17-1-0347.
 \subsection*{Conflict of interest}
@@
 %\nocite{*}% Show all bib entries - both cited and uncited; comment this line to view only cited bib entries;
-\bibliography{wileyNJD-APA}%
-
-\section*{Author Biography}
-
-\begin{biography}{\includegraphics[width=60pt,height=70pt,draft]{empty}}{\textbf{Jesse Islam} A PhD Candidate in Quantitative life sciences, studying complex relationships in data in a survival context.}
-\end{biography}
+\bibliography{CBNNManuscript}%
 

### 2023-05-18 analyses/plotter.Rmd
StatSubmission

@@
   }   
+  #   if(method=="AUC"){
+  #   method=expression("C"["IPCW"])
+  # }   
   metricSummary#<- metricSummary[-which(metricSummary$model=="CBNN_Spline"),]
@@
 #complex[[4]]$layers[[2]]<-NULL
+
+sz<-18
+szx<-15
 pt <- plot_grid( 
-  complex[[1]]+ theme(legend.position="none")+ ggtitle("Complex simulation")+ labs(x = "")+coord_cartesian(ylim=c(-0.3,0.7)),
-  myeloma[[1]]+ theme(legend.position="none")+ ggtitle("MM")+ labs(x = "",y="")+coord_cartesian(ylim=c(-1,.1)),
-  flchain[[1]]+ theme(legend.position="none")+ ggtitle("FLC")+ labs(x = "",y="")+coord_cartesian(ylim=c(-0.025,.45)),
-  prostate[[1]]+ theme(legend.position="none")+ ggtitle("Prostate")+ labs(x = "",y="")+coord_cartesian(ylim=c(-0.02,0.205)),
-  complex[[5]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)")+coord_cartesian(ylim=c(0.68,1)),
-  myeloma[[5]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)",y="")+coord_cartesian(ylim=c(0.00,0.75)),
-  flchain[[5]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)",y="")+coord_cartesian(ylim=c(0.8,0.875)),
-  prostate[[5]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)",y="")+coord_cartesian(ylim=c(0.65,0.82)),
+  complex[[1]]+ theme(legend.position="none")+ ggtitle("Complex simulation")+ labs(x = "")+coord_cartesian(ylim=c(-0.3,0.7))+theme(plot.title = element_text(size=sz)) +theme(axis.title.y=element_text(size=sz))+theme(axis.text=element_text(size=szx)),
+  myeloma[[1]]+ theme(legend.position="none")+ ggtitle("MM")+ labs(x = "",y="")+coord_cartesian(ylim=c(-1,.1)) + theme(plot.title = element_text(size=sz))+theme(axis.text=element_text(size=szx)),
+  flchain[[1]]+ theme(legend.position="none")+ ggtitle("FLC")+ labs(x = "",y="")+coord_cartesian(ylim=c(-0.025,.45)) + theme(plot.title = element_text(size=sz))+theme(axis.text=element_text(size=szx)),
+  prostate[[1]]+ theme(legend.position="none")+ ggtitle("Prostate")+ labs(x = "",y="")+coord_cartesian(ylim=c(-0.02,0.205)) + theme(plot.title = element_text(size=sz))+theme(axis.text=element_text(size=szx)),
+  complex[[5]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)",y=expression("AUC"["IPCW"]))+coord_cartesian(ylim=c(0.68,1)) +theme(axis.title.y=element_text(size=sz),axis.title.x=element_text(size=sz))+theme(axis.text=element_text(size=szx)),
+  myeloma[[5]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)",y="")+coord_cartesian(ylim=c(0.00,0.75)) +theme(axis.title.x=element_text(size=sz))+theme(axis.text=element_text(size=szx)),
+  flchain[[5]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)",y="")+coord_cartesian(ylim=c(0.8,0.875)) +theme(axis.title.x=element_text(size=sz))+theme(axis.text=element_text(size=szx)),
+  prostate[[5]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)",y="")+coord_cartesian(ylim=c(0.65,0.82)) +theme(axis.title.x=element_text(size=sz))+theme(axis.text=element_text(size=szx)),
   # complex[[4]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)")+coord_cartesian(ylim=c(0.5,0.8)),
@@
     theme(legend.position = "bottom")+
-    theme(legend.text=element_text(size=8))
+    theme(legend.text=element_text(size=sz))
 )
@@
 g<-plot_grid(pt, legend,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))
-pdf(file="../StatPaper/figures/Figure2.pdf",width=16,height=9)
+pdf(file="../StatPaper/figures/Figure2.pdf",width=18,height=10)
 g
@@
 
-  ft <- add_header_row(x = ft, values = c(title,"IPA","AUC"),
+  ft <- add_header_row(x = ft, values = c(title,"IPA", "AUC" ),
                        colwidths = c(1,4,4))
-  # ft <- compose(ft, part = "header", i = 1, j = 6,
-  #               value = as_paragraph("C",
-  #                                    as_sub("IPCW")))
+   ft <- compose(ft, part = "header", i = 1, j = 6,
+                 value = as_paragraph("AUC",
+                                      as_sub("IPCW")))
   #ft<-bold(ft, bold = TRUE, part = "header")

### 2023-06-30 manuscript/manuscript.tex
updated manuscript machine learning with applicat

--- 2023-05-18 StatPaper/CBNNManuscript/CBNNManuscript.tex
@@
-\documentclass[APA,LATO1COL]{CBNNManuscript}
-\usepackage{lmodern}
+%% 
+%% Copyright 2007-2020 Elsevier Ltd
+%% 
+%% This file is part of the 'Elsarticle Bundle'.
+%% ---------------------------------------------
+%% 
+%% It may be distributed under the conditions of the LaTeX Project Public
+%% License, either version 1.2 of this license or (at your option) any
+%% later version. The latest version of this license is in
+%% http://www.latex-project.org/lppl.txt
+%% and version 1.2 or later is part of all distributions of LaTeX
+%% version 1999/12/01 or later.
+%% 
+%% The list of all files belonging to the 'Elsarticle Bundle' is
+%% given in the file `manifest.txt'.
+%%
 
-\articletype{Article Type}%
+%% Template article for Elsevier's document class `elsarticle'
+%% with numbered style bibliographic references
+%% SP 2008/03/01
+%%
+%% 
+%%
+%% $Id: elsarticle-template-num.tex 190 2020-11-23 11:12:32Z rishi $
+%%
+%%
+\documentclass[preprint,12pt]{elsarticle}
+\usepackage{hyperref}
+ \usepackage{amsmath}
+ \usepackage{float}
+%% Use the option review to obtain double line spacing
+%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
 
-\received{XXXXXX}
-\revised{XXXXXX}
-\accepted{XXXXXX}
+%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
+%% for a journal layout:
+%% \documentclass[final,1p,times]{elsarticle}
+%% \documentclass[final,1p,times,twocolumn]{elsarticle}
+%% \documentclass[final,3p,times]{elsarticle}
+%% \documentclass[final,3p,times,twocolumn]{elsarticle}
+%% \documentclass[final,5p,times]{elsarticle}
+%% \documentclass[final,5p,times,twocolumn]{elsarticle}
 
-\raggedbottom
+%% For including figures, graphicx.sty has been loaded in
+%% elsarticle.cls. If you prefer to use the old commands
+%% please give \usepackage{epsfig}
+
+%% The amssymb package provides various useful mathematical symbols
+\usepackage{amssymb}
+%% The amsthm package provides extended theorem environments
+%% \usepackage{amsthm}
+
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers.
+%% \usepackage{lineno}
+
+\journal{Machine learning with applications}
 
 \begin{document}
+
+\begin{frontmatter}
+
+%% Title, authors and addresses
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for theassociated footnote;
+%% use the fnref command within \author or \address for footnotes;
+%% use the fntext command for theassociated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for theassociated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \affiliation{organization={},
+%% addressline={},
+%% city={},
+%% postcode={},
+%% state={},
+%% country={}}
+%% \fntext[label3]{}
 
@@
 
-\author[1]{Jesse Islam}
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{}
+%% \affiliation[label1]{organization={},
+%% addressline={},
+%% city={},
+%% postcode={},
+%% state={},
+%% country={}}
+%%
+%% \affiliation[label2]{organization={},
+%% addressline={},
+%% city={},
+%% postcode={},
+%% state={},
+%% country={}}
 
-\author[2]{Maxime Turgeon}
+\author[inst1]{Jesse Islam}
 
-\author[1,3]{Robert Sladek}
+\affiliation[inst1]{organization={McGill University Department of Quantitative Life Sciences},%Department and Organization
+ addressline={805 rue Sherbrooke O}, 
+ city={Montréal},
+ postcode={H3A 0B9}, 
+ state={Quebec},
+ country={Canada}}
 
-\author[4]{Sahir Bhatnagar}
+\author[inst2]{Maxime Turgeon}
+\author[inst1,inst3]{Robert Sladek}
+\author[inst4]{Sahir Bhatnagar}
 
-\address[1]{\orgdiv{Department of Quantitative Life Sciences,} \orgname{McGill University}, \orgaddress{\state{Quebec}, \country{Canada}}}
+\affiliation[inst2]{organization={University of Manitoba Department of Statistics},%Department and Organization
+ addressline={50 Sifton Rd}, 
+ city={Winnipeg},
+ postcode={R3T2N2}, 
+ state={Manitoba},
+ country={Canada}}
 
-\address[2]{\orgdiv{Department of Statistics}, \orgname{University of Manitoba}, \orgaddress{\state{Manitoba}, \country{Canada}}}
+\affiliation[inst3]{organization={McGill University Department of Human Genetics},%Department and Organization
+ addressline={805 rue Sherbrooke O}, 
+ city={Montréal},
+ postcode={H3A 0B9}, 
+ state={Quebec},
+ country={Canada}}
 
-\address[3]{\orgdiv{Department of Human Genetics}, \orgname{McGill University}, \orgaddress{\state{Quebec}, \country{Canada}}}
+\affiliation[inst4]{organization={McGill University Department of Biostatistics},%Department and Organization
+ addressline={805 rue Sherbrooke O}, 
+ city={Montréal},
+ postcode={H3A 0B9}, 
+ state={Quebec},
+ country={Canada}}
 
-\address[4]{\orgdiv{Department of Biostatistics}, \orgname{McGill University}, \orgaddress{\state{Quebec}, \country{Canada}}}
-
-\corres{Jesse Islam \email{jesse.islam@mail.mcgill.ca}}
-
-\presentaddress{McGill Genome Center: 740 Dr Penfield Ave, Montreal, Quebec H3A }
-
-\abstract[Summary]{Neural network-based survival methods can model data-driven covariate interactions.
+\begin{abstract}
+%% Text of abstract
+Neural network-based survival methods can model data-driven covariate interactions.
 While these methods can provide better predictive performance than regression-based approaches, not all can model time-varying interactions and complex baseline hazards.
@@
 Our results highlight the benefit of combining case-base sampling with deep learning to provide a simple and flexible modeling framework for data-driven, time-varying interaction modeling of single event survival outcomes.
-An R package is available at https://github.com/Jesse-Islam/cbnn.}
+An R package is available at https://github.com/Jesse-Islam/cbnn.
+\end{abstract}
 
-\keywords{survival analysis, machine learning, case-base, neural network}
+%%Research highlights
+\begin{highlights}
+\item Case-Base Neural Networks (CBNNs) provide a neural network methodology for survival analysis that leverages a statistical sampling technique, removing the need for complex implementations while providing data-driven flexibility in full hazard function estimation.
+\item CBNNs provide an advantage over competing neural network survival models, as it naturally accounts for censoring, uses binary cross-entropy as the loss function and predicts a smooth in-time risk function.
+\item CBNNs outperform the competing models in a simulation, two studies and shows similar performance in the third.
+\item Our results highlight the benefit of combining case-base sampling with deep learning to provide a simple and flexible modeling
+framework for data-driven, time-varying interaction modeling of single event
+survival outcomes.
+\end{highlights}
 
-\maketitle
+\begin{keyword}
+%% keywords here, in the form: keyword \sep keyword
+survival analysis, \sep machine learning \sep case-base \sep neural network
+%% PACS codes here, in the form: \PACS code \sep code
+%\PACS 0000 \sep 1111
+%% MSC codes here, in the form: \MSC code \sep code
+%% or \MSC[2008] code \sep code (2000 is the default)
+%\MSC 0000 \sep 1111
+\end{keyword}
+
+\end{frontmatter}
+
+%% \linenumbers
 
@@
 accommodate time-varying interactions.
-DeepSurv has shown better performance than Cox proportional hazard models in the Rotterdam and German Breast Cancer Study Group and The Molecular Taxonomy of Breast Cancer International Consortium datasets \citep{katzman2018DeepSurv}.
+DeepSurv performs better than Cox proportional hazard models in the Rotterdam and German Breast Cancer Study Group and The Molecular Taxonomy of Breast Cancer International Consortium datasets \citep{katzman2018DeepSurv}.
 It has also been used to develop personalized prediction models for cancers in a number of studies \citep{ds1} \citep{ds2} \citep{ds3}.
@@
 
-{\centering \includegraphics[width=1\linewidth]{../figures/Figure1.pdf}}
+{\centering \includegraphics[width=1\linewidth]{figures/Figure1.pdf}}
 
@@
 a held out dataset using two metrics: 1) the Index of Prediction Accuracy (IPA) \citep{kattan2018index}; and 2) the Inverse probability
-censoring weights-adjusted time dependent area under the receiver operating characteristic curve ($AUC_{IPCW}$) \citep{auc}. The
+censoring weights-adjusted time-dependent area under the receiver operating characteristic curve ($AUC_{IPCW}$) \citep{auc}. The
 IPA score is a metric for both discrimination and calibration, while the $AUC_{IPCW}$ provides a metric for discrimination only.
@@
 negative values show that the null model performs better. As an extension of BS, the IPA score assesses both model calibration
-and discrimination and show how performance changes over follow-up time \citep{graf1999} \citep{kattan2018index}.
+and discrimination and shows how performance changes over follow-up time \citep{graf1999} \citep{kattan2018index}.
 
 \hypertarget{auc}{%
-\subsubsection{Inverse probability censoring weights-adjusted time dependent area under the receiver operating characteristic curve ($AUC_{IPCW}$) }\label{auc}}
+\subsubsection{Inverse probability censoring weights-adjusted time-dependent area under the receiver operating characteristic curve ($AUC_{IPCW}$)}\label{auc}}
 
@@
 \begin{align*}
+\resizebox{\textwidth}{!}
+{
 AUC_{IPCW}(t)=\frac{\sum^{n}_{i=1}\sum^{n}_{j=1} I \left(\widehat{F}(t|X_{i})>\widehat{F}(t|X_{j})\right) \cdot I(T_{i}\leq t,\delta_{i}=1) \cdot \left( 1-I(T_{j}\leq t,\delta_{j}=1) \right) \cdot W_{i}(t) \cdot W_{j}(t)} { \sum^{n}_{i=1}\sum^{n}_{j=1} I(T_{i}\leq t,\delta_{i}=1) \cdot \left( 1-I(T_{j}\leq t,\delta_{j}=1) \right) \cdot W_{i}(t) \cdot W_{j}(t) },
-\end{align*}where \begin{align*}
+}
+\end{align*}
+where \begin{align*}
 W_{i}(t)=\frac{I(T_{i}\leq t,\delta_{i}=1)}{\widehat{G}(T_{i})} +\frac{I(T_{i}>t)}{\widehat{G}(t)}.
@@
 \[
-Grid\ search: \begin{cases}
+Search \ space: \begin{cases}
 Learning\ rate \sim \{0.001, 0.01\} \\ 
@@
 
-\begin{table}
+DeepHit uses $\alpha$ as a hyperparameter, while DeepSurv and CBNN do not. We track IBS on the validation set for each hyperparameter combination,
+choosing the combination with the lowest score for each method (Table \ref{tab:wins}).
+
+\begin{table}[H]
 \caption{Hyperparameters selected after three-fold cross-validated grid search along with the average IBS for each neural network model in the
@@
 \label{tab:wins}
-\begin{center}\includegraphics[width=1\linewidth]{../figures/Table1.pdf} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{figures/Table1.pdf} \end{center}
 \end{table}
-
-DeepHit uses $\alpha$ as a hyperparameter, while DeepSurv and CBNN do not. We track IBS on the validation set for each hyperparameter combination,
-choosing the combination with the lowest score for each method (Table \ref{tab:wins}).
 
@@
 
-\begin{figure}
+\begin{figure}[H]
 
-{\centering \includegraphics[width=1\linewidth]{../figures/Figure2.pdf}
+{\centering \includegraphics[width=1\linewidth]{figures/Figure2.pdf}
 
@@
 methods. We specify a linear combination of each covariate as the linear predictor in the regression-based methods (Cox, CBLR), which contrasts
-with neural network approaches that allow for approximations of non-linear interactions. The simulated data are based on a complex baseline hazard with time-varying
-interactions, each with 10\% random censoring. We simulate three covariates for 5000 individuals:
+with neural network approaches that allow for approximations of non-linear interactions. We simulate data based on a complex baseline hazard with time-varying
+interactions and 10\% random censoring. We simulate three covariates for 5000 individuals:
 
 \[
+\resizebox{\textwidth}{!}
+{
 z_{1} \sim \textrm{Bernoulli}(0.5) \qquad \qquad %technically a control and treatment group
@@
 z_{3} \sim N(1,0.5).
+}
 \]
+
 In addition to the methods described above, we include the exact functional form of the covariates in a CBLR model (referred to as Optimal for simplicity)
@@
 For discrimination, the Optimal model performs best,
-followed by CBNN, DeepHit, DeepSurv and the linear models (Figure \ref{fig:megaPlot} E). For both   To obtain a more realistic performance assessment, we compared
+followed by CBNN, DeepHit, DeepSurv and the linear models (Figure \ref{fig:megaPlot} E). To obtain a more realistic performance assessment, we compared
 models in three case studies with a time-to-event outcome.
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{../figures/Table2.pdf} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{figures/Table2.pdf} \end{center}
 
@@
 While we apply a full grid search with three-fold cross-validation for all neural network models, there are an infinite number of untested hyperparameters we did not test. A completely
-exhaustive search is computationally infeasible, therefore it is reasonable to expect that there is a set of hyperparameters that is better suited for each model
+exhaustive search is computationally infeasible, therefore it is reasonable to expect that there exists a set of hyperparameters that is better suited for each model
 in each study. However, we test a reasonably large range while accounting for potential over-fitting by including dropout \citep{gulli2017}. We provide the same
@@
 Our study was motivated by the lack of easily implemented survival models based on neural networks that can approximate time-varying interactions.
-This lead us to repurpose the case-base sampling technique, which has previously been used with logistic regression \citep{hanley2009}, to neural network models for a data driven
+This lead us to apply the case-base sampling technique, which has previously been used with logistic regression \citep{hanley2009}, to neural network models for a data-driven
 approach to time-varying interaction modeling. In our paper, We aim to compare CBNNs with existing methods. We assess the discrimination and calibration
@@
 
-\appendix
-%\nocite{*}% Show all bib entries - both cited and uncited; comment this line to view only cited bib entries;
-\bibliography{CBNNManuscript}%
+%% If you have bibdatabase file and want bibtex to generate the
+%% bibitems, please use
+%%
+ \bibliographystyle{elsarticle-num} 
+ \bibliography{cas-refs}
 
+%% else use the following coding to input the bibitems directly in the
+%% TeX file.
+
+% \begin{thebibliography}{00}
+
+% %% \bibitem{label}
+% %% Text of bibliographic item
+
+% \bibitem{}
+
+% \end{thebibliography}
 \end{document}
+\endinput
+%%
+%% End of file `elsarticle-template-num.tex'.

### 2023-06-30 analyses/plotter.Rmd
updatedArxiv, fixed figure size

@@
 library(dplyr)
-library(flextable)
 
@@
 #tf$config$threading$set_inter_op_parallelism_threads(1)
+
+library(flextable)
 
@@
 g<-plot_grid(pt, legend,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))
-pdf(file="../StatPaper/figures/Figure2.pdf",width=18,height=10)
+pdf(file="../manuscript/figures/Figure2.pdf",width=18,height=10)
 g
@@
 
+library(flextable)
 edit_cell <- function(table, row, col, name="core-fg", ...){
@@
 
+convert1<-function(tab){
+  tempNames<-rownames(tab)
+  tab<-as.data.frame(lapply(tab, function(x){gsub("1 ","1.00 ", x,fixed=T)}))
+  tab<-as.data.frame(lapply(tab, function(x){gsub('(1,',"(1.00,", x,fixed=T)}))
+  tab<-as.data.frame(lapply(tab, function(x) gsub(",1)",",1.00)", x,fixed=T)))
+  rownames(tab)<-tempNames
+  return(tab)
+}
+
 flexMaker<-function(tab,boldPicks=NA,title){
@@
   #  colwidths = c(9))
-  ft <- align(ft, i = c(1,2), part = "header", align = "center")
+  ft <- flextable::align(ft, i = c(1,2), part = "header", align = "center")
   ft<-vline(ft, j = c(1,5), border = bord2, part = "body")
@@
   #ft<-hline(ft,i=c(1),border=bord,part="body")
-  ft<-fontsize(ft, i = NULL, j = NULL, size =7, part = "all")
-  ft<-fontsize(ft, i = NULL, j = NULL, size =10, part = "header")
-  ft <- align(ft, i = NULL, part = "body", align = "center")
-  ft$body$styles$text$font.size$data[,1]<-10
+  ft<-fontsize(ft, i = NULL, j = NULL, size =12, part = "all")
+  ft<-fontsize(ft, i = NULL, j = NULL, size =14, part = "header")
+  ft <- flextable::align(ft, i = NULL, part = "body", align = "center")
+  ft$body$styles$text$font.size$data[,1]<-14
 
@@
   Tab<-convert0(Tab)
+  Tab<-convert1(Tab)
   Flex<-flexMaker(tab=Tab,boldPicks = Picks,title=titler)
@@
 
-cFlex<-getFlex(complex,"A: Complex")
-mFlex<-getFlex(myeloma,"B: MM")
-fFlex<-getFlex(flchain,"C: FLC")
-pFlex<-getFlex(prostate,"D: Prostate")
+cFlex<-getFlex(complex,"A:Complex")
+mFlex<-getFlex(myeloma,"B:MM")
+fFlex<-getFlex(flchain,"C:FLC")
+pFlex<-getFlex(prostate,"D:Prostate")
 
-pdf(file="../StatPaper/figures/Table2.pdf",width=14,height=13)
+pdf(file="../manuscript/figures/Table2.pdf",width=19,height=26)
 cowplot::plot_grid(cFlex,mFlex,fFlex,pFlex,ncol=1)
@@
 
-pdf(file="../StatPaper/figures/Table1.pdf",width=12,height=6)
+pdf(file="../manuscript/figures/Table1.pdf",width=12,height=6)
 cowplot::plot_grid(cwtab,mwtab,fwtab,pwtab,ncol=2)
@@
 
-pdf(file="../StatPaper/figures/ComplexPopTime.pdf",width=8,height=6)
+pdf(file="../manuscript/figures/ComplexPopTime.pdf",width=8,height=6)
 a

### 2023-06-30 manuscript/manuscript.tex
updatedArxiv, fixed figure size

@@
 \textrm{IPA}(t) &= 1-\frac{BS_{model}(t)}{BS_{null}(t)}, \nonumber
-\end{align} where \(BS_{model}(t)\) represents the BS over time \(t\) for the model of interest and \(BS_{null}(t)\) represents
+\end{align} where $(BS_{model}(t))$ represents the BS over time $(t)$ for the model of interest and $(BS_{null}(t))$ represents
 the BS if we use an unadjusted Kaplan-Meier (KM) curve as the prediction for all observations \citep{kattan2018index}. The
@@
 
-\hypertarget{auc}{%
-\subsubsection{Inverse probability censoring weights-adjusted time-dependent area under the receiver operating characteristic curve ($AUC_{IPCW}$)}\label{auc}}
+\hypertarget{auc2}{%
+\subsubsection{Inverse probability censoring weights-adjusted time-dependent area under the receiver operating characteristic curve $AUC_{IPCW}$}\label{auc2}}
 
 The IPCW-adjusted AUC ($AUC_{IPCW}$) is a time-dependent metric that considers censoring \citep{auc}. For a given follow-up time of interest,
-\begin{align*}
-\resizebox{\textwidth}{!}
-{
-AUC_{IPCW}(t)=\frac{\sum^{n}_{i=1}\sum^{n}_{j=1} I \left(\widehat{F}(t|X_{i})>\widehat{F}(t|X_{j})\right) \cdot I(T_{i}\leq t,\delta_{i}=1) \cdot \left( 1-I(T_{j}\leq t,\delta_{j}=1) \right) \cdot W_{i}(t) \cdot W_{j}(t)} { \sum^{n}_{i=1}\sum^{n}_{j=1} I(T_{i}\leq t,\delta_{i}=1) \cdot \left( 1-I(T_{j}\leq t,\delta_{j}=1) \right) \cdot W_{i}(t) \cdot W_{j}(t) },
-}
-\end{align*}
+
+\begin{equation*}
+\resizebox{\hsize}{!}{$AUC_{IPCW}(t)=\frac{\sum^{n}_{i=1}\sum^{n}_{j=1} I_{\widehat{F}(t|X_{i})>\widehat{F}(t|X_{j})} \cdot I_{T_{i}\leq t,\delta_{i}=1} \cdot \left( 1-I_{T_{j}\leq t,\delta_{j}=1} \right) \cdot W_{i}(t) \cdot W_{j}(t)} { \sum^{n}_{i=1}\sum^{n}_{j=1} I_{T_{i}\leq t,\delta_{i}=1} \cdot \left( 1-I_{T_{j}\leq t,\delta_{j}=1} \right) \cdot W_{i}(t) \cdot W_{j}(t) },$}
+\end{equation*}
+
 where \begin{align*}
-W_{i}(t)=\frac{I(T_{i}\leq t,\delta_{i}=1)}{\widehat{G}(T_{i})} +\frac{I(T_{i}>t)}{\widehat{G}(t)}.
+W_{i}(t)=\frac{I_{T_{i}\leq t,\delta_{i}=1}}{\widehat{G}(T_{i})} +\frac{I_{T_{i}>t}}{\widehat{G}(t)}.
 \end{align*} The $AUC_{IPCW}$ measures discrimination \citep{auc}. By examining both IPA and $AUC_{IPCW}$,
@@
 
-\[
-\resizebox{\textwidth}{!}
-{
+\begin{equation*}
+\resizebox{\hsize}{!}{$
 z_{1} \sim \textrm{Bernoulli}(0.5) \qquad \qquad %technically a control and treatment group
@@
 \end{cases} \qquad \qquad
-z_{3} \sim N(1,0.5).
-}
-\]
+z_{3} \sim N(1,0.5).$}
+\end{equation*}
 
@@
 \label{tab:megaTable}
-
-\begin{center}\includegraphics[width=1\linewidth]{figures/Table2.pdf} \end{center}
-
+\begin{center}\includegraphics[scale=0.24]{figures/Table2.pdf} \end{center}
 \end{table}
@@
  \bibliographystyle{elsarticle-num} 
- \bibliography{cas-refs}
+ \bibliography{manuscript}
 

### 2023-12-17 analyses/optimFlchain.Rmd
Improved documentation for potential users

@@
 ```{r}
+# First load reticulate , then initialize python environment for CUDA (I turn it off as we only had one gpu and the 4 analyses were run in parallel)
 library(reticulate)
@@
 ##########################
-#bsize=512
-min_delta = 0 #10^-7
-
+min_delta = 0
 epo=as.numeric(params$epo)
@@
 iteration<-as.numeric(params$iteration)
-
-#baselineTrials<-2
-
 reoptimize=T
-
 finalFit=T
@@
 
-#5 fold cross validation to select best hyperparameters, average performance on a test set.
+# 3 fold cross validation to select best hyperparameters, average performance on a test set.
 
 ```{r flchainPrep,eval=TRUE,hide=FALSE}
-
+##############################
+### Process data to prepare for analysis
+##############################
 library(survival)
 flchain<-flchain
-
 preData<-data.frame(age=as.numeric(flchain$age),
@@
 modData<-as.data.frame(model.matrix(~.,preData)[,-1])
-
 data<-as.data.frame(apply(modData,2,as.numeric))
-
 spec = c(train = .85, test = .15)
-
 g = sample(cut(
@@
 ))
-
 res = split(data, g)
@@
 for(i in 1:k){
-
   spec = c(train = .85, val = .15)
-
   g = sample(cut(
@@
 ```{r setFixedHyperparameters,eval=TRUE,hide=FALSE}
+##################################
+###Set up hyperparameter search space
+##################################
 
@@
 ```{r hyperOpDsurv,eval=TRUE,hide=FALSE}
+
+###############################
+### optimize hyperparameters for DeepSurv
+###############################
 klist<-readRDS("cvData/flchainCV.rds")
@@
 ```{r optimDeepHit,eval=TRUE,hide=FALSE}
+
+###############################
+### optimize hyperparameters for DeepHit
+###############################
 klist<-readRDS("cvData/flchainCV.rds")
@@
 ```{r hyperopcbnn,eval=TRUE,hide=FALSE}
+
+###############################
+### optimize hyperparameters for CBNN
+###############################
 klist<-readRDS("cvData/flchainCV.rds")
@@
 ```{r winners,eval=TRUE,hide=FALSE}
+
+###############################
+### save the best performaning hyperparameter sets
+###############################
 winners<-list(cbnn=cbnnPerformance,
@@
 
+#####################################
+### run 100 fold bootstrap on the training set for each method
+#####################################
+
+#prepare data
 winners<-readRDS("cvData/flchainWins.rds")
-
 klist<-readRDS("cvData/flchainCV.rds")
@@
 
+#initialize data
 start_time <- Sys.time()
@@
 
+# split training validation sets
 spec = c(train = .85, validate = .15)
-
 g = sample(cut(
@@
 
+  #prepare data
   trainOG<-res$train
-
   numSamp<-nrow(trainOG)
@@
   train<-trainOG[samp,]
-
   val<-res$validate
-
   sds<-sapply(train, function(x) (sd(x)))
@@
   maxTime<-max(train$time)
-
   fullTest<-readRDS("cvData/flchainTest.rds")
   test<-fullTest[,-c(ncol(data))]
+
+  #normalize data
   test<-normalizer(fullTest,means,sds,maxTime)[,-c(ncol(data))]
   fullTest<-normalizer(fullTest,means,sds,maxTime)
-
   val<-normalizer(res$validate,means,sds,maxTime)
-
   train<-normalizer(train,means,sds,maxTime)
 
+  #prepare data for linear model
   linModelTrain<-as.data.frame(rbind(train,val))
@@
   )
-
   times<-head(times, -1)
@@
 
+#######################
+###plot IPA results
+#######################
+
 BrierIPAList<-readRDS('results/100flchainBrier.rds')
@@
 ```{r auc}
-
+#######################
+###plot AUC results
+#######################
 AUClist<-readRDS('results/100flchainAuc.rds')
@@
 ```{r}
+
+#######################
+###plot C-index results
+#######################
 

### 2023-12-17 analyses/optimMyeloma.Rmd
Improved documentation for potential users

@@
 ##########################
-
 min_delta = 0
-
 epo=as.numeric(params$epo)
@@
 iteration<-as.numeric(params$iteration)
-
-reoptimize=FALSE
-
+reoptimize=T
 finalFit=T
@@
 
-#5 fold cross validation to select best hyperparameters, average performance on a test set.
+# 3 fold cross validation to select best hyperparameters, average performance on a test set.
 
@@
 
+##############################
+### Process data to prepare for analysis
+##############################
+
 library(survival)
-
 data("cancer", package="survival")
-
 preData<-data.frame(year=as.numeric(myeloma$year),
@@
 res = split(data, g)
-
 saveRDS(res$test,"cvData/myelomaTest.rds")
@@
 
+##################################
+###Set up hyperparameter search space
+##################################
+
 lr=c(0.001,0.01)
@@
 ```{r hyperOpDsurv,eval=TRUE,hide=FALSE}
+###############################
+### optimize hyperparameters for DeepSurv
+###############################
 source_python('src/dsurv.py')
@@
 ```{r optimDeepHit,eval=TRUE,hide=FALSE}
+###############################
+### optimize hyperparameters for DeepHit
+###############################
 klist<-readRDS("cvData/myelomaCV.rds")
@@
 ```{r hyperopcbnn,eval=TRUE,hide=FALSE}
+###############################
+### optimize hyperparameters for CBNN
+###############################
 klist<-readRDS("cvData/myelomaCV.rds")
@@
 ```{r winners,eval=TRUE,hide=FALSE}
+###############################
+### save the best performaning hyperparameter sets
+###############################
 winners<-list(cbnn=cbnnPerformance,
@@
 ```{r bestFits,finalFit}
-
+#####################################
+### run 100 fold bootstrap on the training set for each method
+#####################################
 winners<-readRDS("cvData/myelomaWins.rds")
@@
 
+#######################
+###plot IPA results
+#######################
+
 BrierIPAList<-readRDS('results/100myelomaBrier.rds')
@@
 
+#######################
+###plot AUC results
+#######################
+
 AUClist<-readRDS('results/100myelomaAuc.rds')
@@
 
+#######################
+###plot C-index results
+#######################
 cidxList<-readRDS('results/100myelomacidx.rds')
-
 temp<-melt(setDT(as.data.frame(cidxList[[1]])), id.vars = c("times"), variable.name = "model")
 cidxSims=data.frame(model=temp$model,times=temp$times)
-
 count=1

### 2023-12-17 analyses/optimProstate.Rmd
Improved documentation for potential users

@@
 ```{r prostatePrep,eval=TRUE,hide=FALSE}
+##############################
+### Process data to prepare for analysis
+##############################
 
@@
 
+##################################
+###Set up hyperparameter search space
+##################################
 lr=c(0.001,0.01)
@@
 ```{r hyperOpDsurv,eval=TRUE,hide=FALSE}
-
+###############################
+### optimize hyperparameters for DeepSurv
+###############################
 klist<-readRDS("cvData/prostateCV.rds")
@@
 ```{r optimDeepHit,eval=TRUE,hide=FALSE}
+###############################
+### optimize hyperparameters for DeepHit
+###############################
 klist<-readRDS("cvData/prostateCV.rds")
@@
 ```{r hyperopcbnn,eval=TRUE,hide=FALSE}
+###############################
+### optimize hyperparameters for CBNN
+###############################
 klist<-readRDS("cvData/prostateCV.rds")
@@
 
+###############################
+### save the best performaning hyperparameter sets
+###############################
 winners<-list(cbnn=cbnnPerformance,
@@
 
+#####################################
+### run 100 fold bootstrap on the training set for each method
+#####################################
 winners<-readRDS("cvData/prostateWins.rds")
@@
 
+#######################
+###plot IPA results
+#######################
+
 #BrierIPAList<-readRDS('results/100prostateBrier.rds')
@@
 
+#######################
+###plot AUC results
+#######################
 AUClist<-readRDS('results/100prostateAuc.rds')
@@
 
+#######################
+###plot C-index results
+#######################
 cidxList<-readRDS('results/100prostatecidx.rds')

### 2023-12-17 analyses/optimizeComplex.Rmd
Improved documentation for potential users

@@
 ##########################
-
 min_delta = 0 
-
 epo=as.numeric(params$epo)
@@
 iteration<-as.numeric(params$iteration)
-
 numSamp<-5000
-
-#baselineTrials<-2
-
 reoptimize=T
-
 finalFit=T
@@
 ```{r setFixedHyperparameters,eval=TRUE,hide=FALSE}
-
+##################################
+###Set up hyperparameter search space
+##################################
 lr=c(0.001,0.01)
@@
 ```{r hyperOpDsurv,eval=TRUE,hide=FALSE}
+###############################
+### optimize hyperparameters for DeepSurv
+###############################
 klist<-readRDS("cvData/ComplexSimCV.rds")
@@
 ```{r optimDeepHit,eval=TRUE,hide=FALSE}
+###############################
+### optimize hyperparameters for DeepHit
+###############################
 klist<-readRDS("cvData/ComplexSimCV.rds")
@@
 ```{r hyperopcbnn,eval=TRUE,hide=FALSE}
+###############################
+### optimize hyperparameters for CBNN
+###############################
 klist<-readRDS("cvData/ComplexSimCV.rds")
@@
 ```{r,eval=TRUE,hide=FALSE}
+
+###############################
+### save the best performaning hyperparameter sets
+###############################
 winners<-list(cbnn=cbnnPerformance,
@@
 ```{r bestFits,eval=T}
-
+#####################################
+### run 100 fold bootstrap on the training set for each method
+#####################################
+#prepare data
 winners<-readRDS("cvData/complexOptimWins.rds")
-
 klist<-readRDS("cvData/ComplexSimCV.rds")
@@
 
+# split training validation sets
 spec = c(train = .85, validate = .15)
@@
   maxTime<-max(train$time)
-
   fullTest<-readRDS("cvData/ComplexSimTest.rds")
   test<-fullTest[,-c(ncol(data))]
+  #normalize data
   test<-normalizer(fullTest,means,sds,maxTime)[,-c(ncol(data))]
   fullTest<-normalizer(fullTest,means,sds,maxTime)
-
   val<-normalizer(res$validate,means,sds,maxTime)
-
   train<-normalizer(train,means,sds,maxTime)
+  #prepare data for linear model
   linModelTrain<-as.data.frame(rbind(train,val))
@@
 
+#######################
+###plot IPA results
+#######################
 BrierIPAList<-readRDS('results/100complexBrier.rds')
@@
 ```{r auc}
-
+#######################
+###plot AUC results
+#######################
 AUClist<-readRDS('results/100complexAuc.rds')
@@
 
+#######################
+###plot C-index results
+#######################
 cidxList<-readRDS('results/100complexcidx.rds')
-
 temp<-melt(setDT(as.data.frame(cidxList[[1]])), id.vars = c("times"), variable.name = "model")
@@
 ```{r optimDSM,eval=F}
+
+################################
+### This is kept for transparency
+### As DSM did not converge with
+### Our hyperparameters and simulation setting 
+################################
 klist<-readRDS("cvData/ComplexSimCV.rds")

### 2024-01-09 manuscript/manuscript.tex
MLwA update
fixed referencing
Updated manuscript
extended literature review

@@
 %%
-\documentclass[preprint,12pt]{elsarticle}
+\documentclass[preprint,12pt,authoryear]{elsarticle}
+\usepackage[monochrome]{xcolor}
+\usepackage[linesnumbered,ruled]{algorithm2e}
 \usepackage{hyperref}
@@
  \usepackage{float}
-%% Use the option review to obtain double line spacing
-%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
+ \usepackage{xcolor}
 
-%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
-%% for a journal layout:
-%% \documentclass[final,1p,times]{elsarticle}
-%% \documentclass[final,1p,times,twocolumn]{elsarticle}
-%% \documentclass[final,3p,times]{elsarticle}
-%% \documentclass[final,3p,times,twocolumn]{elsarticle}
-%% \documentclass[final,5p,times]{elsarticle}
-%% \documentclass[final,5p,times,twocolumn]{elsarticle}
+ \SetKwComment{Comment}{/* }{ */}
 
-%% For including figures, graphicx.sty has been loaded in
-%% elsarticle.cls. If you prefer to use the old commands
-%% please give \usepackage{epsfig}
-
-%% The amssymb package provides various useful mathematical symbols
 \usepackage{amssymb}
-%% The amsthm package provides extended theorem environments
-%% \usepackage{amsthm}
-
-%% The lineno packages adds line numbers. Start line numbering with
-%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
-%% for the whole article with \linenumbers.
-%% \usepackage{lineno}
 
@@
 
-%% Title, authors and addresses
-
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for theassociated footnote;
-%% use the fnref command within \author or \address for footnotes;
-%% use the fntext command for theassociated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for theassociated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \affiliation{organization={},
-%% addressline={},
-%% city={},
-%% postcode={},
-%% state={},
-%% country={}}
-%% \fntext[label3]{}
-
 \title{Case-Base Neural Network: survival analysis with time-varying,
@@
 
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{}
-%% \affiliation[label1]{organization={},
-%% addressline={},
-%% city={},
-%% postcode={},
-%% state={},
-%% country={}}
-%%
-%% \affiliation[label2]{organization={},
-%% addressline={},
-%% city={},
-%% postcode={},
-%% state={},
-%% country={}}
-
 \author[inst1]{Jesse Islam}
-
+\ead{jesse.islam@mail.mcgill.ca}
 \affiliation[inst1]{organization={McGill University Department of Quantitative Life Sciences},%Department and Organization
@@
 \author[inst2]{Maxime Turgeon}
+\ead{turgeon.maxime@gmail.com}
 \author[inst1,inst3]{Robert Sladek}
+\ead{rob.sladek@mcgill.ca}
 \author[inst4]{Sahir Bhatnagar}
+\ead{sahir.bhatnagar@gmail.com}
 
@@
 %% Text of abstract
-Neural network-based survival methods can model data-driven covariate interactions.
-While these methods can provide better predictive performance than regression-based approaches, not all can model time-varying interactions and complex baseline hazards.
+\textcolor{red}{In the context of survival analysis, data-driven neural network-based methods have been developed to model complex covariate effects}.
+While these methods may provide better predictive performance than regression-based approaches, not all can model time-varying interactions and complex baseline hazards.
 To address this, we propose Case-Base Neural Networks (CBNNs) as a new approach that combines the case-base sampling framework with flexible neural network architectures.
-Using a novel sampling scheme and data augmentation to naturally account for censoring, we construct a feed-forward neural network that may take time as an input.
-CBNNs predict the probability of an event occurring at a given moment to estimate the hazard function.
+Using a novel sampling scheme and data augmentation to naturally account for censoring, we construct a feed-forward neural network \textcolor{red}{that includes} time as an input.
+CBNNs predict the probability of an event occurring at a given moment to estimate the \textcolor{red}{full} hazard function.
 We compare the performance of CBNNs to regression and neural network-based survival methods in a simulation and three case studies using two time-dependent metrics.
@@
 Then, we apply all methods to three real data applications, with CBNNs outperforming the competing models in two studies and showing similar performance in the third.
-Our results highlight the benefit of combining case-base sampling with deep learning to provide a simple and flexible modeling framework for data-driven, time-varying interaction modeling of single event survival outcomes.
-An R package is available at https://github.com/Jesse-Islam/cbnn.
+Our results highlight the benefit of combining case-base sampling with deep learning to provide a simple and flexible framework \textcolor{red}{for data-driven modeling of single event survival outcomes that estimates time-varying effects and a complex baseline hazard by design.} An R package is available at \href{https://github.com/Jesse-Islam/cbnn}{https://github.com/Jesse-Islam/cbnn}.
 \end{abstract}
@@
 \begin{highlights}
-\item Case-Base Neural Networks (CBNNs) provide a neural network methodology for survival analysis that leverages a statistical sampling technique, removing the need for complex implementations while providing data-driven flexibility in full hazard function estimation.
-\item CBNNs provide an advantage over competing neural network survival models, as it naturally accounts for censoring, uses binary cross-entropy as the loss function and predicts a smooth in-time risk function.
-\item CBNNs outperform the competing models in a simulation, two studies and shows similar performance in the third.
-\item Our results highlight the benefit of combining case-base sampling with deep learning to provide a simple and flexible modeling
-framework for data-driven, time-varying interaction modeling of single event
-survival outcomes.
+\item \textcolor{red}{Case-Base Neural Networks (CBNNs) estimate the full hazard function.}
+\item \textcolor{red}{Naturally accounts for censoring and predicts smooth-in-time risk functions.}
+\item \textcolor{red}{Uses a simple objective function, unlike competing methods.}
+\item \textcolor{red}{Models time-varying effects by design, unlike competing methods.}
+\item \textcolor{red}{CBNNs outperform the competing models in a simulation and two studies.}
 \end{highlights}
@@
 %% keywords here, in the form: keyword \sep keyword
-survival analysis, \sep machine learning \sep case-base \sep neural network
-%% PACS codes here, in the form: \PACS code \sep code
-%\PACS 0000 \sep 1111
-%% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code (2000 is the default)
-%\MSC 0000 \sep 1111
+survival analysis \sep machine learning \sep case-base \sep neural network
+
 \end{keyword}
@@
 
-%% \linenumbers
-
 \hypertarget{introduction}{%
@@
 
-A common assumption in survival analysis is that the risk of the event of interest does not vary with
-time (i.e. proportional hazards) \citep{hanley2009}. This simplifying assumption results in using Cox proportional
-hazards models more often than smooth-in-time, accelerated failure time (AFT) models \citep{hanley2009}. This
-preference causes analyses to be based on hazard ratios and relative risks rather than on survival curves and
-absolute risks \citep{hanley2009}. The proportional hazards assumption in a Cox model may be incorrect in studies
-where the disease pathogenesis may change over time \citep{coradini2000time}. For example, studies of breast cancer
-incidence have shown that the risk of developing a tumor involves a time-varying interaction with tumor size
-\citep{coradini2000time}. These time-varying covariates can be incorporated easily into AFT models \citep{royston2002flexible}.
-However, these regression-based models require prior knowledge of potential
-time-varying interactions and their quantitative effects.
+A common assumption in survival analysis is that the risk \textcolor{red}{ratio} of the event of interest does not vary with time \textcolor{red}{which has led to the dominance of proportional hazard models\citep{cox1972regression}} \citep{hanley2009}.
+This simplifying assumption \textcolor{red}{explains the popularity of Cox proportional hazards models over smooth-in-time, accelerated failure time (AFT) models and results in analyses} based on hazard ratios and relative risks rather than on survival curves and absolute risks \citep{hanley2009}.
+The proportional hazards assumption in a Cox model may be incorrect in studies where the disease pathogenesis may change over time.
+\textcolor{red}{For example, age, chronic obstructive pulmonary disease, diabetes mellitus, existing heart disease, prior heart surgery, intra-aortic balloon pump, urgent and emergent surgical priority are all preoperative factors that have demonstrated time-varying associations with long-term mortality following coronary artery bypass graft surgery \citep{gao2006time}.
+A time-varying effect of body mass index has been associated with all-cause mortality in hypertensive patients \citep{zhu2022time} and the effect of long-term statin use on type 2 diabetes onset varies with time \citep{na2020time}.
+The time-varying effect of tumor size has been associated with the risk of cancer recurrence \citep{coradini2000time}.
+Time-varying interactions were identified in appraisals of 28 of 40 targeted and immuno-oncology therapies reported by the National Institute for Health and Care Excellence, highlighting the importance of accounting for time-varying features in cancer treatment studies \citep{salmon2023clinical}.} These time-varying covariate \textcolor{red}{effects} can be incorporated easily into AFT models; however, this requires prior knowledge of potential time-varying interactions and their quantitative effects \citep{royston2002flexible}.
+\textcolor{red}{In this study, we develop a method that accounts for these time-varying effects without user specification, improving risk prediction models.}
 
-Neural networks provide a data-driven approach to approximating interaction terms that benefit prediction. For example,
-DeepSurv is a neural network-based proportional hazards model that implements the Cox partial log-likelihood
-as a custom loss function \citep{katzman2018DeepSurv}. This results in a stepwise absolute risk curve that cannot
-accommodate time-varying interactions.
-DeepSurv performs better than Cox proportional hazard models in the Rotterdam and German Breast Cancer Study Group and The Molecular Taxonomy of Breast Cancer International Consortium datasets \citep{katzman2018DeepSurv}.
-It has also been used to develop personalized prediction models for cancers in a number of studies \citep{ds1} \citep{ds2} \citep{ds3}.
-Previous work suggests we may change the loss function to address
-non-proportional hazards \citep{faraggi1995neural}. In contrast, DeepHit assumes an inverse Gaussian distribution
-as the baseline hazard \citep{lee2018DeepHit} and specifies each follow-up time of interest in the model. DeepHit
-directly estimates survival curves, rather than deriving a hazard function \citep{lee2018DeepHit}. On real datasets,
-DeepHit outperformed DeepSurv \citep{lee2018DeepHit}. We note that DeepSurv and DeepHit both use custom loss
-functions \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}. In addition, DeepHit introduces a hyperparameter
-weighing its two loss functions (negative log-likelihood and ranking losses) \citep{lee2018DeepHit}. Compared
-to these neural network methods, regression-based approaches require prior specification of all interaction terms,
-which makes it challenging to model covariate effects that change with time. The current neural network models
-provide flexibility at the cost of clarity, while regression models provide clarity at the cost of flexibility.
+We propose Case-Base Neural Networks (CBNNs) as a data-driven method for single event survival analysis.
+CBNNs use the case-base sampling technique, which allows probabilistic models to predict survival outcomes \textcolor{red}{by adjusting for a sampling-specific bias with an offset term} \citep{hanley2009}.
+After case-base sampling, we implement a model using common neural network components that use time as a feature to estimate time-varying interactions and a flexible baseline hazard.
 
-We propose Case-Base Neural Networks (CBNNs) as a method for single event survival analysis that models
-time-varying interactions and a flexible baseline hazard. CBNNs use the case-base sampling technique, which
-allows probabilistic models to predict survival outcomes \citep{hanley2009}. After case-base sampling, we
-implement a model using common neural network components that uses transformations of time as a feature
-(covariate) to specify different baseline hazards. For example, by including splines of time as covariates,
-case-base with logistic regression (CBLR) can approximate the Royston-Parmar flexible baseline hazard
-\citep{royston2002flexible} \citep{hanley2009}. However, this still requires explicit specification of
-time-varying interactions. CBNN can model both without extra tuning parameters.
+In this paper, \textcolor{red}{we first review state-of-the-art deep learning methods for survival analysis to position CBNN relative to related works.
+Second,} we describe the case-base sampling \textcolor{red}{procedure} and compare its properties to neural network models, along with our hyperparameter selection \textcolor{red}{criteria}, metrics, and software implementation.
+\textcolor{red}{Third}, we compare the performance of CBNN to neural network and regression-based survival methods on simulated data and describe their performance in three case studies.
+\textcolor{red}{Finally}, we explore the implications of our results and contextualize them within neural network survival analysis in a single event setting.
 
-In this paper, we describe the case-base sampling method and compare its properties to neural
-network models, along with our hyperparameter selection procedure,
-metrics and software implementation. Next, we explore the performance of CBNN, DeepSurv,
-DeepHit, Cox regression and CBLR on simulated data and
-describe their performance in three case studies. Finally, we explore
-the implications of our results and contextualize them within neural network survival analysis in a single
-event setting.
+ \color{red}
+\section{Related works}\label{related}
+ Deep learning approaches for survival analysis can be grouped into four broad categories: models that ignore censoring; Cox partial log-likelihood; discrete survival time; and fully parametric.
+We wish to distinguish models, where the innovation is provided by the neural network architecture, from frameworks, which alter how the survival outcome is interpreted.
+Based on this distinction, any framework can use any neural network architecture.
+
+\subsection{Survival models that ignore censoring}
+
+Survival analysis methods handle censored data, where the event time of interest may not be known.
+Several methods have been developed to ignore these censored individuals during the fitting process, simplifying the survival problem.
+For example, \citet{zadeh2020DeepSurvnet} model survival data in a deep learning context, where the model predicts survival rate categories.
+In contrast, two other studies predict the probability of a bad prognosis by integrating imaging and gene expression data, using either a bilinear network \citep{wang2021gpdbn} or an attention layer with transfer learning for the imaging features and ensemble forests for gene expression \citep{jia2023dccafn}.
+Though these methods predict a simple binary outcome, they are unable to estimate the hazard function.
+These architectures can be inserted into the CBNN framework which provides the added flexibility of estimating complex baseline hazards and time-varying effects.
+Rather than converting time-to-event outcomes into a risk-of-prognosis score, CBNN can model the full hazard function.
+
+\subsection{Cox partial log-likelihood} \label{coxFrame}
+
+Several architectures using the Cox partial log-likelihood have been proposed for conventional tabular datasets.
+\textcolor{black}{DeepSurv is a neural network-based proportional hazards model \citep{katzman2018DeepSurv}, which has been used to develop prediction models} \textcolor{red}{for several clinical scenarios \citep{ds1} \citep{ds2} \citep{ds3} \citep{she2020development}}.
+\citet{ching2018cox} use gene expression data as input to a single hidden layer neural network to model cancer survival; an approach which is outperformed by an autoencoder with Cox regression model \citep{yin2022cox} and a convolutional model with residual networks \citep{huang2020deep}.
+In parallel, \citet{hao2021deep} assess simple Cox neural networks in a high dimensional setting against random forest and penalized regression approaches.
+As a more complex model for gene expression, \citet{meng2022novel} propose to first fit a generative adversarial network (GAN) to the expression data before transferring the weights to a Cox neural network.
+These methods provide data-agnostic architectures in clinical settings that do not take advantage of known relationships between features.
+
+ Other approaches use the prior knowledge of existing structure between features to improve predictive performance. For example,
+ \citet{chen2023pathological} transfer the embeddings of pre-trained convolutional layers on histopathological images to an autoencoder, the central layer of which is passed into a linear Cox regression to predict cervical cancer survival.
+Histopathological images were also used to predict lung cancer survival, using either convolutional layers \citep{zhu2017wsisa} or a convolutional transformer and Siamese network model with a modified Cox partial log-likelihood loss function that accounts for aleatoric uncertainty \citep{tang2023explainable}.
+\citet{kaynar2023pideel} propose a pathway-informed model integrating metabolomic data and known metabolic pathways to predict survival in patients with glioma.
+To assess cardiovascular risk, \citet{barbieri2022predicting} propose a sex-specific neural network model using bidirectional gated recurrent units to represent recent clinical history.
+These architectures focus on structure within unimodal datasets and extend the concepts proposed by data-agnostic architectures.
+
+Other approaches integrate prior knowledge across multiple modalities to improve predictive performance.
+\citet{hao2022joint} use sample similarity and correlations across three genomic data modalities, then integrate them in a graph neural network to model cancer outcomes.
+To integrate genomic and imaging data, \citet{mobadersany2018predicting} make use of convolutional layers and \citet{li2022hfbsurv} propose a factorized bilinear network as an improvement over an unfactorized network.
+Finally, \citet{chen2020pathomic} integrate spatial transcriptomics as well, using both convolutional layers and graphical neural networks.
+Whether data-agnostic, data-specific or multi-modal, all these models make use of the Cox partial log-likelihood and assume time-invariant effects.
+
+\subsection{Discrete-time}
+
+Discrete-time survival models make survival predictions for specified intervals of follow-up time.
+For example, \textcolor{black}{DeepHit directly estimates survival probabilities and assumes an inverse Gaussian distribution as the baseline hazard \citep{lee2018DeepHit}}.
+The same goal is accomplished by \citet{gensheimer2019scalable} using basic hidden layers and \citet{giunchiglia2018rnn} using a recurrent neural network architecture.
+To estimate a piecewise hazard, \citet{kopper2022deeppamm} convert survival data into a piecewise exponential data format and propose a computationally efficient neural network.
+To predict cancer survival, \citet{wulczyn2020deep} incorporate prior knowledge of image motifs using convolutional layers.
+Discrete-time neural network models have also been used with multi-modal data.
+To predict long term cancer survival, \citet{vale2021long} propose a model integrating images along with transcriptomic, genomic and epigenomic data.
+All these methods provide a discrete estimate of survival outcomes dependent on the follow-up times of interest.
+
+\subsection{Fully parametric framework}
+
+ Fully parametric methods provide an estimate of the full hazard function.
+For example, Deep Survival Machines (DSM), uses neural networks with a mixture of distributions to fit a flexible baseline hazard \citep{dsmPaper}.
+This method extends to settings where the covariates are measured multiple times (i.e.
+time-series) \citep{nagpal2021deep}.
+However, the initial reports did not assess performance when time-varying effects of baseline measurements are present \citep{dsmPaper} \citep{nagpal2021deep}.
+
+ clinical datasets often show significant time-varying effects that are relevant to disease-specific outcomes \citep{coradini2000time}\citep{gao2006time} \citep{na2020time} \citep{zhu2022time} \citep{salmon2023clinical}.
+Not limited to these diseases, we expect all models that predict a survival outcome may gain improved performance by accounting for non-proportional hazards.
+Cox partial log-likelihood based models assume the effects are time invariant \citep{cox1972regression}.
+CBNN accounts for the time-varying effects by design.
+For all the discrete-time frameworks and architectures, a selection of smaller intervals of time or more survival times of interest may result in a few or no events per interval, which potentially leads to instability in the optimization function \citep{RJ-2022-052}.
+In contrast, selecting larger intervals may mask nonlinear, time-varying effects in the hazard function \citep{RJ-2022-052}.
+CBNNs can model time-varying interactions and a complex baseline hazard function without the trade-off between long or short interval lengths.
+CBNNs also estimate a full hazard function for all of follow-up time, rather than discrete intervals of follow-up-time.
+Compared to fully parametric approaches like DSM, CBNN explicitly incorporates time into the model, providing a simple way to account for time-varying effects that improve survival prediction.
 
@@
 
+\color{black}
+
 \hypertarget{methods}{%
@@
 
-Case-base sampling is an alternative framework for survival analysis \citep{hanley2009},
-which converts the total survival time into discrete person-moments. In this section, we detail how neural
-networks explicitly incorporate time as a feature while adjusting for the sampling bias in case-base
-sampled data and describe the metrics and hyperparameter selection procedure used to compare
-CBNN with other software packages. An R package to use CBNN is available at
-\url{https://github.com/Jesse-Islam/cbnn}. The entire code base to reproduce the figures
-and empirical results in this paper is available at \url{https://github.com/Jesse-Islam/cbnnManuscript}.
+Case-base sampling is an alternate framework for survival analysis \citep{hanley2009},
+which converts the total survival time into discrete \textcolor{red}{person-time coordinates} (person-moments).
+In this section, we detail how neural networks explicitly incorporate time as a feature while adjusting for the sampling bias in case-base sampled data\textcolor{red}{;} describe the metrics\textcolor{red}{;} and \textcolor{red}{describe the} hyperparameter selection procedure used to compare
+CBNN with other methods.
+An R package to use CBNN is available at \url{https://github.com/Jesse-Islam/cbnn}.
+The entire code base to reproduce the figures and empirical results in this paper is available at \url{https://github.com/Jesse-Islam/cbnnManuscript}.
 
@@
 
-To implement case-base sampling, we divide the total survival time for each individual into discrete person-specific moments (person-moments) and treat each person-moment as a sample.
-This creates a \emph{base series} of
-\emph{person-moments} where an event does not occur. This \emph{base series} complements the \emph{case series}, which contains
-all person-moments at which the event of interest occurs.
+To implement case-base sampling, we divide the total survival time for each individual into discrete person-moments and treat each person-moment as a sample.
+This creates a \emph{base series} of \emph{person-moments} where an event does not occur.
+This \emph{base series} complements the \emph{case series}, which contains all person-moments at which the event of interest occurs.
 
 For each person-moment sampled, let \(X_i\) be the corresponding covariate profile
-\(\left(x_{i1},x_{i2},...,x_{ip} \right)\), \(T_i\) be the time of the person-moment and \(Y_i\) be the indicator
-variable for whether the event of interest occurred at time \(T_i\). We estimate the hazard function
-\(h(t \mid X_i)\) using the sampled person-moments. Recall that \(h(t \mid X_i)\) is the instantaneous
-risk of experiencing the event at time \(t\) for a given set of covariates \(X_i\), assuming \(T_i \geq t\).
+\(\left(x_{i1},x_{i2},...,x_{ip} \right)\), \(T_i\) be the time of the person-moment and \(Y_i\) be the indicator variable for whether the event of interest occurred at time \(T_i\).
+We estimate the hazard function
+\(h(t \mid X_i)\) using the sampled person-moments.
+Recall that \(h(t \mid X_i)\) is the instantaneous risk of experiencing the event at time \(t\) for a given set of covariates \(X_i\), assuming \(T_i \geq t\).
 
-Now, let \(b\) be the (user-defined) size of the \emph{base series} and let \(B\) be the sum of all follow-up
-times for the individuals in the study. Let $c$ be the number of events in the \emph{case series}. A
-reasonable concern is determining how large $b$ should be relative to $c$, since the size of \(b\)
-influences how much information is lost in the sampling process \citep{hanley2009}. The relative information when comparing two averages
-is measured by $\frac{cb}{c+b}=\left[\left( \frac{1}{c}+\frac{1}{b}\right)\right]$, where $b$ is the sample size of the \emph{base series} and $c$
-is the sample size of the \emph{case series} \citep{hanley2009} \citep{mantel1}. If $b=100c$, then we expect learned weights to be at most one
-percent higher than if the entire study base $B$ was used, as they are proportional to
-$\frac{1}{c}+\frac{1}{100c}$ rather than $\frac{1}{c}+\frac{1}{\infty c}$ \citep{hanley2009} \citep{mantel1}.
+Now, let \(b\) be the (user-defined) size of the \emph{base series} and let \(B\) be the sum of all follow-up times for the individuals in the study.
+Let $c$ be the number of events in the \emph{case series}.
+A reasonable concern is determining how large $b$ should be relative to $c$, since the size of \(b\) influences how much information is lost in the sampling process \citep{hanley2009}.
+The relative information when comparing two averages is measured by $\frac{cb}{c+b}=\left[\left( \frac{1}{c}+\frac{1}{b}\right)\right]$, where $b$ is the sample size of the \emph{base series} and $c$ is the sample size of the \emph{case series} \citep{hanley2009} \citep{mantel1}.
+If $b=100c$, then we expect learned weights to be at most one percent higher than if the entire study base $B$ was used, as they are proportional to $\frac{1}{c}+\frac{1}{100c}$ rather than $\frac{1}{c}+\frac{1}{\infty c}$ \citep{hanley2009} \citep{mantel1}.
 
-If we sample the \emph{base series} uniformly across the study base, then the hazard function
-of the sampling process is equal to \(b/B\). Therefore, we have the following equality \citep{saarela2015}
+If we sample the \emph{base series} uniformly across the study base, then the hazard function of the sampling process is equal to \(b/B\). Therefore, we have the following equality \citep{saarela2015}
 [For a rigorous treatment, see Saarela \& Hanley (2015) section 3].:
@@
 
-To estimate the correct hazard function, we adjust for the bias introduced when sampling a fraction of the
-study base (\(\frac{b}{B}\)) by adding the term \(\log\left(\frac{B}{b} \right)\) to offset \(\log\left(\frac{b}{B} \right)\)
-during the fitting process.
+To estimate the correct hazard function, we adjust for the bias introduced when sampling a fraction of the study base (\(\frac{b}{B}\)) by adding the term \(\log\left(\frac{B}{b} \right)\) to offset \(\log\left(\frac{b}{B} \right)\) during the fitting process.
+\textcolor{red}{See Algorithm \ref{alg:1} for a pseudocode implementation of the case-base sampling procedure.
+For a discussion of the connection between our objective function and the full data likelihood, see \cite{saarela2016case}.}
+
+\begin{algorithm}
+\color{red}
+ \SetKwInOut{Input}{Input}
+ \SetKwInOut{Output}{Output}
+ \underline{function sampleCaseBase$(X)$}\;
+ \Input{X = matrix of covariate profiles. \\ $X_{T}$ = Vector of event times for each profile. \\ $X_{Y}$ = Vector of event indicators for each profile.}
+ \Output{Case-base sampled data matrix.}
+ $n$ = length($X_{T}$) \Comment*[r]{Number of subjects}
+ $B$ = Sum($X_{T}$) \Comment*[r]{Total person-time in study base}
+ $c$ = Sum($X_{Y}$) \Comment*[r]{Number of cases}
+ $b$ = 100 * c \Comment*[r]{Number of base series samples}
+ \For{$i = 1, \ldots, n$}{
+ Set $p_i = X_T[i]/B$\;
+}
+\While{$j < b$}{
+ Sample $\ell$ from $\{1, \ldots, n\}$ with probabilities $p_1, \ldots, p_n$\\
+ Set $baseSeries[j, ] = X[\ell, ]$\\
+ Sample $baseSeries_T[j]$ from $\mathrm{Uniform}(0, X_t[\ell])$\\
+}
+ Set $caseSeries = X[X_{Y}==1, ]$.\\
+ return(concatenate($baseSeries$,$caseSeries$))
+ \caption{Pseudocode algorithm of the case-base sampling procedure.}\label{alg:1}
+\end{algorithm}
+
+\color{black}
 
@@
 
-After case-base sampling, we pass all features, including time, into any user-defined feed-forward
-component (Figure \ref{fig:NNarch}). Then, we add an offset term and pass the output through a sigmoid activation function
-(Figure \ref{fig:NNarch}). Since we are interested in predicting the odds of an event occurring, the sigmoid
-activation function is ideal as it is the inverse of the odds and can be used to calculate the hazard. The
-general form for the neural network using CBNN is:
+After case-base sampling, we pass all features, including time, into any user-defined feed-forward component (Figure \ref{fig:NNarch}).
+Then, we add an offset term and pass the output through a sigmoid activation function (Figure \ref{fig:NNarch}).
+Since we are interested in predicting the odds of an event occurring, the sigmoid activation function is ideal as it is the inverse of the odds and can be used to calculate the hazard.
+The general form for \textcolor{red}{any feed-forward} neural network \textcolor{red} {architecture} using CBNN is:
 
@@
 
-where \(T\) is a random variable representing the event time, \(X\) is the random variable for a covariate
-profile, \(f_{\theta}(X, T)\) represents any feed-forward neural network architecture, \(\log\left(\frac{B}{b}\right)\)
-is the offset term to adjust for the bias (\(\log\left(\frac{b}{B}\right)\)) set by case-base sampling, \(\theta\) is the
-set of parameters learned by the neural network and \(\mathrm{sigmoid}(x)=\frac{1}{1+e^{-x}}\). By approximating
-a higher-order polynomial of time using a neural network, the baseline hazard specification is now data-driven,
-while user-defined hyperparameters such as regularization, number of layers and nodes control the flexibility
-of the hazard function.
+\noindent where \(T\) is a random variable representing the event time, \(X\) is the random variable for a covariate profile, \(f_{\theta}(X, T)\) represents any feed-forward neural network architecture, \(\log\left(\frac{B}{b}\right)\) is the offset term to adjust for the bias (\(\log\left(\frac{b}{B}\right)\)) set by case-base sampling, \(\theta\) is the set of parameters learned by the neural network and \(\mathrm{sigmoid}(x)=\frac{1}{1+e^{-x}}\).
+By approximating a higher-order polynomial of time using a neural network, the baseline hazard specification is now data-driven, while user-defined hyperparameters such as regularization, number of layers and nodes control the flexibility of the hazard function.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures/Figure1.pdf}}
+{\centering \includegraphics[width=1\linewidth]{Figure1.pdf}}
 
-\caption{Methodological steps involved in CBNN. The first step, case-base sampling, is completed before training begins.
-Then, we pass this sampled data through a feed-forward neural network and add an offset to adjust for the bias inherent in
-case-base sampling and apply a sigmoid activation function to create a probability. Once the neural network model
-completes its training, we can convert the probability output to a hazard for the survival outcome of interest.}\label{fig:NNarch}
+\caption{Methodological steps involved in CBNN.
+The first step, case-base sampling, is completed before training begins.
+Then, we pass this sampled data through a feed-forward neural network\textcolor{red}{,} add an offset to adjust for the bias inherent in case-base sampling and apply a sigmoid activation function to \textcolor{red}{estimate} a probability.
+Once the neural network model completes its training, we can convert the probability to a hazard for the survival outcome of interest.}\label{fig:NNarch}
 \end{figure}
@@
 
-We use binary cross-entropy as our loss function \citep{gulli2017}:
+We use binary cross-entropy as our loss function \citep{gulli2017}\textcolor{red}{, from which we calculate}:
 \begin{align*}
-L(\theta)=-\frac{1}{N} \sum^{N}_{i=1} y_{i} \cdot \log(\hat{f}_{\theta}(x_{i}, t_{i}) ) + (1-y_{i} )\cdot \log(1-\hat{f}_{\theta}(x_{i}, t_{i}) ),
-\end{align*} where \(\hat{f}_{\theta}(x_{i}, t_{i})\) is our estimate
-for a given covariate profile and time, \(y_{i}\) is our target value specifying whether an event occurred and \(N\) represents
-the number of individuals in our training set.
+L(\theta)=-\frac{1}{N}\left( \sum^{N}_{i=1} y_{i} \cdot \log(\hat{f}_{\theta}(x_{i}, t_{i}) ) + (1-y_{i} )\cdot \log(1-\hat{f}_{\theta}(x_{i}, t_{i}) ) \right),
+\end{align*} where \(\hat{f}_{\theta}(x_{i}, t_{i})\) is our estimate for a given covariate profile and time, \(y_{i}\) is our target value specifying whether an event occurred, and \(N\) represents the number of individuals in our training set.
 
-Backpropagation with an appropriate minimization algorithm (such as Adam, RMSPropagation, stochastic gradient
-descent) is used to optimize the parameters in the model \citep{gulli2017}. For our analysis, we use Adam as implemented
-in Keras \citep{gulli2017}. While the size of the \emph{case series} is fixed as the number of events; the size of the
-\emph{base series} is not restricted. We use a ratio of 100:1 \emph{base series} to \emph{case series} \citep{hanley2009}.
-After fitting our model, we
-convert the output to a hazard. To use CBNN for predictions, we manually set the offset term $\left(\log\left(\frac{B}{b} \right)\right)$
-to 0 in the new data as we already account for the sampling bias during the fitting process.
+Backpropagation with an appropriate minimization algorithm (such as Adam, RMSPropagation, stochastic gradient descent) is used to optimize the parameters in the model \citep{gulli2017}.
+For our analysis, we use Adam \textcolor{red}{\citep{kingma2014adam}}.
+While the size of the \emph{case series} is fixed as the number of events; the size of the \emph{base series} is not restricted.
+We use a ratio of 100:1 \emph{base series} to \emph{case series} \citep{hanley2009}.
+After fitting our model, we convert the output to a hazard.
+To use CBNN for predictions, we manually set the offset term $\left(\log\left(\frac{B}{b} \right)\right)$ to 0 in the new data as we already account for the sampling bias during the fitting process.
 
-Since we are directly modeling the hazard, we can readily estimate the
-risk function (\(F\)) at time \(t\) for a covariate profile \(X\),
+Since we are directly modeling the hazard, we can readily estimate the risk function (\(F\)) at time \(t\) for a covariate profile \(X\),
 \begin{align}\label{eqn:ci2}
 F\left(t\mid X\right)& = 1 - \exp\left(-\int_{0}^{t}h(u|X) \,\textrm du\right).
-\end{align} We use a finite Riemann sum \citep{hughes2020calculus} to
-approximate the integral in \eqref{eqn:ci2}.
+\end{align} We use a finite Riemann sum \citep{hughes2020calculus} to approximate the integral in \eqref{eqn:ci2}.
 
@@
 
-To choose our method-specific hyperparameters, we use the Integrated Brier Score (IBS) \citep{graf1999}, which is based on the
-Brier Score (BS) and provides a summarized assessment of performance for each model. We assess the performance of models on
-a held out dataset using two metrics: 1) the Index of Prediction Accuracy (IPA) \citep{kattan2018index}; and 2) the Inverse probability
-censoring weights-adjusted time-dependent area under the receiver operating characteristic curve ($AUC_{IPCW}$) \citep{auc}. The
-IPA score is a metric for both discrimination and calibration, while the $AUC_{IPCW}$ provides a metric for discrimination only.
+To choose our method-specific hyperparameters, we use the Integrated Brier Score (IBS) \citep{graf1999}, which is based on the Brier Score (BS) and provides a summarized assessment of performance for each model.
+We assess the performance of models on a held-out dataset using two metrics: 1) the Index of Prediction Accuracy (IPA) \citep{kattan2018index}; and 2) the Inverse probability censoring weights-adjusted time-dependent area under the receiver operating characteristic curve ($AUC_{IPCW}$) \citep{auc}.
+The IPA score is a metric for both discrimination and calibration, while the $AUC_{IPCW}$ provides a metric for discrimination only.
 
@@
 The BS \citep{graf1999} is defined as \begin{align}\label{eqn:bs}
+\resizebox{0.9\hsize}{!}{$%
 BS(t)=\frac{1}{N}\sum^{N}_{i=1}\left(\frac{\left(1 - \widehat{F}(t \mid X_{i})\right)^{2}\cdot I(T_{i}\leq t,\delta_{i}=1)}{\widehat{G}(T_{i})} + \frac{\widehat{F}(t\mid X_{i})^{2}\cdot I(T_{i}>t)}{\widehat{G}(t)}\right),
-\end{align} where \(\delta_{i}=1\) shows individuals who have experienced the event, \(N\) represents the number of samples
-in our dataset over which we calculate \(BS(t)\), \(T_{i}\) is the survival or censoring time of an individual, \(\widehat{G}(t)=P[c>t]\) is a non-parametric estimate of the censoring distribution and \(c\) is censoring time.
-The BS provides a score that accounts for the
-information loss because of censoring. Once we fix our \(t\) of interest, the individuals in the dataset can be divided into three
-groups. Individuals who experienced the event before \(t\) are present in the first term of the equation. The second term of the
-equation includes individuals who experience the event or are censored after \(t\). Those censored before \(t\) (the remaining
-individuals) are accounted for by the IPCW adjustment (\(G(\cdot)\)) present on both terms.
+$%
+}%
+\end{align} where \(\delta_{i}=1\) shows individuals who have experienced the event, \(N\) represents the number of samples in our dataset over which we calculate \(BS(t)\), \(T_{i}\) is the survival or censoring time of an individual, \(\widehat{G}(t)=P[c>t]\) is a non-parametric estimate of the censoring distribution and \(c\) is censoring time.
+The BS provides a score that accounts for the information loss because of censoring.
+Once we fix our \(t\) of interest, the individuals in the dataset can be divided into three groups.
+Individuals who experienced the event before \(t\) are present in the first term of the equation.
+The second term of the equation includes individuals who experience the event or are censored after \(t\).
+Those censored before \(t\) (the remaining individuals) are accounted for by the IPCW adjustment (\(G(\cdot)\)) present on both terms.
 
@@
 IBS(t)=\int_{t}^{t_{max}}BS(t)dw(t),
-\end{align*} where $w(t)=\frac{t}{t_{max}}$ and $t_{max}$ is the upper bound for the survival times of interest. As the IBS is calculated for a range of follow-up-times,
-it is a useful metric to track overall performance across all of follow-up time during cross-validation. We use the
-IPA score to get an estimate of performance at each follow-up time for our studies.
+\end{align*} where $w(t)=\frac{t}{t_{max}}$ and $t_{max}$ is the upper bound for the survival times of interest.
+As the IBS is calculated for a range of follow-up times, it is a useful metric to track overall performance across all of follow-up time during cross-validation.
+We use the IPA score to get an estimate of performance at each follow-up time for our studies.
 
@@
 \textrm{IPA}(t) &= 1-\frac{BS_{model}(t)}{BS_{null}(t)}, \nonumber
-\end{align} where $(BS_{model}(t))$ represents the BS over time $(t)$ for the model of interest and $(BS_{null}(t))$ represents
-the BS if we use an unadjusted Kaplan-Meier (KM) curve as the prediction for all observations \citep{kattan2018index}. The
-IPA score has an upper bound of one, where positive values show an increase in performance over the null model and
-negative values show that the null model performs better. As an extension of BS, the IPA score assesses both model calibration
-and discrimination and shows how performance changes over follow-up time \citep{graf1999} \citep{kattan2018index}.
+\end{align} where $BS_{model}(t)$ represents the BS over time $(t)$ for the model of interest and $(S_{null}(t)$ represents the BS if we use an unadjusted Kaplan-Meier (KM) curve as the prediction for all observations \citep{kattan2018index}.
+The IPA score has an upper bound of one, where positive values show an increase in performance over the null model and negative values show that the null model performs better.
+As an extension of BS, the IPA score assesses both model calibration and discrimination and shows how performance changes over follow-up time \citep{graf1999} \citep{kattan2018index}.
 
 \hypertarget{auc2}{%
-\subsubsection{Inverse probability censoring weights-adjusted time-dependent area under the receiver operating characteristic curve $AUC_{IPCW}$}\label{auc2}}
+\subsubsection{Inverse probability censoring weights-adjusted time-dependent area under the receiver operating characteristic curve (\texorpdfstring{$AUC_{IPCW}$}{AUC\_IPCW})}\label{auc2}}
 
@@
 
-Neural networks require external parameters that constrain the learning process (hyperparameters). We apply the following hyperparameter
-optimization procedure to each method for each study. First, we fix a test set with 15\% of the data which we keep aside during hyperparameter selection.
-To determine the best hyperparameters for each neural network method, we use a three-fold cross-validated grid search \citep{gulli2017} on the remaining
-data (85\% training, 15\% validation) for the following range of hyperparameters:
+Neural networks require external parameters that constrain the learning process (hyperparameters).
+We apply the following hyperparameter optimization procedure to each method for each study.
+First, we fix a test set with 15\% of the data which we keep aside during hyperparameter selection.
+To determine the best hyperparameters for each neural network method, we use a three-fold cross-validated grid search on the remaining data (85\% training, 15\% validation) for the following range of hyperparameters:
 \[
-Search \ space: \begin{cases}
-Learning\ rate \sim \{0.001, 0.01\} \\ 
-Dropout \sim \{0.01,0.05, 0.1\} \\
-First\ layer\ nodes \sim \{50, 75, 100\} \\
-Second\ layer\ nodes \sim \{10,25,50\} \\
-Number\ of\ batches \sim \{100, 500\} \\
-Activation\ function \sim \{ReLU, Linear\}\\
-\alpha \sim \{0, 0.5, 1\}. \ (DeepHit \ only)
+\textrm{Search space}: \begin{cases}
+\textrm{Learning rate} \sim \{0.001, 0.01\} \\ 
+\textrm{Dropout} \sim \{0.01,0.05, 0.1\} \\
+\textrm{First layer nodes} \sim \{50, 75, 100\} \\
+\textrm{Second layer nodes} \sim \{10,25,50\} \\
+\textrm{Number of batches} \sim \{100, 500\} \\
+\textrm{Activation function} \sim \{\textrm{ReLU, Linear}\}\\
+\alpha \sim \{0, 0.5, 1\} \ (\textrm{DeepHit only}).
 \end{cases}
@@
 
-DeepHit uses $\alpha$ as a hyperparameter, while DeepSurv and CBNN do not. We track IBS on the validation set for each hyperparameter combination,
-choosing the combination with the lowest score for each method (Table \ref{tab:wins}).
+\textcolor{red}{For all methods, we use the training set to estimate the mean and standard deviation and scale each covariate of interest other than time, which is divided by the training set maximum.
+For CBNN, we perform case-base sampling independently for the training and validation sets, then use the estimated training offset as our validation offset.
+The linear models combine the training set and validation set for training.} DeepHit uses $\alpha$ as a hyperparameter, while DeepSurv and CBNN do not.
+We track IBS on the validation set for each hyperparameter combination, choosing the combination with the lowest score for each method (Table \ref{tab:wins}).
 
 \begin{table}[H]
-\caption{Hyperparameters selected after three-fold cross-validated grid search along with the average IBS for each neural network model in the
-complex simulation (A); multiple myeloma (MM) case study (B); free light chain (FLC) case study (C); and prostate cancer (Prostate) case study (D).}
+\caption{Hyperparameters selected after three-fold cross-validated grid search along with the average IBS for each neural network model in the complex simulation (A), multiple myeloma (MM) case study (B), free light chain (FLC) case study (C) and prostate cancer (Prostate) case study (D).}
 \label{tab:wins}
-\begin{center}\includegraphics[width=1\linewidth]{figures/Table1.pdf} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{Table1.pdf} \end{center}
 \end{table}
@@
 
-R \citep{Rsoft} and Python \citep{py} are used to evaluate methods from both languages. We fit the Cox model using the \textbf{survival} package
-\citep{survpkg} and the CBLR model using the \textbf{casebase} package \citep{cbpkg}. Both DeepSurv and DeepHit are fit using \textbf{pyCox}
+R \citep{Rsoft} and Python \citep{py} are used to evaluate methods from both languages. We fit Cox model\textcolor{red}{s} using the \textbf{survival} package
+\citep{survpkg} and CBLR model\textcolor{red}{s} using the \textbf{casebase} package \citep{cbpkg}. Both DeepSurv and DeepHit are fit using \textbf{pyCox}
 \citep{lee2018DeepHit}. We made the components of CBNN using the \textbf{casebase} package \citep{cbpkg} for the sampling step and the
@@
 \citep{riskRegression} is used to get the IPA and $AUC_{IPCW}$. We modify the \textbf{riskRegression} package to be used with any user supplied
-risk function \(F\). We use the \textbf{reticulate} package \citep{reticulate} to run both R based methods and Python based methods on the same seed.
+risk function \(F\). We use the \textbf{reticulate} package \citep{reticulate} to run both R and Python based methods on the same seed.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures/Figure2.pdf}
+{\centering \includegraphics[width=1\linewidth]{Figure2.pdf}
 
@@
 
-\caption{Performance of each model in the complex simulation (A, E), multiple myeloma (MM) case study (B, F), free light chain (FLC) case study (C, G)
-and prostate cancer (Prostate) case study (D, H). The first row shows the IPA for each model in each study over follow-up time. Negative values mean
-the model performs worse than the null model and positive values mean the model performs better. The second row shows the $AUC_{IPCW}$ for
-each model in each study over follow-up time, where higher is better. Each model-specific metric in each study shows a 95\% confidence interval over 100 iterations. Metrics
-are shown for six models: Case-Base with Logistic Regression (CBLR), Case-Base Neural Network (CBNN), Cox Proportional Hazard (Cox), DeepHit
-and DeepSurv. The Kaplan-Meier (KM) model serves as a baseline, predicting the average curve for all individuals. CBLR and Cox have near identical
-performance, resulting in curves that overlap. The Optimal model (a CBLR model with the exact interaction terms and baseline hazard specified) shows
-the best performance we can expect on the simulated data.}\label{fig:megaPlot}
+\caption{Performance of each model in the complex simulation (A, E), multiple myeloma (MM) case study (B, F), free light chain (FLC) case study (C, G) and prostate cancer (Prostate) case study (D, H).
+The first row shows the IPA for each model in each study over follow-up time.
+Negative values mean the model performs worse than the null model and positive values mean the model performs better.
+The second row shows the $AUC_{IPCW}$ for each model in each study over follow-up time, where higher is better.
+Each model-specific metric in each study shows a 95\% confidence interval over 100 iterations.
+Metrics are shown for six models: Case-Base with Logistic Regression (CBLR), Case-Base Neural Network (CBNN), Cox Proportional Hazard (Cox), DeepHit and DeepSurv.
+The Kaplan-Meier (KM) model serves as a baseline, predicting the average curve for all individuals.
+CBLR and Cox have near identical performance, resulting in curves that overlap.
+The Optimal model (a CBLR model with the exact interaction terms and baseline hazard specified) shows the best performance we can expect on the simulated data.}\label{fig:megaPlot}
 \end{figure}
@@
 
-We simulate data to evaluate the performance of CBNN in comparison with existing regression (Cox, CBLR) and neural network (DeepHit, DeepSurv)
-methods. We specify a linear combination of each covariate as the linear predictor in the regression-based methods (Cox, CBLR), which contrasts
-with neural network approaches that allow for approximations of non-linear interactions. We simulate data based on a complex baseline hazard with time-varying
-interactions and 10\% random censoring. We simulate three covariates for 5000 individuals:
+We simulate data to evaluate the performance of CBNN in comparison with existing regression (Cox, CBLR) and neural network (DeepHit, DeepSurv) methods.
+We specify a linear combination of each covariate as the linear predictor in the regression-based methods (Cox, CBLR), which contrasts with neural network approaches that allow for approximations of non-linear interactions.
+We simulate data based on a complex baseline hazard with time-varying interactions and 10\% random censoring.
+We simulate three covariates for 5000 individuals:
 
@@
 
-In addition to the methods described above, we include the exact functional form of the covariates in a CBLR model (referred to as Optimal for simplicity)
-in the complex simulation. We obtain confidence intervals by conducting 100 bootstrap re-samples on the training data. We keep 15\% of the data for testing before
-hyperparameter selection. We use 15\% of the remaining data for validation and the rest is reserved for training. We predict risk functions for individuals in
-the test set, which are used to calculate our IPA and $AUC_{IPCW}$.
+In addition to the methods described above, we include the exact functional form of the covariates in a CBLR model (referred to as Optimal for simplicity) in the complex simulation.
+We obtain confidence intervals by conducting 100 bootstrap re-samples on the training data.
+We keep 15\% of the data for testing before hyperparameter selection.
+We use 15\% of the remaining data for validation and the rest is reserved for training.
+\textcolor{red}{We do not perform case-base sampling on the testing set and we set the offset term to 0 during prediction because the bias from case-base sampling is appropriately adjusted for during the fitting process.} We predict risk functions for individuals in the test set, which are used to calculate our IPA and $AUC_{IPCW}$.
 
 \hypertarget{complex-simulation-flexible-baseline-hazard-time-varying-interactions}{%
-\subsection{Complex simulation: flexible baseline hazard, time-varying
-interactions}\label{complex-simulation-flexible-baseline-hazard-time-varying-interactions}}
+\subsection{Complex simulation: flexible baseline hazard, time-varying interactions}\label{complex-simulation-flexible-baseline-hazard-time-varying-interactions}}
 
@@
 \end{align} where
-\(\gamma_{1}=3.9, \gamma_{2}=3, \gamma_{3}=-0.43, \gamma_{4}=1.33,\gamma_{5}=-0.86, \beta_{{1}}=-5, \beta_{{2}}=-1, \beta_{{3}}=1, \tau_{1}=0.001, \tau_{2}=-1\)
-and \(\psi_{i}\) are basis splines. The \(\gamma\) coefficients are obtained from an intercept-only cubic splines model with three knots using the \emph{flexsurvspline}
-function from the \textbf{flexsurv} package \citep{flexsurv} on the German Breast Cancer Study Group dataset as it provides a complex baseline hazard from which
-we can simulate. The study comprised 686 women with breast cancer followed between 1984 and 1989 \citep{royston2002flexible}. These $\gamma$, $\beta$ and $\tau$ coefficients are used
-as our baseline hazard parameters and are fixed for the analysis. The \(\beta\) coefficients represent direct effects, \(\tau_{2}\) represents an interaction and
-\(\tau_{1}\) is a time-varying interaction.
+\(\gamma_{1}=3.9, \gamma_{2}=3, \gamma_{3}=-0.43, \gamma_{4}=1.33,\gamma_{5}=-0.86, \beta_{{1}}=-5, \beta_{{2}}=-1, \beta_{{3}}=1, \tau_{1}=0.001, \tau_{2}=-1\) and \(\psi_{i}\) are basis splines.
+The \(\gamma\) coefficients are obtained from an intercept-only cubic splines model with three knots using the \emph{flexsurvspline} function from the \textbf{flexsurv} package \citep{flexsurv} on the German Breast Cancer Study Group dataset.
+\textcolor{red}{These coefficients provide parameters for} a complex baseline hazard from which we can simulate.
+The study comprised \textcolor{red}{of} 686 women with breast cancer followed between 1984 and 1989 \citep{royston2002flexible}.
+These $\gamma$, $\beta$ and $\tau$ coefficients are used as our baseline hazard parameters and are fixed for the analysis.
+The \(\beta\) coefficients represent direct effects, \(\tau_{2}\) represents an interaction and \(\tau_{1}\) is a time-varying interaction.
 
@@
 The Optimal model acts as a reference for ideal performance on the simulated data.
-For discrimination, the Optimal model performs best,
-followed by CBNN, DeepHit, DeepSurv and the linear models (Figure \ref{fig:megaPlot} E). To obtain a more realistic performance assessment, we compared
-models in three case studies with a time-to-event outcome.
+For discrimination, the Optimal model performs best, followed by CBNN, DeepHit, DeepSurv and the linear models (Figure \ref{fig:megaPlot} E).
+To obtain a more realistic performance assessment, we compared models in three case studies with a time-to-event outcome.
 
 \begin{table}[!htbp]
-\caption{Performance at certain percentages of follow-up time in the complex simulation (A), multiple myeloma (MM) case study (B), free light chain (FLC)
-case study (C) and prostate cancer (Prostate) case study (D). Each table shows performance for each method in each study at $25\%$, $50\%$, $75\%$
-and $100\%$ of follow-up time. These tables are included to
-provide exact measures of performance. The models of interest are case-base with logistic regression (CBLR), Cox, Case-Base Neural Network (CBNN),
-DeepHit, DeepSurv, and Optimal (in the complex simulation). The best score at each percent of follow-up time is highlighted in bold. If the average performance
-is tied, then all tied values are highlighted.}
+\caption{Performance at percentages of follow-up time in the complex simulation (A), multiple myeloma (MM) case study (B), free light chain (FLC) case study (C) and prostate cancer (Prostate) case study (D).
+Each table shows performance for each method at $25\%$, $50\%$, $75\%$ and $100\%$ of follow-up time.
+The models of interest are case-base with logistic regression (CBLR), Cox, Case-Base Neural Network (CBNN), DeepHit, DeepSurv, and Optimal (in the complex simulation).
+The best score at each percent of follow-up time is highlighted in bold.
+If tied, then all tied values are highlighted.}
 \label{tab:megaTable}
-\begin{center}\includegraphics[scale=0.24]{figures/Table2.pdf} \end{center}
+\begin{center}\includegraphics[scale=0.225]{Table2.pdf} \end{center}
 \end{table}
@@
 The simulation examines whether a complex baseline hazard and time-varying interactions affect method performance.
-The case study assesses performance in more realistic conditions, where we may not know if a flexible baseline hazard is beneficial, or if time-varying covariates
-exist. For each dataset, we perform a grid search with the same hyperparameter procedure as described in the simulation (three-fold cross-validated grid search).
-We then predict risk functions for everyone in the test set with the selected hyperparameters, which is used to calculate our metrics. We conduct 100-fold
-bootstrap re-samples on the training data to get confidence intervals.
+The case study assesses performance in more realistic conditions, where we may not know if a flexible baseline hazard or time-varying \textcolor{red}{interactions are beneficial for prediction}.
+For each dataset, we perform a grid search with the same hyperparameter procedure as described in the simulation (three-fold cross-validated grid search).
+We then predict risk functions for everyone in the test set with the selected hyperparameters, which is used to calculate our metrics.
+We conduct 100-fold bootstrap re-samples on the training data to get confidence intervals.
 
@@
 \subsection{Performance evaluation on multiple myeloma dataset}\label{pe-multiplemyeloma}}
-We expect cancer may have risk factors that vary with time \citep{coradini2000time}. As such, the first case study examines multiple myeloma (MM),
-using a cohort of 3882 patients seen at the Mayo Clinic from 1947 to 1996 until death \citep{myeloma}. Data are provided by the \textbf{survival} package \citep{survpkg}.
-We use two covariates, year of entry into the study and the time of MM diagnosis \citep{myeloma}. We see 71\% incidence over 23 years \citep{myeloma}.
+We expect cancer may have risk factors that vary with time \citep{coradini2000time} \textcolor{red}{\citep{salmon2023clinical}}.
+As such, the first case study examines multiple myeloma (MM), using a cohort of 3882 patients seen at the Mayo Clinic from 1947 to 1996 until death \citep{myeloma}.
+Data are provided by the \textbf{survival} package \citep{survpkg}.
+We use two covariates, year of entry into the study and the time of MM diagnosis \citep{myeloma}.
+We see 71\% incidence over 23 years \citep{myeloma}.
 
-Figure \ref{fig:megaPlot} B, F and Table \ref{tab:megaTable} B demonstrate the performance over time on a test set. For discrimination, CBNN and DeepHit have a
-similar performance, followed by DeepSurv and finally the linear models performing worst (Figure \ref{fig:megaPlot} F). For both discrimination and calibration, CBNN
-performs substantially better than the linear models and DeepSurv, with DeepHit having the worst performance overall (Figure \ref{fig:megaPlot} B). Together, CBNN
-is the best calibrated model and one of the best at discrimination in the MM case study.
+Figure \ref{fig:megaPlot} B, F and Table \ref{tab:megaTable} B demonstrate the performance over time on a test set.
+For discrimination, CBNN and DeepHit have a similar performance, followed by DeepSurv and finally the linear models performing worst (Figure \ref{fig:megaPlot} F).
+For both discrimination and calibration, CBNN performs substantially better than the linear models and DeepSurv, with DeepHit having the worst performance overall (Figure \ref{fig:megaPlot} B).
+Together, CBNN is the best calibrated model and one of the best at discrimination in the MM case study.
 
@@
 \subsection{Performance evaluation on free light chain dataset}\label{pe-flc}}
-Serum free light chain (FLC) is a known diagnostic tool for assessing MM \citep{mm2flc}. We are interested in whether there is a predictive benefit if the FLC markers
-varies with time. As such the second case study examines the relationship between serum FLC and mortality in a random sample of half the individuals in
-$\frac{2}{3}$ of the residents of Olmsted County over the age of 50 \citep{flc}. Data are provided by the \textbf{survival} package \citep{survpkg} for 7874 subjects tracked
-until death with 27\% incidence over 14 years \citep{flc}. We use five covariates, total serum FLC (sum of kappa and lambda), age, sex, serum creatine and monoclonal gammopathy state \citep{flc}.
+Serum free light chain (FLC) is a known diagnostic tool for assessing MM \citep{mm2flc}.
+We are interested in whether there is a predictive benefit if the FLC markers varies with time.
+As such the second case study examines the relationship between serum FLC and mortality in a random sample of half the individuals in $\frac{2}{3}$ of the residents of Olmsted County over the age of 50 \citep{flc}.
+Data are provided by the \textbf{survival} package \citep{survpkg} for 7874 subjects tracked until death with 27\% incidence over 14 years \citep{flc}.
+We use five covariates, total serum FLC (sum of kappa and lambda), age, sex, serum creatine and monoclonal gammopathy state \citep{flc}.
 
-Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C demonstrate the performance over time on a test set. CBNN, DeepHit and the linear models perform best at
-discrimination, followed by DeepSurv (Figure \ref{fig:megaPlot} G). The rankings remain the same aside from DeepHit where the performance periodically drops to worse
-than DeepSurv (Figure \ref{fig:megaPlot} C). Together, CBNN outperforms the competing models in terms of both calibration and discrimination.
-However, CBNN is only slightly better than the linear ones.
+Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C demonstrate the performance over time on a test set.
+CBNN, DeepHit and the linear models perform best at discrimination, followed by DeepSurv (Figure \ref{fig:megaPlot} G).
+The rankings remain the same aside from DeepHit where the performance periodically drops to worse than DeepSurv (Figure \ref{fig:megaPlot} C).
+Together, CBNN outperforms the competing models in terms of both calibration and discrimination.
+However, CBNN is only slightly better than the linear models.
 
@@
 \subsection{Performance evaluation on prostate cancer dataset}\label{pe-prostate}}
-As the CBNN model saw large predictive benefits on MM and small predictive benefits on FLC, we wish to examine longitudinal data with a more complex risk profile.
+As the CBNN model saw large predictive benefits on MM and small predictive benefits on FLC, we wish to examine longitudinal data with \textcolor{red}{another} complex risk profile.
 The third case study (Prostate) examines prostate cancer survival on a publicly available simulation of the Surveillance, Epidemiology, and End Results (SEER) medicare study \citep{prostate}.
-The Prostate dataset, which is contained in the \textbf{asaur} package \citep{asaur}, has a record of competing risks. As we are only interested in the single event scenario,
-we only keep individuals with prostate cancer death or censoring. This subset tracks 11054 individuals with three covariates: differentiation grade, age group and cancer
-state. There is a 7\% incidence over 10 years \citep{prostate}.
+The Prostate dataset, which is contained in the \textbf{asaur} package \citep{asaur}, has a record of competing risks.
+As we are only interested in the single event scenario, we only keep individuals with prostate cancer death or censoring.
+This subset tracks 11054 individuals with three covariates: differentiation grade, age group and cancer state.
+There is a 7\% incidence over 10 years \citep{prostate}.
 
-Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C demonstrate the performance over time on a test set. In the prostate case study, the linear models outperform the other
-models in both discrimination and calibration, followed by CBNN and DeepHit and finally DeepSurv (Figure \ref{fig:megaPlot} C, G). Aside from DeepSurv, the performance is
-similar across all models.
+Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C demonstrate the performance over time on a test set.
+In the prostate case study, the linear models outperform the other models in both discrimination and calibration, followed by CBNN and DeepHit and finally DeepSurv (Figure \ref{fig:megaPlot} C, G).
+Aside from DeepSurv, the performance is similar across all models.
 
@@
 
-CBNNs model survival outcomes by using neural networks on case-base sampled data. We incorporate follow-up time as a feature, providing a data-driven
-estimate of a flexible baseline hazard and time-varying interactions in our hazard function. The two competing neural network models we evaluated cannot
-model time-varying interactions \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}. The CBNN model performs better due to its ability to model time-varying
-interactions and a complex baseline hazard.
+CBNNs model survival outcomes by using neural networks on case-base sampled data.
+We incorporate follow-up time as a feature, providing a data-driven estimate of a flexible baseline hazard and time-varying interactions in our hazard function.
+The two competing neural network models we evaluated cannot model time-varying interactions \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}.
+The CBNN model performs better due to its ability to model time-varying interactions and a complex baseline hazard.
 
 The complex simulation requires a method that can learn both time-varying interactions and have a flexible baseline hazard.
-Based on our complex simulation results (Figure \ref{fig:megaPlot} A, E and Table \ref{tab:megaTable} A), CBNN
-outperforms the competitors. This simulation shows how all models perform under ideal conditions with minimal noise in the data, while the three
-case studies assess their performance in realistic conditions. In the MM case study, flexibility in both interaction modeling and baseline hazard
-improves the performance of CBNN over the other models, suggesting that this flexibility aids calibration (Figure \ref{fig:megaPlot} B, F and Table
-\ref{tab:megaTable} B). Upon examination of the FLC case study, CBNN demonstrates a small improvement to performance compared to the linear
-models and DeepHit for both IPA and AUC (Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C). In the Prostate case study, the linear
-models outperform the neural network ones, while CBNN and DeepHit alternate their positions depending on the follow-up time of interest and DeepSurv
-maintains last place (Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C). We attribute this to potential over-parameterization in the neural
-network models, as we did not test for fewer nodes in each hidden layer, even with dropout. Though the ranking places the linear models above the
-neural network ones, their overall performance falls within a small range of IPA and AUC values aside from DeepSurv.
+Based on our complex simulation results (Figure \ref{fig:megaPlot} A, E and Table \ref{tab:megaTable} A), CBNN outperforms the competitors.
+This simulation shows how all models perform under ideal conditions with minimal noise in the data, while the three case studies assess their performance in realistic conditions.
+In the MM case study, flexibility in both interaction modeling and baseline hazard improves the performance of CBNN over the other models, suggesting that this flexibility aids calibration (Figure \ref{fig:megaPlot} B, F and Table \ref{tab:megaTable} B).
+Upon examination of the FLC case study, CBNN demonstrates a small improvement to performance compared to the linear models and DeepHit for both IPA and AUC (Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C).
+In the Prostate case study, the linear models outperform the neural network ones, while CBNN and DeepHit alternate their positions depending on the follow-up time of interest and DeepSurv maintains last place (Figure \ref{fig:megaPlot} C, G and Table \ref{tab:megaTable} C).
+We attribute this to potential over-parameterization in the neural network models, as we did not test for fewer nodes in each hidden layer, even with dropout.
+Though the ranking places the linear models above the neural network ones, their overall performance falls within a small range of IPA and AUC values aside from DeepSurv.
 
-Compared to CBNN, the neural network competitors also have limitations. DeepSurv is a proportional hazards model and does not estimate the baseline
-hazard \citep{katzman2018DeepSurv}. DeepHit requires an alpha hyperparameter, assumes a single distribution for the baseline hazard and models the
-survival function directly \citep{lee2018DeepHit}. The alternative neural network methods match on time, while CBNN models time directly. Another method
-caught our interest during our literature review. Deep Survival Machines (DSM), a parametric survival model using neural networks with a mixture of
-distributions that can fit a flexible baseline hazard \citep{dsmPaper}. With our hyperparameter options, DSM did not converge in the complex
-simulation which may have been due to issues in software or method specific limitations. Therefore, we did not include this method in our comparison.
+Compared to CBNN, the neural network competitors have limitations.
+DeepSurv is a proportional hazards model and does not estimate the baseline hazard \citep{katzman2018DeepSurv}.
+DeepHit requires an alpha hyperparameter, assumes a single distribution for the baseline hazard and models the survival function directly \citep{lee2018DeepHit}.
+The alternative neural network methods match on time, while CBNN models time directly.
+\textcolor{red}{By modeling time directly, CBNN can approximate a flexible baseline hazard with whichever time-varying effects may be relevant to the outcome of interest.}
 
-While we apply a full grid search with three-fold cross-validation for all neural network models, there are an infinite number of untested hyperparameters we did not test. A completely
-exhaustive search is computationally infeasible, therefore it is reasonable to expect that there exists a set of hyperparameters that is better suited for each model
-in each study. However, we test a reasonably large range while accounting for potential over-fitting by including dropout \citep{gulli2017}. We provide the same
-options to all models aside from DeepHit, which has a method specific hyperparameter $\alpha$ \citep{lee2018DeepHit}. With these limitations in mind, we
-summarize the differences in performance across all tested methods.
+While we apply a full grid search with three-fold cross-validation for all neural network models, there are an infinite number of untested hyperparameters we did not test.
+A completely exhaustive search is computationally infeasible; therefore, it is reasonable to expect that there exists a set of hyperparameters that is better suited for each model in each study.
+However, we test a reasonably large range while accounting for potential over-fitting by including dropout \citep{srivastava2014dropout}.
+We provide the same options to all models aside from DeepHit, which has a method specific hyperparameter $\alpha$ \citep{lee2018DeepHit}.
+\textcolor{red}{Another method of particular interest is DSM, as it is a fully parametric model \citep{dsmPaper}.
+However,} with our hyperparameter options, \textcolor{red}{DSM} did not converge in the complex simulations, which may have been due to issues in software or method specific limitations.
+Therefore, we did not include this method in our comparison.
+With these limitations in mind, we summarize the differences in performance across all methods in our comparisons.
 
-If prediction is our goal, we suggest CBNN as the best model in the Complex simulation, MM case study and FLC case study. The linear models were
-competitive in the FLC case study and performed best in the Prostate case study. Though neural network interpretability is steadily improving \citep{interpret},
-there is still a trade-off compared to regression models, especially when the predictive performance gain is minimal, like in the FLC case study. We suggest
-that reference models should be included when assessing neural network models. Both a null model (KM curve) and a linear model (either Cox or a flexible
-baseline hazard model like CBLR) provide insight as to whether the neural network model is learning anything useful beyond linear predictors that were
-not accounted for.
+If prediction is our goal, we suggest CBNN as the best model in the Complex simulation, MM case study and FLC case study.
+The linear models were competitive in the FLC case study and performed best in the Prostate case study.
+Though neural network interpretability is steadily improving \citep{interpret}, there is still a trade-off compared to regression models, especially when the predictive performance gain is minimal, like in the FLC case study.
+We suggest that reference models should be included when assessing neural network models.
+Both a null model (KM curve) and a linear model (either Cox or a flexible baseline hazard model like CBLR) provide insight as to whether the neural network model is learning anything useful beyond linear predictors that were not accounted for.
 
@@
 Our study was motivated by the lack of easily implemented survival models based on neural networks that can approximate time-varying interactions.
-This lead us to apply the case-base sampling technique, which has previously been used with logistic regression \citep{hanley2009}, to neural network models for a data-driven
-approach to time-varying interaction modeling. In our paper, We aim to compare CBNNs with existing methods. We assess the discrimination and calibration
-of each method in a simulation with time-varying interactions and a complex baseline hazard, and three case studies. CBNNs outperform all competitors in
-the complex simulation and two case studies while maintaining competitive performance in a final case study. Once we perform case-base sampling and
-adjust for the sampling bias, we can use a sigmoid activation function to predict our hazard function. Our approach provides an alternative to incorporating
-censored individuals, treating survival outcomes as binary ones. Forgoing the requirement of custom loss functions, CBNNs only require the use of standard
-components in machine learning libraries (specifically, the add layer to adjust for sampling bias and the sigmoid activation function) after case-base sampling.
-Due to the simplicity in its implementation and by extension user experience, CBNNs are both a user-friendly approach to data-driven, single event survival analysis and is
-easily extendable to any feed-forward neural network framework.
-
-%\backmatter
+This led us to apply the case-base sampling technique, which has previously been used with logistic regression \citep{hanley2009}, to neural network models for a data-driven approach to time-varying interaction modeling.
+In our paper, \textcolor{red}{we} aim to compare CBNNs with existing methods.
+We assess the discrimination and calibration of each method in a simulation with time-varying interactions and a complex baseline hazard, and three case studies.
+CBNNs outperform all competitors in the complex simulation and two case studies while maintaining competitive performance in a final case study.
+Once we perform case-base sampling and adjust for the sampling bias, we can use a sigmoid activation function to predict our hazard function.
+Our approach provides an alternative to incorporating censored individuals, treating survival outcomes as binary ones.
+Forgoing the requirement of custom loss functions, CBNNs only require the use of standard components in machine learning libraries (specifically, the add layer to adjust for sampling bias and the sigmoid activation function) after case-base sampling.
+Due to the simplicity in its implementation and by extension user experience, CBNNs are both a user-friendly approach to data-driven, single event survival analysis and is easily extendable to any feed-forward neural network framework.
 
@@
 \citep{survpkg}. The Prostate dataset is available as part of the \textbf{asaur} package in R \citep{asaur}.
-The code for this manuscript and its analyses can be found at \url{https://github.com/Jesse-Islam/cbnnManuscript}. The software package making CBNN
-easier to implement can be found at \url{https://github.com/Jesse-Islam/cbnn}.
+The code for this manuscript and its analyses can be found at \url{https://github.com/Jesse-Islam/cbnnManuscript}.
+The software package making CBNNs easier to implement can be found at \url{https://github.com/Jesse-Islam/cbnn}.
 
@@
 
-We would like to thank Dr.~James Meigs, the project leader of UM1DK078616 and R01HL151855, for his support and helpful discussions.
+We thank Dr.~James Meigs, the project leader of UM1DK078616 and R01HL151855, for his support and helpful discussions.
 We would also like to thank Dr.~James Hanley for his support and discussions while extending the case-base methodology.
@@
 
-J.I. conceived the proof of concept, developed the theoretical formalism, developed the neural network code base,
-wrote the majority of the manuscript and performed the simulations and analyses. M.T. wrote the majority of the case-base sampling section \ref{case-base-sampling}.
-M.T. and S.B. provided key insights into the core sampling technique (case-base). R.S. provided multiple edits to the manuscript structure and key insights. All authors
-discussed the results and contributed to the final manuscript.
+J.I.
+conceived the proof of concept, developed the theoretical formalism, developed the neural network code base, wrote the majority of the manuscript, performed the simulations and analyses.
+M.T.
+wrote the majority of the case-base sampling section \ref{case-base-sampling}.
+M.T.
+and S.B.
+provided key insights into the core sampling technique (case-base).
+R.S.
+provided multiple edits to the manuscript structure and key insights.
+All authors discussed the results and contributed to the final manuscript.
 
@@
 
-This work was supported by subcontracts from UM1DK078616 and R01HL151855
-to R.S. The work was also supported as part of the Congressionally Directed Medical Research Programs (CDMRP) award W81XWH-17-1-0347.
+The subcontracts from UM1DK078616 and R01HL151855 to R.S supported this work.
+The work was also supported as part of the Congressionally Directed Medical Research Programs (CDMRP) award W81XWH-17-1-0347.
+
+\subsection*{Declaration of generative AI and AI-assisted technologies in the writing process} During the preparation of this work the author(s) used Microsoft 365 (Word) and ProWritingAid in order to improve grammar, assess typos and improve general sentence structure.
+After using these tools, the authors reviewed and edited the content as needed and take full responsibility for the content of the publication.
+
 \subsection*{Conflict of interest}
@@
 %%
- \bibliographystyle{elsarticle-num} 
+ \bibliographystyle{elsarticle-harv} 
  \bibliography{manuscript}
